### 2017-02-15 doc/manuscript.Rmd
Huge updates to the manuscript.

---
title: Biogeography and Environmental Conditions Shape Phage and Bacteria Interaction Networks Across the Healthy Human Microbiome
author: 
- name: Geoffrey D Hannigan
  affiliation: 1
- name: Melissa B Duhaime
  affiliation: 2
- name: Danai Koutra
  affiliation: 3
- name: Patrick D Schloss
  affiliation: 1,\*
address:
  - code: 1
    address: Department of Microbiology & Immunology, University of Michigan, Ann Arbor, Michigan, 48108
  - code: 2
    address: Department of Ecology and Evolutionary Biology, University of Michigan, Ann Arbor, Michigan, 48108
  - code: 3
    address: Department of Computer Science, University of Michigan, Ann Arbor, Michigan, 48108
  - code: \*
    address: To whom correspondence may be addressed.
output:
  md_document:
    variant: markdown
---

&nbsp;

&nbsp;

| ***Corresponding Author Information***
| Patrick D Schloss, PhD
| 1150 W Medical Center Dr. 1526 MSRB I
| Ann Arbor, Michigan 48109
| Phone: (734) 647-5801
| Email: pschloss@umich.edu

***Journal***: Genome Research ([*Preparation Details*](http://genome.cshlp.org/site/misc/ifora_mspreparation.xhtml))

***Keywords***: Virome, Microbiome, Graph Theory, Machine Learning

***Text Length***: 34,039 / 50,000 Characters

\* *Figures included for internal editing purposes*

\newpage

```{r loadfiles, echo=FALSE}
sigfig = 3
genfigurestats <- read.delim(file = "../rtables/genfigurestats.tsv", header = TRUE, sep = "\t")
nestats <- read.delim(file = "../rtables/nestats.tsv", header = TRUE, sep = "\t")
genmodelper <- read.delim(file = "../rtables/genmodelper.tsv", header = TRUE, sep = "\t")
interstats <- read.delim(file = "../rtables/interstats.tsv", header = TRUE, sep = "\t")
skinboxsig <- read.delim(file = "../rtables/skinboxsig.tsv", header = TRUE, sep = "\t")
```

# Abstract
Viruses and bacteria are critical components to the human microbiome and play important roles in health and disease. Previous work has relied on studying bacteria and phages in isolation (e.g. taxonomic classification, alpha diversity, and beta diversity), reducing them to two separate communities. This approach cannot capture how these communities interact to share information (e.g. horizontal gene transfer) and maintain stability. We developed and implemented a network-based analytical approach to provide an initial understanding of phage-bacteria network diversity across the human body and over time. We built a machine learning algorithm to predict which phages will infect which bacteria in each microbiome. This model was applied to paired viral and bacterial metagenomic sequence sets from three previously published human cohorts. We organized the predicted interactions into networks that allowed us to evaluate the diversity of phage-bacteria connectedness and network structure across the human body. We found that gut and skin network structure was person-specific, and was only weakly conserved among cohabitating family members. High fat diets and obesity were associated with less connected networks. There were significant differences in network structure between skin sites, with those more exposed to the external environment being less connected and more prone to instability. This study characterizes the diversity of microbiome networks across the human body, and provides a baseline for future studies to investigate the role of ecological networks in disease states.

**Word Count**: 231 / 250

\newpage

# Introduction

Viruses and bacteria are critical components to the human microbiome and play important roles in health and disease. Bacterial communities have been associated with diseases including a wide range of skin conditions [@Hannigan:2013im], acute and chronic wound healing conditions [@Hannigan:2014be; @Loesche:2016ev], and gastrointestinal diseases including inflammatory bowel disease [@He:2016ch; @Norman:2015kb], *Clostridium difficile* infections [@Seekatz:2016fz], and colorectal cancer [@Zackular:2014fba; @Baxter:2014hb]. Altered viromes (virus communities consisting primarily of bacteriophages in humans) have also been associated with various diseases and environmental perturbations including inflammatory bowel disease [@Norman:2015kb; @Manrique:2016dx], periodontal disease [@Ly:2014ew], and others [@Monaco:2016ita; @Hannigan:2015fz; @Minot:2011ez; @SantiagoRodriguez:2015gd; @Abeles:2015dy; @Abeles:2014kj]. Individual phages within these communities are capable of lysing their bacterial hosts and modulating their functionality through horizontal gene transfer (i.e. transduction) and altering bacterial host gene expression. At the community level, the human virome has begun to be implicated in promoting antibiotic resistance throughout human-associated microbial communities [@Modi:2013fia; @Hannigan:2015fz]. These virus community shifts are not reflections of the bacterial communities, but rather act in concert with bacteria as one single overall community [@Norman:2015kb; @Haerter:2014ii].

Previous work has relied on studying bacteria and phages in isolation (e.g. taxonomic classification, alpha diversity, and beta diversity), reducing them to two separate communities. In reality, bacteria and phage communities are dynamic and complex, acting in concert to share genetic information (e.g. horizontal gene transfer) and to maintain stable ecosystems, and removal of members can disrupt or even collapse those ecosystems [@Haerter:2014ii; @Harcombe:2005fd; @Middelboe:2001fl; @Poisot:2011jc; @Thompson:2012ki; @Moebus:1981kp; @Flores:2013hc; @Poisot:2012fh; @Poisot071993; @Flores:2011bh; @Jover:2015ev]. Relationship-based network approaches allow us to capture this information, including community resilience to instability and the potential for information flow. The conventional approach to understanding the human microbiome is unable to capture how these communities interact and influence each other. We bridge this knowledge gap by characterizing the human bacterial and phage communities by their *relationships*, leveraging machine learning and graph theory techniques. We focus on characterizing the network diversity of the healthy human microbiome (including the skin and gut), so as to provide a foundation on which we and others can build further studies into the network dynamics associated with disease states.

To begin characterizing the networks of bacteria and phages within the human microbiome, we leveraged three published microbiome datasets with paired virus and bacterial metagenomic sequence sets [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei]. These datasets included gut and skin samples, allowing us to gain broader insights into human microbiome network diversity across different body sites. Our approach built off of previous large-scale phage-bacteria microbiome network analyses by inferring interactions using metagenomic datasets, instead of using culture-based techniques [@Moebus:1981kp; @Flores:2013hc]. Our metagenomic interaction inference model is powered beyond previous models by its inclusion of protein interaction data, inclusion of negative interactions as well as positive, and the use of a more sophisticated machine learning algorithm instead of the linear models used by some groups [@Edwards:2015iz].

Using this approach, we were able to go beyond contemporary isolated methods to uncover a basic understanding of the community network dynamics associated with healthy human phage and bacteria. By building and utilizing a microbiome network, we showed that different people, body sites, and anatomical locations not only support distinct microbiome membership and diversity [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei; @Grice:2009eea; @Findley:2013jf; @Costello:2009im, @Consortium:2012iz], but also support communities with distinct communication structures and propensities toward community instability. By understanding the healthy state of network structures across the human body, we empower future studies to begin investigating how these community structures change in disease or otherwise altered states.

# Results

## Establishing a Model of Phage-Bacteria Infectious Networks
We studied the differences in microbiome phage-bacteria interaction networks across healthy human bodies by leveraging previously published sequence sets containing purified virome samples paired with bacterial metagenomes from whole metagenomic shotgun sequences. Our study contained three datasets, including a study of the impact of diet on the healthy human gut virome [@Minot:2011ez], the impact of anatomical location on the healthy human skin virome [@Hannigan:2015fz], and the viromes of monozygotic twins and their mothers [@Reyes:2010cwa; @Turnbaugh:2009ei]. The viromes associated with these datasets were subjected to virus-like particle (VLP) purification to eliminate other organism DNA including bacteria, fungi, and humans, and allowed us to assess the actively replicating virome.

Bacterial and viral sequences were quality filtered and assembled into contigs. The published datasets we used were conducted over the span of five years using different methods and technologies, and therefore yielded different sequence abundances **(Supplemental Figure \ref{SequenceStats} A-B)**. Because contig assembly most commonly returns genome fragments, we clustered related bacteria and phage contigs by k-mer frequency and co-abundance using the CONCOCT algorithm. These clusters represent operationally defined units of related bacteria and phage genomes that we defined as operational genomic units (OGUs). These OGUs are conceptually similar to the operational taxonomic unit (OTU) and operational protein family (OPF) definitions used for grouping highly similar 16S rRNA gene and open reading frame sequences, respectively [@Schloss:2008ht]. The resulting contigs and OGUs demonstrated high sequence coverage and length **(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})**.

We predicted which phage OGUs infected which bacteria using a random forest model trained on experimentally validated infectious relationships from six previous publications [@Jensen:1998vh;@Malki:2015tm;@Schwarzer:2012ez;@Kim:2012dh;@Matsuzaki:1992gw;@Edwards:2015iz]. This training set contained diverse bacteria and phages, with both broad and specific infectious ranges **(Supplemental Figure \ref{ValidationOverview} A - B)**. Phages with linear and circular genomes, as well as ssDNA and dsDNA genomes, were included in the analysis. Because this was a DNA sequencing study, RNA phages were considered in the analysis **(Supplemental Figure \ref{ValidationOverview} C-D)**. This training set included both positive relationships (a phage infects a bacterium) and negative relationships (a phage does not infect a bacterium). This built on previous work, which focused only on positive relationships, by allowing us to validate the false positive *and* false negative rates associated with our candidate models.

Four phage and bacterial genomic markers were used to predict infectious relationships between bacteria and phages: 1) genome nucleotide similarities, 2) gene amino acid sequence similarities, 3) CRISPR targeting of phages by bacterial CRISPR spacer sequences, and 4) similarity of protein families known to be associated with known protein-protein interactions [@Orchard:2014hq]. The resulting random forest model exhibited high performance with an AUC of `r signif(genmodelper$ROC, digits = sigfig)`, a sensitivity of `r signif(genmodelper$Sens, digits = sigfig)`, and a specificity of `r signif(genmodelper$Spec, digits = sigfig)` **(Figure \ref{RocCurve} A)**. The most important predictor in the model was nucleotide similarity between genes, followed by nucleotide similarity of whole genomes **(Figure \ref{RocCurve} B)**. Protein family interactions were moderately important to the model, while CRISPRs were minimally important, likely because they were redundant with the blast information. Approximately one third of the relationships yielded no score and were automatically classified as being non-infectious **(Figure \ref{RocCurve} C)**.

We used our random forest model to classify the relationships between bacteria and phages in the experimental datasets. The relationships within the three studies were used to construct one master network, containing the three study sub-networks, which themselves each contain sub-networks for each sample **(Figure \ref{RocCurve} D)**. Metadata including study, sample ID, disease, and abundance within the community (based on sequence count) were also stored in the multi-study master network to allow for effective parsing and downstream analysis **(Supplemental Figure \ref{NetworkDiagram})**. The resulting master network was highly connected and contained `r prettyNum(genfigurestats[c(genfigurestats$class %in% "Total"),"Edges"], big.mark = ",")` infectious relationships among `r prettyNum(genfigurestats[c(genfigurestats$class %in% "Total"),"Vertices"], big.mark = ",")` nodes, `r prettyNum(nestats[c(nestats$cats %in% "PhageNodes"),"values"], big.mark = ",")` of which represented phages and `r prettyNum(nestats[c(nestats$cats %in% "BacteriaNodes"),"values"], big.mark = ",")` that represented bacteria. Although the network was highly connected, not all relationships were present in all samples. Furthermore, relationships were weighted by relative abundance of their associated bacteria and phage, meaning that lowly abundant relationships could be present but insignificant compared to those that were more highly abundant. Like the master network, the skin network exhibited a diameter of `r prettyNum(genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Diameter"], big.mark = ",")` (measure of graph size; the greatest number of traversed vertices required between two vertices) and included almost all of the master network nodes and edges **(Figure \ref{RocCurve} E - F)**. The gut diet and twin sample sets each contained less than 150 vertices, less than 20,000 relationships, and diameters of `r prettyNum(genfigurestats[c(genfigurestats$class %in% "DietStudy"),"Diameter"], big.mark = ",")`, suggesting more sparsely related phages and bacteria **(Figure \ref{RocCurve} E - F)**.

## Individuality of Microbial Networks
Previous work has show community membership and diversity are highly personal, with signatures remaining more similar to themselves over time compared to other people [@Grice:2009ee; @Hannigan:2015fz; @Minot:2013ih]. Using our microbiome network model, we determined whether this personal conservation extended to network structure. We calculated the degree of dissimilarity between each subject's network, based on phage and bacteria abundance, as well as centrality. Centrality was evaluated by first calculating the weighted eigenvector centrality of all bacteria and phages within each sample graph. Conceptually, this metric defines central phages as those that are highly abundant and infect many bacteria which themselves are abundant and infected by many other phages. Bacterial centrality was defined in the same way. We then calculated the similarity of community networks using the overall weighted eigenvector centrality of all nodes between all samples. Biologically, samples with similar network structures are interpreted as having similar capacities to influence other microbes within the community, transmit genetic material (e.g. horizontal gene transfer), and maintain stability.

We found that gut microbiome network structure was highly individual specific, with networks clustering significantly by person (anosim p-value = `r signif(interstats[c(interstats$site %in% "DietAnosim"), "prob"], digits = sigfig)`, anosim statistic R = `r signif(interstats[c(interstats$site %in% "DietAnosim"), "abserr"], digits = sigfig)`, **Figure \ref{intradiv} A)**. Network dissimilarity within people over the 8-10 day sampling period was significantly less than the average dissimilarity between that person and others **(Figure \ref{intradiv} B)**. We evaluated the significance of an individual's network being more similar to itself than others by fitting a probability distribution to the reductions in intrapersonal dissimilarity compared to interpersonal dissimilarity (Lines between points in Figure \ref{intradiv} B - D). By taking the integral of the probability distribution from negative infinity to zero, we were able to calculate the probability that an individual's network structure was more similar to itself than others, in a given population. Thus a probability of 1 would mean that all members in a given population will be expected to be more similar to themselves compared to others, and 0.5 means they are no more likely to be more similar to themselves than dissimilar. This is a Bayesian approach that tells us how prevalent we expect this trend to be. The probability of intrapersonal diversity of an individual being less than interpersonal diversity in our dataset was `r signif(interstats[c(interstats$site %in% "Diet"), "prob"], digits = sigfig)` (absolute error = `r signif(interstats[c(interstats$site %in% "Diet"), "abserr"], digits = sigfig)`, **Figure \ref{intradiv} B)**.

Skin microbiome network structure was also conserved within individuals, although not as strongly as the gut. The probability of a given skin site being more similar to itself after a month compared to a different person at the same time was `r signif(interstats[c(interstats$site %in% "Skin"), "prob"], digits = sigfig)` (absolute error = `r signif(interstats[c(interstats$site %in% "Skin"), "abserr"], digits = sigfig)`, **Figure \ref{intradiv} C)**. This distribution was similar when separated by anatomical sites, suggesting this was an accurate representation of overall skin network dissimilarity **(Supplemental Figure \ref{allskin})**.

The conservation of gut network structures was only weakly associated with families. The gut network structures were more similar between twins and their mothers (intrafamily) compared to other families of twins and mothers (inter-family) **(Figure \ref{intradiv} D)**. The probability of a family member being more similar to another family member compared to non-family members was `r signif(interstats[c(interstats$site %in% "Twins"), "prob"], digits = sigfig)` (absolute error = `r signif(interstats[c(interstats$site %in% "Twins"), "abserr"], digits = sigfig)`).

## Role of Diet & Obesity in Gut Microbiome Conectivity
Diet is a major environmental factor that influences resource availability and gut microbiome composition and diversity, including bacteria and phages [@Minot:2011ez; @Turnbaugh:2009hf; @David:2014cl]. Previous work in isolated culture-based systems has suggested that changes in nutrient availability are associated with altered phage - bacteria network structures, although this has yet to be applied to humans [@Poisot:2011jc]. Understanding the impact diet might have on gut microbiome networks is a valuable contribution to the baseline understanding we aim to achieve here. We therefore hypothesized that network structure would be altered by changes in diet. We also described a potential association between gut network structure and obesity, a disease linked to diet and potentially the microbiome [@Sze:2016iw].

We evaluated the differences in gut network structure by quantifying how central each sample's network was on average across all microbes. We accomplished this by utilizing two common centrality metrics: degree centrality and closeness centrality. **Degree centrality**, the simplest centrality metric, was defined as the number of connections each phage made to bacteria, or each bacterium made to phages. Because this metric alone offers only minimal insight, we supplemented it with measurements of closeness centrality. **Closeness centrality** is a metric of how close each phage or bacterium is to all of the other phages and bacteria in the network. A higher closeness centrality suggests that genetic information or the effects of altered abundance would be more impactful to all other microbes in the system. Networks with higher closeness centrality also indicates an overall greater degree of connections, which indicates a greater resilience to instability. Because these values are assigned to each phage and bacterium within each network, we calculated the average connectedness and corrected for the maximum potential degree of connectedness to obtain a single value for the connectedness of each graph.

We found that gut microbiome network structures associated with high fat diets were less connected than those of low fat diets **(Figure \ref{dietnetworks} A-B)**. Tests for statistical differences were not performed, so as to prevent misleading interpretations from the small sample size. High fat diets exhibited less degree centrality **(Figure \ref{dietnetworks} A)**, meaning bacteria were overall targeted by less phages and phage tropism was more specific. High fat diets also exhibited decreased closeness centrality **(Figure \ref{dietnetworks} B)**, meaning the microbes were more distant from other microbes in the community, making genetic information transfer and the impact of altered abundance less capable of impacting other bacteria and phages within the network.

In addition to diet, there was also an association between obesity and network structure **(Figure \ref{dietnetworks} C-D)**. The obesity-associated network demonstrated a higher degree centrality **(Figure \ref{dietnetworks} C)** but a less closeness centrality, compared to the healthy controls **(Figure \ref{dietnetworks} D)**. This meant that the obesity network was overall less connected, having microbes further from all other microbes within the community.

## Variation of Network Structure Across the Human Skin Landscape
Extensive work has illustrated differences in healthy human skin microbiome between anatomical sites, including bacteria, viruses, and fungi [@Grice:2009ee; @Findley:2013jf; @Hannigan:2015fz]. These communities vary by degree of skin moisture, oil, and environmental exposure. We hypothesized that like microbial composition and diversity, microbial network structure differs between anatomical sites. We addressed this hypothesis by evaluating the changes in network structure between anatomical sites within our skin dataset.

We quantified the average centrality of each sample using the weighted eigenvector centrality metric, which yields a larger value for phages and bacteria that are more abundant and have more relationships to other bacteria and phages that themselves have more relationships. We found that intermittently moist skin sites (dynamic sites that fluctuate between being moist and dry) were significantly less connected than the more stable moist and sebaceous environments (p-value < 0.001, **Figure \ref{skinnetwork} A)**. We also found that skin sites that were protected from the environment (occluded) were much more highly connected than those that were constantly exposed to the environment or only intermittently occluded (p-value < 0.001, **Figure \ref{skinnetwork} B)**.

We supplemented this analysis by comparing the network signatures using the centrality dissimilarity approach described above. The dissimilarity between samples was a function of shared relationships, degree of centrality, and bacteria/phage abundance. When using this supplementary approach, we found that network structures significantly clustered by moisture, sebaceous, and intermittently moist status **(Figure \ref{skinnetwork} C,E)**. We also found that occluded sites were significantly different from exposed and intermittently occluded sites, but there was no difference between exposed and intermittently occluded sites **(Figure \ref{skinnetwork} D,F)**.

# Discussion
We developed and implemented a network-based analytical approach that allowed us to begin evaluating the basic properties of the human microbiome through bacteria and phage relationships, instead of membership or diversity of separate communities. The goal of this study was to provide an initial understanding of how phage-bacteria networks differ throughout the human body, so as to provide a baseline for future studies of how microbiome networks differ in disease states. This goal is similar to the initial studies of bacteria and viral diversity across the human body, as well as other environments [@Grice:2009eea; @Findley:2013jf; @Hannigan:2015fz; @Costello:2009im, @Consortium:2012iz; @Schloss:2005hz; @Minot:2011ez]. Our approach was advantageous over previous microbiome network analyses because it did not rely on inappropriate linear correlations of abundance, it was trained on both positive and negative relationships, and it involved utilization of network theory concepts.

The principles of network theory offer extensive analytical opportunities which can be applied to understand complex ecological communities. In this study we focused on utilizing metrics of connectivity to understand the extent to which communities of bacteria and phages interact (e.g. horizontal gene transfer, modulated bacterial gene expression, alterations in abundance), and how resilient they are to instability. Like earlier work laid a foundational understanding of the healthy human microbiome using membership and diversity measurements [@Grice:2009eea; @Findley:2013jf; @Hannigan:2015fz; @Costello:2009im, @Consortium:2012iz], we provide a similar understanding of healthy human microbiome network structures. These properties included variations between individuals and families, different diet types, and different anatomical skin sites.

One major overarching principle we observed was that, just as gut microbiome and virome composition and diversity are conserved in individuals [@Hannigan:2015fz; @Grice:2009eea; @Findley:2013jf; @Minot:2013ih], gut and skin microbiome network structures were also conserved within individuals over time, with the gut being more strongly similar than the skin. Network structure was weakly conserved among family members. One explanation for the weaker intrapersonal network conservation of skin compared to the gut communities is that the skin is in more direct contact with the external environment. This was further supported by the observation that network structure differed between environmentally exposed and occluded skin sites. The exposed and intermittently exposed sites were less connected and therefore are expected to have a higher propensity for instability. Likewise, intermittently moist sites demonstrated less connectedness than the more stable moist and sebaceous sites. This suggests that body sites under greater degrees of fluctuation harbored less connected, potentially less stable microbiomes.

We also observed a link between diet and microbiome network structure. Although our sample size was small, we found that high fat diets appeared to be less connected than low fat diets. This suggested that high fat diets may lead to less stable communities with a decreased ability to influence each other. We supported this finding with the observation that obesity is associated with decreased network connectivity. Further work will be required to further characterize these relationships beyond our small study, but our results allow us to speculate that the food we eat may not only impact what microbes colonize our guts, but may also impact their ability to communicate and maintain stability.

Together this work represents an initial step toward understanding the microbiome through its relationships. While these findings are informative, there are certainly caveats that should be noted. *First*, while our infection classification model is advantageous over existing models, we recognize that, like most classification models, there remains opportunity for improvement. For example, such a model is only as good as its training set, and future large-scale endeavors into infectious relationships (and the associated genomes) will provide more robust training and higher model accuracy. Just as we have improved on previous modeling efforts, we expect that new and creative scoring metrics will likely be integrated into this model to further improve model performance.

*Second*, while informative, this work was done retrospectively and relied on published research from as many as seven years ago. These archived datasets were limited by the technology and costs of the time, meaning the datasets are poorly powered for the statistical analysis we strive for today. While we were able to present initial observations, follow-up studies will be required to validate our observations.

Despite their limitations, these results will be important for informing design and interpretation of future studies of the human microbiome. By showing that microbiome relationship structure, and therefore the community potential for dissemination of information and influence, as well as the potential for maintaining stability, differs significantly between body sites and environmental conditions we demonstrate the diversity of relationships across the human body. We found that other environmental factors, such as diet, also influence relationship structures. This information builds on previous work by suggesting that bacteria and phages not only preferentially colonize different body sites, but also interact differently, allowing for different communication structures and capacities for maintaining stability. This work will also provide a foundation for future studies to begin evaluating how microbiome network dynamics change in disease states, and how the information can be leveraged therapeutically.

# Materials & Methods

## Data Availability
All associated source code is available on GitHub at the following repository:

https://github.com/SchlossLab/Hannigan-2016-ConjunctisViribus.

## Data Acquisition & Quality Control
Raw sequencing data and associated metadata was acquired from the NCBI sequence read archive (SRA). Supplementary metadata was acquired from the same SRA repositories and their associated manuscripts. The gut virome diet study (SRA: `SRP002424`), twin virome studies (SRA: `SRP002523`; `SRP000319`), and skin virome study (SRA: `SRP049645`) were downloaded as `.sra` files. Sequencing files were converted to `fastq` format using the `fastq-dump` tool of the NCBI SRA Toolkit (v2.2.0). Sequences were quality trimmed using the Fastx toolkit (v0.0.14) to exclude bases with quality scores below 33 and shorter than 75 bp [@FASTXToolkit:wr]. Paired end reads were filtered to exclude and sequences missing their corresponding pair using the `get_trimmed_pairs.py` available in the source code.

## Contig Assembly
Contigs were assembled using the Megahit assembly program (v1.0.6) [@Li:2016kd]. A minimum contig length of 1 kb was used. Iterative k-mer stepping began at a minimum length of 21 and progressed by 20 until 101. All other default parameters were used.

## Contig Abundance Calculations
Contigs were concatenated into two master files prior to alignment, one for bacterial contigs and one for phage contigs. Sample sequences were aligned to phage or bacteria contigs using the Bowtie2 global aligner (v2.2.1) [@Langmead:2012jh]. We defined a mismatch threshold of 1 bp and seed length of 25 bp. Sequence abundance was calculated from the Bowtie2 output using the `calculate_abundance_from_sam.pl` script available in the source code.

## Operational Genomic Unit Binning
Contigs often represent large fragments of genomes. In order to reduce redundancy, and the resulting artificially inflated genomic richness within our dataset, it was important to bin contigs into operational units based on their similarity. This approach is conceptually similar to the clustering of related 16S rRNA sequences into operational taxonomic units (OTUs), although here we are clustering contigs into operational genomic units (OGUs) [@Schloss:2005hz].

We clustered contigs using the CONCOCT algorithm (v0.4.0) [@Alneberg:2014fc]. Because of our large dataset and limits in computational efficiency, we randomly subsampled the dataset to include 25% of all samples, and used these to inform contig abundance within the CONCOCT algorithm. CONCOCT was used with a maximum of 500 clusters, a k-mer length of four, a length threshold of 1 kb, 25 iterations, and exclusion of the total coverage variable.

OGU abundance (`Ao`) was obtained as the sum of the abundance of each contig (`Aj`) associated with that OGU. The abundance values were length corrected such that:

$$ { A }_{ O }=\frac { { 10 }^{ 7 }\sum _{ j=1 }^{ k }{ { A }_{ j } }  }{ \sum _{ j=1 }^{ k }{ { L }_{ j } }  } $$

Where `L` is the length of each contig `j` within the OGU.

## Open Reading Frame Prediction
Open reading frames (ORFs) were identified using the Prodigal program (V2.6.2) with the meta mode parameter and default settings [@Hyatt:2012cy].

## Classification Model Creation and Validation
The classification model for predicting interactions was built using experimentally validated bacteria-phage infections or validated lack of infections from six studies [@Jensen:1998vh;@Malki:2015tm;@Schwarzer:2012ez;@Kim:2012dh;@Matsuzaki:1992gw;@Edwards:2015iz]. Associated reference genomes were downloaded from the European Bioinformatics Institute (see details in source code). The model was created based on the four metrics listed below.

The four scores were used as parameters in a random forest model to classify bacteria and bacteriophage pairs as either having infectious interactions or not. The classification model was built using the Caret R package (v6.0.73) [@caretClassificatio:_U2Lit_1]. The model was trained using five-fold cross validation with ten repeats. Pairs without scores were classified as not interacting. The model was optimized using the ROC value. The resulting model performance was plotted using the plotROC R package.

##### Identify Bacterial CRISPRs Targeting Phages
CRISPRs were identified from bacterial genomes using the PilerCR program (v1.06) [@Edgar:2007bh]. Resulting spacer sequences were filtered to exclude spacers shorter than 20 bp and longer than 65 bp. Spacer sequences were aligned to the phage genomes using the nucleotide blast algorithm with default parameters (v2.4.0) [@Camacho:2009fc]. The mean percent identity for each matching pair was recorded for use in our classification model.

##### Detect Matching Prophages within Bacterial Genomes
Temperate bacteriophages infect and integrate into their bacterial host's genome. We detected integrated phage elements within bacterial genomes by aligning phage genomes to bacterial genomes using the nucleotide blast algorithm and a minimum e-value of 1e-10. The resulting bitscore of each alignment was recorded for use in our classification model.

##### Identify Shared Genes Between Bacteria and Phages
Phages may share genes with their bacterial hosts, providing us with evidence of phage-host infectious pairs. We identified shared genes between bacterial and phage genomes by assessing amino acid similarity between the genes using the Diamond protein alignment algorithm (v0.7.11.60) [@Buchfink:2015ki]. The mean alignment bitscores for each genome pair was recorded for use in our classification model.

##### Protein - Protein Interactions
The final method we used for predicting infectious interactions between bacteria and phages was by detecting pairs of genes whose proteins are known to interact. We assigned bacterial and phage genes to protein families by aligning them to the Pfam database using the Diamond protein alignment algorithm. We then identified which pairs of proteins were predicted to interact using the Pfam interaction information within the Intact database [@Orchard:2014hq]. The mean bitscores of the matches between each pair were recorded for use in our classification model.

## Virome Network Construction
The bacteria and phage operational genomic units (OGUs) were scored using the same approach as outlined above. The infectious pairings between bacteria and phage OGUs were classified using the random forest model described above. The predicted infectious pairings and all associated metadata was saved as a graph database using Neo4j graph database software (v2.3.1) [@Neoj:Fuwr6PBN]. This network was used for downstream community analysis.

## Centrality Analysis
Degree and closeness centrality were calculated using the associated functions within the igraph R package (v1.0.1) [@Theigraphsoftware:vh]. Briefly, the **closeness centrality** of Vi was calculated taking the inverse of the average length of the shortest path (`d`) between nodes Vi and the `k` other nodes Vj, such that closeness centrality of node Vi is:

$$ { C }_{ C }\left( { V }_{ i } \right) ={ \left( \sum _{ j=1 }^{ k }{ d\left( { V }_{ i },{ V }_{ j } \right)  }  \right)  }^{ -1 } $$

The distance between nodes (`d`) was calculated as the shortest number of edges required to be traversed to move from one node to another.

The simple metric of **degree centrality** of node Vi was defined as the sum of the number of k edges Ei associated with that node, such that:

$$ { C }_{ D }\left( { V }_{ i } \right) =\sum _{ i=1 }^{ k }{ { E }_{ i } } $$

The Eigenvector centrality was calculated using the values of the first eigen vector of the associated adjacency matrix that are associated with each vertex `Vi`. Conceptually, this function results in a centrality value that reflects the connections of the vertex, as well as the centrality of the connected vertices.

The **centralization** metric was used to assess the average centrality of each sample graph `G`. Centralization was calculated by taking the sum of each vertex `Vi` centrality from the graph maximum centrality `Cw`, such that:

$$ C\left( G \right) =\frac { \sum _{ i=1 }^{ k }{ \max _{ w }{ c\left( w \right)  } -c\left( { v }_{ i } \right)  }  }{ { T } }  $$

The values were corrected for uneven graph sizes by dividing the centralization score by the maximum theoretical centralization (`T`) for a graph with the same number of vertices.

## Network Relationship Dissimilarity
We assessed similarity between graphs by evaluating the shared centrality of their vertices, as has been done previously. More specifically, we calculated the dissimilarity between graphs `Gi` and `Gj` using the Bray-Curtis dissimilarity metric and Eigenvector centrality values such that:

$$ B\left( { G }_{ i },{ G }_{ j } \right) =1-\frac { 2{ C }_{ ij } }{ { C }_{ i }+{ C }_{ j } } $$

Where `Cij` is the sum of the lesser centrality values for those vertices shared between graphs, and `Ci` and `Cj` are are the total number of vertices found in each graph. This allows us calculate the dissimilarity between graphs based on the shared centrality values between the two graphs.

# Acknowledgments
We thank the members of the Schloss lab for their underlying contributions. GDH is supported in part by the University of Michigan Molecular Mechanisms of Microbial Pathogenesis Fellowship.

\newpage

# Figures

![**Summary of Multi-Study Network Model.** *(A) ROC curve resulting from ten iterations used to create the phage - bacteria infection prediction model. (B) Importance scores associated with the metrics used in the random forest model to predict relationships between bacteria and phages. The importance score is defined as the mean decrease in accuracy of the model when a feature (e.g. Pfam) is excluded. (C) Proportions of samples included (gray) and excluded (red) in the model. Samples were excluded from the model because they did not yield any scores. Those interactions without scores were defined as not having interactions. (D) Bipartite visualization of the resulting phage-bacteria network. This network includes information from all three published studies. (E) Network diameter (measure of graph size; the greatest number of traversed vertices required between two vertices), (F) number of vertices, and (G) number of edges (relationships) for the total network (yellow) and the individual study sub-networks (diet study = red, skin study = green, twin study = orange).* \label{RocCurve}](../figures/rocCurves.pdf){ width=90% }

\newpage

![**Intrapersonal vs Interpersonal Network Dissimilarity Across Different Human Systems.** *(A) NMDS ordination illustrating network dissimilarity between subjects over time. Each sample is colored by subject, with each sample pair collected 8-10 days apart. Dissimilarity was calculated using the Bray-Curtis metric based on abundance weighted Eigenvector centrality signatures, with a greater distance representing greater dissimilarity in bacteria and phage centrality and abundance. (B) Quantification of gut network dissimilarity within the same subject over time (intrapersonal) and the mean dissimilarity between the subject of interest and all other subjects (interpersonal). The density curve is the probability distribution of the changes between intrapersonal and interpersonal diversity, representing the probability that intrapersonal dissimilarity will be lower than interpersonal dissimilarity (lower intrapersonal dissimilarity defined as change less than zero). (C) Quantification of skin network dissimilarity within the same subject and anatomical location over time (intrapersonal) and the mean dissimilarity between the subject of interest and all other subjects at the same time and the same anatomical location (interpersonal). Probability distribution the same as for panel B. (D) Quantification of gut network dissimilarity within subjects from the same family (intrafamily) and the mean dissimilarity between subjects within a family and those of other families (interfamily). Probability distribution the same as for panel B.*\label{intradiv}](../figures/intrapersonal_diversity.pdf){ width=90% }

\newpage

![**Impact of Diet and Obesity on Gut Network Structure.** *(A) Quantification of average degree centrality (number of edges per node) and (B) closeness centrality (average distance from each node to every other node) of gut microbiome networks of subjects limited to exclusively high fat or low fat diets. Lines represent the mean degree of centrality for each diet. (C) Quantification of average degree centrality and (D) closeness centrality between obese and healthy adult women.*\label{dietnetworks}](../figures/dietnetworks.pdf){ width=90% }

\newpage

![**Impact of Skin Micro-Environment on Microbiome Network Structure.**  *(A) Notched box-plot depicting differences in average Eigenvector centrality between moist, intermittently moist, and sebaceous skin sites and (B) occluded, intermittently occluded, and exposed sites. Notched box-plots were created using ggplot2 and show the median (center line), the inter-quartile range (IQR; upper and lower boxes), the highest and lowest value within 1.5 * IQR (whiskers), outliers (dots), and the notch which provides an approximate 95% confidence interval as defined by 1.58 * IQR / sqrt(n). (C) NMDS ordination depicting the differences in skin microbiome network structure between skin moisture levels and (D) occlusion. Samples are colored by their environment and their dissimilarity to other samples was calculated as described in figure \ref{intradiv}. (E) The statistical differences of networks between moisture and (F) occlusion status were quantified with an anova and post hoc Tukey test. Cluster centroids are represented by dots and the extended lines represent the associated 95% confidence intervals. Significant comparisons (p-value < 0.05) are colored in red, and non-significant comparisons are gray.*\label{skinnetwork}](../figures/skinplotresults.pdf){ width=75% }

\newpage

# Supplemental Figures

\beginsupplement

![**Sequencing Depth Summary.** *Number of sequences that aligned to (A) Phage and (B) Bacteria operational genomic units. Sequencing count aligned to OGUs was length corrected.*\label{SequenceStats}](../figures/SequenceAbund.pdf){ width=80% }

\newpage

![**Contig Summary Statistics.** *Scatter plot heat map with each hexagon representing an abundance of contigs. Contigs are organized by length on the x-axis and the number of aligned sequences on the y-axis.*\label{ContigStats}](../figures/ContigStats.pdf){ width=50% }

\newpage

![**Operational Genomic Unit Summary Statistics.** *Scatter plot with operational genomic unit clusters organized by average contig length within the cluster on the x-axis and the number of contigs in the cluster on the y-axis. Operational genomic units of (A) bacteriophages and (B) bacteria are shown.*\label{ClusterStats}](../figures/ClusterStats.pdf){ width=80% }

\newpage

![**Summary information of validation dataset used in the interaction predictive model.** *A) Categorical heat-map highlighting the experimentally validated positive and negative interactions. Only bacteria species are shown, which represent multiple reference strains. Phages are labeled on the x-axis and bacteria are labeled on the y-axis. B) World map illustrating the sampling locations used in the study (red dots). C) Quantification of bacterial host strains known to exist for each phage. D) Genome strandedness and E) linearity of the phage reference genomes used for the dataset.*\label{ValidationOverview}](../figures/BenchmarkDataset.pdf){ width=100% }

\newpage

![**Structure of the interactive network.** *Metadata relationships to samples (Phage Sample ID and Bacteria Sample ID) included the associated time point, the study, the subject the sample was taken from, and the associated disease. Infectious interactions were recorded between phage and bacteria operational genomic units (OGUs). Sequence count abundance for each OGU within each sample was also recorded.*\label{NetworkDiagram}](../figures/graphdatabasediagram.pdf){ width=50% }

\newpage

![**Intrapersonal vs Interpersonal Dissimilarity of the Skin.** *Quantification of skin network dissimilarity within the same subject and anatomical location over time (intrapersonal) and the mean dissimilarity between the subject of interest and all other subjects at the same time and the same anatomical location (interpersonal), separated by each anatomical site (forehead [Fh], palm [Pa], toe web [Tw], umbilicus [Um], antecubital fossa [Ac], axilla [Ax], and retroauricular crease [Ra]). Below is the probability distribution of the changes between intrapersonal and interpersonal diversity, representing the probability that intrapersonal dissimilarity will be lower than interpersonal dissimilarity (intrapersonal change less than zero). The probability that the slope will be less than zero (integral from negative infinity to zero) is provided in the top left corner.*\label{allskin}](../figures/intraallskin.pdf){ width=75% }

\newpage

# References

### 2017-02-15 doc/manuscript.Rmd
Updated word counts.
Updated the smaller format things.
Made font and organization updates.

@@
     variant: markdown
+fontsize: 11pt
+mainfont: 'Helvetica'
 ---
+
+&nbsp;
 
@@
 
+***Running Title***: Network Diversity of the Healthy Human Microbiome
+
 ***Journal***: Genome Research ([*Preparation Details*](http://genome.cshlp.org/site/misc/ifora_mspreparation.xhtml))
@@
 
-***Text Length***: 34,039 / 50,000 Characters
+***Text Length***: 35,640 / 50,000 Characters
 
-\* *Figures included for internal editing purposes*
+\* *Figures at the end of the document for internal review only.*
 
@@
 
+# Disclosure Declaration
+The authors report no conflicts of interest.
+
 \newpage
@@
 
-\newpage
-
 ![**Intrapersonal vs Interpersonal Network Dissimilarity Across Different Human Systems.** *(A) NMDS ordination illustrating network dissimilarity between subjects over time.
@@
 
-\newpage
-
 ![**Impact of Diet and Obesity on Gut Network Structure.** *(A) Quantification of average degree centrality (number of edges per node) and (B) closeness centrality (average distance from each node to every other node) of gut microbiome networks of subjects limited to exclusively high fat or low fat diets.
@@
 (C) Quantification of average degree centrality and (D) closeness centrality between obese and healthy adult women.*\label{dietnetworks}](../figures/dietnetworks.pdf){ width=90% }
-
-\newpage
 

### 2017-03-26 doc/manuscript.Rmd
Some minor updates.

@@
   - code: 1
-    address: Department of Microbiology & Immunology, University of Michigan, Ann Arbor, Michigan, 48108
+    address: Department of Microbiology & Immunology, University of Michigan, Ann Arbor, Michigan, 48109
   - code: 2
-    address: Department of Ecology and Evolutionary Biology, University of Michigan, Ann Arbor, Michigan, 48108
+    address: Department of Ecology and Evolutionary Biology, University of Michigan, Ann Arbor, Michigan, 48109
   - code: 3
-    address: Department of Computer Science, University of Michigan, Ann Arbor, Michigan, 48108
+    address: Department of Computer Science, University of Michigan, Ann Arbor, Michigan, 48109
   - code: \*
@@
 # Abstract
-Viruses and bacteria are critical components to the human microbiome and play important roles in health and disease.
-Previous work has relied on studying bacteria and phages in isolation (e.g.
-taxonomic classification, alpha diversity, and beta diversity), reducing them to two separate communities.
-This approach cannot capture how these communities interact to share information (e.g.
+Viruses and bacteria are critical components of the human microbiome and play important roles in health and disease.
+Previous work has relied on studying bacteria and phages in isolation, reducing them to two separate communities.
+These approaches have not captured how these communities interact to share information (e.g.
 horizontal gene transfer) and maintain stability.
-We developed and implemented a network-based analytical approach to provide an initial understanding of phage-bacteria network diversity across the human body and over time.
-We built a machine learning algorithm to predict which phages will infect which bacteria in each microbiome.
-This model was applied to paired viral and bacterial metagenomic sequence sets from three previously published human cohorts.
-We organized the predicted interactions into networks that allowed us to evaluate the diversity of phage-bacteria connectedness and network structure across the human body.
-We found that gut and skin network structure was person-specific, and was only weakly conserved among cohabitating family members.
+We developed and implemented a network-based analytical approach to provide an initial understanding of phage-bacteria network diversity throughout the human body.
+We accomplished this by building a machine learning algorithm to predict which phages infected which bacteria in a given microbiome.
+This algorithm was applied to paired viral and bacterial metagenomic sequence sets from three previously published human cohorts.
+We organized the predicted interactions into networks that allowed us to evaluate the diversity of phage-bacteria connectedness across the human body.
+We focused on establishing a foundational understanding of microbiome network structures across the human body.
+We found that gut and skin network structure was person-specific, and was not conserved among cohabitating family members.
 High fat diets and obesity were associated with less connected networks.
-There were significant differences in network structure between skin sites, with those more exposed to the external environment being less connected and more prone to instability.
-This study characterizes the diversity of microbiome networks across the human body, and provides a baseline for future studies to investigate the role of ecological networks in disease states.
+There were significant differences in network structure between skin sites, with those exposed to the external environment being less connected and more prone to instability.
+Altogether this study characterizes the diversity of microbiome networks across the human body, and provides a baseline for future studies to investigate the role of ecological networks in disease states.
 
-**Word Count**: 231 / 250
+**Word Count**: 236 / 250
 
@@
 Bacterial communities have been associated with diseases including a wide range of skin conditions [@Hannigan:2013im], acute and chronic wound healing conditions [@Hannigan:2014be; @Loesche:2016ev], and gastrointestinal diseases including inflammatory bowel disease [@He:2016ch; @Norman:2015kb], *Clostridium difficile* infections [@Seekatz:2016fz], and colorectal cancer [@Zackular:2014fba; @Baxter:2014hb].
-Altered viromes (virus communities consisting primarily of bacteriophages in humans) have also been associated with various diseases and environmental perturbations including inflammatory bowel disease [@Norman:2015kb; @Manrique:2016dx], periodontal disease [@Ly:2014ew], and others [@Monaco:2016ita; @Hannigan:2015fz; @Minot:2011ez; @SantiagoRodriguez:2015gd; @Abeles:2015dy; @Abeles:2014kj].
+Altered viromes (virus communities consisting primarily of bacteriophages in humans) have also been associated with various diseases and environmental perturbations including inflammatory bowel disease [@Norman:2015kb; @Manrique:2016dx], periodontal disease [@Ly:2014ew], spread of antibiotic resistance [@Modi:2013fia; @Hannigan:2015fz], and others [@Monaco:2016ita; @Hannigan:2015fz; @Minot:2011ez; @SantiagoRodriguez:2015gd; @Abeles:2015dy; @Abeles:2014kj].
+These shifts in virus community composition are not reflections of the bacterial communities, but rather act in concert with bacteria as one single overall community [@Norman:2015kb; @Haerter:2014ii].
 Individual phages within these communities are capable of lysing their bacterial hosts and modulating their functionality through horizontal gene transfer (i.e.
 transduction) and altering bacterial host gene expression.
-At the community level, the human virome has begun to be implicated in promoting antibiotic resistance throughout human-associated microbial communities [@Modi:2013fia; @Hannigan:2015fz].
-These virus community shifts are not reflections of the bacterial communities, but rather act in concert with bacteria as one single overall community [@Norman:2015kb; @Haerter:2014ii].
 
-Previous work has relied on studying bacteria and phages in isolation (e.g.
-taxonomic classification, alpha diversity, and beta diversity), reducing them to two separate communities.
-In reality, bacteria and phage communities are dynamic and complex, acting in concert to share genetic information (e.g.
-horizontal gene transfer) and to maintain stable ecosystems, and removal of members can disrupt or even collapse those ecosystems [@Haerter:2014ii; @Harcombe:2005fd; @Middelboe:2001fl; @Poisot:2011jc; @Thompson:2012ki; @Moebus:1981kp; @Flores:2013hc; @Poisot:2012fh; @Poisot071993; @Flores:2011bh; @Jover:2015ev].
-Relationship-based network approaches allow us to capture this information, including community resilience to instability and the potential for information flow.
-The conventional approach to understanding the human microbiome is unable to capture how these communities interact and influence each other.
-We bridge this knowledge gap by characterizing the human bacterial and phage communities by their *relationships*, leveraging machine learning and graph theory techniques.
-We focus on characterizing the network diversity of the healthy human microbiome (including the skin and gut), so as to provide a foundation on which we and others can build further studies into the network dynamics associated with disease states.
+Previous work has relied on studying bacterial and phage communities in isolation, reducing them to two separate communities.
+In reality, bacteria and phage communities are dynamic and complex, acting in concert to share genetic information and to maintain stable ecosystems, and removal of members can disrupt or even collapse those ecosystems [@Haerter:2014ii; @Harcombe:2005fd; @Middelboe:2001fl; @Poisot:2011jc; @Thompson:2012ki; @Moebus:1981kp; @Flores:2013hc; @Poisot:2012fh; @Poisot071993; @Flores:2011bh; @Jover:2015ev].
+The conventional approach to understanding the human microbiome is unable to capture how these communities interact and influence each other, but relationship-based network approaches allow us to capture this information.
+We aimed to bridge this knowledge gap by characterizing human bacterial and phage communities by their *relationships*, leveraging machine learning and graph theory techniques.
+We focused on characterizing the network diversity of the healthy human microbiome (including the skin and gut), so as to provide a foundation on which we and others can build further studies into the network dynamics associated with disease states.
 
-To begin characterizing the networks of bacteria and phages within the human microbiome, we leveraged three published microbiome datasets with paired virus and bacterial metagenomic sequence sets [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei].
+We began characterizing human microbiome phage and bacteria networks by leveraged three published microbiome datasets with paired virus and bacterial metagenomic sequence sets [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei].
 These datasets included gut and skin samples, allowing us to gain broader insights into human microbiome network diversity across different body sites.
 Our approach built off of previous large-scale phage-bacteria microbiome network analyses by inferring interactions using metagenomic datasets, instead of using culture-based techniques [@Moebus:1981kp; @Flores:2013hc].
-Our metagenomic interaction inference model is powered beyond previous models by its inclusion of protein interaction data, inclusion of negative interactions as well as positive, and the use of a more sophisticated machine learning algorithm instead of the linear models used by some groups [@Edwards:2015iz].
+Our metagenomic interaction inference model was powered beyond previous models by its inclusion of protein interaction data, inclusion of negative interactions as well as positive, and the use of a more sophisticated machine learning algorithm instead of linear models [@Edwards:2015iz].
 
-Using this approach, we were able to go beyond contemporary isolated methods to uncover a basic understanding of the community network dynamics associated with healthy human phage and bacteria.
-By building and utilizing a microbiome network, we showed that different people, body sites, and anatomical locations not only support distinct microbiome membership and diversity [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei; @Grice:2009eea; @Findley:2013jf; @Costello:2009im, @Consortium:2012iz], but also support communities with distinct communication structures and propensities toward community instability.
-By understanding the healthy state of network structures across the human body, we empower future studies to begin investigating how these community structures change in disease or otherwise altered states.
+Using this approach, we were able to go beyond contemporary isolated methods to provide a basic understanding of the community network dynamics associated with healthy human phages and bacteria.
+By building and utilizing a microbiome network, we found that different people, body sites, and anatomical locations not only support distinct microbiome membership and diversity [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei; @Grice:2009eea; @Findley:2013jf; @Costello:2009im, @Consortium:2012iz], but also support communities with distinct communication structures and propensities toward community instability.
+By understanding the healthy state of network structures across the human body, we will empower future studies to begin investigating how these community structures change in disease or otherwise altered states.
 
@@
 
-## Establishing a Model of Phage-Bacteria Infectious Networks
+## Cohort Curation and Sample Processing
 We studied the differences in microbiome phage-bacteria interaction networks across healthy human bodies by leveraging previously published sequence sets containing purified virome samples paired with bacterial metagenomes from whole metagenomic shotgun sequences.
 Our study contained three datasets, including a study of the impact of diet on the healthy human gut virome [@Minot:2011ez], the impact of anatomical location on the healthy human skin virome [@Hannigan:2015fz], and the viromes of monozygotic twins and their mothers [@Reyes:2010cwa; @Turnbaugh:2009ei].
-The viromes associated with these datasets were subjected to virus-like particle (VLP) purification to eliminate other organism DNA including bacteria, fungi, and humans, and allowed us to assess the actively replicating virome.
+The viromes associated with these datasets were subjected to virus-like particle (VLP) purification to eliminate other organismal DNA including bacteria, fungi, and humans, and allowed us to assess the actively replicating virome.
 
-Bacterial and viral sequences were quality filtered and assembled into contigs.
-The published datasets we used were conducted over the span of five years using different methods and technologies, and therefore yielded different sequence abundances **(Supplemental Figure \ref{SequenceStats} A-B)**.
+The publications we utilized were conducted over the span of five years using different methods and technologies, and therefore yielded different sequence abundances **(Supplemental Figure \ref{SequenceStats} A-B)**.
+The bacterial and viral sequences from those publications were quality filtered and assembled into contigs.
 Because contig assembly most commonly returns genome fragments, we clustered related bacteria and phage contigs by k-mer frequency and co-abundance using the CONCOCT algorithm.
@@
 
+## Establishing a Model of Phage-Bacteria Interactions
 We predicted which phage OGUs infected which bacteria using a random forest model trained on experimentally validated infectious relationships from six previous publications [@Jensen:1998vh;@Malki:2015tm;@Schwarzer:2012ez;@Kim:2012dh;@Matsuzaki:1992gw;@Edwards:2015iz].
@@
 Phages with linear and circular genomes, as well as ssDNA and dsDNA genomes, were included in the analysis.
-Because this was a DNA sequencing study, RNA phages were considered in the analysis **(Supplemental Figure \ref{ValidationOverview} C-D)**.
+Because this was a DNA sequencing study, RNA phages were not considered in the analysis **(Supplemental Figure \ref{ValidationOverview} C-D)**.
 This training set included both positive relationships (a phage infects a bacterium) and negative relationships (a phage does not infect a bacterium).
-This built on previous work, which focused only on positive relationships, by allowing us to validate the false positive *and* false negative rates associated with our candidate models.
+This built on previous work, which focused only on positive relationships, by allowing us to validate the false positive *and* false negative rates associated with our candidate models [@Edwards:2015iz].
 
 Four phage and bacterial genomic markers were used to predict infectious relationships between bacteria and phages: 1) genome nucleotide similarities, 2) gene amino acid sequence similarities, 3) CRISPR targeting of phages by bacterial CRISPR spacer sequences, and 4) similarity of protein families known to be associated with known protein-protein interactions [@Orchard:2014hq].
-The resulting random forest model exhibited high performance with an AUC of `r signif(genmodelper$ROC, digits = sigfig)`, a sensitivity of `r signif(genmodelper$Sens, digits = sigfig)`, and a specificity of `r signif(genmodelper$Spec, digits = sigfig)` **(Figure \ref{RocCurve} A)**.
+The resulting random forest model performed well, with an AUC of `r signif(genmodelper$ROC, digits = sigfig)`, a sensitivity of `r signif(genmodelper$Sens, digits = sigfig)`, and a specificity of `r signif(genmodelper$Spec, digits = sigfig)` **(Figure \ref{RocCurve} A)**.
 The most important predictor in the model was nucleotide similarity between genes, followed by nucleotide similarity of whole genomes **(Figure \ref{RocCurve} B)**.
 Protein family interactions were moderately important to the model, while CRISPRs were minimally important, likely because they were redundant with the blast information.
-Approximately one third of the relationships yielded no score and were automatically classified as being non-infectious **(Figure \ref{RocCurve} C)**.
+Approximately one third of the relationships yielded no score and were therefore unknown, unable to be assigned an interaction **(Figure \ref{RocCurve} C)**.
 
 We used our random forest model to classify the relationships between bacteria and phages in the experimental datasets.
-The relationships within the three studies were used to construct one master network, containing the three study sub-networks, which themselves each contain sub-networks for each sample **(Figure \ref{RocCurve} D)**.
+The relationships within the three studies were used to construct one master network, containing the three study sub-networks, which themselves each contained sub-networks for each sample **(Figure \ref{RocCurve} D)**.
 Metadata including study, sample ID, disease, and abundance within the community (based on sequence count) were also stored in the multi-study master network to allow for effective parsing and downstream analysis **(Supplemental Figure \ref{NetworkDiagram})**.
@@
 Although the network was highly connected, not all relationships were present in all samples.
-Furthermore, relationships were weighted by relative abundance of their associated bacteria and phage, meaning that lowly abundant relationships could be present but insignificant compared to those that were more highly abundant.
+Furthermore, relationships were weighted by relative abundance of their associated bacteria and phages, meaning lowly abundant relationships could be present but insignificant compared to those that were highly abundant.
 Like the master network, the skin network exhibited a diameter of `r prettyNum(genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Diameter"], big.mark = ",")` (measure of graph size; the greatest number of traversed vertices required between two vertices) and included almost all of the master network nodes and edges **(Figure \ref{RocCurve} E - F)**.
-The gut diet and twin sample sets each contained less than 150 vertices, less than 20,000 relationships, and diameters of `r prettyNum(genfigurestats[c(genfigurestats$class %in% "DietStudy"),"Diameter"], big.mark = ",")`, suggesting more sparsely related phages and bacteria **(Figure \ref{RocCurve} E - F)**.
+The gut diet and twin sample sets each contained fewer than 150 vertices, fewer than 20,000 relationships, and diameters of `r prettyNum(genfigurestats[c(genfigurestats$class %in% "DietStudy"),"Diameter"], big.mark = ",")`, suggesting more sparsely related phages and bacteria **(Figure \ref{RocCurve} E - F)**.
 
 ## Individuality of Microbial Networks
-Previous work has show community membership and diversity are highly personal, with signatures remaining more similar to themselves over time compared to other people [@Grice:2009ee; @Hannigan:2015fz; @Minot:2013ih].
-Using our microbiome network model, we determined whether this personal conservation extended to network structure.
+Skin and gut community membership and diversity are highly personal, with people remaining more similar to themselves over time compared to other people [@Grice:2009ee; @Hannigan:2015fz; @Minot:2013ih].
+We determined whether this personal conservation extended to network structure using our microbiome network model.
 We calculated the degree of dissimilarity between each subject's network, based on phage and bacteria abundance, as well as centrality.
@@
 Conceptually, this metric defines central phages as those that are highly abundant and infect many bacteria which themselves are abundant and infected by many other phages.
-Bacterial centrality was defined in the same way.
-We then calculated the similarity of community networks using the overall weighted eigenvector centrality of all nodes between all samples.
+Bacterial centrality was defined using the same metric.
+We then calculated the similarity of community networks using the weighted eigenvector centrality of all nodes between all samples.
 Biologically, samples with similar network structures are interpreted as having similar capacities to influence other microbes within the community, transmit genetic material (e.g.
@@
 
-We found that gut microbiome network structure was highly individual specific, with networks clustering significantly by person (anosim p-value = `r signif(interstats[c(interstats$site %in% "DietAnosim"), "prob"], digits = sigfig)`, anosim statistic R = `r signif(interstats[c(interstats$site %in% "DietAnosim"), "abserr"], digits = sigfig)`, **Figure \ref{intradiv} A)**.
-Network dissimilarity within people over the 8-10 day sampling period was significantly less than the average dissimilarity between that person and others **(Figure \ref{intradiv} B)**.
-We evaluated the significance of an individual's network being more similar to itself than others by fitting a probability distribution to the reductions in intrapersonal dissimilarity compared to interpersonal dissimilarity (Lines between points in Figure \ref{intradiv} B - D).
-By taking the integral of the probability distribution from negative infinity to zero, we were able to calculate the probability that an individual's network structure was more similar to itself than others, in a given population.
-Thus a probability of 1 would mean that all members in a given population will be expected to be more similar to themselves compared to others, and 0.5 means they are no more likely to be more similar to themselves than dissimilar.
-This is a Bayesian approach that tells us how prevalent we expect this trend to be.
-The probability of intrapersonal diversity of an individual being less than interpersonal diversity in our dataset was `r signif(interstats[c(interstats$site %in% "Diet"), "prob"], digits = sigfig)` (absolute error = `r signif(interstats[c(interstats$site %in% "Diet"), "abserr"], digits = sigfig)`, **Figure \ref{intradiv} B)**.
+We used this network dissimilarity metric to test whether microbiome network structures were more similar within people over time compared to other people.
+We found that gut microbiome network structure clustered by person (ANOSIM p-value = `r signif(interstats[c(interstats$site %in% "DietAnosim"), "prob"], digits = sigfig)`, R = `r signif(interstats[c(interstats$site %in% "AS"), "prob"], digits = sigfig)`, **Figure \ref{intradiv} A**).
+We found evidence that network dissimilarity within people over the 8-10 day sampling period was less than the average dissimilarity between that person and others, although this was not statistically significant (p-value = `r signif(interstats[c(interstats$site %in% "Diet"), "prob"], digits = sigfig)`, **Figure \ref{intradiv} B**).
+The lack of statistical confidence was likely due to the small sample size of this dataset.
+Although there was evidence for gut network conservation among individuals, we found no evidence for conservation of gut network structures within families.
+The gut network structures were not more similar between twins and their mothers (intrafamily) compared to other families of twins and mothers (inter-family) (p-value = `r signif(interstats[c(interstats$site %in% "Skin"), "prob"], digits = sigfig)`, **Figure \ref{intradiv} C**).
 
-Skin microbiome network structure was also conserved within individuals, although not as strongly as the gut.
-The probability of a given skin site being more similar to itself after a month compared to a different person at the same time was `r signif(interstats[c(interstats$site %in% "Skin"), "prob"], digits = sigfig)` (absolute error = `r signif(interstats[c(interstats$site %in% "Skin"), "abserr"], digits = sigfig)`, **Figure \ref{intradiv} C)**.
-This distribution was similar when separated by anatomical sites, suggesting this was an accurate representation of overall skin network dissimilarity **(Supplemental Figure \ref{allskin})**.
-
-The conservation of gut network structures was only weakly associated with families.
-The gut network structures were more similar between twins and their mothers (intrafamily) compared to other families of twins and mothers (inter-family) **(Figure \ref{intradiv} D)**.
-The probability of a family member being more similar to another family member compared to non-family members was `r signif(interstats[c(interstats$site %in% "Twins"), "prob"], digits = sigfig)` (absolute error = `r signif(interstats[c(interstats$site %in% "Twins"), "abserr"], digits = sigfig)`).
+Skin microbiome network structure was strongly conserved within individuals, and more adequately confirmed by our larger skin dataset (p-value = `r signif(interstats[c(interstats$site %in% "Twins"), "prob"], digits = sigfig)`, **Figure \ref{intradiv} D**).
+This distribution was similar when separated by anatomical sites, with most sites also being significantly more conserved within individuals **(Supplemental Figure \ref{allskin})**.
 
@@
 Diet is a major environmental factor that influences resource availability and gut microbiome composition and diversity, including bacteria and phages [@Minot:2011ez; @Turnbaugh:2009hf; @David:2014cl].
-Previous work in isolated culture-based systems has suggested that changes in nutrient availability are associated with altered phage - bacteria network structures, although this has yet to be applied to humans [@Poisot:2011jc].
-Understanding the impact diet might have on gut microbiome networks is a valuable contribution to the baseline understanding we aim to achieve here.
-We therefore hypothesized that network structure would be altered by changes in diet.
-We also described a potential association between gut network structure and obesity, a disease linked to diet and potentially the microbiome [@Sze:2016iw].
+Previous work in isolated culture-based systems has suggested that changes in nutrient availability are associated with altered phage - bacteria network structures, although this has yet to be tested in humans [@Poisot:2011jc].
+We hypothesized that a change in diet would be associated with a change in network structure.
 
-We evaluated the differences in gut network structure by quantifying how central each sample's network was on average across all microbes.
+We evaluated the diet-associated differences in gut network structure by quantifying how central each sample's network was on average across all microbes.
 We accomplished this by utilizing two common centrality metrics: degree centrality and closeness centrality.
@@
 A higher closeness centrality suggests that genetic information or the effects of altered abundance would be more impactful to all other microbes in the system.
-Networks with higher closeness centrality also indicates an overall greater degree of connections, which indicates a greater resilience to instability.
+A networks with higher closeness centrality also indicates an overall greater degree of connections, which suggests a greater resilience to instability.
 Because these values are assigned to each phage and bacterium within each network, we calculated the average connectedness and corrected for the maximum potential degree of connectedness to obtain a single value for the connectedness of each graph.
@@
 We found that gut microbiome network structures associated with high fat diets were less connected than those of low fat diets **(Figure \ref{dietnetworks} A-B)**.
-Tests for statistical differences were not performed, so as to prevent misleading interpretations from the small sample size.
-High fat diets exhibited less degree centrality **(Figure \ref{dietnetworks} A)**, meaning bacteria were overall targeted by less phages and phage tropism was more specific.
-High fat diets also exhibited decreased closeness centrality **(Figure \ref{dietnetworks} B)**, meaning the microbes were more distant from other microbes in the community, making genetic information transfer and the impact of altered abundance less capable of impacting other bacteria and phages within the network.
+Tests for statistical differences were not performed due to the small sample size.
+High fat diets exhibited less degree centrality **(Figure \ref{dietnetworks} A)**.
+This suggests that bacteria were targeted by less phages and that phage tropism was more specific.
+High fat diets also exhibited decreased closeness centrality **(Figure \ref{dietnetworks} B)**, indicating that microbes were more distant from other microbes in the community.
+This would make genetic transfer and altered abundance less capable of impacting other bacteria and phages within the network.
 
@@
 The obesity-associated network demonstrated a higher degree centrality **(Figure \ref{dietnetworks} C)** but a less closeness centrality, compared to the healthy controls **(Figure \ref{dietnetworks} D)**.
-This meant that the obesity network was overall less connected, having microbes further from all other microbes within the community.
+This suggested that the obesity network was overall less connected, having microbes further from all other microbes within the community.
 
@@
 These communities vary by degree of skin moisture, oil, and environmental exposure.
-We hypothesized that like microbial composition and diversity, microbial network structure differs between anatomical sites.
-We addressed this hypothesis by evaluating the changes in network structure between anatomical sites within our skin dataset.
+We hypothesized that like microbial composition and diversity, microbial network structure differed between anatomical sites.
+We tested this hypothesis by evaluating the changes in network structure between anatomical sites within our skin dataset.
 
-We quantified the average centrality of each sample using the weighted eigenvector centrality metric, which yields a larger value for phages and bacteria that are more abundant and have more relationships to other bacteria and phages that themselves have more relationships.
+We quantified the average centrality of each sample again using the weighted eigenvector centrality metric.
 We found that intermittently moist skin sites (dynamic sites that fluctuate between being moist and dry) were significantly less connected than the more stable moist and sebaceous environments (p-value < 0.001, **Figure \ref{skinnetwork} A)**.
@@
 # Discussion
-We developed and implemented a network-based analytical approach that allowed us to begin evaluating the basic properties of the human microbiome through bacteria and phage relationships, instead of membership or diversity of separate communities.
+We developed and implemented a network-based analytical approach that allowed us to begin evaluating the basic properties of the human microbiome through bacteria and phage *relationships*, instead of membership or diversity.
 The goal of this study was to provide an initial understanding of how phage-bacteria networks differ throughout the human body, so as to provide a baseline for future studies of how microbiome networks differ in disease states.
-This goal is similar to the initial studies of bacteria and viral diversity across the human body, as well as other environments [@Grice:2009eea; @Findley:2013jf; @Hannigan:2015fz; @Costello:2009im, @Consortium:2012iz; @Schloss:2005hz; @Minot:2011ez].
-Our approach was advantageous over previous microbiome network analyses because it did not rely on inappropriate linear correlations of abundance, it was trained on both positive and negative relationships, and it involved utilization of network theory concepts.
+This goal was similar to the initial studies of bacteria and viral diversity across the human body, as well as other environments [@Grice:2009eea; @Findley:2013jf; @Hannigan:2015fz; @Costello:2009im, @Consortium:2012iz; @Schloss:2005hz; @Minot:2011ez].
+The use of network theory offered extensive analytical opportunities which could be applied to understand complex ecological communities.
+We focused on utilizing metrics of connectivity to understand the extent to which communities of bacteria and phages interact (e.g.
+horizontal gene transfer, modulated bacterial gene expression, alterations in abundance).
+By accomplishing this goal, we aimed to provide an initial understanding of human microbiome networks that will power fututure studies to use similar approaches in other setting such as disease states.
 
-The principles of network theory offer extensive analytical opportunities which can be applied to understand complex ecological communities.
-In this study we focused on utilizing metrics of connectivity to understand the extent to which communities of bacteria and phages interact (e.g.
-horizontal gene transfer, modulated bacterial gene expression, alterations in abundance), and how resilient they are to instability.
-Like earlier work laid a foundational understanding of the healthy human microbiome using membership and diversity measurements [@Grice:2009eea; @Findley:2013jf; @Hannigan:2015fz; @Costello:2009im, @Consortium:2012iz], we provide a similar understanding of healthy human microbiome network structures.
-These properties included variations between individuals and families, different diet types, and different anatomical skin sites.
+We reported that, just as gut microbiome and virome composition and diversity are conserved in individuals [@Hannigan:2015fz; @Grice:2009eea; @Findley:2013jf; @Minot:2013ih], gut and skin microbiome network structures were also conserved within individuals over time.
+Gut network structure was not conserved among family members.
+These findings suggest that microbiome properties such as stability and the potential for horizontal gene transfer are personal, and potentially impacted by personal factors ranging from the body's immune system to environemntal conditions such as climate and diet.
+The ability of environmental conditions to shape gut and skin microbiome network structure was supported by our finding that diet and skin location were assocaited with altered network structures.
 
-One major overarching principle we observed was that, just as gut microbiome and virome composition and diversity are conserved in individuals [@Hannigan:2015fz; @Grice:2009eea; @Findley:2013jf; @Minot:2013ih], gut and skin microbiome network structures were also conserved within individuals over time, with the gut being more strongly similar than the skin.
-Network structure was weakly conserved among family members.
-One explanation for the weaker intrapersonal network conservation of skin compared to the gut communities is that the skin is in more direct contact with the external environment.
-This was further supported by the observation that network structure differed between environmentally exposed and occluded skin sites.
-The exposed and intermittently exposed sites were less connected and therefore are expected to have a higher propensity for instability.
-Likewise, intermittently moist sites demonstrated less connectedness than the more stable moist and sebaceous sites.
-This suggests that body sites under greater degrees of fluctuation harbored less connected, potentially less stable microbiomes.
-
-We also observed a link between diet and microbiome network structure.
+We found evidence that diet was sufficient to alter gut microbiome network connectivity.
 Although our sample size was small, we found that high fat diets appeared to be less connected than low fat diets.
@@
 We supported this finding with the observation that obesity is associated with decreased network connectivity.
-Further work will be required to further characterize these relationships beyond our small study, but our results allow us to speculate that the food we eat may not only impact what microbes colonize our guts, but may also impact their ability to communicate and maintain stability.
+Further work will be required to characterize these relationships with a larger cohort, but our results allow us to speculate that the food we eat may not only impact what microbes colonize our guts, but may also impact interactions.
 
-Together this work represents an initial step toward understanding the microbiome through its relationships.
+We also found evidence that skin microbiome network structure varied by skin environments.
+We observed that network structure differed between environmentally exposed and occluded skin sites.
+The sites under greater environmental fluctuation and exposure, the exposed and intermittently exposed sites, were less connected and therefore were expected to have a higher propensity for instability.
+Likewise, intermittently moist sites demonstrated less connectedness than the more stable moist and sebaceous sites.
+Together this suggested that body sites under greater degrees of fluctuation harbored less connected, potentially less stable microbiomes.
+
+This work represents an initial step toward understanding the microbiome through its relationships.
 While these findings are informative, there are certainly caveats that should be noted.
@@
 Just as we have improved on previous modeling efforts, we expect that new and creative scoring metrics will likely be integrated into this model to further improve model performance.
-
 *Second*, while informative, this work was done retrospectively and relied on published research from as many as seven years ago.
@@
 While we were able to present initial observations, follow-up studies will be required to validate our observations.
+Despite their limitations, these results will be important for informing design and interpretation of future studies of the human microbiome.
 
-Despite their limitations, these results will be important for informing design and interpretation of future studies of the human microbiome.
 By showing that microbiome relationship structure, and therefore the community potential for dissemination of information and influence, as well as the potential for maintaining stability, differs significantly between body sites and environmental conditions we demonstrate the diversity of relationships across the human body.
@@
 (B) Quantification of gut network dissimilarity within the same subject over time (intrapersonal) and the mean dissimilarity between the subject of interest and all other subjects (interpersonal).
-The density curve is the probability distribution of the changes between intrapersonal and interpersonal diversity, representing the probability that intrapersonal dissimilarity will be lower than interpersonal dissimilarity (lower intrapersonal dissimilarity defined as change less than zero).
-(C) Quantification of skin network dissimilarity within the same subject and anatomical location over time (intrapersonal) and the mean dissimilarity between the subject of interest and all other subjects at the same time and the same anatomical location (interpersonal).
-Probability distribution the same as for panel B.
-(D) Quantification of gut network dissimilarity within subjects from the same family (intrafamily) and the mean dissimilarity between subjects within a family and those of other families (interfamily).
-Probability distribution the same as for panel B.*\label{intradiv}](../figures/intrapersonal_diversity.pdf){ width=90% }
+The p-value is also provided.
+(C) Quantification of gut network dissimilarity within subjects from the same family (intrafamily) and the mean dissimilarity between subjects within a family and those of other families (interfamily).
+The p-value is also provided.
+(D) Quantification of skin network dissimilarity within the same subject and anatomical location over time (intrapersonal) and the mean dissimilarity between the subject of interest and all other subjects at the same time and the same anatomical location (interpersonal).
+The p-value is also provided.
+*\label{intradiv}](../figures/intrapersonal_diversity.pdf){ width=90% }
 

### 2017-03-28 doc/manuscript.Rmd

@@
 These findings suggest that microbiome properties such as stability and the potential for horizontal gene transfer are personal, and potentially impacted by personal factors ranging from the body's immune system to environemntal conditions such as climate and diet.
-The ability of environmental conditions to shape gut and skin microbiome network structure was supported by our finding that diet and skin location were assocaited with altered network structures.
+The ability of environmental conditions to shape gut and skin microbiome network structure was supported by our finding that diet and skin location were associated with altered network structures.
 
@@
 ## Centrality Analysis
-Degree and closeness centrality were calculated using the associated functions within the igraph R package (v1.0.1) [@Theigraphsoftware:vh].
-Briefly, the **closeness centrality** of Vi was calculated taking the inverse of the average length of the shortest path (`d`) between nodes Vi and the `k` other nodes Vj, such that closeness centrality of node Vi is:
+Let $G(V,E)$ be an undirected, unweighted graph with $|V|=n$ nodes and $|E|=m$ edges.
+Also, let $\mathbf{A}$ be its corresponding adjacency matrix with entries $a_{ij} = 1$ if nodes $V_i$ and $V_j$ are connected via an edge, and $a_{ij} = 0$ otherwise.
 
-$$ { C }_{ C }\left( { V }_{ i } \right) ={ \left( \sum _{ j=1 }^{ k }{ d\left( { V }_{ i },{ V }_{ j } \right)  }  \right)  }^{ -1 } $$
+Briefly, the **closeness centrality** of node $V_i$ is calculated taking the inverse of the average length of the shortest paths (`d`) between nodes $V_i$ and all the other nodes $V_j$.
+Mathematically, the closeness centrality of node $V_i$ is given as:
 
-The distance between nodes (`d`) was calculated as the shortest number of edges required to be traversed to move from one node to another.
+$$ { C }_{ C }\left( { V }_{ i } \right) ={ \left( \sum _{ j=1 }^{ n }{ d\left( { V }_{ i },{ V }_{ j } \right)  }  \right)  }^{ -1 } $$
 
-The simple metric of **degree centrality** of node Vi was defined as the sum of the number of k edges Ei associated with that node, such that:
+[//]: <> The distance between nodes (`d`) was calculated as the shortest number of edges required to be traversed to move from one node to another.
 
-$$ { C }_{ D }\left( { V }_{ i } \right) =\sum _{ i=1 }^{ k }{ { E }_{ i } } $$
+Intuitively, the **degree centrality** of node $V_i$ is defined as the number of edges that are adjacent to that node:
 
-The Eigenvector centrality was calculated using the values of the first eigen vector of the associated adjacency matrix that are associated with each vertex `Vi`.
-Conceptually, this function results in a centrality value that reflects the connections of the vertex, as well as the centrality of the connected vertices.
+$$ { C }_{ D }\left( { V }_{ i } \right) =\sum _{ j=1 }^{ n }{ { a }_{ ij } } $$
+
+where $a_{ij}$ is the $ij^{th}$ entry in the adjacency matrix $\mathbf{A}$.
+
+[//]: <> The simple metric of degree centrality of node Vi was defined as the sum of the number of k edges Ei associated with that node, such that:
+
+The Eigenvector centrality of node $V_i$ is defined as the $i^{th}$ value in the first eigenvector of the associated adjacency matrix $\mathbf{A}$.
+Conceptually, this function results in a centrality value that reflects the connections of the vertex, as well as the centrality of its neighboring vertices.
 
 The **centralization** metric was used to assess the average centrality of each sample graph `G`.
-Centralization was calculated by taking the sum of each vertex `Vi` centrality from the graph maximum centrality `Cw`, such that:
+Centralization was calculated by taking the sum of each vertex `Vi`'s centrality from the graph maximum centrality `Cw`, such that:
 
-$$ C\left( G \right) =\frac { \sum _{ i=1 }^{ k }{ \max _{ w }{ c\left( w \right)  } -c\left( { v }_{ i } \right)  }  }{ { T } }  $$
+$$ C\left( G \right) =\frac { \sum _{ i=1 }^{ n }{ Cw -c\left( { v }_{ i } \right)  }  }{ { T } }  $$
 
 The values were corrected for uneven graph sizes by dividing the centralization score by the maximum theoretical centralization (`T`) for a graph with the same number of vertices.
+
+Degree and closeness centrality were calculated using the associated functions within the igraph R package (v1.0.1) [@Theigraphsoftware:vh].
 
@@
 
-Where `Cij` is the sum of the lesser centrality values for those vertices shared between graphs, and `Ci` and `Cj` are are the total number of vertices found in each graph.
-This allows us calculate the dissimilarity between graphs based on the shared centrality values between the two graphs.
+Where `Cij` is the sum of the lesser centrality values for those vertices shared between graphs, and `Ci` and `Cj` are the total number of vertices found in each graph.
+This allows us to calculate the dissimilarity between graphs based on the shared centrality values between the two graphs.
 

### 2017-03-30 doc/manuscript.Rmd
More updates.

@@
 We found that gut and skin network structure was person-specific, and was not conserved among cohabitating family members.
-High fat diets and obesity were associated with less connected networks.
+High-fat diets and obesity were associated with less connected networks.
 There were significant differences in network structure between skin sites, with those exposed to the external environment being less connected and more prone to instability.
@@
 The conventional approach to understanding the human microbiome is unable to capture how these communities interact and influence each other, but relationship-based network approaches allow us to capture this information.
-We aimed to bridge this knowledge gap by characterizing human bacterial and phage communities by their *relationships*, leveraging machine learning and graph theory techniques.
-We focused on characterizing the network diversity of the healthy human microbiome (including the skin and gut), so as to provide a foundation on which we and others can build further studies into the network dynamics associated with disease states.
+We aimed to bridge this knowledge gap by characterizing human bacterial and phage communities by their relationships, leveraging machine learning and graph theory techniques.
+We focused on characterizing the network diversity of the human microbiome (including skin and gut) to provide a foundation for further studies of disease network dynamics.
 
@@
 Our study contained three datasets, including a study of the impact of diet on the healthy human gut virome [@Minot:2011ez], the impact of anatomical location on the healthy human skin virome [@Hannigan:2015fz], and the viromes of monozygotic twins and their mothers [@Reyes:2010cwa; @Turnbaugh:2009ei].
-The viromes associated with these datasets were subjected to virus-like particle (VLP) purification to eliminate other organismal DNA including bacteria, fungi, and humans, and allowed us to assess the actively replicating virome.
+The viromes associated with these datasets were all subjected to virus-like particle (VLP) purification to eliminate other organismal DNA including bacteria, fungi, and humans, thus allowing the original authors and ourselves to assess the actively replicating virome.
 
-The publications we utilized were conducted over the span of five years using different methods and technologies, and therefore yielded different sequence abundances **(Supplemental Figure \ref{SequenceStats} A-B)**.
+The publications we utilized were published over the span of five years using different methods and technologies, and therefore yielded different sequence abundances **(Supplemental Figure \ref{SequenceStats} A-B)**.
 The bacterial and viral sequences from those publications were quality filtered and assembled into contigs.
@@
 Because this was a DNA sequencing study, RNA phages were not considered in the analysis **(Supplemental Figure \ref{ValidationOverview} C-D)**.
-This training set included both positive relationships (a phage infects a bacterium) and negative relationships (a phage does not infect a bacterium).
+This training set included both positive relationships (phage infects a bacterium) and negative relationships (phage does not infect a bacterium).
 This built on previous work, which focused only on positive relationships, by allowing us to validate the false positive *and* false negative rates associated with our candidate models [@Edwards:2015iz].
@@
 The most important predictor in the model was nucleotide similarity between genes, followed by nucleotide similarity of whole genomes **(Figure \ref{RocCurve} B)**.
-Protein family interactions were moderately important to the model, while CRISPRs were minimally important, likely because they were redundant with the blast information.
+Protein family interactions were moderately important to the model, while CRISPRs were minimally important, likely because they were redundant with the BLAST information.
 Approximately one third of the relationships yielded no score and were therefore unknown, unable to be assigned an interaction **(Figure \ref{RocCurve} C)**.
@@
 We determined whether this personal conservation extended to network structure using our microbiome network model.
-We calculated the degree of dissimilarity between each subject's network, based on phage and bacteria abundance, as well as centrality.
+We calculated the degree of dissimilarity between each subject's network based on phage and bacteria abundance, as well as centrality.
 Centrality was evaluated by first calculating the weighted eigenvector centrality of all bacteria and phages within each sample graph.
@@
 Although there was evidence for gut network conservation among individuals, we found no evidence for conservation of gut network structures within families.
-The gut network structures were not more similar between twins and their mothers (intrafamily) compared to other families of twins and mothers (inter-family) (p-value = `r signif(interstats[c(interstats$site %in% "Skin"), "prob"], digits = sigfig)`, **Figure \ref{intradiv} C**).
+The gut network structures were not more similar between twins and their mothers (intrafamily) compared to other families of twins and mothers (inter-family) (p-value = `r signif(interstats[c(interstats$site %in% "Twins"), "prob"], digits = sigfig)`, **Figure \ref{intradiv} C**).
 
-Skin microbiome network structure was strongly conserved within individuals, and more adequately confirmed by our larger skin dataset (p-value = `r signif(interstats[c(interstats$site %in% "Twins"), "prob"], digits = sigfig)`, **Figure \ref{intradiv} D**).
+Skin microbiome network structure was strongly conserved within individuals, and more adequately confirmed by our larger skin dataset (p-value = `r signif(interstats[c(interstats$site %in% "Skin"), "prob"], digits = sigfig)`, **Figure \ref{intradiv} D**).
 This distribution was similar when separated by anatomical sites, with most sites also being significantly more conserved within individuals **(Supplemental Figure \ref{allskin})**.
@@
 
-We evaluated the diet-associated differences in gut network structure by quantifying how central each sample's network was on average across all microbes.
+We evaluated the diet-associated differences in gut network structure by quantifying how central each sample's network was on average.
 We accomplished this by utilizing two common centrality metrics: degree centrality and closeness centrality.
-**Degree centrality**, the simplest centrality metric, was defined as the number of connections each phage made to bacteria, or each bacterium made to phages.
+Degree centrality, the simplest centrality metric, was defined as the number of connections each phage made to bacteria, or each bacterium made to phages.
 Because this metric alone offers only minimal insight, we supplemented it with measurements of closeness centrality.
-**Closeness centrality** is a metric of how close each phage or bacterium is to all of the other phages and bacteria in the network.
+Closeness centrality is a metric of how close each phage or bacterium is to all of the other phages and bacteria in the network.
 A higher closeness centrality suggests that genetic information or the effects of altered abundance would be more impactful to all other microbes in the system.
-A networks with higher closeness centrality also indicates an overall greater degree of connections, which suggests a greater resilience to instability.
+A network with higher average closeness centrality also indicates an overall greater degree of connections, which suggests a greater resilience to instability.
 Because these values are assigned to each phage and bacterium within each network, we calculated the average connectedness and corrected for the maximum potential degree of connectedness to obtain a single value for the connectedness of each graph.
 
-We found that gut microbiome network structures associated with high fat diets were less connected than those of low fat diets **(Figure \ref{dietnetworks} A-B)**.
+We found that gut microbiome network structures associated with high-fat diets were less connected than those of low-fat diets **(Figure \ref{dietnetworks} A-B)**.
 Tests for statistical differences were not performed due to the small sample size.
-High fat diets exhibited less degree centrality **(Figure \ref{dietnetworks} A)**.
-This suggests that bacteria were targeted by less phages and that phage tropism was more specific.
-High fat diets also exhibited decreased closeness centrality **(Figure \ref{dietnetworks} B)**, indicating that microbes were more distant from other microbes in the community.
+High-fat diets exhibited less degree centrality **(Figure \ref{dietnetworks} A)**.
+This finding suggests that bacteria were targeted by less phages and that phage tropism was more specific.
+High-fat diets also exhibited decreased closeness centrality **(Figure \ref{dietnetworks} B)**, indicating that microbes were more distant from other microbes in the community.
 This would make genetic transfer and altered abundance less capable of impacting other bacteria and phages within the network.
 
-In addition to diet, there was also an association between obesity and network structure **(Figure \ref{dietnetworks} C-D)**.
-The obesity-associated network demonstrated a higher degree centrality **(Figure \ref{dietnetworks} C)** but a less closeness centrality, compared to the healthy controls **(Figure \ref{dietnetworks} D)**.
-This suggested that the obesity network was overall less connected, having microbes further from all other microbes within the community.
+In addition to diet, we also observed an association between obesity and network structure **(Figure \ref{dietnetworks} C-D)**.
+The obesity-associated network demonstrated a higher degree centrality **(Figure \ref{dietnetworks} C)** but less closeness centrality compared to the healthy controls **(Figure \ref{dietnetworks} D)**.
+These results suggest that the obesity network was overall less connected, having microbes further from all other microbes within the community.
 
@@
 
-We quantified the average centrality of each sample again using the weighted eigenvector centrality metric.
+We quantified the average centrality of each sample using the weighted eigenvector centrality metric.
 We found that intermittently moist skin sites (dynamic sites that fluctuate between being moist and dry) were significantly less connected than the more stable moist and sebaceous environments (p-value < 0.001, **Figure \ref{skinnetwork} A)**.
@@
 We also found that occluded sites were significantly different from exposed and intermittently occluded sites, but there was no difference between exposed and intermittently occluded sites **(Figure \ref{skinnetwork} D,F)**.
+These findings provide further support that skin microbiome network structure differs significantly between skin sites.
 
 # Discussion
-We developed and implemented a network-based analytical approach that allowed us to begin evaluating the basic properties of the human microbiome through bacteria and phage *relationships*, instead of membership or diversity.
+We developed and implemented a network-based analytical approach which we used to begin evaluating the basic properties of the human microbiome through bacteria and phage *relationships*, instead of membership or diversity.
 The goal of this study was to provide an initial understanding of how phage-bacteria networks differ throughout the human body, so as to provide a baseline for future studies of how microbiome networks differ in disease states.
-This goal was similar to the initial studies of bacteria and viral diversity across the human body, as well as other environments [@Grice:2009eea; @Findley:2013jf; @Hannigan:2015fz; @Costello:2009im, @Consortium:2012iz; @Schloss:2005hz; @Minot:2011ez].
-The use of network theory offered extensive analytical opportunities which could be applied to understand complex ecological communities.
+This goal of providing a baseline understanding of the human microbiome was similar to the goals of initial studies of bacteria and viral diversity across the human body, as well as other environments [@Grice:2009eea; @Findley:2013jf; @Hannigan:2015fz; @Costello:2009im, @Consortium:2012iz; @Schloss:2005hz; @Minot:2011ez].
+Through the use of network theory, we leveraged extensive analytical opportunities which could be applied to understand complex ecological communities.
 We focused on utilizing metrics of connectivity to understand the extent to which communities of bacteria and phages interact (e.g.
 horizontal gene transfer, modulated bacterial gene expression, alterations in abundance).
-By accomplishing this goal, we aimed to provide an initial understanding of human microbiome networks that will power fututure studies to use similar approaches in other setting such as disease states.
+By pursuing this goal, we aimed to provide an initial understanding of human microbiome networks that will power future studies to use similar approaches in other setting such as disease states.
 
@@
 Gut network structure was not conserved among family members.
-These findings suggest that microbiome properties such as stability and the potential for horizontal gene transfer are personal, and potentially impacted by personal factors ranging from the body's immune system to environemntal conditions such as climate and diet.
+These findings suggest that microbiome properties such as stability and the potential for horizontal gene transfer are personal, and potentially impacted by personal factors ranging from the body's immune system to environmental conditions such as climate and diet.
 The ability of environmental conditions to shape gut and skin microbiome network structure was supported by our finding that diet and skin location were associated with altered network structures.
@@
 We found evidence that diet was sufficient to alter gut microbiome network connectivity.
-Although our sample size was small, we found that high fat diets appeared to be less connected than low fat diets.
-This suggested that high fat diets may lead to less stable communities with a decreased ability to influence each other.
-We supported this finding with the observation that obesity is associated with decreased network connectivity.
-Further work will be required to characterize these relationships with a larger cohort, but our results allow us to speculate that the food we eat may not only impact what microbes colonize our guts, but may also impact interactions.
+Although our sample size was small, we found evidence that high-fat diets were less connected than low-fat diets.
+This suggested that high-fat diets may lead to less stable communities with a decreased ability to influence each other.
+We supported this finding with the observation that obesity was associated with decreased network connectivity.
+Our results allow us to speculate that the food we eat may not only impact what microbes colonize our guts, but may also impact their interactions.
+Further work will be required to characterize these relationships with a larger cohort.
 
-We also found evidence that skin microbiome network structure varied by skin environments.
+In addition to diet, we found evidence that skin microbiome network structure varied by skin environment.
 We observed that network structure differed between environmentally exposed and occluded skin sites.
@@
 Likewise, intermittently moist sites demonstrated less connectedness than the more stable moist and sebaceous sites.
-Together this suggested that body sites under greater degrees of fluctuation harbored less connected, potentially less stable microbiomes.
+Thus body sites under greater degrees of fluctuation harbored less connected, potentially less stable microbiomes.
 
-This work represents an initial step toward understanding the microbiome through its relationships.
+Our results represent a step toward understanding the microbiome through interspecies relationships.
 While these findings are informative, there are certainly caveats that should be noted.
@@
 These archived datasets were limited by the technology and costs of the time, meaning the datasets are poorly powered for the statistical analysis we strive for today.
-While we were able to present initial observations, follow-up studies will be required to validate our observations.
-Despite their limitations, these results will be important for informing design and interpretation of future studies of the human microbiome.
+While we were able to present initial conclusions, follow-up studies will be required to validate our findings.
+Despite their limitations, these methods and results will be important for informing design and interpretation of future studies of the human microbiome.
 
-By showing that microbiome relationship structure, and therefore the community potential for dissemination of information and influence, as well as the potential for maintaining stability, differs significantly between body sites and environmental conditions we demonstrate the diversity of relationships across the human body.
+We demonstrated the diversity of relationships across the human body by showing that microbiome relationship structure, and therefore the potential for genetic transfer and stability maintenance, differs significantly between body sites and environmental conditions.
 We found that other environmental factors, such as diet, also influence relationship structures.
 This information builds on previous work by suggesting that bacteria and phages not only preferentially colonize different body sites, but also interact differently, allowing for different communication structures and capacities for maintaining stability.
-This work will also provide a foundation for future studies to begin evaluating how microbiome network dynamics change in disease states, and how the information can be leveraged therapeutically.
+Our results will also provide a foundation for future studies to begin evaluating how microbiome network dynamics change in disease states, and how the information can be leveraged therapeutically.
 
@@
 
-OGU abundance (`Ao`) was obtained as the sum of the abundance of each contig (`Aj`) associated with that OGU. The abundance values were length corrected such that:
+OGU abundance ($A_{O}$) was obtained as the sum of the abundance of each contig ($A_{j}$) associated with that OGU. The abundance values were length corrected such that:
 
@@
 Resulting spacer sequences were filtered to exclude spacers shorter than 20 bp and longer than 65 bp.
-Spacer sequences were aligned to the phage genomes using the nucleotide blast algorithm with default parameters (v2.4.0) [@Camacho:2009fc].
+Spacer sequences were aligned to the phage genomes using the nucleotide BLAST algorithm with default parameters (v2.4.0) [@Camacho:2009fc].
 The mean percent identity for each matching pair was recorded for use in our classification model.
@@
 Temperate bacteriophages infect and integrate into their bacterial host's genome.
-We detected integrated phage elements within bacterial genomes by aligning phage genomes to bacterial genomes using the nucleotide blast algorithm and a minimum e-value of 1e-10.
+We detected integrated phage elements within bacterial genomes by aligning phage genomes to bacterial genomes using the nucleotide BLAST algorithm and a minimum e-value of 1e-10.
 The resulting bitscore of each alignment was recorded for use in our classification model.
@@
 
-[//]: <> The distance between nodes (`d`) was calculated as the shortest number of edges required to be traversed to move from one node to another.
+The distance between nodes (`d`) was calculated as the shortest number of edges required to be traversed to move from one node to another.
 
@@
 
-[//]: <> The simple metric of degree centrality of node Vi was defined as the sum of the number of k edges Ei associated with that node, such that:
-
 The Eigenvector centrality of node $V_i$ is defined as the $i^{th}$ value in the first eigenvector of the associated adjacency matrix $\mathbf{A}$.
@@
 
-The **centralization** metric was used to assess the average centrality of each sample graph `G`.
-Centralization was calculated by taking the sum of each vertex `Vi`'s centrality from the graph maximum centrality `Cw`, such that:
+The **centralization** metric was used to assess the average centrality of each sample graph $\mathbf{G}$.
+Centralization was calculated by taking the sum of each vertex $V_{i}$'s centrality from the graph maximum centrality $C_{w}$, such that:
 
@@
 We assessed similarity between graphs by evaluating the shared centrality of their vertices, as has been done previously.
-More specifically, we calculated the dissimilarity between graphs `Gi` and `Gj` using the Bray-Curtis dissimilarity metric and Eigenvector centrality values such that:
+More specifically, we calculated the dissimilarity between graphs $G_{i}$ and $G_{j}$ using the Bray-Curtis dissimilarity metric and Eigenvector centrality values such that:
 
@@
 
-Where `Cij` is the sum of the lesser centrality values for those vertices shared between graphs, and `Ci` and `Cj` are the total number of vertices found in each graph.
+Where $C_{ij}$ is the sum of the lesser centrality values for those vertices shared between graphs, and $C_{i}$ and $C_{j}$ are the total number of vertices found in each graph.
 This allows us to calculate the dissimilarity between graphs based on the shared centrality values between the two graphs.
@@
 
-![**Summary of Multi-Study Network Model.** *(A) ROC curve resulting from ten iterations used to create the phage - bacteria infection prediction model.
+![**Summary of Multi-Study Network Model.** *(A) Average ROC curve resulting from ten iterations used to create the phage - bacteria infection prediction model.
 (B) Importance scores associated with the metrics used in the random forest model to predict relationships between bacteria and phages.
@@
 
-![**Impact of Diet and Obesity on Gut Network Structure.** *(A) Quantification of average degree centrality (number of edges per node) and (B) closeness centrality (average distance from each node to every other node) of gut microbiome networks of subjects limited to exclusively high fat or low fat diets.
+![**Impact of Diet and Obesity on Gut Network Structure.** *(A) Quantification of average degree centrality (number of edges per node) and (B) closeness centrality (average distance from each node to every other node) of gut microbiome networks of subjects limited to exclusively high-fat or low-fat diets.
 Lines represent the mean degree of centrality for each diet.

### 2017-03-31 doc/manuscript.Rmd
Changed repo name.
Updated with Jeniors edits.

@@
 Previous work has relied on studying bacterial and phage communities in isolation, reducing them to two separate communities.
-In reality, bacteria and phage communities are dynamic and complex, acting in concert to share genetic information and to maintain stable ecosystems, and removal of members can disrupt or even collapse those ecosystems [@Haerter:2014ii; @Harcombe:2005fd; @Middelboe:2001fl; @Poisot:2011jc; @Thompson:2012ki; @Moebus:1981kp; @Flores:2013hc; @Poisot:2012fh; @Poisot071993; @Flores:2011bh; @Jover:2015ev].
-The conventional approach to understanding the human microbiome is unable to capture how these communities interact and influence each other, but relationship-based network approaches allow us to capture this information.
+In reality, bacteria and phage communities are dynamic and complex.
+They frequently share genetic information and work together to maintain stable ecosystems.
+Removal of members can disrupt or even collapse those ecosystems [@Haerter:2014ii; @Harcombe:2005fd; @Middelboe:2001fl; @Poisot:2011jc; @Thompson:2012ki; @Moebus:1981kp; @Flores:2013hc; @Poisot:2012fh; @Poisot071993; @Flores:2011bh; @Jover:2015ev].
+The conventional approach to understanding the human microbiome has been unable to capture how these communities interact and influence each other, but relationship-based network approaches would allow us to capture this information.
 We aimed to bridge this knowledge gap by characterizing human bacterial and phage communities by their relationships, leveraging machine learning and graph theory techniques.
-We focused on characterizing the network diversity of the human microbiome (including skin and gut) to provide a foundation for further studies of disease network dynamics.
+We focused on characterizing the network diversity of the human skin and gut microbiomes to provide a foundation for further studies of disease network dynamics and to gain broader insights into human microbiome network diversity across different body sites.
 
-We began characterizing human microbiome phage and bacteria networks by leveraged three published microbiome datasets with paired virus and bacterial metagenomic sequence sets [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei].
-These datasets included gut and skin samples, allowing us to gain broader insights into human microbiome network diversity across different body sites.
+We first investigated human microbiome phage and bacteria networks by utilizing three published microbiome datasets with paired virus and bacterial metagenomic sequence sets [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei].
 Our approach built off of previous large-scale phage-bacteria microbiome network analyses by inferring interactions using metagenomic datasets, instead of using culture-based techniques [@Moebus:1981kp; @Flores:2013hc].
-Our metagenomic interaction inference model was powered beyond previous models by its inclusion of protein interaction data, inclusion of negative interactions as well as positive, and the use of a more sophisticated machine learning algorithm instead of linear models [@Edwards:2015iz].
+This metagenomic interaction inference model was powered beyond previous models by its inclusion of protein interaction data, negative as well as positive interactions, and the use of a more sophisticated machine learning algorithm instead of simple linear models [@Edwards:2015iz].
 
-Using this approach, we were able to go beyond contemporary isolated methods to provide a basic understanding of the community network dynamics associated with healthy human phages and bacteria.
+Through this approach, we were able to go beyond contemporary isolated methods to provide a basic understanding of the community network dynamics associated with healthy human phages and bacteria.
 By building and utilizing a microbiome network, we found that different people, body sites, and anatomical locations not only support distinct microbiome membership and diversity [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei; @Grice:2009eea; @Findley:2013jf; @Costello:2009im, @Consortium:2012iz], but also support communities with distinct communication structures and propensities toward community instability.
-By understanding the healthy state of network structures across the human body, we will empower future studies to begin investigating how these community structures change in disease or otherwise altered states.
+By understanding the healthy state of network structures across the human body, we will empower future studies to begin investigating how these community structures change in disease and their overall impact on human health.
 
@@
 Our study contained three datasets, including a study of the impact of diet on the healthy human gut virome [@Minot:2011ez], the impact of anatomical location on the healthy human skin virome [@Hannigan:2015fz], and the viromes of monozygotic twins and their mothers [@Reyes:2010cwa; @Turnbaugh:2009ei].
-The viromes associated with these datasets were all subjected to virus-like particle (VLP) purification to eliminate other organismal DNA including bacteria, fungi, and humans, thus allowing the original authors and ourselves to assess the actively replicating virome.
+The viromes associated with these datasets were all subjected to virus-like particle (VLP) purification to eliminate other organismal DNA and allow for direct assessment of the actively replicating virome.
 
@@
 Because contig assembly most commonly returns genome fragments, we clustered related bacteria and phage contigs by k-mer frequency and co-abundance using the CONCOCT algorithm.
-These clusters represent operationally defined units of related bacteria and phage genomes that we defined as operational genomic units (OGUs).
+We interpreted these clusters as operational units of related bacteria and phage genomes that we defined as operational genomic units (OGUs).
 These OGUs are conceptually similar to the operational taxonomic unit (OTU) and operational protein family (OPF) definitions used for grouping highly similar 16S rRNA gene and open reading frame sequences, respectively [@Schloss:2008ht].
@@
 The most important predictor in the model was nucleotide similarity between genes, followed by nucleotide similarity of whole genomes **(Figure \ref{RocCurve} B)**.
-Protein family interactions were moderately important to the model, while CRISPRs were minimally important, likely because they were redundant with the BLAST information.
+Protein family interactions were moderately important to the model, while CRISPRs were minimally important.
 Approximately one third of the relationships yielded no score and were therefore unknown, unable to be assigned an interaction **(Figure \ref{RocCurve} C)**.
 
-We used our random forest model to classify the relationships between bacteria and phages in the experimental datasets.
+Next we used our random forest model to classify the relationships between bacteria and phages in the experimental datasets.
 The relationships within the three studies were used to construct one master network, containing the three study sub-networks, which themselves each contained sub-networks for each sample **(Figure \ref{RocCurve} D)**.
@@
 Like the master network, the skin network exhibited a diameter of `r prettyNum(genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Diameter"], big.mark = ",")` (measure of graph size; the greatest number of traversed vertices required between two vertices) and included almost all of the master network nodes and edges **(Figure \ref{RocCurve} E - F)**.
-The gut diet and twin sample sets each contained fewer than 150 vertices, fewer than 20,000 relationships, and diameters of `r prettyNum(genfigurestats[c(genfigurestats$class %in% "DietStudy"),"Diameter"], big.mark = ",")`, suggesting more sparsely related phages and bacteria **(Figure \ref{RocCurve} E - F)**.
+The gut diet and twin sample sets each contained fewer than 150 vertices, fewer than 20,000 relationships, and diameters of `r prettyNum(genfigurestats[c(genfigurestats$class %in% "DietStudy"),"Diameter"], big.mark = ",")`, which indicated more sparsely related phages and bacteria **(Figure \ref{RocCurve} E - F)**.
 
@@
 We determined whether this personal conservation extended to network structure using our microbiome network model.
-We calculated the degree of dissimilarity between each subject's network based on phage and bacteria abundance, as well as centrality.
-Centrality was evaluated by first calculating the weighted eigenvector centrality of all bacteria and phages within each sample graph.
-Conceptually, this metric defines central phages as those that are highly abundant and infect many bacteria which themselves are abundant and infected by many other phages.
-Bacterial centrality was defined using the same metric.
+To accomplish this, we calculated the degree of dissimilarity between each subject's network based on phage and bacteria abundance, as well as centrality.
+Weighted eigenvector centrality was calculated for all bacteria and phages within each sample graph.
+This metric defines central phages as those that are highly abundant and infect many distinct bacteria which themselves are abundant and infected by many other phages.
+Similarly, bacterial centrality was defined as those bacteria that were both abundant and susceptible to numerous phages which infected many bacteria.
 We then calculated the similarity of community networks using the weighted eigenvector centrality of all nodes between all samples.
-Biologically, samples with similar network structures are interpreted as having similar capacities to influence other microbes within the community, transmit genetic material (e.g.
-horizontal gene transfer), and maintain stability.
+Samples with similar network structures were interpreted as having similar capacities for maintaining stability and transmitting genetic material.
 
@@
 Degree centrality, the simplest centrality metric, was defined as the number of connections each phage made to bacteria, or each bacterium made to phages.
-Because this metric alone offers only minimal insight, we supplemented it with measurements of closeness centrality.
+We more completely quantified the network properties by supplementing degree centrality with measurements of closeness centrality.
 Closeness centrality is a metric of how close each phage or bacterium is to all of the other phages and bacteria in the network.
@@
 This information builds on previous work by suggesting that bacteria and phages not only preferentially colonize different body sites, but also interact differently, allowing for different communication structures and capacities for maintaining stability.
-Our results will also provide a foundation for future studies to begin evaluating how microbiome network dynamics change in disease states, and how the information can be leveraged therapeutically.
+Our results will also provide a foundation for future studies to begin evaluating how microbiome network dynamics change in disease states, and how the information can be applied therapeutically.
 
@@
 
-https://github.com/SchlossLab/Hannigan-2016-ConjunctisViribus.
+https://github.com/SchlossLab/Hannigan_ConjunctisViribus_GenRes_2017
 

### 2017-04-18 doc/manuscript.Rmd
Most of Melissas edits at this point. A couple more to go though.

@@
 Viruses and bacteria are critical components of the human microbiome and play important roles in health and disease.
-Previous work has relied on studying bacteria and phages in isolation, reducing them to two separate communities.
-These approaches have not captured how these communities interact to share information (e.g.
-horizontal gene transfer) and maintain stability.
+Most previous work has relied on studying bacteria and phages in isolation, reducing them to two separate communities.
+Such approaches are not able to capture how these communities interact to maintain stability and share information (i.e.
+horizontal gene transfer).
 We developed and implemented a network-based analytical approach to provide an initial understanding of phage-bacteria network diversity throughout the human body.
@@
 We organized the predicted interactions into networks that allowed us to evaluate the diversity of phage-bacteria connectedness across the human body.
-We focused on establishing a foundational understanding of microbiome network structures across the human body.
-We found that gut and skin network structure was person-specific, and was not conserved among cohabitating family members.
+We found that gut and skin network structures were person-specific, and were not conserved among cohabitating family members.
 High-fat diets and obesity were associated with less connected networks.
 There were significant differences in network structure between skin sites, with those exposed to the external environment being less connected and more prone to instability.
-Altogether this study characterizes the diversity of microbiome networks across the human body, and provides a baseline for future studies to investigate the role of ecological networks in disease states.
+Together this study characterized the diversity of virome-microbiome networks across the human body and illustrated how environmental factors may influence phage-bacteria interactive dynamics.
+This work provides an baseline for future studies to investigate the role of ecological networks in disease states.
 
-**Word Count**: 236 / 250
+**Word Count**: 238 / 250
 
@@
 Bacterial communities have been associated with diseases including a wide range of skin conditions [@Hannigan:2013im], acute and chronic wound healing conditions [@Hannigan:2014be; @Loesche:2016ev], and gastrointestinal diseases including inflammatory bowel disease [@He:2016ch; @Norman:2015kb], *Clostridium difficile* infections [@Seekatz:2016fz], and colorectal cancer [@Zackular:2014fba; @Baxter:2014hb].
-Altered viromes (virus communities consisting primarily of bacteriophages in humans) have also been associated with various diseases and environmental perturbations including inflammatory bowel disease [@Norman:2015kb; @Manrique:2016dx], periodontal disease [@Ly:2014ew], spread of antibiotic resistance [@Modi:2013fia; @Hannigan:2015fz], and others [@Monaco:2016ita; @Hannigan:2015fz; @Minot:2011ez; @SantiagoRodriguez:2015gd; @Abeles:2015dy; @Abeles:2014kj].
-These shifts in virus community composition are not reflections of the bacterial communities, but rather act in concert with bacteria as one single overall community [@Norman:2015kb; @Haerter:2014ii].
-Individual phages within these communities are capable of lysing their bacterial hosts and modulating their functionality through horizontal gene transfer (i.e.
-transduction) and altering bacterial host gene expression.
+Altered viromes (virus communities consisting primarily of bacteriophages in humans) have also been associated with various diseases and environmental perturbations including inflammatory bowel disease [@Norman:2015kb; @Manrique:2016dx], periodontal disease [@Ly:2014ew], spread of antibiotic resistance [@Modi:2013fia], and others [@Monaco:2016ita; @Hannigan:2015fz; @Minot:2011ez; @SantiagoRodriguez:2015gd; @Abeles:2015dy; @Abeles:2014kj].
+These virus communities act in concert with bacteria as one single overall community [@Haerter:2014ii].
+Individual phages within these communities are capable of lysing their bacterial hosts, modulating their functionality through processes such as horizontal gene transfer (i.e.
+transduction) [@Moon:2015fa, @Modi:2013fi, @Ogg:1981th, @Frost:2005dn], and altering bacterial host gene expression [@Lindell:2005gz, @Tyler:2013fl, @Hargreaves:2014ja].
 
-Previous work has relied on studying bacterial and phage communities in isolation, reducing them to two separate communities.
+Most previous work has relied on studying bacterial and phage communities in isolation, reducing them to two separate communities.
 In reality, bacteria and phage communities are dynamic and complex.
@@
 Removal of members can disrupt or even collapse those ecosystems [@Haerter:2014ii; @Harcombe:2005fd; @Middelboe:2001fl; @Poisot:2011jc; @Thompson:2012ki; @Moebus:1981kp; @Flores:2013hc; @Poisot:2012fh; @Poisot071993; @Flores:2011bh; @Jover:2015ev].
-The conventional approach to understanding the human microbiome has been unable to capture how these communities interact and influence each other, but relationship-based network approaches would allow us to capture this information.
-We aimed to bridge this knowledge gap by characterizing human bacterial and phage communities by their relationships, leveraging machine learning and graph theory techniques.
+The most common approach of studying communities in isolation has been unable to capture how these communities interact and influence one another, but relationship-based network approaches allow us to capture this information.
+We aimed to bridge this knowledge gap by leveraging machine learning and graph theory techniques to characterize human bacterial and phage communities by their relationships.
 We focused on characterizing the network diversity of the human skin and gut microbiomes to provide a foundation for further studies of disease network dynamics and to gain broader insights into human microbiome network diversity across different body sites.
 
-We first investigated human microbiome phage and bacteria networks by utilizing three published microbiome datasets with paired virus and bacterial metagenomic sequence sets [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei].
-Our approach built off of previous large-scale phage-bacteria microbiome network analyses by inferring interactions using metagenomic datasets, instead of using culture-based techniques [@Moebus:1981kp; @Flores:2013hc].
-This metagenomic interaction inference model was powered beyond previous models by its inclusion of protein interaction data, negative as well as positive interactions, and the use of a more sophisticated machine learning algorithm instead of simple linear models [@Edwards:2015iz].
+First we investigated human microbiome phage and bacteria networks by utilizing three published microbiome datasets with paired virus and bacterial metagenomic sequence sets [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei].
+Our approach built off of previous large-scale phage-bacteria microbiome network analyses by inferring interactions using metagenomic datasets, instead of using culture-based techniques [@Flores:2013hc].
+This metagenomic interaction inference model was powered beyond previous models by its inclusion of protein interaction data as well as both negative and positive interactions [@Edwards:2015iz, @Roux:2016cc].
 
@@
 By building and utilizing a microbiome network, we found that different people, body sites, and anatomical locations not only support distinct microbiome membership and diversity [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei; @Grice:2009eea; @Findley:2013jf; @Costello:2009im, @Consortium:2012iz], but also support communities with distinct communication structures and propensities toward community instability.
-By understanding the healthy state of network structures across the human body, we will empower future studies to begin investigating how these community structures change in disease and their overall impact on human health.
+Through an improved understanding of the healthy state of network structures across the human body, we will empower future studies to begin investigating how these community structures change in disease and their overall impact on human health.
 
@@
 Our study contained three datasets, including a study of the impact of diet on the healthy human gut virome [@Minot:2011ez], the impact of anatomical location on the healthy human skin virome [@Hannigan:2015fz], and the viromes of monozygotic twins and their mothers [@Reyes:2010cwa; @Turnbaugh:2009ei].
-The viromes associated with these datasets were all subjected to virus-like particle (VLP) purification to eliminate other organismal DNA and allow for direct assessment of the actively replicating virome.
+The viromes associated with these datasets were all subjected to virus-like particle (VLP) purification to eliminate other organismal DNA and allow for direct assessment of both the extracellular and actively replicating intracellular virome.
 
-The publications we utilized were published over the span of five years using different methods and technologies, and therefore yielded different sequence abundances **(Supplemental Figure \ref{SequenceStats} A-B)**.
+We utilized data that was published over the span of five years using different methods and technologies, and therefore yielded different sequence abundances **(Supplemental Figure \ref{SequenceStats} A-B)** [@Minot:2011ez, @Hannigan:2015fz, @Reyes:2010cwa; @Turnbaugh:2009ei].
 The bacterial and viral sequences from those publications were quality filtered and assembled into contigs.
-Because contig assembly most commonly returns genome fragments, we clustered related bacteria and phage contigs by k-mer frequency and co-abundance using the CONCOCT algorithm.
+We generate clusters of related related bacterial and phage contigs by k-mer frequency and co-abundance using the CONCOCT algorithm.
 We interpreted these clusters as operational units of related bacteria and phage genomes that we defined as operational genomic units (OGUs).
@@
 We predicted which phage OGUs infected which bacteria using a random forest model trained on experimentally validated infectious relationships from six previous publications [@Jensen:1998vh;@Malki:2015tm;@Schwarzer:2012ez;@Kim:2012dh;@Matsuzaki:1992gw;@Edwards:2015iz].
-This training set contained diverse bacteria and phages, with both broad and specific infectious ranges **(Supplemental Figure \ref{ValidationOverview} A - B)**.
+This training set contained 43 diverse bacterial species and 30 diverse phage strains, with both broad and specific infectious ranges **(Supplemental Figure \ref{ValidationOverview} A - B)**.
 Phages with linear and circular genomes, as well as ssDNA and dsDNA genomes, were included in the analysis.
-Because this was a DNA sequencing study, RNA phages were not considered in the analysis **(Supplemental Figure \ref{ValidationOverview} C-D)**.
+Because this was a DNA sequencing study, RNA phages were not considered **(Supplemental Figure \ref{ValidationOverview} C-D)**.
 This training set included both positive relationships (phage infects a bacterium) and negative relationships (phage does not infect a bacterium).
-This built on previous work, which focused only on positive relationships, by allowing us to validate the false positive *and* false negative rates associated with our candidate models [@Edwards:2015iz].
+This allowed us to validate the false positive and false negative rates associated with our candidate models, thereby building upon previous work that only considered positive relationship [@Edwards:2015iz].
 
-Four phage and bacterial genomic markers were used to predict infectious relationships between bacteria and phages: 1) genome nucleotide similarities, 2) gene amino acid sequence similarities, 3) CRISPR targeting of phages by bacterial CRISPR spacer sequences, and 4) similarity of protein families known to be associated with known protein-protein interactions [@Orchard:2014hq].
-The resulting random forest model performed well, with an AUC of `r signif(genmodelper$ROC, digits = sigfig)`, a sensitivity of `r signif(genmodelper$Sens, digits = sigfig)`, and a specificity of `r signif(genmodelper$Spec, digits = sigfig)` **(Figure \ref{RocCurve} A)**.
+Four phage and bacterial genomic features were used to predict infectious relationships between bacteria and phages: 1) genome nucleotide similarities, 2) gene amino acid sequence similarities, 3) CRISPR targeting of phages by bacterial CRISPR spacer sequences, and 4) similarity of protein families associated with experimentally identified protein-protein interactions [@Orchard:2014hq].
+The resulting random forest model performed with an AUC of `r signif(genmodelper$ROC, digits = sigfig)`, a sensitivity of `r signif(genmodelper$Sens, digits = sigfig)`, and a specificity of `r signif(genmodelper$Spec, digits = sigfig)` **(Figure \ref{RocCurve} A)**.
 The most important predictor in the model was nucleotide similarity between genes, followed by nucleotide similarity of whole genomes **(Figure \ref{RocCurve} B)**.
-Protein family interactions were moderately important to the model, while CRISPRs were minimally important.
-Approximately one third of the relationships yielded no score and were therefore unknown, unable to be assigned an interaction **(Figure \ref{RocCurve} C)**.
+Protein family interactions were moderately important to the model, while CRISPRs were not informative important **(Figure \ref{RocCurve} B)**.
+Approximately one third of the relationships yielded no score and were therefore unknown, unable to be assigned an interaction prediction **(Figure \ref{RocCurve} C)**.
 
-Next we used our random forest model to classify the relationships between bacteria and phages in the experimental datasets.
-The relationships within the three studies were used to construct one master network, containing the three study sub-networks, which themselves each contained sub-networks for each sample **(Figure \ref{RocCurve} D)**.
-Metadata including study, sample ID, disease, and abundance within the community (based on sequence count) were also stored in the multi-study master network to allow for effective parsing and downstream analysis **(Supplemental Figure \ref{NetworkDiagram})**.
-The resulting master network was highly connected and contained `r prettyNum(genfigurestats[c(genfigurestats$class %in% "Total"),"Edges"], big.mark = ",")` infectious relationships among `r prettyNum(genfigurestats[c(genfigurestats$class %in% "Total"),"Vertices"], big.mark = ",")` nodes, `r prettyNum(nestats[c(nestats$cats %in% "PhageNodes"),"values"], big.mark = ",")` of which represented phages and `r prettyNum(nestats[c(nestats$cats %in% "BacteriaNodes"),"values"], big.mark = ",")` that represented bacteria.
+Next we used our random forest model to classify the relationships between bacteria and phage operational genomic units in the experimental datasets.
+The relationships predicted within the three published datasets were used to construct one master network, which was built to contain the three studies as sub-networks, which themselves each contained sub-networks for each sample **(Figure \ref{RocCurve} D)**.
+Metadata including study, sample ID, disease, and OGU abundance within the community were stored in the master network for downstream analysis **(Supplemental Figure \ref{NetworkDiagram})**.
+The master network was highly connected and contained `r prettyNum(genfigurestats[c(genfigurestats$class %in% "Total"),"Edges"], big.mark = ",")` infectious relationships among `r prettyNum(genfigurestats[c(genfigurestats$class %in% "Total"),"Vertices"], big.mark = ",")` nodes, `r prettyNum(nestats[c(nestats$cats %in% "PhageNodes"),"values"], big.mark = ",")` phages and `r prettyNum(nestats[c(nestats$cats %in% "BacteriaNodes"),"values"], big.mark = ",")` bacteria.
 Although the network was highly connected, not all relationships were present in all samples.
-Furthermore, relationships were weighted by relative abundance of their associated bacteria and phages, meaning lowly abundant relationships could be present but insignificant compared to those that were highly abundant.
+As relationships were weighted by the relative abundances of their associated bacteria and phages, lowly abundant relationships could be present but not highly abundant.
 Like the master network, the skin network exhibited a diameter of `r prettyNum(genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Diameter"], big.mark = ",")` (measure of graph size; the greatest number of traversed vertices required between two vertices) and included almost all of the master network nodes and edges **(Figure \ref{RocCurve} E - F)**.
-The gut diet and twin sample sets each contained fewer than 150 vertices, fewer than 20,000 relationships, and diameters of `r prettyNum(genfigurestats[c(genfigurestats$class %in% "DietStudy"),"Diameter"], big.mark = ",")`, which indicated more sparsely related phages and bacteria **(Figure \ref{RocCurve} E - F)**.
+The phages and bacteria in the gut diet and twin sample sets were more sparsely related, each contained fewer than 150 vertices, fewer than 20,000 relationships, and diameters of `r prettyNum(genfigurestats[c(genfigurestats$class %in% "DietStudy"),"Diameter"], big.mark = ",")`, which indicated more sparsely related phages and bacteria **(Figure \ref{RocCurve} E - F)**.
 
@@
 Skin and gut community membership and diversity are highly personal, with people remaining more similar to themselves over time compared to other people [@Grice:2009ee; @Hannigan:2015fz; @Minot:2013ih].
-We determined whether this personal conservation extended to network structure using our microbiome network model.
-To accomplish this, we calculated the degree of dissimilarity between each subject's network based on phage and bacteria abundance, as well as centrality.
-Weighted eigenvector centrality was calculated for all bacteria and phages within each sample graph.
+We therefore hypothesized that this personal conservation extended to microbiome network structure.
+To answer this hypothesis, we calculated the degree of dissimilarity between each subject's network based on phage and bacteria abundance, as well as centrality.
+We quantified phage and bacteria centrality within each sample graph using the weighted eigenvector centrality metric.
 This metric defines central phages as those that are highly abundant and infect many distinct bacteria which themselves are abundant and infected by many other phages.
-Similarly, bacterial centrality was defined as those bacteria that were both abundant and susceptible to numerous phages which infected many bacteria.
+Similarly, bacterial centrality was defined as those bacteria that were both abundant and connected to numerous phages that were themselves connected to many bacteria.
 We then calculated the similarity of community networks using the weighted eigenvector centrality of all nodes between all samples.
@@
 
-We used this network dissimilarity metric to test whether microbiome network structures were more similar within people over time compared to other people.
+We used this network dissimilarity metric to test whether microbiome network structures were more similar within people than between people over time.
 We found that gut microbiome network structure clustered by person (ANOSIM p-value = `r signif(interstats[c(interstats$site %in% "DietAnosim"), "prob"], digits = sigfig)`, R = `r signif(interstats[c(interstats$site %in% "AS"), "prob"], digits = sigfig)`, **Figure \ref{intradiv} A**).
-We found evidence that network dissimilarity within people over the 8-10 day sampling period was less than the average dissimilarity between that person and others, although this was not statistically significant (p-value = `r signif(interstats[c(interstats$site %in% "Diet"), "prob"], digits = sigfig)`, **Figure \ref{intradiv} B**).
-The lack of statistical confidence was likely due to the small sample size of this dataset.
+Network dissimilarity within people over the 8-10 day sampling period was less than the average dissimilarity between that person and others, although this was not statistically significant (p-value = `r signif(interstats[c(interstats$site %in% "Diet"), "prob"], digits = sigfig)`, **Figure \ref{intradiv} B**).
+The lack of statistical confidence was likely due to the small gut microbiome sample size of this dataset.
 Although there was evidence for gut network conservation among individuals, we found no evidence for conservation of gut network structures within families.
@@
 
-Skin microbiome network structure was strongly conserved within individuals, and more adequately confirmed by our larger skin dataset (p-value = `r signif(interstats[c(interstats$site %in% "Skin"), "prob"], digits = sigfig)`, **Figure \ref{intradiv} D**).
-This distribution was similar when separated by anatomical sites, with most sites also being significantly more conserved within individuals **(Supplemental Figure \ref{allskin})**.
+<!-- Maybe add twin to twin data here? -->
+
+Skin microbiome network structure was strongly conserved within individuals (p-value = `r signif(interstats[c(interstats$site %in% "Skin"), "prob"], digits = sigfig)`, **Figure \ref{intradiv} D**) and more adequately confirmed by our larger skin dataset.
+This distribution was similar when separated by anatomical sites, with most sites also being statistically significantly more conserved within individuals **(Supplemental Figure \ref{allskin})**.
 
 ## Role of Diet & Obesity in Gut Microbiome Conectivity
+
+<!-- Could maybe move this section up before the previous section -->
+
 Diet is a major environmental factor that influences resource availability and gut microbiome composition and diversity, including bacteria and phages [@Minot:2011ez; @Turnbaugh:2009hf; @David:2014cl].
-Previous work in isolated culture-based systems has suggested that changes in nutrient availability are associated with altered phage - bacteria network structures, although this has yet to be tested in humans [@Poisot:2011jc].
-We hypothesized that a change in diet would be associated with a change in network structure.
+Previous work in isolated culture-based systems has suggested that changes in nutrient availability are associated with altered phage - bacteria network structures [@Poisot:2011jc], although this has yet to be tested in humans.
+We hypothesized that a change in diet would be associated with a change in network structure in these human systems.
 
@@
 Closeness centrality is a metric of how close each phage or bacterium is to all of the other phages and bacteria in the network.
-A higher closeness centrality suggests that genetic information or the effects of altered abundance would be more impactful to all other microbes in the system.
-A network with higher average closeness centrality also indicates an overall greater degree of connections, which suggests a greater resilience to instability.
+A higher closeness centrality suggests that the effects of genetic information or altered abundance would be more impactful to all other microbes in the system.
+A network with higher average closeness centrality also indicates an overall greater degree of connections, which suggests a greater resilience against instability.
 Because these values are assigned to each phage and bacterium within each network, we calculated the average connectedness and corrected for the maximum potential degree of connectedness to obtain a single value for the connectedness of each graph.
 
-We found that gut microbiome network structures associated with high-fat diets were less connected than those of low-fat diets **(Figure \ref{dietnetworks} A-B)**.
+We found that the gut microbiome network structures associated with high-fat diets were less connected than those of low-fat diets **(Figure \ref{dietnetworks} A-B)**.
 Tests for statistical differences were not performed due to the small sample size.
 High-fat diets exhibited less degree centrality **(Figure \ref{dietnetworks} A)**.
-This finding suggests that bacteria were targeted by less phages and that phage tropism was more specific.
-High-fat diets also exhibited decreased closeness centrality **(Figure \ref{dietnetworks} B)**, indicating that microbes were more distant from other microbes in the community.
-This would make genetic transfer and altered abundance less capable of impacting other bacteria and phages within the network.
+This finding suggests that bacteria in high-fat environments were targeted by fewer phages and that phage tropism was more specific.
+High-fat diets also exhibited decreased closeness centrality **(Figure \ref{dietnetworks} B)**, indicating that bacteria and phages were more distant from other bacteria and phages in the community.
+This would make genetic transfer and altered abundance of a given phage or bacterium less capable of impacting other bacteria and phages within the network.
 
@@
 
-## Variation of Network Structure Across the Human Skin Landscape
+## Association Between Environmental Stability and Network Structure Across the Human Skin Landscape
 Extensive work has illustrated differences in healthy human skin microbiome between anatomical sites, including bacteria, viruses, and fungi [@Grice:2009ee; @Findley:2013jf; @Hannigan:2015fz].
 These communities vary by degree of skin moisture, oil, and environmental exposure.
-We hypothesized that like microbial composition and diversity, microbial network structure differed between anatomical sites.
+We hypothesized that like microbial composition and diversity, microbial network structure would be specific to anatomical sites.
 We tested this hypothesis by evaluating the changes in network structure between anatomical sites within our skin dataset.
@@
 # Discussion
-We developed and implemented a network-based analytical approach which we used to begin evaluating the basic properties of the human microbiome through bacteria and phage *relationships*, instead of membership or diversity.
-The goal of this study was to provide an initial understanding of how phage-bacteria networks differ throughout the human body, so as to provide a baseline for future studies of how microbiome networks differ in disease states.
-This goal of providing a baseline understanding of the human microbiome was similar to the goals of initial studies of bacteria and viral diversity across the human body, as well as other environments [@Grice:2009eea; @Findley:2013jf; @Hannigan:2015fz; @Costello:2009im, @Consortium:2012iz; @Schloss:2005hz; @Minot:2011ez].
-Through the use of network theory, we leveraged extensive analytical opportunities which could be applied to understand complex ecological communities.
-We focused on utilizing metrics of connectivity to understand the extent to which communities of bacteria and phages interact (e.g.
+Early, foundational microbiome work has provided a baseline understanding of the human microbiome by characterizing bacteria and viral diversity across the human body, as well as other environments [@Grice:2009eea; @Findley:2013jf; @Hannigan:2015fz; @Costello:2009im, @Consortium:2012iz; @Schloss:2005hz; @Minot:2011ez].
+We similarily aimed to provide an initial understanding of how phage-bacteria networks differ throughout the human body, so as to provide a baseline for future studies of how microbiome networks differ in disease states.
+We developed and implemented a network-based analytical approach that we used to evaluate the basic properties of the human microbiome through bacteria and phage *relationships*, instead of membership or diversity.
+Through the use of network theory, we leveraged extensive analytical opportunities that could be applied to understand complex ecological communities.
+We utilized metrics of connectivity to model the extent to which communities of bacteria and phages interact (e.g.
 horizontal gene transfer, modulated bacterial gene expression, alterations in abundance).
-By pursuing this goal, we aimed to provide an initial understanding of human microbiome networks that will power future studies to use similar approaches in other setting such as disease states.
+By pursuing this goal, we aimed to provide an initial understanding of human microbiome networks that will power future studies to use similar approaches in other settings, such as disease states.
 
-We reported that, just as gut microbiome and virome composition and diversity are conserved in individuals [@Hannigan:2015fz; @Grice:2009eea; @Findley:2013jf; @Minot:2013ih], gut and skin microbiome network structures were also conserved within individuals over time.
+We found that, just as gut microbiome and virome composition and diversity are conserved in individuals [@Hannigan:2015fz; @Grice:2009eea; @Findley:2013jf; @Minot:2013ih], gut and skin microbiome network structures were also conserved within individuals over time.
 Gut network structure was not conserved among family members.
-These findings suggest that microbiome properties such as stability and the potential for horizontal gene transfer are personal, and potentially impacted by personal factors ranging from the body's immune system to environmental conditions such as climate and diet.
+These findings suggest that microbiome network properties, such as stability and the potential for horizontal gene transfer, are personal and may be impacted by personal factors ranging from the body's immune system to environmental conditions such as climate and diet.
 The ability of environmental conditions to shape gut and skin microbiome network structure was supported by our finding that diet and skin location were associated with altered network structures.
@@
 Although our sample size was small, we found evidence that high-fat diets were less connected than low-fat diets.
-This suggested that high-fat diets may lead to less stable communities with a decreased ability to influence each other.
+This suggested that high-fat diets may lead to less stable communities with a decreased ability for individuals to influence one other.
 We supported this finding with the observation that obesity was associated with decreased network connectivity.
-Our results allow us to speculate that the food we eat may not only impact what microbes colonize our guts, but may also impact their interactions.
+Together these findings suggest the food we eat may not only impact which microbes colonize our guts, but may also impact their interactions.
 Further work will be required to characterize these relationships with a larger cohort.
 
-In addition to diet, we found evidence that skin microbiome network structure varied by skin environment.
-We observed that network structure differed between environmentally exposed and occluded skin sites.
-The sites under greater environmental fluctuation and exposure, the exposed and intermittently exposed sites, were less connected and therefore were expected to have a higher propensity for instability.
+In addition to diet, we found that skin microbiome network structure varied by skin environment.
+Network structure differed between environmentally exposed and occluded skin sites.
+The sites under greater environmental fluctuation and exposure, the exposed and intermittently exposed sites, were less connected and therefore were predicted to have a higher propensity for instability.
 Likewise, intermittently moist sites demonstrated less connectedness than the more stable moist and sebaceous sites.
@@
 *First*, while our infection classification model is advantageous over existing models, we recognize that, like most classification models, there remains opportunity for improvement.
-For example, such a model is only as good as its training set, and future large-scale endeavors into infectious relationships (and the associated genomes) will provide more robust training and higher model accuracy.
-Just as we have improved on previous modeling efforts, we expect that new and creative scoring metrics will likely be integrated into this model to further improve model performance.
-*Second*, while informative, this work was done retrospectively and relied on published research from as many as seven years ago.
-These archived datasets were limited by the technology and costs of the time, meaning the datasets are poorly powered for the statistical analysis we strive for today.
+For example, such a model is only as good as its training set.
+Large scale experimental screens for phage and bacteria infectious interactions will provide more robust model training, and therefore improved model performance.
+Just as we have improved on previous modeling efforts, we expect that new and creative scoring metrics will be integrated into this model to further improve model performance.
+
+*Second*, while a successful and informative proof of concept, this work was done retrospectively and relied on published research from as many as seven years ago.
+These archived datasets were limited by the technology and costs of the time, meaning the datasets were poorly powered for the statistical analysis we strive for today.
 While we were able to present initial conclusions, follow-up studies will be required to validate our findings.
-Despite their limitations, these methods and results will be important for informing design and interpretation of future studies of the human microbiome.
+Despite their limitations, these methods and results will be important for informing design and interpretation of future studies and for highlighting the importance of bacteria-virus interaction profiles as a characteristic of human-associated microbial communities.
 
-We demonstrated the diversity of relationships across the human body by showing that microbiome relationship structure, and therefore the potential for genetic transfer and stability maintenance, differs significantly between body sites and environmental conditions.
-We found that other environmental factors, such as diet, also influence relationship structures.
+We demonstrated that microbiome relationship structure, and therefore the potential for microbiome genetic transfer and stability maintenance, differed significantly between body sites.
+Other environmental factors, such as diet and individuality, also influenced relationship structures of the microbiome.
 This information builds on previous work by suggesting that bacteria and phages not only preferentially colonize different body sites, but also interact differently, allowing for different communication structures and capacities for maintaining stability.
-Our results will also provide a foundation for future studies to begin evaluating how microbiome network dynamics change in disease states, and how the information can be applied therapeutically.
+Our results provide a foundation for future studies to begin evaluating how microbiome network dynamics change in disease states, and how the information can be applied therapeutically.
 
@@
 Contigs were concatenated into two master files prior to alignment, one for bacterial contigs and one for phage contigs.
-Sample sequences were aligned to phage or bacteria contigs using the Bowtie2 global aligner (v2.2.1) [@Langmead:2012jh].
+Sample sequences were aligned to phage or bacterial contigs using the Bowtie2 global aligner (v2.2.1) [@Langmead:2012jh].
 We defined a mismatch threshold of 1 bp and seed length of 25 bp.

### 2017-04-18 doc/manuscript.Rmd
Most edits done. Only a couple of small things.

@@
 
-**Word Count**: 238 / 250
+**Word Count**: 237 / 250
 
@@
 
-Most previous work has relied on studying bacterial and phage communities in isolation, reducing them to two separate communities.
+Most previous work has relied on studying bacterial and phage communities in isolation, reducing them to two separate communities [@Norman:2015kb; @Manrique:2016dx; @Ly:2014ew; @Monaco:2016ita; @Hannigan:2015fz; @Minot:2011ez; @SantiagoRodriguez:2015gd; @Abeles:2015dy; @Abeles:2014kj].
 In reality, bacteria and phage communities are dynamic and complex.
@@
 Removal of members can disrupt or even collapse those ecosystems [@Haerter:2014ii; @Harcombe:2005fd; @Middelboe:2001fl; @Poisot:2011jc; @Thompson:2012ki; @Moebus:1981kp; @Flores:2013hc; @Poisot:2012fh; @Poisot071993; @Flores:2011bh; @Jover:2015ev].
-The most common approach of studying communities in isolation has been unable to capture how these communities interact and influence one another, but relationship-based network approaches allow us to capture this information.
+The most common approach to studying communities in isolation has been unable to capture how these communities interact and influence one another, but relationship-based network approaches allow us to capture this information.
 We aimed to bridge this knowledge gap by leveraging machine learning and graph theory techniques to characterize human bacterial and phage communities by their relationships.
-We focused on characterizing the network diversity of the human skin and gut microbiomes to provide a foundation for further studies of disease network dynamics and to gain broader insights into human microbiome network diversity across different body sites.
+We focused on characterizing the human skin and gut virome-microbiome network diversity to provide a foundation for further studies of disease network dynamics and to gain broader insights into human microbiome network diversity across different body sites.
 
@@
 ## Cohort Curation and Sample Processing
-We studied the differences in microbiome phage-bacteria interaction networks across healthy human bodies by leveraging previously published sequence sets containing purified virome samples paired with bacterial metagenomes from whole metagenomic shotgun sequences.
+We studied the differences in virome-microbiome interaction networks across healthy human bodies by leveraging previously published sequence sets containing purified virome samples paired with bacterial metagenomes from whole metagenomic shotgun sequences.
 Our study contained three datasets, including a study of the impact of diet on the healthy human gut virome [@Minot:2011ez], the impact of anatomical location on the healthy human skin virome [@Hannigan:2015fz], and the viromes of monozygotic twins and their mothers [@Reyes:2010cwa; @Turnbaugh:2009ei].
-The viromes associated with these datasets were all subjected to virus-like particle (VLP) purification to eliminate other organismal DNA and allow for direct assessment of both the extracellular and actively replicating intracellular virome.
+The viromes associated with these datasets were selected because they were subjected to virus-like particle (VLP) purification to eliminate other organismal DNA and allow for direct assessment of both the extracellular and actively replicating intracellular virome.
 
@@
 The bacterial and viral sequences from those publications were quality filtered and assembled into contigs.
-We generate clusters of related related bacterial and phage contigs by k-mer frequency and co-abundance using the CONCOCT algorithm.
+We generated clusters of related related bacterial and phage contigs by k-mer frequency and co-abundance using the CONCOCT algorithm **(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})**.
 We interpreted these clusters as operational units of related bacteria and phage genomes that we defined as operational genomic units (OGUs).
 These OGUs are conceptually similar to the operational taxonomic unit (OTU) and operational protein family (OPF) definitions used for grouping highly similar 16S rRNA gene and open reading frame sequences, respectively [@Schloss:2008ht].
-The resulting contigs and OGUs demonstrated high sequence coverage and length **(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})**.
 
 ## Establishing a Model of Phage-Bacteria Interactions
-We predicted which phage OGUs infected which bacteria using a random forest model trained on experimentally validated infectious relationships from six previous publications [@Jensen:1998vh;@Malki:2015tm;@Schwarzer:2012ez;@Kim:2012dh;@Matsuzaki:1992gw;@Edwards:2015iz].
+We predicted which phage OGUs infected which bacterial OGUs using a random forest model trained on experimentally validated infectious relationships from six previous publications [@Jensen:1998vh;@Malki:2015tm;@Schwarzer:2012ez;@Kim:2012dh;@Matsuzaki:1992gw;@Edwards:2015iz].
 This training set contained 43 diverse bacterial species and 30 diverse phage strains, with both broad and specific infectious ranges **(Supplemental Figure \ref{ValidationOverview} A - B)**.
 Phages with linear and circular genomes, as well as ssDNA and dsDNA genomes, were included in the analysis.
-Because this was a DNA sequencing study, RNA phages were not considered **(Supplemental Figure \ref{ValidationOverview} C-D)**.
+Because we used DNA sequencing studies, RNA phages were not considered **(Supplemental Figure \ref{ValidationOverview} C-D)**.
 This training set included both positive relationships (phage infects a bacterium) and negative relationships (phage does not infect a bacterium).
-This allowed us to validate the false positive and false negative rates associated with our candidate models, thereby building upon previous work that only considered positive relationship [@Edwards:2015iz].
+This allowed us to validate the false positive and false negative rates associated with our candidate models, thereby building upon previous work that only considered positive relationships [@Edwards:2015iz].
 
@@
 The most important predictor in the model was nucleotide similarity between genes, followed by nucleotide similarity of whole genomes **(Figure \ref{RocCurve} B)**.
-Protein family interactions were moderately important to the model, while CRISPRs were not informative important **(Figure \ref{RocCurve} B)**.
-Approximately one third of the relationships yielded no score and were therefore unknown, unable to be assigned an interaction prediction **(Figure \ref{RocCurve} C)**.
+Protein family interactions were moderately important to the model, while CRISPRs were not informative **(Figure \ref{RocCurve} B)**.
+Approximately one third of the relationships yielded no score and were therefore unable to be assigned an interaction prediction **(Figure \ref{RocCurve} C)**.
 
-Next we used our random forest model to classify the relationships between bacteria and phage operational genomic units in the experimental datasets.
+Once our random forest model was built, we used it to classify the relationships between bacteria and phage operational genomic units in the experimental datasets.
 The relationships predicted within the three published datasets were used to construct one master network, which was built to contain the three studies as sub-networks, which themselves each contained sub-networks for each sample **(Figure \ref{RocCurve} D)**.
@@
 As relationships were weighted by the relative abundances of their associated bacteria and phages, lowly abundant relationships could be present but not highly abundant.
-Like the master network, the skin network exhibited a diameter of `r prettyNum(genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Diameter"], big.mark = ",")` (measure of graph size; the greatest number of traversed vertices required between two vertices) and included almost all of the master network nodes and edges **(Figure \ref{RocCurve} E - F)**.
+Like the master network, the skin network exhibited a diameter of `r prettyNum(genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Diameter"], big.mark = ",")` (measure of graph size; the greatest number of traversed vertices required between two vertices) and included `r round(100 * genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Vertices"] / genfigurestats[c(genfigurestats$class %in% "Total"),"Vertices"], digits = 1)`% and `r round(100 * genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Edges"] / genfigurestats[c(genfigurestats$class %in% "Total"),"Edges"], digits = 1)`% of the master network nodes and edges, respectively **(Figure \ref{RocCurve} E - F)**.
 The phages and bacteria in the gut diet and twin sample sets were more sparsely related, each contained fewer than 150 vertices, fewer than 20,000 relationships, and diameters of `r prettyNum(genfigurestats[c(genfigurestats$class %in% "DietStudy"),"Diameter"], big.mark = ",")`, which indicated more sparsely related phages and bacteria **(Figure \ref{RocCurve} E - F)**.
+
+## Role of Diet & Obesity in Gut Microbiome Connectivity
+
+Diet is a major environmental factor that influences resource availability and gut microbiome composition and diversity, including bacteria and phages [@Minot:2011ez; @Turnbaugh:2009hf; @David:2014cl].
+Previous work in isolated culture-based systems has suggested that changes in nutrient availability are associated with altered phage - bacteria network structures [@Poisot:2011jc], although this has yet to be tested in humans.
+We hypothesized that a change in diet would be associated with a change in virome-microbiome network structure in these human systems.
+
+We evaluated the diet-associated differences in gut virome-microbiome network structure by quantifying how central each sample's network was on average.
+We accomplished this by utilizing two common centrality metrics: degree centrality and closeness centrality.
+Degree centrality, the simplest centrality metric, was defined as the number of connections each phage made to bacteria, or each bacterium made to phages.
+We also quantified the network properties by supplementing degree centrality with measurements of closeness centrality.
+Closeness centrality is a metric of how close each phage or bacterium is to all of the other phages and bacteria in the network.
+A higher closeness centrality suggests that the effects of genetic information or altered abundance would be more impactful to all other microbes in the system.
+A network with higher average closeness centrality also indicates an overall greater degree of connections, which suggests a greater resilience against instability.
+Because these values are assigned to each phage and bacterium within each network, we calculated the average connectedness and corrected for the maximum potential degree of connectedness to obtain a single value for the connectedness of each graph.
+
+We found that the gut microbiome network structures associated with high-fat diets were less connected than those of low-fat diets **(Figure \ref{dietnetworks} A-B)**.
+Tests for statistical differences were not performed due to the small sample size.
+High-fat diets exhibited less degree centrality **(Figure \ref{dietnetworks} A)**.
+This finding suggested that bacteria in high-fat environments were targeted by fewer phages and that phage tropism was more specific.
+High-fat diets also exhibited decreased closeness centrality **(Figure \ref{dietnetworks} B)**, indicating that bacteria and phages were more distant from other bacteria and phages in the community.
+This would make genetic transfer and altered abundance of a given phage or bacterium less capable of impacting other bacteria and phages within the network.
+
+In addition to diet, we also observed an association between obesity and network structure **(Figure \ref{dietnetworks} C-D)**.
+The obesity-associated network demonstrated a higher degree centrality **(Figure \ref{dietnetworks} C)** but less closeness centrality compared to the healthy controls **(Figure \ref{dietnetworks} D)**.
+These results suggest that the obesity network was overall less connected, having microbes further from all other microbes within the community.
 
@@
 We found that gut microbiome network structure clustered by person (ANOSIM p-value = `r signif(interstats[c(interstats$site %in% "DietAnosim"), "prob"], digits = sigfig)`, R = `r signif(interstats[c(interstats$site %in% "AS"), "prob"], digits = sigfig)`, **Figure \ref{intradiv} A**).
-Network dissimilarity within people over the 8-10 day sampling period was less than the average dissimilarity between that person and others, although this was not statistically significant (p-value = `r signif(interstats[c(interstats$site %in% "Diet"), "prob"], digits = sigfig)`, **Figure \ref{intradiv} B**).
-The lack of statistical confidence was likely due to the small gut microbiome sample size of this dataset.
+Network dissimilarity within each person over the 8-10 day sampling period was less than the average dissimilarity between that person and others, although this difference was not statistically significant (p-value = `r signif(interstats[c(interstats$site %in% "Diet"), "prob"], digits = sigfig)`, **Figure \ref{intradiv} B**).
+The lack of statistical confidence was likely due to the small sample size of this dataset.
 Although there was evidence for gut network conservation among individuals, we found no evidence for conservation of gut network structures within families.
-The gut network structures were not more similar between twins and their mothers (intrafamily) compared to other families of twins and mothers (inter-family) (p-value = `r signif(interstats[c(interstats$site %in% "Twins"), "prob"], digits = sigfig)`, **Figure \ref{intradiv} C**).
-
-<!-- Maybe add twin to twin data here? -->
+The gut network structures were not more similar within families (twins and their mothers; intrafamily) compared to other families (inter-family) (p-value = `r signif(interstats[c(interstats$site %in% "Twins"), "prob"], digits = sigfig)`, **Figure \ref{intradiv} C**).
 
@@
 This distribution was similar when separated by anatomical sites, with most sites also being statistically significantly more conserved within individuals **(Supplemental Figure \ref{allskin})**.
-
-## Role of Diet & Obesity in Gut Microbiome Conectivity
-
-<!-- Could maybe move this section up before the previous section -->
-
-Diet is a major environmental factor that influences resource availability and gut microbiome composition and diversity, including bacteria and phages [@Minot:2011ez; @Turnbaugh:2009hf; @David:2014cl].
-Previous work in isolated culture-based systems has suggested that changes in nutrient availability are associated with altered phage - bacteria network structures [@Poisot:2011jc], although this has yet to be tested in humans.
-We hypothesized that a change in diet would be associated with a change in network structure in these human systems.
-
-We evaluated the diet-associated differences in gut network structure by quantifying how central each sample's network was on average.
-We accomplished this by utilizing two common centrality metrics: degree centrality and closeness centrality.
-Degree centrality, the simplest centrality metric, was defined as the number of connections each phage made to bacteria, or each bacterium made to phages.
-We more completely quantified the network properties by supplementing degree centrality with measurements of closeness centrality.
-Closeness centrality is a metric of how close each phage or bacterium is to all of the other phages and bacteria in the network.
-A higher closeness centrality suggests that the effects of genetic information or altered abundance would be more impactful to all other microbes in the system.
-A network with higher average closeness centrality also indicates an overall greater degree of connections, which suggests a greater resilience against instability.
-Because these values are assigned to each phage and bacterium within each network, we calculated the average connectedness and corrected for the maximum potential degree of connectedness to obtain a single value for the connectedness of each graph.
-
-We found that the gut microbiome network structures associated with high-fat diets were less connected than those of low-fat diets **(Figure \ref{dietnetworks} A-B)**.
-Tests for statistical differences were not performed due to the small sample size.
-High-fat diets exhibited less degree centrality **(Figure \ref{dietnetworks} A)**.
-This finding suggests that bacteria in high-fat environments were targeted by fewer phages and that phage tropism was more specific.
-High-fat diets also exhibited decreased closeness centrality **(Figure \ref{dietnetworks} B)**, indicating that bacteria and phages were more distant from other bacteria and phages in the community.
-This would make genetic transfer and altered abundance of a given phage or bacterium less capable of impacting other bacteria and phages within the network.
-
-In addition to diet, we also observed an association between obesity and network structure **(Figure \ref{dietnetworks} C-D)**.
-The obesity-associated network demonstrated a higher degree centrality **(Figure \ref{dietnetworks} C)** but less closeness centrality compared to the healthy controls **(Figure \ref{dietnetworks} D)**.
-These results suggest that the obesity network was overall less connected, having microbes further from all other microbes within the community.
 
@@
 Gut network structure was not conserved among family members.
-These findings suggest that microbiome network properties, such as stability and the potential for horizontal gene transfer, are personal and may be impacted by personal factors ranging from the body's immune system to environmental conditions such as climate and diet.
+These findings suggested that microbiome network properties, such as stability and the potential for horizontal gene transfer, were personal and may be impacted by personal factors ranging from the body's immune system to environmental conditions such as climate and diet.
 The ability of environmental conditions to shape gut and skin microbiome network structure was supported by our finding that diet and skin location were associated with altered network structures.
@@
 This suggested that high-fat diets may lead to less stable communities with a decreased ability for individuals to influence one other.
-We supported this finding with the observation that obesity was associated with decreased network connectivity.
+We supported this finding with the observation that obesity was associated with decreased network connectivity (closeness of bacterial and phage vertices).
 Together these findings suggest the food we eat may not only impact which microbes colonize our guts, but may also impact their interactions.
@@
 
+![**Impact of Diet and Obesity on Gut Network Structure.** *(A) Quantification of average degree centrality (number of edges per node) and (B) closeness centrality (average distance from each node to every other node) of gut microbiome networks of subjects limited to exclusively high-fat or low-fat diets.
+Lines represent the mean degree of centrality for each diet.
+(C) Quantification of average degree centrality and (D) closeness centrality between obese and healthy adult women.*\label{dietnetworks}](../figures/dietnetworks.pdf){ width=90% }
+
 ![**Intrapersonal vs Interpersonal Network Dissimilarity Across Different Human Systems.** *(A) NMDS ordination illustrating network dissimilarity between subjects over time.
@@
 *\label{intradiv}](../figures/intrapersonal_diversity.pdf){ width=90% }
-
-![**Impact of Diet and Obesity on Gut Network Structure.** *(A) Quantification of average degree centrality (number of edges per node) and (B) closeness centrality (average distance from each node to every other node) of gut microbiome networks of subjects limited to exclusively high-fat or low-fat diets.
-Lines represent the mean degree of centrality for each diet.
-(C) Quantification of average degree centrality and (D) closeness centrality between obese and healthy adult women.*\label{dietnetworks}](../figures/dietnetworks.pdf){ width=90% }
 
@@
 Phages are labeled on the x-axis and bacteria are labeled on the y-axis.
-B) World map illustrating the sampling locations used in the study (red dots).
-C) Quantification of bacterial host strains known to exist for each phage.
-D) Genome strandedness and E) linearity of the phage reference genomes used for the dataset.*\label{ValidationOverview}](../figures/BenchmarkDataset.pdf){ width=100% }
+B) Quantification of bacterial host strains known to exist for each phage.
+C) Genome strandedness and E) linearity of the phage reference genomes used for the dataset.*\label{ValidationOverview}](../figures/BenchmarkDataset.pdf){ width=100% }
 

### 2017-04-18 doc/manuscript.Rmd
A couple more edits.

@@
 
+## Statistics and Comparisons
+Differences in intrapersonal and interpersonal network structure diversity, based on multivariate data, were calculated using an analysis of similarity (ANOSIM).
+Statistical significance of univariate Eigenvector centrality differences were calculated using a paired Wilcoxon test.
+
+Statistical significance of differences in univariate eigenvector centrality measurements of skin virome-microbiome networks were calculated using a pairwise Wilcoxon test, corrected for multiple hypothesis tests using the Holm correction method.
+Multivariate eigenvector centrality was measured as the mean differences between cluster centroids, with statistical significance measured using an ANOVA and post hoc Tukey test.
+
 # Acknowledgments
@@
 ![**Impact of Skin Micro-Environment on Microbiome Network Structure.**  *(A) Notched box-plot depicting differences in average Eigenvector centrality between moist, intermittently moist, and sebaceous skin sites and (B) occluded, intermittently occluded, and exposed sites.
-Notched box-plots were created using ggplot2 and show the median (center line), the inter-quartile range (IQR; upper and lower boxes), the highest and lowest value within 1.5 * IQR (whiskers), outliers (dots), and the notch which provides an approximate 95% confidence interval as defined by 1.58 * IQR / sqrt(n).
+Notched box-plots were created using ggplot2 and show the median (center line), the inter-quartile range (IQR; upper and lower boxes), the highest and lowest value within 1.5 \* IQR (whiskers), outliers (dots), and the notch which provides an approximate 95% confidence interval as defined by 1.58 \* IQR / sqrt(n).
 (C) NMDS ordination depicting the differences in skin microbiome network structure between skin moisture levels and (D) occlusion.

### 2017-04-19 doc/manuscript.Rmd
Alright, updates updates updates.

@@
 Together this study characterized the diversity of virome-microbiome networks across the human body and illustrated how environmental factors may influence phage-bacteria interactive dynamics.
-This work provides an baseline for future studies to investigate the role of ecological networks in disease states.
+This work provides a baseline for future studies to investigate the role of ecological networks in conditions such as disease states.
 
@@
 We aimed to bridge this knowledge gap by leveraging machine learning and graph theory techniques to characterize human bacterial and phage communities by their relationships.
-We focused on characterizing the human skin and gut virome-microbiome network diversity to provide a foundation for further studies of disease network dynamics and to gain broader insights into human microbiome network diversity across different body sites.
+We focused on characterizing human skin and gut virome-microbiome network diversity to provide a foundation for further studies of disease network dynamics and to gain broader insights into human microbiome network diversity across different body sites.
 
-First we investigated human microbiome phage and bacteria networks by utilizing three published microbiome datasets with paired virus and bacterial metagenomic sequence sets [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei].
+First we investigated human microbiome-virome networks by utilizing three published microbiome datasets with paired virus and bacterial metagenomic sequence sets [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei].
 Our approach built off of previous large-scale phage-bacteria microbiome network analyses by inferring interactions using metagenomic datasets, instead of using culture-based techniques [@Flores:2013hc].
@@
 Through this approach, we were able to go beyond contemporary isolated methods to provide a basic understanding of the community network dynamics associated with healthy human phages and bacteria.
-By building and utilizing a microbiome network, we found that different people, body sites, and anatomical locations not only support distinct microbiome membership and diversity [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei; @Grice:2009eea; @Findley:2013jf; @Costello:2009im, @Consortium:2012iz], but also support communities with distinct communication structures and propensities toward community instability.
+By building and utilizing a microbiome-virome network, we found that different people, body sites, and anatomical locations not only support distinct microbiome membership and diversity [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei; @Grice:2009eea; @Findley:2013jf; @Costello:2009im, @Consortium:2012iz], but also support communities with distinct communication structures and propensities toward community instability.
 Through an improved understanding of the healthy state of network structures across the human body, we will empower future studies to begin investigating how these community structures change in disease and their overall impact on human health.
@@
 
-We utilized data that was published over the span of five years using different methods and technologies, and therefore yielded different sequence abundances **(Supplemental Figure \ref{SequenceStats} A-B)** [@Minot:2011ez, @Hannigan:2015fz, @Reyes:2010cwa; @Turnbaugh:2009ei].
+We utilized data that was published over the span of five years, which used different methods and technologies, and therefore yielded different sequence abundances **(Supplemental Figure \ref{SequenceStats} A-B)** [@Minot:2011ez, @Hannigan:2015fz, @Reyes:2010cwa; @Turnbaugh:2009ei].
 The bacterial and viral sequences from those publications were quality filtered and assembled into contigs.
-We generated clusters of related related bacterial and phage contigs by k-mer frequency and co-abundance using the CONCOCT algorithm **(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})**.
-We interpreted these clusters as operational units of related bacteria and phage genomes that we defined as operational genomic units (OGUs).
+We generated clusters of related bacterial and phage contigs by k-mer frequency and co-abundance using the CONCOCT algorithm **(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})**.
+We interpreted these clusters as operational units of related bacteria and phage genomes, which we defined as operational genomic units (OGUs).
 These OGUs are conceptually similar to the operational taxonomic unit (OTU) and operational protein family (OPF) definitions used for grouping highly similar 16S rRNA gene and open reading frame sequences, respectively [@Schloss:2008ht].
@@
 Because we used DNA sequencing studies, RNA phages were not considered **(Supplemental Figure \ref{ValidationOverview} C-D)**.
-This training set included both positive relationships (phage infects a bacterium) and negative relationships (phage does not infect a bacterium).
+This training set included both positive relationships (a phage infects a bacterium) and negative relationships (a phage does not infect a bacterium).
 This allowed us to validate the false positive and false negative rates associated with our candidate models, thereby building upon previous work that only considered positive relationships [@Edwards:2015iz].
 
-Four phage and bacterial genomic features were used to predict infectious relationships between bacteria and phages: 1) genome nucleotide similarities, 2) gene amino acid sequence similarities, 3) CRISPR targeting of phages by bacterial CRISPR spacer sequences, and 4) similarity of protein families associated with experimentally identified protein-protein interactions [@Orchard:2014hq].
+Four phage and bacterial genomic features were used to predict infectious relationships between bacteria and phages: 1) genome nucleotide similarities, 2) gene amino acid sequence similarities, 3) Clustered Regularly Interspaced Short Palindromic Repeat (CRISPR) targeting of phages by bacterial CRISPR spacer sequences, and 4) similarity of protein families associated with experimentally identified protein-protein interactions [@Orchard:2014hq].
 The resulting random forest model performed with an AUC of `r signif(genmodelper$ROC, digits = sigfig)`, a sensitivity of `r signif(genmodelper$Sens, digits = sigfig)`, and a specificity of `r signif(genmodelper$Spec, digits = sigfig)` **(Figure \ref{RocCurve} A)**.
@@
 Protein family interactions were moderately important to the model, while CRISPRs were not informative **(Figure \ref{RocCurve} B)**.
-Approximately one third of the relationships yielded no score and were therefore unable to be assigned an interaction prediction **(Figure \ref{RocCurve} C)**.
+Approximately one third of the training set relationships yielded no score and were therefore unable to be assigned an interaction prediction **(Figure \ref{RocCurve} C)**.
 
@@
 Diet is a major environmental factor that influences resource availability and gut microbiome composition and diversity, including bacteria and phages [@Minot:2011ez; @Turnbaugh:2009hf; @David:2014cl].
-Previous work in isolated culture-based systems has suggested that changes in nutrient availability are associated with altered phage - bacteria network structures [@Poisot:2011jc], although this has yet to be tested in humans.
-We hypothesized that a change in diet would be associated with a change in virome-microbiome network structure in these human systems.
+Previous work in isolated culture-based systems has suggested that changes in nutrient availability are associated with altered phage-bacteria network structures [@Poisot:2011jc], although this has yet to be tested in humans.
+We hypothesized that a change in diet would also be associated with a change in virome-microbiome network structure in the human gut.
 
@@
 Degree centrality, the simplest centrality metric, was defined as the number of connections each phage made to bacteria, or each bacterium made to phages.
-We also quantified the network properties by supplementing degree centrality with measurements of closeness centrality.
+We supplemented measurements of degree centrality with measurements of closeness centrality.
 Closeness centrality is a metric of how close each phage or bacterium is to all of the other phages and bacteria in the network.
@@
 Tests for statistical differences were not performed due to the small sample size.
-High-fat diets exhibited less degree centrality **(Figure \ref{dietnetworks} A)**.
-This finding suggested that bacteria in high-fat environments were targeted by fewer phages and that phage tropism was more specific.
+High-fat diets exhibited reduced degree centrality **(Figure \ref{dietnetworks} A)**, suggesting bacteria in high-fat environments were targeted by fewer phages and that phage tropism was more restricted.
 High-fat diets also exhibited decreased closeness centrality **(Figure \ref{dietnetworks} B)**, indicating that bacteria and phages were more distant from other bacteria and phages in the community.
@@
 We therefore hypothesized that this personal conservation extended to microbiome network structure.
-To answer this hypothesis, we calculated the degree of dissimilarity between each subject's network based on phage and bacteria abundance, as well as centrality.
+To answer this hypothesis, we calculated the degree of dissimilarity between each subject's network, based on phage and bacteria abundance as well as centrality.
 We quantified phage and bacteria centrality within each sample graph using the weighted eigenvector centrality metric.
@@
 
-Skin microbiome network structure was strongly conserved within individuals (p-value = `r signif(interstats[c(interstats$site %in% "Skin"), "prob"], digits = sigfig)`, **Figure \ref{intradiv} D**) and more adequately confirmed by our larger skin dataset.
+Skin microbiome network structure was strongly conserved within individuals (p-value < 0.001, **Figure \ref{intradiv} D**) and more adequately confirmed by our larger skin dataset.
 This distribution was similar when separated by anatomical sites, with most sites also being statistically significantly more conserved within individuals **(Supplemental Figure \ref{allskin})**.
@@
 ## Association Between Environmental Stability and Network Structure Across the Human Skin Landscape
-Extensive work has illustrated differences in healthy human skin microbiome between anatomical sites, including bacteria, viruses, and fungi [@Grice:2009ee; @Findley:2013jf; @Hannigan:2015fz].
+Extensive work has illustrated differences in the healthy human skin microbiome between anatomical sites, including bacteria, viruses, and fungi [@Grice:2009ee; @Findley:2013jf; @Hannigan:2015fz].
 These communities vary by degree of skin moisture, oil, and environmental exposure.
@@
 We similarily aimed to provide an initial understanding of how phage-bacteria networks differ throughout the human body, so as to provide a baseline for future studies of how microbiome networks differ in disease states.
-We developed and implemented a network-based analytical approach that we used to evaluate the basic properties of the human microbiome through bacteria and phage *relationships*, instead of membership or diversity.
+We developed and implemented a network-based analytical approach that we used to evaluate the basic properties of the human microbiome through bacteria and phage relationships, instead of membership or diversity.
 Through the use of network theory, we leveraged extensive analytical opportunities that could be applied to understand complex ecological communities.
@@
 We found evidence that diet was sufficient to alter gut microbiome network connectivity.
-Although our sample size was small, we found evidence that high-fat diets were less connected than low-fat diets.
+Although our sample size was small, our findings suggested that high-fat diets were less connected than low-fat diets.
 This suggested that high-fat diets may lead to less stable communities with a decreased ability for individuals to influence one other.
@@
 While these findings are informative, there are certainly caveats that should be noted.
-*First*, while our infection classification model is advantageous over existing models, we recognize that, like most classification models, there remains opportunity for improvement.
+First, while our infection classification model is advantageous over existing models, we recognize that, like most classification models, there remains opportunity for improvement.
 For example, such a model is only as good as its training set.
@@
 
-*Second*, while a successful and informative proof of concept, this work was done retrospectively and relied on published research from as many as seven years ago.
+Second, while an informative proof of concept, this work was done retrospectively and relied on published research from as long as seven years ago.
 These archived datasets were limited by the technology and costs of the time, meaning the datasets were poorly powered for the statistical analysis we strive for today.
@@
 
-We demonstrated that microbiome relationship structure, and therefore the potential for microbiome genetic transfer and stability maintenance, differed significantly between body sites.
+We overall demonstrated that microbiome relationship structure, and therefore the potential for microbiome genetic transfer and stability maintenance, differed significantly between body sites.
 Other environmental factors, such as diet and individuality, also influenced relationship structures of the microbiome.
@@
 ##### Identify Bacterial CRISPRs Targeting Phages
-CRISPRs were identified from bacterial genomes using the PilerCR program (v1.06) [@Edgar:2007bh].
+Clustered Regularly Interspaced Short Palindromic Repeats (CRISPRs) were identified from bacterial genomes using the PilerCR program (v1.06) [@Edgar:2007bh].
 Resulting spacer sequences were filtered to exclude spacers shorter than 20 bp and longer than 65 bp.
@@
 ##### Identify Shared Genes Between Bacteria and Phages
-Phages may share genes with their bacterial hosts, providing us with evidence of phage-host infectious pairs.
+As a result of gene transfer or phage genome integration during infection, phages may share genes with their bacterial hosts, providing us with evidence of phage-host pairing.
 We identified shared genes between bacterial and phage genomes by assessing amino acid similarity between the genes using the Diamond protein alignment algorithm (v0.7.11.60) [@Buchfink:2015ki].
@@
 The infectious pairings between bacteria and phage OGUs were classified using the random forest model described above.
-The predicted infectious pairings and all associated metadata was saved as a graph database using Neo4j graph database software (v2.3.1) [@Neoj:Fuwr6PBN].
+The predicted infectious pairings and all associated metadata were saved as a graph database using Neo4j graph database software (v2.3.1) [@Neoj:Fuwr6PBN].
 This network was used for downstream community analysis.
@@
 ## Centrality Analysis
-Let $G(V,E)$ be an undirected, unweighted graph with $|V|=n$ nodes and $|E|=m$ edges.
+We quantified the centrality of graph vertices using three different metrics, each of which provided different information graph structure.
+When calculating these values, let $G(V,E)$ be an undirected, unweighted graph with $|V|=n$ nodes and $|E|=m$ edges.
 Also, let $\mathbf{A}$ be its corresponding adjacency matrix with entries $a_{ij} = 1$ if nodes $V_i$ and $V_j$ are connected via an edge, and $a_{ij} = 0$ otherwise.
@@
 
-The Eigenvector centrality of node $V_i$ is defined as the $i^{th}$ value in the first eigenvector of the associated adjacency matrix $\mathbf{A}$.
+The eigenvector centrality of node $V_i$ is defined as the $i^{th}$ value in the first eigenvector of the associated adjacency matrix $\mathbf{A}$.
 Conceptually, this function results in a centrality value that reflects the connections of the vertex, as well as the centrality of its neighboring vertices.
@@
 We assessed similarity between graphs by evaluating the shared centrality of their vertices, as has been done previously.
-More specifically, we calculated the dissimilarity between graphs $G_{i}$ and $G_{j}$ using the Bray-Curtis dissimilarity metric and Eigenvector centrality values such that:
+More specifically, we calculated the dissimilarity between graphs $G_{i}$ and $G_{j}$ using the Bray-Curtis dissimilarity metric and eigenvector centrality values such that:
 
@@
 
-![**Summary of Multi-Study Network Model.** *(A) Average ROC curve resulting from ten iterations used to create the phage - bacteria infection prediction model.
+![**Summary of Multi-Study Network Model.** *(A) Average ROC curve used to create the microbiome-virome infection prediction model.
 (B) Importance scores associated with the metrics used in the random forest model to predict relationships between bacteria and phages.
@@
 Each sample is colored by subject, with each sample pair collected 8-10 days apart.
-Dissimilarity was calculated using the Bray-Curtis metric based on abundance weighted Eigenvector centrality signatures, with a greater distance representing greater dissimilarity in bacteria and phage centrality and abundance.
+Dissimilarity was calculated using the Bray-Curtis metric based on abundance weighted eigenvector centrality signatures, with a greater distance representing greater dissimilarity in bacteria and phage centrality and abundance.
 (B) Quantification of gut network dissimilarity within the same subject over time (intrapersonal) and the mean dissimilarity between the subject of interest and all other subjects (interpersonal).
@@
 (D) Quantification of skin network dissimilarity within the same subject and anatomical location over time (intrapersonal) and the mean dissimilarity between the subject of interest and all other subjects at the same time and the same anatomical location (interpersonal).
-The p-value is also provided.
-*\label{intradiv}](../figures/intrapersonal_diversity.pdf){ width=90% }
+P-value was calculated using a paired Wilcoxon test.*\label{intradiv}](../figures/intrapersonal_diversity.pdf){ width=90% }
 
-![**Impact of Skin Micro-Environment on Microbiome Network Structure.**  *(A) Notched box-plot depicting differences in average Eigenvector centrality between moist, intermittently moist, and sebaceous skin sites and (B) occluded, intermittently occluded, and exposed sites.
+![**Impact of Skin Micro-Environment on Microbiome Network Structure.**  *(A) Notched box-plot depicting differences in average eigenvector centrality between moist, intermittently moist, and sebaceous skin sites and (B) occluded, intermittently occluded, and exposed sites.
 Notched box-plots were created using ggplot2 and show the median (center line), the inter-quartile range (IQR; upper and lower boxes), the highest and lowest value within 1.5 \* IQR (whiskers), outliers (dots), and the notch which provides an approximate 95% confidence interval as defined by 1.58 \* IQR / sqrt(n).
@@
 
-![**Sequencing Depth Summary.** *Number of sequences that aligned to (A) Phage and (B) Bacteria operational genomic units.
-Sequencing count aligned to OGUs was length corrected.*\label{SequenceStats}](../figures/SequenceAbund.pdf){ width=80% }
+![**Sequencing Depth Summary.** *Number of sequences that aligned to (A) Phage and (B) Bacteria operational genomic units per sample and colored by study.*\label{SequenceStats}](../figures/SequenceAbund.pdf){ width=80% }
 
@@
 
-![**Contig Summary Statistics.** *Scatter plot heat map with each hexagon representing an abundance of contigs.
+![**Contig Summary Statistics.** *Scatter plot heat map with each hexagon representing the abundance of contigs.
 Contigs are organized by length on the x-axis and the number of aligned sequences on the y-axis.*\label{ContigStats}](../figures/ContigStats.pdf){ width=50% }
@@
 B) Quantification of bacterial host strains known to exist for each phage.
-C) Genome strandedness and E) linearity of the phage reference genomes used for the dataset.*\label{ValidationOverview}](../figures/BenchmarkDataset.pdf){ width=100% }
+C) Genome strandedness and D) linearity of the phage reference genomes used for the dataset.*\label{ValidationOverview}](../figures/BenchmarkDataset.pdf){ width=100% }
 
@@
 ![**Intrapersonal vs Interpersonal Dissimilarity of the Skin.** *Quantification of skin network dissimilarity within the same subject and anatomical location over time (intrapersonal) and the mean dissimilarity between the subject of interest and all other subjects at the same time and the same anatomical location (interpersonal), separated by each anatomical site (forehead [Fh], palm [Pa], toe web [Tw], umbilicus [Um], antecubital fossa [Ac], axilla [Ax], and retroauricular crease [Ra]).
-Below is the probability distribution of the changes between intrapersonal and interpersonal diversity, representing the probability that intrapersonal dissimilarity will be lower than interpersonal dissimilarity (intrapersonal change less than zero).
-The probability that the slope will be less than zero (integral from negative infinity to zero) is provided in the top left corner.*\label{allskin}](../figures/intraallskin.pdf){ width=75% }
+P-value was calculated using a paired Wilcoxon test.*\label{allskin}](../figures/intraallskin.pdf){ width=75% }
 

### 2017-04-24 doc/manuscript.Rmd

@@
 They frequently share genetic information and work together to maintain stable ecosystems.
-Removal of members can disrupt or even collapse those ecosystems [@Haerter:2014ii; @Harcombe:2005fd; @Middelboe:2001fl; @Poisot:2011jc; @Thompson:2012ki; @Moebus:1981kp; @Flores:2013hc; @Poisot:2012fh; @Poisot071993; @Flores:2011bh; @Jover:2015ev].
-The most common approach to studying communities in isolation has been unable to capture how these communities interact and influence one another, but relationship-based network approaches allow us to capture this information.
+Removal of bacteria or phage can disrupt or even collapse those ecosystems [@Haerter:2014ii; @Harcombe:2005fd; @Middelboe:2001fl; @Poisot:2011jc; @Thompson:2012ki; @Moebus:1981kp; @Flores:2013hc; @Poisot:2012fh; @Poisot071993; @Flores:2011bh; @Jover:2015ev].
+The common approach of studying phage and bacterial communities separately has been unable to capture how these communities interact and influence one another; however relationship-based network approaches allow us to capture this information.
 We aimed to bridge this knowledge gap by leveraging machine learning and graph theory techniques to characterize human bacterial and phage communities by their relationships.
@@
 
-First we investigated human microbiome-virome networks by utilizing three published microbiome datasets with paired virus and bacterial metagenomic sequence sets [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei].
+We investigated human microbiome-virome networks by utilizing three published microbiome datasets with paired virus and bacterial metagenomic sequence sets [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei].
 Our approach built off of previous large-scale phage-bacteria microbiome network analyses by inferring interactions using metagenomic datasets, instead of using culture-based techniques [@Flores:2013hc].
 This metagenomic interaction inference model was powered beyond previous models by its inclusion of protein interaction data as well as both negative and positive interactions [@Edwards:2015iz, @Roux:2016cc].
-
 Through this approach, we were able to go beyond contemporary isolated methods to provide a basic understanding of the community network dynamics associated with healthy human phages and bacteria.
@@
 We studied the differences in virome-microbiome interaction networks across healthy human bodies by leveraging previously published sequence sets containing purified virome samples paired with bacterial metagenomes from whole metagenomic shotgun sequences.
-Our study contained three datasets, including a study of the impact of diet on the healthy human gut virome [@Minot:2011ez], the impact of anatomical location on the healthy human skin virome [@Hannigan:2015fz], and the viromes of monozygotic twins and their mothers [@Reyes:2010cwa; @Turnbaugh:2009ei].
-The viromes associated with these datasets were selected because they were subjected to virus-like particle (VLP) purification to eliminate other organismal DNA and allow for direct assessment of both the extracellular and actively replicating intracellular virome.
+Our study contained three datasets that studied impact of diet on the healthy human gut virome [@Minot:2011ez], the impact of anatomical location on the healthy human skin virome [@Hannigan:2015fz], and the viromes of monozygotic twins and their mothers [@Reyes:2010cwa; @Turnbaugh:2009ei].
+These datasets were selected because they were subjected to virus-like particle (VLP) purification to eliminate other organismal DNA and allow for direct assessment of both the extracellular and actively replicating intracellular virome **(Supplemental Figure \ref{SequenceStats} A-B)** [@Minot:2011ez, @Hannigan:2015fz, @Reyes:2010cwa; @Turnbaugh:2009ei].
 
-We utilized data that was published over the span of five years, which used different methods and technologies, and therefore yielded different sequence abundances **(Supplemental Figure \ref{SequenceStats} A-B)** [@Minot:2011ez, @Hannigan:2015fz, @Reyes:2010cwa; @Turnbaugh:2009ei].
-The bacterial and viral sequences from those publications were quality filtered and assembled into contigs.
+The bacterial and viral sequences from these publications were quality filtered and assembled into contigs.
 We generated clusters of related bacterial and phage contigs by k-mer frequency and co-abundance using the CONCOCT algorithm **(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})**.
@@
 Once our random forest model was built, we used it to classify the relationships between bacteria and phage operational genomic units in the experimental datasets.
-The relationships predicted within the three published datasets were used to construct one master network, which was built to contain the three studies as sub-networks, which themselves each contained sub-networks for each sample **(Figure \ref{RocCurve} D)**.
+The relationships predicted within the three published datasets were used to construct one master network.
+The master network contained the three studies as sub-networks, which themselves each contained sub-networks for each sample **(Figure \ref{RocCurve} D)**.
 Metadata including study, sample ID, disease, and OGU abundance within the community were stored in the master network for downstream analysis **(Supplemental Figure \ref{NetworkDiagram})**.
@@
 # Discussion
-Early, foundational microbiome work has provided a baseline understanding of the human microbiome by characterizing bacteria and viral diversity across the human body, as well as other environments [@Grice:2009eea; @Findley:2013jf; @Hannigan:2015fz; @Costello:2009im, @Consortium:2012iz; @Schloss:2005hz; @Minot:2011ez].
+Foundational microbiome work has provided a baseline understanding of the human microbiome by characterizing bacteria and viral diversity across the human body, as well as other environments [@Grice:2009eea; @Findley:2013jf; @Hannigan:2015fz; @Costello:2009im, @Consortium:2012iz; @Schloss:2005hz; @Minot:2011ez].
 We similarily aimed to provide an initial understanding of how phage-bacteria networks differ throughout the human body, so as to provide a baseline for future studies of how microbiome networks differ in disease states.
@@
 
-We found that, just as gut microbiome and virome composition and diversity are conserved in individuals [@Hannigan:2015fz; @Grice:2009eea; @Findley:2013jf; @Minot:2013ih], gut and skin microbiome network structures were also conserved within individuals over time.
+Just as gut microbiome and virome composition and diversity are conserved in individuals [@Hannigan:2015fz; @Grice:2009eea; @Findley:2013jf; @Minot:2013ih], gut and skin microbiome network structures were also conserved within individuals over time.
 Gut network structure was not conserved among family members.
@@
 
-In addition to diet, we found that skin microbiome network structure varied by skin environment.
+In addition to diet, the skin microbiome network structure varied by skin environment.
 Network structure differed between environmentally exposed and occluded skin sites.
@@
 Our results represent a step toward understanding the microbiome through interspecies relationships.
-While these findings are informative, there are certainly caveats that should be noted.
-First, while our infection classification model is advantageous over existing models, we recognize that, like most classification models, there remains opportunity for improvement.
-For example, such a model is only as good as its training set.
-Large scale experimental screens for phage and bacteria infectious interactions will provide more robust model training, and therefore improved model performance.
+Although these findings are informative, there are certainly caveats that should be noted.
+First, although our infection classification model is advantageous over existing models, we recognize that, like most classification models, ours is only as good as its training set.
+Large scale experimental screens for phage and bacteria infectious interactions will provide more robust model training and therefore improved model performance.
 Just as we have improved on previous modeling efforts, we expect that new and creative scoring metrics will be integrated into this model to further improve model performance.
+Second, although our analyses offer an informative proof of concept, this work was done retrospectively and relied on published research from as long as seven years ago.
+These archived datasets were limited by the technology and costs of the time, meaning the datasets were poorly powered for the statistical analysis we strive for today.
+Despite these limitations, we were able to present initial conclusions; however, follow-up studies will be required to validate our findings.
+These methods and results will be important for informing design and interpretation of future studies and for highlighting the importance of bacteria-virus interaction profiles as a characteristic of human-associated microbial communities.
 
-Second, while an informative proof of concept, this work was done retrospectively and relied on published research from as long as seven years ago.
-These archived datasets were limited by the technology and costs of the time, meaning the datasets were poorly powered for the statistical analysis we strive for today.
-While we were able to present initial conclusions, follow-up studies will be required to validate our findings.
-Despite their limitations, these methods and results will be important for informing design and interpretation of future studies and for highlighting the importance of bacteria-virus interaction profiles as a characteristic of human-associated microbial communities.
-
-We overall demonstrated that microbiome relationship structure, and therefore the potential for microbiome genetic transfer and stability maintenance, differed significantly between body sites.
+The virome-microbiome relationship structure, and therefore the potential for microbiome genetic transfer and stability maintenance, differed significantly between body sites.
 Other environmental factors, such as diet and individuality, also influenced relationship structures of the microbiome.
@@
 # Acknowledgments
-We thank the members of the Schloss lab for their underlying contributions. GDH is supported in part by the University of Michigan Molecular Mechanisms of Microbial Pathogenesis Fellowship.
+We thank the members of the Schloss lab for their underlying contributions.
+GDH is supported in part by the University of Michigan Molecular Mechanisms of Microbial Pathogenesis Fellowship.
+GDH and PDS were supported in part by funding from the NIH (P30DK034933, U19AI09087, and U01AI124255).
 

### 2017-04-25 doc/manuscript.Rmd
Updated with Pats suggestions.
Updates updates updates.

@@
 ---
-
-&nbsp;
 
@@
 This algorithm was applied to paired viral and bacterial metagenomic sequence sets from three previously published human cohorts.
-We organized the predicted interactions into networks that allowed us to evaluate the diversity of phage-bacteria connectedness across the human body.
+We organized the predicted interactions into networks that allowed us to evaluate phage-bacteria connectedness across the human body.
 We found that gut and skin network structures were person-specific, and were not conserved among cohabitating family members.
@@
 This metagenomic interaction inference model was powered beyond previous models by its inclusion of protein interaction data as well as both negative and positive interactions [@Edwards:2015iz, @Roux:2016cc].
-Through this approach, we were able to go beyond contemporary isolated methods to provide a basic understanding of the community network dynamics associated with healthy human phages and bacteria.
+Through this approach, we were able to go beyond contemporary isolated methods to provide a basic understanding of the network dynamics associated with healthy human phage and bacteria communities.
 By building and utilizing a microbiome-virome network, we found that different people, body sites, and anatomical locations not only support distinct microbiome membership and diversity [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei; @Grice:2009eea; @Findley:2013jf; @Costello:2009im, @Consortium:2012iz], but also support communities with distinct communication structures and propensities toward community instability.
@@
 We generated clusters of related bacterial and phage contigs by k-mer frequency and co-abundance using the CONCOCT algorithm **(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})**.
+Contig lengths within clusters ranged between $10_{3}$ and $10_{5.5}$, confirming the majority of contigs did not represent whole genomes alone **(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})**.
 We interpreted these clusters as operational units of related bacteria and phage genomes, which we defined as operational genomic units (OGUs).
@@
 The resulting random forest model performed with an AUC of `r signif(genmodelper$ROC, digits = sigfig)`, a sensitivity of `r signif(genmodelper$Sens, digits = sigfig)`, and a specificity of `r signif(genmodelper$Spec, digits = sigfig)` **(Figure \ref{RocCurve} A)**.
-The most important predictor in the model was nucleotide similarity between genes, followed by nucleotide similarity of whole genomes **(Figure \ref{RocCurve} B)**.
-Protein family interactions were moderately important to the model, while CRISPRs were not informative **(Figure \ref{RocCurve} B)**.
+The most important predictor in the model was amino acid similarity between genes (blastx), followed by nucleotide similarity of whole genomes **(Figure \ref{RocCurve} B)**.
+Protein family interactions were moderately important to the model, and CRISPRs were not informative **(Figure \ref{RocCurve} B)**.
 Approximately one third of the training set relationships yielded no score and were therefore unable to be assigned an interaction prediction **(Figure \ref{RocCurve} C)**.
@@
 Like the master network, the skin network exhibited a diameter of `r prettyNum(genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Diameter"], big.mark = ",")` (measure of graph size; the greatest number of traversed vertices required between two vertices) and included `r round(100 * genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Vertices"] / genfigurestats[c(genfigurestats$class %in% "Total"),"Vertices"], digits = 1)`% and `r round(100 * genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Edges"] / genfigurestats[c(genfigurestats$class %in% "Total"),"Edges"], digits = 1)`% of the master network nodes and edges, respectively **(Figure \ref{RocCurve} E - F)**.
-The phages and bacteria in the gut diet and twin sample sets were more sparsely related, each contained fewer than 150 vertices, fewer than 20,000 relationships, and diameters of `r prettyNum(genfigurestats[c(genfigurestats$class %in% "DietStudy"),"Diameter"], big.mark = ",")`, which indicated more sparsely related phages and bacteria **(Figure \ref{RocCurve} E - F)**.
+The phages and bacteria in the gut diet and twin sample sets were more sparsely related, with each containing fewer than 150 vertices, fewer than 20,000 relationships, and diameters of `r prettyNum(genfigurestats[c(genfigurestats$class %in% "DietStudy"),"Diameter"], big.mark = ",")`, which indicated more sparsely related phages and bacteria **(Figure \ref{RocCurve} E - F)**.
 
@@
 We accomplished this by utilizing two common centrality metrics: degree centrality and closeness centrality.
-Degree centrality, the simplest centrality metric, was defined as the number of connections each phage made to bacteria, or each bacterium made to phages.
+Degree centrality, the simplest centrality metric, was defined as the number of connections each phage made with each bacterium.
 We supplemented measurements of degree centrality with measurements of closeness centrality.
@@
 A network with higher average closeness centrality also indicates an overall greater degree of connections, which suggests a greater resilience against instability.
-Because these values are assigned to each phage and bacterium within each network, we calculated the average connectedness and corrected for the maximum potential degree of connectedness to obtain a single value for the connectedness of each graph.
+Because these values are assigned to each phage and bacterium within each sample network, we calculated the average connectedness per sample and corrected for the maximum potential degree of connectedness to obtain a single value for sample graph connectedness.
 
@@
 ## Individuality of Microbial Networks
-Skin and gut community membership and diversity are highly personal, with people remaining more similar to themselves over time compared to other people [@Grice:2009ee; @Hannigan:2015fz; @Minot:2013ih].
+Skin and gut community membership and diversity are highly personal, with people remaining more similar to themselves than to other people over time [@Grice:2009ee; @Hannigan:2015fz; @Minot:2013ih].
 We therefore hypothesized that this personal conservation extended to microbiome network structure.
-To answer this hypothesis, we calculated the degree of dissimilarity between each subject's network, based on phage and bacteria abundance as well as centrality.
+we addressed this hypothesis by calculating the degree of dissimilarity between each subject's network, based on phage and bacteria abundance as well as centrality.
 We quantified phage and bacteria centrality within each sample graph using the weighted eigenvector centrality metric.
@@
 We used this network dissimilarity metric to test whether microbiome network structures were more similar within people than between people over time.
-We found that gut microbiome network structure clustered by person (ANOSIM p-value = `r signif(interstats[c(interstats$site %in% "DietAnosim"), "prob"], digits = sigfig)`, R = `r signif(interstats[c(interstats$site %in% "AS"), "prob"], digits = sigfig)`, **Figure \ref{intradiv} A**).
+We found that gut microbiome network structures clustered by person (ANOSIM p-value = `r signif(interstats[c(interstats$site %in% "DietAnosim"), "prob"], digits = sigfig)`, R = `r signif(interstats[c(interstats$site %in% "AS"), "prob"], digits = sigfig)`, **Figure \ref{intradiv} A**).
 Network dissimilarity within each person over the 8-10 day sampling period was less than the average dissimilarity between that person and others, although this difference was not statistically significant (p-value = `r signif(interstats[c(interstats$site %in% "Diet"), "prob"], digits = sigfig)`, **Figure \ref{intradiv} B**).
@@
 Although our sample size was small, our findings suggested that high-fat diets were less connected than low-fat diets.
-This suggested that high-fat diets may lead to less stable communities with a decreased ability for individuals to influence one other.
+This suggested that high-fat diets may lead to less stable communities with a decreased ability for microbes to directly influence one other.
 We supported this finding with the observation that obesity was associated with decreased network connectivity (closeness of bacterial and phage vertices).
@@
 Likewise, intermittently moist sites demonstrated less connectedness than the more stable moist and sebaceous sites.
-Thus body sites under greater degrees of fluctuation harbored less connected, potentially less stable microbiomes.
+Togteher this data suggested body sites under greater degrees of fluctuation harbored less connected, potentially less stable microbiomes.
 
@@
 Large scale experimental screens for phage and bacteria infectious interactions will provide more robust model training and therefore improved model performance.
-Just as we have improved on previous modeling efforts, we expect that new and creative scoring metrics will be integrated into this model to further improve model performance.
+Just as we have improved on previous modeling efforts, we expect that new and creative scoring metrics will be integrated into this model to further improve performance.
 Second, although our analyses offer an informative proof of concept, this work was done retrospectively and relied on published research from as long as seven years ago.
@@
 We thank the members of the Schloss lab for their underlying contributions.
-GDH is supported in part by the University of Michigan Molecular Mechanisms of Microbial Pathogenesis Fellowship.
+GDH was supported in part by the Molecular Mechanisms in Microbial Pathogenesis Training Program (T32 AI007528).
 GDH and PDS were supported in part by funding from the NIH (P30DK034933, U19AI09087, and U01AI124255).

### 2017-05-05 doc/manuscript.Rmd
Went back and moved changed from md to rmd. I need to make it clearer that md is an int file

@@
 Most previous work has relied on studying bacteria and phages in isolation, reducing them to two separate communities.
-Such approaches are not able to capture how these communities interact to maintain stability and share information (i.e.
-horizontal gene transfer).
-We developed and implemented a network-based analytical approach to provide an initial understanding of phage-bacteria network diversity throughout the human body.
+Such approaches are not able to capture how these communities interact to maintain stability and share information (e.g., horizontal gene transfer <!-- as this isn't the only way to share information, use eg instaed of ie.
+Also, is there work that demonstrates how these communities interact to maintain stability?
+By stability, do we mean maintain diversity?
+What are the other outcomes of interactions?
+e.g., (antagonistic) co-evolution, etc.
+I know these aren't elements to fit into an Abstract, so staying tuned as to whether they appear later, as well-->).
+We developed and implemented a network-based analytical approach to describe phage-bacteria network diversity throughout the human body.
 We accomplished this by building a machine learning algorithm to predict which phages infected which bacteria in a given microbiome.
@@
 We organized the predicted interactions into networks that allowed us to evaluate phage-bacteria connectedness across the human body.
-We found that gut and skin network structures were person-specific, and were not conserved among cohabitating family members.
+We found that gut and skin network structures were person-specific and were not conserved among cohabitating family members.
 High-fat diets and obesity were associated with less connected networks.
 There were significant differences in network structure between skin sites, with those exposed to the external environment being less connected and more prone to instability.
-Together this study characterized the diversity of virome-microbiome networks across the human body and illustrated how environmental factors may influence phage-bacteria interactive dynamics.
-This work provides a baseline for future studies to investigate the role of ecological networks in conditions such as disease states.
+This study quantified and contrasted the diversity of virome-microbiome networks across the human body and illustrated how environmental factors may influence phage-bacteria interactive dynamics.
+This work provides a baseline for future studies to investigate the role of virus-microbe ecological networks in advancing our understanding of conditions, such as disease states.
 
@@
 Bacterial communities have been associated with diseases including a wide range of skin conditions [@Hannigan:2013im], acute and chronic wound healing conditions [@Hannigan:2014be; @Loesche:2016ev], and gastrointestinal diseases including inflammatory bowel disease [@He:2016ch; @Norman:2015kb], *Clostridium difficile* infections [@Seekatz:2016fz], and colorectal cancer [@Zackular:2014fba; @Baxter:2014hb].
-Altered viromes (virus communities consisting primarily of bacteriophages in humans) have also been associated with various diseases and environmental perturbations including inflammatory bowel disease [@Norman:2015kb; @Manrique:2016dx], periodontal disease [@Ly:2014ew], spread of antibiotic resistance [@Modi:2013fia], and others [@Monaco:2016ita; @Hannigan:2015fz; @Minot:2011ez; @SantiagoRodriguez:2015gd; @Abeles:2015dy; @Abeles:2014kj].
-These virus communities act in concert with bacteria as one single overall community [@Haerter:2014ii].
-Individual phages within these communities are capable of lysing their bacterial hosts, modulating their functionality through processes such as horizontal gene transfer (i.e.
-transduction) [@Moon:2015fa, @Modi:2013fi, @Ogg:1981th, @Frost:2005dn], and altering bacterial host gene expression [@Lindell:2005gz, @Tyler:2013fl, @Hargreaves:2014ja].
+Altered viromes (virus communities consisting primarily of bacteriophages) have also been associated with various diseases and perturbations, including inflammatory bowel disease [@Norman:2015kb; @Manrique:2016dx], periodontal disease [@Ly:2014ew], spread of antibiotic resistance [@Modi:2013fia], and others [@Monaco:2016ita; @Hannigan:2015fz; @Minot:2011ez; @SantiagoRodriguez:2015gd; @Abeles:2015dy; @Abeles:2014kj].
+Viruses interact with bacteria and archaea and together they comprise one total microbial community [@Haerter:2014ii].
+Viruses influence the living microbes of these communities by, e.g., lysing their microbial hosts, modulating their functionality through altered host gene expression [@Lindell:2005gz, @Tyler:2013fl, @Hargreaves:2014ja], influencing evolutionary processes, e.g., through horizontal gene transfer  [@Moon:2015fa, @Modi:2013fi, @Ogg:1981th, @Frost:2005dn] or co-evolution <!-- Koskella, B.
+& Brockhurst, M.A., 2014, Bacteria-phage coevolution as a driver of ecological and evolutionary processes in microbial communities, FEMS microbiology reviews    IMPORTANT, check out the abstract and intro of this paper--I think well-reviewed here.
+Bringing in a few of these examples to demistify what we are suggesting our network edges to imply will benefit this work and help reviewers appreciate the relevance.
+-->, and by altering ecosystem processes and elemental stoichiometry <!-- Jover, L.F., Effler, T.C., Buchan, A., Wilhelm, S.W.
+& Weitz, J.S., 2014, The elemental composition of virus particles: implications for marine biogeochemical cycles, Nature reviews.
+Microbiology, 12(7), pp.
+519-28 -->.
 
-Most previous work has relied on studying bacterial and phage communities in isolation, reducing them to two separate communities [@Norman:2015kb; @Manrique:2016dx; @Ly:2014ew; @Monaco:2016ita; @Hannigan:2015fz; @Minot:2011ez; @SantiagoRodriguez:2015gd; @Abeles:2015dy; @Abeles:2014kj].
+Most previous work has relied on studying bacterial <!-- before I go through and change all instances of bacteria to something that might include archaea too, at least where shotgun metagenomics is used, I'll pose "is this intentional?" If not, I suggest changing it --> and phage communities in isolation, reducing them to two separate communities [@Norman:2015kb; @Manrique:2016dx; @Ly:2014ew; @Monaco:2016ita; @Hannigan:2015fz; @Minot:2011ez; @SantiagoRodriguez:2015gd; @Abeles:2015dy; @Abeles:2014kj].
 In reality, bacteria and phage communities are dynamic and complex.
@@
 Removal of bacteria or phage can disrupt or even collapse those ecosystems [@Haerter:2014ii; @Harcombe:2005fd; @Middelboe:2001fl; @Poisot:2011jc; @Thompson:2012ki; @Moebus:1981kp; @Flores:2013hc; @Poisot:2012fh; @Poisot071993; @Flores:2011bh; @Jover:2015ev].
-The common approach of studying phage and bacterial communities separately has been unable to capture how these communities interact and influence one another; however relationship-based network approaches allow us to capture this information.
-We aimed to bridge this knowledge gap by leveraging machine learning and graph theory techniques to characterize human bacterial and phage communities by their relationships.
+Relationship-based network approaches allow us to capture this information.
+We leveraged machine learning and graph theory techniques to characterize human bacterial and phage communities by their relationships.
 We focused on characterizing human skin and gut virome-microbiome network diversity to provide a foundation for further studies of disease network dynamics and to gain broader insights into human microbiome network diversity across different body sites.
@@
 Our approach built off of previous large-scale phage-bacteria microbiome network analyses by inferring interactions using metagenomic datasets, instead of using culture-based techniques [@Flores:2013hc].
-This metagenomic interaction inference model was powered beyond previous models by its inclusion of protein interaction data as well as both negative and positive interactions [@Edwards:2015iz, @Roux:2016cc].
-Through this approach, we were able to go beyond contemporary isolated methods to provide a basic understanding of the network dynamics associated with healthy human phage and bacteria communities.
+This metagenomic interaction inference model was powered beyond previous models, e.g., [@Edwards:2015iz, @Roux:2016cc], by its inclusion of protein interaction data as well as both negative and positive interactions.
+Through this approach, we were able to provide a basic understanding of the network dynamics associated with healthy human phage and bacterial <!-- archaeal or "their microbial host"--> communities.
 By building and utilizing a microbiome-virome network, we found that different people, body sites, and anatomical locations not only support distinct microbiome membership and diversity [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei; @Grice:2009eea; @Findley:2013jf; @Costello:2009im, @Consortium:2012iz], but also support communities with distinct communication structures and propensities toward community instability.
-Through an improved understanding of the healthy state of network structures across the human body, we will empower future studies to begin investigating how these community structures change in disease and their overall impact on human health.
+Through an improved understanding of the healthy state of network structures across the human body, we empower future studies to begin investigating how these community structures are influenced by disease state and their overall impact on human health.
 
@@
 We studied the differences in virome-microbiome interaction networks across healthy human bodies by leveraging previously published sequence sets containing purified virome samples paired with bacterial metagenomes from whole metagenomic shotgun sequences.
-Our study contained three datasets that studied impact of diet on the healthy human gut virome [@Minot:2011ez], the impact of anatomical location on the healthy human skin virome [@Hannigan:2015fz], and the viromes of monozygotic twins and their mothers [@Reyes:2010cwa; @Turnbaugh:2009ei].
-These datasets were selected because they were subjected to virus-like particle (VLP) purification to eliminate other organismal DNA and allow for direct assessment of both the extracellular and actively replicating intracellular virome **(Supplemental Figure \ref{SequenceStats} A-B)** [@Minot:2011ez, @Hannigan:2015fz, @Reyes:2010cwa; @Turnbaugh:2009ei].
+Our study contained three datasets that explored the impact of diet on the healthy human gut virome [@Minot:2011ez], the impact of anatomical location on the healthy human skin virome [@Hannigan:2015fz], and the viromes of monozygotic twins and their mothers [@Reyes:2010cwa; @Turnbaugh:2009ei].
+These datasets were selected because they were subjected to virus-like particle (VLP) purification.
+They employed a chloroform treatment <!--did they all do this?
+did they use the same purification treatment?
+--> to eliminate organismal DNA and thereby allowing for direct assessment of both the extracellular and fully-assembled intracellular virome **(Supplemental Figure \ref{SequenceStats} A-B)** [@Minot:2011ez, @Hannigan:2015fz, @Reyes:2010cwa; @Turnbaugh:2009ei].
 
-The bacterial and viral sequences from these publications were quality filtered and assembled into contigs.
-We generated clusters of related bacterial and phage contigs by k-mer frequency and co-abundance using the CONCOCT algorithm **(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})**.
-Contig lengths within clusters ranged between $10_{3}$ and $10_{5.5}$, confirming the majority of contigs did not represent whole genomes alone **(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})**.
+Here, the bacterial and viral sequences from these studies were quality filtered and assembled into contigs.
+We generated clusters of related bacterial and phage contigs based on their k-mer frequencies and co-abundance patterns using the CONCOCT algorithm **(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})**.
+Contig lengths within clusters ranged between $10_{3}$ and $10_{5.5}$, confirming the majority of contigs represented genome fragments, rather than complete genomes  **(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})**.
 We interpreted these clusters as operational units of related bacteria and phage genomes, which we defined as operational genomic units (OGUs).
@@
 
-Four phage and bacterial genomic features were used to predict infectious relationships between bacteria and phages: 1) genome nucleotide similarities, 2) gene amino acid sequence similarities, 3) Clustered Regularly Interspaced Short Palindromic Repeat (CRISPR) targeting of phages by bacterial CRISPR spacer sequences, and 4) similarity of protein families associated with experimentally identified protein-protein interactions [@Orchard:2014hq].
+Four phage and bacterial genomic features were used to predict infectious relationships between bacteria and phages: 1) genome nucleotide similarities, 2) gene amino acid sequence similarities, 3) bacterial Clustered Regularly Interspaced Short Palindromic Repeat (CRISPR) spacer sequences that target phages, and 4) similarity of protein families associated with experimentally identified protein-protein interactions [@Orchard:2014hq].
 The resulting random forest model performed with an AUC of `r signif(genmodelper$ROC, digits = sigfig)`, a sensitivity of `r signif(genmodelper$Sens, digits = sigfig)`, and a specificity of `r signif(genmodelper$Spec, digits = sigfig)` **(Figure \ref{RocCurve} A)**.
 The most important predictor in the model was amino acid similarity between genes (blastx), followed by nucleotide similarity of whole genomes **(Figure \ref{RocCurve} B)**.
-Protein family interactions were moderately important to the model, and CRISPRs were not informative **(Figure \ref{RocCurve} B)**.
+Protein family interactions were moderately important to the model, and CRISPRs were not informative <!-- how many of the bacteria had Cas proteins?
+At least 50%?
+How informative did you expect CRISPRs to be in your training set?
+And how does this relate to how informative we expect them to be across bacteria/archaea?
+Most work predicting phage-host interactions from metagenomic data have relied on CRISPR linkages, so we might want to contextualize why it was not informative, as those readers and community members will be wondering about this.--> **(Figure \ref{RocCurve} B)**.
 Approximately one third of the training set relationships yielded no score and were therefore unable to be assigned an interaction prediction **(Figure \ref{RocCurve} C)**.
@@
 In addition to diet, we also observed an association between obesity and network structure **(Figure \ref{dietnetworks} C-D)**.
-The obesity-associated network demonstrated a higher degree centrality **(Figure \ref{dietnetworks} C)** but less closeness centrality compared to the healthy controls **(Figure \ref{dietnetworks} D)**.
+The obesity-associated network demonstrated a higher degree centrality **(Figure \ref{dietnetworks} C)**, but less closeness centrality compared to the healthy controls **(Figure \ref{dietnetworks} D)**.
 These results suggest that the obesity network was overall less connected, having microbes further from all other microbes within the community.
@@
 We therefore hypothesized that this personal conservation extended to microbiome network structure.
-we addressed this hypothesis by calculating the degree of dissimilarity between each subject's network, based on phage and bacteria abundance as well as centrality.
+We addressed this hypothesis by calculating the degree of dissimilarity between each subject's network, based on phage and bacteria abundance and centrality.
 We quantified phage and bacteria centrality within each sample graph using the weighted eigenvector centrality metric.
-This metric defines central phages as those that are highly abundant and infect many distinct bacteria which themselves are abundant and infected by many other phages.
+This metric defines central phages as those that are highly abundant <!-- how was abundance determined?
+read mapping to contigs, "size" of OGUs, as in how many contigs belonged to an OGU group?
+a combination --> and infect many distinct bacteria which themselves are abundant and infected by many other phages.
 Similarly, bacterial centrality was defined as those bacteria that were both abundant and connected to numerous phages that were themselves connected to many bacteria.
@@
 
-Skin microbiome network structure was strongly conserved within individuals (p-value < 0.001, **Figure \ref{intradiv} D**) and more adequately confirmed by our larger skin dataset.
-This distribution was similar when separated by anatomical sites, with most sites also being statistically significantly more conserved within individuals **(Supplemental Figure \ref{allskin})**.
+Skin microbiome network structure was strongly conserved within individuals (p-value < 0.001, **Figure \ref{intradiv} D**).
+This distribution was similar when separated by anatomical sites.
+Most sites were statistically significantly more conserved within individuals **(Supplemental Figure \ref{allskin})**.
 
 ## Association Between Environmental Stability and Network Structure Across the Human Skin Landscape
-Extensive work has illustrated differences in the healthy human skin microbiome between anatomical sites, including bacteria, viruses, and fungi [@Grice:2009ee; @Findley:2013jf; @Hannigan:2015fz].
+Extensive work has illustrated differences in XYZ <!-- diversity?
+community structure?
+--> in the healthy human skin microbiome between anatomical sites, including bacteria, viruses, and fungi [@Grice:2009ee; @Findley:2013jf; @Hannigan:2015fz].
 These communities vary by degree of skin moisture, oil, and environmental exposure.
-We hypothesized that like microbial composition and diversity, microbial network structure would be specific to anatomical sites.
-We tested this hypothesis by evaluating the changes in network structure between anatomical sites within our skin dataset.
+As viruses are known to influence microbial XYZ <!-- above, maybe community structure and diversity --> , we hypothesized that microbe-virus network structure would be specific to anatomical sites, as well.
+To test this, we evaluated the changes in network structure between anatomical sites within the skin dataset.
 
-We quantified the average centrality of each sample using the weighted eigenvector centrality metric.
-We found that intermittently moist skin sites (dynamic sites that fluctuate between being moist and dry) were significantly less connected than the more stable moist and sebaceous environments (p-value < 0.001, **Figure \ref{skinnetwork} A)**.
-We also found that skin sites that were protected from the environment (occluded) were much more highly connected than those that were constantly exposed to the environment or only intermittently occluded (p-value < 0.001, **Figure \ref{skinnetwork} B)**.
+The average centrality of each sample was quantified using the weighted eigenvector centrality metric.
+Intermittently moist skin sites (dynamic sites that fluctuate between being moist and dry) were significantly less connected than the more stable moist and sebaceous environments (p-value < 0.001, **Figure \ref{skinnetwork} A)**.
+Also, skin sites that were protected from the environment (occluded) were much more highly connected than those that were constantly exposed to the environment or only intermittently occluded (p-value < 0.001, **Figure \ref{skinnetwork} B)**.
 
-We supplemented this analysis by comparing the network signatures using the centrality dissimilarity approach described above.
+To supplement this analysis, we compared the network signatures using the centrality dissimilarity approach described above.
 The dissimilarity between samples was a function of shared relationships, degree of centrality, and bacteria/phage abundance.
 When using this supplementary approach, we found that network structures significantly clustered by moisture, sebaceous, and intermittently moist status **(Figure \ref{skinnetwork} C,E)**.
-We also found that occluded sites were significantly different from exposed and intermittently occluded sites, but there was no difference between exposed and intermittently occluded sites **(Figure \ref{skinnetwork} D,F)**.
+Occluded sites were significantly different from exposed and intermittently occluded sites, but there was no difference between exposed and intermittently occluded sites **(Figure \ref{skinnetwork} D,F)**.
 These findings provide further support that skin microbiome network structure differs significantly between skin sites.
@@
 # Discussion
+<!-- Much of the discussion repeats the statements in the results, it felt a bit redundant.
+At times, the discussion even repeated itself (you'll see I deleted most of the last paragraph).
+There are probably ways to be less repetitive and more provocative or to bring in discussion of more recent work?
+Or to reduce it even more by removing redundancies.
+Alternatively, could remove the redundancies in the results...as in only state the results, and nt reference previous work.
+Then could discuss how findings relate to previous work only in the discussion, rather than the results...which does happen at times in current form.-->
 Foundational microbiome work has provided a baseline understanding of the human microbiome by characterizing bacteria and viral diversity across the human body, as well as other environments [@Grice:2009eea; @Findley:2013jf; @Hannigan:2015fz; @Costello:2009im, @Consortium:2012iz; @Schloss:2005hz; @Minot:2011ez].
-We similarily aimed to provide an initial understanding of how phage-bacteria networks differ throughout the human body, so as to provide a baseline for future studies of how microbiome networks differ in disease states.
-We developed and implemented a network-based analytical approach that we used to evaluate the basic properties of the human microbiome through bacteria and phage relationships, instead of membership or diversity.
-Through the use of network theory, we leveraged extensive analytical opportunities that could be applied to understand complex ecological communities.
-We utilized metrics of connectivity to model the extent to which communities of bacteria and phages interact (e.g.
-horizontal gene transfer, modulated bacterial gene expression, alterations in abundance).
-By pursuing this goal, we aimed to provide an initial understanding of human microbiome networks that will power future studies to use similar approaches in other settings, such as disease states.
+Here, we offer an initial understanding of how phage-bacteria networks differ throughout the human body, so as to provide a baseline for future studies of how and why microbiome networks differ in disease states.
+We developed and implemented a network-based analytical approach to evaluate the basic properties of the human microbiome through bacteria and phage relationships, instead of membership or diversity alone.
+This enabled the application of network theory to provide a new perspective on complex ecological communities.
+We utilized metrics of connectivity to model the extent to which communities of bacteria and phages interact (e.g., horizontal gene transfer, modulated bacterial gene expression, alterations in abundance <!--again, I would keep building on this, or just build on it once in the intro, but then be sure not to parenthetically distill it to only these three examples in the discussion here-->).
+<!-- what I deleted here is starting to be overstated throughout, no need to say it again here -->
 
-Just as gut microbiome and virome composition and diversity are conserved in individuals [@Hannigan:2015fz; @Grice:2009eea; @Findley:2013jf; @Minot:2013ih], gut and skin microbiome network structures were also conserved within individuals over time.
-Gut network structure was not conserved among family members.
-These findings suggested that microbiome network properties, such as stability and the potential for horizontal gene transfer, were personal and may be impacted by personal factors ranging from the body's immune system to environmental conditions such as climate and diet.
+Just as gut microbiome and virome composition and diversity are conserved in individuals [@Hannigan:2015fz; @Grice:2009eea; @Findley:2013jf; @Minot:2013ih], gut and skin microbiome network structures were conserved within individuals over time.
+Yet, gut network structure was not conserved among family members.
+These findings suggested that microbiome network properties, such as stability and the potential for horizontal gene transfer <!-- and lots of other things related to host and virus eco and evo processes -->, were personal and may be impacted by personal factors ranging from the body's immune system to environmental conditions, such as climate and diet <!-- though I'd argue that the immune system is a component/control on the environmental condition of the microbial communities -->.
 The ability of environmental conditions to shape gut and skin microbiome network structure was supported by our finding that diet and skin location were associated with altered network structures.
@@
 Likewise, intermittently moist sites demonstrated less connectedness than the more stable moist and sebaceous sites.
-Togteher this data suggested body sites under greater degrees of fluctuation harbored less connected, potentially less stable microbiomes.
+Together these data suggested body sites under greater degrees of fluctuation harbored less connected, potentially less stable microbiomes.
 
-Our results represent a step toward understanding the microbiome through interspecies relationships.
-Although these findings are informative, there are certainly caveats that should be noted.
+While these findings take us an important step closer to understanding the microbiome through interspecies relationships, there are caveats to the approach that should be noted.
 First, although our infection classification model is advantageous over existing models, we recognize that, like most classification models, ours is only as good as its training set.
-Large scale experimental screens for phage and bacteria infectious interactions will provide more robust model training and therefore improved model performance.
+Large-scale experimental screens for phage and bacteria infectious interactions that report high-confidence negative interactions (i.e., no infection) will provide more robust model training and improved model performance.
 Just as we have improved on previous modeling efforts, we expect that new and creative scoring metrics will be integrated into this model to further improve performance.
 Second, although our analyses offer an informative proof of concept, this work was done retrospectively and relied on published research from as long as seven years ago.
-These archived datasets were limited by the technology and costs of the time, meaning the datasets were poorly powered for the statistical analysis we strive for today.
-Despite these limitations, we were able to present initial conclusions; however, follow-up studies will be required to validate our findings.
-These methods and results will be important for informing design and interpretation of future studies and for highlighting the importance of bacteria-virus interaction profiles as a characteristic of human-associated microbial communities.
+These archived datasets were limited by the technology and costs of the time, meaning the datasets were suboptimally powered for the statistical analysis we strive for today.
+Despite these limitations, we were able to present initial conclusions.
+Follow-up studies will validate our findings and inform the design and interpretation of future studies.
 
-The virome-microbiome relationship structure, and therefore the potential for microbiome genetic transfer and stability maintenance, differed significantly between body sites.
-Other environmental factors, such as diet and individuality, also influenced relationship structures of the microbiome.
-This information builds on previous work by suggesting that bacteria and phages not only preferentially colonize different body sites, but also interact differently, allowing for different communication structures and capacities for maintaining stability.
-Our results provide a foundation for future studies to begin evaluating how microbiome network dynamics change in disease states, and how the information can be applied therapeutically.
+Our work takes a first step towards defining bacteria-virus interaction profiles as a characteristic of human-associated microbial communities.
+We provide a foundation for future studies to begin evaluating how microbiome network dynamics change in disease states, and how the information can be applied therapeutically.
 

### 2017-05-08 doc/manuscript.Rmd
Edit update.

@@
 Most previous work has relied on studying bacteria and phages in isolation, reducing them to two separate communities.
-Such approaches are not able to capture how these communities interact to maintain stability and share information (e.g., horizontal gene transfer <!-- as this isn't the only way to share information, use eg instaed of ie.
+Such approaches are not able to capture how these communities interact to maintain stability/diversity, co-evolve, and share genetic information (e.g.
+horizontal gene transfer <!--X as this isn't the only way to share information, use eg instaed of ie.
 Also, is there work that demonstrates how these communities interact to maintain stability?
@@
 Altered viromes (virus communities consisting primarily of bacteriophages) have also been associated with various diseases and perturbations, including inflammatory bowel disease [@Norman:2015kb; @Manrique:2016dx], periodontal disease [@Ly:2014ew], spread of antibiotic resistance [@Modi:2013fia], and others [@Monaco:2016ita; @Hannigan:2015fz; @Minot:2011ez; @SantiagoRodriguez:2015gd; @Abeles:2015dy; @Abeles:2014kj].
-Viruses interact with bacteria and archaea and together they comprise one total microbial community [@Haerter:2014ii].
-Viruses influence the living microbes of these communities by, e.g., lysing their microbial hosts, modulating their functionality through altered host gene expression [@Lindell:2005gz, @Tyler:2013fl, @Hargreaves:2014ja], influencing evolutionary processes, e.g., through horizontal gene transfer  [@Moon:2015fa, @Modi:2013fi, @Ogg:1981th, @Frost:2005dn] or co-evolution <!-- Koskella, B.
-& Brockhurst, M.A., 2014, Bacteria-phage coevolution as a driver of ecological and evolutionary processes in microbial communities, FEMS microbiology reviews    IMPORTANT, check out the abstract and intro of this paper--I think well-reviewed here.
-Bringing in a few of these examples to demistify what we are suggesting our network edges to imply will benefit this work and help reviewers appreciate the relevance.
--->, and by altering ecosystem processes and elemental stoichiometry <!-- Jover, L.F., Effler, T.C., Buchan, A., Wilhelm, S.W.
-& Weitz, J.S., 2014, The elemental composition of virus particles: implications for marine biogeochemical cycles, Nature reviews.
-Microbiology, 12(7), pp.
-519-28 -->.
+Viruses interact with bacteria, which together comprise a major portion of total microbial communities [@Haerter:2014ii].
+Viruses influence the living microbes of these communities by, for example, lysing their microbial hosts, modulating their functionality through altered host gene expression [@Lindell:2005gz, @Tyler:2013fl, @Hargreaves:2014ja], influencing evolutionary processes through, for example, horizontal gene transfer [@Moon:2015fa, @Modi:2013fi, @Ogg:1981th, @Frost:2005dn] or co-evolution [@Koskella:2014ds], and by altering ecosystem processes and elemental stoichiometry [@Jover:2014gq].
 
-Most previous work has relied on studying bacterial <!-- before I go through and change all instances of bacteria to something that might include archaea too, at least where shotgun metagenomics is used, I'll pose "is this intentional?" If not, I suggest changing it --> and phage communities in isolation, reducing them to two separate communities [@Norman:2015kb; @Manrique:2016dx; @Ly:2014ew; @Monaco:2016ita; @Hannigan:2015fz; @Minot:2011ez; @SantiagoRodriguez:2015gd; @Abeles:2015dy; @Abeles:2014kj].
+Most previous human microbiome work has focused on bacterial and viral communities, but have relied on studying them in isolation by reducing them to two separate communities [@Norman:2015kb; @Manrique:2016dx; @Ly:2014ew; @Monaco:2016ita; @Hannigan:2015fz; @Minot:2011ez; @SantiagoRodriguez:2015gd; @Abeles:2015dy; @Abeles:2014kj].
 In reality, bacteria and phage communities are dynamic and complex.
@@
 We leveraged machine learning and graph theory techniques to characterize human bacterial and phage communities by their relationships.
-We focused on characterizing human skin and gut virome-microbiome network diversity to provide a foundation for further studies of disease network dynamics and to gain broader insights into human microbiome network diversity across different body sites.
+We focused on characterizing human skin and gut virus-bacteria network diversity to provide a foundation for further studies of disease network dynamics and to gain broader insights into human microbiome network diversity across different body sites.
 
-We investigated human microbiome-virome networks by utilizing three published microbiome datasets with paired virus and bacterial metagenomic sequence sets [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei].
+We investigated human bacterial-phage networks by utilizing three published microbiome datasets which contained paired virus and total metagenomic sequence sets, which primarily consisted of bacteria [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei].
 Our approach built off of previous large-scale phage-bacteria microbiome network analyses by inferring interactions using metagenomic datasets, instead of using culture-based techniques [@Flores:2013hc].
-This metagenomic interaction inference model was powered beyond previous models, e.g., [@Edwards:2015iz, @Roux:2016cc], by its inclusion of protein interaction data as well as both negative and positive interactions.
-Through this approach, we were able to provide a basic understanding of the network dynamics associated with healthy human phage and bacterial <!-- archaeal or "their microbial host"--> communities.
+This metagenomic interaction inference model was powered beyond previous models [@Edwards:2015iz, @Roux:2016cc] by its inclusion of protein interaction data as well as the inclusion of both negative and positive interactions.
+Through this approach, we were able to provide a basic understanding of the network dynamics associated with healthy human phage and bacterial communities.
 By building and utilizing a microbiome-virome network, we found that different people, body sites, and anatomical locations not only support distinct microbiome membership and diversity [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei; @Grice:2009eea; @Findley:2013jf; @Costello:2009im, @Consortium:2012iz], but also support communities with distinct communication structures and propensities toward community instability.
@@
 ## Cohort Curation and Sample Processing
-We studied the differences in virome-microbiome interaction networks across healthy human bodies by leveraging previously published sequence sets containing purified virome samples paired with bacterial metagenomes from whole metagenomic shotgun sequences.
-Our study contained three datasets that explored the impact of diet on the healthy human gut virome [@Minot:2011ez], the impact of anatomical location on the healthy human skin virome [@Hannigan:2015fz], and the viromes of monozygotic twins and their mothers [@Reyes:2010cwa; @Turnbaugh:2009ei].
-These datasets were selected because they were subjected to virus-like particle (VLP) purification.
-They employed a chloroform treatment <!--did they all do this?
-did they use the same purification treatment?
---> to eliminate organismal DNA and thereby allowing for direct assessment of both the extracellular and fully-assembled intracellular virome **(Supplemental Figure \ref{SequenceStats} A-B)** [@Minot:2011ez, @Hannigan:2015fz, @Reyes:2010cwa; @Turnbaugh:2009ei].
+We studied the differences in virus-bacteria interaction networks across healthy human bodies by leveraging previously published sequence sets containing purified virome samples paired with bacterial metagenomes from whole metagenomic shotgun sequences.
+Our study contained three datasets, which explored the impact of diet on the healthy human gut virome [@Minot:2011ez], the impact of anatomical location on the healthy human skin virome [@Hannigan:2015fz], and the viromes of monozygotic twins and their mothers [@Reyes:2010cwa; @Turnbaugh:2009ei].
+These datasets were selected because they included virome samples subjected to virus-like particle (VLP) purification.
+They employed combinations of filtration, chloroform/DNase treatment, and Cesium Chloride treatments to eliminate organismal DNA and thereby allow for direct assessment of both the extracellular and fully-assembled intracellular virome **(Supplemental Figure \ref{SequenceStats} A-B)** [@Minot:2011ez, @Hannigan:2015fz, @Reyes:2010cwa; @Turnbaugh:2009ei].
+While the whole metagenomic shotgun sequence samples were not subjected to purification, they primarily consisted of bacteria and other microbiome members (e.g.
+fungi, archaea, *Demodex* mites, etc) were not considered in this initial study.
 
@@
 Contig lengths within clusters ranged between $10_{3}$ and $10_{5.5}$, confirming the majority of contigs represented genome fragments, rather than complete genomes  **(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})**.
+We also confirmed that the majority of bacterial clusters were indeed identified as bacteria (CITE FIGURE HERE).
 We interpreted these clusters as operational units of related bacteria and phage genomes, which we defined as operational genomic units (OGUs).
@@
 The most important predictor in the model was amino acid similarity between genes (blastx), followed by nucleotide similarity of whole genomes **(Figure \ref{RocCurve} B)**.
-Protein family interactions were moderately important to the model, and CRISPRs were not informative <!-- how many of the bacteria had Cas proteins?
-At least 50%?
-How informative did you expect CRISPRs to be in your training set?
-And how does this relate to how informative we expect them to be across bacteria/archaea?
-Most work predicting phage-host interactions from metagenomic data have relied on CRISPR linkages, so we might want to contextualize why it was not informative, as those readers and community members will be wondering about this.--> **(Figure \ref{RocCurve} B)**.
+Protein family interactions were moderately important to the model, and CRISPRs were largely uninformative, due to the lack of identifiable CRISPRs in the dataset and their redundancy with the nucleotide similarity methods **(Figure \ref{RocCurve} B)**.
 Approximately one third of the training set relationships yielded no score and were therefore unable to be assigned an interaction prediction **(Figure \ref{RocCurve} C)**.
@@
 We quantified phage and bacteria centrality within each sample graph using the weighted eigenvector centrality metric.
-This metric defines central phages as those that are highly abundant <!-- how was abundance determined?
-read mapping to contigs, "size" of OGUs, as in how many contigs belonged to an OGU group?
-a combination --> and infect many distinct bacteria which themselves are abundant and infected by many other phages.
+This metric defines central phages as those that are highly abundant ($A_{O}$ as defined in the methods) and infect many distinct bacteria which themselves are abundant and infected by many other phages.
 Similarly, bacterial centrality was defined as those bacteria that were both abundant and connected to numerous phages that were themselves connected to many bacteria.
@@
 ## Association Between Environmental Stability and Network Structure Across the Human Skin Landscape
-Extensive work has illustrated differences in XYZ <!-- diversity?
-community structure?
---> in the healthy human skin microbiome between anatomical sites, including bacteria, viruses, and fungi [@Grice:2009ee; @Findley:2013jf; @Hannigan:2015fz].
+Extensive work has illustrated differences in diversity and composition of the healthy human skin microbiome between anatomical sites, including bacteria, virus, and fungal communities [@Grice:2009ee; @Findley:2013jf; @Hannigan:2015fz].
 These communities vary by degree of skin moisture, oil, and environmental exposure.
-As viruses are known to influence microbial XYZ <!-- above, maybe community structure and diversity --> , we hypothesized that microbe-virus network structure would be specific to anatomical sites, as well.
+As viruses are known to influence microbial diversity and community composition, we hypothesized that microbe-virus network structure would be specific to anatomical sites, as well.
 To test this, we evaluated the changes in network structure between anatomical sites within the skin dataset.
@@
 Or to reduce it even more by removing redundancies.
-Alternatively, could remove the redundancies in the results...as in only state the results, and nt reference previous work.
+Alternatively, could remove the redundancies in the results...as in only state the results, and not reference previous work.
 Then could discuss how findings relate to previous work only in the discussion, rather than the results...which does happen at times in current form.-->
@@
 This enabled the application of network theory to provide a new perspective on complex ecological communities.
-We utilized metrics of connectivity to model the extent to which communities of bacteria and phages interact (e.g., horizontal gene transfer, modulated bacterial gene expression, alterations in abundance <!--again, I would keep building on this, or just build on it once in the intro, but then be sure not to parenthetically distill it to only these three examples in the discussion here-->).
-<!-- what I deleted here is starting to be overstated throughout, no need to say it again here -->
+We utilized metrics of connectivity to model the extent to which communities of bacteria and phages interact, through mechanisms including horizontal gene transfer, modulated bacterial gene expression, and alterations in abundance.
 
@@
 Yet, gut network structure was not conserved among family members.
-These findings suggested that microbiome network properties, such as stability and the potential for horizontal gene transfer <!-- and lots of other things related to host and virus eco and evo processes -->, were personal and may be impacted by personal factors ranging from the body's immune system to environmental conditions, such as climate and diet <!-- though I'd argue that the immune system is a component/control on the environmental condition of the microbial communities -->.
+These findings suggested that microbiome network properties, including stability and the potential for horizontal gene transfer <!-- and lots of other things related to host and virus eco and evo processes -->, were personal and may be impacted by personal factors ranging from the body's immune system to environmental conditions, such as climate and diet <!-- though I'd argue that the immune system is a component/control on the environmental condition of the microbial communities -->.
 The ability of environmental conditions to shape gut and skin microbiome network structure was supported by our finding that diet and skin location were associated with altered network structures.

### 2017-05-08 bin/CalculatePredModel.R
Meh a lot of stuff. Added getting longest rep contig.

@@ -9,6 +9,7 @@
 list.of.packages <- c("RNeo4j", "ggplot2", "C50", "caret", "wesanderson", "plotROC")
 new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
 if(length(new.packages)) install.packages(new.packages)
+lapply(list.of.packages, library, character.only = TRUE)
 
 #################
 # Set Libraries #
@@ -131,21 +132,20 @@ excludedgraph <- ggplot(Exclusiondf[order(Exclusiondf$ExclusionStatus, decreasin
   ylab("Percent of Total Samples")
 
 # Save the model to a file so that it can be used later
-save(outmodel, file="./data/rfinteractionmodel.RData")
+save(outmodel, excludedgraph, file="./data/rfinteractionmodel.RData")
+load(file="./data/rfinteractionmodel.RData")
 
 #############
 # Plot Data #
 #############
 roclobster <- ggplot(outmodel$pred, aes(d = obs, m = NotInteracts)) +
   geom_roc(n.cuts = 0, color = wes_palette("Royal1")[2]) +
-  theme_classic() +
+  style_roc() +
   theme(
     axis.line.x = element_line(colour = "black"),
     axis.line.y = element_line(colour = "black")
   ) +
-  geom_segment(aes(x = 0, y = 0, xend = 1, yend = 1), linetype=2, colour=wes_palette("Royal1")[1]) +
-  ylab("Sensitivity") +
-  xlab(paste("Inverse Specificity"))
+  geom_segment(aes(x = 0, y = 0, xend = 1, yend = 1), linetype=2, colour=wes_palette("Royal1")[1])
 
 # Plot prediction points
 rocdensity <- ggplot(outmodel$pred, aes(x=Interacts, group=obs, fill=obs)) +
@@ -176,6 +176,8 @@ importanceplot <- ggplot(vardf, aes(x=categories, y=Overall)) +
 
 barplots <- plot_grid(importanceplot, excludedgraph, labels=c("C", "D"), nrow=1)
 horizontal <- plot_grid(rocdensity, barplots, labels=c("B", ""), nrow=2)
+
+save(roclobster, rocdensity, vardf, importanceplot, file="./data/modelplots.Rdata")
 
 pdf(file="./figures/rocCurves.pdf",
 height=6,

### 2017-05-08 bin/ClusterStats.R

@@ -1,0 +1,59 @@
+# CluserStats.R
+# Geoffrey Hannigan
+# Schloss Lab
+# University of Michigan
+
+write("Performing final contig stats processes.", stderr())
+
+library("optparse")
+library("plyr")
+library("ggplot2")
+library("wesanderson")
+library("cowplot")
+
+contiglength <- read.delim("./data/PhageContigStats/ContigLength.tsv", head = FALSE, sep = "\t")
+bcc <- read.delim(file = "./data/ContigClustersBacteria/clustering_gt1000.csv", head = FALSE, sep = ",")
+pcc <- read.delim(file = "./data/ContigClustersPhage/clustering_gt1000.csv", head = FALSE, sep = ",")
+
+mpcc <- merge(pcc, contiglength, by = "V1")
+mpcc <- ddply(mpcc, "V2.x", summarize, abund = length(V1), alen = mean(V2.y))
+mbcc <- merge(bcc, contiglength, by = "V1")
+mbcc <- ddply(mbcc, "V2.x", summarize, abund = length(V1), alen = mean(V2.y))
+
+phageclust <- ggplot(mpcc, aes(x = alen, y = abund)) +
+    theme_classic() +
+    scale_x_log10(
+       breaks = scales::trans_breaks("log10", function(x) 10^x),
+       labels = scales::trans_format("log10", scales::math_format(10^.x))
+     ) +
+     scale_y_log10(
+       breaks = scales::trans_breaks("log10", function(x) 10^x),
+       labels = scales::trans_format("log10", scales::math_format(10^.x))) +
+     annotation_logticks() +
+     geom_point() +
+     xlab("Average Cluster Length (bp)") +
+     ylab("Cluster Contig Count") +
+     ggtitle("Phage Operational Genomic Units")
+
+bacclust <- ggplot(mbcc, aes(x = alen, y = abund)) +
+    theme_classic() +
+    scale_x_log10(
+       breaks = scales::trans_breaks("log10", function(x) 10^x),
+       labels = scales::trans_format("log10", scales::math_format(10^.x))
+     ) +
+     scale_y_log10(
+       breaks = scales::trans_breaks("log10", function(x) 10^x),
+       labels = scales::trans_format("log10", scales::math_format(10^.x))) +
+     annotation_logticks() +
+     geom_point() +
+     xlab("Average Cluster Length (bp)") +
+     ylab("Cluster Contig Count") +
[…]

### 2017-05-08 bin/CompareSkin.R

@@ -45,11 +45,11 @@ totalgraph <- rbind(graphdfTP2, graphdfTP3)
 format(object.size(totalgraph), units = "MB")
 
 # Run subsampling
-uniquephagegraph <- unique(totalgraph[-c(2,7)])
+uniquephagegraph <- unique(totalgraph[-c(2,7:9)])
 phageminseq <- quantile(ddply(uniquephagegraph, c("PatientID", "Location", "TimePoint"), summarize, sum = sum(as.numeric(PhageAbundance)))$sum, 0.05)
 print(format(object.size(uniquephagegraph), units = "MB"))
 
-uniquebacteriagraph <- unique(totalgraph[-c(1,6)])
+uniquebacteriagraph <- unique(totalgraph[-c(1,6,7,9)])
 bacminseq <- quantile(ddply(uniquebacteriagraph, c("PatientID", "Location", "TimePoint"), summarize, sum = sum(as.numeric(BacteriaAbundance)))$sum, 0.05)
 print(format(object.size(uniquephagegraph), units = "MB"))
 
@@ -121,7 +121,7 @@ rdfbacteria <- rdfbacteria[-c(1:4)]
 totalgraphcombo <- totalgraph
 totalgraphcombo$combophage <- paste(totalgraphcombo$from, totalgraphcombo$PatientID, totalgraphcombo$Location, totalgraphcombo$TimePoint, sep = "__")
 totalgraphcombo$combobacteria <- paste(totalgraphcombo$to, totalgraphcombo$PatientID, totalgraphcombo$Location, totalgraphcombo$TimePoint, sep = "__")
-totalgraphcombo <- totalgraphcombo[-c(1:7)]
+totalgraphcombo <- totalgraphcombo[-c(1:9)]
 
 format(object.size(totalgraphcombo), units = "MB")
 format(object.size(rdfphage), units = "KB")
@@ -200,7 +200,7 @@ rcentraldf$dg <- as.numeric(as.character(rcentraldf$dg))
 rcentraldf$di <- as.numeric(as.character(rcentraldf$di))
 
 
-ggplot(rcentraldf, aes(x = factor(location), y = ec)) +
+ggplot(rcentraldf, aes(x = factor(location), y = cl)) +
 	theme_classic() +
 	geom_boxplot(notch = TRUE, fill="gray") +
 	ylab("Alpha Centrality") +
[…]

### 2017-05-08 bin/FinalizeContigStats.R

@@ -9,6 +9,7 @@ library("optparse")
 library("plyr")
 library("ggplot2")
 library("wesanderson")
+library("hexbin")
 
 option_list <- list(
   make_option(c("-l", "--lengths"),
@@ -27,8 +28,8 @@ opt_parser <- OptionParser(option_list = option_list);
 opt <- parse_args(opt_parser);
 
 # Import files
-contiglength <- read.delim("./data/PhageContigStats/ContigLength.tsv", head = FALSE, sep = "\t")
-contigcounts <- read.delim("./data/PhageContigStats/FinalContigCounts.tsv", head = TRUE, sep = "\t")
+contiglength <- read.delim(opt$lengths, head = FALSE, sep = "\t")
+contigcounts <- read.delim(opt$counts, head = TRUE, sep = "\t")
 
 head(contiglength)
 head(contigcounts)
@@ -48,7 +49,7 @@ contigstatsplot <- ggplot(lengthcount, aes(x = Length, y = Count)) +
        breaks = scales::trans_breaks("log10", function(x) 10^x),
        labels = scales::trans_format("log10", scales::math_format(10^.x))) +
      annotation_logticks() +
-     geom_point() +
+     stat_binhex(bins = 100) +
      xlab("Length (bp)") +
      ylab("Sequencing Depth")
 

### 2017-05-08 bin/GetLongestContig.R

@@ -1,0 +1,56 @@
+# GetLongestContig.R
+# Geoffrey Hannigan
+# Schloss Lab
+# University of Michigan
+
+library("optparse")
+library("ggplot2")
+library("cowplot")
+library("plyr")
+
+option_list <- list(
+  make_option(c("-i", "--input"),
+    type = "character",
+    default = NULL,
+    help = "Contig length table.",
+    metavar = "character"),
+  make_option(c("-t", "--toplength"),
+    type = "integer",
+    default = 1,
+    help = "Amount of top lengths to report.",
+    metavar = "character"),
+  make_option(c("-c", "--clusters"),
+    type = "character",
+    default = NULL,
+    help = "Contig cluster id table.",
+    metavar = "character"),
+  make_option(c("-o", "--out"),
+    type = "character",
+    default = NULL,
+    help = "Output name.",
+    metavar = "character")
+)
+
+opt_parser <- OptionParser(option_list=option_list);
+opt <- parse_args(opt_parser);
+
+# Virus
+
+inputfile <- read.delim(opt$input, head = FALSE, sep = "\t")
+colnames(inputfile) <- c("ContigID", "Length")
+
+clustersfile <- read.delim(opt$clusters, head = FALSE, sep = ",")
+colnames(clustersfile) <- c("ContigID", "Cluster")
+
+mergeclust <- merge(inputfile, clustersfile, by = "ContigID")
+
+# Get top lengths
+topcontigsbylength <- ddply(mergeclust, "Cluster", function(x) head(x[order(x$Length, decreasing = TRUE) , ], opt$toplength))
+
+write.table(
+  x = topcontigsbylength,
+  file = opt$out,
+  quote = FALSE,
+  sep = "\t",
+  row.names = FALSE
+)

### 2017-05-08 bin/PlotBenchmarkInteractions.R

@@ -25,7 +25,7 @@ input <- read.delim(file="./data/ValidationSet/Interactions.tsv",
   header=TRUE)
 genometypes <- read.delim(file="./data/ValidationSet/PhageGenomeTypes.tsv",
   sep="\t",
-  header=FALSE)
+  header = FALSE)
 genometypes$V1 <- gsub("_", " ", genometypes$V1, perl=TRUE)
 samplemetadata <- read.delim(file="./data/PublishedDatasets/metadatatable.tsv",
   sep="\t",
@@ -97,7 +97,7 @@ sitegeocodes <- geocode(locationsites)
 
 #Using GGPLOT, plot the Base World Map
 mp <- NULL
-mapWorld <- borders("world", colour=wes_palette("Royal1")[1], fill=wes_palette("Royal1")[1]) # create a layer of borders
+mapWorld <- borders("usa", colour=wes_palette("Royal1")[1], fill=wes_palette("Royal1")[1]) # create a layer of borders
 mapplot <-  ggplot(sitegeocodes, aes(x=lon, y=lat)) +
   theme_classic() +
   theme(axis.title.x=element_blank(),
@@ -110,8 +110,8 @@ mapplot <-  ggplot(sitegeocodes, aes(x=lon, y=lat)) +
   geom_point(color=wes_palette("Royal1")[2], size=3)
 
 
-barside <- plot_grid(dnatype, circulartype, labels = c("D", "E"), ncol = 1)
-buildingalmost <- plot_grid(mapplot, countbar, barside, labels = c("B", "C"), ncol = 1, rel_heights=c(2,2,1))
+barside <- plot_grid(dnatype, circulartype, labels = c("C", "D"), ncol = 1)
+buildingalmost <- plot_grid(countbar, barside, labels = c("B"), ncol = 1, rel_heights=c(2,1))
 donebuild <- plot_grid(heatmap, buildingalmost, labels = c("A"), ncol = 2)
 donebuild
 
@@ -120,20 +120,10 @@ donebuild
 ###############
 
 width <- 16
-height <- 10
+height <- 8
 
 pdf(file="./figures/BenchmarkDataset.pdf",
 width=width,
 height=height)
-  a <- dev.cur()
-  png(file="./figures/BenchmarkDataset.png",
-  width=width,
-  height=height,
-  units="in",
-  res=800)
-    dev.control("enable")
-    # ggplot heatmap
-    donebuild
-    dev.copy(which=a)
-  dev.off()
+  donebuild
 dev.off()

### 2017-05-08 bin/VisualizeNetwork.R

@@ -14,6 +14,15 @@ library("ggraph")
 library("grid")
 library("stringr")
 
+suppressMessages(c(
+library("RNeo4j"),
+library("ggplot2"),
+library("C50"),
+library("caret"),
+library("wesanderson"),
+library("plotROC"),
+library("cowplot")
+))
 
 ###################
 # Set Subroutines #
@@ -56,10 +65,13 @@ plotnetwork <- function (nodeframe=nodeout, edgeframe=edgeout) {
   # Create the plot
   fres <- ggraph(ig, 'igraph', algorithm = 'bipartite') + 
         geom_edge_link0(edge_alpha = 0.0065) +
-        geom_node_point(aes(color = label), size = 1.5) +
+        geom_node_point(aes(color = label), size = 1.5, show.legend = FALSE) +
         ggforce::theme_no_axes() +
         scale_color_manual(values = wes_palette("Royal1")[c(2,4)]) +
-        coord_flip()
+        coord_flip() +
+        annotate("text", x = 475, y = 0.85, label = "Phage", size = 6, color = wes_palette("Royal1")[c(4)]) +
+        annotate("text", x = 25, y = 0.17, label = "Bacteria", size = 6, color = wes_palette("Royal1")[c(2)]) +
+        theme_graph(border = FALSE)
 
   return(fres)
 }
@@ -97,7 +109,7 @@ graphDiameter <- function (nodeframe=nodeout, edgeframe=edgeout) {
   edge <- ecount(ig)
   finaldf <- t(as.data.frame(c(connectionresult, vert, edge)))
   rownames(finaldf) <- NULL
-  colnames(finaldf) <- c("diameter", "vert", "edge")
+  colnames(finaldf) <- c("Diameter", "Vertices", "Edges")
 
   return(finaldf)
 }
@@ -149,10 +161,11 @@ write(connectionstrength(), stderr())
 totalstats <- as.data.frame(graphDiameter())
 totalstats$class <- "Total"
 
-length(grep("Phage", nodeout[,1]))
-length(grep("Bacteria", nodeout[,1]))
-
-length(edgeout[,1])
+# Collect some stats for the data table
+phagenodes <- length(grep("Phage", nodeout[,1]))
+bactnodes <- length(grep("Bacteria", nodeout[,1]))
+totaledges <- length(edgeout[,1])
[…]

### 2017-05-08 bin/alteredDiet.R

@@ -145,29 +145,33 @@ rcentraldf$di <- as.numeric(rcentraldf$di)
 binlength <- c(1:2) + 0.5
 didgbox <- ggplot(rcentraldf[c(rcentraldf$time %in% "TP10" | rcentraldf$time %in% "TP8"),], aes(x = patientdiet, y = dg)) +
 	theme_classic() +
-	geom_dotplot(fill=wes_palette("Royal1")[2], binaxis = "y", binwidth = 0.0025, stackdir = "center") +
+	geom_dotplot(fill=wes_palette("Royal1")[2], binaxis = "y", stackdir = "center", dotsize = 0.5, stackratio = 1) +
 	ylab("Degree Centrality") +
 	xlab("") +
 	theme(
 	    axis.line.x = element_line(colour = "black"),
 	    axis.line.y = element_line(colour = "black")
 	) +
-	stat_summary(fun.y = mean, fun.ymin = mean, fun.ymax = mean, geom = "crossbar", width = 0.5) +
-	geom_vline(xintercept=binlength,color="grey")
+	# stat_summary(fun.y = mean, fun.ymin = mean, fun.ymax = mean, geom = "crossbar", width = 0.5) +
+	# geom_vline(xintercept=binlength,color="grey") +
+	geom_vline(xintercept=binlength,color="grey") +
+	ylim(0, NA) +
+	scale_x_discrete(labels = c("High Fat", "Low Fat"))
 
 diclbox <- ggplot(rcentraldf[c(rcentraldf$time %in% "TP10" | rcentraldf$time %in% "TP8"),], aes(x = patientdiet, y = cl)) +
 	theme_classic() +
-	geom_dotplot(fill=wes_palette("Royal1")[2], binaxis = "y", binwidth = 0.0025, stackdir = "center") +
+	geom_dotplot(fill=wes_palette("Royal1")[2], binaxis = "y", stackdir = "center", dotsize = 0.5, stackratio = 1) +
 	ylab("Closeness Centrality") +
 	xlab("") +
 	theme(
 	    axis.line.x = element_line(colour = "black"),
 	    axis.line.y = element_line(colour = "black")
 	) +
-	stat_summary(fun.y = mean, fun.ymin = mean, fun.ymax = mean, geom = "crossbar", width = 0.5) +
-	geom_vline(xintercept=binlength,color="grey")
-
[…]

### 2017-05-08 bin/interpersonaldiversity.R

@@ -1,7 +1,7 @@
 ##################
 # Load Libraries #
 ##################
-gcinfo(TRUE)
+gcinfo(FALSE)
 packagelist <- c("RNeo4j", "ggplot2", "wesanderson", "igraph", "visNetwork", "scales", "plyr", "cowplot", "vegan", "reshape2", "stringr")
 new.packages <- packagelist[!(packagelist %in% installed.packages()[,"Package"])]
 if(length(new.packages)) install.packages(new.packages, repos='http://cran.us.r-project.org')
@@ -137,6 +137,8 @@ fx <- approxfun(pdf$x, pdf$y, yleft=0, yright=0)
 cdfdiet <- integrate(fx, -Inf, 0)
 cdfdiet
 
+pdietpvalue <- wilcox.test(data = ravg, avg ~ class, paired = TRUE)$p.value
+
 linediet <- ggplot(ravg, aes(x = class, y = avg, group = patient1)) +
 	theme_classic() +
 	theme(
@@ -146,7 +148,9 @@ linediet <- ggplot(ravg, aes(x = class, y = avg, group = patient1)) +
 	geom_line(colour = wes_palette("Royal1")[2]) +
 	geom_point(colour = "black") +
 	ylab("EV Centrality Distance") +
-	xlab("")
+	xlab("") +
+	ylim(0, NA) +
+	annotate("text", x = 1.5, y = 0, label = paste("p-value = ", signif(pdietpvalue, digits = 3), sep = ""))
 
 densitydiet <- ggplot(ravgslope, aes(y)) +
 	theme_classic() +
@@ -177,9 +181,10 @@ plotnmds_dietstudy <- ggplot(ORD_FIT, aes(x=MDS1, y=MDS2, colour=factor(V1))) +
 	  legend.position = "bottom"
 	) +
     geom_point() +
-    scale_colour_manual(values = wes_palette("Royal2"), name = "Subject")
+    scale_colour_manual(values = wes_palette("Darjeeling"), name = "Subject")
+plotnmds_dietstudy
 
-anosim(rdist, ORD_FIT$V1)
+anosimstat <- anosim(rdist, ORD_FIT$V1)
 
 ##############
 # Skin Graph #
@@ -211,6 +216,12 @@ for (i in skinsites) {
 rm(i)
 
 totalgraph <- rbind(graphdfTP2, graphdfTP3)
+
+# Correct the lengths
[…]

### 2017-05-08 bin/sequencedepth.R

@@ -1,0 +1,152 @@
+# sequencedepth.R
+# Geoffrey Hannigan
+# Schloss Lab
+# University of Michigan
+
+gcinfo(TRUE)
+packagelist <- c("RNeo4j", "ggplot2", "wesanderson", "igraph", "visNetwork", "scales", "plyr", "cowplot", "vegan", "reshape2", "stringr")
+new.packages <- packagelist[!(packagelist %in% installed.packages()[,"Package"])]
+if(length(new.packages)) install.packages(new.packages, repos='http://cran.us.r-project.org')
+lapply(packagelist, library, character.only = TRUE)
+
+graph <- startGraph("http://localhost:7474/db/data/", "neo4j", "root")
+
+### Diet ###
+
+# Get list of the sample IDs
+sampleidquery <- "
+MATCH
+	(x:SRP002424)-->(y)-[d]->(z:Phage)-->(a:Bacterial_Host)<-[e]-(b),
+	(b)<--(i:PatientID)-->(y),
+	(b)<--(t:TimePoint)-->(y),
+	(k:Disease)-->(y)
+WHERE toInt(d.Abundance) > 0
+OR toInt(e.Abundance) > 0
+RETURN DISTINCT
+	z.Name AS from,
+	a.Name AS to,
+	i.Name AS PatientID,
+	t.Name AS TimePoint,
+	k.Name AS Diet,
+	toInt(d.Abundance) AS PhageAbundance,
+	toInt(z.Length) AS PhageLength,
+	toInt(e.Abundance) AS BacteriaAbundance,
+	toInt(a.Length) AS BacteriaLength;
+"
+
+sampletable <- as.data.frame(cypher(graph, sampleidquery))
+
+# Correct the lengths
+sampletable$PhageAbundance <- round(1e7 * sampletable$PhageAbundance / sampletable$PhageLength)
+sampletable$BacteriaAbundance <- round(1e7 * sampletable$BacteriaAbundance / sampletable$BacteriaLength)
+sampletable <- sampletable[,-9]
+sampletablediet <- sampletable[,-7]
+sampletablediet$PatientID <- paste(sampletablediet$PatientID, sampletablediet$TimePoint, sep = "_")
+sampletablediet <- sampletablediet[,c("PatientID", "PhageAbundance", "BacteriaAbundance")]
+sampletablediet$class <- "Diet"
+
+### Skin ###
+
+# Import graphs into a list
[…]

### 2017-05-08 bin/ContigLengthTable.pl
Added my script for contig length.

@@ -1,0 +1,62 @@
+#!/usr/local/bin/perl -w
+# ContigLengthTable.pl
+# Geoffrey Hannigan
+# Pat Schloss Lab
+# University of Michigan
+
+# Set use
+use strict;
+use warnings;
+use Getopt::Long;
+use Pod::Usage;
+
+print STDERR "Creating sequence length table.\n";
+
+# Set Variables
+## Options
+my $opt_help;
+my $input;
+my $output;
+
+## Files
+my $IN;
+my $OUT;
+
+## Misc
+my $flag = 0;
+my $LineLength;
+my $lineName;
+
+# Set options
+GetOptions(
+	'h|help' => \$opt_help,
+	'i|input=s' => \$input,
+	'o|output=s' => \$output,
+);
+
+pod2usage(-verbose => 1) && exit if defined $opt_help;
+
+open($IN, "<", "$input") || die "Unable to read $input: $!";
+open($OUT, ">", "$output") || die "Unable to write to $output: $!";
+
+# Now run the filter
+while (my $line = <$IN>) {
+	chomp $line;
+	if ($line =~ /^\>/ && $flag==0) {
+		($lineName = $line) =~ s/^\>//;
+		$flag = 1;
+		next;
+	} elsif ($line !~ /^\>/ && $flag==1) {
+		$LineLength = length $line;
+		print $OUT "$lineName\t$LineLength\n";
+		$flag = 0;
+		$lineName = '';
+	} else {
+		die "Error in the fasta format: $!";
+	}
+}
+
+close($IN);
+close($OUT);
+
+print STDERR "Finished creating sequence length table.\n";

### 2017-05-08 bin/GetLongestContig.R
Removed depends that did not need to be there.

@@ -4,8 +4,6 @@
 # University of Michigan
 
 library("optparse")
-library("ggplot2")
-library("cowplot")
 library("plyr")
 
 option_list <- list(

### 2017-05-08 bin/IdentifyContigsBac.sh
Fixed depends.

@@ -1,0 +1,91 @@
+#! /bin/bash
+# IdentifyContigs.sh
+# Geoffrey Hannigan
+# Pat Schloss Lab
+# University of Michigan
+
+export fastafile=$1
+export referencefile=$2
+export idfile=$3
+export outputfile=$4
+export BlastPath=$5
+
+# Get the contig representative sequences
+mkdir -p ./data/tmpidbac
+
+RunBlast () {
+	# 1 = Query Seqs
+	# 2 = ReferenceSeqs
+
+	echo Making blast database...
+	${BlastPath}makeblastdb \
+		-dbtype nucl \
+		-in "${2}" \
+		-out ./data/tmpid/ReferenceGenomes
+
+	echo Running tblastx...
+	${BlastPath}tblastx \
+    	-query "${1}" \
+    	-out ./data/tmpid/blastout.tsv \
+    	-db ./data/tmpid/ReferenceGenomes \
+    	-evalue 1e-25 \
+    	-num_threads 8 \
+    	-max_target_seqs 1 \
+    	-outfmt 6
+}
+
+export -f RunBlast
+
+cut -f 1 ${idfile} | tail -n +2 > ./data/tmpid/contiglist.tsv
+grep -A 1 -f ./data/tmpid/contiglist.tsv ${fastafile} \
+	| egrep -v "\-\-" \
+	> ./data/tmpid/contigrepset.fa
+
+# RunBlast ./data/tmpid/contigrepset.fa ${referencefile}
+
+# Add cluster ID to the table
+cut -f 1,2 ./data/tmpid/blastout.tsv | sort | uniq > ./data/tmpid/cutblastout.tsv
+grep --file=./data/tmpid/contiglist.tsv ./data/ContigClustersBacteria/clustering_gt2000.csv \
+	| sed 's/\,/\t/' \
+	> ./data/tmpid/clusterids.tsv
+awk -F "\t" 'FNR==NR { a[$1] = $2; next } { for( i in a ) if($1 ~ i) {print a[$1]"\t"$2} }' \
+	./data/tmpid/clusterids.tsv \
+	./data/tmpid/cutblastout.tsv \
+	| sed 's/ENA|\(.\+\)|.*/\1/' \
+	> ./data/tmpid/clustform.tsv
+
+# Make list of IDs
[…]

### 2017-05-08 bin/IdentifyContigsBac.sh
Fixed path.

@@ -11,7 +11,7 @@ export outputfile=$4
 export BlastPath=$5
 
 # Get the contig representative sequences
-mkdir -p ./data/tmpidbac
+mkdir -p ./data/tmpid
 
 RunBlast () {
 	# 1 = Query Seqs

### 2017-05-08 bin/IdentifyContigsBac.sh
Added blast to the mix.

@@ -41,7 +41,7 @@ grep -A 1 -f ./data/tmpid/contiglist.tsv ${fastafile} \
 	| egrep -v "\-\-" \
 	> ./data/tmpid/contigrepset.fa
 
-# RunBlast ./data/tmpid/contigrepset.fa ${referencefile}
+RunBlast ./data/tmpid/contigrepset.fa ${referencefile}
 
 # Add cluster ID to the table
 cut -f 1,2 ./data/tmpid/blastout.tsv | sort | uniq > ./data/tmpid/cutblastout.tsv

### 2017-05-08 doc/manuscript.Rmd
Updates.

@@
 Such approaches are not able to capture how these communities interact to maintain stability/diversity, co-evolve, and share genetic information (e.g.
-horizontal gene transfer <!--X as this isn't the only way to share information, use eg instaed of ie.
-Also, is there work that demonstrates how these communities interact to maintain stability?
-By stability, do we mean maintain diversity?
-What are the other outcomes of interactions?
-e.g., (antagonistic) co-evolution, etc.
-I know these aren't elements to fit into an Abstract, so staying tuned as to whether they appear later, as well-->).
+horizontal gene transfer).
 We developed and implemented a network-based analytical approach to describe phage-bacteria network diversity throughout the human body.
@@
 
-We investigated human bacterial-phage networks by utilizing three published microbiome datasets which contained paired virus and total metagenomic sequence sets, which primarily consisted of bacteria [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei].
+We investigated human bacterial-phage networks by utilizing three published microbiome datasets that contained paired virus and total metagenomic sequence sets, which primarily consisted of bacteria [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei].
 Our approach built off of previous large-scale phage-bacteria microbiome network analyses by inferring interactions using metagenomic datasets, instead of using culture-based techniques [@Flores:2013hc].
@@
 They employed combinations of filtration, chloroform/DNase treatment, and Cesium Chloride treatments to eliminate organismal DNA and thereby allow for direct assessment of both the extracellular and fully-assembled intracellular virome **(Supplemental Figure \ref{SequenceStats} A-B)** [@Minot:2011ez, @Hannigan:2015fz, @Reyes:2010cwa; @Turnbaugh:2009ei].
-While the whole metagenomic shotgun sequence samples were not subjected to purification, they primarily consisted of bacteria and other microbiome members (e.g.
-fungi, archaea, *Demodex* mites, etc) were not considered in this initial study.
+While the whole metagenomic shotgun sequence samples were not subjected to purification, they primarily consisted of bacteria and had minimal representation from other groups [@Minot:2011ez, @Hannigan:2015fz, @Reyes:2010cwa; @Turnbaugh:2009ei].
+Other, minimally represented microbiome members (e.g.
+fungi, archaea, eukaryotes including *Demodex* mites, etc) were not considered in this study.
 
@@
 We generated clusters of related bacterial and phage contigs based on their k-mer frequencies and co-abundance patterns using the CONCOCT algorithm **(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})**.
-Contig lengths within clusters ranged between $10_{3}$ and $10_{5.5}$, confirming the majority of contigs represented genome fragments, rather than complete genomes  **(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})**.
-We also confirmed that the majority of bacterial clusters were indeed identified as bacteria (CITE FIGURE HERE).
+Contig lengths within clusters ranged between $10^{3}$ and $10^{5.5}$, confirming the majority of contigs represented genome fragments, rather than complete genomes  **(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})**.
+We also confirmed that the majority of bacterial clusters were indeed identified as bacteria (CITE FIGURE HERE), confirming the previous reports [@Minot:2011ez, @Hannigan:2015fz, @Turnbaugh:2009ei].
 We interpreted these clusters as operational units of related bacteria and phage genomes, which we defined as operational genomic units (OGUs).
@@
 We predicted which phage OGUs infected which bacterial OGUs using a random forest model trained on experimentally validated infectious relationships from six previous publications [@Jensen:1998vh;@Malki:2015tm;@Schwarzer:2012ez;@Kim:2012dh;@Matsuzaki:1992gw;@Edwards:2015iz].
-This training set contained 43 diverse bacterial species and 30 diverse phage strains, with both broad and specific infectious ranges **(Supplemental Figure \ref{ValidationOverview} A - B)**.
+Only bacteria and phages were used in the model.
+The training set contained 43 diverse bacterial species and 30 diverse phage strains, with both broad and specific infectious ranges **(Supplemental Figure \ref{ValidationOverview} A - B)**.
 Phages with linear and circular genomes, as well as ssDNA and dsDNA genomes, were included in the analysis.
@@
 # Discussion
-<!-- Much of the discussion repeats the statements in the results, it felt a bit redundant.
-At times, the discussion even repeated itself (you'll see I deleted most of the last paragraph).
-There are probably ways to be less repetitive and more provocative or to bring in discussion of more recent work?
-Or to reduce it even more by removing redundancies.
-Alternatively, could remove the redundancies in the results...as in only state the results, and not reference previous work.
-Then could discuss how findings relate to previous work only in the discussion, rather than the results...which does happen at times in current form.-->
 Foundational microbiome work has provided a baseline understanding of the human microbiome by characterizing bacteria and viral diversity across the human body, as well as other environments [@Grice:2009eea; @Findley:2013jf; @Hannigan:2015fz; @Costello:2009im, @Consortium:2012iz; @Schloss:2005hz; @Minot:2011ez].
@@
 Just as gut microbiome and virome composition and diversity are conserved in individuals [@Hannigan:2015fz; @Grice:2009eea; @Findley:2013jf; @Minot:2013ih], gut and skin microbiome network structures were conserved within individuals over time.
-Yet, gut network structure was not conserved among family members.
-These findings suggested that microbiome network properties, including stability and the potential for horizontal gene transfer <!-- and lots of other things related to host and virus eco and evo processes -->, were personal and may be impacted by personal factors ranging from the body's immune system to environmental conditions, such as climate and diet <!-- though I'd argue that the immune system is a component/control on the environmental condition of the microbial communities -->.
-The ability of environmental conditions to shape gut and skin microbiome network structure was supported by our finding that diet and skin location were associated with altered network structures.
+Gut network structure was not conserved among family members.
+These findings suggested that microbiome network properties, including stability, the potential for horizontal gene transfer, co-evolution, etc, were personal and may be impacted by personal factors ranging from the body's immune system to external environmental conditions, such as climate and diet.
+The ability of environmental conditions to shape gut and skin microbiome network structure was further supported by our finding that diet and skin location were associated with altered network structures.
 
 We found evidence that diet was sufficient to alter gut microbiome network connectivity.
-Although our sample size was small, our findings suggested that high-fat diets were less connected than low-fat diets.
-This suggested that high-fat diets may lead to less stable communities with a decreased ability for microbes to directly influence one other.
-We supported this finding with the observation that obesity was associated with decreased network connectivity (closeness of bacterial and phage vertices).
+Although our sample size was small, our findings provided evidence that high-fat diets were less connected than low-fat diets.
+This suggested that high-fat diets may lead to less stable communities with a decreased ability for microbes to directly influence one another.
+We supported this finding with the observation that obesity may have been associated with decreased network connectivity.
 Together these findings suggest the food we eat may not only impact which microbes colonize our guts, but may also impact their interactions.
-Further work will be required to characterize these relationships with a larger cohort.
+Further work will certainly be required to characterize these relationships with a larger cohort.
 
@@
 Network structure differed between environmentally exposed and occluded skin sites.
-The sites under greater environmental fluctuation and exposure, the exposed and intermittently exposed sites, were less connected and therefore were predicted to have a higher propensity for instability.
+The sites under greater environmental fluctuation and exposure (the exposed and intermittently exposed sites) were less connected and therefore were predicted to have a higher propensity for instability.
 Likewise, intermittently moist sites demonstrated less connectedness than the more stable moist and sebaceous sites.
-Together these data suggested body sites under greater degrees of fluctuation harbored less connected, potentially less stable microbiomes.
+Together these data suggested that body sites under greater degrees of fluctuation harbored less connected, potentially less stable microbiomes.
 
@@
 Large-scale experimental screens for phage and bacteria infectious interactions that report high-confidence negative interactions (i.e., no infection) will provide more robust model training and improved model performance.
-Just as we have improved on previous modeling efforts, we expect that new and creative scoring metrics will be integrated into this model to further improve performance.
+Furthermore, just as we have improved on previous modeling efforts, we expect that new and creative scoring metrics will be integrated into this model to improve performance.
+
 Second, although our analyses offer an informative proof of concept, this work was done retrospectively and relied on published research from as long as seven years ago.
@@
 
-Our work takes a first step towards defining bacteria-virus interaction profiles as a characteristic of human-associated microbial communities.
-We provide a foundation for future studies to begin evaluating how microbiome network dynamics change in disease states, and how the information can be applied therapeutically.
+It is important to note that the microbiome consists of bacteria and viruses, as well as fungi, archaea, eukaryotes such as *Demodex* mites, etc.
+These microbiome components also interact with human immune cells and other non-microbial community members.
+Our study only represents an initial step toward network analysis, and future work will include more components of the microbiome to gain a more complete understanding of the community through its interactions.
+In addition to these microbiome components, additional relationship types will also be added so as to provide a more robust network for analysis.
+
+Together, our work takes a first step towards defining bacteria-virus interaction profiles as a characteristic of human-associated microbial communities.
+By focusing on relationships between bacterial and viral communities, we can begin studying them as interacting cohorts they are, instead of isolated entities.
+This work will also provide a foundation for future studies to begin evaluating how microbiome network dynamics change in disease states, and how the information can be applied therapeutically.
 

### 2017-05-09 bin/IdentifyContigsBac.sh
Fixed to blastn over tblastx, which made no sense.

@@ -23,12 +23,12 @@ RunBlast () {
 		-in "${2}" \
 		-out ./data/tmpid/ReferenceGenomes
 
-	echo Running tblastx...
-	${BlastPath}tblastx \
+	echo Running blastn...
+	${BlastPath}blastn \
     	-query "${1}" \
     	-out ./data/tmpid/blastout.tsv \
     	-db ./data/tmpid/ReferenceGenomes \
-    	-evalue 1e-25 \
+    	-evalue 1e-10 \
     	-num_threads 8 \
     	-max_target_seqs 1 \
     	-outfmt 6

### 2017-05-09 bin/IdentifyContigsBac.sh
Just try running blast for a minute.

@@ -17,11 +17,11 @@ RunBlast () {
 	# 1 = Query Seqs
 	# 2 = ReferenceSeqs
 
-	echo Making blast database...
-	${BlastPath}makeblastdb \
-		-dbtype nucl \
-		-in "${2}" \
-		-out ./data/tmpid/ReferenceGenomes
+	# echo Making blast database...
+	# ${BlastPath}makeblastdb \
+	# 	-dbtype nucl \
+	# 	-in "${2}" \
+	# 	-out ./data/tmpid/ReferenceGenomes
 
 	echo Running blastn...
 	${BlastPath}blastn \
@@ -43,49 +43,49 @@ grep -A 1 -f ./data/tmpid/contiglist.tsv ${fastafile} \
 
 RunBlast ./data/tmpid/contigrepset.fa ${referencefile}
 
-# Add cluster ID to the table
-cut -f 1,2 ./data/tmpid/blastout.tsv | sort | uniq > ./data/tmpid/cutblastout.tsv
-grep --file=./data/tmpid/contiglist.tsv ./data/ContigClustersBacteria/clustering_gt2000.csv \
-	| sed 's/\,/\t/' \
-	> ./data/tmpid/clusterids.tsv
-awk -F "\t" 'FNR==NR { a[$1] = $2; next } { for( i in a ) if($1 ~ i) {print a[$1]"\t"$2} }' \
-	./data/tmpid/clusterids.tsv \
-	./data/tmpid/cutblastout.tsv \
-	| sed 's/ENA|\(.\+\)|.*/\1/' \
-	> ./data/tmpid/clustform.tsv
+# # Add cluster ID to the table
+# cut -f 1,2 ./data/tmpid/blastout.tsv | sort | uniq > ./data/tmpid/cutblastout.tsv
+# grep --file=./data/tmpid/contiglist.tsv ./data/ContigClustersBacteria/clustering_gt2000.csv \
+# 	| sed 's/\,/\t/' \
+# 	> ./data/tmpid/clusterids.tsv
+# awk -F "\t" 'FNR==NR { a[$1] = $2; next } { for( i in a ) if($1 ~ i) {print a[$1]"\t"$2} }' \
+# 	./data/tmpid/clusterids.tsv \
+# 	./data/tmpid/cutblastout.tsv \
+# 	| sed 's/ENA|\(.\+\)|.*/\1/' \
+# 	> ./data/tmpid/clustform.tsv
[…]

### 2017-05-09 bin/IdentifyContigsBac.sh
Curious to change threshold.

@@ -28,7 +28,7 @@ RunBlast () {
     	-query "${1}" \
     	-out ./data/tmpid/blastout.tsv \
     	-db ./data/tmpid/ReferenceGenomes \
-    	-evalue 1e-10 \
+    	-evalue 1e-5 \
     	-num_threads 8 \
     	-max_target_seqs 1 \
     	-outfmt 6

### 2017-05-09 doc/manuscript.Rmd
Updated manuscript.

@@
 
-**Word Count**: 237 / 250
+**Word Count**: 242 / 250
 
@@
 Bacterial communities have been associated with diseases including a wide range of skin conditions [@Hannigan:2013im], acute and chronic wound healing conditions [@Hannigan:2014be; @Loesche:2016ev], and gastrointestinal diseases including inflammatory bowel disease [@He:2016ch; @Norman:2015kb], *Clostridium difficile* infections [@Seekatz:2016fz], and colorectal cancer [@Zackular:2014fba; @Baxter:2014hb].
-Altered viromes (virus communities consisting primarily of bacteriophages) have also been associated with various diseases and perturbations, including inflammatory bowel disease [@Norman:2015kb; @Manrique:2016dx], periodontal disease [@Ly:2014ew], spread of antibiotic resistance [@Modi:2013fia], and others [@Monaco:2016ita; @Hannigan:2015fz; @Minot:2011ez; @SantiagoRodriguez:2015gd; @Abeles:2015dy; @Abeles:2014kj].
+Altered viromes (virus communities consisting primarily of bacteriophages) have also been associated with various diseases and perturbations including inflammatory bowel disease [@Norman:2015kb; @Manrique:2016dx], periodontal disease [@Ly:2014ew], spread of antibiotic resistance [@Modi:2013fia], and others [@Monaco:2016ita; @Hannigan:2015fz; @Minot:2011ez; @SantiagoRodriguez:2015gd; @Abeles:2015dy; @Abeles:2014kj].
 Viruses interact with bacteria, which together comprise a major portion of total microbial communities [@Haerter:2014ii].
@@
 Through this approach, we were able to provide a basic understanding of the network dynamics associated with healthy human phage and bacterial communities.
-By building and utilizing a microbiome-virome network, we found that different people, body sites, and anatomical locations not only support distinct microbiome membership and diversity [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei; @Grice:2009eea; @Findley:2013jf; @Costello:2009im, @Consortium:2012iz], but also support communities with distinct communication structures and propensities toward community instability.
+By building and utilizing a microbiome network, we found that different people, body sites, and anatomical locations not only support distinct microbiome membership and diversity [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei; @Grice:2009eea; @Findley:2013jf; @Costello:2009im, @Consortium:2012iz], but also support communities with distinct communication structures and propensities toward community instability.
 Through an improved understanding of the healthy state of network structures across the human body, we empower future studies to begin investigating how these community structures are influenced by disease state and their overall impact on human health.
@@
 These datasets were selected because they included virome samples subjected to virus-like particle (VLP) purification.
-They employed combinations of filtration, chloroform/DNase treatment, and Cesium Chloride treatments to eliminate organismal DNA and thereby allow for direct assessment of both the extracellular and fully-assembled intracellular virome **(Supplemental Figure \ref{SequenceStats} A-B)** [@Minot:2011ez, @Hannigan:2015fz, @Reyes:2010cwa; @Turnbaugh:2009ei].
-While the whole metagenomic shotgun sequence samples were not subjected to purification, they primarily consisted of bacteria and had minimal representation from other groups [@Minot:2011ez, @Hannigan:2015fz, @Reyes:2010cwa; @Turnbaugh:2009ei].
-Other, minimally represented microbiome members (e.g.
-fungi, archaea, eukaryotes including *Demodex* mites, etc) were not considered in this study.
+To this end, they employed combinations of filtration, chloroform/DNase treatment, and cesium chloride gradients to eliminate organismal DNA and thereby allow for direct assessment of both the extracellular and fully-assembled intracellular virome **(Supplemental Figure \ref{SequenceStats} A-B)** [@Minot:2011ez, @Hannigan:2015fz, @Reyes:2010cwa; @Turnbaugh:2009ei].
+While the whole metagenomic shotgun sequence samples were not subjected to purification, they primarily consisted of bacteria [@Minot:2011ez, @Hannigan:2015fz, @Reyes:2010cwa; @Turnbaugh:2009ei].
 
-Here, the bacterial and viral sequences from these studies were quality filtered and assembled into contigs.
+The bacterial and viral sequences from these studies were quality filtered and assembled into contigs.
 We generated clusters of related bacterial and phage contigs based on their k-mer frequencies and co-abundance patterns using the CONCOCT algorithm **(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})**.
 Contig lengths within clusters ranged between $10^{3}$ and $10^{5.5}$, confirming the majority of contigs represented genome fragments, rather than complete genomes  **(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})**.
-We also confirmed that the majority of bacterial clusters were indeed identified as bacteria (CITE FIGURE HERE), confirming the previous reports [@Minot:2011ez, @Hannigan:2015fz, @Turnbaugh:2009ei].
+We were also able to identify the majority (79%) of bacteria clusters, agreeing with previous reports [@Minot:2011ez, @Hannigan:2015fz, @Turnbaugh:2009ei].
 We interpreted these clusters as operational units of related bacteria and phage genomes, which we defined as operational genomic units (OGUs).
@@
 
-In addition to diet, we also observed an association between obesity and network structure **(Figure \ref{dietnetworks} C-D)**.
+In addition to diet, we also observed a potential association between obesity and network structure **(Figure \ref{dietnetworks} C-D)**.
 The obesity-associated network demonstrated a higher degree centrality **(Figure \ref{dietnetworks} C)**, but less closeness centrality compared to the healthy controls **(Figure \ref{dietnetworks} D)**.
-These results suggest that the obesity network was overall less connected, having microbes further from all other microbes within the community.
+These results suggest that the obesity network may have been overall less connected, having microbes further from all other microbes within the community.
 
@@
 
-Together, our work takes a first step towards defining bacteria-virus interaction profiles as a characteristic of human-associated microbial communities.
+Together our work takes an initial step towards defining bacteria-virus interaction profiles as a characteristic of human-associated microbial communities.
 By focusing on relationships between bacterial and viral communities, we can begin studying them as interacting cohorts they are, instead of isolated entities.
@@
 Where `L` is the length of each contig `j` within the OGU.
+
+## Bacterial OGU Identification
+The longest contig from each OGU was used as its representative sequence for identification.
+Represetnative sequences were aligned to the European Nucleotide Archive (ENA) bacterial reference database using the blastn algorithm (e-value < $10^{-5}$).
 
@@
 
-Intuitively, the **degree centrality** of node $V_i$ is defined as the number of edges that are adjacent to that node:
+Intuitively, the **degree centrality** of node $V_i$ is defined as the number of edges that are incident to that node:
 
@@
 
-$$ C\left( G \right) =\frac { \sum _{ i=1 }^{ n }{ Cw -c\left( { v }_{ i } \right)  }  }{ { T } }  $$
+$$ C\left( G \right) =\frac { \sum _{ i=1 }^{ n }{ Cw -c\left( { V }_{ i } \right)  }  }{ { T } }  $$
 

### 2017-05-16 bin/IdentifyContigs.sh
Added contig ID script.

@@ -17,18 +17,18 @@ RunBlast () {
 	# 1 = Query Seqs
 	# 2 = ReferenceSeqs
 
-	# echo Making blast database...
-	# ${BlastPath}makeblastdb \
-	# 	-dbtype nucl \
-	# 	-in "${2}" \
-	# 	-out ./data/tmpid/ReferenceGenomes
+	echo Making blast database...
+	${BlastPath}makeblastdb \
+		-dbtype nucl \
+		-in "${2}" \
+		-out ./data/tmpid/ReferenceGenomes
 
-	echo Running blastn...
-	${BlastPath}blastn \
+	echo Running tblastx...
+	${BlastPath}tblastx \
     	-query "${1}" \
     	-out ./data/tmpid/blastout.tsv \
     	-db ./data/tmpid/ReferenceGenomes \
-    	-evalue 1e-5 \
+    	-evalue 1e-25 \
     	-num_threads 8 \
     	-max_target_seqs 1 \
     	-outfmt 6
@@ -43,21 +43,19 @@ grep -A 1 -f ./data/tmpid/contiglist.tsv ${fastafile} \
 
 RunBlast ./data/tmpid/contigrepset.fa ${referencefile}
 
-# # Add cluster ID to the table
-# cut -f 1,2 ./data/tmpid/blastout.tsv | sort | uniq > ./data/tmpid/cutblastout.tsv
-# grep --file=./data/tmpid/contiglist.tsv ./data/ContigClustersBacteria/clustering_gt2000.csv \
-# 	| sed 's/\,/\t/' \
-# 	> ./data/tmpid/clusterids.tsv
-# awk -F "\t" 'FNR==NR { a[$1] = $2; next } { for( i in a ) if($1 ~ i) {print a[$1]"\t"$2} }' \
-# 	./data/tmpid/clusterids.tsv \
-# 	./data/tmpid/cutblastout.tsv \
-# 	| sed 's/ENA|\(.\+\)|.*/\1/' \
-# 	> ./data/tmpid/clustform.tsv
+# Add cluster ID to the table
+cut -f 1,2 ./data/tmpid/blastout.tsv | sort | uniq > ./data/tmpid/cutblastout.tsv
+sed 's/\,/\t/' ./data/ContigClustersVirus/clustering_gt1000.csv > ./data/tmpid/clusterids.tsv
[…]

### 2017-05-16 bin/IdentifyContigs.sh
Fixed to blastn

@@ -23,8 +23,8 @@ RunBlast () {
 		-in "${2}" \
 		-out ./data/tmpid/ReferenceGenomes
 
-	echo Running tblastx...
-	${BlastPath}tblastx \
+	echo Running blastn...
+	${BlastPath}blastn \
     	-query "${1}" \
     	-out ./data/tmpid/blastout.tsv \
     	-db ./data/tmpid/ReferenceGenomes \

### 2017-05-16 doc/manuscript.Rmd
Updated with edits.

@@
 Viruses and bacteria are critical components of the human microbiome and play important roles in health and disease.
-Most previous work has relied on studying bacteria and phages in isolation, reducing them to two separate communities.
-Such approaches are not able to capture how these communities interact to maintain stability/diversity, co-evolve, and share genetic information (e.g.
-horizontal gene transfer).
+Most previous work has relied on studying microbes and viruses in isolation, thereby reducing them to two separate communities.
+Such approaches are unable to capture how these microbial communities interact, including processes such as the mechanisms that maintain community stability or allow phage-host populations to co-evolve.
 We developed and implemented a network-based analytical approach to describe phage-bacteria network diversity throughout the human body.
@@
 This study quantified and contrasted the diversity of virome-microbiome networks across the human body and illustrated how environmental factors may influence phage-bacteria interactive dynamics.
-This work provides a baseline for future studies to investigate the role of virus-microbe ecological networks in advancing our understanding of conditions, such as disease states.
+This work provides a baseline for future studies to better understand system perturbations, such as disease states, through ecological networks.
 
-**Word Count**: 242 / 250
+**Word Count**: 243 / 250
 
@@
 Viruses and bacteria are critical components to the human microbiome and play important roles in health and disease.
-Bacterial communities have been associated with diseases including a wide range of skin conditions [@Hannigan:2013im], acute and chronic wound healing conditions [@Hannigan:2014be; @Loesche:2016ev], and gastrointestinal diseases including inflammatory bowel disease [@He:2016ch; @Norman:2015kb], *Clostridium difficile* infections [@Seekatz:2016fz], and colorectal cancer [@Zackular:2014fba; @Baxter:2014hb].
+Bacterial communities have been associated with diseases including a range of skin conditions [@Hannigan:2013im], acute and chronic wound healing conditions [@Hannigan:2014be; @Loesche:2016ev], and gastrointestinal diseases including inflammatory bowel disease [@He:2016ch; @Norman:2015kb], *Clostridium difficile* infections [@Seekatz:2016fz], and colorectal cancer [@Zackular:2014fba; @Baxter:2014hb].
 Altered viromes (virus communities consisting primarily of bacteriophages) have also been associated with various diseases and perturbations including inflammatory bowel disease [@Norman:2015kb; @Manrique:2016dx], periodontal disease [@Ly:2014ew], spread of antibiotic resistance [@Modi:2013fia], and others [@Monaco:2016ita; @Hannigan:2015fz; @Minot:2011ez; @SantiagoRodriguez:2015gd; @Abeles:2015dy; @Abeles:2014kj].
-Viruses interact with bacteria, which together comprise a major portion of total microbial communities [@Haerter:2014ii].
-Viruses influence the living microbes of these communities by, for example, lysing their microbial hosts, modulating their functionality through altered host gene expression [@Lindell:2005gz, @Tyler:2013fl, @Hargreaves:2014ja], influencing evolutionary processes through, for example, horizontal gene transfer [@Moon:2015fa, @Modi:2013fi, @Ogg:1981th, @Frost:2005dn] or co-evolution [@Koskella:2014ds], and by altering ecosystem processes and elemental stoichiometry [@Jover:2014gq].
+The viruses act in concert with their microbial hosts as a single ecological community [@Haerter:2014ii].
+Viruses influence their living microbial host communities through processes including lysing their hosts, modulating host gene expression [@Lindell:2005gz, @Tyler:2013fl, @Hargreaves:2014ja], influencing evolutionary processes, such as horizontal gene transfer [@Moon:2015fa, @Modi:2013fi, @Ogg:1981th, @Frost:2005dn] or antagonistic co-evolution [@Koskella:2014ds], and by altering ecosystem processes and elemental stoichiometry [@Jover:2014gq].
 
@@
 We leveraged machine learning and graph theory techniques to characterize human bacterial and phage communities by their relationships.
-We focused on characterizing human skin and gut virus-bacteria network diversity to provide a foundation for further studies of disease network dynamics and to gain broader insights into human microbiome network diversity across different body sites.
+We characterized human skin and gut virus-bacteria network diversity to provide a foundation for further studies of disease network dynamics and to gain broader insights into human microbiome network diversity across different body sites.
 
@@
 We studied the differences in virus-bacteria interaction networks across healthy human bodies by leveraging previously published sequence sets containing purified virome samples paired with bacterial metagenomes from whole metagenomic shotgun sequences.
-Our study contained three datasets, which explored the impact of diet on the healthy human gut virome [@Minot:2011ez], the impact of anatomical location on the healthy human skin virome [@Hannigan:2015fz], and the viromes of monozygotic twins and their mothers [@Reyes:2010cwa; @Turnbaugh:2009ei].
+Our study contained three datasets which explored the impact of diet on the healthy human gut virome [@Minot:2011ez], the impact of anatomical location on the healthy human skin virome [@Hannigan:2015fz], and the viromes of monozygotic twins and their mothers [@Reyes:2010cwa; @Turnbaugh:2009ei].
+The twin and diet studies utilized multiple displacement amplification methods in their library preparations.
 These datasets were selected because they included virome samples subjected to virus-like particle (VLP) purification.
@@
 The bacterial and viral sequences from these studies were quality filtered and assembled into contigs.
-We generated clusters of related bacterial and phage contigs based on their k-mer frequencies and co-abundance patterns using the CONCOCT algorithm **(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})**.
-Contig lengths within clusters ranged between $10^{3}$ and $10^{5.5}$, confirming the majority of contigs represented genome fragments, rather than complete genomes  **(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})**.
-We were also able to identify the majority (79%) of bacteria clusters, agreeing with previous reports [@Minot:2011ez, @Hannigan:2015fz, @Turnbaugh:2009ei].
-We interpreted these clusters as operational units of related bacteria and phage genomes, which we defined as operational genomic units (OGUs).
-These OGUs are conceptually similar to the operational taxonomic unit (OTU) and operational protein family (OPF) definitions used for grouping highly similar 16S rRNA gene and open reading frame sequences, respectively [@Schloss:2008ht].
+We further grouped the related bacterial and phage contigs into operationally defined units which were based on their k-mer frequencies and co-abundance patterns, similar to previous reports **(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})** [@Roux:2016cc].
+We referred to these operationally defined groups of related contigs as operational genomic units (OGUs).
+Each OGU represented either a genomically similar sub-population of bacteria or phages.
+Contig lengths within clusters ranged between $10^{3}$ and $10^{5.5}$ bp **(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})**.
 
-## Establishing a Model of Phage-Bacteria Interactions
+## Evaluating the Model to Predict Phage-Bacteria Interactions
 We predicted which phage OGUs infected which bacterial OGUs using a random forest model trained on experimentally validated infectious relationships from six previous publications [@Jensen:1998vh;@Malki:2015tm;@Schwarzer:2012ez;@Kim:2012dh;@Matsuzaki:1992gw;@Edwards:2015iz].
@@
 This training set included both positive relationships (a phage infects a bacterium) and negative relationships (a phage does not infect a bacterium).
-This allowed us to validate the false positive and false negative rates associated with our candidate models, thereby building upon previous work that only considered positive relationships [@Edwards:2015iz].
+This allowed us to validate the false positive and false negative rates associated with our candidate models, thereby building upon much previous work that only considered positive relationships [@Edwards:2015iz].
 
@@
 The resulting random forest model performed with an AUC of `r signif(genmodelper$ROC, digits = sigfig)`, a sensitivity of `r signif(genmodelper$Sens, digits = sigfig)`, and a specificity of `r signif(genmodelper$Spec, digits = sigfig)` **(Figure \ref{RocCurve} A)**.
-The most important predictor in the model was amino acid similarity between genes (blastx), followed by nucleotide similarity of whole genomes **(Figure \ref{RocCurve} B)**.
+The most important predictor in the model was amino acid similarity between genes, followed by nucleotide similarity of whole genomes **(Figure \ref{RocCurve} B)**.
 Protein family interactions were moderately important to the model, and CRISPRs were largely uninformative, due to the lack of identifiable CRISPRs in the dataset and their redundancy with the nucleotide similarity methods **(Figure \ref{RocCurve} B)**.
@@
 
-Once our random forest model was built, we used it to classify the relationships between bacteria and phage operational genomic units in the experimental datasets.
-The relationships predicted within the three published datasets were used to construct one master network.
+We used the random forest model to classify the relationships between bacteria and phage operational genomic units, which were then used to build the interactive network.
 The master network contained the three studies as sub-networks, which themselves each contained sub-networks for each sample **(Figure \ref{RocCurve} D)**.
@@
 Like the master network, the skin network exhibited a diameter of `r prettyNum(genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Diameter"], big.mark = ",")` (measure of graph size; the greatest number of traversed vertices required between two vertices) and included `r round(100 * genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Vertices"] / genfigurestats[c(genfigurestats$class %in% "Total"),"Vertices"], digits = 1)`% and `r round(100 * genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Edges"] / genfigurestats[c(genfigurestats$class %in% "Total"),"Edges"], digits = 1)`% of the master network nodes and edges, respectively **(Figure \ref{RocCurve} E - F)**.
-The phages and bacteria in the gut diet and twin sample sets were more sparsely related, with each containing fewer than 150 vertices, fewer than 20,000 relationships, and diameters of `r prettyNum(genfigurestats[c(genfigurestats$class %in% "DietStudy"),"Diameter"], big.mark = ",")`, which indicated more sparsely related phages and bacteria **(Figure \ref{RocCurve} E - F)**.
+The phages and bacteria in the gut diet and twin sample sets were more sparsely related: each contained fewer than 150 vertices, fewer than 20,000 relationships, and diameters of `r prettyNum(genfigurestats[c(genfigurestats$class %in% "DietStudy"),"Diameter"], big.mark = ",")` **(Figure \ref{RocCurve} E - F)**.
 
@@
 Previous work in isolated culture-based systems has suggested that changes in nutrient availability are associated with altered phage-bacteria network structures [@Poisot:2011jc], although this has yet to be tested in humans.
-We hypothesized that a change in diet would also be associated with a change in virome-microbiome network structure in the human gut.
+We therefore hypothesized that a change in diet would also be associated with a change in virome-microbiome network structure in the human gut.
 
@@
 A network with higher average closeness centrality also indicates an overall greater degree of connections, which suggests a greater resilience against instability.
-Because these values are assigned to each phage and bacterium within each sample network, we calculated the average connectedness per sample and corrected for the maximum potential degree of connectedness to obtain a single value for sample graph connectedness.
+We used this information to calculate the average connectedness per sample, which was corrected for the maximum potential degree of connectedness.
 
@@
 
-In addition to diet, we also observed a potential association between obesity and network structure **(Figure \ref{dietnetworks} C-D)**.
-The obesity-associated network demonstrated a higher degree centrality **(Figure \ref{dietnetworks} C)**, but less closeness centrality compared to the healthy controls **(Figure \ref{dietnetworks} D)**.
-These results suggest that the obesity network may have been overall less connected, having microbes further from all other microbes within the community.
+In addition to diet, obesity was found to influence network structure.
+Obesity-associated networks demonstrated a higher degree centrality **(Figure \ref{dietnetworks} C)**, but less closeness centrality than the healthy-associated networks **(Figure \ref{dietnetworks} D)**.
+These results suggested that the obesity-associated networks are less connected, having microbes further from all other microbes within the community.
 
@@
 We found evidence that diet was sufficient to alter gut microbiome network connectivity.
-Although our sample size was small, our findings provided evidence that high-fat diets were less connected than low-fat diets.
-This suggested that high-fat diets may lead to less stable communities with a decreased ability for microbes to directly influence one another.
+Although our sample size was small, our findings provided evidence that high-fat diets were less connected than low-fat diets, and high-fat diets may therefore lead to less stable communities with a decreased ability for microbes to directly influence one another.
 We supported this finding with the observation that obesity may have been associated with decreased network connectivity.
-Together these findings suggest the food we eat may not only impact which microbes colonize our guts, but may also impact their interactions.
+Together these findings suggest the food we eat may not only impact which microbes colonize our guts, but may also impact their infectious interactions.
 Further work will certainly be required to characterize these relationships with a larger cohort.
@@
 Together these data suggested that body sites under greater degrees of fluctuation harbored less connected, potentially less stable microbiomes.
+This points to a link between microbiome and environmental stability, and warrants further investigation.
 
@@
 
-It is important to note that the microbiome consists of bacteria and viruses, as well as fungi, archaea, eukaryotes such as *Demodex* mites, etc.
+It is important to note that the networks in this study were built using operational genomic units, which represent groups of highly genomically similar bacteria or phages as clustered sub-populations.
+This operationally defined approach allows us to study whole community networks, but limits our ability to make conclusions about interactions among specific phage or bacterial entities, such as species or strains.
+Although this approach lacks the resolution for drawing conclusions about specific bacteria or phages, future work could begin addressing these questions through improved binning methods and deeper metagenomic shotgun sequencing.
+
+It is also important to consider the multiple displacement amplification (MDA) methods used in the library preparations of the gut diet and twin studies.
+Multiple displacement amplification was widely used to amplify shotgun metagenomic DNA before sequencing.
+This amplification process was unfortunately not random and resulted in significant biases toward ssDNA viral genomes within viromes [@Kim:2008to; @Kim:2011hp].
+Therefore there are likely biases in the relative abundace values within those studies.
+However, by comparing samples within studies, which were all treated the same, we were able to provide the fairest possible comparison for drawing conclusions.
+Future work will build on these findings by avoiding the use of MDA while implementing contemporary methods.
+
+The microbiome consists of bacteria and viruses, as well as fungi, archaea, eukaryotes such as *Demodex* mites, etc.
 These microbiome components also interact with human immune cells and other non-microbial community members.
@@
 By focusing on relationships between bacterial and viral communities, we can begin studying them as interacting cohorts they are, instead of isolated entities.
-This work will also provide a foundation for future studies to begin evaluating how microbiome network dynamics change in disease states, and how the information can be applied therapeutically.
+By highlighting the impacts different human environments (the skin and gut) can have on general microbiome connectivity, this work will inform future work into the underlying evolutionary and ecological mechanisms.
 
@@
 Represetnative sequences were aligned to the European Nucleotide Archive (ENA) bacterial reference database using the blastn algorithm (e-value < $10^{-5}$).
+79% of the bacterial OGUs were able to be identified.
+
+## Phage OGU Identification
+To confirm a lack of phage sequences in the bacterial OGU dataset, we performed a relatively strict blast nucleotide alignment of the bacterial OGU representative sequences using an e-value < $10^{-25}$.
+We used a stricter threshold because we already know there are genomic similarities between bacteria and phage OGUs from the interactive model, but we were interested in contigs with high enough similarity to references that they may indeed be from phages.
+2% of the OGUs had nucleotide similarities to known bacteriophage genomes, although the alignments were short (a couple of kb) and represented a small fraction of the alignment sequences.
 

### 2017-05-18 doc/manuscript.Rmd
Fixed the paper format.
edits
 Please enter the commit message for your changes. Lines starting

@@
 Viruses and bacteria are critical components of the human microbiome and play important roles in health and disease.
-Most previous work has relied on studying microbes and viruses in isolation, thereby reducing them to two separate communities.
-Such approaches are unable to capture how these microbial communities interact, including processes such as the mechanisms that maintain community stability or allow phage-host populations to co-evolve.
+Most previous work has relied on studying microbes and viruses independently <!--I'd changed this previously, as "in isolation" can conjucture a whole slew of other things -->, thereby reducing them to two separate communities.
+Such approaches are unable to capture how these microbial communities interact, such as through processes that maintain community stability or allow phage-host populations to co-evolve.
 We developed and implemented a network-based analytical approach to describe phage-bacteria network diversity throughout the human body.
-We accomplished this by building a machine learning algorithm to predict which phages infected which bacteria in a given microbiome.
+We accomplished this by building a machine learning algorithm to predict which phages could infect which bacteria in a given microbiome.
 This algorithm was applied to paired viral and bacterial metagenomic sequence sets from three previously published human cohorts.
@@
 Bacterial communities have been associated with diseases including a range of skin conditions [@Hannigan:2013im], acute and chronic wound healing conditions [@Hannigan:2014be; @Loesche:2016ev], and gastrointestinal diseases including inflammatory bowel disease [@He:2016ch; @Norman:2015kb], *Clostridium difficile* infections [@Seekatz:2016fz], and colorectal cancer [@Zackular:2014fba; @Baxter:2014hb].
-Altered viromes (virus communities consisting primarily of bacteriophages) have also been associated with various diseases and perturbations including inflammatory bowel disease [@Norman:2015kb; @Manrique:2016dx], periodontal disease [@Ly:2014ew], spread of antibiotic resistance [@Modi:2013fia], and others [@Monaco:2016ita; @Hannigan:2015fz; @Minot:2011ez; @SantiagoRodriguez:2015gd; @Abeles:2015dy; @Abeles:2014kj].
+Altered viromes (virus communities consisting primarily of bacteriophages) have been also <!-- that was a split infinitive, you can't do that, also has to come after have been --> associated with various diseases and perturbations including inflammatory bowel disease [@Norman:2015kb; @Manrique:2016dx], periodontal disease [@Ly:2014ew], spread of antibiotic resistance [@Modi:2013fia], and others [@Monaco:2016ita; @Hannigan:2015fz; @Minot:2011ez; @SantiagoRodriguez:2015gd; @Abeles:2015dy; @Abeles:2014kj].
 The viruses act in concert with their microbial hosts as a single ecological community [@Haerter:2014ii].
-Viruses influence their living microbial host communities through processes including lysing their hosts, modulating host gene expression [@Lindell:2005gz, @Tyler:2013fl, @Hargreaves:2014ja], influencing evolutionary processes, such as horizontal gene transfer [@Moon:2015fa, @Modi:2013fi, @Ogg:1981th, @Frost:2005dn] or antagonistic co-evolution [@Koskella:2014ds], and by altering ecosystem processes and elemental stoichiometry [@Jover:2014gq].
+Viruses influence their living microbial host communities through processes including lysis, modulating host gene expression [@Lindell:2005gz, @Tyler:2013fl, @Hargreaves:2014ja], influencing evolutionary processes, such as horizontal gene transfer [@Moon:2015fa, @Modi:2013fi, @Ogg:1981th, @Frost:2005dn] or antagonistic co-evolution [@Koskella:2014ds], and by altering ecosystem processes and elemental stoichiometry [@Jover:2014gq].
 
-Most previous human microbiome work has focused on bacterial and viral communities, but have relied on studying them in isolation by reducing them to two separate communities [@Norman:2015kb; @Manrique:2016dx; @Ly:2014ew; @Monaco:2016ita; @Hannigan:2015fz; @Minot:2011ez; @SantiagoRodriguez:2015gd; @Abeles:2015dy; @Abeles:2014kj].
+Previous human microbiome work has focused on bacterial and viral communities, but have reduced them to two separate communities by studying them independently  [@Norman:2015kb; @Manrique:2016dx; @Ly:2014ew; @Monaco:2016ita; @Hannigan:2015fz; @Minot:2011ez; @SantiagoRodriguez:2015gd; @Abeles:2015dy; @Abeles:2014kj].
 In reality, bacteria and phage communities are dynamic and complex.
@@
 Removal of bacteria or phage can disrupt or even collapse those ecosystems [@Haerter:2014ii; @Harcombe:2005fd; @Middelboe:2001fl; @Poisot:2011jc; @Thompson:2012ki; @Moebus:1981kp; @Flores:2013hc; @Poisot:2012fh; @Poisot071993; @Flores:2011bh; @Jover:2015ev].
-Relationship-based network approaches allow us to capture this information.
-We leveraged machine learning and graph theory techniques to characterize human bacterial and phage communities by their relationships.
+Relationship-based network approaches allow us to capture this interaction information.
+We leveraged machine learning and graph theory techniques to characterize human bacterial and phage communities by their interferred relationships.
+<!--this is redundant to the previous sentence and distracting...not a good end of this paragraph-->
+
 We characterized human skin and gut virus-bacteria network diversity to provide a foundation for further studies of disease network dynamics and to gain broader insights into human microbiome network diversity across different body sites.
 
-We investigated human bacterial-phage networks by utilizing three published microbiome datasets that contained paired virus and total metagenomic sequence sets, which primarily consisted of bacteria [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei].
+Human bacterial-phage network diversity was established and explored by utilizing three published microbiome datasets that contained paired virus and whole community metagenomic sequence sets dominated by bacteria [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei].
 Our approach built off of previous large-scale phage-bacteria microbiome network analyses by inferring interactions using metagenomic datasets, instead of using culture-based techniques [@Flores:2013hc].
-This metagenomic interaction inference model was powered beyond previous models [@Edwards:2015iz, @Roux:2016cc] by its inclusion of protein interaction data as well as the inclusion of both negative and positive interactions.
-Through this approach, we were able to provide a basic understanding of the network dynamics associated with healthy human phage and bacterial communities.
-By building and utilizing a microbiome network, we found that different people, body sites, and anatomical locations not only support distinct microbiome membership and diversity [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei; @Grice:2009eea; @Findley:2013jf; @Costello:2009im, @Consortium:2012iz], but also support communities with distinct communication structures and propensities toward community instability.
-Through an improved understanding of the healthy state of network structures across the human body, we empower future studies to begin investigating how these community structures are influenced by disease state and their overall impact on human health.
+This metagenomic interaction inference model was powered beyond previous models [@Edwards:2015iz, <!--Lima-Mendez et al., Science 2015; her Tara paper, they predicted phage-host interactions and described phage-host ecology/host range distributions and compared to the Flores work of nestedness/modularity of host ranges-->, @Roux:2016cc] by its inclusion of protein interaction data, as well as experimentally verified negative and positive interactions, i.e., "who does and does not infect whom." Through this approach, we were able to provide a basic understanding of the network dynamics associated with phage and bacterial communities of healthy humans.
+By building and utilizing a microbiome network, we found that different people, body sites, and anatomical locations not only support distinct microbiome membership and diversity [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei; @Grice:2009eea; @Findley:2013jf; @Costello:2009im, @Consortium:2012iz], but also support ecological communities with distinct communication structures and propensities toward community instability.
+Through an improved understanding of the healthy state of network structures across the human body, we empower future studies to  investigate how these community structures are influenced by disease state and their overall impact on human health.
 
@@
 We studied the differences in virus-bacteria interaction networks across healthy human bodies by leveraging previously published sequence sets containing purified virome samples paired with bacterial metagenomes from whole metagenomic shotgun sequences.
-Our study contained three datasets which explored the impact of diet on the healthy human gut virome [@Minot:2011ez], the impact of anatomical location on the healthy human skin virome [@Hannigan:2015fz], and the viromes of monozygotic twins and their mothers [@Reyes:2010cwa; @Turnbaugh:2009ei].
+Our study contained three datasets that explored the impact of diet on the healthy human gut virome [@Minot:2011ez], the impact of anatomical location on the healthy human skin virome [@Hannigan:2015fz], and the viromes of monozygotic twins and their mothers [@Reyes:2010cwa; @Turnbaugh:2009ei].
 The twin and diet studies utilized multiple displacement amplification methods in their library preparations.
-These datasets were selected because they included virome samples subjected to virus-like particle (VLP) purification.
+<!--this is weird here, I would just bring it up in in the discussion of these studies or when you talk about the caveats of the overall approach/interpretation, cautioning about the fact that MDA was used in these two studies, and it will bias the representation (which nodes exist) and relative node quantification (node/cluster size) --> These datasets were selected because they included virome samples subjected to virus-like particle (VLP) purification.
 To this end, they employed combinations of filtration, chloroform/DNase treatment, and cesium chloride gradients to eliminate organismal DNA and thereby allow for direct assessment of both the extracellular and fully-assembled intracellular virome **(Supplemental Figure \ref{SequenceStats} A-B)** [@Minot:2011ez, @Hannigan:2015fz, @Reyes:2010cwa; @Turnbaugh:2009ei].
@@
 The bacterial and viral sequences from these studies were quality filtered and assembled into contigs.
-We further grouped the related bacterial and phage contigs into operationally defined units which were based on their k-mer frequencies and co-abundance patterns, similar to previous reports **(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})** [@Roux:2016cc].
+We further grouped the related bacterial and phage contigs into operationally defined units based on their k-mer frequencies and co-abundance patterns, similar to previous reports **(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})** [@Roux:2016cc].
 We referred to these operationally defined groups of related contigs as operational genomic units (OGUs).
-Each OGU represented either a genomically similar sub-population of bacteria or phages.
+Each OGU represented a genomically similar sub-population of either bacteria or phages.
 Contig lengths within clusters ranged between $10^{3}$ and $10^{5.5}$ bp **(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})**.
@@
 Only bacteria and phages were used in the model.
-The training set contained 43 diverse bacterial species and 30 diverse phage strains, with both broad and specific infectious ranges **(Supplemental Figure \ref{ValidationOverview} A - B)**.
+The training set contained 43 diverse bacterial species and 30 diverse phage strains, including both broad and specific infectious ranges **(Supplemental Figure \ref{ValidationOverview} A - B)**.
 Phages with linear and circular genomes, as well as ssDNA and dsDNA genomes, were included in the analysis.
@@
 This training set included both positive relationships (a phage infects a bacterium) and negative relationships (a phage does not infect a bacterium).
-This allowed us to validate the false positive and false negative rates associated with our candidate models, thereby building upon much previous work that only considered positive relationships [@Edwards:2015iz].
+This allowed us to validate the false positive and false negative rates associated with our candidate models, thereby building upon previous work that only considered positive relationships [@Edwards:2015iz].
 
@@
 Protein family interactions were moderately important to the model, and CRISPRs were largely uninformative, due to the lack of identifiable CRISPRs in the dataset and their redundancy with the nucleotide similarity methods **(Figure \ref{RocCurve} B)**.
-Approximately one third of the training set relationships yielded no score and were therefore unable to be assigned an interaction prediction **(Figure \ref{RocCurve} C)**.
+Approximately one third of the training set relationships yielded no score and therefore were unable to be assigned an interaction prediction **(Figure \ref{RocCurve} C)**.
 
@@
 # Discussion
-Foundational microbiome work has provided a baseline understanding of the human microbiome by characterizing bacteria and viral diversity across the human body, as well as other environments [@Grice:2009eea; @Findley:2013jf; @Hannigan:2015fz; @Costello:2009im, @Consortium:2012iz; @Schloss:2005hz; @Minot:2011ez].
-Here, we offer an initial understanding of how phage-bacteria networks differ throughout the human body, so as to provide a baseline for future studies of how and why microbiome networks differ in disease states.
+Foundational work has provided a baseline understanding of the human microbiome by characterizing bacterial and viral diversity across the human body [@Grice:2009eea; @Findley:2013jf; @Hannigan:2015fz; @Costello:2009im, @Consortium:2012iz; @Schloss:2005hz; @Minot:2011ez]<!--I think it ok to reframe/reset the discussion to focus on human-->.
+Here, we offer a novel approach to describe how phage-bacteria networks differ throughout the human body to provide a baseline for future studies of how and why microbiome networks differ in disease states.
 We developed and implemented a network-based analytical approach to evaluate the basic properties of the human microbiome through bacteria and phage relationships, instead of membership or diversity alone.
 This enabled the application of network theory to provide a new perspective on complex ecological communities.
-We utilized metrics of connectivity to model the extent to which communities of bacteria and phages interact, through mechanisms including horizontal gene transfer, modulated bacterial gene expression, and alterations in abundance.
+We utilized metrics of connectivity to model the extent to which communities of bacteria and phages interact through mechanisms such as horizontal gene transfer, modulated bacterial gene expression, and alterations in abundance.
 
@@
 Gut network structure was not conserved among family members.
-These findings suggested that microbiome network properties, including stability, the potential for horizontal gene transfer, co-evolution, etc, were personal and may be impacted by personal factors ranging from the body's immune system to external environmental conditions, such as climate and diet.
-The ability of environmental conditions to shape gut and skin microbiome network structure was further supported by our finding that diet and skin location were associated with altered network structures.
+These findings suggested that microbiome network properties, such as stability <!--this is a network property -->, the potential for horizontal gene transfer between members, co-evolution, etc.<!--but these are not network properties...so this sentence doesn't make sense.
+But do you mean the "potential for community interactions inferred from network properties" were personal?
+So could change this sentence to: "These findings suggested that the community properties inferred from microbiome interaction network structures, such as stability, the potential for horizontal gene transfer between members, co-evolution of populations, etc., were person-specific.
+These properties may be impacted by personal factors ranging from the body's immune system to external environmental conditions, such as climate and diet."-->, were personal and may be impacted by personal factors ranging from the body's immune system to external environmental conditions, such as climate and diet.
 
-We found evidence that diet was sufficient to alter gut microbiome network connectivity.
-Although our sample size was small, our findings provided evidence that high-fat diets were less connected than low-fat diets, and high-fat diets may therefore lead to less stable communities with a decreased ability for microbes to directly influence one another.
+The ability of environmental conditions to shape gut and skin microbiome interaction network structure was further supported by our finding that diet and skin location were associated with altered network structures.
+<!-- This is a more appropriate leader sentence for the next two anecdotes.--> We found evidence that diet was sufficient to alter gut microbiome network connectivity.
+Although our sample size was small, our findings provided evidence that high-fat diets were less connected than low-fat diets, and high-fat diets therefore may lead <!-- another spit infinitive--> to less stable communities with a decreased ability for microbes to directly influence one another.
 We supported this finding with the observation that obesity may have been associated with decreased network connectivity.
 Together these findings suggest the food we eat may not only impact which microbes colonize our guts, but may also impact their infectious interactions.
-Further work will certainly be required to characterize these relationships with a larger cohort.
+Further work will be required to characterize these relationships with a larger cohort.
 
-In addition to diet, the skin microbiome network structure varied by skin environment.
+In addition to diet, the skin environment influenced the microbiome interaction network structure, as well.
 Network structure differed between environmentally exposed and occluded skin sites.
@@
 
-While these findings take us an important step closer to understanding the microbiome through interspecies relationships, there are caveats to the approach that should be noted.
-First, although our infection classification model is advantageous over existing models, we recognize that, like most classification models, ours is only as good as its training set.
+While these findings take us an important step closer to understanding the microbiome through interspecies relationships, there are caveats to and considerations regarding the approach.
+First, <!--"although our infection classification model is advantageous over existing models," ...this was not tested, so wouldn't say it, but it is also not needed--> as with most classification models, the infection classification model developed and applied is only as good as its training set--in this case, the collection of experimentally-verified positive and negative infection data, where genomes of all members are fully sequenced.
 Large-scale experimental screens for phage and bacteria infectious interactions that report high-confidence negative interactions (i.e., no infection) will provide more robust model training and improved model performance.
@@
 
-Second, although our analyses offer an informative proof of concept, this work was done retrospectively and relied on published research from as long as seven years ago.
-These archived datasets were limited by the technology and costs of the time, meaning the datasets were suboptimally powered for the statistical analysis we strive for today.
-Despite these limitations, we were able to present initial conclusions.
-Follow-up studies will validate our findings and inform the design and interpretation of future studies.
+Second, although our analyses offer an informative proof of concept, this work was done retrospectively and relied on existing data up to seven years old.
+These archived datasets were limited by the technology and costs of the time <!--the MDA issue falls under this, in my mind, as issue of using old datsets.
+So integrated it here, also to bring down the focus a bit/not give it it's own paragraph-->.
+This resulted in small sequencing effort (as compared to today's dataset sizes) and thus datasets that were suboptimally powered for the statistical analyses we strive for today.
+Further, two studies, the diet and twin studies, relied on multiple displacement amplification (MDA) in their library preparations--an approach used to overcome the large nucleic acids requirements, especially of older sequencing library generation protocols.
+MDA results in significant biases in microbobial community composition <!--Yilmaz, S., Allgaier, M.
+& Hugenholtz, P., 2010, Multiple displacement amplification compromises quantitative analysis of metagenomes, Nat Methods, 7(12), pp.
+943-4-->, as well as toward ssDNA viral genomes [@Kim:2008to; @Kim:2011hp], thus rendering the resulting microbial and viral metagenomes non-quantitative.
+<!--because Yilmaz et al., show that the bias depends on the initial community structure, I think this means that an MDA bias would also compromise your ability to compare samples even within studies, so I removed this: "However, by comparing samples within studies, which were all treated the same, we were able to provide the fairest possible comparison for drawing conclusions."--> Future work that employs larger sequence datasets and that avoids the use of MDA will build on and validate our findings, as well as inform the design and interpretation of further studies.
 
-It is important to note that the networks in this study were built using operational genomic units, which represent groups of highly genomically similar bacteria or phages as clustered sub-populations.
-This operationally defined approach allows us to study whole community networks, but limits our ability to make conclusions about interactions among specific phage or bacterial entities, such as species or strains.
-Although this approach lacks the resolution for drawing conclusions about specific bacteria or phages, future work could begin addressing these questions through improved binning methods and deeper metagenomic shotgun sequencing.
-
-It is also important to consider the multiple displacement amplification (MDA) methods used in the library preparations of the gut diet and twin studies.
-Multiple displacement amplification was widely used to amplify shotgun metagenomic DNA before sequencing.
-This amplification process was unfortunately not random and resulted in significant biases toward ssDNA viral genomes within viromes [@Kim:2008to; @Kim:2011hp].
-Therefore there are likely biases in the relative abundace values within those studies.
-However, by comparing samples within studies, which were all treated the same, we were able to provide the fairest possible comparison for drawing conclusions.
-Future work will build on these findings by avoiding the use of MDA while implementing contemporary methods.
-
-The microbiome consists of bacteria and viruses, as well as fungi, archaea, eukaryotes such as *Demodex* mites, etc.
-These microbiome components also interact with human immune cells and other non-microbial community members.
-Our study only represents an initial step toward network analysis, and future work will include more components of the microbiome to gain a more complete understanding of the community through its interactions.
-In addition to these microbiome components, additional relationship types will also be added so as to provide a more robust network for analysis.
+Finally, the networks in this study were built using operational genomic units, which represent groups of highly similar bacteria or phage genomes or genome fragments as clustered sub-populations.
+This operationally defined approach allows us to study whole community networks, but limits our ability to make conclusions about interactions among specific phage or bacterial species or populations.
+Although this approach lacks the resolution for drawing conclusions about specific bacteria or phages, future work could begin addressing these questions through improved binning methods and deeper metagenomic shotgun sequencing, as well as an improved understanding of what defines ecologically and evolutionarily cohesive units for both phage and bacteria, e.g., Polz et al., 2006 <!--Polz, M.F., Hunt, D.E., Preheim, S.P.
+& Weinreich, D.M., 2006, Patterns and mechanisms of genetic and phenotypic differentiation in marine microbes, Philos Trans R Soc Lond B Biol Sci, 361(1475), pp.
+2009-21.
+-->
 
 Together our work takes an initial step towards defining bacteria-virus interaction profiles as a characteristic of human-associated microbial communities.
-By focusing on relationships between bacterial and viral communities, we can begin studying them as interacting cohorts they are, instead of isolated entities.
-By highlighting the impacts different human environments (the skin and gut) can have on general microbiome connectivity, this work will inform future work into the underlying evolutionary and ecological mechanisms.
+This approach revealed the impacts different human environments (e.g., the skin and gut) can have on microbiome connectivity.
+By focusing on relationships between bacterial and viral communities, they are studied as the interacting cohorts they are, rather than as independent entities.
+While our developed bacteria-phage interaction framework is a novel conceptual advance, the microbiome also consists of archaea and small eukaryotes, such as fungi and *Demodex* mites <!--reference?-->.
+Further, the components of the microbiome can interact with human immune cells and other non-microbial community members <!--reference?-->.
+Future work will build from our approach and include these additional community members and their diverse interactions and relationships (e.g., beyond phage-bacteria).
+This will result in a more robust network and a more holistic understanding of the human-associated microbiome and the evolutionary and ecological processes that drive its assembly and function <!--there were a lot of redunant thougths here I streamlined.
+Also you'll see I merged the final paragraphs to make the shared forward looking elements stronger.-->
 
@@
 ## Data Availability
-All associated source code is available on GitHub at the following repository:
+All associated source code is available on GitHub at the following repository <!--don't know if you want to change this to reflect journal change?-->:
 
@@
 ## Data Acquisition & Quality Control
-Raw sequencing data and associated metadata was acquired from the NCBI sequence read archive (SRA).
-Supplementary metadata was acquired from the same SRA repositories and their associated manuscripts.
+Raw sequencing data and associated metadata were acquired from the NCBI sequence read archive (SRA).
+Supplementary metadata were acquired from the same SRA repositories and their associated manuscripts.
 The gut virome diet study (SRA: `SRP002424`), twin virome studies (SRA: `SRP002523`; `SRP000319`), and skin virome study (SRA: `SRP049645`) were downloaded as `.sra` files.
@@
 Sequences were quality trimmed using the Fastx toolkit (v0.0.14) to exclude bases with quality scores below 33 and shorter than 75 bp [@FASTXToolkit:wr].
-Paired end reads were filtered to exclude and sequences missing their corresponding pair using the `get_trimmed_pairs.py` available in the source code.
+Paired end reads were filtered to exclude sequences missing their corresponding pair using the `get_trimmed_pairs.py` script available in the source code.
 
@@
 Contigs often represent large fragments of genomes.
-In order to reduce redundancy, and the resulting artificially inflated genomic richness within our dataset, it was important to bin contigs into operational units based on their similarity.
+In order to reduce redundancy and the resulting artificially inflated genomic richness within our dataset, it was important to bin contigs into operational units based on their similarity.
 This approach is conceptually similar to the clustering of related 16S rRNA sequences into operational taxonomic units (OTUs), although here we are clustering contigs into operational genomic units (OGUs) [@Schloss:2005hz].
 
-We clustered contigs using the CONCOCT algorithm (v0.4.0) [@Alneberg:2014fc].
+Contigs were clustered using the CONCOCT algorithm (v0.4.0) [@Alneberg:2014fc].
 Because of our large dataset and limits in computational efficiency, we randomly subsampled the dataset to include 25% of all samples, and used these to inform contig abundance within the CONCOCT algorithm.
@@
 The longest contig from each OGU was used as its representative sequence for identification.
-Represetnative sequences were aligned to the European Nucleotide Archive (ENA) bacterial reference database using the blastn algorithm (e-value < $10^{-5}$).
+Representative sequences were aligned to the European Nucleotide Archive (ENA) bacterial reference database using the blastn algorithm (e-value < $10^{-5}$).
 79% of the bacterial OGUs were able to be identified.
+<!--what was this ID used for?
+i don't recall any discussion of inferred cluster taxonomic identity-->
 
 ## Phage OGU Identification
-To confirm a lack of phage sequences in the bacterial OGU dataset, we performed a relatively strict blast nucleotide alignment of the bacterial OGU representative sequences using an e-value < $10^{-25}$.
-We used a stricter threshold because we already know there are genomic similarities between bacteria and phage OGUs from the interactive model, but we were interested in contigs with high enough similarity to references that they may indeed be from phages.
+To confirm a lack of phage sequences in the bacterial OGU dataset, we performed a relatively strict <!--indeed, this is relative.
+can you relate it?
+What is the distribution of evalues or bitscores from this blastp?
+any anecdotes that it was a useful cut-off?--> blast nucleotide alignment of the bacterial OGU representative sequences using an e-value < $10^{-25}$.
+We used a stricter threshold because we know there are genomic similarities between bacteria and phage OGUs from the interactive model, but we were interested in contigs with high enough similarity to references that they may indeed be from phages.
 2% of the OGUs had nucleotide similarities to known bacteriophage genomes, although the alignments were short (a couple of kb) and represented a small fraction of the alignment sequences.
+<!-- shouldn't you also consider a relative length component?
+Meaning, contigs with only phage similarity along its entire length are more likely phage than contigs with only small portion similar to phage, while rest has similarity to microbes...thus more likley prophage and ok to stay classified as bacteria.
+Why not just run already created probabilitic models that use many criteria, like VirSorter?
+Then you get predictions of phage (cat 1-3) versus propahge (cat 4-6) and the confidence of the prediction.
+-->
 
@@
 The resulting bitscore of each alignment was recorded for use in our classification model.
+<!-- I get suspicious whenever e-value is used as a threshold.
+Do you have anecdotal evidence this a is a good criterion?
+Should be easy to check-->
 
@@
 We identified shared genes between bacterial and phage genomes by assessing amino acid similarity between the genes using the Diamond protein alignment algorithm (v0.7.11.60) [@Buchfink:2015ki].
-The mean alignment bitscores for each genome pair was recorded for use in our classification model.
+The mean alignment bitscores for each genome pair were recorded for use in our classification model.
 
 ##### Protein - Protein Interactions
-The final method we used for predicting infectious interactions between bacteria and phages was by detecting pairs of genes whose proteins are known to interact.
+The final method used for predicting infectious interactions between bacteria and phages was the detection of pairs of genes whose proteins are known to interact.
 We assigned bacterial and phage genes to protein families by aligning them to the Pfam database using the Diamond protein alignment algorithm.
 We then identified which pairs of proteins were predicted to interact using the Pfam interaction information within the Intact database [@Orchard:2014hq].
-The mean bitscores of the matches between each pair were recorded for use in our classification model.
+The mean bitscores of the matches between each pair were recorded for use in the classification model.
 
-## Virome Network Construction
+## Interaction Network Construction
 The bacteria and phage operational genomic units (OGUs) were scored using the same approach as outlined above.
 The infectious pairings between bacteria and phage OGUs were classified using the random forest model described above.
-The predicted infectious pairings and all associated metadata were saved as a graph database using Neo4j graph database software (v2.3.1) [@Neoj:Fuwr6PBN].
+The predicted infectious pairings and all associated metadata were used to populate a graph database using Neo4j graph database software (v2.3.1) [@Neoj:Fuwr6PBN].
 This network was used for downstream community analysis.
@@
 
+\newpage
+
 ![**Impact of Diet and Obesity on Gut Network Structure.** *(A) Quantification of average degree centrality (number of edges per node) and (B) closeness centrality (average distance from each node to every other node) of gut microbiome networks of subjects limited to exclusively high-fat or low-fat diets.
@@
 (C) Quantification of average degree centrality and (D) closeness centrality between obese and healthy adult women.*\label{dietnetworks}](../figures/dietnetworks.pdf){ width=90% }
+
+\newpage
 

### 2017-05-23 doc/manuscript.Rmd
Updated from Duhaime suggestions.

@@
 
-***Journal***: Genome Research ([*Preparation Details*](http://genome.cshlp.org/site/misc/ifora_mspreparation.xhtml))
+***Journal***: mSystems
 
@@
 Viruses and bacteria are critical components of the human microbiome and play important roles in health and disease.
-Most previous work has relied on studying microbes and viruses independently <!--I'd changed this previously, as "in isolation" can conjucture a whole slew of other things -->, thereby reducing them to two separate communities.
+Most previous work has relied on studying microbes and viruses independently, thereby reducing them to two separate communities.
 Such approaches are unable to capture how these microbial communities interact, such as through processes that maintain community stability or allow phage-host populations to co-evolve.
@@
 Bacterial communities have been associated with diseases including a range of skin conditions [@Hannigan:2013im], acute and chronic wound healing conditions [@Hannigan:2014be; @Loesche:2016ev], and gastrointestinal diseases including inflammatory bowel disease [@He:2016ch; @Norman:2015kb], *Clostridium difficile* infections [@Seekatz:2016fz], and colorectal cancer [@Zackular:2014fba; @Baxter:2014hb].
-Altered viromes (virus communities consisting primarily of bacteriophages) have been also <!-- that was a split infinitive, you can't do that, also has to come after have been --> associated with various diseases and perturbations including inflammatory bowel disease [@Norman:2015kb; @Manrique:2016dx], periodontal disease [@Ly:2014ew], spread of antibiotic resistance [@Modi:2013fia], and others [@Monaco:2016ita; @Hannigan:2015fz; @Minot:2011ez; @SantiagoRodriguez:2015gd; @Abeles:2015dy; @Abeles:2014kj].
+Altered viromes (virus communities consisting primarily of bacteriophages) have been also associated with various diseases and perturbations including inflammatory bowel disease [@Norman:2015kb; @Manrique:2016dx], periodontal disease [@Ly:2014ew], spread of antibiotic resistance [@Modi:2013fia], and others [@Monaco:2016ita; @Hannigan:2015fz; @Minot:2011ez; @SantiagoRodriguez:2015gd; @Abeles:2015dy; @Abeles:2014kj].
 The viruses act in concert with their microbial hosts as a single ecological community [@Haerter:2014ii].
@@
 Relationship-based network approaches allow us to capture this interaction information.
-We leveraged machine learning and graph theory techniques to characterize human bacterial and phage communities by their interferred relationships.
-<!--this is redundant to the previous sentence and distracting...not a good end of this paragraph-->
-
-We characterized human skin and gut virus-bacteria network diversity to provide a foundation for further studies of disease network dynamics and to gain broader insights into human microbiome network diversity across different body sites.
+We therefore leveraged machine learning and graph theory techniques to characterize human bacterial and phage communities by their inferred relationships.
+By doing so, we provide a foundation for further studies of disease network dynamics and gain broader insights into human microbiome network diversity across different body sites.
 
 Human bacterial-phage network diversity was established and explored by utilizing three published microbiome datasets that contained paired virus and whole community metagenomic sequence sets dominated by bacteria [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei].
-Our approach built off of previous large-scale phage-bacteria microbiome network analyses by inferring interactions using metagenomic datasets, instead of using culture-based techniques [@Flores:2013hc].
-This metagenomic interaction inference model was powered beyond previous models [@Edwards:2015iz, <!--Lima-Mendez et al., Science 2015; her Tara paper, they predicted phage-host interactions and described phage-host ecology/host range distributions and compared to the Flores work of nestedness/modularity of host ranges-->, @Roux:2016cc] by its inclusion of protein interaction data, as well as experimentally verified negative and positive interactions, i.e., "who does and does not infect whom." Through this approach, we were able to provide a basic understanding of the network dynamics associated with phage and bacterial communities of healthy humans.
+Our approach built off of previous large-scale phage-bacteria microbiome network analyses by inferring interactions using metagenomic datasets, instead of using culture-based techniques, which limit the scale of the experiments and analyses [@Flores:2013hc].
+This metagenomic interaction inference model built on previous models, which have utilized a variety of techniques including taxonomic-based co-occurrence models using linear relationships between bacteria and phages [@LimaMendez:2015hw], as well as nucleotide similarity (including CRISPR targeting) models using OGUs [@@Roux:2016cc] and whole genomes [@Edwards:2015iz].
+Our approach uniquely included protein interaction data, and was validated on experimentally determined negative and positive interactions, i.e., "who does and does not infect whom." Through this approach, we were able to provide a basic understanding of the network dynamics associated with phage and bacterial communities on the human body.
 By building and utilizing a microbiome network, we found that different people, body sites, and anatomical locations not only support distinct microbiome membership and diversity [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei; @Grice:2009eea; @Findley:2013jf; @Costello:2009im, @Consortium:2012iz], but also support ecological communities with distinct communication structures and propensities toward community instability.
@@
 The twin and diet studies utilized multiple displacement amplification methods in their library preparations.
-<!--this is weird here, I would just bring it up in in the discussion of these studies or when you talk about the caveats of the overall approach/interpretation, cautioning about the fact that MDA was used in these two studies, and it will bias the representation (which nodes exist) and relative node quantification (node/cluster size) --> These datasets were selected because they included virome samples subjected to virus-like particle (VLP) purification.
+These datasets were selected because they included virome samples subjected to virus-like particle (VLP) purification.
 To this end, they employed combinations of filtration, chloroform/DNase treatment, and cesium chloride gradients to eliminate organismal DNA and thereby allow for direct assessment of both the extracellular and fully-assembled intracellular virome **(Supplemental Figure \ref{SequenceStats} A-B)** [@Minot:2011ez, @Hannigan:2015fz, @Reyes:2010cwa; @Turnbaugh:2009ei].
@@
 # Discussion
-Foundational work has provided a baseline understanding of the human microbiome by characterizing bacterial and viral diversity across the human body [@Grice:2009eea; @Findley:2013jf; @Hannigan:2015fz; @Costello:2009im, @Consortium:2012iz; @Schloss:2005hz; @Minot:2011ez]<!--I think it ok to reframe/reset the discussion to focus on human-->.
-Here, we offer a novel approach to describe how phage-bacteria networks differ throughout the human body to provide a baseline for future studies of how and why microbiome networks differ in disease states.
-We developed and implemented a network-based analytical approach to evaluate the basic properties of the human microbiome through bacteria and phage relationships, instead of membership or diversity alone.
+Foundational work has provided a baseline understanding of the human microbiome by characterizing bacterial and viral diversity across the human body [@Grice:2009eea; @Findley:2013jf; @Hannigan:2015fz; @Costello:2009im, @Consortium:2012iz; @Schloss:2005hz; @Minot:2011ez].
+Here, we offer an approach to describe how phage-bacteria networks differ throughout the human body to provide a baseline for future studies of how and why microbiome networks differ in disease states.
+We developed and implemented a network-based analytical model to evaluate the basic properties of the human microbiome through bacteria and phage relationships, instead of membership or diversity alone.
 This enabled the application of network theory to provide a new perspective on complex ecological communities.
@@
 Gut network structure was not conserved among family members.
-These findings suggested that microbiome network properties, such as stability <!--this is a network property -->, the potential for horizontal gene transfer between members, co-evolution, etc.<!--but these are not network properties...so this sentence doesn't make sense.
-But do you mean the "potential for community interactions inferred from network properties" were personal?
-So could change this sentence to: "These findings suggested that the community properties inferred from microbiome interaction network structures, such as stability, the potential for horizontal gene transfer between members, co-evolution of populations, etc., were person-specific.
-These properties may be impacted by personal factors ranging from the body's immune system to external environmental conditions, such as climate and diet."-->, were personal and may be impacted by personal factors ranging from the body's immune system to external environmental conditions, such as climate and diet.
+These findings suggested that the community properties inferred from microbiome interaction network structures, such as stability, the potential for horizontal gene transfer between members, co-evolution of populations, etc., were person-specific.
+These properties may be impacted by personal factors ranging from the body's immune system to external environmental conditions, such as climate and diet.
 
 The ability of environmental conditions to shape gut and skin microbiome interaction network structure was further supported by our finding that diet and skin location were associated with altered network structures.
-<!-- This is a more appropriate leader sentence for the next two anecdotes.--> We found evidence that diet was sufficient to alter gut microbiome network connectivity.
-Although our sample size was small, our findings provided evidence that high-fat diets were less connected than low-fat diets, and high-fat diets therefore may lead <!-- another spit infinitive--> to less stable communities with a decreased ability for microbes to directly influence one another.
+We found evidence that diet was sufficient to alter gut microbiome network connectivity.
+Although our sample size was small, our findings provided evidence that high-fat diets were less connected than low-fat diets, and high-fat diets therefore may lead to less stable communities with a decreased ability for microbes to directly influence one another.
 We supported this finding with the observation that obesity may have been associated with decreased network connectivity.
@@
 
-In addition to diet, the skin environment influenced the microbiome interaction network structure, as well.
+In addition to diet, the skin environment influenced the microbiome interaction network structure as well.
 Network structure differed between environmentally exposed and occluded skin sites.
@@
 While these findings take us an important step closer to understanding the microbiome through interspecies relationships, there are caveats to and considerations regarding the approach.
-First, <!--"although our infection classification model is advantageous over existing models," ...this was not tested, so wouldn't say it, but it is also not needed--> as with most classification models, the infection classification model developed and applied is only as good as its training set--in this case, the collection of experimentally-verified positive and negative infection data, where genomes of all members are fully sequenced.
+First, as with most classification models, the infection classification model developed and applied is only as good as its training set -- in this case, the collection of experimentally-verified positive and negative infection data, where genomes of all members are fully sequenced.
 Large-scale experimental screens for phage and bacteria infectious interactions that report high-confidence negative interactions (i.e., no infection) will provide more robust model training and improved model performance.
@@
 Second, although our analyses offer an informative proof of concept, this work was done retrospectively and relied on existing data up to seven years old.
-These archived datasets were limited by the technology and costs of the time <!--the MDA issue falls under this, in my mind, as issue of using old datsets.
-So integrated it here, also to bring down the focus a bit/not give it it's own paragraph-->.
-This resulted in small sequencing effort (as compared to today's dataset sizes) and thus datasets that were suboptimally powered for the statistical analyses we strive for today.
+These archived datasets were limited by the technology and costs of the time.
+This resulted in small sequencing effort (as compared to today's dataset sizes) and thus datasets that were sub-optimally powered for the statistical analyses we strive for today.
 Further, two studies, the diet and twin studies, relied on multiple displacement amplification (MDA) in their library preparations--an approach used to overcome the large nucleic acids requirements, especially of older sequencing library generation protocols.
-MDA results in significant biases in microbobial community composition <!--Yilmaz, S., Allgaier, M.
-& Hugenholtz, P., 2010, Multiple displacement amplification compromises quantitative analysis of metagenomes, Nat Methods, 7(12), pp.
-943-4-->, as well as toward ssDNA viral genomes [@Kim:2008to; @Kim:2011hp], thus rendering the resulting microbial and viral metagenomes non-quantitative.
-<!--because Yilmaz et al., show that the bias depends on the initial community structure, I think this means that an MDA bias would also compromise your ability to compare samples even within studies, so I removed this: "However, by comparing samples within studies, which were all treated the same, we were able to provide the fairest possible comparison for drawing conclusions."--> Future work that employs larger sequence datasets and that avoids the use of MDA will build on and validate our findings, as well as inform the design and interpretation of further studies.
+MDA results in significant biases in microbial community composition [@Yilmaz:2010jb], as well as toward ssDNA viral genomes [@Kim:2008to; @Kim:2011hp], thus rendering the resulting microbial and viral metagenomes non-quantitative.
+Future work that employs larger sequence datasets and that avoids the use of MDA will build on and validate our findings, as well as inform the design and interpretation of further studies.
 
@@
 This operationally defined approach allows us to study whole community networks, but limits our ability to make conclusions about interactions among specific phage or bacterial species or populations.
-Although this approach lacks the resolution for drawing conclusions about specific bacteria or phages, future work could begin addressing these questions through improved binning methods and deeper metagenomic shotgun sequencing, as well as an improved understanding of what defines ecologically and evolutionarily cohesive units for both phage and bacteria, e.g., Polz et al., 2006 <!--Polz, M.F., Hunt, D.E., Preheim, S.P.
-& Weinreich, D.M., 2006, Patterns and mechanisms of genetic and phenotypic differentiation in marine microbes, Philos Trans R Soc Lond B Biol Sci, 361(1475), pp.
-2009-21.
--->
+Although this approach lacks the resolution for drawing conclusions about specific bacteria or phages, future work could begin addressing these questions through improved binning methods and deeper metagenomic shotgun sequencing, as well as an improved understanding of what defines ecologically and evolutionarily cohesive units for both phage and bacteria, e.g., Polz et al., 2006 [@Polz:2006fi].
+
+Such work will include more sophisticated approaches to defining operational genomic units and their taxonomic underpinnings (e.g.
+OGUs defined as clustering at the genus or species level).
+Previous work has implemented more sophisticated clustering and validation approaches that allow for such refined operational genomic unit clustering, such as the phylogenomic analyses performed to cluster cyanophage isolate genomes into informative groups using shared gene content, average nucleotide identity of shared genes, and pairwise differences between genomes [@Gregory:2016cg].
+Similar clustering definition and validation methods, both computational and experimental, have been implemented in some metagenomic sequencing studies [@Roux:2016cc; @Brum:2015iaa; @Deng:2014eb; @Minot:2012ed] and could offer yet another level of sophistication to our network-based analyses, especially by allowing us to make conclusions about the specific virus ad bacterial interactions within the networks.
 
@@
 By focusing on relationships between bacterial and viral communities, they are studied as the interacting cohorts they are, rather than as independent entities.
-While our developed bacteria-phage interaction framework is a novel conceptual advance, the microbiome also consists of archaea and small eukaryotes, such as fungi and *Demodex* mites <!--reference?-->.
-Further, the components of the microbiome can interact with human immune cells and other non-microbial community members <!--reference?-->.
+While our developed bacteria-phage interaction framework is a novel conceptual advance, the microbiome also consists of archaea and small eukaryotes, including fungi and *Demodex* mites [@Hannigan:2013im; @Grice:2011gy].
+Further, the components of the microbiome can interact with human immune cells and other non-microbial community members [@Round:2009bz].
 Future work will build from our approach and include these additional community members and their diverse interactions and relationships (e.g., beyond phage-bacteria).
-This will result in a more robust network and a more holistic understanding of the human-associated microbiome and the evolutionary and ecological processes that drive its assembly and function <!--there were a lot of redunant thougths here I streamlined.
-Also you'll see I merged the final paragraphs to make the shared forward looking elements stronger.-->
+This will result in a more robust network and a more holistic understanding of the human-associated microbiome and the evolutionary and ecological processes that drive its assembly and function.
 
@@
 ## Data Availability
-All associated source code is available on GitHub at the following repository <!--don't know if you want to change this to reflect journal change?-->:
+All associated source code is available on GitHub at the following repository:
 
@@
 
-## Bacterial OGU Identification
-The longest contig from each OGU was used as its representative sequence for identification.
-Representative sequences were aligned to the European Nucleotide Archive (ENA) bacterial reference database using the blastn algorithm (e-value < $10^{-5}$).
-79% of the bacterial OGUs were able to be identified.
-<!--what was this ID used for?
-i don't recall any discussion of inferred cluster taxonomic identity-->
-
 ## Phage OGU Identification
-To confirm a lack of phage sequences in the bacterial OGU dataset, we performed a relatively strict <!--indeed, this is relative.
-can you relate it?
-What is the distribution of evalues or bitscores from this blastp?
-any anecdotes that it was a useful cut-off?--> blast nucleotide alignment of the bacterial OGU representative sequences using an e-value < $10^{-25}$.
+To confirm a lack of phage sequences in the bacterial OGU dataset, we performed blast nucleotide alignment of the bacterial OGU representative sequences using an e-value < $10^{-25}$, which was stricter than the $10^{-10}$ threshold used in the random forest model below.
 We used a stricter threshold because we know there are genomic similarities between bacteria and phage OGUs from the interactive model, but we were interested in contigs with high enough similarity to references that they may indeed be from phages.
 2% of the OGUs had nucleotide similarities to known bacteriophage genomes, although the alignments were short (a couple of kb) and represented a small fraction of the alignment sequences.
-<!-- shouldn't you also consider a relative length component?
-Meaning, contigs with only phage similarity along its entire length are more likely phage than contigs with only small portion similar to phage, while rest has similarity to microbes...thus more likley prophage and ok to stay classified as bacteria.
-Why not just run already created probabilitic models that use many criteria, like VirSorter?
-Then you get predictions of phage (cat 1-3) versus propahge (cat 4-6) and the confidence of the prediction.
--->
 

### 2017-05-30 doc/manuscript.Rmd
Added the comments.

@@
 
-Viruses and bacteria are critical components to the human microbiome and play important roles in health and disease.
-Bacterial communities have been associated with diseases including a range of skin conditions [@Hannigan:2013im], acute and chronic wound healing conditions [@Hannigan:2014be; @Loesche:2016ev], and gastrointestinal diseases including inflammatory bowel disease [@He:2016ch; @Norman:2015kb], *Clostridium difficile* infections [@Seekatz:2016fz], and colorectal cancer [@Zackular:2014fba; @Baxter:2014hb].
-Altered viromes (virus communities consisting primarily of bacteriophages) have been also associated with various diseases and perturbations including inflammatory bowel disease [@Norman:2015kb; @Manrique:2016dx], periodontal disease [@Ly:2014ew], spread of antibiotic resistance [@Modi:2013fia], and others [@Monaco:2016ita; @Hannigan:2015fz; @Minot:2011ez; @SantiagoRodriguez:2015gd; @Abeles:2015dy; @Abeles:2014kj].
-The viruses act in concert with their microbial hosts as a single ecological community [@Haerter:2014ii].
+Viruses and bacteria are critical components of the human microbiome and play important roles in health and disease.
+Bacterial communities have been associated with disease states, including a range of skin conditions [@Hannigan:2013im], acute and chronic wound healing conditions [@Hannigan:2014be; @Loesche:2016ev], and gastrointestinal diseases, such as inflammatory bowel disease [@He:2016ch; @Norman:2015kb], *Clostridium difficile* infections [@Seekatz:2016fz] and colorectal cancer [@Zackular:2014fba; @Baxter:2014hb].
+Altered human viromes (virus communities consisting primarily of bacteriophages) also have been  associated with diseases and perturbations, including inflammatory bowel disease [@Norman:2015kb; @Manrique:2016dx], periodontal disease [@Ly:2014ew], spread of antibiotic resistance [@Modi:2013fia], and others [@Monaco:2016ita; @Hannigan:2015fz; @Minot:2011ez; @SantiagoRodriguez:2015gd; @Abeles:2015dy; @Abeles:2014kj].
+Viruses act in concert with their microbial hosts as a single ecological community [@Haerter:2014ii].
 Viruses influence their living microbial host communities through processes including lysis, modulating host gene expression [@Lindell:2005gz, @Tyler:2013fl, @Hargreaves:2014ja], influencing evolutionary processes, such as horizontal gene transfer [@Moon:2015fa, @Modi:2013fi, @Ogg:1981th, @Frost:2005dn] or antagonistic co-evolution [@Koskella:2014ds], and by altering ecosystem processes and elemental stoichiometry [@Jover:2014gq].
@@
 Removal of bacteria or phage can disrupt or even collapse those ecosystems [@Haerter:2014ii; @Harcombe:2005fd; @Middelboe:2001fl; @Poisot:2011jc; @Thompson:2012ki; @Moebus:1981kp; @Flores:2013hc; @Poisot:2012fh; @Poisot071993; @Flores:2011bh; @Jover:2015ev].
-Relationship-based network approaches allow us to capture this interaction information.
-We therefore leveraged machine learning and graph theory techniques to characterize human bacterial and phage communities by their inferred relationships.
-By doing so, we provide a foundation for further studies of disease network dynamics and gain broader insights into human microbiome network diversity across different body sites.
+Studying such bacteria-phage interactions through community-wide networks built from inferred relationships could, e.g., offer insights into the drivers of human microbiome diversity across body sites and enable the study of human microbiome network dynamics overall.
+<!--This is not the right place to put a concluding clause (it is both illogical paragraph/idea flow, but also redundant to the closing sentence of the last paragraph...where concluding clauses fit well), I tried to morph it into more forward looking wording ("what do we need?").
+Then the next paragraph starts with "and here is what we provided towards what we need." -->
 
-Human bacterial-phage network diversity was established and explored by utilizing three published microbiome datasets that contained paired virus and whole community metagenomic sequence sets dominated by bacteria [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei].
-Our approach built off of previous large-scale phage-bacteria microbiome network analyses by inferring interactions using metagenomic datasets, instead of using culture-based techniques, which limit the scale of the experiments and analyses [@Flores:2013hc].
-This metagenomic interaction inference model built on previous models, which have utilized a variety of techniques including taxonomic-based co-occurrence models using linear relationships between bacteria and phages [@LimaMendez:2015hw], as well as nucleotide similarity (including CRISPR targeting) models using OGUs [@@Roux:2016cc] and whole genomes [@Edwards:2015iz].
-Our approach uniquely included protein interaction data, and was validated on experimentally determined negative and positive interactions, i.e., "who does and does not infect whom." Through this approach, we were able to provide a basic understanding of the network dynamics associated with phage and bacterial communities on the human body.
+In this study, human-associated bacterial and phage communities were characterized by their inferred relationships using three published paired virus and bacteria-dominated whole community metagenomic datasets [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei].
+Machine learning and graph theory techniques were leveraged to establish and explore the human bacteria-phage network diversity therein.
+This approach built upon previous large-scale phage-bacteria network analyses by inferring interactions from metagenomic datasets, rather than culture-dependent data [@Flores:2013hc], which is limited in the scale of possible experiments and analyses.
+Our metagenomic interaction inference model improves upon previous models of phage-host predictions that have utilized a variety of techniques, such as linear models to predict bacteria-phage co-occurence based on taxonomy assignments <!-- does this capture your understanding of their approach?--> [@LimaMendez:2015hw], and nucleotide similarity models applied to both whole virus genomes [@Edwards:2015iz] and related clusters of whole and partial virus genomes [@@Roux:2016cc]<!--neither CRISPRs nor OGUs had been introduced yet, so I simplified and took those out here.
+Otherwise, this is the type of overview I'd envisioned, thanks...there is no need to mention that these are in ocean systems, only focus on giving credit to the approaches and methods.-->.
+Our approach uniquely included protein interaction data and was validated based on experimentally determined positive and negative interactions, i.e., "who does and does not infect whom." Through this approach, we were able to provide a basic understanding of the network dynamics associated with phage and bacterial communities on and in the human body.
 By building and utilizing a microbiome network, we found that different people, body sites, and anatomical locations not only support distinct microbiome membership and diversity [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei; @Grice:2009eea; @Findley:2013jf; @Costello:2009im, @Consortium:2012iz], but also support ecological communities with distinct communication structures and propensities toward community instability.
-Through an improved understanding of the healthy state of network structures across the human body, we empower future studies to  investigate how these community structures are influenced by disease state and their overall impact on human health.
+Through an improved understanding of network structures across the human body, we empower future studies to investigate how these communities and interaction dynamics are influenced by disease state and the overall impact they may have on human health.
 
@@
 ## Cohort Curation and Sample Processing
-We studied the differences in virus-bacteria interaction networks across healthy human bodies by leveraging previously published sequence sets containing purified virome samples paired with bacterial metagenomes from whole metagenomic shotgun sequences.
+We studied the differences in virus-bacteria interaction networks across healthy human bodies by leveraging previously published shotgun sequence datasets of purified viral metagenomes (viromes) paired with bacteria-dominated whole community metagenomes.
 Our study contained three datasets that explored the impact of diet on the healthy human gut virome [@Minot:2011ez], the impact of anatomical location on the healthy human skin virome [@Hannigan:2015fz], and the viromes of monozygotic twins and their mothers [@Reyes:2010cwa; @Turnbaugh:2009ei].
-The twin and diet studies utilized multiple displacement amplification methods in their library preparations.
-These datasets were selected because they included virome samples subjected to virus-like particle (VLP) purification.
+<!--MDA reference not needed/relevant here, distracting-->These datasets were selected because their virome samples were subjected to virus-like particle (VLP) purification.
 To this end, they employed combinations of filtration, chloroform/DNase treatment, and cesium chloride gradients to eliminate organismal DNA and thereby allow for direct assessment of both the extracellular and fully-assembled intracellular virome **(Supplemental Figure \ref{SequenceStats} A-B)** [@Minot:2011ez, @Hannigan:2015fz, @Reyes:2010cwa; @Turnbaugh:2009ei].
@@
 Only bacteria and phages were used in the model.
-The training set contained 43 diverse bacterial species and 30 diverse phage strains, including both broad and specific infectious ranges **(Supplemental Figure \ref{ValidationOverview} A - B)**.
+The training set contained 43 diverse bacterial species and 30 diverse phage strains, including both broad and specific ranges of infection **(Supplemental Figure \ref{ValidationOverview} A - B)**.
 Phages with linear and circular genomes, as well as ssDNA and dsDNA genomes, were included in the analysis.
@@
 Four phage and bacterial genomic features were used to predict infectious relationships between bacteria and phages: 1) genome nucleotide similarities, 2) gene amino acid sequence similarities, 3) bacterial Clustered Regularly Interspaced Short Palindromic Repeat (CRISPR) spacer sequences that target phages, and 4) similarity of protein families associated with experimentally identified protein-protein interactions [@Orchard:2014hq].
-The resulting random forest model performed with an AUC of `r signif(genmodelper$ROC, digits = sigfig)`, a sensitivity of `r signif(genmodelper$Sens, digits = sigfig)`, and a specificity of `r signif(genmodelper$Spec, digits = sigfig)` **(Figure \ref{RocCurve} A)**.
+The resulting random forest model was assessed and the area under its receiver operating characteristic (ROC) curve was `r signif(genmodelper$ROC, digits = sigfig)`, the model sensitivity was `r signif(genmodelper$Sens, digits = sigfig)`, and specificity was `r signif(genmodelper$Spec, digits = sigfig)` **(Figure \ref{RocCurve} A)**.
 The most important predictor in the model was amino acid similarity between genes, followed by nucleotide similarity of whole genomes **(Figure \ref{RocCurve} B)**.
@@
 Metadata including study, sample ID, disease, and OGU abundance within the community were stored in the master network for downstream analysis **(Supplemental Figure \ref{NetworkDiagram})**.
-The master network was highly connected and contained `r prettyNum(genfigurestats[c(genfigurestats$class %in% "Total"),"Edges"], big.mark = ",")` infectious relationships among `r prettyNum(genfigurestats[c(genfigurestats$class %in% "Total"),"Vertices"], big.mark = ",")` nodes, `r prettyNum(nestats[c(nestats$cats %in% "PhageNodes"),"values"], big.mark = ",")` phages and `r prettyNum(nestats[c(nestats$cats %in% "BacteriaNodes"),"values"], big.mark = ",")` bacteria.
+The master network was highly connected and contained `r prettyNum(genfigurestats[c(genfigurestats$class %in% "Total"),"Edges"], big.mark = ",")` infectious relationships among `r prettyNum(genfigurestats[c(genfigurestats$class %in% "Total"),"Vertices"], big.mark = ",")` nodes, representing `r prettyNum(nestats[c(nestats$cats %in% "PhageNodes"),"values"], big.mark = ",")` phages and `r prettyNum(nestats[c(nestats$cats %in% "BacteriaNodes"),"values"], big.mark = ",")` bacteria.
 Although the network was highly connected, not all relationships were present in all samples.
@@
 We found evidence that diet was sufficient to alter gut microbiome network connectivity.
-Although our sample size was small, our findings provided evidence that high-fat diets were less connected than low-fat diets, and high-fat diets therefore may lead to less stable communities with a decreased ability for microbes to directly influence one another.
+Although our sample size was small, our findings provided evidence that high-fat diets were less connected than low-fat diets and that high-fat diets therefore may lead to less stable communities with a decreased ability for microbes to directly influence one another.
 We supported this finding with the observation that obesity may have been associated with decreased network connectivity.
-Together these findings suggest the food we eat may not only impact which microbes colonize our guts, but may also impact their infectious interactions.
+Together these findings suggest the food we eat may not only impact which microbes colonize our guts, but may also impact their  interactions with infecting phages.
 Further work will be required to characterize these relationships with a larger cohort.
 
-In addition to diet, the skin environment influenced the microbiome interaction network structure as well.
+In addition to diet, the skin environment also influenced the microbiome interaction network structure.
 Network structure differed between environmentally exposed and occluded skin sites.
@@
 Together these data suggested that body sites under greater degrees of fluctuation harbored less connected, potentially less stable microbiomes.
-This points to a link between microbiome and environmental stability, and warrants further investigation.
+This points to a link between microbiome and environmental stability and warrants further investigation.
 
@@
 First, as with most classification models, the infection classification model developed and applied is only as good as its training set -- in this case, the collection of experimentally-verified positive and negative infection data, where genomes of all members are fully sequenced.
-Large-scale experimental screens for phage and bacteria infectious interactions that report high-confidence negative interactions (i.e., no infection) will provide more robust model training and improved model performance.
-Furthermore, just as we have improved on previous modeling efforts, we expect that new and creative scoring metrics will be integrated into this model to improve performance.
+Large-scale experimental screens for phage and bacteria infectious interactions that report high-confidence negative interactions (i.e., no infection) are desperately needed, as they would provide more robust model training and improved model performance.
+Furthermore, just as we have improved on previous modeling efforts, we expect that new and creative scoring metrics will be integrated into this model to improve future performance.
 
@@
 These archived datasets were limited by the technology and costs of the time.
-This resulted in small sequencing effort (as compared to today's dataset sizes) and thus datasets that were sub-optimally powered for the statistical analyses we strive for today.
-Further, two studies, the diet and twin studies, relied on multiple displacement amplification (MDA) in their library preparations--an approach used to overcome the large nucleic acids requirements, especially of older sequencing library generation protocols.
-MDA results in significant biases in microbial community composition [@Yilmaz:2010jb], as well as toward ssDNA viral genomes [@Kim:2008to; @Kim:2011hp], thus rendering the resulting microbial and viral metagenomes non-quantitative.
-Future work that employs larger sequence datasets and that avoids the use of MDA will build on and validate our findings, as well as inform the design and interpretation of further studies.
+This resulted in small sequencing effort (as compared to today's dataset sizes) and thus datasets that were sub-optimally powered for statistical analyses.
+Further, two studies, the diet and twin studies, relied on multiple displacement amplification (MDA) in their library preparations--an approach used to overcome the large nucleic acids requirements typical of older sequencing library generation protocols.
+It is now known that MDA results in significant biases in microbial community composition [@Yilmaz:2010jb], as well as toward ssDNA viral genomes [@Kim:2008to; @Kim:2011hp], thus rendering the resulting microbial and viral metagenomes non-quantitative.
+Future work that employs larger sequence datasets and that avoids the use of bias-inducing amplification steps will build on and validate our findings, as well as inform the design and interpretation of further studies.
 
-Finally, the networks in this study were built using operational genomic units, which represent groups of highly similar bacteria or phage genomes or genome fragments as clustered sub-populations.
-This operationally defined approach allows us to study whole community networks, but limits our ability to make conclusions about interactions among specific phage or bacterial species or populations.
-Although this approach lacks the resolution for drawing conclusions about specific bacteria or phages, future work could begin addressing these questions through improved binning methods and deeper metagenomic shotgun sequencing, as well as an improved understanding of what defines ecologically and evolutionarily cohesive units for both phage and bacteria, e.g., Polz et al., 2006 [@Polz:2006fi].
-
-Such work will include more sophisticated approaches to defining operational genomic units and their taxonomic underpinnings (e.g.
-OGUs defined as clustering at the genus or species level).
-Previous work has implemented more sophisticated clustering and validation approaches that allow for such refined operational genomic unit clustering, such as the phylogenomic analyses performed to cluster cyanophage isolate genomes into informative groups using shared gene content, average nucleotide identity of shared genes, and pairwise differences between genomes [@Gregory:2016cg].
-Similar clustering definition and validation methods, both computational and experimental, have been implemented in some metagenomic sequencing studies [@Roux:2016cc; @Brum:2015iaa; @Deng:2014eb; @Minot:2012ed] and could offer yet another level of sophistication to our network-based analyses, especially by allowing us to make conclusions about the specific virus ad bacterial interactions within the networks.
+Finally, the networks in this study were built using operational genomic units (OGUs), which represented groups of highly similar bacteria or phage genomes or genome fragments as clustered sub-populations.
+Similar clustering definition and validation methods, both computational and experimental, have been implemented in other metagenomic sequencing studies, as well [@Minot:2012ed; @Deng:2014eb; @Brum:2015iaa; @Roux:2016cc].
+These approaches could offer yet another level of sophistication to our network-based analyses.
+While this operationally defined clustering approach allows us to study whole community networks, our ability to make conclusions about interactions among specific phage or bacterial species or populations is inherently limited.
+Future work must address this limitation, e.g., through improved binning methods and deeper metagenomic shotgun sequencing, but most importantly through an improved conceptual framing of what defines ecologically and evolutionarily cohesive units for both phage and bacteria, sensu Polz et al., 2006 [@Polz:2006fi].
+Defining operational genomic units and their taxonomic underpinnings (e.g., whether OGU clusters represent genera or species) is an active area of work critical to the utility of this approach.
+As a first step, phylogenomic analyses have been performed to cluster cyanophage isolate genomes into informative groups using shared gene content, average nucleotide identity of shared genes, and pairwise differences between genomes [@Gregory:2016cg].
+Such population-genetic assessment of phage evolution, coupled with the ecological implications of genome heterogeneity, will inform how to define nodes in future iterations of the ecological network developed here.
 
 Together our work takes an initial step towards defining bacteria-virus interaction profiles as a characteristic of human-associated microbial communities.
-This approach revealed the impacts different human environments (e.g., the skin and gut) can have on microbiome connectivity.
+This approach revealed the impacts that different human environments (e.g., the skin and gut) can have on microbiome connectivity.
 By focusing on relationships between bacterial and viral communities, they are studied as the interacting cohorts they are, rather than as independent entities.
-While our developed bacteria-phage interaction framework is a novel conceptual advance, the microbiome also consists of archaea and small eukaryotes, including fungi and *Demodex* mites [@Hannigan:2013im; @Grice:2011gy].
-Further, the components of the microbiome can interact with human immune cells and other non-microbial community members [@Round:2009bz].
+While our developed bacteria-phage interaction framework is a novel conceptual advance, the microbiome also consists of archaea and small eukaryotes, including fungi and *Demodex* mites [@Hannigan:2013im; @Grice:2011gy]--all of which can interact with human immune cells and other non-microbial community members [@Round:2009bz].
 Future work will build from our approach and include these additional community members and their diverse interactions and relationships (e.g., beyond phage-bacteria).
-This will result in a more robust network and a more holistic understanding of the human-associated microbiome and the evolutionary and ecological processes that drive its assembly and function.
+This will result in a more robust network and a more holistic understanding of the evolutionary and ecological processes that drive the assembly and function of the human-associated microbiome.
 

### 2017-05-30 doc/manuscript.Rmd
Updated with modifications, edits, and format for journal.

@@
 
-&nbsp;
-
 | ***Corresponding Author Information***
@@
 
-***Text Length***: 35,640 / 50,000 Characters
+***Abstract Length***: 236 / 250
+
+***Importance Length***: 145 / 150
+
+***Text Length***: 4,987 / 5,000 Words
 
@@
 We organized the predicted interactions into networks that allowed us to evaluate phage-bacteria connectedness across the human body.
-We found that gut and skin network structures were person-specific and were not conserved among cohabitating family members.
+We found that gut and skin network structures were person-specific and not conserved among cohabitating family members.
 High-fat diets and obesity were associated with less connected networks.
-There were significant differences in network structure between skin sites, with those exposed to the external environment being less connected and more prone to instability.
+Network structure differed between skin sites, with those exposed to the external environment being less connected and more prone to instability.
 This study quantified and contrasted the diversity of virome-microbiome networks across the human body and illustrated how environmental factors may influence phage-bacteria interactive dynamics.
@@
 
-**Word Count**: 243 / 250
+# Importance
+
+The human microbiome, the collection of microbial communities that colonize the human body, is a crucial component to health and disease.
+Two major components to the human microbiome are the bacterial and viral communities.
+These communities have primarily been studied separately using metrics of community composition and diversity.
+These approaches have failed to capture the complex dynamics of interacting bacteria and phage communities, which frequently share genetic information and work together to maintain stable ecosystems.
+Removal of bacteria or phage can disrupt or even collapse those ecosystems.
+Relationship-based network approaches allow us to capture this interaction information.
+Using this network-based approach with three independent human cohorts, we were able to present an initial understanding of how phage-bacteria networks differ throughout the human body, so as to provide a baseline for future studies of how and why microbiome networks differ in disease states.
 
@@
 Viruses act in concert with their microbial hosts as a single ecological community [@Haerter:2014ii].
-Viruses influence their living microbial host communities through processes including lysis, modulating host gene expression [@Lindell:2005gz, @Tyler:2013fl, @Hargreaves:2014ja], influencing evolutionary processes, such as horizontal gene transfer [@Moon:2015fa, @Modi:2013fi, @Ogg:1981th, @Frost:2005dn] or antagonistic co-evolution [@Koskella:2014ds], and by altering ecosystem processes and elemental stoichiometry [@Jover:2014gq].
+Viruses influence their living microbial host communities through processes including lysis, host gene expression modulation [@Lindell:2005gz, @Tyler:2013fl, @Hargreaves:2014ja], influence on evolutionary processes such as horizontal gene transfer [@Moon:2015fa, @Modi:2013fi, @Ogg:1981th, @Frost:2005dn] or antagonistic co-evolution [@Koskella:2014ds], and alteration of ecosystem processes and elemental stoichiometry [@Jover:2014gq].
 
 Previous human microbiome work has focused on bacterial and viral communities, but have reduced them to two separate communities by studying them independently  [@Norman:2015kb; @Manrique:2016dx; @Ly:2014ew; @Monaco:2016ita; @Hannigan:2015fz; @Minot:2011ez; @SantiagoRodriguez:2015gd; @Abeles:2015dy; @Abeles:2014kj].
-In reality, bacteria and phage communities are dynamic and complex.
-They frequently share genetic information and work together to maintain stable ecosystems.
+This approach fails to capture the complex dynamics of interacting bacteria and phage communities, which frequently share genetic information and work together to maintain stable ecosystems.
 Removal of bacteria or phage can disrupt or even collapse those ecosystems [@Haerter:2014ii; @Harcombe:2005fd; @Middelboe:2001fl; @Poisot:2011jc; @Thompson:2012ki; @Moebus:1981kp; @Flores:2013hc; @Poisot:2012fh; @Poisot071993; @Flores:2011bh; @Jover:2015ev].
-Studying such bacteria-phage interactions through community-wide networks built from inferred relationships could, e.g., offer insights into the drivers of human microbiome diversity across body sites and enable the study of human microbiome network dynamics overall.
-<!--This is not the right place to put a concluding clause (it is both illogical paragraph/idea flow, but also redundant to the closing sentence of the last paragraph...where concluding clauses fit well), I tried to morph it into more forward looking wording ("what do we need?").
-Then the next paragraph starts with "and here is what we provided towards what we need." -->
+Relationship-based network approaches allow us to capture this interaction information.
+Studying such bacteria-phage interactions through community-wide networks built from inferred relationships could offer further insights into the drivers of human microbiome diversity across body sites and enable the study of human microbiome network dynamics overall.
 
-In this study, human-associated bacterial and phage communities were characterized by their inferred relationships using three published paired virus and bacteria-dominated whole community metagenomic datasets [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei].
-Machine learning and graph theory techniques were leveraged to establish and explore the human bacteria-phage network diversity therein.
+In this study, we characterized human-associated bacterial and phage communities by their inferred relationships using three published paired virus and bacteria-dominated whole community metagenomic datasets [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei].
+We leveraged machine learning and graph theory techniques to establish and explore the human bacteria-phage network diversity therein.
 This approach built upon previous large-scale phage-bacteria network analyses by inferring interactions from metagenomic datasets, rather than culture-dependent data [@Flores:2013hc], which is limited in the scale of possible experiments and analyses.
-Our metagenomic interaction inference model improves upon previous models of phage-host predictions that have utilized a variety of techniques, such as linear models to predict bacteria-phage co-occurence based on taxonomy assignments <!-- does this capture your understanding of their approach?--> [@LimaMendez:2015hw], and nucleotide similarity models applied to both whole virus genomes [@Edwards:2015iz] and related clusters of whole and partial virus genomes [@@Roux:2016cc]<!--neither CRISPRs nor OGUs had been introduced yet, so I simplified and took those out here.
-Otherwise, this is the type of overview I'd envisioned, thanks...there is no need to mention that these are in ocean systems, only focus on giving credit to the approaches and methods.-->.
-Our approach uniquely included protein interaction data and was validated based on experimentally determined positive and negative interactions, i.e., "who does and does not infect whom." Through this approach, we were able to provide a basic understanding of the network dynamics associated with phage and bacterial communities on and in the human body.
+Our metagenomic interaction inference model improved upon previous models of phage-host predictions that have utilized a variety of techniques, such as linear models to predict bacteria-phage co-occurrence using taxonomic assignments [@LimaMendez:2015hw], and nucleotide similarity models that were applied to both whole virus genomes [@Edwards:2015iz] and related clusters of whole and partial virus genomes [@@Roux:2016cc].
+Our approach uniquely included protein interaction data and was validated based on experimentally determined positive and negative interactions (i.e.
+who does and does not infect whom).
+Through this approach we were able to provide a basic understanding of the network dynamics associated with phage and bacterial communities on and in the human body.
 By building and utilizing a microbiome network, we found that different people, body sites, and anatomical locations not only support distinct microbiome membership and diversity [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei; @Grice:2009eea; @Findley:2013jf; @Costello:2009im, @Consortium:2012iz], but also support ecological communities with distinct communication structures and propensities toward community instability.
-Through an improved understanding of network structures across the human body, we empower future studies to investigate how these communities and interaction dynamics are influenced by disease state and the overall impact they may have on human health.
+Through an improved understanding of network structures across the human body, we empower future studies to investigate how these communities dynamics are influenced by disease states and the overall impact they may have on human health.
 
@@
 Our study contained three datasets that explored the impact of diet on the healthy human gut virome [@Minot:2011ez], the impact of anatomical location on the healthy human skin virome [@Hannigan:2015fz], and the viromes of monozygotic twins and their mothers [@Reyes:2010cwa; @Turnbaugh:2009ei].
-<!--MDA reference not needed/relevant here, distracting-->These datasets were selected because their virome samples were subjected to virus-like particle (VLP) purification.
+We selected these datasets because their virome samples were subjected to virus-like particle (VLP) purification.
 To this end, they employed combinations of filtration, chloroform/DNase treatment, and cesium chloride gradients to eliminate organismal DNA and thereby allow for direct assessment of both the extracellular and fully-assembled intracellular virome **(Supplemental Figure \ref{SequenceStats} A-B)** [@Minot:2011ez, @Hannigan:2015fz, @Reyes:2010cwa; @Turnbaugh:2009ei].
-While the whole metagenomic shotgun sequence samples were not subjected to purification, they primarily consisted of bacteria [@Minot:2011ez, @Hannigan:2015fz, @Reyes:2010cwa; @Turnbaugh:2009ei].
+While the whole metagenomic shotgun sequence samples were not subjected to purification, they primarily consisted of bacteria [@Minot:2011ez; @Hannigan:2015fz; @Reyes:2010cwa; @Turnbaugh:2009ei].
 
@@
 
-Four phage and bacterial genomic features were used to predict infectious relationships between bacteria and phages: 1) genome nucleotide similarities, 2) gene amino acid sequence similarities, 3) bacterial Clustered Regularly Interspaced Short Palindromic Repeat (CRISPR) spacer sequences that target phages, and 4) similarity of protein families associated with experimentally identified protein-protein interactions [@Orchard:2014hq].
+Four phage and bacterial genomic features were used in a random forest model to predict infectious relationships between bacteria and phages: 1) genome nucleotide similarities, 2) gene amino acid sequence similarities, 3) bacterial Clustered Regularly Interspaced Short Palindromic Repeat (CRISPR) spacer sequences that target phages, and 4) similarity of protein families associated with experimentally identified protein-protein interactions [@Orchard:2014hq].
 The resulting random forest model was assessed and the area under its receiver operating characteristic (ROC) curve was `r signif(genmodelper$ROC, digits = sigfig)`, the model sensitivity was `r signif(genmodelper$Sens, digits = sigfig)`, and specificity was `r signif(genmodelper$Spec, digits = sigfig)` **(Figure \ref{RocCurve} A)**.
 The most important predictor in the model was amino acid similarity between genes, followed by nucleotide similarity of whole genomes **(Figure \ref{RocCurve} B)**.
-Protein family interactions were moderately important to the model, and CRISPRs were largely uninformative, due to the lack of identifiable CRISPRs in the dataset and their redundancy with the nucleotide similarity methods **(Figure \ref{RocCurve} B)**.
+Protein family interactions were moderately important to the model, and CRISPRs were largely uninformative, due to the minimal amount of identifiable CRISPRs in the dataset and their redundancy with the nucleotide similarity methods **(Figure \ref{RocCurve} B)**.
 Approximately one third of the training set relationships yielded no score and therefore were unable to be assigned an interaction prediction **(Figure \ref{RocCurve} C)**.
 
-We used the random forest model to classify the relationships between bacteria and phage operational genomic units, which were then used to build the interactive network.
+We used our random forest model to classify the relationships between bacteria and phage operational genomic units, which were then used to build the interactive network.
 The master network contained the three studies as sub-networks, which themselves each contained sub-networks for each sample **(Figure \ref{RocCurve} D)**.
-Metadata including study, sample ID, disease, and OGU abundance within the community were stored in the master network for downstream analysis **(Supplemental Figure \ref{NetworkDiagram})**.
+Metadata including study, sample ID, disease, and OGU abundance within the community were stored in the master network for parsing in downstream analyses **(Supplemental Figure \ref{NetworkDiagram})**.
 The master network was highly connected and contained `r prettyNum(genfigurestats[c(genfigurestats$class %in% "Total"),"Edges"], big.mark = ",")` infectious relationships among `r prettyNum(genfigurestats[c(genfigurestats$class %in% "Total"),"Vertices"], big.mark = ",")` nodes, representing `r prettyNum(nestats[c(nestats$cats %in% "PhageNodes"),"values"], big.mark = ",")` phages and `r prettyNum(nestats[c(nestats$cats %in% "BacteriaNodes"),"values"], big.mark = ",")` bacteria.
@@
 Intermittently moist skin sites (dynamic sites that fluctuate between being moist and dry) were significantly less connected than the more stable moist and sebaceous environments (p-value < 0.001, **Figure \ref{skinnetwork} A)**.
-Also, skin sites that were protected from the environment (occluded) were much more highly connected than those that were constantly exposed to the environment or only intermittently occluded (p-value < 0.001, **Figure \ref{skinnetwork} B)**.
+Also, skin sites that were occluded from the environment were much more highly connected than those that were constantly exposed to the environment or only intermittently occluded (p-value < 0.001, **Figure \ref{skinnetwork} B)**.
 
@@
 Foundational work has provided a baseline understanding of the human microbiome by characterizing bacterial and viral diversity across the human body [@Grice:2009eea; @Findley:2013jf; @Hannigan:2015fz; @Costello:2009im, @Consortium:2012iz; @Schloss:2005hz; @Minot:2011ez].
-Here, we offer an approach to describe how phage-bacteria networks differ throughout the human body to provide a baseline for future studies of how and why microbiome networks differ in disease states.
+Here, we offer an initial understanding of how phage-bacteria networks differ throughout the human body, so as to provide a baseline for future studies of how and why microbiome networks differ in disease states.
 We developed and implemented a network-based analytical model to evaluate the basic properties of the human microbiome through bacteria and phage relationships, instead of membership or diversity alone.
@@
 Gut network structure was not conserved among family members.
-These findings suggested that the community properties inferred from microbiome interaction network structures, such as stability, the potential for horizontal gene transfer between members, co-evolution of populations, etc., were person-specific.
+These findings suggested that the community properties inferred from microbiome interaction network structures, such as stability, the potential for horizontal gene transfer between members, and co-evolution of populations, were person-specific.
 These properties may be impacted by personal factors ranging from the body's immune system to external environmental conditions, such as climate and diet.
@@
 While this operationally defined clustering approach allows us to study whole community networks, our ability to make conclusions about interactions among specific phage or bacterial species or populations is inherently limited.
-Future work must address this limitation, e.g., through improved binning methods and deeper metagenomic shotgun sequencing, but most importantly through an improved conceptual framing of what defines ecologically and evolutionarily cohesive units for both phage and bacteria, sensu Polz et al., 2006 [@Polz:2006fi].
+Future work must address this limitation, e.g., through improved binning methods and deeper metagenomic shotgun sequencing, but most importantly through an improved conceptual framing of what defines ecologically and evolutionarily cohesive units for both phage and bacteria [@Polz:2006fi].
 Defining operational genomic units and their taxonomic underpinnings (e.g., whether OGU clusters represent genera or species) is an active area of work critical to the utility of this approach.
@@
 
-https://github.com/SchlossLab/Hannigan_ConjunctisViribus_GenRes_2017
+https://github.com/SchlossLab/Hannigan_ConjunctisViribus_mSystems_2017
 
@@
 The resulting bitscore of each alignment was recorded for use in our classification model.
-<!-- I get suspicious whenever e-value is used as a threshold.
-Do you have anecdotal evidence this a is a good criterion?
-Should be easy to check-->
 
@@
 We thank the members of the Schloss lab for their underlying contributions.
+
+# Funding Information
 GDH was supported in part by the Molecular Mechanisms in Microbial Pathogenesis Training Program (T32 AI007528).

### 2017-05-31 doc/manuscript.Rmd
Removed internal review statement.
Final fixes.

@@
 ---
-title: Biogeography and Environmental Conditions Shape Phage and Bacteria Interaction Networks Across the Healthy Human Microbiome
+title: Biogeography & Environmental Conditions Shape Phage & Bacteria Interaction Networks Across the Human Microbiome
 author:
@@
   - code: 2
-    address: Department of Ecology and Evolutionary Biology, University of Michigan, Ann Arbor, Michigan, 48109
+    address: Department of Ecology & Evolutionary Biology, University of Michigan, Ann Arbor, Michigan, 48109
   - code: 3
@@
     variant: markdown
-fontsize: 11pt
+fontsize: 12pt
 mainfont: 'Helvetica'
@@
 &nbsp;
+
+&nbsp;
+
+&nbsp;
+
+***Running Title***: Network Diversity of the Healthy Human Microbiome
 
@@
 | Email: pschloss@umich.edu
-
-***Running Title***: Network Diversity of the Healthy Human Microbiome
-
-***Journal***: mSystems
-
-***Keywords***: Virome, Microbiome, Graph Theory, Machine Learning
-
-***Abstract Length***: 236 / 250
-
-***Importance Length***: 145 / 150
-
-***Text Length***: 4,987 / 5,000 Words
-
-\* *Figures at the end of the document for internal review only.*
 
@@
 This work provides a baseline for future studies to better understand system perturbations, such as disease states, through ecological networks.
+
+\newpage
 
@@
 The gut network structures were not more similar within families (twins and their mothers; intrafamily) compared to other families (inter-family) (p-value = `r signif(interstats[c(interstats$site %in% "Twins"), "prob"], digits = sigfig)`, **Figure \ref{intradiv} C**).
-
-Skin microbiome network structure was strongly conserved within individuals (p-value < 0.001, **Figure \ref{intradiv} D**).
+In addition to the gut, skin microbiome network structure was strongly conserved within individuals (p-value < 0.001, **Figure \ref{intradiv} D**).
 This distribution was similar when separated by anatomical sites.
@@
 
-Second, although our analyses offer an informative proof of concept, this work was done retrospectively and relied on existing data up to seven years old.
+Second, although our analyses utilized the best datasets currently avilable for our study, this work was done retrospectively and relied on existing data up to seven years old.
 These archived datasets were limited by the technology and costs of the time.
-This resulted in small sequencing effort (as compared to today's dataset sizes) and thus datasets that were sub-optimally powered for statistical analyses.
-Further, two studies, the diet and twin studies, relied on multiple displacement amplification (MDA) in their library preparations--an approach used to overcome the large nucleic acids requirements typical of older sequencing library generation protocols.
-It is now known that MDA results in significant biases in microbial community composition [@Yilmaz:2010jb], as well as toward ssDNA viral genomes [@Kim:2008to; @Kim:2011hp], thus rendering the resulting microbial and viral metagenomes non-quantitative.
+For example, the diet and twin studies, relied on multiple displacement amplification (MDA) in their library preparations--an approach used to overcome the large nucleic acids requirements typical of older sequencing library generation protocols.
+It is now known that MDA results in biases in microbial community composition [@Yilmaz:2010jb], as well as toward ssDNA viral genomes [@Kim:2008to; @Kim:2011hp], thus rendering the resulting microbial and viral metagenomes largely non-quantitative.
 Future work that employs larger sequence datasets and that avoids the use of bias-inducing amplification steps will build on and validate our findings, as well as inform the design and interpretation of further studies.
@@
 
-## Data Availability
-All associated source code is available on GitHub at the following repository:
+## Code Availability
+A reproducible version of this manuscript written in R markdown and all of the code used to obtain and process the sequencing data is available at the following GitHub repository:
 
@@
 This network includes information from all three published studies.
-(E) Network diameter (measure of graph size; the greatest number of traversed vertices required between two vertices), (F) number of vertices, and (G) number of edges (relationships) for the total network (yellow) and the individual study sub-networks (diet study = red, skin study = green, twin study = orange).* \label{RocCurve}](../figures/rocCurves.pdf){ width=90% }
+(E) Network diameter (measure of graph size; the greatest number of traversed vertices required between two vertices), (F) number of vertices, and (G) number of edges (relationships) for the total network (yellow) and the individual study sub-networks (diet study = red, skin study = green, twin study = orange).* \label{RocCurve}](../figures/rocCurves.pdf){ width=85% }
 

### 2017-07-26 doc/manuscript.Rmd
A lot of manuscript updated.

@@
 nestats <- read.delim(file = "../rtables/nestats.tsv", header = TRUE, sep = "\t")
+sitestats <- read.delim(file = "../rtables/sitestattable.tsv", header = TRUE, sep = "\t")
 genmodelper <- read.delim(file = "../rtables/genmodelper.tsv", header = TRUE, sep = "\t")
@@
 We found that gut and skin network structures were person-specific and not conserved among cohabitating family members.
-High-fat diets and obesity were associated with less connected networks.
+High-fat diets and obesity appeared to be associated with less connected networks.
 Network structure differed between skin sites, with those exposed to the external environment being less connected and more prone to instability.
@@
 This approach fails to capture the complex dynamics of interacting bacteria and phage communities, which frequently share genetic information and work together to maintain stable ecosystems.
-Removal of bacteria or phage can disrupt or even collapse those ecosystems [@Haerter:2014ii; @Harcombe:2005fd; @Middelboe:2001fl; @Poisot:2011jc; @Thompson:2012ki; @Moebus:1981kp; @Flores:2013hc; @Poisot:2012fh; @Poisot071993; @Flores:2011bh; @Jover:2015ev].
+Removal of bacteria or phages can disrupt or even collapse those ecosystems [@Haerter:2014ii; @Harcombe:2005fd; @Middelboe:2001fl; @Poisot:2011jc; @Thompson:2012ki; @Moebus:1981kp; @Flores:2013hc; @Poisot:2012fh; @Poisot071993; @Flores:2011bh; @Jover:2015ev].
 Relationship-based network approaches allow us to capture this interaction information.
@@
 Our study contained three datasets that explored the impact of diet on the healthy human gut virome [@Minot:2011ez], the impact of anatomical location on the healthy human skin virome [@Hannigan:2015fz], and the viromes of monozygotic twins and their mothers [@Reyes:2010cwa; @Turnbaugh:2009ei].
-We selected these datasets because their virome samples were subjected to virus-like particle (VLP) purification.
-To this end, they employed combinations of filtration, chloroform/DNase treatment, and cesium chloride gradients to eliminate organismal DNA and thereby allow for direct assessment of both the extracellular and fully-assembled intracellular virome **(Supplemental Figure \ref{SequenceStats} A-B)** [@Minot:2011ez, @Hannigan:2015fz, @Reyes:2010cwa; @Turnbaugh:2009ei].
-While the whole metagenomic shotgun sequence samples were not subjected to purification, they primarily consisted of bacteria [@Minot:2011ez; @Hannigan:2015fz; @Reyes:2010cwa; @Turnbaugh:2009ei].
+We selected these datasets because their virome samples were subjected to virus-like particle (VLP) purification, which removed contaminating DNA from human cells, bacteria, etc.
+To this end, the publishing authors employed combinations of filtration, chloroform/DNase treatment, and cesium chloride gradients to eliminate organismal DNA (bacteria, human, fungal, etc) and thereby allow for direct assessment of both the extracellular and fully-assembled intracellular virome **(Supplemental Figure \ref{SequenceStats} A-B)** [@Minot:2011ez, @Hannigan:2015fz, @Reyes:2010cwa; @Turnbaugh:2009ei].
+Each research group reported quality control measures to ensure the purity of the virome sequence datasets, using both computational and molecular techniques, including, for example, 16S rRNA gene qPCR **(Table S1)**.
+These reports confirmed that the virome libraries consisted of highly purified virus genomic DNA.
+We supplemented these findings and found that, in light of the rigorous purification and quality control during sample collection and preparation, 77% (228 / 298 operational genomic units) still had some nucleotide similarity to a given bacterial reference genome (e-value < $10^{-25}$).
+These findings agree with the previous reports from which we obtained our data, in that the majority of the gut and skin bacteriophages were temperate and shared elements with bacterial reference genomes, primarily because their replication cycles include bacterial genome integration [@Minot:2011ez, @Hannigan:2015fz, @Reyes:2010cwa].
+We additionally confidently identified two OGUs as being complete phages using the stringent Virsorter phage identification algorithm (class 1 confidence group) [@Roux:2015dh].
+
+While the whole metagenomic shotgun sequence samples were not subjected to purification, they primarily consisted of bacteria, quantified with an average viral relative abundance of 0.4% **(Table S1)** [@Minot:2011ez; @Hannigan:2015fz; @Reyes:2010cwa; @Turnbaugh:2009ei].
+We again supplemented these findings and found that only 2% (6 / 280 operational genomic units) of bacteria had significantly strong nucleotide similarity to phage reference genomes (e-value < $10^{-25}$) [@Minot:2011ez; @Hannigan:2015fz; @Reyes:2010cwa; @Turnbaugh:2009ei].
+No OGUs were confidently identified as lytic or temperate phage OGUs in the bacterial dataset using the Virsorter algorithm [@Roux:2015dh].
+Not only does this suggest a low level of "contamination", but also highlights future needs to more deeply sequence the bacterial metagenomes, whose genomes are order of magnitude longer than phages and other viruses.
 
@@
 The training set contained 43 diverse bacterial species and 30 diverse phage strains, including both broad and specific ranges of infection **(Supplemental Figure \ref{ValidationOverview} A - B)**.
+While it is true that there are more known phages that infect bacteria, we were limited by the information confirming which phages do not infect certain bacteria and attempted to keep the numbers of positive and negative interactions similar.
 Phages with linear and circular genomes, as well as ssDNA and dsDNA genomes, were included in the analysis.
@@
 Metadata including study, sample ID, disease, and OGU abundance within the community were stored in the master network for parsing in downstream analyses **(Supplemental Figure \ref{NetworkDiagram})**.
+Bacterial and phage relative abundance was recorded in each sample for each OGU, and the weight of the edge connecting those OGUs was calculated as a function of those relative abundance values.
+The separate extraction of the phage and bacterial libraries ensured a more accurate measurement of the microbial communities, as has been outlined previously [@Kleiner:2015kd; @Meisel:2016iu].
 The master network was highly connected and contained `r prettyNum(genfigurestats[c(genfigurestats$class %in% "Total"),"Edges"], big.mark = ",")` infectious relationships among `r prettyNum(genfigurestats[c(genfigurestats$class %in% "Total"),"Vertices"], big.mark = ",")` nodes, representing `r prettyNum(nestats[c(nestats$cats %in% "PhageNodes"),"values"], big.mark = ",")` phages and `r prettyNum(nestats[c(nestats$cats %in% "BacteriaNodes"),"values"], big.mark = ",")` bacteria.
@@
 As relationships were weighted by the relative abundances of their associated bacteria and phages, lowly abundant relationships could be present but not highly abundant.
-Like the master network, the skin network exhibited a diameter of `r prettyNum(genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Diameter"], big.mark = ",")` (measure of graph size; the greatest number of traversed vertices required between two vertices) and included `r round(100 * genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Vertices"] / genfigurestats[c(genfigurestats$class %in% "Total"),"Vertices"], digits = 1)`% and `r round(100 * genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Edges"] / genfigurestats[c(genfigurestats$class %in% "Total"),"Edges"], digits = 1)`% of the master network nodes and edges, respectively **(Figure \ref{RocCurve} E - F)**.
-The phages and bacteria in the gut diet and twin sample sets were more sparsely related: each contained fewer than 150 vertices, fewer than 20,000 relationships, and diameters of `r prettyNum(genfigurestats[c(genfigurestats$class %in% "DietStudy"),"Diameter"], big.mark = ",")` **(Figure \ref{RocCurve} E - F)**.
+Like the master network, the skin network exhibited a diameter of `r prettyNum(genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Diameter"], big.mark = ",")` (measure of graph size; the greatest number of traversed vertices required between two vertices) and included `r prettyNum(genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Vertices"], big.mark = ",")` (`r prettyNum(sitestats[c(sitestats$class %in% "skin"),"phagenodes"], big.mark = ",")` phages, `r prettyNum(sitestats[c(sitestats$class %in% "skin"),"bacterianodes"], big.mark = ",")` bacteria, `r round(100 * genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Vertices"] / genfigurestats[c(genfigurestats$class %in% "Total"),"Vertices"], digits = 1)`% total) and `r prettyNum(genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Edges"], big.mark = ",")` (`r round(100 * genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Edges"] / genfigurestats[c(genfigurestats$class %in% "Total"),"Edges"], digits = 1)`%) of the master network nodes and edges, respectively **(Figure \ref{RocCurve} E - F)**.
+The phages and bacteria in the gut diet and twin sample sets were more sparsely related, with the diet study consisting of `r prettyNum(genfigurestats[c(genfigurestats$class %in% "DietStudy"),"Vertices"], big.mark = ",")` (`r prettyNum(sitestats[c(sitestats$class %in% "diet"),"phagenodes"], big.mark = ",")` phages, `r prettyNum(sitestats[c(sitestats$class %in% "diet"),"bacterianodes"], big.mark = ",")` bacteria) nodes and `r prettyNum(genfigurestats[c(genfigurestats$class %in% "DietStudy"),"Edges"], big.mark = ",")` relationships, and the twin study containing `r prettyNum(genfigurestats[c(genfigurestats$class %in% "TwinStudy"),"Vertices"], big.mark = ",")` (`r prettyNum(sitestats[c(sitestats$class %in% "twin"),"phagenodes"], big.mark = ",")` phages, `r prettyNum(sitestats[c(sitestats$class %in% "twin"),"bacterianodes"], big.mark = ",")` bacteria) nodes and `r prettyNum(genfigurestats[c(genfigurestats$class %in% "TwinStudy"),"Edges"], big.mark = ",")` relationships **(Figure \ref{RocCurve} E - F)**.
 
@@
 A network with higher average closeness centrality also indicates an overall greater degree of connections, which suggests a greater resilience against instability.
+This is because more highly connected networks are more stable as a result of pathway dependencies and the unlikelihood that a randomly removed bacteria or phage would cause major divisions across the network [@Jeong:2000im; @Poisot:2011jc].
 We used this information to calculate the average connectedness per sample, which was corrected for the maximum potential degree of connectedness.
+Unfortunately our dataset was insufficiently powered to make strong conclusions toward this hypothesis, but the data still provides an interesting "case observation" that warrants further investigation.
 
-We found that the gut microbiome network structures associated with high-fat diets were less connected than those of low-fat diets **(Figure \ref{dietnetworks} A-B)**.
-Tests for statistical differences were not performed due to the small sample size.
-High-fat diets exhibited reduced degree centrality **(Figure \ref{dietnetworks} A)**, suggesting bacteria in high-fat environments were targeted by fewer phages and that phage tropism was more restricted.
-High-fat diets also exhibited decreased closeness centrality **(Figure \ref{dietnetworks} B)**, indicating that bacteria and phages were more distant from other bacteria and phages in the community.
+Using our limited sample set, we observed that the gut microbiome network structures associated with high-fat diets appeared less connected than those of low-fat diets, although a greater sample size will be required to more properly evaluate this trend **(Figure \ref{dietnetworks} A-B)**.
+Five subjects were available for use, all of which had matching bacteria and virome datasets and samples from 8-10 days following the initiation of their diets.
+High-fat diets appeared to exhibit reduced degree centrality **(Figure \ref{dietnetworks} A)**, suggesting bacteria in high-fat environments were targeted by fewer phages and that phage tropism was more restricted.
+High-fat diets also appeared to exhibit decreased closeness centrality **(Figure \ref{dietnetworks} B)**, indicating that bacteria and phages were more distant from other bacteria and phages in the community.
 This would make genetic transfer and altered abundance of a given phage or bacterium less capable of impacting other bacteria and phages within the network.
 
-In addition to diet, obesity was found to influence network structure.
-Obesity-associated networks demonstrated a higher degree centrality **(Figure \ref{dietnetworks} C)**, but less closeness centrality than the healthy-associated networks **(Figure \ref{dietnetworks} D)**.
+In addition to diet, we found preliminary evidence that obesity influenced network structure.
+This was done using the three mother samples available from the twin sample set, all of which had matching bacteria and phage samples and confirmed BMI information.
+The obesity-associated network appeared to have a higher degree centrality **(Figure \ref{dietnetworks} C)**, but less closeness centrality than the healthy-associated networks **(Figure \ref{dietnetworks} D)**.
 These results suggested that the obesity-associated networks are less connected, having microbes further from all other microbes within the community.
+This again comes with the caveat that this is a case observation with too small of a sample size to make more substantial claims.
+This is however an area for future investigation.
 
@@
 Network dissimilarity within each person over the 8-10 day sampling period was less than the average dissimilarity between that person and others, although this difference was not statistically significant (p-value = `r signif(interstats[c(interstats$site %in% "Diet"), "prob"], digits = sigfig)`, **Figure \ref{intradiv} B**).
-The lack of statistical confidence was likely due to the small sample size of this dataset.
+Four of the five available subjects were used here because one of the subjects was not sampled at the initial time point.
+The lack of statistical confidence was again due to the small sample size of this dataset.
+
 Although there was evidence for gut network conservation among individuals, we found no evidence for conservation of gut network structures within families.
-The gut network structures were not more similar within families (twins and their mothers; intrafamily) compared to other families (inter-family) (p-value = `r signif(interstats[c(interstats$site %in% "Twins"), "prob"], digits = sigfig)`, **Figure \ref{intradiv} C**).
+The gut network structures were not more similar within families (twins and their mothers; intrafamily) compared to other families (other twins and mothers; inter-family) (p-value = `r signif(interstats[c(interstats$site %in% "Twins"), "prob"], digits = sigfig)`, **Figure \ref{intradiv} C**).
 In addition to the gut, skin microbiome network structure was strongly conserved within individuals (p-value < 0.001, **Figure \ref{intradiv} D**).
@@
 Gut network structure was not conserved among family members.
-These findings suggested that the community properties inferred from microbiome interaction network structures, such as stability, the potential for horizontal gene transfer between members, and co-evolution of populations, were person-specific.
+These findings suggested that the community properties inferred from microbiome interaction network structures, such as stability (meaning a more highly connected network is more stable because a randomly removed bacteria or phage node is less likely to divide or disintegrate [@Jeong:2000im; @Poisot:2011jc] the overall network), the potential for horizontal gene transfer between members, and co-evolution of populations, were person-specific.
 These properties may be impacted by personal factors ranging from the body's immune system to external environmental conditions, such as climate and diet.
 
-The ability of environmental conditions to shape gut and skin microbiome interaction network structure was further supported by our finding that diet and skin location were associated with altered network structures.
-We found evidence that diet was sufficient to alter gut microbiome network connectivity.
-Although our sample size was small, our findings provided evidence that high-fat diets were less connected than low-fat diets and that high-fat diets therefore may lead to less stable communities with a decreased ability for microbes to directly influence one another.
+We observed evidence supporting the ability of environmental conditions to shape gut and skin microbiome interaction network structure by observing that diet and skin location were associated with altered network structures.
+We found evidence that diet was sufficient to alter gut microbiome network connectivity, although this needs to be interpreted as a case observation, due to the small sample size.
+Although our sample size was small, our findings provided some preliminary evidence that high-fat diets were less connected than low-fat diets and that high-fat diets therefore may lead to less stable communities with a decreased ability for microbes to directly influence one another.
 We supported this finding with the observation that obesity may have been associated with decreased network connectivity.
@@
 
-Second, although our analyses utilized the best datasets currently avilable for our study, this work was done retrospectively and relied on existing data up to seven years old.
+Second, although our analyses utilized the best datasets currently available for our study, this work was done retrospectively and relied on existing data up to seven years old.
 These archived datasets were limited by the technology and costs of the time.
@@
 These approaches could offer yet another level of sophistication to our network-based analyses.
-While this operationally defined clustering approach allows us to study whole community networks, our ability to make conclusions about interactions among specific phage or bacterial species or populations is inherently limited.
+While this operationally defined clustering approach allows us to study whole community networks, our ability to make conclusions about interactions among specific phage or bacterial species or populations is inherently limited, compared to more focused, culture-based studies such as the work by Malki *et al* [@Malki:2015dw].
 Future work must address this limitation, e.g., through improved binning methods and deeper metagenomic shotgun sequencing, but most importantly through an improved conceptual framing of what defines ecologically and evolutionarily cohesive units for both phage and bacteria [@Polz:2006fi].
@@
 Such population-genetic assessment of phage evolution, coupled with the ecological implications of genome heterogeneity, will inform how to define nodes in future iterations of the ecological network developed here.
+Even though we are hesitant to speculate on phage host ranges at low taxonomic levels in our dataset, the data does aggree with previous reports of instances of broad phage host range [@Malki:2015dw; @PaezEspino:2016gi].
 
@@
 
-## Phage OGU Identification
-To confirm a lack of phage sequences in the bacterial OGU dataset, we performed blast nucleotide alignment of the bacterial OGU representative sequences using an e-value < $10^{-25}$, which was stricter than the $10^{-10}$ threshold used in the random forest model below.
+## Operational Genomic Unit Identification
+To confirm a lack of phage sequences in the bacterial OGU dataset, we performed blast nucleotide alignment of the bacterial OGU representative sequences using an e-value < $10^{-25}$, which was stricter than the $10^{-10}$ threshold used in the random forest model below, against all of the phage reference genomes available in the EMBL database.
 We used a stricter threshold because we know there are genomic similarities between bacteria and phage OGUs from the interactive model, but we were interested in contigs with high enough similarity to references that they may indeed be from phages.
-2% of the OGUs had nucleotide similarities to known bacteriophage genomes, although the alignments were short (a couple of kb) and represented a small fraction of the alignment sequences.
+We also performed the converse analysis of aligning phage OGU representative sequences to EMBL bacterial reference genomes.
+Finally, we ran both the phage and bacteria OGU representative sequences through the Virsorter program (1.0.3) to identify phages (all default parameters were used), using only those in the high confidence identification category "class 1" [@Roux:2015dh].
 
@@
 We thank the members of the Schloss lab for their underlying contributions.
+We thank the authors of the original studies for making their data and metadata publicly available and understandable.
+We also thank the participants in the studies.
 
@@
 Pfam) is excluded.
+Features include the local gene alignments between bacteria and phage genes (denoted `blastx`; the blastx algorithm in Diamond aligner), local genome nucleotide alignments between bacteria and phage OGUs, presence of experimentally validated protein family domains (Pfams) between phage and bacteria OGUs, and CRISPR targeting of bacteria toward phages (CRISPR).
 (C) Proportions of samples included (gray) and excluded (red) in the model.
 Samples were excluded from the model because they did not yield any scores.
-Those interactions without scores were defined as not having interactions.
+Those interactions without scores were automatically classified as not having interactions.
 (D) Bipartite visualization of the resulting phage-bacteria network.
+Phage OGUs are presented in orange, bacteria OGUs in red, and their interaction edges are represented as connecting lines.
 This network includes information from all three published studies.
-(E) Network diameter (measure of graph size; the greatest number of traversed vertices required between two vertices), (F) number of vertices, and (G) number of edges (relationships) for the total network (yellow) and the individual study sub-networks (diet study = red, skin study = green, twin study = orange).* \label{RocCurve}](../figures/rocCurves.pdf){ width=85% }
+(E) Network diameter (measure of graph size; the greatest number of traversed vertices required between two vertices), (F) number of vertices, and (G) number of edges (relationships) for the total network (orange) and the individual study sub-networks (diet study = red, skin study = yellow, twin study = green).* \label{RocCurve}](../figures/rocCurves.pdf){ width=85% }
 
@@
 ![**Impact of Diet and Obesity on Gut Network Structure.** *(A) Quantification of average degree centrality (number of edges per node) and (B) closeness centrality (average distance from each node to every other node) of gut microbiome networks of subjects limited to exclusively high-fat or low-fat diets.
-Lines represent the mean degree of centrality for each diet.
-(C) Quantification of average degree centrality and (D) closeness centrality between obese and healthy adult women.*\label{dietnetworks}](../figures/dietnetworks.pdf){ width=90% }
+Each point represents the centrality from a human subject stool sample that was collected 8-10 days following the beginning of their defined diet.
+There are five samples here, compared to the four in figure \ref{intradiv}, because one of the was only sampled post-diet, providing us data for this analysis but not allowing us to compare to a baseline for figure \ref{intradiv}.
+(C) Quantification of average degree centrality and (D) closeness centrality between obese and healthy adult women from the Twin gut study.
+Each point represents a stool sample taken from one of the three adult woman confirmed as obese or healthy and with matching virus and bacteria data.
+*\label{dietnetworks}](../figures/dietnetworks.pdf){ width=90% }
 
@@
 ![**Intrapersonal vs Interpersonal Network Dissimilarity Across Different Human Systems.** *(A) NMDS ordination illustrating network dissimilarity between subjects over time.
-Each sample is colored by subject, with each sample pair collected 8-10 days apart.
+Each sample is colored by subject, with each colored sample pair collected 8-10 days apart.
 Dissimilarity was calculated using the Bray-Curtis metric based on abundance weighted eigenvector centrality signatures, with a greater distance representing greater dissimilarity in bacteria and phage centrality and abundance.
+Only four subjects were included here, compared to the five used in figure \ref{dietnetworks}, because one of the subjects was missing the inital sampling time point and therefore lacked temporal sampling.
 (B) Quantification of gut network dissimilarity within the same subject over time (intrapersonal) and the mean dissimilarity between the subject of interest and all other subjects (interpersonal).
-The p-value is also provided.
+The p-value is provided near the bottom of the figure.
 (C) Quantification of gut network dissimilarity within subjects from the same family (intrafamily) and the mean dissimilarity between subjects within a family and those of other families (interfamily).
-The p-value is also provided.
+Each point represents the inter-family and intra-family dissimilarity of a twin or mother that was sampled over time.
 (D) Quantification of skin network dissimilarity within the same subject and anatomical location over time (intrapersonal) and the mean dissimilarity between the subject of interest and all other subjects at the same time and the same anatomical location (interpersonal).
-P-value was calculated using a paired Wilcoxon test.*\label{intradiv}](../figures/intrapersonal_diversity.pdf){ width=90% }
+All p-values were calculated using a paired Wilcoxon test.*\label{intradiv}](../figures/intrapersonal_diversity.pdf){ width=90% }
 
@@
 Notched box-plots were created using ggplot2 and show the median (center line), the inter-quartile range (IQR; upper and lower boxes), the highest and lowest value within 1.5 \* IQR (whiskers), outliers (dots), and the notch which provides an approximate 95% confidence interval as defined by 1.58 \* IQR / sqrt(n).
+Sample sizes for each group were: Moist = 81, Sebaceous = 56, IntMoist = 56, Occluded = 106, Exposed = 61, IntOccluded = 26.
 (C) NMDS ordination depicting the differences in skin microbiome network structure between skin moisture levels and (D) occlusion.
@@
 
+# Supplemental Tables
+
+--------------------------
+Study                     Citation          Virome Quality Control Measures
+
+------------------------- ----------------- -----------------------------------------------------
+   Diet & the Gut Virome     Minot, 2011    • 16S rRNA gene qPCR revealed reduction in bacterial DNA of at least 10,000X.
+
+                                            • Alignment of shotgun sequences revealed 35X reduction in 16S rRNA gene alignments in virome compared to bacteria shotgun.
+
+                                            • Electron microscopy and nucleic acid stain techniques visually confirmed lack of bacteria in virome samples.
+
+   Skin Virome             Hannigan, 2015    • Significant reduction in reads mapping to 16S rRNA gene sequence, compared to bacteria shotgun dataset.
+
+                                             • Significant reduction in reads mapping to human genome, compared to bacteria shotgun dataset.
+
+                                             • Average viral relative abundance of 0.4% in bacterial shotgun dataset.
+
+  Twin Gut Virome           Reyes, 2010      • Confirmation that 2.5% of bacterial shotgun reads mapped to virome, and 76% of virome reads matched the shotgun 2.5%.
+--------------------------
+
+Table: Summary of the primary quality control measures reported in the original publications of the viromes used in this current study.
+
+\newpage
+
 # References

### 2017-07-26 revfiles/ReviewerResponse.md
Draft reviewer comments.

# Reviewer #1 (Comments for the Author):

Hannigan et al. describe a machine learning-based approach for predicting phage-bacterial host interactions and use it to build ecological networks for human microbiome samples based on published data. They analyze differences in network features as a function of host nutrition (low fat vs high fat), health status (lean vs obese twins) and body regions (diverse skin sites). The authors built their prediction model using data from previous publications as training set, containing 43 diverse bacterial species and 30 diverse phage strains, where the infectiveness relationship had been established experimentally (Fig 1A). The evaluation of the resulting predictive model shows that the amino acid similarity between phage and bacteria was the strongest predictor of infectiveness, followed by nucleotide similarity, shared protein families, and to a negligible degree CRISPRs (Fig 1B). They attribute the lack of predictive power of the latter to the minimal amounts of CRISPRs identified.

The authors applied their model to three publicly available datasets (skin microbiome, twin gut microbiome of lean and obese twins and their mothers, and effect of diet on the gut microbiome) that contain sequences from full metagenomics as well as from purified viral-like particle fractions. Relationships between 298 phages operational genomic units (OGU) and 280 bacteria OGU were predicted from the combined taxa of all datasets and overall network features quantified (Fig1 D-G). Using the phage-bacteria network analysis, they explore the impact of diet and obesity in the gut microbiome structure (Fig 2), compare similarities of intrapersonal and interpersonal networks of skin and gut microbiomes (Fig 3) and test for differences between skin sites with different environmental qualities (Fig 4). All these comparisons are performed by either using measurements of network dissimilarity and network topology (eigen centrality, closeness centrality and degree centrality).

Based on their results, the authors conclude that changes in diet, health state and body site topology can influence the phage-bacterial network stability. As an example they show that high-fat diet and obesity as well as external exposure of skin microbiota to reduce network connectivity. They also find network similarities to agree with the well-accepted notion of differences to be smaller within than between individuals.

The approach taken by Hannigan et al. to construct phage-bacteria networks to analyze host-associated communities is original and will be relevant to many researchers within and also outside of the field of human microbiome research. One missing point the description on how the authors assessed the amount of bacterial DNA in the viral-like particle (VLP) fraction. If bacterial DNA were wrongly labeled as being of viral origin, it would lead to the formation of false phage OGU and result in relationships based on bacterial co-occurrences rather than phage infectivity. Also, the reader is not informed how abundances in the VLP fraction where integrated with the abundances of microbial fractions. In other words, if the DNA of the whole sample was not extracted together, how can the relative abundance of each fraction be compared to each other?

The presented results on the impact of diet and obesity, however, are severely underpowered, as the authors also point out (159, 187, 236). In particular, only one or two data points per group were used to compare the influence of diet and obesity (Figure 2). Although intra- vs. inter-individual and intra- vs. inter-family differences cannot be confirmed by statistical tests, they are claimed to exist. This is unfortunate, since the analytical approach using appropriately powered statistics reveals significant differences between skin sites, highlighting the potential power and impact of the methodology.

Taken together, the machine leaning-based approach would only require some additional tests to exclude or estimate the degree of bacterial contamination in the VLP fraction and some explanations on how abundances in VLP and microbial fractions were combined and normalized. With regard to the application of the methodology, two of the three datasets are severely underpowered to support major claims on the impact of diet and health state. I would thus recommend to put the focus of the manuscript on the analysis of the well-powered skin data set and reduce the diet and obesity-related part as potentially promising case-examples while clearly describing the limitations of the analysis (which is done) and toning down or removing some of the respective claims.

## Major Points

1.  Conceptual distinction between infectiveness and community stability: For the development of the network approach, the authors use infectiveness data as evidence for interactions between phage and bacteria. For the analysis part, such interactions are used to describe community stability (i.e., more interactions per node, higher stability). If phages are considered as agents grazing on their microbial hosts, is it justified to use degree of infectiveness analogous to community stability? Some discussion or background information would be helpful.

*Thank you for pointing this out. This is a good point for further discussion. We clarified in the text why we feel this is justified, as the degree of stability refers to the reduced likelihood that a randomly removed node will cause a major divide or lack of connection between agents within the network. We also provided citations to work that further explains this in metabolic and phage-bacteria systems, if the readers are interested in a more in-depth discussion of the topic.*

2.  Testing for bacterial DNA in the phage fraction: The authors do not describe how bacterial sequences in the VLP fraction were eliminated. If bacterial sequences were present, they would be labeled as phages and lead to false predictions.

*This is a good point and we agree that this needed to be better outlined in the text. It is important that we make clear that we are not violating our assumption that that phage OGU pool represents phages and the bacterial OGU pool represented bacteria. We updated the manuscript to address this concern in multiple ways. First, because this was a "data mining" endeavor, we do not have access to the samples for performing further molecular QC. The original study authors did however perform and report a variety of QC measures to confirm the purity of their virome sample sets. To this end, we summarized those QC results into Table S1, which we added to the manuscript and referenced in the results. Furthermore, we performed our own local alignment (blast) QC protocol to assess the degree to which our OGUs were contaminated. We found that many phages had genomic elements similar to reference bacteria genomes, which agrees with the previous findings that the majority of the phages in these systems are temperate. We also identified two complete phages using the stringent Virsorter algorithm. Conversely, there were minimal phage reference genome elements identified in the bacterial OGUs, which also agrees with the previous findings that these sequences are mostly bacterial (cited in Table S1). We also did not identify any phages in the bacterial dataset using Virsorter. The lack of prophage identification, both here and in the previous study analyses, speaks to the need to sequence the bacteria more deeply and create more robust assemblies (their larger size prevents assembly completion at the level that can be done with viruses).*

3.  Combining data from VLPs and microbial fractions: Information on how data from these two individual samples were matched and normalized is not provided. How are potentially different amounts of VLPs in different samples accounted for?

*We are not sure we understand this comment? Our goal was to use the edge weights to reflect samples that were highly relative abundant in both the bacterial metagenome and the phage metagenome. The edge weight used was therefore a multiplication function of the two relative abundance values, yielding a high value for those that were highly abundant in both, and low values for those that were not. This is described in the methods section of the manuscript. These values were recorded on a per-sample basis (within the subgraphs) so that each potential combination is accounted for, within each sampling pair. Furthermore, it may be worth noting that the bacterial and viral DNA extractions were done separately so as to maximize purity of the samples, but this approach allows us to more accurately measure the relative abundance in the bacterial and especially viral samples, as has been noted previously. We made this clearer in the text and added more  relevant citations.*

4.  Datasets underpowered to sustain claims on diet and obesity: Although the authors point out that the data are underpowered, they still make strong claims and the analysis of these data as major points of this manuscript.

*We appreciate this point referring to figure 2, and while we attempted to avoid too strong of claims, we realize we fell short. Our goal was to present these as interesting observations, somewhat like a case study. We addressed this in the manuscript by further minimizing the effects of our claims and adding relevant caveats.*

For example, 
a.  the authors describe (158-164) how diet affects the network structure, without using sufficient data points to perform any statistical tests.
b.  the authors claim (166-169) a difference between obese and healthy individuals by using one or two data points per group.
With respect to these analyses, the authors write for example, "We found evidence that diet was sufficient to alter gut microbiome network connectivity" (235), which is a clear overstatement and such claims should be modified here and elsewhere in the manuscript.

## Minor Points

* Throughout the manuscript, it is hard to follow how which and how many data points were used for each analysis. A supplementary table listing the samples for each of the studies and providing sample sizes in each figure would be very helpful.

*We agree that this needed to be clarified in the manuscript. We added the relevant sample size information to the text and figure legends to make clearer what the data actually represent.*

* Figure 1D is not informative and could be removed since the information is provided in the main text.

*It is true that the information is in the text, but we feel this is still an important panel to the figure because it illustrates the degree of connectance between the phage and bacterial communities, and provides a visualization to the text included in the results section. Even more importantly, we intended this panel to act as a visual validation that there are no erroneous connections among phage and bacterial nodes. We made these reasons clearer in the text.*

* Figure 1 EFG: color labels in legend are wrong and not since y-axis labels are present.

*Thank you for pointing this mistake out. We have fixed it.*

* Figure 2 could use headers above the plots to indicate which one relate to impact of diet (AB) and which one to obesity (CD).

*This was especially confusing because of the mislabeled axes. We fixed the axis labels to make the data clearer.*

* Figure 2 panels C and D have wrong x-axis labels. Should be "obese" and "healthy". Here it is not clear why only 2 and 1 data points are available. The original data in Reyes et al. seem to consist of 32 fecal samples from 13 individuals from 5 families with at least two time points.

*We fixed the error, and made the causes for the sample size clearer in the text as well as the figure legends. Thank you for pointing this out.*

* Figure 3A. Not clear why only 8 data points are available if Minot et al. report 15 samples from 5 individuals.

*We only used the subset of data points that had temporal data spaced approximately the same distance apart. We made this clearer in the text.*

* Figure 3B. Not clear how data were computed. If the comparison of within individuals (i.e. 1 value if same data as in 3A) with mean of all other individuals, then standard deviations should be shown. Suggest to show only the smallest inter-individual distance, which would greatly strengthen the point on individuality.

*Yes, this is a comparison between the interpersonal distance, and the distance between the person's samples and all other samples. We considered calculating the distance differences using standard deviation and pairing the samples as every interpersonal comparison instead of the average. The problem we found was that that pairwise comparison inflated the significance values by artificially creating a large n, all of which were not truly independent. We therefore collapsed the data points into a single average instead of pairing the same sample to many other samples. Showing the smallest distance is also an interesting idea, but we are worried that this may also skew the data by failing to account for the variation in the sample distributions.*

* Figure 3C. Not clear which data were used, in particular if same data as in Figure 2 were used. E.g., were only the twins included, or also the mothers?

*We made this information clearer in the text.*

* Figure 3D. As described for 4B, smallest inter-individual distance should be used if showing individuality should be demonstrated, which is very different from showing on average higher similarity within than between individuals.

*This is also a good point, thank you. We hit this point in the related bullet point above.*

* Figure 4. Please provide sample numbers for each of the boxplots and data in 4E.

*We added this information to the figure legend.*

# Reviewer #2 (Comments for the Author):

The work presented in the manuscript entitled "Biogeography & Environmental Conditions Shape Phage & Bacteria Interaction Networks Across the Human Microbiome" by Hannigan et al. consists on prediction of host-bacteriophage interactions from skin and gut microbiome public datasets and graph-based analysis of the underlying phage-host networks, assessing the effects of associated metadata (obesity, diet, type of skin site). For the prediction of interactions, the authors trained a random forest classifier using known phage-bacteria interactions using the following features: CRISPR-spacer detection, prophage prediction based on nucleotide similarity (blastn, e-value <=1e-10.), shared gene identification based on aminoacid similarity (Diamond) and detection of pairs of genes whose proteins interact (Pfam interaction information within the Intact database). My main concern with this paper is on the prediction of phage-host interactions. The authors report 72287 infectious relationships among 578 nodes, representing 298 phages and 280 bacteria. If all phages would infect all bacteria, the network would have 83440 links (complete network = total possible interactions). The reported network contains 86,6% of the number of possible interactions. This essentially means that most phages are generalists and can infect any bacteria...The authors fail to report the number of phages and bacteria per microbiome: skin or gut. This number would allow us to estimate the total number of possible interactions within an environment (gut or skin), a more relevant number to calculate the density of the network. One explanation for the high density of the network is the use of aminoacid similarity between phage and bacteria as predictor. Shared genes have been used to link phages to their host, but at nucleotide rather than at aminoacid level (Edwards et al. FEMS Microbiol Rev. 2016). Independently from the classifier performance, when confronted with this high network density, the shared genes should have been further analysed and the procedure revised.

## Major comments:

1. Network inference: operational genomic units were linked sample-wise, microbiome type wise or all microbiome together? Aminoacid similarity is not specific enough and is likely the cause of the high density of the network, which suggests that almost all phages infect all bacteria (86% density), meaning the resulting network is (likely) no statistically different from a random network with equal number of links. In the absence of further evidence and controls, I am not convinced with this result.

*We would like to thank the reviewer for pointing out this area of clarification. The operational genomic units were linked sample-wise, by linking the OGUs found between each sample. The source of these samples were recorded, so that we knew whether they originated from the gut or the skin, etc. We also agree that amino acid similarity of genes along is not enough, and that is why we incorporated other similarity metrics in our random forest model. Even though the amino acid similarity metric was the most "important" to the model, it was used within the context of the other metrics in the random forest. It is true that this is a dense network, and we hypothesize that that may be because each OGU does not necessarily represent a specific phage or bacterial strain, but is an operationally defined group of similar genomes, similar to the operation taxonomic unit (OTU) often used in microbiome studies.*

2. Number of samples per study and metadata groups are not reported but the authors state about the diet effect: "Tests for statistical differences were not performed due to the small sample size ". If sample sizes are small for statistical analysis, no comparison should be made in the first place. The plots of Figure 2 feature 2 vs. 3 comparison for high fat vs. low fat and 2 vs. 1 for obese vs. healthy...are these the nb. of samples per group? Bacterial or phage diversity are likely confounders for network connectivity comparisons and this is overlooked here.

*We appreciate this point and while we attempted to report this as more of a case trial observation (as outlined for the related comment above), we unintentionally used too strong of language, thus overstating the finding beyond our intentions. We also needed to make the analysis being performed clearer in the text. It is also true that the network analysis is a function of the diversity of the two communities, and we do not wish to make it sound like they are separate. We would be hesitant however to call this dependence a confounder since it is more building off of the diversity data and incorporating other additional information from the community. We made this clearer in the text.*

3. Graph-analysis: Analysis of network nestedness and modularity (Weitz et al. Trends in Micr. 2012) may be more relevant than diameter, degree and closeness for phage-host networks.

*We chose to omit such an analysis as was done in Weitz et al because, as we attempted to describe in the manuscript, we are not using the OGUs as individual phage strains as was done in the Weitz and other papers. Instead these OGUs are operational units that allow us to condense our dataset by contig similarity. Making statements with as high of phylogenetic resolution as Weitz et al would be unfounded at this point.*

## Other comments:

1.  Training set contained on 43 diverse bacterial species and 30 diverse phage strains. Given than these numbers are (much) lower than the ones published in references 41, 47-51, from which the data on known interactions was extracted, I wonder how the curation/filtering was done.

*Our limitation was not on how many known interactions that could be found, but instead on how many non-interactions could be found. In other words, we needed to utilize known infectious pairs, as well as phages that have been validated as not infecting certain bacteria. Because we were limited to only a small number of validated non-interactions, we decided to also limit the number of positive interactions included, so as to prevent a large bias in the ratio between the two interaction types in the training dataset. We appologize for this not being clearer in the text and have clarified that text.*

2.  The sensitivity of the classifier is reported as 0.846. However, "Approximately one third of the training set relationships yielded no score and therefore were unable to be assigned an interaction prediction". This means the sensitivity can not be higher than 0.67...

*Here we are reporting the accuracy for those values that were scored. Pairs without scores were omitted from this analysis, and we instead only focused on those that were scored. We made this clearer in the text.*

3.  Dataset: The authors should report the number of samples per study and associated metadata (e.g low /high fat diet, obese/lean).

*This is a very good point and we added this information.*

4.  OGU construction, why the OGUs were constructed with CONCOCT without coverage information (CONCOCT-NC)?

*We maintained the coverage information but omitted the feature of total sequence counts since we worried this would confound the results. This means that the feature set matrix used by the CONCOCT clustering algorithm only included k-mer frequencies and sequence count (relative abundance) values per sample.*

5.  Network construction: The master network contained the three studies as sub-networks, which themselves each contained sub-networks for each sample. The authors provide only numbers of phage and bacterial nodes for the master network. For predicting the associations, were all phages and bacteria considered disregarding the sample type (skin vs. gut)?

*Yes.*

6.  The phages and bacteria in the gut diet and twin sample sets were more sparsely related: each contained fewer than 150 vertices, fewer than 20,000 relationships". This is a bipartite network, what should be reported is the number of vertices of each type rather than the total nb of vertices (this is relevant to calculate the network density, we cannot know here what is the nb of possible interactions).

*This is a good point that makes the manuscript clearer. We added these values to the results section of the manuscript so that the reader can access this information.*

7.  Comparison with Paez-Espino et al. (Nature 2016, doi:10.1038/nature19094) phage-host human associated subnetworks should be done.

*This is a cool suggestion. It sounds like this comment is a reference to the host range inferrences within the suggested manuscript. Paez-Espino et al used CRISPR and tRNA comparisons to identify bacterial hosts for the bacteriophage genomes and genome fragments. The group's work suggested that phage host range is broader than has been traditionally appreciated. In our discussion section, we mentioned a comparison of these and related findings to our own dataset. This information was included in the operational genomic unit section of the discussion.*

8.  Figure 1. panel B: methods include blastx, in the network inference Diamond was used instead; Panel E-F, colors in figure do not correspond to those in legend.

*It is true that we used the Diamond aligner, but we used the blastx algorithm within the Diamond aligner (written as `diamond blastx`). We made this information clearer in the text and figure legend. We also fixed the color references.*

9.  Figure 2 legend: it is not clear if/that the dots represent sample-wise networks. "Lines represent the mean degree of centrality for each diet", the only lines in the plots are parallel to the centrality axis, they are just splitting high fat vs. low fat diet fields in the x axis. X-axis from panels C and D are mislabeled.

*Thank you for pointing these out as well. We clarified the dot information in the text and legend, and also fixed the incorrect labels.*

# Reviewer #3 (Comments for the Author):

This research article by Hannigan et.al., focuses on an important and indispensable part of the microbiome- "Virome". Interplay within inter and intra microbial community members is important to get the complete picture. This paper focuses on such interplay and understands that microbiome operates with both the viral and bacterial community members. For this study they worked on previously published metagenomics data, to not only infer based on nucleotide information but also based on protein interactions.

## Main Concerns:

The data from studies referenced 13 & 14 even though may have bias due to sequencing, have similarity in approach wherein collection of VLP and bacterial whole genome shotgun sequencing was performed. VLP data in reference 14 was based on Genomiphi amplification of VLP DNA prior to sequencing that is bound to introduce amplification and sequencing errors which would further lead to clustering bias in the current study. The concern is also regarding data from references 38 and 39 that are not only performed by 454 technology. Further on, reference 39 doesn't have VLP data available. The study is based on 16S and whole genome shotgun of gut microbiome.

*We referenced this statement in reference 38: "We had previously performed shotgun sequencing of total fecal DNA isolated from the 12 frozen samples obtained at the first time point that were now used to prepare purified VLPs (Table S3)." This reference (12) led to our reference 39, which contained the data in the related SRA submission and was cited accordingly. Between these two published datasets, we were able to pair the whole shotgun sequencing data with the purified virome data.*

The question is how were the above biases (including the kitome) taken into consideration while conducting these analyses. The authors may be better off carrying out the analysis only using data from references 13 and 14 for skin and only data from 38 for gut?

*This is a good point that there certainly could have been extraction, kit, or other related differences between the 38 and 39 dataset. We feel these findings, which we attempt to present more as case observations that will require more dedicated study, are worth noting in the paper. We altered the text to better clarify our position. We also avoided analyses in which we directly compared datasets, because we are unable to discern between true biological differences and differences between the studies. Such analyses would require further and properly controlled sampling.*

Even though the authors have acknowledged some of these biases including MDA, in their discussion, these concerns should be explicitly discussed and measures they have taken to overcome these.

*This is a good point. We included further discussion about the biases associated with these methods. As far as measures to overcome these biases, we acknoledge the biases and only performed analyses within studies where samples were treated the same. The goal is for this work to use these findings as a jumping off point, and performing future studies with more limited biases from techniques such as MDA.*

### 2017-07-27 revfiles/ReviewerResponse.md

@@
+# Point-By-Point Reviewer Comment Response
+
+We would like to thank the reviewers and editor for taking the time to read our manuscript and offer constructive feedback.
+We incorporated the suggested edits and feel that this great improved the manusript.
+Below we included a point-by-point response to the reviewer comments, with our comments highlighted in *italics* and the relevant manuscript change locations highlighted as **bold**.
+
 # Reviewer #1 (Comments for the Author):
@@
 We clarified in the text why we feel this is justified, as the degree of stability refers to the reduced likelihood that a randomly removed node will cause a major divide or lack of connection between agents within the network.
-We also provided citations to work that further explains this in metabolic and phage-bacteria systems, if the readers are interested in a more in-depth discussion of the topic.*
+We also provided citations to work that further explains this in metabolic and phage-bacteria systems, if the readers are interested in a more in-depth discussion of the topic.* **LINES 177 - 184**
 
@@
 We also did not identify any phages in the bacterial dataset using Virsorter.
-The lack of prophage identification, both here and in the previous study analyses, speaks to the need to sequence the bacteria more deeply and create more robust assemblies (their larger size prevents assembly completion at the level that can be done with viruses).*
+The lack of prophage identification, both here and in the previous study analyses, speaks to the need to sequence the bacteria more deeply and create more robust assemblies (their larger size prevents assembly completion at the level that can be done with viruses).* **TABLE S1; LINES 91 - 114**
 
@@
 Furthermore, it may be worth noting that the bacterial and viral DNA extractions were done separately so as to maximize purity of the samples, but this approach allows us to more accurately measure the relative abundance in the bacterial and especially viral samples, as has been noted previously.
-We made this clearer in the text and added more  relevant citations.*
+We made this clearer in the text and added more relevant citations.* **LINES 150 - 154**
 
@@
 Our goal was to present these as interesting observations, somewhat like a case study.
-We addressed this in the manuscript by further minimizing the effects of our claims and adding relevant caveats.*
+We addressed this in the manuscript by further minimizing the effects of our claims and adding relevant caveats.* **LINES 29 - 31; 182 - 184; 185 - 202; 269 - 278**
 
@@
 
-*We agree that this needed to be clarified in the manuscript. We added the relevant sample size information to the text and figure legends to make clearer what the data actually represent.*
+*We agree that this needed to be clarified in the manuscript.
+We added the relevant sample size information to the text and figure legends to make clearer what the data actually represent.* **LINES 187; 195 - 197; 220 - 222; 224 - 226; FIGURE LEGENDS 2,3,4**
 
@@
 *It is true that the information is in the text, but we feel this is still an important panel to the figure because it illustrates the degree of connectance between the phage and bacterial communities, and provides a visualization to the text included in the results section.
-Even more importantly, we intended this panel to act as a visual validation that there are no erroneous connections among phage and bacterial nodes.
-We made these reasons clearer in the text.*
+Even more importantly, we intended this panel to act as a visual validation that there are no erroneous connections among phage and bacterial nodes.*
 
@@
 
-*Thank you for pointing this mistake out. We have fixed it.*
+*Thank you for pointing this mistake out. We have fixed it.* **FIGURE LEGEND 1**
 
@@
 
-*This was especially confusing because of the mislabeled axes. We fixed the axis labels to make the data clearer.*
+*This was especially confusing because of the mislabeled axes. We fixed the axis labels to make the data clearer.* ** FIGURE 2**
 
@@
 
-*We fixed the error, and made the causes for the sample size clearer in the text as well as the figure legends. Thank you for pointing this out.*
+*We fixed the error, and made the causes for the sample size clearer in the text as well as the figure legends. Thank you for pointing this out.* **LINES 195 - 196; FIGURE LEGEND 2**
 
@@
 
-*We only used the subset of data points that had temporal data spaced approximately the same distance apart. We made this clearer in the text.*
+*We only used the subset of data points that had temporal data spaced approximately the same distance apart. We made this clearer in the text.* **LINES 187 - 188; 220 - 222; FIGURE LEGEND 1,2**
 
@@
 
-*We made this information clearer in the text.*
+*We made this information clearer in the text.* **LINES 224 - 226; FIGURE LEGEND 3**
 
@@
 
-*We added this information to the figure legend.*
+*We added this information to the figure legend.* **FIGURE LEGEND 4**.
 
@@
 We also agree that amino acid similarity of genes along is not enough, and that is why we incorporated other similarity metrics in our random forest model.
-Even though the amino acid similarity metric was the most "important" to the model, it was used within the context of the other metrics in the random forest.
-It is true that this is a dense network, and we hypothesize that that may be because each OGU does not necessarily represent a specific phage or bacterial strain, but is an operationally defined group of similar genomes, similar to the operation taxonomic unit (OTU) often used in microbiome studies.*
+As we described, even though the amino acid similarity metric was the most "important" to the model, it was used within the context of the other metrics in the random forest.
+It is true that this is a dense network, and we hypothesize that that may be because each OGU does not necessarily represent a specific phage or bacterial strain, but is an operationally defined group of similar genomes, similar to the operation taxonomic unit (OTU) often used in microbiome studies.* **SEE LINES 117 - 119; 305 - 323**
 
@@
 We would be hesitant however to call this dependence a confounder since it is more building off of the diversity data and incorporating other additional information from the community.
-We made this clearer in the text.*
+We made this clearer in the text.* **LINES 187; 195 - 197; 220 - 222; 224 - 226; FIGURE LEGENDS 2,3,4**
 
@@
 Instead these OGUs are operational units that allow us to condense our dataset by contig similarity.
-Making statements with as high of phylogenetic resolution as Weitz et al would be unfounded at this point.*
+We feel that making statements with as high of phylogenetic resolution as Weitz et al would be unfounded at this point.*
 
@@
 Because we were limited to only a small number of validated non-interactions, we decided to also limit the number of positive interactions included, so as to prevent a large bias in the ratio between the two interaction types in the training dataset.
-We appologize for this not being clearer in the text and have clarified that text.*
+We appologize for this not being clearer in the text and have clarified that text.* **LINES 125 - 128**
 
@@
 Pairs without scores were omitted from this analysis, and we instead only focused on those that were scored.
-We made this clearer in the text.*
+We made this clearer in the text.* **LINES 144 - 145; FIGURE LEGEND 1**
 
@@
 
-*This is a very good point and we added this information.*
+*This is a very good point and we added this information.* **LINES 187; 195 - 197; 220 - 222; 224 - 226; FIGURE LEGENDS 2,3,4**
 
@@
 
-*Yes.*
+*Yes, as outlined in the results and methods sections.*
 
@@
 
-*This is a good point that makes the manuscript clearer. We added these values to the results section of the manuscript so that the reader can access this information.*
+*This is a good point that makes the manuscript clearer. We added these values to the results section of the manuscript so that the reader can access this information.* **LINES 154 - 163**
 
@@
 In our discussion section, we mentioned a comparison of these and related findings to our own dataset.
-This information was included in the operational genomic unit section of the discussion.*
+This information was included in the operational genomic unit section of the discussion.* **LINES 321 - 323**
 
@@
 *It is true that we used the Diamond aligner, but we used the blastx algorithm within the Diamond aligner (written as `diamond blastx`).
-We made this information clearer in the text and figure legend.
-We also fixed the color references.*
+We made this information clearer in the text.
+We also fixed the color references.* **FIGURE 1**
 
@@
 
-*Thank you for pointing these out as well. We clarified the dot information in the text and legend, and also fixed the incorrect labels.*
+*Thank you for pointing these out as well. We clarified the dot information, and also fixed the incorrect labels.* **FIGURE 2**
 
@@
 *This is a good point.
-We included further discussion about the biases associated with these methods.
+We included of the biases associated with these methods.
 As far as measures to overcome these biases, we acknoledge the biases and only performed analyses within studies where samples were treated the same.
-The goal is for this work to use these findings as a jumping off point, and performing future studies with more limited biases from techniques such as MDA.*
+The goal is for this work to use these findings as a jumping off point, and performing future studies with more limited biases from techniques such as MDA.* **LINES 296 - 304**

### 2017-08-06 revfiles/ReviewerResponse.md
Added a few more small-ish edits to the document and reviewer comments.

@@
-# Point-By-Point Reviewer Comment Response
+---
+title: Point-By-Point Reviewer Comment Response
+author: Hannigan *et al*.
+---
 
 We would like to thank the reviewers and editor for taking the time to read our manuscript and offer constructive feedback.
-We incorporated the suggested edits and feel that this great improved the manusript.
+We incorporated the suggested edits and feel that this greatly improved the manuscript.
 Below we included a point-by-point response to the reviewer comments, with our comments highlighted in *italics* and the relevant manuscript change locations highlighted as **bold**.
@@
 *Thank you for pointing this out.
-This is a good point for further discussion.
+This is a good point warranting further discussion.
 We clarified in the text why we feel this is justified, as the degree of stability refers to the reduced likelihood that a randomly removed node will cause a major divide or lack of connection between agents within the network.
-We also provided citations to work that further explains this in metabolic and phage-bacteria systems, if the readers are interested in a more in-depth discussion of the topic.* **LINES 177 - 184**
+The infectiveness refers to the link we drew between the bacteria and phages (how we connected phages and bacteria in our networks).
+We also provided citations to work that further explains this in metabolic and phage-bacteria systems, if the readers are interested in a more in-depth discussion of the topic.* **LINES 176 - 183**
 
@@
 We also did not identify any phages in the bacterial dataset using Virsorter.
-The lack of prophage identification, both here and in the previous study analyses, speaks to the need to sequence the bacteria more deeply and create more robust assemblies (their larger size prevents assembly completion at the level that can be done with viruses).* **TABLE S1; LINES 91 - 114**
+The lack of prophage identification, both here and in the previous study analyses, speaks to the need to sequence the bacteria more deeply and create more robust assemblies (their larger size prevents assembly completion at the level that can be done with viruses).* **TABLE S1; LINES 93 - 119**
 
@@
 These values were recorded on a per-sample basis (within the subgraphs) so that each potential combination is accounted for, within each sampling pair.
-Furthermore, it may be worth noting that the bacterial and viral DNA extractions were done separately so as to maximize purity of the samples, but this approach allows us to more accurately measure the relative abundance in the bacterial and especially viral samples, as has been noted previously.
-We made this clearer in the text and added more relevant citations.* **LINES 150 - 154**
+Furthermore, it may be worth noting that the bacterial and viral DNA extractions were done separately so as to maximize purity of the samples.
+This approach allows us to more accurately measure the relative abundance in the bacterial and especially viral samples, as has been noted previously.
+We made this clearer in the text and added more relevant citations.* **LINES 149 - 153**
 
@@
 Our goal was to present these as interesting observations, somewhat like a case study.
-We addressed this in the manuscript by further minimizing the effects of our claims and adding relevant caveats.* **LINES 29 - 31; 182 - 184; 185 - 202; 269 - 278**
+We addressed this in the manuscript by further minimizing the effects of our claims and adding relevant caveats.* **LINES 29 - 31; 184 - 201; 268 - 277**
 
@@
 *We agree that this needed to be clarified in the manuscript.
-We added the relevant sample size information to the text and figure legends to make clearer what the data actually represent.* **LINES 187; 195 - 197; 220 - 222; 224 - 226; FIGURE LEGENDS 2,3,4**
+We added the relevant sample size information to the text and figure legends to make clearer what the data actually represent.* **LINES 186; 194 - 196; 219 - 221; 223 - 225; FIGURE LEGENDS 2,3,4**
 
@@
 
-*Thank you for pointing this mistake out. We have fixed it.* **FIGURE LEGEND 1**
+*Thank you for noting this mistake. We have fixed it.* **FIGURE LEGEND 1**
 
@@
 
-*This was especially confusing because of the mislabeled axes. We fixed the axis labels to make the data clearer.* ** FIGURE 2**
+*This was especially confusing because of the mislabeled axes. We fixed the axis labels to make the data clearer.* **FIGURE 2**
 
@@
 
-*We fixed the error, and made the causes for the sample size clearer in the text as well as the figure legends. Thank you for pointing this out.* **LINES 195 - 196; FIGURE LEGEND 2**
+*We fixed the error, and made the causes for the sample size clearer in the text as well as the figure legends. Thank you for pointing this out.* **LINES 194 - 195; FIGURE LEGEND 2**
 
@@
 
-*We only used the subset of data points that had temporal data spaced approximately the same distance apart. We made this clearer in the text.* **LINES 187 - 188; 220 - 222; FIGURE LEGEND 1,2**
+*We only used the subset of data points that had temporal data spaced approximately the same distance apart. We made this clearer in the text.* **LINES 186 - 187; 219 - 221; FIGURE LEGEND 1,2**
 
@@
 
-*We made this information clearer in the text.* **LINES 224 - 226; FIGURE LEGEND 3**
+*We made this information clearer in the text.* **LINES 223 - 225; FIGURE LEGEND 3**
 
@@
 As we described, even though the amino acid similarity metric was the most "important" to the model, it was used within the context of the other metrics in the random forest.
-It is true that this is a dense network, and we hypothesize that that may be because each OGU does not necessarily represent a specific phage or bacterial strain, but is an operationally defined group of similar genomes, similar to the operation taxonomic unit (OTU) often used in microbiome studies.* **SEE LINES 117 - 119; 305 - 323**
+It is true that this is a dense network, and we hypothesize that that may be because each OGU does not necessarily represent a specific phage or bacterial strain, but is an operationally defined group of similar genomes, similar to the operation taxonomic unit (OTU) often used in microbiome studies.
+In other words, we are not using the OGUs as individual phage strains, but instead these OGUs are operational units that allow us to condense our dataset by contig similarity.* **SEE LINES 114 - 119; 304 - 322**
 
@@
 We would be hesitant however to call this dependence a confounder since it is more building off of the diversity data and incorporating other additional information from the community.
-We made this clearer in the text.* **LINES 187; 195 - 197; 220 - 222; 224 - 226; FIGURE LEGENDS 2,3,4**
+We made this clearer in the text.* **LINES 186; 194 - 196; 219 - 221; 223 - 225; FIGURE LEGENDS 2,3,4**
 
@@
 Because we were limited to only a small number of validated non-interactions, we decided to also limit the number of positive interactions included, so as to prevent a large bias in the ratio between the two interaction types in the training dataset.
-We appologize for this not being clearer in the text and have clarified that text.* **LINES 125 - 128**
+We apologize for this not being clearer in the text and have clarified that text.* **LINES 124 - 127**
 
@@
 Pairs without scores were omitted from this analysis, and we instead only focused on those that were scored.
-We made this clearer in the text.* **LINES 144 - 145; FIGURE LEGEND 1**
+We made this clearer in the text.* **LINES 143 - 144; FIGURE LEGEND 1**
 
@@
 
-*This is a very good point and we added this information.* **LINES 187; 195 - 197; 220 - 222; 224 - 226; FIGURE LEGENDS 2,3,4**
+*This is a very good point and we added this information.* **LINES 186; 194 - 196; 219 - 221; 223 - 225; FIGURE LEGENDS 2,3,4**
 
@@
 
-*Yes, as outlined in the results and methods sections.*
+*Yes, this is described correctly by the reviewer.*
 
@@
 *This is a cool suggestion.
-It sounds like this comment is a reference to the host range inferrences within the suggested manuscript.
+It sounds like this comment is a reference to the host range inferences within the suggested manuscript.
 Paez-Espino et al used CRISPR and tRNA comparisons to identify bacterial hosts for the bacteriophage genomes and genome fragments.
@@
 In our discussion section, we mentioned a comparison of these and related findings to our own dataset.
-This information was included in the operational genomic unit section of the discussion.* **LINES 321 - 323**
+This information was included in the operational genomic unit section of the discussion.* **LINES 318 - 322**
 
@@
 *This is a good point.
-We included of the biases associated with these methods.
-As far as measures to overcome these biases, we acknoledge the biases and only performed analyses within studies where samples were treated the same.
-The goal is for this work to use these findings as a jumping off point, and performing future studies with more limited biases from techniques such as MDA.* **LINES 296 - 304**
+We included a description of these biases associated with these methods.
+As far as measures to overcome these biases, we acknowledge the biases and only performed analyses within studies where samples were treated the same.
+The goal is for this work to use these findings as a jumping off point, and performing future studies with more limited biases from techniques such as MDA.* **LINES 295 - 303**

### 2017-08-06 doc/manuscript.Rmd
Added a few more small-ish edits to the document and reviewer comments.

@@
 These reports confirmed that the virome libraries consisted of highly purified virus genomic DNA.
-We supplemented these findings and found that, in light of the rigorous purification and quality control during sample collection and preparation, 77% (228 / 298 operational genomic units) still had some nucleotide similarity to a given bacterial reference genome (e-value < $10^{-25}$).
-These findings agree with the previous reports from which we obtained our data, in that the majority of the gut and skin bacteriophages were temperate and shared elements with bacterial reference genomes, primarily because their replication cycles include bacterial genome integration [@Minot:2011ez, @Hannigan:2015fz, @Reyes:2010cwa].
-We additionally confidently identified two OGUs as being complete phages using the stringent Virsorter phage identification algorithm (class 1 confidence group) [@Roux:2015dh].
-
-While the whole metagenomic shotgun sequence samples were not subjected to purification, they primarily consisted of bacteria, quantified with an average viral relative abundance of 0.4% **(Table S1)** [@Minot:2011ez; @Hannigan:2015fz; @Reyes:2010cwa; @Turnbaugh:2009ei].
-We again supplemented these findings and found that only 2% (6 / 280 operational genomic units) of bacteria had significantly strong nucleotide similarity to phage reference genomes (e-value < $10^{-25}$) [@Minot:2011ez; @Hannigan:2015fz; @Reyes:2010cwa; @Turnbaugh:2009ei].
-No OGUs were confidently identified as lytic or temperate phage OGUs in the bacterial dataset using the Virsorter algorithm [@Roux:2015dh].
-Not only does this suggest a low level of "contamination", but also highlights future needs to more deeply sequence the bacterial metagenomes, whose genomes are order of magnitude longer than phages and other viruses.
 
@@
 Contig lengths within clusters ranged between $10^{3}$ and $10^{5.5}$ bp **(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})**.
+
+We supplemented the previous study findings **(Table S1)** and found that, in light of the rigorous purification and quality control during sample collection and preparation, 77% (228 / 298 operational genomic units) still had some nucleotide similarity to a given bacterial reference genome (e-value < $10^{-25}$).
+As absence of bacterial contamination has been confirmed **(Table S1)**, we interpret this as evidence that the majority of the gut and skin bacteriophages were temperate and thereby shared elements with bacterial reference genomes that contained similar integrated phages when sequenced--a trend previously reported [@Minot:2011ez, @Hannigan:2015fz, @Reyes:2010cwa].
+Additionally, we identified two OGUs as being complete phages using the stringent Virsorter phage identification algorithm (class 1 confidence group) [@Roux:2015dh].
+
+While the whole metagenomic shotgun sequence samples were not subjected to purification, they primarily consisted of bacteria, quantified with an average viral relative abundance of 0.4% **(Table S1)** [@Minot:2011ez; @Hannigan:2015fz; @Reyes:2010cwa; @Turnbaugh:2009ei].
+We again supplemented these findings and found that only 2% (6 / 280 operational genomic units) of bacteria had significantly strong nucleotide similarity to phage reference genomes (e-value < $10^{-25}$) [@Minot:2011ez; @Hannigan:2015fz; @Reyes:2010cwa; @Turnbaugh:2009ei].
+No OGUs were confidently identified as lytic or temperate phage OGUs in the bacterial dataset using the Virsorter algorithm [@Roux:2015dh].
+Together this suggests only minimal "contamination" of the bacterial OGUs.
 
@@
 Although the network was highly connected, not all relationships were present in all samples.
-As relationships were weighted by the relative abundances of their associated bacteria and phages, lowly abundant relationships could be present but not highly abundant.
+Relationships were weighted by the relative abundances of their associated bacteria and phages.
 Like the master network, the skin network exhibited a diameter of `r prettyNum(genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Diameter"], big.mark = ",")` (measure of graph size; the greatest number of traversed vertices required between two vertices) and included `r prettyNum(genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Vertices"], big.mark = ",")` (`r prettyNum(sitestats[c(sitestats$class %in% "skin"),"phagenodes"], big.mark = ",")` phages, `r prettyNum(sitestats[c(sitestats$class %in% "skin"),"bacterianodes"], big.mark = ",")` bacteria, `r round(100 * genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Vertices"] / genfigurestats[c(genfigurestats$class %in% "Total"),"Vertices"], digits = 1)`% total) and `r prettyNum(genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Edges"], big.mark = ",")` (`r round(100 * genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Edges"] / genfigurestats[c(genfigurestats$class %in% "Total"),"Edges"], digits = 1)`%) of the master network nodes and edges, respectively **(Figure \ref{RocCurve} E - F)**.
@@
 
-Finally, the networks in this study were built using operational genomic units (OGUs), which represented groups of highly similar bacteria or phage genomes or genome fragments as clustered sub-populations.
+Finally, the networks in this study were built using operational genomic units (OGUs), which represented groups of highly similar bacteria or phage genomes or clustered genome fragments.
 Similar clustering definition and validation methods, both computational and experimental, have been implemented in other metagenomic sequencing studies, as well [@Minot:2012ed; @Deng:2014eb; @Brum:2015iaa; @Roux:2016cc].

### 2017-08-09 doc/manuscript.Rmd
Created doc with diffs highlighted.

@@
 (C) Quantification of average degree centrality and (D) closeness centrality between obese and healthy adult women from the Twin gut study.
-Each point represents a stool sample taken from one of the three adult woman confirmed as obese or healthy and with matching virus and bacteria data.
-*\label{dietnetworks}](../figures/dietnetworks.pdf){ width=90% }
-
-\newpage
+Each point represents a stool sample taken from one of the three adult woman confirmed as obese or healthy and with matching virus and bacteria data.* \label{dietnetworks}](../figures/dietnetworks.pdf){ width=90% }
 

### 2017-08-09 doc/manuscript.tex
Created doc with diffs highlighted.

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
    \setmainfont[]{Helvetica}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Biogeography \& Environmental Conditions Shape Phage \& Bacteria Interaction Networks Across the Human Microbiome},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\usepackage{setspace}
\doublespacing
\usepackage[vmargin=1in,hmargin=1in]{geometry}
\usepackage{lineno}
\linenumbers
\usepackage{parskip}
\setlength{\parskip}{7.5pt}
\exhyphenpenalty=10000
\hyphenpenalty=10000

\newcommand{\beginsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }

\title{Biogeography \& Environmental Conditions Shape Phage \& Bacteria
Interaction Networks Across the Human Microbiome}
    \usepackage{authblk}
    \font\myfont=Helvetica at 12pt
    \font\nextfont=Helvetica at 10pt
                        \author[1]{\myfont Geoffrey D Hannigan}
                    \author[2]{\myfont Melissa B Duhaime}
                    \author[3]{\myfont Danai Koutra}
                    \author[1,*]{\myfont Patrick D Schloss}
                            \affil[1]{\nextfont Department of Microbiology \& Immunology, University of Michigan, Ann
Arbor, Michigan, 48109}
                    \affil[2]{\nextfont Department of Ecology \& Evolutionary Biology, University of Michigan,
Ann Arbor, Michigan, 48109}
                    \affil[3]{\nextfont Department of Computer Science, University of Michigan, Ann Arbor,
Michigan, 48109}
                    \affil[*]{\nextfont To whom correspondence may be addressed.}
            \date{}

\begin{document}
\maketitle

~

~

~

\textbf{\emph{Running Title}}: Network Diversity of the Healthy Human
Microbiome

\textbf{\emph{Corresponding Author Information}}\\
Patrick D Schloss, PhD\\
1150 W Medical Center Dr.~1526 MSRB I\\
Ann Arbor, Michigan 48109\\
Phone: (734) 647-5801\\
Email: pschloss@umich.edu

\newpage

\section{Abstract}\label{abstract}

Viruses and bacteria are critical components of the human microbiome and
play important roles in health and disease. Most previous work has
relied on studying microbes and viruses independently, thereby reducing
them to two separate communities. Such approaches are unable to capture
how these microbial communities interact, such as through processes that
maintain community stability or allow phage-host populations to
co-evolve. We developed and implemented a network-based analytical
approach to describe phage-bacteria network diversity throughout the
human body. We accomplished this by building a machine learning
algorithm to predict which phages could infect which bacteria in a given
microbiome. This algorithm was applied to paired viral and bacterial
metagenomic sequence sets from three previously published human cohorts.
We organized the predicted interactions into networks that allowed us to
evaluate phage-bacteria connectedness across the human body. We found
that gut and skin network structures were person-specific and not
conserved among cohabitating family members. High-fat diets and obesity
appeared to be associated with less connected networks. Network
structure differed between skin sites, with those exposed to the
external environment being less connected and more prone to instability.
This study quantified and contrasted the diversity of virome-microbiome
networks across the human body and illustrated how environmental factors
may influence phage-bacteria interactive dynamics. This work provides a
baseline for future studies to better understand system perturbations,
such as disease states, through ecological networks.

\newpage

\section{Importance}\label{importance}

The human microbiome, the collection of microbial communities that
colonize the human body, is a crucial component to health and disease.
Two major components to the human microbiome are the bacterial and viral
communities. These communities have primarily been studied separately
using metrics of community composition and diversity. These approaches
have failed to capture the complex dynamics of interacting bacteria and
phage communities, which frequently share genetic information and work
together to maintain stable ecosystems. Removal of bacteria or phage can
disrupt or even collapse those ecosystems. Relationship-based network
approaches allow us to capture this interaction information. Using this
network-based approach with three independent human cohorts, we were
able to present an initial understanding of how phage-bacteria networks
differ throughout the human body, so as to provide a baseline for future
studies of how and why microbiome networks differ in disease states.

\newpage

\section{Introduction}\label{introduction}

Viruses and bacteria are critical components of the human microbiome and
play important roles in health and disease. Bacterial communities have
been associated with disease states, including a range of skin
conditions (1), acute and chronic wound healing conditions (2, 3), and
gastrointestinal diseases, such as inflammatory bowel disease (4, 5),
\emph{Clostridium difficile} infections (6) and colorectal cancer (7,
8). Altered human viromes (virus communities consisting primarily of
bacteriophages) also have been associated with diseases and
perturbations, including inflammatory bowel disease (5, 9), periodontal
disease (10), spread of antibiotic resistance (11), and others (12--17).
Viruses act in concert with their microbial hosts as a single ecological
community (18). Viruses influence their living microbial host
communities through processes including lysis, host gene expression
modulation (19), influence on evolutionary processes such as horizontal
gene transfer (22) or antagonistic co-evolution (26), and alteration of
ecosystem processes and elemental stoichiometry (27).

Previous human microbiome work has focused on bacterial and viral
communities, but have reduced them to two separate communities by
studying them independently (5, 9, 10, 12--17). This approach fails to
capture the complex dynamics of interacting bacteria and phage
communities, which frequently share genetic information and work
together to maintain stable ecosystems. Removal of bacteria or phages
can disrupt or even collapse those ecosystems (18, 28--37).
Relationship-based network approaches allow us to capture this
interaction information. Studying such bacteria-phage interactions
through community-wide networks built from inferred relationships could
offer further insights into the drivers of human microbiome diversity
across body sites and enable the study of human microbiome network
dynamics overall.

In this study, we characterized human-associated bacterial and phage
communities by their inferred relationships using three published paired
virus and bacteria-dominated whole community metagenomic datasets (13,
14, 38, 39). We leveraged machine learning and graph theory techniques
to establish and explore the human bacteria-phage network diversity
therein. This approach built upon previous large-scale phage-bacteria
network analyses by inferring interactions from metagenomic datasets,
rather than culture-dependent data (33), which is limited in the scale
of possible experiments and analyses. Our metagenomic interaction
inference model improved upon previous models of phage-host predictions
that have utilized a variety of techniques, such as linear models to
predict bacteria-phage co-occurrence using taxonomic assignments (40),
and nucleotide similarity models that were applied to both whole virus
genomes (41) and related clusters of whole and partial virus genomes
(42). Our approach uniquely included protein interaction data and was
validated based on experimentally determined positive and negative
interactions (i.e.~who does and does not infect whom). Through this
approach we were able to provide a basic understanding of the network
dynamics associated with phage and bacterial communities on and in the
human body. By building and utilizing a microbiome network, we found
that different people, body sites, and anatomical locations not only
support distinct microbiome membership and diversity (13, 14, 38, 39,
43--45), but also support ecological communities with distinct
communication structures and propensities toward community instability.
Through an improved understanding of network structures across the human
body, we empower future studies to investigate how these communities
dynamics are influenced by disease states and the overall impact they
may have on human health.

\section{Results}\label{results}

\subsection{Cohort Curation and Sample
Processing}\label{cohort-curation-and-sample-processing}

We studied the differences in virus-bacteria interaction networks across
healthy human bodies by leveraging previously published shotgun sequence
datasets of purified viral metagenomes (viromes) paired with
bacteria-dominated whole community metagenomes. Our study contained
three datasets that explored the impact of diet on the healthy human gut
virome (14), the impact of anatomical location on the healthy human skin
virome (13), and the viromes of monozygotic twins and their mothers (38,
39). We selected these datasets because their virome samples were
subjected to virus-like particle (VLP) purification, which removed
contaminating DNA from human cells, bacteria, etc. To this end, the
publishing authors employed combinations of filtration, chloroform/DNase
treatment, and cesium chloride gradients to eliminate organismal DNA
(bacteria, human, fungal, etc) and thereby allow for direct assessment
of both the extracellular and fully-assembled intracellular virome
\textbf{(Supplemental Figure \ref{SequenceStats} A-B)} (14, 39). Each
research group reported quality control measures to ensure the purity of
the virome sequence datasets, using both computational and molecular
techniques, including, for example, 16S rRNA gene qPCR \textbf{(Table
S1)}. These reports confirmed that the virome libraries consisted of
highly purified virus genomic DNA.

The bacterial and viral sequences from these studies were quality
filtered and assembled into contigs. We further grouped the related
bacterial and phage contigs into operationally defined units based on
their k-mer frequencies and co-abundance patterns, similar to previous
reports \textbf{(Supplemental Figure \ref{ContigStats} -
\ref{ClusterStats})} (42). We referred to these operationally defined
groups of related contigs as operational genomic units (OGUs). Each OGU
represented a genomically similar sub-population of either bacteria or
phages. Contig lengths within clusters ranged between \(10^{3}\) and
\(10^{5.5}\) bp \textbf{(Supplemental Figure \ref{ContigStats} -
\ref{ClusterStats})}.

We supplemented the previous study findings \textbf{(Table S1)} and
found that, in light of the rigorous purification and quality control
during sample collection and preparation, 77\% (228 / 298 operational
genomic units) still had some nucleotide similarity to a given bacterial
reference genome (e-value \textless{} \(10^{-25}\)). As absence of
bacterial contamination has been confirmed \textbf{(Table S1)}, we
interpret this as evidence that the majority of the gut and skin
bacteriophages were temperate and thereby shared elements with bacterial
reference genomes that contained similar integrated phages when
sequenced--a trend previously reported (14). Additionally, we identified
two OGUs as being complete phages using the stringent Virsorter phage
identification algorithm (class 1 confidence group) (47).

While the whole metagenomic shotgun sequence samples were not subjected
to purification, they primarily consisted of bacteria, quantified with
an average viral relative abundance of 0.4\% \textbf{(Table S1)} (13,
14, 38, 39). We again supplemented these findings and found that only
2\% (6 / 280 operational genomic units) of bacteria had significantly
strong nucleotide similarity to phage reference genomes (e-value
\textless{} \(10^{-25}\)) (13, 14, 38, 39). No OGUs were confidently
identified as lytic or temperate phage OGUs in the bacterial dataset
using the Virsorter algorithm (47). Together this suggests only minimal
``contamination'' of the bacterial OGUs.

\subsection{Evaluating the Model to Predict Phage-Bacteria
Interactions}\label{evaluating-the-model-to-predict-phage-bacteria-interactions}

We predicted which phage OGUs infected which bacterial OGUs using a
random forest model trained on experimentally validated infectious
relationships from six previous publications (41, 48--52). Only bacteria
and phages were used in the model. The training set contained 43 diverse
bacterial species and 30 diverse phage strains, including both broad and
specific ranges of infection \textbf{(Supplemental Figure
\ref{ValidationOverview} A - B)}. While it is true that there are more
known phages that infect bacteria, we were limited by the information
confirming which phages do not infect certain bacteria and attempted to
keep the numbers of positive and negative interactions similar. Phages
with linear and circular genomes, as well as ssDNA and dsDNA genomes,
were included in the analysis. Because we used DNA sequencing studies,
RNA phages were not considered \textbf{(Supplemental Figure
\ref{ValidationOverview} C-D)}. This training set included both positive
relationships (a phage infects a bacterium) and negative relationships
(a phage does not infect a bacterium). This allowed us to validate the
false positive and false negative rates associated with our candidate
models, thereby building upon previous work that only considered
positive relationships (41).

Four phage and bacterial genomic features were used in a random forest
model to predict infectious relationships between bacteria and phages:
1) genome nucleotide similarities, 2) gene amino acid sequence
similarities, 3) bacterial Clustered Regularly Interspaced Short
Palindromic Repeat (CRISPR) spacer sequences that target phages, and 4)
similarity of protein families associated with experimentally identified
protein-protein interactions (53). The resulting random forest model was
assessed and the area under its receiver operating characteristic (ROC)
curve was 0.846, the model sensitivity was 0.829, and specificity was
0.767 \textbf{(Figure \ref{RocCurve} A)}. The most important predictor
in the model was amino acid similarity between genes, followed by
nucleotide similarity of whole genomes \textbf{(Figure \ref{RocCurve}
B)}. Protein family interactions were moderately important to the model,
and CRISPRs were largely uninformative, due to the minimal amount of
identifiable CRISPRs in the dataset and their redundancy with the
nucleotide similarity methods \textbf{(Figure \ref{RocCurve} B)}.
Approximately one third of the training set relationships yielded no
score and therefore were unable to be assigned an interaction prediction
\textbf{(Figure \ref{RocCurve} C)}.

We used our random forest model to classify the relationships between
bacteria and phage operational genomic units, which were then used to
build the interactive network. The master network contained the three
studies as sub-networks, which themselves each contained sub-networks
for each sample \textbf{(Figure \ref{RocCurve} D)}. Metadata including
study, sample ID, disease, and OGU abundance within the community were
stored in the master network for parsing in downstream analyses
\textbf{(Supplemental Figure \ref{NetworkDiagram})}. Bacterial and phage
relative abundance was recorded in each sample for each OGU, and the
weight of the edge connecting those OGUs was calculated as a function of
those relative abundance values. The separate extraction of the phage
and bacterial libraries ensured a more accurate measurement of the
microbial communities, as has been outlined previously (54, 55). The
master network was highly connected and contained 72,287 infectious
relationships among 578 nodes, representing 298 phages and 280 bacteria.
Although the network was highly connected, not all relationships were
present in all samples. Relationships were weighted by the relative
abundances of their associated bacteria and phages. Like the master
network, the skin network exhibited a diameter of 4 (measure of graph
size; the greatest number of traversed vertices required between two
vertices) and included 576 (297 phages, 279 bacteria, 99.7\% total) and
72,127 (99.8\%) of the master network nodes and edges, respectively
\textbf{(Figure \ref{RocCurve} E - F)}. The phages and bacteria in the
gut diet and twin sample sets were more sparsely related, with the diet
study consisting of 89 (41 phages, 48 bacteria) nodes and 5,566
relationships, and the twin study containing 137 (36 phages, 101
bacteria) nodes and 17,250 relationships \textbf{(Figure \ref{RocCurve}
E - F)}.

\subsection{Role of Diet \& Obesity in Gut Microbiome
Connectivity}\label{role-of-diet-obesity-in-gut-microbiome-connectivity}

Diet is a major environmental factor that influences resource
availability and gut microbiome composition and diversity, including
bacteria and phages (14, 56, 57). Previous work in isolated
culture-based systems has suggested that changes in nutrient
availability are associated with altered phage-bacteria network
structures (30), although this has yet to be tested in humans. We
therefore hypothesized that a change in diet would also be associated
with a change in virome-microbiome network structure in the human gut.

We evaluated the diet-associated differences in gut virome-microbiome
network structure by quantifying how central each sample's network was
on average. We accomplished this by utilizing two common centrality
metrics: degree centrality and closeness centrality. Degree centrality,
the simplest centrality metric, was defined as the number of connections
each phage made with each bacterium. We supplemented measurements of
degree centrality with measurements of closeness centrality. Closeness
centrality is a metric of how close each phage or bacterium is to all of
the other phages and bacteria in the network. A higher closeness
centrality suggests that the effects of genetic information or altered
abundance would be more impactful to all other microbes in the system. A
network with higher average closeness centrality also indicates an
overall greater degree of connections, which suggests a greater
resilience against instability. This is because more highly connected
networks are more stable as a result of pathway dependencies and the
unlikelihood that a randomly removed bacteria or phage would cause major
divisions across the network (30, 58). We used this information to
calculate the average connectedness per sample, which was corrected for
the maximum potential degree of connectedness. Unfortunately our dataset
was insufficiently powered to make strong conclusions toward this
hypothesis, but the data still provides an interesting ``case
observation'' that warrants further investigation.

Using our limited sample set, we observed that the gut microbiome
network structures associated with high-fat diets appeared less
connected than those of low-fat diets, although a greater sample size
will be required to more properly evaluate this trend \textbf{(Figure
\ref{dietnetworks} A-B)}. Five subjects were available for use, all of
which had matching bacteria and virome datasets and samples from 8-10
days following the initiation of their diets. High-fat diets appeared to
exhibit reduced degree centrality \textbf{(Figure \ref{dietnetworks}
A)}, suggesting bacteria in high-fat environments were targeted by fewer
phages and that phage tropism was more restricted. High-fat diets also
appeared to exhibit decreased closeness centrality \textbf{(Figure
\ref{dietnetworks} B)}, indicating that bacteria and phages were more
distant from other bacteria and phages in the community. This would make
genetic transfer and altered abundance of a given phage or bacterium
less capable of impacting other bacteria and phages within the network.

In addition to diet, we found preliminary evidence that obesity
influenced network structure. This was done using the three mother
samples available from the twin sample set, all of which had matching
bacteria and phage samples and confirmed BMI information. The
obesity-associated network appeared to have a higher degree centrality
\textbf{(Figure \ref{dietnetworks} C)}, but less closeness centrality
than the healthy-associated networks \textbf{(Figure \ref{dietnetworks}
D)}. These results suggested that the obesity-associated networks are
less connected, having microbes further from all other microbes within
the community. This again comes with the caveat that this is a case
observation with too small of a sample size to make more substantial
claims. This is however an area for future investigation.

\subsection{Individuality of Microbial
Networks}\label{individuality-of-microbial-networks}

Skin and gut community membership and diversity are highly personal,
with people remaining more similar to themselves than to other people
over time (13, 59, 60). We therefore hypothesized that this personal
conservation extended to microbiome network structure. We addressed this
hypothesis by calculating the degree of dissimilarity between each
subject's network, based on phage and bacteria abundance and centrality.
We quantified phage and bacteria centrality within each sample graph
using the weighted eigenvector centrality metric. This metric defines
central phages as those that are highly abundant (\(A_{O}\) as defined
in the methods) and infect many distinct bacteria which themselves are
abundant and infected by many other phages. Similarly, bacterial
centrality was defined as those bacteria that were both abundant and
connected to numerous phages that were themselves connected to many
bacteria. We then calculated the similarity of community networks using
the weighted eigenvector centrality of all nodes between all samples.
Samples with similar network structures were interpreted as having
similar capacities for maintaining stability and transmitting genetic
material.

We used this network dissimilarity metric to test whether microbiome
network structures were more similar within people than between people
over time. We found that gut microbiome network structures clustered by
person (ANOSIM p-value = 0.005, R = 0.958, \textbf{Figure \ref{intradiv}
A}). Network dissimilarity within each person over the 8-10 day sampling
period was less than the average dissimilarity between that person and
others, although this difference was not statistically significant
(p-value = 0.125, \textbf{Figure \ref{intradiv} B}). Four of the five
available subjects were used here because one of the subjects was not
sampled at the initial time point. The lack of statistical confidence
was again due to the small sample size of this dataset.

Although there was evidence for gut network conservation among
individuals, we found no evidence for conservation of gut network
structures within families. The gut network structures were not more
similar within families (twins and their mothers; intrafamily) compared
to other families (other twins and mothers; inter-family) (p-value =
0.312, \textbf{Figure \ref{intradiv} C}). In addition to the gut, skin
microbiome network structure was strongly conserved within individuals
(p-value \textless{} 0.001, \textbf{Figure \ref{intradiv} D}). This
distribution was similar when separated by anatomical sites. Most sites
were statistically significantly more conserved within individuals
\textbf{(Supplemental Figure \ref{allskin})}.

\subsection{Association Between Environmental Stability and Network
Structure Across the Human Skin
Landscape}\label{association-between-environmental-stability-and-network-structure-across-the-human-skin-landscape}

Extensive work has illustrated differences in diversity and composition
of the healthy human skin microbiome between anatomical sites, including
bacteria, virus, and fungal communities (13, 44, 59). These communities
vary by degree of skin moisture, oil, and environmental exposure. As
viruses are known to influence microbial diversity and community
composition, we hypothesized that microbe-virus network structure would
be specific to anatomical sites, as well. To test this, we evaluated the
changes in network structure between anatomical sites within the skin
dataset.

The average centrality of each sample was quantified using the weighted
eigenvector centrality metric. Intermittently moist skin sites (dynamic
sites that fluctuate between being moist and dry) were significantly
less connected than the more stable moist and sebaceous environments
(p-value \textless{} 0.001, \textbf{Figure \ref{skinnetwork} A)}. Also,
skin sites that were occluded from the environment were much more highly
connected than those that were constantly exposed to the environment or
only intermittently occluded (p-value \textless{} 0.001, \textbf{Figure
\ref{skinnetwork} B)}.

To supplement this analysis, we compared the network signatures using
the centrality dissimilarity approach described above. The dissimilarity
between samples was a function of shared relationships, degree of
centrality, and bacteria/phage abundance. When using this supplementary
approach, we found that network structures significantly clustered by
moisture, sebaceous, and intermittently moist status \textbf{(Figure
\ref{skinnetwork} C,E)}. Occluded sites were significantly different
from exposed and intermittently occluded sites, but there was no
difference between exposed and intermittently occluded sites
\textbf{(Figure \ref{skinnetwork} D,F)}. These findings provide further
support that skin microbiome network structure differs significantly
between skin sites.

\section{Discussion}\label{discussion}

Foundational work has provided a baseline understanding of the human
microbiome by characterizing bacterial and viral diversity across the
human body (13, 14, 43--45, 61). Here, we offer an initial understanding
of how phage-bacteria networks differ throughout the human body, so as
to provide a baseline for future studies of how and why microbiome
networks differ in disease states. We developed and implemented a
network-based analytical model to evaluate the basic properties of the
human microbiome through bacteria and phage relationships, instead of
membership or diversity alone. This enabled the application of network
theory to provide a new perspective on complex ecological communities.
We utilized metrics of connectivity to model the extent to which
communities of bacteria and phages interact through mechanisms such as
horizontal gene transfer, modulated bacterial gene expression, and
alterations in abundance.

Just as gut microbiome and virome composition and diversity are
conserved in individuals (13, 43, 44, 60), gut and skin microbiome
network structures were conserved within individuals over time. Gut
network structure was not conserved among family members. These findings
suggested that the community properties inferred from microbiome
interaction network structures, such as stability (meaning a more highly
connected network is more stable because a randomly removed bacteria or
phage node is less likely to divide or disintegrate (30, 58) the overall
network), the potential for horizontal gene transfer between members,
and co-evolution of populations, were person-specific. These properties
may be impacted by personal factors ranging from the body's immune
system to external environmental conditions, such as climate and diet.

We observed evidence supporting the ability of environmental conditions
to shape gut and skin microbiome interaction network structure by
observing that diet and skin location were associated with altered
network structures. We found evidence that diet was sufficient to alter
gut microbiome network connectivity, although this needs to be
interpreted as a case observation, due to the small sample size.
Although our sample size was small, our findings provided some
preliminary evidence that high-fat diets were less connected than
low-fat diets and that high-fat diets therefore may lead to less stable
communities with a decreased ability for microbes to directly influence
one another. We supported this finding with the observation that obesity
may have been associated with decreased network connectivity. Together
these findings suggest the food we eat may not only impact which
microbes colonize our guts, but may also impact their interactions with
infecting phages. Further work will be required to characterize these
relationships with a larger cohort.

In addition to diet, the skin environment also influenced the microbiome
interaction network structure. Network structure differed between
environmentally exposed and occluded skin sites. The sites under greater
environmental fluctuation and exposure (the exposed and intermittently
exposed sites) were less connected and therefore were predicted to have
a higher propensity for instability. Likewise, intermittently moist
sites demonstrated less connectedness than the more stable moist and
sebaceous sites. Together these data suggested that body sites under
greater degrees of fluctuation harbored less connected, potentially less
stable microbiomes. This points to a link between microbiome and
environmental stability and warrants further investigation.

While these findings take us an important step closer to understanding
the microbiome through interspecies relationships, there are caveats to
and considerations regarding the approach. First, as with most
classification models, the infection classification model developed and
applied is only as good as its training set -- in this case, the
collection of experimentally-verified positive and negative infection
data, where genomes of all members are fully sequenced. Large-scale
experimental screens for phage and bacteria infectious interactions that
report high-confidence negative interactions (i.e., no infection) are
desperately needed, as they would provide more robust model training and
improved model performance. Furthermore, just as we have improved on
previous modeling efforts, we expect that new and creative scoring
metrics will be integrated into this model to improve future
performance.

Second, although our analyses utilized the best datasets currently
available for our study, this work was done retrospectively and relied
on existing data up to seven years old. These archived datasets were
limited by the technology and costs of the time. For example, the diet
and twin studies, relied on multiple displacement amplification (MDA) in
their library preparations--an approach used to overcome the large
nucleic acids requirements typical of older sequencing library
generation protocols. It is now known that MDA results in biases in
microbial community composition (62), as well as toward ssDNA viral
genomes (63, 64), thus rendering the resulting microbial and viral
metagenomes largely non-quantitative. Future work that employs larger
sequence datasets and that avoids the use of bias-inducing amplification
steps will build on and validate our findings, as well as inform the
design and interpretation of further studies.

Finally, the networks in this study were built using operational genomic
units (OGUs), which represented groups of highly similar bacteria or
phage genomes or clustered genome fragments. Similar clustering
definition and validation methods, both computational and experimental,
have been implemented in other metagenomic sequencing studies, as well
(42, 65--67). These approaches could offer yet another level of
sophistication to our network-based analyses. While this operationally
defined clustering approach allows us to study whole community networks,
our ability to make conclusions about interactions among specific phage
or bacterial species or populations is inherently limited, compared to
more focused, culture-based studies such as the work by Malki \emph{et
al} (68). Future work must address this limitation, e.g., through
improved binning methods and deeper metagenomic shotgun sequencing, but
most importantly through an improved conceptual framing of what defines
ecologically and evolutionarily cohesive units for both phage and
bacteria (69). Defining operational genomic units and their taxonomic
underpinnings (e.g., whether OGU clusters represent genera or species)
is an active area of work critical to the utility of this approach. As a
first step, phylogenomic analyses have been performed to cluster
cyanophage isolate genomes into informative groups using shared gene
content, average nucleotide identity of shared genes, and pairwise
differences between genomes (70). Such population-genetic assessment of
phage evolution, coupled with the ecological implications of genome
heterogeneity, will inform how to define nodes in future iterations of
the ecological network developed here. Even though we are hesitant to
speculate on phage host ranges at low taxonomic levels in our dataset,
the data does aggree with previous reports of instances of broad phage
host range (68, 71).

Together our work takes an initial step towards defining bacteria-virus
interaction profiles as a characteristic of human-associated microbial
communities. This approach revealed the impacts that different human
environments (e.g., the skin and gut) can have on microbiome
connectivity. By focusing on relationships between bacterial and viral
communities, they are studied as the interacting cohorts they are,
rather than as independent entities. While our developed bacteria-phage
interaction framework is a novel conceptual advance, the microbiome also
consists of archaea and small eukaryotes, including fungi and
\emph{Demodex} mites (1, 72)--all of which can interact with human
immune cells and other non-microbial community members (73). Future work
will build from our approach and include these additional community
members and their diverse interactions and relationships (e.g., beyond
phage-bacteria). This will result in a more robust network and a more
holistic understanding of the evolutionary and ecological processes that
drive the assembly and function of the human-associated microbiome.

\section{Materials \& Methods}\label{materials-methods}

\subsection{Code Availability}\label{code-availability}

A reproducible version of this manuscript written in R markdown and all
of the code used to obtain and process the sequencing data is available
at the following GitHub repository:

https://github.com/SchlossLab/Hannigan\_ConjunctisViribus\_mSystems\_2017

\subsection{Data Acquisition \& Quality
Control}\label{data-acquisition-quality-control}

Raw sequencing data and associated metadata were acquired from the NCBI
sequence read archive (SRA). Supplementary metadata were acquired from
the same SRA repositories and their associated manuscripts. The gut
virome diet study (SRA: \texttt{SRP002424}), twin virome studies (SRA:
\texttt{SRP002523}; \texttt{SRP000319}), and skin virome study (SRA:
\texttt{SRP049645}) were downloaded as \texttt{.sra} files. Sequencing
files were converted to \texttt{fastq} format using the
\texttt{fastq-dump} tool of the NCBI SRA Toolkit (v2.2.0). Sequences
were quality trimmed using the Fastx toolkit (v0.0.14) to exclude bases
with quality scores below 33 and shorter than 75 bp (74). Paired end
reads were filtered to exclude sequences missing their corresponding
pair using the \texttt{get\_trimmed\_pairs.py} script available in the
source code.

\subsection{Contig Assembly}\label{contig-assembly}

Contigs were assembled using the Megahit assembly program (v1.0.6) (75).
A minimum contig length of 1 kb was used. Iterative k-mer stepping began
at a minimum length of 21 and progressed by 20 until 101. All other
default parameters were used.

\subsection{Contig Abundance
Calculations}\label{contig-abundance-calculations}

Contigs were concatenated into two master files prior to alignment, one
for bacterial contigs and one for phage contigs. Sample sequences were
aligned to phage or bacterial contigs using the Bowtie2 global aligner
(v2.2.1) (76). We defined a mismatch threshold of 1 bp and seed length
of 25 bp. Sequence abundance was calculated from the Bowtie2 output
using the \texttt{calculate\_abundance\_from\_sam.pl} script available
in the source code.

\subsection{Operational Genomic Unit
Binning}\label{operational-genomic-unit-binning}

Contigs often represent large fragments of genomes. In order to reduce
redundancy and the resulting artificially inflated genomic richness
within our dataset, it was important to bin contigs into operational
units based on their similarity. This approach is conceptually similar
to the clustering of related 16S rRNA sequences into operational
taxonomic units (OTUs), although here we are clustering contigs into
operational genomic units (OGUs) (61).

Contigs were clustered using the CONCOCT algorithm (v0.4.0) (77).
Because of our large dataset and limits in computational efficiency, we
randomly subsampled the dataset to include 25\% of all samples, and used
these to inform contig abundance within the CONCOCT algorithm. CONCOCT
was used with a maximum of 500 clusters, a k-mer length of four, a
length threshold of 1 kb, 25 iterations, and exclusion of the total
coverage variable.

OGU abundance (\(A_{O}\)) was obtained as the sum of the abundance of
each contig (\(A_{j}\)) associated with that OGU. The abundance values
were length corrected such that:

\[ { A }_{ O }=\frac { { 10 }^{ 7 }\sum _{ j=1 }^{ k }{ { A }_{ j } }  }{ \sum _{ j=1 }^{ k }{ { L }_{ j } }  } \]

Where \texttt{L} is the length of each contig \texttt{j} within the OGU.

\subsection{Operational Genomic Unit
Identification}\label{operational-genomic-unit-identification}

To confirm a lack of phage sequences in the bacterial OGU dataset, we
performed blast nucleotide alignment of the bacterial OGU representative
sequences using an e-value \textless{} \(10^{-25}\), which was stricter
than the \(10^{-10}\) threshold used in the random forest model below,
against all of the phage reference genomes available in the EMBL
database. We used a stricter threshold because we know there are genomic
similarities between bacteria and phage OGUs from the interactive model,
but we were interested in contigs with high enough similarity to
references that they may indeed be from phages. We also performed the
converse analysis of aligning phage OGU representative sequences to EMBL
bacterial reference genomes. Finally, we ran both the phage and bacteria
OGU representative sequences through the Virsorter program (1.0.3) to
identify phages (all default parameters were used), using only those in
the high confidence identification category ``class 1'' (47).

\subsection{Open Reading Frame
Prediction}\label{open-reading-frame-prediction}

Open reading frames (ORFs) were identified using the Prodigal program
(V2.6.2) with the meta mode parameter and default settings (78).

\subsection{Classification Model Creation and
Validation}\label{classification-model-creation-and-validation}

The classification model for predicting interactions was built using
experimentally validated bacteria-phage infections or validated lack of
infections from six studies (41, 48--52). Associated reference genomes
were downloaded from the European Bioinformatics Institute (see details
in source code). The model was created based on the four metrics listed
below.

The four scores were used as parameters in a random forest model to
classify bacteria and bacteriophage pairs as either having infectious
interactions or not. The classification model was built using the Caret
R package (v6.0.73) (79). The model was trained using five-fold cross
validation with ten repeats. Pairs without scores were classified as not
interacting. The model was optimized using the ROC value. The resulting
model performance was plotted using the plotROC R package.

\subparagraph{Identify Bacterial CRISPRs Targeting
Phages}\label{identify-bacterial-crisprs-targeting-phages}

Clustered Regularly Interspaced Short Palindromic Repeats (CRISPRs) were
identified from bacterial genomes using the PilerCR program (v1.06)
(80). Resulting spacer sequences were filtered to exclude spacers
shorter than 20 bp and longer than 65 bp. Spacer sequences were aligned
to the phage genomes using the nucleotide BLAST algorithm with default
parameters (v2.4.0) (81). The mean percent identity for each matching
pair was recorded for use in our classification model.

\subparagraph{Detect Matching Prophages within Bacterial
Genomes}\label{detect-matching-prophages-within-bacterial-genomes}

Temperate bacteriophages infect and integrate into their bacterial
host's genome. We detected integrated phage elements within bacterial
genomes by aligning phage genomes to bacterial genomes using the
nucleotide BLAST algorithm and a minimum e-value of 1e-10. The resulting
bitscore of each alignment was recorded for use in our classification
model.

\subparagraph{Identify Shared Genes Between Bacteria and
Phages}\label{identify-shared-genes-between-bacteria-and-phages}

As a result of gene transfer or phage genome integration during
infection, phages may share genes with their bacterial hosts, providing
us with evidence of phage-host pairing. We identified shared genes
between bacterial and phage genomes by assessing amino acid similarity
between the genes using the Diamond protein alignment algorithm
(v0.7.11.60) (82). The mean alignment bitscores for each genome pair
were recorded for use in our classification model.

\subparagraph{Protein - Protein
Interactions}\label{protein---protein-interactions}

The final method used for predicting infectious interactions between
bacteria and phages was the detection of pairs of genes whose proteins
are known to interact. We assigned bacterial and phage genes to protein
families by aligning them to the Pfam database using the Diamond protein
alignment algorithm. We then identified which pairs of proteins were
predicted to interact using the Pfam interaction information within the
Intact database (53). The mean bitscores of the matches between each
pair were recorded for use in the classification model.

\subsection{Interaction Network
Construction}\label{interaction-network-construction}

The bacteria and phage operational genomic units (OGUs) were scored
using the same approach as outlined above. The infectious pairings
between bacteria and phage OGUs were classified using the random forest
model described above. The predicted infectious pairings and all
associated metadata were used to populate a graph database using Neo4j
graph database software (v2.3.1) (83). This network was used for
downstream community analysis.

\subsection{Centrality Analysis}\label{centrality-analysis}

We quantified the centrality of graph vertices using three different
metrics, each of which provided different information graph structure.
When calculating these values, let \(G(V,E)\) be an undirected,
unweighted graph with \(|V|=n\) nodes and \(|E|=m\) edges. Also, let
\(\mathbf{A}\) be its corresponding adjacency matrix with entries
\(a_{ij} = 1\) if nodes \(V_i\) and \(V_j\) are connected via an edge,
and \(a_{ij} = 0\) otherwise.

Briefly, the \textbf{closeness centrality} of node \(V_i\) is calculated
taking the inverse of the average length of the shortest paths
(\texttt{d}) between nodes \(V_i\) and all the other nodes \(V_j\).
Mathematically, the closeness centrality of node \(V_i\) is given as:

\[ { C }_{ C }\left( { V }_{ i } \right) ={ \left( \sum _{ j=1 }^{ n }{ d\left( { V }_{ i },{ V }_{ j } \right)  }  \right)  }^{ -1 } \]

The distance between nodes (\texttt{d}) was calculated as the shortest
number of edges required to be traversed to move from one node to
another.

Intuitively, the \textbf{degree centrality} of node \(V_i\) is defined
as the number of edges that are incident to that node:

\[ { C }_{ D }\left( { V }_{ i } \right) =\sum _{ j=1 }^{ n }{ { a }_{ ij } } \]

where \(a_{ij}\) is the \(ij^{th}\) entry in the adjacency matrix
\(\mathbf{A}\).

The eigenvector centrality of node \(V_i\) is defined as the \(i^{th}\)
value in the first eigenvector of the associated adjacency matrix
\(\mathbf{A}\). Conceptually, this function results in a centrality
value that reflects the connections of the vertex, as well as the
centrality of its neighboring vertices.

The \textbf{centralization} metric was used to assess the average
centrality of each sample graph \(\mathbf{G}\). Centralization was
calculated by taking the sum of each vertex \(V_{i}\)'s centrality from
the graph maximum centrality \(C_{w}\), such that:

\[ C\left( G \right) =\frac { \sum _{ i=1 }^{ n }{ Cw -c\left( { V }_{ i } \right)  }  }{ { T } }  \]

The values were corrected for uneven graph sizes by dividing the
centralization score by the maximum theoretical centralization
(\texttt{T}) for a graph with the same number of vertices.

Degree and closeness centrality were calculated using the associated
functions within the igraph R package (v1.0.1) (84).

\subsection{Network Relationship
Dissimilarity}\label{network-relationship-dissimilarity}

We assessed similarity between graphs by evaluating the shared
centrality of their vertices, as has been done previously. More
specifically, we calculated the dissimilarity between graphs \(G_{i}\)
and \(G_{j}\) using the Bray-Curtis dissimilarity metric and eigenvector
centrality values such that:

\[ B\left( { G }_{ i },{ G }_{ j } \right) =1-\frac { 2{ C }_{ ij } }{ { C }_{ i }+{ C }_{ j } } \]

Where \(C_{ij}\) is the sum of the lesser centrality values for those
vertices shared between graphs, and \(C_{i}\) and \(C_{j}\) are the
total number of vertices found in each graph. This allows us to
calculate the dissimilarity between graphs based on the shared
centrality values between the two graphs.

\subsection{Statistics and
Comparisons}\label{statistics-and-comparisons}

Differences in intrapersonal and interpersonal network structure
diversity, based on multivariate data, were calculated using an analysis
of similarity (ANOSIM). Statistical significance of univariate
Eigenvector centrality differences were calculated using a paired
Wilcoxon test.

Statistical significance of differences in univariate eigenvector
centrality measurements of skin virome-microbiome networks were
calculated using a pairwise Wilcoxon test, corrected for multiple
hypothesis tests using the Holm correction method. Multivariate
eigenvector centrality was measured as the mean differences between
cluster centroids, with statistical significance measured using an ANOVA
and post hoc Tukey test.

\section{Acknowledgments}\label{acknowledgments}

We thank the members of the Schloss lab for their underlying
contributions. We thank the authors of the original studies for making
their data and metadata publicly available and understandable. We also
thank the participants in the studies.

\section{Funding Information}\label{funding-information}

GDH was supported in part by the Molecular Mechanisms in Microbial
Pathogenesis Training Program (T32 AI007528). GDH and PDS were supported
in part by funding from the NIH (P30DK034933, U19AI09087, and
U01AI124255).

\section{Disclosure Declaration}\label{disclosure-declaration}

The authors report no conflicts of interest.

\newpage

\section{Figures}\label{figures}

\begin{figure}[htbp]
\centering
\includegraphics[width=0.85000\textwidth]{../figures/rocCurves.pdf}
\caption{\textbf{Summary of Multi-Study Network Model.} \emph{(A)
Average ROC curve used to create the microbiome-virome infection
prediction model. (B) Importance scores associated with the metrics used
in the random forest model to predict relationships between bacteria and
phages. The importance score is defined as the mean decrease in accuracy
of the model when a feature (e.g.~Pfam) is excluded. Features include
the local gene alignments between bacteria and phage genes (denoted
\texttt{blastx}; the blastx algorithm in Diamond aligner), local genome
nucleotide alignments between bacteria and phage OGUs, presence of
experimentally validated protein family domains (Pfams) between phage
and bacteria OGUs, and CRISPR targeting of bacteria toward phages
(CRISPR). (C) Proportions of samples included (gray) and excluded (red)
in the model. Samples were excluded from the model because they did not
yield any scores. Those interactions without scores were automatically
classified as not having interactions. (D) Bipartite visualization of
the resulting phage-bacteria network. Phage OGUs are presented in
orange, bacteria OGUs in red, and their interaction edges are
represented as connecting lines. This network includes information from
all three published studies. (E) Network diameter (measure of graph
size; the greatest number of traversed vertices required between two
vertices), (F) number of vertices, and (G) number of edges
(relationships) for the total network (orange) and the individual study
sub-networks (diet study = red, skin study = yellow, twin study =
green).} \label{RocCurve}}
\end{figure}

\newpage

\begin{figure}[htbp]
\centering
\includegraphics[width=0.90000\textwidth]{../figures/dietnetworks.pdf}
\caption{\textbf{Impact of Diet and Obesity on Gut Network Structure.}
\emph{(A) Quantification of average degree centrality (number of edges
per node) and (B) closeness centrality (average distance from each node
to every other node) of gut microbiome networks of subjects limited to
exclusively high-fat or low-fat diets. Each point represents the
centrality from a human subject stool sample that was collected 8-10
days following the beginning of their defined diet. There are five
samples here, compared to the four in figure \ref{intradiv}, because one
of the was only sampled post-diet, providing us data for this analysis
but not allowing us to compare to a baseline for figure \ref{intradiv}.
(C) Quantification of average degree centrality and (D) closeness
centrality between obese and healthy adult women from the Twin gut
study. Each point represents a stool sample taken from one of the three
adult woman confirmed as obese or healthy and with matching virus and
bacteria data.} \label{dietnetworks}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[width=0.90000\textwidth]{../figures/intrapersonal_diversity.pdf}
\caption{\textbf{Intrapersonal vs Interpersonal Network Dissimilarity
Across Different Human Systems.} \emph{(A) NMDS ordination illustrating
network dissimilarity between subjects over time. Each sample is colored
by subject, with each colored sample pair collected 8-10 days apart.
Dissimilarity was calculated using the Bray-Curtis metric based on
abundance weighted eigenvector centrality signatures, with a greater
distance representing greater dissimilarity in bacteria and phage
centrality and abundance. Only four subjects were included here,
compared to the five used in figure \ref{dietnetworks}, because one of
the subjects was missing the inital sampling time point and therefore
lacked temporal sampling. (B) Quantification of gut network
dissimilarity within the same subject over time (intrapersonal) and the
mean dissimilarity between the subject of interest and all other
subjects (interpersonal). The p-value is provided near the bottom of the
figure. (C) Quantification of gut network dissimilarity within subjects
from the same family (intrafamily) and the mean dissimilarity between
subjects within a family and those of other families (interfamily). Each
point represents the inter-family and intra-family dissimilarity of a
twin or mother that was sampled over time. (D) Quantification of skin
network dissimilarity within the same subject and anatomical location
over time (intrapersonal) and the mean dissimilarity between the subject
of interest and all other subjects at the same time and the same
anatomical location (interpersonal). All p-values were calculated using
a paired Wilcoxon test.}\label{intradiv}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[width=0.75000\textwidth]{../figures/skinplotresults.pdf}
\caption{\textbf{Impact of Skin Micro-Environment on Microbiome Network
Structure.} \emph{(A) Notched box-plot depicting differences in average
eigenvector centrality between moist, intermittently moist, and
sebaceous skin sites and (B) occluded, intermittently occluded, and
exposed sites. Notched box-plots were created using ggplot2 and show the
median (center line), the inter-quartile range (IQR; upper and lower
boxes), the highest and lowest value within 1.5 * IQR (whiskers),
outliers (dots), and the notch which provides an approximate 95\%
confidence interval as defined by 1.58 * IQR / sqrt(n). Sample sizes for
each group were: Moist = 81, Sebaceous = 56, IntMoist = 56, Occluded =
106, Exposed = 61, IntOccluded = 26. (C) NMDS ordination depicting the
differences in skin microbiome network structure between skin moisture
levels and (D) occlusion. Samples are colored by their environment and
their dissimilarity to other samples was calculated as described in
figure \ref{intradiv}. (E) The statistical differences of networks
between moisture and (F) occlusion status were quantified with an anova
and post hoc Tukey test. Cluster centroids are represented by dots and
the extended lines represent the associated 95\% confidence intervals.
Significant comparisons (p-value \textless{} 0.05) are colored in red,
and non-significant comparisons are gray.}\label{skinnetwork}}
\end{figure}

\newpage

\section{Supplemental Figures}\label{supplemental-figures}

\beginsupplement

\begin{figure}[htbp]
\centering
\includegraphics[width=0.80000\textwidth]{../figures/SequenceAbund.pdf}
\caption{\textbf{Sequencing Depth Summary.} \emph{Number of sequences
that aligned to (A) Phage and (B) Bacteria operational genomic units per
sample and colored by study.}\label{SequenceStats}}
\end{figure}

\newpage

\begin{figure}[htbp]
\centering
\includegraphics[width=0.50000\textwidth]{../figures/ContigStats.pdf}
\caption{\textbf{Contig Summary Statistics.} \emph{Scatter plot heat map
with each hexagon representing the abundance of contigs. Contigs are
organized by length on the x-axis and the number of aligned sequences on
the y-axis.}\label{ContigStats}}
\end{figure}

\newpage

\begin{figure}[htbp]
\centering
\includegraphics[width=0.80000\textwidth]{../figures/ClusterStats.pdf}
\caption{\textbf{Operational Genomic Unit Summary Statistics.}
\emph{Scatter plot with operational genomic unit clusters organized by
average contig length within the cluster on the x-axis and the number of
contigs in the cluster on the y-axis. Operational genomic units of (A)
bacteriophages and (B) bacteria are shown.}\label{ClusterStats}}
\end{figure}

\newpage

\begin{figure}[htbp]
\centering
\includegraphics[width=1.00000\textwidth]{../figures/BenchmarkDataset.pdf}
\caption{\textbf{Summary information of validation dataset used in the
interaction predictive model.} \emph{A) Categorical heat-map
highlighting the experimentally validated positive and negative
interactions. Only bacteria species are shown, which represent multiple
reference strains. Phages are labeled on the x-axis and bacteria are
labeled on the y-axis. B) Quantification of bacterial host strains known
to exist for each phage. C) Genome strandedness and D) linearity of the
phage reference genomes used for the
dataset.}\label{ValidationOverview}}
\end{figure}

\newpage

\begin{figure}[htbp]
\centering
\includegraphics[width=0.50000\textwidth]{../figures/graphdatabasediagram.pdf}
\caption{\textbf{Structure of the interactive network.} \emph{Metadata
relationships to samples (Phage Sample ID and Bacteria Sample ID)
included the associated time point, the study, the subject the sample
was taken from, and the associated disease. Infectious interactions were
recorded between phage and bacteria operational genomic units (OGUs).
Sequence count abundance for each OGU within each sample was also
recorded.}\label{NetworkDiagram}}
\end{figure}

\newpage

\begin{figure}[htbp]
\centering
\includegraphics[width=0.75000\textwidth]{../figures/intraallskin.pdf}
\caption{\textbf{Intrapersonal vs Interpersonal Dissimilarity of the
Skin.} \emph{Quantification of skin network dissimilarity within the
same subject and anatomical location over time (intrapersonal) and the
mean dissimilarity between the subject of interest and all other
subjects at the same time and the same anatomical location
(interpersonal), separated by each anatomical site (forehead {[}Fh{]},
palm {[}Pa{]}, toe web {[}Tw{]}, umbilicus {[}Um{]}, antecubital fossa
{[}Ac{]}, axilla {[}Ax{]}, and retroauricular crease {[}Ra{]}). P-value
was calculated using a paired Wilcoxon test.}\label{allskin}}
\end{figure}

\newpage

\section{Supplemental Tables}\label{supplemental-tables}

\begin{longtable}[]{@{}lll@{}}
\caption{Summary of the primary quality control measures reported in the
original publications of the viromes used in this current
study.}\tabularnewline
\toprule
\begin{minipage}[b]{0.24\columnwidth}\raggedright\strut
Study
\strut\end{minipage} &
\begin{minipage}[b]{0.17\columnwidth}\raggedright\strut
Citation
\strut\end{minipage} &
\begin{minipage}[b]{0.50\columnwidth}\raggedright\strut
Virome Quality Control Measures
\strut\end{minipage}\tabularnewline
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{0.24\columnwidth}\raggedright\strut
Study
\strut\end{minipage} &
\begin{minipage}[b]{0.17\columnwidth}\raggedright\strut
Citation
\strut\end{minipage} &
\begin{minipage}[b]{0.50\columnwidth}\raggedright\strut
Virome Quality Control Measures
\strut\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
Diet \& the Gut Virome
\strut\end{minipage} &
\begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
Minot, 2011
\strut\end{minipage} &
\begin{minipage}[t]{0.50\columnwidth}\raggedright\strut
• 16S rRNA gene qPCR revealed reduction in bacterial DNA of at least
10,000X.
\strut\end{minipage}\tabularnewline
\begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
\strut\end{minipage} &
\begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
\strut\end{minipage} &
\begin{minipage}[t]{0.50\columnwidth}\raggedright\strut
• Alignment of shotgun sequences revealed 35X reduction in 16S rRNA gene
alignments in virome compared to bacteria shotgun.
\strut\end{minipage}\tabularnewline
\begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
\strut\end{minipage} &
\begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
\strut\end{minipage} &
\begin{minipage}[t]{0.50\columnwidth}\raggedright\strut
• Electron microscopy and nucleic acid stain techniques visually
confirmed lack of bacteria in virome samples.
\strut\end{minipage}\tabularnewline
\begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
Skin Virome
\strut\end{minipage} &
\begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
Hannigan, 2015
\strut\end{minipage} &
\begin{minipage}[t]{0.50\columnwidth}\raggedright\strut
• Significant reduction in reads mapping to 16S rRNA gene sequence,
compared to bacteria shotgun dataset.
\strut\end{minipage}\tabularnewline
\begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
\strut\end{minipage} &
\begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
\strut\end{minipage} &
\begin{minipage}[t]{0.50\columnwidth}\raggedright\strut
• Significant reduction in reads mapping to human genome, compared to
bacteria shotgun dataset.
\strut\end{minipage}\tabularnewline
\begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
\strut\end{minipage} &
\begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
\strut\end{minipage} &
\begin{minipage}[t]{0.50\columnwidth}\raggedright\strut
• Average viral relative abundance of 0.4\% in bacterial shotgun
dataset.
\strut\end{minipage}\tabularnewline
\begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
Twin Gut Virome
\strut\end{minipage} &
\begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
Reyes, 2010
\strut\end{minipage} &
\begin{minipage}[t]{0.50\columnwidth}\raggedright\strut
• Confirmation that 2.5\% of bacterial shotgun reads mapped to virome,
and 76\% of virome reads matched the shotgun 2.5\%.
\strut\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

\newpage

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\hypertarget{ref-Hannigan:2013im}{}
1. \textbf{Hannigan GD}, \textbf{Grice EA}. 2013. Microbial Ecology of
the Skin in the Era of Metagenomics and Molecular Microbiology. Cold
Spring Harbor Perspectives in Medicine \textbf{3}:a015362--a015362.

\hypertarget{ref-Hannigan:2014be}{}
2. \textbf{Hannigan GD}, \textbf{Hodkinson BP}, \textbf{McGinnis K},
\textbf{Tyldsley AS}, \textbf{Anari JB}, \textbf{Horan AD},
\textbf{Grice EA}, \textbf{Mehta S}. 2014. Culture-independent pilot
study of microbiota colonizing open fractures and association with
severity, mechanism, location, and complication from presentation to
early outpatient follow-up. Journal of Orthopaedic Research
\textbf{32}:597--605.

\hypertarget{ref-Loesche:2016ev}{}
3. \textbf{Loesche M}, \textbf{Gardner SE}, \textbf{Kalan L},
\textbf{Horwinski J}, \textbf{Zheng Q}, \textbf{Hodkinson BP},
\textbf{Tyldsley AS}, \textbf{Franciscus CL}, \textbf{Hillis SL},
\textbf{Mehta S}, \textbf{Margolis DJ}, \textbf{Grice EA}. 2016.
Temporal stability in chronic wound microbiota is associated with poor
healing. Journal of Investigative Dermatology.

\hypertarget{ref-He:2016ch}{}
4. \textbf{He Q}, \textbf{Li X}, \textbf{Liu C}, \textbf{Su L},
\textbf{Xia Z}, \textbf{Li X}, \textbf{Li Y}, \textbf{Li L}, \textbf{Yan
T}, \textbf{Feng Q}, \textbf{Xiao L}. 2016. Dysbiosis of the fecal
microbiota in the TNBS-induced Crohn's disease mouse model. Applied
Microbiology and Biotechnology 1--10.

\hypertarget{ref-Norman:2015kb}{}
5. \textbf{Norman JM}, \textbf{Handley SA}, \textbf{Baldridge MT},
\textbf{Droit L}, \textbf{Liu CY}, \textbf{Keller BC}, \textbf{Kambal
A}, \textbf{Monaco CL}, \textbf{Zhao G}, \textbf{Fleshner P},
\textbf{Stappenbeck TS}, \textbf{McGovern DPB}, \textbf{Keshavarzian A},
\textbf{Mutlu EA}, \textbf{Sauk J}, \textbf{Gevers D}, \textbf{Xavier
RJ}, \textbf{Wang D}, \textbf{Parkes M}, \textbf{Virgin HW}. 2015.
Disease-specific alterations in the enteric virome in inflammatory bowel
disease. Cell \textbf{160}:447--460.

\hypertarget{ref-Seekatz:2016fz}{}
6. \textbf{Seekatz AM}, \textbf{Rao K}, \textbf{Santhosh K},
\textbf{Young VB}. 2016. Dynamics of the fecal microbiome in patients
with recurrent and nonrecurrent Clostridium difficile infection. Genome
medicine \textbf{8}:47.

\hypertarget{ref-Zackular:2014fba}{}
7. \textbf{Zackular JP}, \textbf{Rogers MAM}, \textbf{Ruffin MT},
\textbf{Schloss PD}. 2014. The human gut microbiome as a screening tool
for colorectal cancer. Cancer prevention research (Philadelphia, Pa)
\textbf{7}:1112--1121.

\hypertarget{ref-Baxter:2014hb}{}
8. \textbf{Baxter NT}, \textbf{Zackular JP}, \textbf{Chen GY},
\textbf{Schloss PD}. 2014. Structure of the gut microbiome following
colonization with human feces determines colonic tumor burden.
Microbiome \textbf{2}:20.

\hypertarget{ref-Manrique:2016dx}{}
9. \textbf{Manrique P}, \textbf{Bolduc B}, \textbf{Walk ST},
\textbf{Oost J van der}, \textbf{Vos WM de}, \textbf{Young MJ}. 2016.
Healthy human gut phageome. Proceedings of the National Academy of
Sciences of the United States of America 201601060.

\hypertarget{ref-Ly:2014ew}{}
10. \textbf{Ly M}, \textbf{Abeles SR}, \textbf{Boehm TK},
\textbf{Robles-Sikisaka R}, \textbf{Naidu M}, \textbf{Santiago-Rodriguez
T}, \textbf{Pride DT}. 2014. Altered Oral Viral Ecology in Association
with Periodontal Disease. mBio \textbf{5}:e01133--14--e01133--14.

\hypertarget{ref-Modi:2013fia}{}
11. \textbf{Modi SR}, \textbf{Lee HH}, \textbf{Spina CS},
\textbf{Collins JJ}. 2013. Antibiotic treatment expands the resistance
reservoir and ecological network of the phage metagenome. Nature
\textbf{499}:219--222.

\hypertarget{ref-Monaco:2016ita}{}
12. \textbf{Monaco CL}, \textbf{Gootenberg DB}, \textbf{Zhao G},
\textbf{Handley SA}, \textbf{Ghebremichael MS}, \textbf{Lim ES},
\textbf{Lankowski A}, \textbf{Baldridge MT}, \textbf{Wilen CB},
\textbf{Flagg M}, \textbf{Norman JM}, \textbf{Keller BC},
\textbf{Luévano JM}, \textbf{Wang D}, \textbf{Boum Y}, \textbf{Martin
JN}, \textbf{Hunt PW}, \textbf{Bangsberg DR}, \textbf{Siedner MJ},
\textbf{Kwon DS}, \textbf{Virgin HW}. 2016. Altered Virome and Bacterial
Microbiome in Human Immunodeficiency Virus-Associated Acquired
Immunodeficiency Syndrome. Cell Host and Microbe \textbf{19}:311--322.

\hypertarget{ref-Hannigan:2015fz}{}
13. \textbf{Hannigan GD}, \textbf{Meisel JS}, \textbf{Tyldsley AS},
\textbf{Zheng Q}, \textbf{Hodkinson BP}, \textbf{SanMiguel AJ},
\textbf{Minot S}, \textbf{Bushman FD}, \textbf{Grice EA}. 2015. The
Human Skin Double-Stranded DNA Virome: Topographical and Temporal
Diversity, Genetic Enrichment, and Dynamic Associations with the Host
Microbiome. mBio \textbf{6}:e01578--15.

\hypertarget{ref-Minot:2011ez}{}
14. \textbf{Minot S}, \textbf{Sinha R}, \textbf{Chen J}, \textbf{Li H},
\textbf{Keilbaugh SA}, \textbf{Wu GD}, \textbf{Lewis JD},
\textbf{Bushman FD}. 2011. The human gut virome: Inter-individual
variation and dynamic response to diet. Genome Research
\textbf{21}:1616--1625.

\hypertarget{ref-SantiagoRodriguez:2015gd}{}
15. \textbf{Santiago-Rodriguez TM}, \textbf{Ly M}, \textbf{Bonilla N},
\textbf{Pride DT}. 2015. The human urine virome in association with
urinary tract infections. Frontiers in Microbiology \textbf{6}:14.

\hypertarget{ref-Abeles:2015dy}{}
16. \textbf{Abeles SR}, \textbf{Ly M}, \textbf{Santiago-Rodriguez TM},
\textbf{Pride DT}. 2015. Effects of Long Term Antibiotic Therapy on
Human Oral and Fecal Viromes. PLOS ONE \textbf{10}:e0134941.

\hypertarget{ref-Abeles:2014kj}{}
17. \textbf{Abeles SR}, \textbf{Robles-Sikisaka R}, \textbf{Ly M},
\textbf{Lum AG}, \textbf{Salzman J}, \textbf{Boehm TK}, \textbf{Pride
DT}. 2014. Human oral viruses are personal, persistent and
gender-consistent 1--15.

\hypertarget{ref-Haerter:2014ii}{}
18. \textbf{Haerter JO}, \textbf{Mitarai N}, \textbf{Sneppen K}. 2014.
Phage and bacteria support mutual diversity in a narrowing staircase of
coexistence. The ISME Journal \textbf{8}:2317--2326.

\hypertarget{ref-Lindell:2005gz}{}
19. \textbf{Lindell D}, \textbf{Jaffe JD}, \textbf{Johnson ZI},
\textbf{Church GM}, \textbf{Chisholm SW}. 2005. Photosynthesis genes in
marine viruses yield proteins during host infection. Nature
\textbf{438}:86--89.

\hypertarget{ref-Tyler:2013fl}{}
20. \textbf{Tyler JS}, \textbf{Beeri K}, \textbf{Reynolds JL},
\textbf{Alteri CJ}, \textbf{Skinner KG}, \textbf{Friedman JH},
\textbf{Eaton KA}, \textbf{Friedman DI}. 2013. Prophage induction is
enhanced and required for renal disease and lethality in an EHEC mouse
model. PLoS Pathogens \textbf{9}:e1003236.

\hypertarget{ref-Hargreaves:2014ja}{}
21. \textbf{Hargreaves KR}, \textbf{Kropinski AM}, \textbf{Clokie MR}.
2014. Bacteriophage behavioral ecology: How phages alter their bacterial
host's habits. Bacteriophage \textbf{4}:e29866.

\hypertarget{ref-Moon:2015fa}{}
22. \textbf{Moon BY}, \textbf{Park JY}, \textbf{Hwang SY},
\textbf{Robinson DA}, \textbf{Thomas JC}, \textbf{Fitzgerald JR},
\textbf{Park YH}, \textbf{Seo KS}. 2015. Phage-mediated horizontal
transfer of a Staphylococcus aureus virulence-associated genomic island.
Scientific Reports \textbf{5}:9784.

\hypertarget{ref-Modi:2013fi}{}
23. \textbf{Modi SR}, \textbf{Lee HH}, \textbf{Spina CS},
\textbf{Collins JJ}. 2013. Antibiotic treatment expands the resistance
reservoir and ecological network of the phage metagenome. Nature
\textbf{499}:219--222.

\hypertarget{ref-Ogg:1981th}{}
24. \textbf{Ogg JE}, \textbf{Timme TL}, \textbf{Alemohammad MM}. 1981.
General Transduction in Vibrio cholerae. Infection and Immunity
\textbf{31}:737--741.

\hypertarget{ref-Frost:2005dn}{}
25. \textbf{Frost LS}, \textbf{Leplae R}, \textbf{Summers AO},
\textbf{Toussaint A}. 2005. Mobile genetic elements: the agents of open
source evolution. Nature Reviews Microbiology \textbf{3}:722--732.

\hypertarget{ref-Koskella:2014ds}{}
26. \textbf{Koskella B}, \textbf{Brockhurst MA}. 2014. Bacteria-phage
coevolution as a driver of ecological and evolutionary processes in
microbial communities. FEMS Microbiology Reviews \textbf{38}:916--931.

\hypertarget{ref-Jover:2014gq}{}
27. \textbf{Jover LF}, \textbf{Effler TC}, \textbf{Buchan A},
\textbf{Wilhelm SW}, \textbf{Weitz JS}. 2014. The elemental composition
of virus particles: implications for marine biogeochemical cycles.
Nature Reviews Microbiology \textbf{12}:519--528.

\hypertarget{ref-Harcombe:2005fd}{}
28. \textbf{Harcombe WR}, \textbf{Bull JJ}. 2005. Impact of phages on
two-species bacterial communities. Applied and Environmental
Microbiology \textbf{71}:5254--5259.

\hypertarget{ref-Middelboe:2001fl}{}
29. \textbf{Middelboe M}, \textbf{Hagström A}, \textbf{Blackburn N},
\textbf{Sinn B}, \textbf{Fischer U}, \textbf{Borch NH}, \textbf{Pinhassi
J}, \textbf{Simu K}, \textbf{Lorenz MG}. 2001. Effects of Bacteriophages
on the Population Dynamics of Four Strains of Pelagic Marine Bacteria.
Microbial Ecology \textbf{42}:395--406.

\hypertarget{ref-Poisot:2011jc}{}
30. \textbf{Poisot T}, \textbf{Lepennetier G}, \textbf{Martinez E},
\textbf{Ramsayer J}, \textbf{Hochberg ME}. 2011. Resource availability
affects the structure of a natural bacteriabacteriophage community.
Biology letters \textbf{7}:201--204.

\hypertarget{ref-Thompson:2012ki}{}
31. \textbf{Thompson RM}, \textbf{Brose U}, \textbf{Dunne JA},
\textbf{Hall RO}, \textbf{Hladyz S}, \textbf{Kitching RL},
\textbf{Martinez ND}, \textbf{Rantala H}, \textbf{Romanuk TN},
\textbf{Stouffer DB}, \textbf{Tylianakis JM}. 2012. Food webs:
reconciling the structure and function of biodiversity. Trends in
ecology \& evolution \textbf{27}:689--697.

\hypertarget{ref-Moebus:1981kp}{}
32. \textbf{Moebus K}, \textbf{Nattkemper H}. 1981. Bacteriophage
sensitivity patterns among bacteria isolated from marine waters.
Helgoländer Meeresuntersuchungen \textbf{34}:375--385.

\hypertarget{ref-Flores:2013hc}{}
33. \textbf{Flores CO}, \textbf{Valverde S}, \textbf{Weitz JS}. 2013.
Multi-scale structure and geographic drivers of cross-infection within
marine bacteria and phages. The ISME Journal \textbf{7}:520--532.

\hypertarget{ref-Poisot:2012fh}{}
34. \textbf{Poisot T}, \textbf{Canard E}, \textbf{Mouillot D},
\textbf{Mouquet N}, \textbf{Gravel D}. 2012. The dissimilarity of
species interaction networks. Ecology letters \textbf{15}:1353--1361.

\hypertarget{ref-Poisot071993}{}
35. \textbf{Poisot T}, \textbf{Stouffer D}. 2016. How ecological
networks evolve. bioRxiv.

\hypertarget{ref-Flores:2011bh}{}
36. \textbf{Flores CO}, \textbf{Meyer JR}, \textbf{Valverde S},
\textbf{Farr L}, \textbf{Weitz JS}. 2011. Statistical structure of
host-phage interactions. Proceedings of the National Academy of Sciences
of the United States of America \textbf{108}:E288--97.

\hypertarget{ref-Jover:2015ev}{}
37. \textbf{Jover LF}, \textbf{Flores CO}, \textbf{Cortez MH},
\textbf{Weitz JS}. 2015. Multiple regimes of robust patterns between
network structure and biodiversity. Scientific Reports \textbf{5}:17856.

\hypertarget{ref-Reyes:2010cwa}{}
38. \textbf{Reyes A}, \textbf{Haynes M}, \textbf{Hanson N},
\textbf{Angly FE}, \textbf{Heath AC}, \textbf{Rohwer F}, \textbf{Gordon
JI}. 2010. Viruses in the faecal microbiota of monozygotic twins and
their mothers. Nature \textbf{466}:334--338.

\hypertarget{ref-Turnbaugh:2009ei}{}
39. \textbf{Turnbaugh PJ}, \textbf{Hamady M}, \textbf{Yatsunenko T},
\textbf{Cantarel BL}, \textbf{Duncan A}, \textbf{Ley RE}, \textbf{Sogin
ML}, \textbf{Jones WJ}, \textbf{Roe BA}, \textbf{Affourtit JP},
\textbf{Egholm M}, \textbf{Henrissat B}, \textbf{Heath AC},
\textbf{Knight R}, \textbf{Gordon JI}. 2009. A core gut microbiome in
obese and lean twins. Nature \textbf{457}:480--484.

\hypertarget{ref-LimaMendez:2015hw}{}
40. \textbf{Lima-Mendez G}, \textbf{Faust K}, \textbf{Henry N},
\textbf{Decelle J}, \textbf{Colin S}, \textbf{Carcillo F},
\textbf{Chaffron S}, \textbf{Ignacio-Espinosa JC}, \textbf{Roux S},
\textbf{Vincent F}, \textbf{Bittner L}, \textbf{Darzi Y}, \textbf{Wang
J}, \textbf{Audic S}, \textbf{Berline L}, \textbf{Bontempi G},
\textbf{Cabello AM}, \textbf{Coppola L}, \textbf{Cornejo-Castillo FM},
\textbf{d'Ovidio F}, \textbf{De Meester L}, \textbf{Ferrera I},
\textbf{Garet-Delmas M-J}, \textbf{Guidi L}, \textbf{Lara E},
\textbf{Pesant S}, \textbf{Royo-Llonch M}, \textbf{Salazar G},
\textbf{Sánchez P}, \textbf{Sebastian M}, \textbf{Souffreau C},
\textbf{Dimier C}, \textbf{Picheral M}, \textbf{Searson S},
\textbf{Kandels-Lewis S}, \textbf{Tara Oceans Coordinators},
\textbf{Gorsky G}, \textbf{Not F}, \textbf{Ogata H}, \textbf{Speich S},
\textbf{Stemmann L}, \textbf{Weissenbach J}, \textbf{Wincker P},
\textbf{Acinas SG}, \textbf{Sunagawa S}, \textbf{Bork P},
\textbf{Sullivan MB}, \textbf{Karsenti E}, \textbf{Bowler C},
\textbf{Vargas C de}, \textbf{Raes J}. 2015. Ocean plankton.
Determinants of community structure in the global plankton interactome.
Science \textbf{348}:1262073--1262073.

\hypertarget{ref-Edwards:2015iz}{}
41. \textbf{Edwards RA}, \textbf{McNair K}, \textbf{Faust K},
\textbf{Raes J}, \textbf{Dutilh BE}. 2015. Computational approaches to
predict bacteriophage-host relationships. FEMS Microbiology Reviews
\textbf{40}:258--272.

\hypertarget{ref-Roux:2016cc}{}
42. \textbf{Roux S}, \textbf{Brum JR}, \textbf{Dutilh BE},
\textbf{Sunagawa S}, \textbf{Duhaime MB}, \textbf{Loy A}, \textbf{Poulos
BT}, \textbf{Solonenko N}, \textbf{Lara E}, \textbf{Poulain J},
\textbf{Pesant S}, \textbf{Kandels-Lewis S}, \textbf{Dimier C},
\textbf{Picheral M}, \textbf{Searson S}, \textbf{Cruaud C},
\textbf{Alberti A}, \textbf{Duarte CM}, \textbf{Gasol JM}, \textbf{Vaqué
D}, \textbf{Tara Oceans Coordinators}, \textbf{Bork P}, \textbf{Acinas
SG}, \textbf{Wincker P}, \textbf{Sullivan MB}. 2016. Ecogenomics and
potential biogeochemical impacts of globally abundant ocean viruses.
Nature \textbf{537}:689--693.

\hypertarget{ref-Grice:2009eea}{}
43. \textbf{Grice EA}, \textbf{Kong HH}, \textbf{Conlan S},
\textbf{Deming CB}, \textbf{Davis J}, \textbf{Young AC}, \textbf{NISC
Comparative Sequencing Program}, \textbf{Bouffard GG}, \textbf{Blakesley
RW}, \textbf{Murray PR}, \textbf{Green ED}, \textbf{Turner ML},
\textbf{Segre JA}. 2009. Topographical and Temporal Diversity of the
Human Skin Microbiome. Science \textbf{324}:1190--1192.

\hypertarget{ref-Findley:2013jf}{}
44. \textbf{Findley K}, \textbf{Oh J}, \textbf{Yang J}, \textbf{Conlan
S}, \textbf{Deming C}, \textbf{Meyer JA}, \textbf{Schoenfeld D},
\textbf{Nomicos E}, \textbf{Park M}, \textbf{NIH Intramural Sequencing
Center Comparative Sequencing Program}, \textbf{Kong HH}, \textbf{Segre
JA}. 2013. Topographic diversity of fungal and bacterial communities in
human skin. Nature 1--6.

\hypertarget{ref-Costello:2009im}{}
45. \textbf{Costello EK}, \textbf{Lauber CL}, \textbf{Hamady M},
\textbf{Fierer N}, \textbf{Gordon JI}, \textbf{Knight R}. 2009.
Bacterial community variation in human body habitats across space and
time. Science \textbf{326}:1694--1697.

\hypertarget{ref-Consortium:2012iz}{}
46. \textbf{Consortium THMP}. 2012. A framework for human microbiome
research. Nature \textbf{486}:215--221.

\hypertarget{ref-Roux:2015dh}{}
47. \textbf{Roux S}, \textbf{Enault F}, \textbf{Hurwitz BL},
\textbf{Sullivan MB}. 2015. VirSorter: mining viral signal from
microbial genomic data. PeerJ \textbf{3}:e985--20.

\hypertarget{ref-Jensen:1998vh}{}
48. \textbf{Jensen EC}, \textbf{Schrader HS}, \textbf{Rieland B},
\textbf{Thompson TL}, \textbf{Lee KW}, \textbf{Nickerson KW},
\textbf{Kokjohn TA}. 1998. Prevalence of broad-host-range lytic
bacteriophages of Sphaerotilus natans, Escherichia coli, and Pseudomonas
aeruginosa. Applied and Environmental Microbiology \textbf{64}:575--580.

\hypertarget{ref-Malki:2015tm}{}
49. \textbf{Malki K}, \textbf{Kula A}, \textbf{Bruder K}, \textbf{Sible
E}. 2015. Bacteriophages isolated from Lake Michigan demonstrate broad
host-range across several bacterial phyla. Virology.

\hypertarget{ref-Schwarzer:2012ez}{}
50. \textbf{Schwarzer D}, \textbf{Buettner FFR}, \textbf{Browning C},
\textbf{Nazarov S}, \textbf{Rabsch W}, \textbf{Bethe A},
\textbf{Oberbeck A}, \textbf{Bowman VD}, \textbf{Stummeyer K},
\textbf{Mühlenhoff M}, \textbf{Leiman PG}, \textbf{Gerardy-Schahn R}.
2012. A multivalent adsorption apparatus explains the broad host range
of phage phi92: a comprehensive genomic and structural analysis. Journal
of Virology \textbf{86}:10384--10398.

\hypertarget{ref-Kim:2012dh}{}
51. \textbf{Kim S}, \textbf{Rahman M}, \textbf{Seol SY}, \textbf{Yoon
SS}, \textbf{Kim J}. 2012. Pseudomonas aeruginosa bacteriophage PA1Ø
requires type IV pili for infection and shows broad bactericidal and
biofilm removal activities. Applied and Environmental Microbiology
\textbf{78}:6380--6385.

\hypertarget{ref-Matsuzaki:1992gw}{}
52. \textbf{Matsuzaki S}, \textbf{Tanaka S}, \textbf{Koga T},
\textbf{Kawata T}. 1992. A Broad-Host-Range Vibriophage, KVP40, Isolated
from Sea Water. Microbiology and Immunology \textbf{36}:93--97.

\hypertarget{ref-Orchard:2014hq}{}
53. \textbf{Orchard S}, \textbf{Ammari M}, \textbf{Aranda B},
\textbf{Breuza L}, \textbf{Briganti L}, \textbf{Broackes-Carter F},
\textbf{Campbell NH}, \textbf{Chavali G}, \textbf{Chen C},
\textbf{del-Toro N}, \textbf{Duesbury M}, \textbf{Dumousseau M},
\textbf{Galeota E}, \textbf{Hinz U}, \textbf{Iannuccelli M},
\textbf{Jagannathan S}, \textbf{Jimenez R}, \textbf{Khadake J},
\textbf{Lagreid A}, \textbf{Licata L}, \textbf{Lovering RC},
\textbf{Meldal B}, \textbf{Melidoni AN}, \textbf{Milagros M},
\textbf{Peluso D}, \textbf{Perfetto L}, \textbf{Porras P},
\textbf{Raghunath A}, \textbf{Ricard-Blum S}, \textbf{Roechert B},
\textbf{Stutz A}, \textbf{Tognolli M}, \textbf{Roey K van},
\textbf{Cesareni G}, \textbf{Hermjakob H}. 2014. The MIntAct
project--IntAct as a common curation platform for 11 molecular
interaction databases. Nucleic Acids Research \textbf{42}:D358--63.

\hypertarget{ref-Kleiner:2015kd}{}
54. \textbf{Kleiner M}, \textbf{Hooper LV}, \textbf{Duerkop BA}. 2015.
Evaluation of methods to purify virus-like particles for metagenomic
sequencing of intestinal viromes. BMC Genomics \textbf{16}:7.

\hypertarget{ref-Meisel:2016iu}{}
55. \textbf{Meisel JS}, \textbf{Hannigan GD}, \textbf{Tyldsley AS},
\textbf{SanMiguel AJ}, \textbf{Hodkinson BP}, \textbf{Zheng Q},
\textbf{Grice EA}. 2016. Skin microbiome surveys are strongly influenced
by experimental design. Journal of Investigative Dermatology.

\hypertarget{ref-Turnbaugh:2009hf}{}
56. \textbf{Turnbaugh PJ}, \textbf{Ridaura VK}, \textbf{Faith JJ},
\textbf{Rey FE}, \textbf{Knight R}, \textbf{Gordon JI}. 2009. The effect
of diet on the human gut microbiome: a metagenomic analysis in humanized
gnotobiotic mice. Science Translational Medicine
\textbf{1}:6ra14--6ra14.

\hypertarget{ref-David:2014cl}{}
57. \textbf{David LA}, \textbf{Maurice CF}, \textbf{Carmody RN},
\textbf{Gootenberg DB}, \textbf{Button JE}, \textbf{Wolfe BE},
\textbf{Ling AV}, \textbf{Devlin AS}, \textbf{Varma Y},
\textbf{Fischbach MA}, \textbf{Biddinger SB}, \textbf{Dutton RJ},
\textbf{Turnbaugh PJ}. 2014. Diet rapidly and reproducibly alters the
human gut microbiome. Nature \textbf{505}:559--563.

\hypertarget{ref-Jeong:2000im}{}
58. \textbf{Jeong H}, \textbf{Tombor B}, \textbf{Albert R},
\textbf{Oltvai ZN}, \textbf{Barabási AL}. 2000. The large-scale
organization of metabolic networks. Nature \textbf{407}:651--654.

\hypertarget{ref-Grice:2009ee}{}
59. \textbf{Grice EA}, \textbf{Kong HH}, \textbf{Conlan S},
\textbf{Deming CB}, \textbf{Davis J}, \textbf{Young AC}, \textbf{NISC
Comparative Sequencing Program}, \textbf{Bouffard GG}, \textbf{Blakesley
RW}, \textbf{Murray PR}, \textbf{Green ED}, \textbf{Turner ML},
\textbf{Segre JA}. 2009. Topographical and Temporal Diversity of the
Human Skin Microbiome. Science \textbf{324}:1190--1192.

\hypertarget{ref-Minot:2013ih}{}
60. \textbf{Minot S}, \textbf{Bryson A}, \textbf{Chehoud C}, \textbf{Wu
GD}, \textbf{Lewis JD}, \textbf{Bushman FD}. 2013. Rapid evolution of
the human gut virome. Proceedings of the National Academy of Sciences of
the United States of America \textbf{110}:12450--12455.

\hypertarget{ref-Schloss:2005hz}{}
61. \textbf{Schloss PD}, \textbf{Handelsman J}. 2005. Introducing DOTUR,
a computer program for defining operational taxonomic units and
estimating species richness. Applied and Environmental Microbiology
\textbf{71}:1501--1506.

\hypertarget{ref-Yilmaz:2010jb}{}
62. \textbf{Yilmaz S}, \textbf{Allgaier M}, \textbf{Hugenholtz P}. 2010.
Multiple displacement amplification compromises quantitative analysis of
metagenomes. Nature Methods \textbf{7}:943--944.

\hypertarget{ref-Kim:2008to}{}
63. \textbf{Kim KH}, \textbf{Chang HW}, \textbf{Nam YD}, \textbf{Roh
SW}. 2008. Amplification of uncultured single-stranded DNA viruses from
rice paddy soil. Applied and \ldots{}.

\hypertarget{ref-Kim:2011hp}{}
64. \textbf{Kim K-H}, \textbf{Bae J-W}. 2011. Amplification methods bias
metagenomic libraries of uncultured single-stranded and double-stranded
DNA viruses. Applied and Environmental Microbiology
\textbf{77}:7663--7668.

\hypertarget{ref-Minot:2012ed}{}
65. \textbf{Minot S}, \textbf{Wu GD}, \textbf{Lewis JD}, \textbf{Bushman
FD}. 2012. Conservation of gene cassettes among diverse viruses of the
human gut. PLOS ONE \textbf{7}:e42342.

\hypertarget{ref-Deng:2014eb}{}
66. \textbf{Deng L}, \textbf{Ignacio-Espinoza JC}, \textbf{Gregory AC},
\textbf{Poulos BT}, \textbf{Weitz JS}, \textbf{Hugenholtz P},
\textbf{Sullivan MB}. 2014. Viral tagging reveals discrete populations
in Synechococcus viral genome sequence space. Nature
\textbf{513}:242--245.

\hypertarget{ref-Brum:2015iaa}{}
67. \textbf{Brum JR}, \textbf{Ignacio-Espinoza JC}, \textbf{Roux S},
\textbf{Doulcier G}, \textbf{Acinas SG}, \textbf{Alberti A},
\textbf{Chaffron S}, \textbf{Cruaud C}, \textbf{Vargas C de},
\textbf{Gasol JM}, \textbf{Gorsky G}, \textbf{Gregory AC}, \textbf{Guidi
L}, \textbf{Hingamp P}, \textbf{Iudicone D}, \textbf{Not F},
\textbf{Ogata H}, \textbf{Pesant S}, \textbf{Poulos BT},
\textbf{Schwenck SM}, \textbf{Speich S}, \textbf{Dimier C},
\textbf{Kandels-Lewis S}, \textbf{Picheral M}, \textbf{Searson S},
\textbf{Tara Oceans Coordinators}, \textbf{Bork P}, \textbf{Bowler C},
\textbf{Sunagawa S}, \textbf{Wincker P}, \textbf{Karsenti E},
\textbf{Sullivan MB}. 2015. Ocean plankton. Patterns and ecological
drivers of ocean viral communities. Science
\textbf{348}:1261498--1261498.

\hypertarget{ref-Malki:2015dw}{}
68. \textbf{Malki K}, \textbf{Kula A}, \textbf{Bruder K}, \textbf{Sible
E}, \textbf{Hatzopoulos T}, \textbf{Steidel S}, \textbf{Watkins SC},
\textbf{Putonti C}. 2015. Bacteriophages isolated from Lake Michigan
demonstrate broad host-range across several bacterial phyla. Virology
Journal \textbf{12}:164.

\hypertarget{ref-Polz:2006fi}{}
69. \textbf{Polz MF}, \textbf{Hunt DE}, \textbf{Preheim SP},
\textbf{Weinreich DM}. 2006. Patterns and mechanisms of genetic and
phenotypic differentiation in marine microbes. Philosophical
Transactions of the Royal Society B: Biological Sciences
\textbf{361}:2009--2021.

\hypertarget{ref-Gregory:2016cg}{}
70. \textbf{Gregory AC}, \textbf{Solonenko SA}, \textbf{Ignacio-Espinoza
JC}, \textbf{LaButti K}, \textbf{Copeland A}, \textbf{Sudek S},
\textbf{Maitland A}, \textbf{Chittick L}, \textbf{Dos Santos F},
\textbf{Weitz JS}, \textbf{Worden AZ}, \textbf{Woyke T},
\textbf{Sullivan MB}. 2016. Genomic differentiation among wild
cyanophages despite widespread horizontal gene transfer. BMC Genomics
\textbf{17}:930.

\hypertarget{ref-PaezEspino:2016gi}{}
71. \textbf{Paez-Espino D}, \textbf{Eloe-Fadrosh EA},
\textbf{Pavlopoulos GA}, \textbf{Thomas AD}, \textbf{Huntemann M},
\textbf{Mikhailova N}, \textbf{Rubin E}, \textbf{Ivanova NN},
\textbf{Kyrpides NC}. 2016. Uncovering Earth's virome. Nature.

\hypertarget{ref-Grice:2011gy}{}
72. \textbf{Grice EA}, \textbf{Segre JA}. 2011. The skin microbiome.
Nature Reviews Microbiology \textbf{9}:244--253.

\hypertarget{ref-Round:2009bz}{}
73. \textbf{Round JL}, \textbf{Mazmanian SK}. 2009. The gut microbiota
shapes intestinal immune responses during health and disease. Nature
reviews Immunology \textbf{9}:313--323.

\hypertarget{ref-FASTXToolkit:wr}{}
74. \textbf{Hannon GJ}. FASTX-Toolkit GNU Affero General Public License.

\hypertarget{ref-Li:2016kd}{}
75. \textbf{Li D}, \textbf{Luo R}, \textbf{Liu C-M}, \textbf{Leung C-M},
\textbf{Ting H-F}, \textbf{Sadakane K}, \textbf{Yamashita H},
\textbf{Lam T-W}. 2016. MEGAHIT v1.0: A fast and scalable metagenome
assembler driven by advanced methodologies and community practices.
METHODS \textbf{102}:3--11.

\hypertarget{ref-Langmead:2012jh}{}
76. \textbf{Langmead B}, \textbf{Salzberg SL}. 2012. Fast gapped-read
alignment with Bowtie 2. Nature Methods \textbf{9}:357--359.

\hypertarget{ref-Alneberg:2014fc}{}
77. \textbf{Alneberg J}, \textbf{Bjarnason BS aacute ri}, \textbf{Bruijn
I de}, \textbf{Schirmer M}, \textbf{Quick J}, \textbf{Ijaz UZ},
\textbf{Lahti L}, \textbf{Loman NJ}, \textbf{Andersson AF},
\textbf{Quince C}. 2014. Binning metagenomic contigs by coverage and
composition. Nature Methods 1--7.

\hypertarget{ref-Hyatt:2012cy}{}
78. \textbf{Hyatt D}, \textbf{LoCascio PF}, \textbf{Hauser LJ},
\textbf{Uberbacher EC}. 2012. Gene and translation initiation site
prediction in metagenomic sequences. Bioinformatics
\textbf{28}:2223--2230.

\hypertarget{ref-caretClassificatio:ux5fU2Litux5f1}{}
79. \textbf{Kuhn M}. caret: Classification and Regression Training.

\hypertarget{ref-Edgar:2007bh}{}
80. \textbf{Edgar RC}. 2007. PILER-CR: fast and accurate identification
of CRISPR repeats. BMC Bioinformatics \textbf{8}:18.

\hypertarget{ref-Camacho:2009fc}{}
81. \textbf{Camacho C}, \textbf{Coulouris G}, \textbf{Avagyan V},
\textbf{Ma N}, \textbf{Papadopoulos J}, \textbf{Bealer K},
\textbf{Madden TL}. 2009. BLAST+: architecture and applications. BMC
Bioinformatics \textbf{10}:1.

\hypertarget{ref-Buchfink:2015ki}{}
82. \textbf{Buchfink B}, \textbf{Xie C}, \textbf{Huson DH}. 2015. Fast
and sensitive protein alignment using DIAMOND. Nature Methods
\textbf{12}:59--60.

\hypertarget{ref-Neoj:Fuwr6PBN}{}
83. Neo4j.

\hypertarget{ref-Theigraphsoftware:vh}{}
84. \textbf{Csardi G}, \textbf{Nepusz T}. The igraph software package
for complex network research.

\end{document}

### 2017-08-11 revfiles/ReviewerResponse.md
Updated formatting.
Updated docs for resubmission.

@@
 Below we included a point-by-point response to the reviewer comments, with our comments highlighted in *italics* and the relevant manuscript change locations highlighted as **bold**.
+Line numbers were included to guide the reviewers to the relevant sections.
+Additions to the text are highlighted in **blue** and removed text is highlighted in **red**.
 
@@
 
-*Thank you for pointing this out.
-This is a good point warranting further discussion.
+*Thank you for raising this point which warrants further discussion.
 We clarified in the text why we feel this is justified, as the degree of stability refers to the reduced likelihood that a randomly removed node will cause a major divide or lack of connection between agents within the network.
@@
 
-*This is a good point and we agree that this needed to be better outlined in the text.
+*We agree that this needed to be better outlined in the text.
 It is important that we make clear that we are not violating our assumption that that phage OGU pool represents phages and the bacterial OGU pool represented bacteria.
 We updated the manuscript to address this concern in multiple ways.
-First, because this was a "data mining" endeavor, we do not have access to the samples for performing further molecular QC.
+First, because this was a post hoc analysis of published data, we do not have access to the samples for performing further molecular QC.
 The original study authors did however perform and report a variety of QC measures to confirm the purity of their virome sample sets.
@@
 We also did not identify any phages in the bacterial dataset using Virsorter.
-The lack of prophage identification, both here and in the previous study analyses, speaks to the need to sequence the bacteria more deeply and create more robust assemblies (their larger size prevents assembly completion at the level that can be done with viruses).* **TABLE S1; LINES 93 - 119**
+The lack of prophage identification, both here and in the previous study analyses, speaks to the need to sequence the bacterial fraction more deeply and create more robust assemblies (their larger size prevents assembly completion at the level that can be done with viruses).* **TABLE S1; LINES 93 - 119**
 
@@
 
-*We are not sure we understand this comment?
+*We are unclear what the reviewer was asking.
 Our goal was to use the edge weights to reflect samples that were highly relative abundant in both the bacterial metagenome and the phage metagenome.
@@
 This approach allows us to more accurately measure the relative abundance in the bacterial and especially viral samples, as has been noted previously.
-We made this clearer in the text and added more relevant citations.* **LINES 149 - 153**
+We clarified this in the text and added more relevant citations.* **LINES 149 - 153**
 
@@
 Datasets underpowered to sustain claims on diet and obesity: Although the authors point out that the data are underpowered, they still make strong claims and the analysis of these data as major points of this manuscript.
-
-*We appreciate this point referring to figure 2, and while we attempted to avoid too strong of claims, we realize we fell short.
-Our goal was to present these as interesting observations, somewhat like a case study.
-We addressed this in the manuscript by further minimizing the effects of our claims and adding relevant caveats.* **LINES 29 - 31; 184 - 201; 268 - 277**
 
@@
 
+*We appreciate this point referring to figure 2, and while we attempted to avoid too strong of claims, we realize we fell short.
+Our goal was to present these as interesting observations, somewhat like a case study.
+We addressed this in the manuscript by further minimizing the effects of our claims and adding relevant caveats.* **LINES 29 - 31; 184 - 201; 268 - 277**
+
 ## Minor Points
@@
 
-*We agree that this needed to be clarified in the manuscript.
-We added the relevant sample size information to the text and figure legends to make clearer what the data actually represent.* **LINES 186; 194 - 196; 219 - 221; 223 - 225; FIGURE LEGENDS 2,3,4**
+*We added the relevant sample size information to the text and figure legends to make clearer what the data actually represent.* **LINES 186; 194 - 196; 219 - 221; 223 - 225; FIGURE LEGENDS 2,3,4**
 
@@
 
-*This is also a good point, thank you. We hit this point in the related bullet point above.*
+*We addressed this point in the related bullet point above.*
 
@@
 
-*We would like to thank the reviewer for pointing out this area of clarification.
+*We would like to thank the reviewer for pointing out that this area needed clarification.
 The operational genomic units were linked sample-wise, by linking the OGUs found between each sample.
@@
 *We appreciate this point and while we attempted to report this as more of a case trial observation (as outlined for the related comment above), we unintentionally used too strong of language, thus overstating the finding beyond our intentions.
-We also needed to make the analysis being performed clearer in the text.
+We also needed to clarify the analysis being performed.
 It is also true that the network analysis is a function of the diversity of the two communities, and we do not wish to make it sound like they are separate.
@@
 Because we were limited to only a small number of validated non-interactions, we decided to also limit the number of positive interactions included, so as to prevent a large bias in the ratio between the two interaction types in the training dataset.
-We apologize for this not being clearer in the text and have clarified that text.* **LINES 124 - 127**
+We clarified that text.* **LINES 124 - 127**
 
@@
 Pairs without scores were omitted from this analysis, and we instead only focused on those that were scored.
-We made this clearer in the text.* **LINES 143 - 144; FIGURE LEGEND 1**
+We have clarified the text.* **LINES 143 - 144; FIGURE LEGEND 1**
 
@@
 
-*This is a very good point and we added this information.* **LINES 186; 194 - 196; 219 - 221; 223 - 225; FIGURE LEGENDS 2,3,4**
+*We have added this information.* **LINES 186; 194 - 196; 219 - 221; 223 - 225; FIGURE LEGENDS 2,3,4**
 
@@
 
-*This is a good point that makes the manuscript clearer. We added these values to the results section of the manuscript so that the reader can access this information.* **LINES 154 - 163**
+*We added these values to the results section of the manuscript so that the reader can access this information.* **LINES 154 - 163**
 
@@
 
-*This is a cool suggestion.
+*This is an intriguing suggestion.
 It sounds like this comment is a reference to the host range inferences within the suggested manuscript.
@@
 
-*Thank you for pointing these out as well. We clarified the dot information, and also fixed the incorrect labels.* **FIGURE 2**
+*We clarified the dot information, and also fixed the incorrect labels.* **FIGURE 2**
 
@@
 
-*This is a good point that there certainly could have been extraction, kit, or other related differences between the 38 and 39 dataset.
+*There certainly could have been extraction, kit, or other related differences between the 38 and 39 dataset.
 We feel these findings, which we attempt to present more as case observations that will require more dedicated study, are worth noting in the paper.
@@
 
-*This is a good point.
-We included a description of these biases associated with these methods.
+*We included a description of these biases associated with these methods.
 As far as measures to overcome these biases, we acknowledge the biases and only performed analyses within studies where samples were treated the same.

### 2017-08-11 doc/manuscript.Rmd
Updated docs for resubmission.

--- 2017-08-06 doc/manuscript.Rmd
@@
 We selected these datasets because their virome samples were subjected to virus-like particle (VLP) purification, which removed contaminating DNA from human cells, bacteria, etc.
-To this end, the publishing authors employed combinations of filtration, chloroform/DNase treatment, and cesium chloride gradients to eliminate organismal DNA (bacteria, human, fungal, etc) and thereby allow for direct assessment of both the extracellular and fully-assembled intracellular virome **(Supplemental Figure \ref{SequenceStats} A-B)** [@Minot:2011ez, @Hannigan:2015fz, @Reyes:2010cwa; @Turnbaugh:2009ei].
-Each research group reported quality control measures to ensure the purity of the virome sequence datasets, using both computational and molecular techniques, including, for example, 16S rRNA gene qPCR **(Table S1)**.
+To this end, the publishing authors employed combinations of filtration, chloroform/DNase treatment, and cesium chloride gradients to eliminate organismal DNA (e.g.
+bacteria, human, fungi, etc) and thereby allow for direct assessment of both the extracellular and fully-assembled intracellular virome **(Supplemental Figure \ref{SequenceStats} A-B)** [@Minot:2011ez, @Hannigan:2015fz, @Reyes:2010cwa; @Turnbaugh:2009ei].
+Each research group reported quality control measures to ensure the purity of the virome sequence datasets, using both computational and molecular techniques (e.g.
+16S rRNA gene qPCR) **(Table S1)**.
 These reports confirmed that the virome libraries consisted of highly purified virus genomic DNA.
@@
 
-We supplemented the previous study findings **(Table S1)** and found that, in light of the rigorous purification and quality control during sample collection and preparation, 77% (228 / 298 operational genomic units) still had some nucleotide similarity to a given bacterial reference genome (e-value < $10^{-25}$).
-As absence of bacterial contamination has been confirmed **(Table S1)**, we interpret this as evidence that the majority of the gut and skin bacteriophages were temperate and thereby shared elements with bacterial reference genomes that contained similar integrated phages when sequenced--a trend previously reported [@Minot:2011ez, @Hannigan:2015fz, @Reyes:2010cwa].
+While supplementing the previous quality control measures **(Table S1)** we found that, in light of the rigorous purification and quality control during sample collection and preparation, 77% (228 / 298 operational genomic units) still had some nucleotide similarity to a given bacterial reference genome (e-value < $10^{-25}$).
+As absence of bacterial contamination had been confirmed by sensitive molecular methods **(Table S1)**, we interpreted this as evidence that the majority of the gut and skin bacteriophages were temperate and thereby shared elements with bacterial reference genomes that contained similar integrated phages when sequenced--a trend previously reported [@Minot:2011ez, @Hannigan:2015fz, @Reyes:2010cwa].
 Additionally, we identified two OGUs as being complete phages using the stringent Virsorter phage identification algorithm (class 1 confidence group) [@Roux:2015dh].
 
-While the whole metagenomic shotgun sequence samples were not subjected to purification, they primarily consisted of bacteria, quantified with an average viral relative abundance of 0.4% **(Table S1)** [@Minot:2011ez; @Hannigan:2015fz; @Reyes:2010cwa; @Turnbaugh:2009ei].
-We again supplemented these findings and found that only 2% (6 / 280 operational genomic units) of bacteria had significantly strong nucleotide similarity to phage reference genomes (e-value < $10^{-25}$) [@Minot:2011ez; @Hannigan:2015fz; @Reyes:2010cwa; @Turnbaugh:2009ei].
+The whole metagenomic shotgun sequence samples primarily consisted of bacteria, with an average viral relative abundance of 0.4% **(Table S1)** [@Minot:2011ez; @Hannigan:2015fz; @Reyes:2010cwa; @Turnbaugh:2009ei].
+We found that only 2% (6 / 280 OGUs) of bacterial OGUs had significantly strong nucleotide similarity to phage reference genomes (e-value < $10^{-25}$) [@Minot:2011ez; @Hannigan:2015fz; @Reyes:2010cwa; @Turnbaugh:2009ei].
 No OGUs were confidently identified as lytic or temperate phage OGUs in the bacterial dataset using the Virsorter algorithm [@Roux:2015dh].
@@
 Metadata including study, sample ID, disease, and OGU abundance within the community were stored in the master network for parsing in downstream analyses **(Supplemental Figure \ref{NetworkDiagram})**.
-Bacterial and phage relative abundance was recorded in each sample for each OGU, and the weight of the edge connecting those OGUs was calculated as a function of those relative abundance values.
+Bacterial and phage relative abundance was recorded in each sample for each OGU and the weight of the edge connecting those OGUs was calculated as a function of those relative abundance values.
 The separate extraction of the phage and bacterial libraries ensured a more accurate measurement of the microbial communities, as has been outlined previously [@Kleiner:2015kd; @Meisel:2016iu].
@@
 Like the master network, the skin network exhibited a diameter of `r prettyNum(genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Diameter"], big.mark = ",")` (measure of graph size; the greatest number of traversed vertices required between two vertices) and included `r prettyNum(genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Vertices"], big.mark = ",")` (`r prettyNum(sitestats[c(sitestats$class %in% "skin"),"phagenodes"], big.mark = ",")` phages, `r prettyNum(sitestats[c(sitestats$class %in% "skin"),"bacterianodes"], big.mark = ",")` bacteria, `r round(100 * genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Vertices"] / genfigurestats[c(genfigurestats$class %in% "Total"),"Vertices"], digits = 1)`% total) and `r prettyNum(genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Edges"], big.mark = ",")` (`r round(100 * genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Edges"] / genfigurestats[c(genfigurestats$class %in% "Total"),"Edges"], digits = 1)`%) of the master network nodes and edges, respectively **(Figure \ref{RocCurve} E - F)**.
-The phages and bacteria in the gut diet and twin sample sets were more sparsely related, with the diet study consisting of `r prettyNum(genfigurestats[c(genfigurestats$class %in% "DietStudy"),"Vertices"], big.mark = ",")` (`r prettyNum(sitestats[c(sitestats$class %in% "diet"),"phagenodes"], big.mark = ",")` phages, `r prettyNum(sitestats[c(sitestats$class %in% "diet"),"bacterianodes"], big.mark = ",")` bacteria) nodes and `r prettyNum(genfigurestats[c(genfigurestats$class %in% "DietStudy"),"Edges"], big.mark = ",")` relationships, and the twin study containing `r prettyNum(genfigurestats[c(genfigurestats$class %in% "TwinStudy"),"Vertices"], big.mark = ",")` (`r prettyNum(sitestats[c(sitestats$class %in% "twin"),"phagenodes"], big.mark = ",")` phages, `r prettyNum(sitestats[c(sitestats$class %in% "twin"),"bacterianodes"], big.mark = ",")` bacteria) nodes and `r prettyNum(genfigurestats[c(genfigurestats$class %in% "TwinStudy"),"Edges"], big.mark = ",")` relationships **(Figure \ref{RocCurve} E - F)**.
+The phages and bacteria in the diet and twin sample sets were more sparsely related, with the diet study consisting of `r prettyNum(genfigurestats[c(genfigurestats$class %in% "DietStudy"),"Vertices"], big.mark = ",")` (`r prettyNum(sitestats[c(sitestats$class %in% "diet"),"phagenodes"], big.mark = ",")` phages, `r prettyNum(sitestats[c(sitestats$class %in% "diet"),"bacterianodes"], big.mark = ",")` bacteria) nodes and `r prettyNum(genfigurestats[c(genfigurestats$class %in% "DietStudy"),"Edges"], big.mark = ",")` relationships, and the twin study containing `r prettyNum(genfigurestats[c(genfigurestats$class %in% "TwinStudy"),"Vertices"], big.mark = ",")` (`r prettyNum(sitestats[c(sitestats$class %in% "twin"),"phagenodes"], big.mark = ",")` phages, `r prettyNum(sitestats[c(sitestats$class %in% "twin"),"bacterianodes"], big.mark = ",")` bacteria) nodes and `r prettyNum(genfigurestats[c(genfigurestats$class %in% "TwinStudy"),"Edges"], big.mark = ",")` relationships **(Figure \ref{RocCurve} E - F)**.
 
@@
 We used this information to calculate the average connectedness per sample, which was corrected for the maximum potential degree of connectedness.
-Unfortunately our dataset was insufficiently powered to make strong conclusions toward this hypothesis, but the data still provides an interesting "case observation" that warrants further investigation.
+Unfortunately our dataset was insufficiently powered to make strong conclusions toward this hypothesis, but this is an interesting observation that warrants further investigation.
 
@@
 These results suggested that the obesity-associated networks are less connected, having microbes further from all other microbes within the community.
-This again comes with the caveat that this is a case observation with too small of a sample size to make more substantial claims.
-This is however an area for future investigation.
+This again comes with the caveat that this is a preliminary observation with too small of a sample size to make more substantial claims.
 
@@
 Network dissimilarity within each person over the 8-10 day sampling period was less than the average dissimilarity between that person and others, although this difference was not statistically significant (p-value = `r signif(interstats[c(interstats$site %in% "Diet"), "prob"], digits = sigfig)`, **Figure \ref{intradiv} B**).
-Four of the five available subjects were used here because one of the subjects was not sampled at the initial time point.
-The lack of statistical confidence was again due to the small sample size of this dataset.
+Four of the five available subjects were used because one of the subjects was not sampled at the initial time point.
+The lack of statistical confidence was likely due to the small sample size of this dataset.
 
@@
 
-# Figures
+# Figure Legends
 
@@
 This network includes information from all three published studies.
-(E) Network diameter (measure of graph size; the greatest number of traversed vertices required between two vertices), (F) number of vertices, and (G) number of edges (relationships) for the total network (orange) and the individual study sub-networks (diet study = red, skin study = yellow, twin study = green).* \label{RocCurve}](../figures/rocCurves.pdf){ width=85% }
-
-\newpage
+(E) Network diameter (measure of graph size; the greatest number of traversed vertices required between two vertices), (F) number of vertices, and (G) number of edges (relationships) for the total network (orange) and the individual study sub-networks (diet study = red, skin study = yellow, twin study = green).* \label{RocCurve}](){ width=85% }
 
@@
 (C) Quantification of average degree centrality and (D) closeness centrality between obese and healthy adult women from the Twin gut study.
-Each point represents a stool sample taken from one of the three adult woman confirmed as obese or healthy and with matching virus and bacteria data.
-*\label{dietnetworks}](../figures/dietnetworks.pdf){ width=90% }
-
-\newpage
+Each point represents a stool sample taken from one of the three adult woman confirmed as obese or healthy and with matching virus and bacteria data.* \label{dietnetworks}](){ width=90% }
 
@@
 (D) Quantification of skin network dissimilarity within the same subject and anatomical location over time (intrapersonal) and the mean dissimilarity between the subject of interest and all other subjects at the same time and the same anatomical location (interpersonal).
-All p-values were calculated using a paired Wilcoxon test.*\label{intradiv}](../figures/intrapersonal_diversity.pdf){ width=90% }
+All p-values were calculated using a paired Wilcoxon test.*\label{intradiv}](){ width=90% }
 
@@
 Cluster centroids are represented by dots and the extended lines represent the associated 95% confidence intervals.
-Significant comparisons (p-value < 0.05) are colored in red, and non-significant comparisons are gray.*\label{skinnetwork}](../figures/skinplotresults.pdf){ width=75% }
+Significant comparisons (p-value < 0.05) are colored in red, and non-significant comparisons are gray.*\label{skinnetwork}](){ width=75% }
 
@@
 
-# Supplemental Figures
+# Supplemental Figure Legends
 
@@
 
-![**Sequencing Depth Summary.** *Number of sequences that aligned to (A) Phage and (B) Bacteria operational genomic units per sample and colored by study.*\label{SequenceStats}](../figures/SequenceAbund.pdf){ width=80% }
+![**Sequencing Depth Summary.** *Number of sequences that aligned to (A) Phage and (B) Bacteria operational genomic units per sample and colored by study.*\label{SequenceStats}](){ width=80% }
 
@@
 ![**Contig Summary Statistics.** *Scatter plot heat map with each hexagon representing the abundance of contigs.
-Contigs are organized by length on the x-axis and the number of aligned sequences on the y-axis.*\label{ContigStats}](../figures/ContigStats.pdf){ width=50% }
+Contigs are organized by length on the x-axis and the number of aligned sequences on the y-axis.*\label{ContigStats}](){ width=50% }
 
@@
 ![**Operational Genomic Unit Summary Statistics.** *Scatter plot with operational genomic unit clusters organized by average contig length within the cluster on the x-axis and the number of contigs in the cluster on the y-axis.
-Operational genomic units of (A) bacteriophages and (B) bacteria are shown.*\label{ClusterStats}](../figures/ClusterStats.pdf){ width=80% }
+Operational genomic units of (A) bacteriophages and (B) bacteria are shown.*\label{ClusterStats}](){ width=80% }
 
@@
 B) Quantification of bacterial host strains known to exist for each phage.
-C) Genome strandedness and D) linearity of the phage reference genomes used for the dataset.*\label{ValidationOverview}](../figures/BenchmarkDataset.pdf){ width=100% }
+C) Genome strandedness and D) linearity of the phage reference genomes used for the dataset.*\label{ValidationOverview}](){ width=100% }
 
@@
 Infectious interactions were recorded between phage and bacteria operational genomic units (OGUs).
-Sequence count abundance for each OGU within each sample was also recorded.*\label{NetworkDiagram}](../figures/graphdatabasediagram.pdf){ width=50% }
+Sequence count abundance for each OGU within each sample was also recorded.*\label{NetworkDiagram}](){ width=50% }
 
@@
 ![**Intrapersonal vs Interpersonal Dissimilarity of the Skin.** *Quantification of skin network dissimilarity within the same subject and anatomical location over time (intrapersonal) and the mean dissimilarity between the subject of interest and all other subjects at the same time and the same anatomical location (interpersonal), separated by each anatomical site (forehead [Fh], palm [Pa], toe web [Tw], umbilicus [Um], antecubital fossa [Ac], axilla [Ax], and retroauricular crease [Ra]).
-P-value was calculated using a paired Wilcoxon test.*\label{allskin}](../figures/intraallskin.pdf){ width=75% }
+P-value was calculated using a paired Wilcoxon test.*\label{allskin}](){ width=75% }
 
@@
 
-# Supplemental Tables
-
---------------------------
-Study                     Citation          Virome Quality Control Measures
-
-------------------------- ----------------- -----------------------------------------------------
-   Diet & the Gut Virome     Minot, 2011    • 16S rRNA gene qPCR revealed reduction in bacterial DNA of at least 10,000X.
-
-                                            • Alignment of shotgun sequences revealed 35X reduction in 16S rRNA gene alignments in virome compared to bacteria shotgun.
-
-                                            • Electron microscopy and nucleic acid stain techniques visually confirmed lack of bacteria in virome samples.
-
-   Skin Virome             Hannigan, 2015    • Significant reduction in reads mapping to 16S rRNA gene sequence, compared to bacteria shotgun dataset.
-
-                                             • Significant reduction in reads mapping to human genome, compared to bacteria shotgun dataset.
-
-                                             • Average viral relative abundance of 0.4% in bacterial shotgun dataset.
-
-  Twin Gut Virome           Reyes, 2010      • Confirmation that 2.5% of bacterial shotgun reads mapped to virome, and 76% of virome reads matched the shotgun 2.5%.
---------------------------
+# Supplemental Table Legend
 

### 2017-08-11 doc/manuscript.tex
Updated docs for resubmission.

@@
 \urlstyle{same}  % don't use monospace font for urls
-\usepackage{longtable,booktabs}
 \usepackage{graphicx,grffile}
@@
 treatment, and cesium chloride gradients to eliminate organismal DNA
-(bacteria, human, fungal, etc) and thereby allow for direct assessment
-of both the extracellular and fully-assembled intracellular virome
-\textbf{(Supplemental Figure \ref{SequenceStats} A-B)} (14, 39). Each
-research group reported quality control measures to ensure the purity of
-the virome sequence datasets, using both computational and molecular
-techniques, including, for example, 16S rRNA gene qPCR \textbf{(Table
-S1)}. These reports confirmed that the virome libraries consisted of
-highly purified virus genomic DNA.
+(e.g.~bacteria, human, fungi, etc) and thereby allow for direct
+assessment of both the extracellular and fully-assembled intracellular
+virome \textbf{(Supplemental Figure \ref{SequenceStats} A-B)} (14, 39).
+Each research group reported quality control measures to ensure the
+purity of the virome sequence datasets, using both computational and
+molecular techniques (e.g.~16S rRNA gene qPCR) \textbf{(Table S1)}.
+These reports confirmed that the virome libraries consisted of highly
+purified virus genomic DNA.
 
@@
 
-We supplemented the previous study findings \textbf{(Table S1)} and
-found that, in light of the rigorous purification and quality control
-during sample collection and preparation, 77\% (228 / 298 operational
-genomic units) still had some nucleotide similarity to a given bacterial
-reference genome (e-value \textless{} \(10^{-25}\)). As absence of
-bacterial contamination has been confirmed \textbf{(Table S1)}, we
-interpret this as evidence that the majority of the gut and skin
-bacteriophages were temperate and thereby shared elements with bacterial
-reference genomes that contained similar integrated phages when
-sequenced--a trend previously reported (14). Additionally, we identified
-two OGUs as being complete phages using the stringent Virsorter phage
-identification algorithm (class 1 confidence group) (47).
-
-While the whole metagenomic shotgun sequence samples were not subjected
-to purification, they primarily consisted of bacteria, quantified with
-an average viral relative abundance of 0.4\% \textbf{(Table S1)} (13,
-14, 38, 39). We again supplemented these findings and found that only
-2\% (6 / 280 operational genomic units) of bacteria had significantly
-strong nucleotide similarity to phage reference genomes (e-value
-\textless{} \(10^{-25}\)) (13, 14, 38, 39). No OGUs were confidently
-identified as lytic or temperate phage OGUs in the bacterial dataset
-using the Virsorter algorithm (47). Together this suggests only minimal
-``contamination'' of the bacterial OGUs.
+While supplementing the previous quality control measures \textbf{(Table
+S1)} we found that, in light of the rigorous purification and quality
+control during sample collection and preparation, 77\% (228 / 298
+operational genomic units) still had some nucleotide similarity to a
+given bacterial reference genome (e-value \textless{} \(10^{-25}\)). As
+absence of bacterial contamination had been confirmed by sensitive
+molecular methods \textbf{(Table S1)}, we interpreted this as evidence
+that the majority of the gut and skin bacteriophages were temperate and
+thereby shared elements with bacterial reference genomes that contained
+similar integrated phages when sequenced--a trend previously reported
+(14). Additionally, we identified two OGUs as being complete phages
+using the stringent Virsorter phage identification algorithm (class 1
+confidence group) (47).
+
+The whole metagenomic shotgun sequence samples primarily consisted of
+bacteria, with an average viral relative abundance of 0.4\%
+\textbf{(Table S1)} (13, 14, 38, 39). We found that only 2\% (6 / 280
+OGUs) of bacterial OGUs had significantly strong nucleotide similarity
+to phage reference genomes (e-value \textless{} \(10^{-25}\)) (13, 14,
+38, 39). No OGUs were confidently identified as lytic or temperate phage
+OGUs in the bacterial dataset using the Virsorter algorithm (47).
+Together this suggests only minimal ``contamination'' of the bacterial
+OGUs.
 
@@
 \textbf{(Supplemental Figure \ref{NetworkDiagram})}. Bacterial and phage
-relative abundance was recorded in each sample for each OGU, and the
+relative abundance was recorded in each sample for each OGU and the
 weight of the edge connecting those OGUs was calculated as a function of
@@
 \textbf{(Figure \ref{RocCurve} E - F)}. The phages and bacteria in the
-gut diet and twin sample sets were more sparsely related, with the diet
+diet and twin sample sets were more sparsely related, with the diet
 study consisting of 89 (41 phages, 48 bacteria) nodes and 5,566
@@
 was insufficiently powered to make strong conclusions toward this
-hypothesis, but the data still provides an interesting ``case
-observation'' that warrants further investigation.
+hypothesis, but this is an interesting observation that warrants further
+investigation.
 
@@
 less connected, having microbes further from all other microbes within
-the community. This again comes with the caveat that this is a case
-observation with too small of a sample size to make more substantial
-claims. This is however an area for future investigation.
+the community. This again comes with the caveat that this is a
+preliminary observation with too small of a sample size to make more
+substantial claims.
 
@@
 (p-value = 0.125, \textbf{Figure \ref{intradiv} B}). Four of the five
-available subjects were used here because one of the subjects was not
-sampled at the initial time point. The lack of statistical confidence
-was again due to the small sample size of this dataset.
+available subjects were used because one of the subjects was not sampled
+at the initial time point. The lack of statistical confidence was likely
+due to the small sample size of this dataset.
 
@@
 
-\section{Figures}\label{figures}
+\section{Figure Legends}\label{figure-legends}
 
@@
 \centering
-\includegraphics[width=0.85000\textwidth]{../figures/rocCurves.pdf}
+\includegraphics[width=0.85000\textwidth]{}
 \caption{\textbf{Summary of Multi-Study Network Model.} \emph{(A)
@@
 
-\newpage
-
 \begin{figure}[htbp]
 \centering
-\includegraphics[width=0.90000\textwidth]{../figures/dietnetworks.pdf}
+\includegraphics[width=0.90000\textwidth]{}
 \caption{\textbf{Impact of Diet and Obesity on Gut Network Structure.}
@@
 \centering
-\includegraphics[width=0.90000\textwidth]{../figures/intrapersonal_diversity.pdf}
+\includegraphics[width=0.90000\textwidth]{}
 \caption{\textbf{Intrapersonal vs Interpersonal Network Dissimilarity
@@
 \centering
-\includegraphics[width=0.75000\textwidth]{../figures/skinplotresults.pdf}
+\includegraphics[width=0.75000\textwidth]{}
 \caption{\textbf{Impact of Skin Micro-Environment on Microbiome Network
@@
 
-\section{Supplemental Figures}\label{supplemental-figures}
+\section{Supplemental Figure Legends}\label{supplemental-figure-legends}
 
@@
 \centering
-\includegraphics[width=0.80000\textwidth]{../figures/SequenceAbund.pdf}
+\includegraphics[width=0.80000\textwidth]{}
 \caption{\textbf{Sequencing Depth Summary.} \emph{Number of sequences
@@
 \centering
-\includegraphics[width=0.50000\textwidth]{../figures/ContigStats.pdf}
+\includegraphics[width=0.50000\textwidth]{}
 \caption{\textbf{Contig Summary Statistics.} \emph{Scatter plot heat map
@@
 \centering
-\includegraphics[width=0.80000\textwidth]{../figures/ClusterStats.pdf}
+\includegraphics[width=0.80000\textwidth]{}
 \caption{\textbf{Operational Genomic Unit Summary Statistics.}
@@
 \centering
-\includegraphics[width=1.00000\textwidth]{../figures/BenchmarkDataset.pdf}
+\includegraphics[width=1.00000\textwidth]{}
 \caption{\textbf{Summary information of validation dataset used in the
@@
 \centering
-\includegraphics[width=0.50000\textwidth]{../figures/graphdatabasediagram.pdf}
+\includegraphics[width=0.50000\textwidth]{}
 \caption{\textbf{Structure of the interactive network.} \emph{Metadata
@@
 \centering
-\includegraphics[width=0.75000\textwidth]{../figures/intraallskin.pdf}
+\includegraphics[width=0.75000\textwidth]{}
 \caption{\textbf{Intrapersonal vs Interpersonal Dissimilarity of the
@@
 
-\section{Supplemental Tables}\label{supplemental-tables}
-
-\begin{longtable}[]{@{}lll@{}}
-\caption{Summary of the primary quality control measures reported in the
-original publications of the viromes used in this current
-study.}\tabularnewline
-\toprule
-\begin{minipage}[b]{0.24\columnwidth}\raggedright\strut
-Study
-\strut\end{minipage} &
-\begin{minipage}[b]{0.17\columnwidth}\raggedright\strut
-Citation
-\strut\end{minipage} &
-\begin{minipage}[b]{0.50\columnwidth}\raggedright\strut
-Virome Quality Control Measures
-\strut\end{minipage}\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-\begin{minipage}[b]{0.24\columnwidth}\raggedright\strut
-Study
-\strut\end{minipage} &
-\begin{minipage}[b]{0.17\columnwidth}\raggedright\strut
-Citation
-\strut\end{minipage} &
-\begin{minipage}[b]{0.50\columnwidth}\raggedright\strut
-Virome Quality Control Measures
-\strut\end{minipage}\tabularnewline
-\midrule
-\endhead
-\begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
-Diet \& the Gut Virome
-\strut\end{minipage} &
-\begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
-Minot, 2011
-\strut\end{minipage} &
-\begin{minipage}[t]{0.50\columnwidth}\raggedright\strut
-• 16S rRNA gene qPCR revealed reduction in bacterial DNA of at least
-10,000X.
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
-\strut\end{minipage} &
-\begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
-\strut\end{minipage} &
-\begin{minipage}[t]{0.50\columnwidth}\raggedright\strut
-• Alignment of shotgun sequences revealed 35X reduction in 16S rRNA gene
-alignments in virome compared to bacteria shotgun.
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
-\strut\end{minipage} &
-\begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
-\strut\end{minipage} &
-\begin{minipage}[t]{0.50\columnwidth}\raggedright\strut
-• Electron microscopy and nucleic acid stain techniques visually
-confirmed lack of bacteria in virome samples.
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
-Skin Virome
-\strut\end{minipage} &
-\begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
-Hannigan, 2015
-\strut\end{minipage} &
-\begin{minipage}[t]{0.50\columnwidth}\raggedright\strut
-• Significant reduction in reads mapping to 16S rRNA gene sequence,
-compared to bacteria shotgun dataset.
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
-\strut\end{minipage} &
-\begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
-\strut\end{minipage} &
-\begin{minipage}[t]{0.50\columnwidth}\raggedright\strut
-• Significant reduction in reads mapping to human genome, compared to
-bacteria shotgun dataset.
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
-\strut\end{minipage} &
-\begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
-\strut\end{minipage} &
-\begin{minipage}[t]{0.50\columnwidth}\raggedright\strut
-• Average viral relative abundance of 0.4\% in bacterial shotgun
-dataset.
-\strut\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.24\columnwidth}\raggedright\strut
-Twin Gut Virome
-\strut\end{minipage} &
-\begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
-Reyes, 2010
-\strut\end{minipage} &
-\begin{minipage}[t]{0.50\columnwidth}\raggedright\strut
-• Confirmation that 2.5\% of bacterial shotgun reads mapped to virome,
-and 76\% of virome reads matched the shotgun 2.5\%.
-\strut\end{minipage}\tabularnewline
-\bottomrule
-\end{longtable}
+\section{Supplemental Table Legend}\label{supplemental-table-legend}
+
+Table: Summary of the primary quality control measures reported in the
+original publications of the viromes used in this current study.
 

### 2017-08-11 revfiles/ReviewerResponse.md
Updated line references.

@@
 The infectiveness refers to the link we drew between the bacteria and phages (how we connected phages and bacteria in our networks).
-We also provided citations to work that further explains this in metabolic and phage-bacteria systems, if the readers are interested in a more in-depth discussion of the topic.* **LINES 176 - 183**
+We also provided citations to work that further explains this in metabolic and phage-bacteria systems, if the readers are interested in a more in-depth discussion of the topic.* **LINES 180 - 187; 268 - 270**
 
@@
 We also did not identify any phages in the bacterial dataset using Virsorter.
-The lack of prophage identification, both here and in the previous study analyses, speaks to the need to sequence the bacterial fraction more deeply and create more robust assemblies (their larger size prevents assembly completion at the level that can be done with viruses).* **TABLE S1; LINES 93 - 119**
+The lack of prophage identification, both here and in the previous study analyses, speaks to the need to sequence the bacterial fraction more deeply and create more robust assemblies (their larger size prevents assembly completion at the level that can be done with viruses).* **TABLE S1; LINES 92 - 121; 379 - 390**
 
@@
 This approach allows us to more accurately measure the relative abundance in the bacterial and especially viral samples, as has been noted previously.
-We clarified this in the text and added more relevant citations.* **LINES 149 - 153**
+We clarified this in the text and added more relevant citations.* **LINES 151 - 155; 157 - 159**
 
@@
 Our goal was to present these as interesting observations, somewhat like a case study.
-We addressed this in the manuscript by further minimizing the effects of our claims and adding relevant caveats.* **LINES 29 - 31; 184 - 201; 268 - 277**
+We addressed this in the manuscript by further minimizing the effects of our claims and adding relevant caveats.* **LINES 29; 185 - 206; 274 - 278**
 
@@
 
-*We added the relevant sample size information to the text and figure legends to make clearer what the data actually represent.* **LINES 186; 194 - 196; 219 - 221; 223 - 225; FIGURE LEGENDS 2,3,4**
+*We added the relevant sample size information to the text and figure legends to make clearer what the data actually represent.* **LINES 191 - 192; 200 - 202; 219 - 221; 224 - 225; 229; FIGURE LEGENDS 2,3,4**
 
@@
 
-*We fixed the error, and made the causes for the sample size clearer in the text as well as the figure legends. Thank you for pointing this out.* **LINES 194 - 195; FIGURE LEGEND 2**
+*We fixed the error, and made the causes for the sample size clearer in the text as well as the figure legends. Thank you for pointing this out.* **LINES 200 - 202; FIGURE LEGEND 2**
 
@@
 
-*We only used the subset of data points that had temporal data spaced approximately the same distance apart. We made this clearer in the text.* **LINES 186 - 187; 219 - 221; FIGURE LEGEND 1,2**
+*We only used the subset of data points that had temporal data spaced approximately the same distance apart. We made this clearer in the text.* **LINES 191 - 192; 224 - 225; FIGURE LEGEND 1,2**
 
@@
 
-*We made this information clearer in the text.* **LINES 223 - 225; FIGURE LEGEND 3**
+*We made this information clearer in the text.* **LINES 229 - 230; FIGURE LEGEND 3**
 
@@
 It is true that this is a dense network, and we hypothesize that that may be because each OGU does not necessarily represent a specific phage or bacterial strain, but is an operationally defined group of similar genomes, similar to the operation taxonomic unit (OTU) often used in microbiome studies.
-In other words, we are not using the OGUs as individual phage strains, but instead these OGUs are operational units that allow us to condense our dataset by contig similarity.* **SEE LINES 114 - 119; 304 - 322**
+In other words, we are not using the OGUs as individual phage strains, but instead these OGUs are operational units that allow us to condense our dataset by contig similarity.* **SEE LINES 103 - 108; 311 - 329**
 
@@
 We also needed to clarify the analysis being performed.
-It is also true that the network analysis is a function of the diversity of the two communities, and we do not wish to make it sound like they are separate.
+It is also true that the network analysis is a function of the diversity of the two communities, and we did not wish to make it sound like they are separate.
 We would be hesitant however to call this dependence a confounder since it is more building off of the diversity data and incorporating other additional information from the community.
-We made this clearer in the text.* **LINES 186; 194 - 196; 219 - 221; 223 - 225; FIGURE LEGENDS 2,3,4**
+We made this clearer in the text.* **LINES 191 - 192; 200 - 202; 219 - 221; 224 - 225; 229; FIGURE LEGENDS 2,3,4**
 
@@
 Because we were limited to only a small number of validated non-interactions, we decided to also limit the number of positive interactions included, so as to prevent a large bias in the ratio between the two interaction types in the training dataset.
-We clarified that text.* **LINES 124 - 127**
+We clarified that text.* **LINES 127 - 129**
 
@@
 Pairs without scores were omitted from this analysis, and we instead only focused on those that were scored.
-We have clarified the text.* **LINES 143 - 144; FIGURE LEGEND 1**
+We have clarified this point.* **SEE LINES 145 - 146; FIGURE LEGEND 1**
 
@@
 
-*We have added this information.* **LINES 186; 194 - 196; 219 - 221; 223 - 225; FIGURE LEGENDS 2,3,4**
+*We have added this information.* **LINES 191 - 192; 200 - 202; 219 - 221; 224 - 225; 229; FIGURE LEGENDS 2,3,4**
 
@@
 
-*We added these values to the results section of the manuscript so that the reader can access this information.* **LINES 154 - 163**
+*We added these values to the results section of the manuscript so that the reader can access this information.* **LINES 155 - 165**
 
@@
 In our discussion section, we mentioned a comparison of these and related findings to our own dataset.
-This information was included in the operational genomic unit section of the discussion.* **LINES 318 - 322**
+A reference to this information was included in the operational genomic unit section of the discussion.* **LINES 318**
 
@@
 As far as measures to overcome these biases, we acknowledge the biases and only performed analyses within studies where samples were treated the same.
-The goal is for this work to use these findings as a jumping off point, and performing future studies with more limited biases from techniques such as MDA.* **LINES 295 - 303**
+The goal is for this work to use these findings as a jumping off point, and performing future studies with more limited biases from techniques such as MDA.* **SEE LINES 302 - 310**

### 2017-08-11 revfiles/ReviewerResponse.md
Fixed two line numbers

@@
 
-*We added the relevant sample size information to the text and figure legends to make clearer what the data actually represent.* **LINES 191 - 192; 200 - 202; 219 - 221; 224 - 225; 229; FIGURE LEGENDS 2,3,4**
+*We added the relevant sample size information to the text and figure legends to make clearer what the data actually represent.* **LINES 191 - 192; 200 - 202; 224 - 225; 229; FIGURE LEGENDS 2,3,4**
 
@@
 We would be hesitant however to call this dependence a confounder since it is more building off of the diversity data and incorporating other additional information from the community.
-We made this clearer in the text.* **LINES 191 - 192; 200 - 202; 219 - 221; 224 - 225; 229; FIGURE LEGENDS 2,3,4**
+We made this clearer in the text.* **LINES 191 - 192; 200 - 202; 224 - 225; 229; FIGURE LEGENDS 2,3,4**
 
@@
 
-*We have added this information.* **LINES 191 - 192; 200 - 202; 219 - 221; 224 - 225; 229; FIGURE LEGENDS 2,3,4**
+*We have added this information.* **LINES 191 - 192; 200 - 202; 224 - 225; 229; FIGURE LEGENDS 2,3,4**
 

### 2017-08-29 doc/manuscript.Rmd
A couple minor updates.
Updates for resubmission.

@@
 skinboxsig <- read.delim(file = "../rtables/skinboxsig.tsv", header = TRUE, sep = "\t")
+averageauc <- read.delim(file = "../data/avgaucnested.tsv", header = TRUE, sep = "\t")
 ```
@@
 Four phage and bacterial genomic features were used in a random forest model to predict infectious relationships between bacteria and phages: 1) genome nucleotide similarities, 2) gene amino acid sequence similarities, 3) bacterial Clustered Regularly Interspaced Short Palindromic Repeat (CRISPR) spacer sequences that target phages, and 4) similarity of protein families associated with experimentally identified protein-protein interactions [@Orchard:2014hq].
-The resulting random forest model was assessed and the area under its receiver operating characteristic (ROC) curve was `r signif(genmodelper$ROC, digits = sigfig)`, the model sensitivity was `r signif(genmodelper$Sens, digits = sigfig)`, and specificity was `r signif(genmodelper$Spec, digits = sigfig)` **(Figure \ref{RocCurve} A)**.
+The resulting random forest model was assessed using nested cross validation, and the median area under its receiver operating characteristic (ROC) curve was `r signif(averageauc[c(averageauc$type %in% "AUC"), 2], digits = sigfig)`, the median model sensitivity was `r signif(averageauc[c(averageauc$type %in% "SENS"), 2], digits = sigfig)`, and median specificity was `r signif(averageauc[c(averageauc$type %in% "SPEC"), 2], digits = sigfig)` **(Figure \ref{RocCurve} A)**.
 The most important predictor in the model was amino acid similarity between genes, followed by nucleotide similarity of whole genomes **(Figure \ref{RocCurve} B)**.
@@
 The classification model was built using the Caret R package (v6.0.73) [@caretClassificatio:_U2Lit_1].
-The model was trained using five-fold cross validation with ten repeats.
+The model was trained using five-fold cross validation with ten repeats, and the median model performance was evaluated by training the model on 80% of the dataset and testing performance on the remaining 20%.
 Pairs without scores were classified as not interacting.
@@
 
-![**Summary of Multi-Study Network Model.** *(A) Average ROC curve used to create the microbiome-virome infection prediction model.
+![**Summary of Multi-Study Network Model.** *(A) Median ROC curve (dark red) used to create the microbiome-virome infection prediction model, based on nested cross validation over 25 random iterations.
+The maximum and minimum performance are shown in light red.
 (B) Importance scores associated with the metrics used in the random forest model to predict relationships between bacteria and phages.
@@
 
-Table: Summary of the primary quality control measures reported in the original publications of the viromes used in this current study.
+Table S1: Summary of the primary quality control measures reported in the original publications of the viromes used in this current study.
 

### 2017-08-29 doc/manuscript.tex
A couple minor updates.
Updates for resubmission.

@@
 protein-protein interactions (53). The resulting random forest model was
-assessed and the area under its receiver operating characteristic (ROC)
-curve was 0.846, the model sensitivity was 0.829, and specificity was
-0.767 \textbf{(Figure \ref{RocCurve} A)}. The most important predictor
-in the model was amino acid similarity between genes, followed by
-nucleotide similarity of whole genomes \textbf{(Figure \ref{RocCurve}
-B)}. Protein family interactions were moderately important to the model,
-and CRISPRs were largely uninformative, due to the minimal amount of
-identifiable CRISPRs in the dataset and their redundancy with the
-nucleotide similarity methods \textbf{(Figure \ref{RocCurve} B)}.
-Approximately one third of the training set relationships yielded no
-score and therefore were unable to be assigned an interaction prediction
-\textbf{(Figure \ref{RocCurve} C)}.
+assessed using nested cross validation, and the median area under its
+receiver operating characteristic (ROC) curve was 0.788, the median
+model sensitivity was 0.952, and median specificity was 0.615
+\textbf{(Figure \ref{RocCurve} A)}. The most important predictor in the
+model was amino acid similarity between genes, followed by nucleotide
+similarity of whole genomes \textbf{(Figure \ref{RocCurve} B)}. Protein
+family interactions were moderately important to the model, and CRISPRs
+were largely uninformative, due to the minimal amount of identifiable
+CRISPRs in the dataset and their redundancy with the nucleotide
+similarity methods \textbf{(Figure \ref{RocCurve} B)}. Approximately one
+third of the training set relationships yielded no score and therefore
+were unable to be assigned an interaction prediction \textbf{(Figure
+\ref{RocCurve} C)}.
 
@@
 R package (v6.0.73) (79). The model was trained using five-fold cross
-validation with ten repeats. Pairs without scores were classified as not
-interacting. The model was optimized using the ROC value. The resulting
-model performance was plotted using the plotROC R package.
+validation with ten repeats, and the median model performance was
+evaluated by training the model on 80\% of the dataset and testing
+performance on the remaining 20\%. Pairs without scores were classified
+as not interacting. The model was optimized using the ROC value. The
+resulting model performance was plotted using the plotROC R package.
 
@@
 \includegraphics[width=0.85000\textwidth]{}
-\caption{\textbf{Summary of Multi-Study Network Model.} \emph{(A)
-Average ROC curve used to create the microbiome-virome infection
-prediction model. (B) Importance scores associated with the metrics used
-in the random forest model to predict relationships between bacteria and
-phages. The importance score is defined as the mean decrease in accuracy
-of the model when a feature (e.g.~Pfam) is excluded. Features include
-the local gene alignments between bacteria and phage genes (denoted
+\caption{\textbf{Summary of Multi-Study Network Model.} \emph{(A) Median
+ROC curve (dark red) used to create the microbiome-virome infection
+prediction model, based on nested cross validation over 25 random
+iterations. The maximum and minimum performance are shown in light red.
+(B) Importance scores associated with the metrics used in the random
+forest model to predict relationships between bacteria and phages. The
+importance score is defined as the mean decrease in accuracy of the
+model when a feature (e.g.~Pfam) is excluded. Features include the local
+gene alignments between bacteria and phage genes (denoted
 \texttt{blastx}; the blastx algorithm in Diamond aligner), local genome
@@
 
-Table: Summary of the primary quality control measures reported in the
-original publications of the viromes used in this current study.
+Table S1: Summary of the primary quality control measures reported in
+the original publications of the viromes used in this current study.
 

### 2017-10-01 doc/manuscript.Rmd
Manuscript updates.
Updated manusript with format for new submission.

@@
 ---
-title: Biogeography & Environmental Conditions Shape Phage & Bacteria Interaction Networks Across the Human Microbiome
+title: Biogeography & environmental conditions shape bacteriophage & bacteria interaction networks across the human microbiome
 author:
@@
   - code: 1
-    address: Department of Microbiology & Immunology, University of Michigan, Ann Arbor, Michigan, 48109
+    address: Department of Microbiology & Immunology, University of Michigan, Ann Arbor, Michigan, 48109, USA
   - code: 2
-    address: Department of Ecology & Evolutionary Biology, University of Michigan, Ann Arbor, Michigan, 48109
+    address: Department of Ecology & Evolutionary Biology, University of Michigan, Ann Arbor, Michigan, 48109, USA
   - code: 3
-    address: Department of Computer Science, University of Michigan, Ann Arbor, Michigan, 48109
+    address: Department of Computer Science, University of Michigan, Ann Arbor, Michigan, 48109, USA
   - code: \*
@@
 
-***Running Title***: Network Diversity of the Healthy Human Microbiome
+***Short Title***: Network Diversity of the Healthy Human Microbiome
 
@@
 
-# Importance
+# Author Summary
 
@@
 
+![**Summary of Multi-Study Network Model.** *(A) Median ROC curve (dark red) used to create the microbiome-virome infection prediction model, based on nested cross validation over 25 random iterations.
+The maximum and minimum performance are shown in light red.
+(B) Importance scores associated with the metrics used in the random forest model to predict relationships between bacteria and phages.
+The importance score is defined as the mean decrease in accuracy of the model when a feature (e.g.
+Pfam) is excluded.
+Features include the local gene alignments between bacteria and phage genes (denoted `blastx`; the blastx algorithm in Diamond aligner), local genome nucleotide alignments between bacteria and phage OGUs, presence of experimentally validated protein family domains (Pfams) between phage and bacteria OGUs, and CRISPR targeting of bacteria toward phages (CRISPR).
+(C) Proportions of samples included (gray) and excluded (red) in the model.
+Samples were excluded from the model because they did not yield any scores.
+Those interactions without scores were automatically classified as not having interactions.
+(D) Bipartite visualization of the resulting phage-bacteria network.
+Phage OGUs are presented in orange, bacteria OGUs in red, and their interaction edges are represented as connecting lines.
+This network includes information from all three published studies.
+(E) Network diameter (measure of graph size; the greatest number of traversed vertices required between two vertices), (F) number of vertices, and (G) number of edges (relationships) for the total network (orange) and the individual study sub-networks (diet study = red, skin study = yellow, twin study = green).* \label{RocCurve}](){ width=85% }
+
 We used our random forest model to classify the relationships between bacteria and phage operational genomic units, which were then used to build the interactive network.
@@
 
+![**Impact of Diet and Obesity on Gut Network Structure.** *(A) Quantification of average degree centrality (number of edges per node) and (B) closeness centrality (average distance from each node to every other node) of gut microbiome networks of subjects limited to exclusively high-fat or low-fat diets.
+Each point represents the centrality from a human subject stool sample that was collected 8-10 days following the beginning of their defined diet.
+There are five samples here, compared to the four in figure \ref{intradiv}, because one of the was only sampled post-diet, providing us data for this analysis but not allowing us to compare to a baseline for figure \ref{intradiv}.
+(C) Quantification of average degree centrality and (D) closeness centrality between obese and healthy adult women from the Twin gut study.
+Each point represents a stool sample taken from one of the three adult woman confirmed as obese or healthy and with matching virus and bacteria data.* \label{dietnetworks}](){ width=90% }
+
 In addition to diet, we found preliminary evidence that obesity influenced network structure.
@@
 
+![**Intrapersonal vs Interpersonal Network Dissimilarity Across Different Human Systems.** *(A) NMDS ordination illustrating network dissimilarity between subjects over time.
+Each sample is colored by subject, with each colored sample pair collected 8-10 days apart.
+Dissimilarity was calculated using the Bray-Curtis metric based on abundance weighted eigenvector centrality signatures, with a greater distance representing greater dissimilarity in bacteria and phage centrality and abundance.
+Only four subjects were included here, compared to the five used in figure \ref{dietnetworks}, because one of the subjects was missing the inital sampling time point and therefore lacked temporal sampling.
+(B) Quantification of gut network dissimilarity within the same subject over time (intrapersonal) and the mean dissimilarity between the subject of interest and all other subjects (interpersonal).
+The p-value is provided near the bottom of the figure.
+(C) Quantification of gut network dissimilarity within subjects from the same family (intrafamily) and the mean dissimilarity between subjects within a family and those of other families (interfamily).
+Each point represents the inter-family and intra-family dissimilarity of a twin or mother that was sampled over time.
+(D) Quantification of skin network dissimilarity within the same subject and anatomical location over time (intrapersonal) and the mean dissimilarity between the subject of interest and all other subjects at the same time and the same anatomical location (interpersonal).
+All p-values were calculated using a paired Wilcoxon test.*\label{intradiv}](){ width=90% }
+
 ## Association Between Environmental Stability and Network Structure Across the Human Skin Landscape
@@
 Also, skin sites that were occluded from the environment were much more highly connected than those that were constantly exposed to the environment or only intermittently occluded (p-value < 0.001, **Figure \ref{skinnetwork} B)**.
+
+![**Impact of Skin Micro-Environment on Microbiome Network Structure.**  *(A) Notched box-plot depicting differences in average eigenvector centrality between moist, intermittently moist, and sebaceous skin sites and (B) occluded, intermittently occluded, and exposed sites.
+Notched box-plots were created using ggplot2 and show the median (center line), the inter-quartile range (IQR; upper and lower boxes), the highest and lowest value within 1.5 \* IQR (whiskers), outliers (dots), and the notch which provides an approximate 95% confidence interval as defined by 1.58 \* IQR / sqrt(n).
+Sample sizes for each group were: Moist = 81, Sebaceous = 56, IntMoist = 56, Occluded = 106, Exposed = 61, IntOccluded = 26.
+(C) NMDS ordination depicting the differences in skin microbiome network structure between skin moisture levels and (D) occlusion.
+Samples are colored by their environment and their dissimilarity to other samples was calculated as described in figure \ref{intradiv}.
+(E) The statistical differences of networks between moisture and (F) occlusion status were quantified with an anova and post hoc Tukey test.
+Cluster centroids are represented by dots and the extended lines represent the associated 95% confidence intervals.
+Significant comparisons (p-value < 0.05) are colored in red, and non-significant comparisons are gray.*\label{skinnetwork}](){ width=75% }
 
@@
 
-https://github.com/SchlossLab/Hannigan_ConjunctisViribus_mSystems_2017
+https://github.com/SchlossLab/Hannigan_ConjunctisViribus_ploscompbio_2017
 
@@
 
+# Author Contributions
+*Conceptualization*: GDH, MBD, DK, PDS.
+*Data Curation*: GDH.
+*Formal Analysis*: GDH.
+*Funding Acquisition*: GDH, PDS.
+*Writing – Original Draft Preparation*: GDH, PDS.
+*Writing – Review & Editing*: GDH, MBD, DK, PDS.
+
 # Funding Information
@@
 
-# Disclosure Declaration
+# Competing interests
 The authors report no conflicts of interest.
@@
 
-# Figure Legends
-
-![**Summary of Multi-Study Network Model.** *(A) Median ROC curve (dark red) used to create the microbiome-virome infection prediction model, based on nested cross validation over 25 random iterations.
-The maximum and minimum performance are shown in light red.
-(B) Importance scores associated with the metrics used in the random forest model to predict relationships between bacteria and phages.
-The importance score is defined as the mean decrease in accuracy of the model when a feature (e.g.
-Pfam) is excluded.
-Features include the local gene alignments between bacteria and phage genes (denoted `blastx`; the blastx algorithm in Diamond aligner), local genome nucleotide alignments between bacteria and phage OGUs, presence of experimentally validated protein family domains (Pfams) between phage and bacteria OGUs, and CRISPR targeting of bacteria toward phages (CRISPR).
-(C) Proportions of samples included (gray) and excluded (red) in the model.
-Samples were excluded from the model because they did not yield any scores.
-Those interactions without scores were automatically classified as not having interactions.
-(D) Bipartite visualization of the resulting phage-bacteria network.
-Phage OGUs are presented in orange, bacteria OGUs in red, and their interaction edges are represented as connecting lines.
-This network includes information from all three published studies.
-(E) Network diameter (measure of graph size; the greatest number of traversed vertices required between two vertices), (F) number of vertices, and (G) number of edges (relationships) for the total network (orange) and the individual study sub-networks (diet study = red, skin study = yellow, twin study = green).* \label{RocCurve}](){ width=85% }
-
-![**Impact of Diet and Obesity on Gut Network Structure.** *(A) Quantification of average degree centrality (number of edges per node) and (B) closeness centrality (average distance from each node to every other node) of gut microbiome networks of subjects limited to exclusively high-fat or low-fat diets.
-Each point represents the centrality from a human subject stool sample that was collected 8-10 days following the beginning of their defined diet.
-There are five samples here, compared to the four in figure \ref{intradiv}, because one of the was only sampled post-diet, providing us data for this analysis but not allowing us to compare to a baseline for figure \ref{intradiv}.
-(C) Quantification of average degree centrality and (D) closeness centrality between obese and healthy adult women from the Twin gut study.
-Each point represents a stool sample taken from one of the three adult woman confirmed as obese or healthy and with matching virus and bacteria data.* \label{dietnetworks}](){ width=90% }
-
-![**Intrapersonal vs Interpersonal Network Dissimilarity Across Different Human Systems.** *(A) NMDS ordination illustrating network dissimilarity between subjects over time.
-Each sample is colored by subject, with each colored sample pair collected 8-10 days apart.
-Dissimilarity was calculated using the Bray-Curtis metric based on abundance weighted eigenvector centrality signatures, with a greater distance representing greater dissimilarity in bacteria and phage centrality and abundance.
-Only four subjects were included here, compared to the five used in figure \ref{dietnetworks}, because one of the subjects was missing the inital sampling time point and therefore lacked temporal sampling.
-(B) Quantification of gut network dissimilarity within the same subject over time (intrapersonal) and the mean dissimilarity between the subject of interest and all other subjects (interpersonal).
-The p-value is provided near the bottom of the figure.
-(C) Quantification of gut network dissimilarity within subjects from the same family (intrafamily) and the mean dissimilarity between subjects within a family and those of other families (interfamily).
-Each point represents the inter-family and intra-family dissimilarity of a twin or mother that was sampled over time.
-(D) Quantification of skin network dissimilarity within the same subject and anatomical location over time (intrapersonal) and the mean dissimilarity between the subject of interest and all other subjects at the same time and the same anatomical location (interpersonal).
-All p-values were calculated using a paired Wilcoxon test.*\label{intradiv}](){ width=90% }
-
-![**Impact of Skin Micro-Environment on Microbiome Network Structure.**  *(A) Notched box-plot depicting differences in average eigenvector centrality between moist, intermittently moist, and sebaceous skin sites and (B) occluded, intermittently occluded, and exposed sites.
-Notched box-plots were created using ggplot2 and show the median (center line), the inter-quartile range (IQR; upper and lower boxes), the highest and lowest value within 1.5 \* IQR (whiskers), outliers (dots), and the notch which provides an approximate 95% confidence interval as defined by 1.58 \* IQR / sqrt(n).
-Sample sizes for each group were: Moist = 81, Sebaceous = 56, IntMoist = 56, Occluded = 106, Exposed = 61, IntOccluded = 26.
-(C) NMDS ordination depicting the differences in skin microbiome network structure between skin moisture levels and (D) occlusion.
-Samples are colored by their environment and their dissimilarity to other samples was calculated as described in figure \ref{intradiv}.
-(E) The statistical differences of networks between moisture and (F) occlusion status were quantified with an anova and post hoc Tukey test.
-Cluster centroids are represented by dots and the extended lines represent the associated 95% confidence intervals.
-Significant comparisons (p-value < 0.05) are colored in red, and non-significant comparisons are gray.*\label{skinnetwork}](){ width=75% }
-
-\newpage
-
-# Supplemental Figure Legends
+# Supplemental Figure Captions
 
@@
 
-# Supplemental Table Legend
+# Supplemental Table Captions
 

### 2017-10-01 doc/manuscript.tex
Manuscript updates.
Updated manusript with format for new submission.

@@
 \hypersetup{unicode=true,
-            pdftitle={Biogeography \& Environmental Conditions Shape Phage \& Bacteria Interaction Networks Across the Human Microbiome},
+            pdftitle={Biogeography \& environmental conditions shape bacteriophage \& bacteria interaction networks across the human microbiome},
             pdfborder={0 0 0},
@@
 
-\title{Biogeography \& Environmental Conditions Shape Phage \& Bacteria
-Interaction Networks Across the Human Microbiome}
+\title{Biogeography \& environmental conditions shape bacteriophage \& bacteria
+interaction networks across the human microbiome}
     \usepackage{authblk}
@@
                             \affil[1]{\nextfont Department of Microbiology \& Immunology, University of Michigan, Ann
-Arbor, Michigan, 48109}
+Arbor, Michigan, 48109, USA}
                     \affil[2]{\nextfont Department of Ecology \& Evolutionary Biology, University of Michigan,
-Ann Arbor, Michigan, 48109}
+Ann Arbor, Michigan, 48109, USA}
                     \affil[3]{\nextfont Department of Computer Science, University of Michigan, Ann Arbor,
-Michigan, 48109}
+Michigan, 48109, USA}
                     \affil[*]{\nextfont To whom correspondence may be addressed.}
@@
 
-\textbf{\emph{Running Title}}: Network Diversity of the Healthy Human
+\textbf{\emph{Short Title}}: Network Diversity of the Healthy Human
 Microbiome
@@
 
-\section{Importance}\label{importance}
+\section{Author Summary}\label{author-summary}
 
@@
 been associated with disease states, including a range of skin
-conditions (1), acute and chronic wound healing conditions (2, 3), and
-gastrointestinal diseases, such as inflammatory bowel disease (4, 5),
-\emph{Clostridium difficile} infections (6) and colorectal cancer (7,
-8). Altered human viromes (virus communities consisting primarily of
-bacteriophages) also have been associated with diseases and
-perturbations, including inflammatory bowel disease (5, 9), periodontal
-disease (10), spread of antibiotic resistance (11), and others (12--17).
-Viruses act in concert with their microbial hosts as a single ecological
-community (18). Viruses influence their living microbial host
-communities through processes including lysis, host gene expression
-modulation (19), influence on evolutionary processes such as horizontal
-gene transfer (22) or antagonistic co-evolution (26), and alteration of
-ecosystem processes and elemental stoichiometry (27).
+conditions {[}1{]}, acute and chronic wound healing conditions
+{[}2,3{]}, and gastrointestinal diseases, such as inflammatory bowel
+disease {[}4,5{]}, \emph{Clostridium difficile} infections {[}6{]} and
+colorectal cancer {[}7,8{]}. Altered human viromes (virus communities
+consisting primarily of bacteriophages) also have been associated with
+diseases and perturbations, including inflammatory bowel disease
+{[}5,9{]}, periodontal disease {[}10{]}, spread of antibiotic resistance
+{[}11{]}, and others {[}12--17{]}. Viruses act in concert with their
+microbial hosts as a single ecological community {[}18{]}. Viruses
+influence their living microbial host communities through processes
+including lysis, host gene expression modulation {[}19{]}, influence on
+evolutionary processes such as horizontal gene transfer {[}22{]} or
+antagonistic co-evolution {[}26{]}, and alteration of ecosystem
+processes and elemental stoichiometry {[}27{]}.
 
@@
 communities, but have reduced them to two separate communities by
-studying them independently (5, 9, 10, 12--17). This approach fails to
+studying them independently {[}5,9,10,12--17{]}. This approach fails to
 capture the complex dynamics of interacting bacteria and phage
@@
 together to maintain stable ecosystems. Removal of bacteria or phages
-can disrupt or even collapse those ecosystems (18, 28--37).
+can disrupt or even collapse those ecosystems {[}18,28--37{]}.
 Relationship-based network approaches allow us to capture this
@@
 communities by their inferred relationships using three published paired
-virus and bacteria-dominated whole community metagenomic datasets (13,
-14, 38, 39). We leveraged machine learning and graph theory techniques
-to establish and explore the human bacteria-phage network diversity
-therein. This approach built upon previous large-scale phage-bacteria
-network analyses by inferring interactions from metagenomic datasets,
-rather than culture-dependent data (33), which is limited in the scale
-of possible experiments and analyses. Our metagenomic interaction
-inference model improved upon previous models of phage-host predictions
-that have utilized a variety of techniques, such as linear models to
-predict bacteria-phage co-occurrence using taxonomic assignments (40),
-and nucleotide similarity models that were applied to both whole virus
-genomes (41) and related clusters of whole and partial virus genomes
-(42). Our approach uniquely included protein interaction data and was
-validated based on experimentally determined positive and negative
-interactions (i.e.~who does and does not infect whom). Through this
-approach we were able to provide a basic understanding of the network
-dynamics associated with phage and bacterial communities on and in the
-human body. By building and utilizing a microbiome network, we found
-that different people, body sites, and anatomical locations not only
-support distinct microbiome membership and diversity (13, 14, 38, 39,
-43--45), but also support ecological communities with distinct
-communication structures and propensities toward community instability.
-Through an improved understanding of network structures across the human
-body, we empower future studies to investigate how these communities
-dynamics are influenced by disease states and the overall impact they
-may have on human health.
+virus and bacteria-dominated whole community metagenomic datasets
+{[}13,14,38,39{]}. We leveraged machine learning and graph theory
+techniques to establish and explore the human bacteria-phage network
+diversity therein. This approach built upon previous large-scale
+phage-bacteria network analyses by inferring interactions from
+metagenomic datasets, rather than culture-dependent data {[}33{]}, which
+is limited in the scale of possible experiments and analyses. Our
+metagenomic interaction inference model improved upon previous models of
+phage-host predictions that have utilized a variety of techniques, such
+as linear models to predict bacteria-phage co-occurrence using taxonomic
+assignments {[}40{]}, and nucleotide similarity models that were applied
+to both whole virus genomes {[}41{]} and related clusters of whole and
+partial virus genomes {[}42{]}. Our approach uniquely included protein
+interaction data and was validated based on experimentally determined
+positive and negative interactions (i.e.~who does and does not infect
+whom). Through this approach we were able to provide a basic
+understanding of the network dynamics associated with phage and
+bacterial communities on and in the human body. By building and
+utilizing a microbiome network, we found that different people, body
+sites, and anatomical locations not only support distinct microbiome
+membership and diversity {[}13,14,38,39,43--45{]}, but also support
+ecological communities with distinct communication structures and
+propensities toward community instability. Through an improved
+understanding of network structures across the human body, we empower
+future studies to investigate how these communities dynamics are
+influenced by disease states and the overall impact they may have on
+human health.
 
@@
 three datasets that explored the impact of diet on the healthy human gut
-virome (14), the impact of anatomical location on the healthy human skin
-virome (13), and the viromes of monozygotic twins and their mothers (38,
-39). We selected these datasets because their virome samples were
-subjected to virus-like particle (VLP) purification, which removed
-contaminating DNA from human cells, bacteria, etc. To this end, the
-publishing authors employed combinations of filtration, chloroform/DNase
-treatment, and cesium chloride gradients to eliminate organismal DNA
-(e.g.~bacteria, human, fungi, etc) and thereby allow for direct
-assessment of both the extracellular and fully-assembled intracellular
-virome \textbf{(Supplemental Figure \ref{SequenceStats} A-B)} (14, 39).
-Each research group reported quality control measures to ensure the
-purity of the virome sequence datasets, using both computational and
-molecular techniques (e.g.~16S rRNA gene qPCR) \textbf{(Table S1)}.
-These reports confirmed that the virome libraries consisted of highly
-purified virus genomic DNA.
+virome {[}14{]}, the impact of anatomical location on the healthy human
+skin virome {[}13{]}, and the viromes of monozygotic twins and their
+mothers {[}38,39{]}. We selected these datasets because their virome
+samples were subjected to virus-like particle (VLP) purification, which
+removed contaminating DNA from human cells, bacteria, etc. To this end,
+the publishing authors employed combinations of filtration,
+chloroform/DNase treatment, and cesium chloride gradients to eliminate
+organismal DNA (e.g.~bacteria, human, fungi, etc) and thereby allow for
+direct assessment of both the extracellular and fully-assembled
+intracellular virome \textbf{(Supplemental Figure \ref{SequenceStats}
+A-B)} {[}14,39{]}. Each research group reported quality control measures
+to ensure the purity of the virome sequence datasets, using both
+computational and molecular techniques (e.g.~16S rRNA gene qPCR)
+\textbf{(Table S1)}. These reports confirmed that the virome libraries
+consisted of highly purified virus genomic DNA.
 
@@
 reports \textbf{(Supplemental Figure \ref{ContigStats} -
-\ref{ClusterStats})} (42). We referred to these operationally defined
-groups of related contigs as operational genomic units (OGUs). Each OGU
-represented a genomically similar sub-population of either bacteria or
-phages. Contig lengths within clusters ranged between \(10^{3}\) and
-\(10^{5.5}\) bp \textbf{(Supplemental Figure \ref{ContigStats} -
-\ref{ClusterStats})}.
+\ref{ClusterStats})} {[}42{]}. We referred to these operationally
+defined groups of related contigs as operational genomic units (OGUs).
+Each OGU represented a genomically similar sub-population of either
+bacteria or phages. Contig lengths within clusters ranged between
+\(10^{3}\) and \(10^{5.5}\) bp \textbf{(Supplemental Figure
+\ref{ContigStats} - \ref{ClusterStats})}.
 
@@
 similar integrated phages when sequenced--a trend previously reported
-(14). Additionally, we identified two OGUs as being complete phages
+{[}14{]}. Additionally, we identified two OGUs as being complete phages
 using the stringent Virsorter phage identification algorithm (class 1
-confidence group) (47).
+confidence group) {[}47{]}.
 
@@
 bacteria, with an average viral relative abundance of 0.4\%
-\textbf{(Table S1)} (13, 14, 38, 39). We found that only 2\% (6 / 280
+\textbf{(Table S1)} {[}13,14,38,39{]}. We found that only 2\% (6 / 280
 OGUs) of bacterial OGUs had significantly strong nucleotide similarity
-to phage reference genomes (e-value \textless{} \(10^{-25}\)) (13, 14,
-38, 39). No OGUs were confidently identified as lytic or temperate phage
-OGUs in the bacterial dataset using the Virsorter algorithm (47).
-Together this suggests only minimal ``contamination'' of the bacterial
-OGUs.
+to phage reference genomes (e-value \textless{} \(10^{-25}\))
+{[}13,14,38,39{]}. No OGUs were confidently identified as lytic or
+temperate phage OGUs in the bacterial dataset using the Virsorter
+algorithm {[}47{]}. Together this suggests only minimal
+``contamination'' of the bacterial OGUs.
 
@@
 random forest model trained on experimentally validated infectious
-relationships from six previous publications (41, 48--52). Only bacteria
-and phages were used in the model. The training set contained 43 diverse
-bacterial species and 30 diverse phage strains, including both broad and
-specific ranges of infection \textbf{(Supplemental Figure
+relationships from six previous publications {[}41,48--52{]}. Only
+bacteria and phages were used in the model. The training set contained
+43 diverse bacterial species and 30 diverse phage strains, including
+both broad and specific ranges of infection \textbf{(Supplemental Figure
 \ref{ValidationOverview} A - B)}. While it is true that there are more
@@
 models, thereby building upon previous work that only considered
-positive relationships (41).
+positive relationships {[}41{]}.
 
@@
 similarity of protein families associated with experimentally identified
-protein-protein interactions (53). The resulting random forest model was
-assessed using nested cross validation, and the median area under its
-receiver operating characteristic (ROC) curve was 0.788, the median
+protein-protein interactions {[}53{]}. The resulting random forest model
+was assessed using nested cross validation, and the median area under
+its receiver operating characteristic (ROC) curve was 0.788, the median
 model sensitivity was 0.952, and median specificity was 0.615
@@
 \ref{RocCurve} C)}.
-
-We used our random forest model to classify the relationships between
-bacteria and phage operational genomic units, which were then used to
-build the interactive network. The master network contained the three
-studies as sub-networks, which themselves each contained sub-networks
-for each sample \textbf{(Figure \ref{RocCurve} D)}. Metadata including
-study, sample ID, disease, and OGU abundance within the community were
-stored in the master network for parsing in downstream analyses
-\textbf{(Supplemental Figure \ref{NetworkDiagram})}. Bacterial and phage
-relative abundance was recorded in each sample for each OGU and the
-weight of the edge connecting those OGUs was calculated as a function of
-those relative abundance values. The separate extraction of the phage
-and bacterial libraries ensured a more accurate measurement of the
-microbial communities, as has been outlined previously (54, 55). The
-master network was highly connected and contained 72,287 infectious
-relationships among 578 nodes, representing 298 phages and 280 bacteria.
-Although the network was highly connected, not all relationships were
-present in all samples. Relationships were weighted by the relative
-abundances of their associated bacteria and phages. Like the master
-network, the skin network exhibited a diameter of 4 (measure of graph
-size; the greatest number of traversed vertices required between two
-vertices) and included 576 (297 phages, 279 bacteria, 99.7\% total) and
-72,127 (99.8\%) of the master network nodes and edges, respectively
-\textbf{(Figure \ref{RocCurve} E - F)}. The phages and bacteria in the
-diet and twin sample sets were more sparsely related, with the diet
-study consisting of 89 (41 phages, 48 bacteria) nodes and 5,566
-relationships, and the twin study containing 137 (36 phages, 101
-bacteria) nodes and 17,250 relationships \textbf{(Figure \ref{RocCurve}
-E - F)}.
-
-\subsection{Role of Diet \& Obesity in Gut Microbiome
-Connectivity}\label{role-of-diet-obesity-in-gut-microbiome-connectivity}
-
-Diet is a major environmental factor that influences resource
-availability and gut microbiome composition and diversity, including
-bacteria and phages (14, 56, 57). Previous work in isolated
-culture-based systems has suggested that changes in nutrient
-availability are associated with altered phage-bacteria network
-structures (30), although this has yet to be tested in humans. We
-therefore hypothesized that a change in diet would also be associated
-with a change in virome-microbiome network structure in the human gut.
-
-We evaluated the diet-associated differences in gut virome-microbiome
-network structure by quantifying how central each sample's network was
-on average. We accomplished this by utilizing two common centrality
-metrics: degree centrality and closeness centrality. Degree centrality,
-the simplest centrality metric, was defined as the number of connections
-each phage made with each bacterium. We supplemented measurements of
-degree centrality with measurements of closeness centrality. Closeness
-centrality is a metric of how close each phage or bacterium is to all of
-the other phages and bacteria in the network. A higher closeness
-centrality suggests that the effects of genetic information or altered
-abundance would be more impactful to all other microbes in the system. A
-network with higher average closeness centrality also indicates an
-overall greater degree of connections, which suggests a greater
-resilience against instability. This is because more highly connected
-networks are more stable as a result of pathway dependencies and the
-unlikelihood that a randomly removed bacteria or phage would cause major
-divisions across the network (30, 58). We used this information to
-calculate the average connectedness per sample, which was corrected for
-the maximum potential degree of connectedness. Unfortunately our dataset
-was insufficiently powered to make strong conclusions toward this
-hypothesis, but this is an interesting observation that warrants further
-investigation.
-
-Using our limited sample set, we observed that the gut microbiome
-network structures associated with high-fat diets appeared less
-connected than those of low-fat diets, although a greater sample size
-will be required to more properly evaluate this trend \textbf{(Figure
-\ref{dietnetworks} A-B)}. Five subjects were available for use, all of
-which had matching bacteria and virome datasets and samples from 8-10
-days following the initiation of their diets. High-fat diets appeared to
-exhibit reduced degree centrality \textbf{(Figure \ref{dietnetworks}
-A)}, suggesting bacteria in high-fat environments were targeted by fewer
-phages and that phage tropism was more restricted. High-fat diets also
-appeared to exhibit decreased closeness centrality \textbf{(Figure
-\ref{dietnetworks} B)}, indicating that bacteria and phages were more
-distant from other bacteria and phages in the community. This would make
-genetic transfer and altered abundance of a given phage or bacterium
-less capable of impacting other bacteria and phages within the network.
-
-In addition to diet, we found preliminary evidence that obesity
-influenced network structure. This was done using the three mother
-samples available from the twin sample set, all of which had matching
-bacteria and phage samples and confirmed BMI information. The
-obesity-associated network appeared to have a higher degree centrality
-\textbf{(Figure \ref{dietnetworks} C)}, but less closeness centrality
-than the healthy-associated networks \textbf{(Figure \ref{dietnetworks}
-D)}. These results suggested that the obesity-associated networks are
-less connected, having microbes further from all other microbes within
-the community. This again comes with the caveat that this is a
-preliminary observation with too small of a sample size to make more
-substantial claims.
-
-\subsection{Individuality of Microbial
-Networks}\label{individuality-of-microbial-networks}
-
-Skin and gut community membership and diversity are highly personal,
-with people remaining more similar to themselves than to other people
-over time (13, 59, 60). We therefore hypothesized that this personal
-conservation extended to microbiome network structure. We addressed this
-hypothesis by calculating the degree of dissimilarity between each
-subject's network, based on phage and bacteria abundance and centrality.
-We quantified phage and bacteria centrality within each sample graph
-using the weighted eigenvector centrality metric. This metric defines
-central phages as those that are highly abundant (\(A_{O}\) as defined
-in the methods) and infect many distinct bacteria which themselves are
-abundant and infected by many other phages. Similarly, bacterial
-centrality was defined as those bacteria that were both abundant and
-connected to numerous phages that were themselves connected to many
-bacteria. We then calculated the similarity of community networks using
-the weighted eigenvector centrality of all nodes between all samples.
-Samples with similar network structures were interpreted as having
-similar capacities for maintaining stability and transmitting genetic
-material.
-
-We used this network dissimilarity metric to test whether microbiome
-network structures were more similar within people than between people
-over time. We found that gut microbiome network structures clustered by
-person (ANOSIM p-value = 0.005, R = 0.958, \textbf{Figure \ref{intradiv}
-A}). Network dissimilarity within each person over the 8-10 day sampling
-period was less than the average dissimilarity between that person and
-others, although this difference was not statistically significant
-(p-value = 0.125, \textbf{Figure \ref{intradiv} B}). Four of the five
-available subjects were used because one of the subjects was not sampled
-at the initial time point. The lack of statistical confidence was likely
-due to the small sample size of this dataset.
-
-Although there was evidence for gut network conservation among
-individuals, we found no evidence for conservation of gut network
-structures within families. The gut network structures were not more
-similar within families (twins and their mothers; intrafamily) compared
-to other families (other twins and mothers; inter-family) (p-value =
-0.312, \textbf{Figure \ref{intradiv} C}). In addition to the gut, skin
-microbiome network structure was strongly conserved within individuals
-(p-value \textless{} 0.001, \textbf{Figure \ref{intradiv} D}). This
-distribution was similar when separated by anatomical sites. Most sites
-were statistically significantly more conserved within individuals
-\textbf{(Supplemental Figure \ref{allskin})}.
-
-\subsection{Association Between Environmental Stability and Network
-Structure Across the Human Skin
-Landscape}\label{association-between-environmental-stability-and-network-structure-across-the-human-skin-landscape}
-
-Extensive work has illustrated differences in diversity and composition
-of the healthy human skin microbiome between anatomical sites, including
-bacteria, virus, and fungal communities (13, 44, 59). These communities
-vary by degree of skin moisture, oil, and environmental exposure. As
-viruses are known to influence microbial diversity and community
-composition, we hypothesized that microbe-virus network structure would
-be specific to anatomical sites, as well. To test this, we evaluated the
-changes in network structure between anatomical sites within the skin
-dataset.
-
-The average centrality of each sample was quantified using the weighted
-eigenvector centrality metric. Intermittently moist skin sites (dynamic
-sites that fluctuate between being moist and dry) were significantly
-less connected than the more stable moist and sebaceous environments
-(p-value \textless{} 0.001, \textbf{Figure \ref{skinnetwork} A)}. Also,
-skin sites that were occluded from the environment were much more highly
-connected than those that were constantly exposed to the environment or
-only intermittently occluded (p-value \textless{} 0.001, \textbf{Figure
-\ref{skinnetwork} B)}.
-
-To supplement this analysis, we compared the network signatures using
-the centrality dissimilarity approach described above. The dissimilarity
-between samples was a function of shared relationships, degree of
-centrality, and bacteria/phage abundance. When using this supplementary
-approach, we found that network structures significantly clustered by
-moisture, sebaceous, and intermittently moist status \textbf{(Figure
-\ref{skinnetwork} C,E)}. Occluded sites were significantly different
-from exposed and intermittently occluded sites, but there was no
-difference between exposed and intermittently occluded sites
-\textbf{(Figure \ref{skinnetwork} D,F)}. These findings provide further
-support that skin microbiome network structure differs significantly
-between skin sites.
-
-\section{Discussion}\label{discussion}
-
-Foundational work has provided a baseline understanding of the human
-microbiome by characterizing bacterial and viral diversity across the
-human body (13, 14, 43--45, 61). Here, we offer an initial understanding
-of how phage-bacteria networks differ throughout the human body, so as
-to provide a baseline for future studies of how and why microbiome
-networks differ in disease states. We developed and implemented a
-network-based analytical model to evaluate the basic properties of the
-human microbiome through bacteria and phage relationships, instead of
-membership or diversity alone. This enabled the application of network
-theory to provide a new perspective on complex ecological communities.
-We utilized metrics of connectivity to model the extent to which
-communities of bacteria and phages interact through mechanisms such as
-horizontal gene transfer, modulated bacterial gene expression, and
-alterations in abundance.
-
-Just as gut microbiome and virome composition and diversity are
-conserved in individuals (13, 43, 44, 60), gut and skin microbiome
-network structures were conserved within individuals over time. Gut
-network structure was not conserved among family members. These findings
-suggested that the community properties inferred from microbiome
-interaction network structures, such as stability (meaning a more highly
-connected network is more stable because a randomly removed bacteria or
-phage node is less likely to divide or disintegrate (30, 58) the overall
-network), the potential for horizontal gene transfer between members,
-and co-evolution of populations, were person-specific. These properties
-may be impacted by personal factors ranging from the body's immune
-system to external environmental conditions, such as climate and diet.
-
-We observed evidence supporting the ability of environmental conditions
-to shape gut and skin microbiome interaction network structure by
-observing that diet and skin location were associated with altered
-network structures. We found evidence that diet was sufficient to alter
-gut microbiome network connectivity, although this needs to be
-interpreted as a case observation, due to the small sample size.
-Although our sample size was small, our findings provided some
-preliminary evidence that high-fat diets were less connected than
-low-fat diets and that high-fat diets therefore may lead to less stable
-communities with a decreased ability for microbes to directly influence
-one another. We supported this finding with the observation that obesity
-may have been associated with decreased network connectivity. Together
-these findings suggest the food we eat may not only impact which
-microbes colonize our guts, but may also impact their interactions with
-infecting phages. Further work will be required to characterize these
-relationships with a larger cohort.
-
-In addition to diet, the skin environment also influenced the microbiome
-interaction network structure. Network structure differed between
-environmentally exposed and occluded skin sites. The sites under greater
-environmental fluctuation and exposure (the exposed and intermittently
-exposed sites) were less connected and therefore were predicted to have
-a higher propensity for instability. Likewise, intermittently moist
-sites demonstrated less connectedness than the more stable moist and
-sebaceous sites. Together these data suggested that body sites under
-greater degrees of fluctuation harbored less connected, potentially less
-stable microbiomes. This points to a link between microbiome and
-environmental stability and warrants further investigation.
-
-While these findings take us an important step closer to understanding
-the microbiome through interspecies relationships, there are caveats to
-and considerations regarding the approach. First, as with most
-classification models, the infection classification model developed and
-applied is only as good as its training set -- in this case, the
-collection of experimentally-verified positive and negative infection
-data, where genomes of all members are fully sequenced. Large-scale
-experimental screens for phage and bacteria infectious interactions that
-report high-confidence negative interactions (i.e., no infection) are
-desperately needed, as they would provide more robust model training and
-improved model performance. Furthermore, just as we have improved on
-previous modeling efforts, we expect that new and creative scoring
-metrics will be integrated into this model to improve future
-performance.
-
-Second, although our analyses utilized the best datasets currently
-available for our study, this work was done retrospectively and relied
-on existing data up to seven years old. These archived datasets were
-limited by the technology and costs of the time. For example, the diet
-and twin studies, relied on multiple displacement amplification (MDA) in
-their library preparations--an approach used to overcome the large
-nucleic acids requirements typical of older sequencing library
-generation protocols. It is now known that MDA results in biases in
-microbial community composition (62), as well as toward ssDNA viral
-genomes (63, 64), thus rendering the resulting microbial and viral
-metagenomes largely non-quantitative. Future work that employs larger
-sequence datasets and that avoids the use of bias-inducing amplification
-steps will build on and validate our findings, as well as inform the
-design and interpretation of further studies.
-
-Finally, the networks in this study were built using operational genomic
-units (OGUs), which represented groups of highly similar bacteria or
-phage genomes or clustered genome fragments. Similar clustering
-definition and validation methods, both computational and experimental,
-have been implemented in other metagenomic sequencing studies, as well
-(42, 65--67). These approaches could offer yet another level of
-sophistication to our network-based analyses. While this operationally
-defined clustering approach allows us to study whole community networks,
-our ability to make conclusions about interactions among specific phage
-or bacterial species or populations is inherently limited, compared to
-more focused, culture-based studies such as the work by Malki \emph{et
-al} (68). Future work must address this limitation, e.g., through
-improved binning methods and deeper metagenomic shotgun sequencing, but
-most importantly through an improved conceptual framing of what defines
-ecologically and evolutionarily cohesive units for both phage and
-bacteria (69). Defining operational genomic units and their taxonomic
-underpinnings (e.g., whether OGU clusters represent genera or species)
-is an active area of work critical to the utility of this approach. As a
-first step, phylogenomic analyses have been performed to cluster
-cyanophage isolate genomes into informative groups using shared gene
-content, average nucleotide identity of shared genes, and pairwise
-differences between genomes (70). Such population-genetic assessment of
-phage evolution, coupled with the ecological implications of genome
-heterogeneity, will inform how to define nodes in future iterations of
-the ecological network developed here. Even though we are hesitant to
-speculate on phage host ranges at low taxonomic levels in our dataset,
-the data does aggree with previous reports of instances of broad phage
-host range (68, 71).
-
-Together our work takes an initial step towards defining bacteria-virus
-interaction profiles as a characteristic of human-associated microbial
-communities. This approach revealed the impacts that different human
-environments (e.g., the skin and gut) can have on microbiome
-connectivity. By focusing on relationships between bacterial and viral
-communities, they are studied as the interacting cohorts they are,
-rather than as independent entities. While our developed bacteria-phage
-interaction framework is a novel conceptual advance, the microbiome also
-consists of archaea and small eukaryotes, including fungi and
-\emph{Demodex} mites (1, 72)--all of which can interact with human
-immune cells and other non-microbial community members (73). Future work
-will build from our approach and include these additional community
-members and their diverse interactions and relationships (e.g., beyond
-phage-bacteria). This will result in a more robust network and a more
-holistic understanding of the evolutionary and ecological processes that
-drive the assembly and function of the human-associated microbiome.
-
-\section{Materials \& Methods}\label{materials-methods}
-
-\subsection{Code Availability}\label{code-availability}
-
-A reproducible version of this manuscript written in R markdown and all
-of the code used to obtain and process the sequencing data is available
-at the following GitHub repository:
-
-https://github.com/SchlossLab/Hannigan\_ConjunctisViribus\_mSystems\_2017
-
-\subsection{Data Acquisition \& Quality
-Control}\label{data-acquisition-quality-control}
-
-Raw sequencing data and associated metadata were acquired from the NCBI
-sequence read archive (SRA). Supplementary metadata were acquired from
-the same SRA repositories and their associated manuscripts. The gut
-virome diet study (SRA: \texttt{SRP002424}), twin virome studies (SRA:
-\texttt{SRP002523}; \texttt{SRP000319}), and skin virome study (SRA:
-\texttt{SRP049645}) were downloaded as \texttt{.sra} files. Sequencing
-files were converted to \texttt{fastq} format using the
-\texttt{fastq-dump} tool of the NCBI SRA Toolkit (v2.2.0). Sequences
-were quality trimmed using the Fastx toolkit (v0.0.14) to exclude bases
-with quality scores below 33 and shorter than 75 bp (74). Paired end
-reads were filtered to exclude sequences missing their corresponding
-pair using the \texttt{get\_trimmed\_pairs.py} script available in the
-source code.
-
-\subsection{Contig Assembly}\label{contig-assembly}
-
-Contigs were assembled using the Megahit assembly program (v1.0.6) (75).
-A minimum contig length of 1 kb was used. Iterative k-mer stepping began
-at a minimum length of 21 and progressed by 20 until 101. All other
-default parameters were used.
-
-\subsection{Contig Abundance
-Calculations}\label{contig-abundance-calculations}
-
-Contigs were concatenated into two master files prior to alignment, one
-for bacterial contigs and one for phage contigs. Sample sequences were
-aligned to phage or bacterial contigs using the Bowtie2 global aligner
-(v2.2.1) (76). We defined a mismatch threshold of 1 bp and seed length
-of 25 bp. Sequence abundance was calculated from the Bowtie2 output
-using the \texttt{calculate\_abundance\_from\_sam.pl} script available
-in the source code.
-
-\subsection{Operational Genomic Unit
-Binning}\label{operational-genomic-unit-binning}
-
-Contigs often represent large fragments of genomes. In order to reduce
-redundancy and the resulting artificially inflated genomic richness
-within our dataset, it was important to bin contigs into operational
-units based on their similarity. This approach is conceptually similar
-to the clustering of related 16S rRNA sequences into operational
-taxonomic units (OTUs), although here we are clustering contigs into
-operational genomic units (OGUs) (61).
-
-Contigs were clustered using the CONCOCT algorithm (v0.4.0) (77).
-Because of our large dataset and limits in computational efficiency, we
-randomly subsampled the dataset to include 25\% of all samples, and used
-these to inform contig abundance within the CONCOCT algorithm. CONCOCT
-was used with a maximum of 500 clusters, a k-mer length of four, a
-length threshold of 1 kb, 25 iterations, and exclusion of the total
-coverage variable.
-
-OGU abundance (\(A_{O}\)) was obtained as the sum of the abundance of
-each contig (\(A_{j}\)) associated with that OGU. The abundance values
-were length corrected such that:
-
-\[ { A }_{ O }=\frac { { 10 }^{ 7 }\sum _{ j=1 }^{ k }{ { A }_{ j } }  }{ \sum _{ j=1 }^{ k }{ { L }_{ j } }  } \]
-
-Where \texttt{L} is the length of each contig \texttt{j} within the OGU.
-
-\subsection{Operational Genomic Unit
-Identification}\label{operational-genomic-unit-identification}
-
-To confirm a lack of phage sequences in the bacterial OGU dataset, we
-performed blast nucleotide alignment of the bacterial OGU representative
-sequences using an e-value \textless{} \(10^{-25}\), which was stricter
-than the \(10^{-10}\) threshold used in the random forest model below,
-against all of the phage reference genomes available in the EMBL
-database. We used a stricter threshold because we know there are genomic
-similarities between bacteria and phage OGUs from the interactive model,
-but we were interested in contigs with high enough similarity to
-references that they may indeed be from phages. We also performed the
-converse analysis of aligning phage OGU representative sequences to EMBL
-bacterial reference genomes. Finally, we ran both the phage and bacteria
-OGU representative sequences through the Virsorter program (1.0.3) to
-identify phages (all default parameters were used), using only those in
-the high confidence identification category ``class 1'' (47).
-
-\subsection{Open Reading Frame
-Prediction}\label{open-reading-frame-prediction}
-
-Open reading frames (ORFs) were identified using the Prodigal program
-(V2.6.2) with the meta mode parameter and default settings (78).
-
-\subsection{Classification Model Creation and
-Validation}\label{classification-model-creation-and-validation}
-
-The classification model for predicting interactions was built using
-experimentally validated bacteria-phage infections or validated lack of
-infections from six studies (41, 48--52). Associated reference genomes
-were downloaded from the European Bioinformatics Institute (see details
-in source code). The model was created based on the four metrics listed
-below.
-
-The four scores were used as parameters in a random forest model to
-classify bacteria and bacteriophage pairs as either having infectious
-interactions or not. The classification model was built using the Caret
-R package (v6.0.73) (79). The model was trained using five-fold cross
-validation with ten repeats, and the median model performance was
-evaluated by training the model on 80\% of the dataset and testing
-performance on the remaining 20\%. Pairs without scores were classified
-as not interacting. The model was optimized using the ROC value. The
-resulting model performance was plotted using the plotROC R package.
-
-\subparagraph{Identify Bacterial CRISPRs Targeting
-Phages}\label{identify-bacterial-crisprs-targeting-phages}
-
-Clustered Regularly Interspaced Short Palindromic Repeats (CRISPRs) were
-identified from bacterial genomes using the PilerCR program (v1.06)
-(80). Resulting spacer sequences were filtered to exclude spacers
-shorter than 20 bp and longer than 65 bp. Spacer sequences were aligned
-to the phage genomes using the nucleotide BLAST algorithm with default
-parameters (v2.4.0) (81). The mean percent identity for each matching
-pair was recorded for use in our classification model.
-
-\subparagraph{Detect Matching Prophages within Bacterial
-Genomes}\label{detect-matching-prophages-within-bacterial-genomes}
-
-Temperate bacteriophages infect and integrate into their bacterial
-host's genome. We detected integrated phage elements within bacterial
-genomes by aligning phage genomes to bacterial genomes using the
-nucleotide BLAST algorithm and a minimum e-value of 1e-10. The resulting
-bitscore of each alignment was recorded for use in our classification
-model.
-
-\subparagraph{Identify Shared Genes Between Bacteria and
-Phages}\label{identify-shared-genes-between-bacteria-and-phages}
-
-As a result of gene transfer or phage genome integration during
-infection, phages may share genes with their bacterial hosts, providing
-us with evidence of phage-host pairing. We identified shared genes
-between bacterial and phage genomes by assessing amino acid similarity
-between the genes using the Diamond protein alignment algorithm
-(v0.7.11.60) (82). The mean alignment bitscores for each genome pair
-were recorded for use in our classification model.
-
-\subparagraph{Protein - Protein
-Interactions}\label{protein---protein-interactions}
-
-The final method used for predicting infectious interactions between
-bacteria and phages was the detection of pairs of genes whose proteins
-are known to interact. We assigned bacterial and phage genes to protein
-families by aligning them to the Pfam database using the Diamond protein
-alignment algorithm. We then identified which pairs of proteins were
-predicted to interact using the Pfam interaction information within the
-Intact database (53). The mean bitscores of the matches between each
-pair were recorded for use in the classification model.
-
-\subsection{Interaction Network
-Construction}\label{interaction-network-construction}
-
-The bacteria and phage operational genomic units (OGUs) were scored
-using the same approach as outlined above. The infectious pairings
-between bacteria and phage OGUs were classified using the random forest
-model described above. The predicted infectious pairings and all
-associated metadata were used to populate a graph database using Neo4j
-graph database software (v2.3.1) (83). This network was used for
-downstream community analysis.
-
-\subsection{Centrality Analysis}\label{centrality-analysis}
-
-We quantified the centrality of graph vertices using three different
-metrics, each of which provided different information graph structure.
-When calculating these values, let \(G(V,E)\) be an undirected,
-unweighted graph with \(|V|=n\) nodes and \(|E|=m\) edges. Also, let
-\(\mathbf{A}\) be its corresponding adjacency matrix with entries
-\(a_{ij} = 1\) if nodes \(V_i\) and \(V_j\) are connected via an edge,
-and \(a_{ij} = 0\) otherwise.
-
-Briefly, the \textbf{closeness centrality} of node \(V_i\) is calculated
-taking the inverse of the average length of the shortest paths
-(\texttt{d}) between nodes \(V_i\) and all the other nodes \(V_j\).
-Mathematically, the closeness centrality of node \(V_i\) is given as:
-
-\[ { C }_{ C }\left( { V }_{ i } \right) ={ \left( \sum _{ j=1 }^{ n }{ d\left( { V }_{ i },{ V }_{ j } \right)  }  \right)  }^{ -1 } \]
-
-The distance between nodes (\texttt{d}) was calculated as the shortest
-number of edges required to be traversed to move from one node to
-another.
-
-Intuitively, the \textbf{degree centrality} of node \(V_i\) is defined
-as the number of edges that are incident to that node:
-
-\[ { C }_{ D }\left( { V }_{ i } \right) =\sum _{ j=1 }^{ n }{ { a }_{ ij } } \]
-
-where \(a_{ij}\) is the \(ij^{th}\) entry in the adjacency matrix
-\(\mathbf{A}\).
-
-The eigenvector centrality of node \(V_i\) is defined as the \(i^{th}\)
-value in the first eigenvector of the associated adjacency matrix
-\(\mathbf{A}\). Conceptually, this function results in a centrality
-value that reflects the connections of the vertex, as well as the
-centrality of its neighboring vertices.
-
-The \textbf{centralization} metric was used to assess the average
-centrality of each sample graph \(\mathbf{G}\). Centralization was
-calculated by taking the sum of each vertex \(V_{i}\)'s centrality from
-the graph maximum centrality \(C_{w}\), such that:
-
-\[ C\left( G \right) =\frac { \sum _{ i=1 }^{ n }{ Cw -c\left( { V }_{ i } \right)  }  }{ { T } }  \]
-
-The values were corrected for uneven graph sizes by dividing the
-centralization score by the maximum theoretical centralization
-(\texttt{T}) for a graph with the same number of vertices.
-
-Degree and closeness centrality were calculated using the associated
-functions within the igraph R package (v1.0.1) (84).
-
-\subsection{Network Relationship
-Dissimilarity}\label{network-relationship-dissimilarity}
-
-We assessed similarity between graphs by evaluating the shared
-centrality of their vertices, as has been done previously. More
-specifically, we calculated the dissimilarity between graphs \(G_{i}\)
-and \(G_{j}\) using the Bray-Curtis dissimilarity metric and eigenvector
-centrality values such that:
-
-\[ B\left( { G }_{ i },{ G }_{ j } \right) =1-\frac { 2{ C }_{ ij } }{ { C }_{ i }+{ C }_{ j } } \]
-
-Where \(C_{ij}\) is the sum of the lesser centrality values for those
-vertices shared between graphs, and \(C_{i}\) and \(C_{j}\) are the
-total number of vertices found in each graph. This allows us to
-calculate the dissimilarity between graphs based on the shared
-centrality values between the two graphs.
-
-\subsection{Statistics and
-Comparisons}\label{statistics-and-comparisons}
-
-Differences in intrapersonal and interpersonal network structure
-diversity, based on multivariate data, were calculated using an analysis
-of similarity (ANOSIM). Statistical significance of univariate
-Eigenvector centrality differences were calculated using a paired
-Wilcoxon test.
-
-Statistical significance of differences in univariate eigenvector
-centrality measurements of skin virome-microbiome networks were
-calculated using a pairwise Wilcoxon test, corrected for multiple
-hypothesis tests using the Holm correction method. Multivariate
-eigenvector centrality was measured as the mean differences between
-cluster centroids, with statistical significance measured using an ANOVA
-and post hoc Tukey test.
-
-\section{Acknowledgments}\label{acknowledgments}
-
-We thank the members of the Schloss lab for their underlying
-contributions. We thank the authors of the original studies for making
-their data and metadata publicly available and understandable. We also
-thank the participants in the studies.
-
-\section{Funding Information}\label{funding-information}
-
-GDH was supported in part by the Molecular Mechanisms in Microbial
-Pathogenesis Training Program (T32 AI007528). GDH and PDS were supported
-in part by funding from the NIH (P30DK034933, U19AI09087, and
-U01AI124255).
-
-\section{Disclosure Declaration}\label{disclosure-declaration}
-
-The authors report no conflicts of interest.
-
-\newpage
-
-\section{Figure Legends}\label{figure-legends}
 
@@
 
+We used our random forest model to classify the relationships between
+bacteria and phage operational genomic units, which were then used to
+build the interactive network. The master network contained the three
+studies as sub-networks, which themselves each contained sub-networks
+for each sample \textbf{(Figure \ref{RocCurve} D)}. Metadata including
+study, sample ID, disease, and OGU abundance within the community were
+stored in the master network for parsing in downstream analyses
+\textbf{(Supplemental Figure \ref{NetworkDiagram})}. Bacterial and phage
+relative abundance was recorded in each sample for each OGU and the
+weight of the edge connecting those OGUs was calculated as a function of
+those relative abundance values. The separate extraction of the phage
+and bacterial libraries ensured a more accurate measurement of the
+microbial communities, as has been outlined previously {[}54,55{]}. The
+master network was highly connected and contained 72,287 infectious
+relationships among 578 nodes, representing 298 phages and 280 bacteria.
+Although the network was highly connected, not all relationships were
+present in all samples. Relationships were weighted by the relative
+abundances of their associated bacteria and phages. Like the master
+network, the skin network exhibited a diameter of 4 (measure of graph
+size; the greatest number of traversed vertices required between two
+vertices) and included 576 (297 phages, 279 bacteria, 99.7\% total) and
+72,127 (99.8\%) of the master network nodes and edges, respectively
+\textbf{(Figure \ref{RocCurve} E - F)}. The phages and bacteria in the
+diet and twin sample sets were more sparsely related, with the diet
+study consisting of 89 (41 phages, 48 bacteria) nodes and 5,566
+relationships, and the twin study containing 137 (36 phages, 101
+bacteria) nodes and 17,250 relationships \textbf{(Figure \ref{RocCurve}
+E - F)}.
+
+\subsection{Role of Diet \& Obesity in Gut Microbiome
+Connectivity}\label{role-of-diet-obesity-in-gut-microbiome-connectivity}
+
+Diet is a major environmental factor that influences resource
+availability and gut microbiome composition and diversity, including
+bacteria and phages {[}14,56,57{]}. Previous work in isolated
+culture-based systems has suggested that changes in nutrient
+availability are associated with altered phage-bacteria network
+structures {[}30{]}, although this has yet to be tested in humans. We
+therefore hypothesized that a change in diet would also be associated
+with a change in virome-microbiome network structure in the human gut.
+
+We evaluated the diet-associated differences in gut virome-microbiome
+network structure by quantifying how central each sample's network was
+on average. We accomplished this by utilizing two common centrality
+metrics: degree centrality and closeness centrality. Degree centrality,
+the simplest centrality metric, was defined as the number of connections
+each phage made with each bacterium. We supplemented measurements of
+degree centrality with measurements of closeness centrality. Closeness
+centrality is a metric of how close each phage or bacterium is to all of
+the other phages and bacteria in the network. A higher closeness
+centrality suggests that the effects of genetic information or altered
+abundance would be more impactful to all other microbes in the system. A
+network with higher average closeness centrality also indicates an
+overall greater degree of connections, which suggests a greater
+resilience against instability. This is because more highly connected
+networks are more stable as a result of pathway dependencies and the
+unlikelihood that a randomly removed bacteria or phage would cause major
+divisions across the network {[}30,58{]}. We used this information to
+calculate the average connectedness per sample, which was corrected for
+the maximum potential degree of connectedness. Unfortunately our dataset
+was insufficiently powered to make strong conclusions toward this
+hypothesis, but this is an interesting observation that warrants further
+investigation.
+
+Using our limited sample set, we observed that the gut microbiome
+network structures associated with high-fat diets appeared less
+connected than those of low-fat diets, although a greater sample size
+will be required to more properly evaluate this trend \textbf{(Figure
+\ref{dietnetworks} A-B)}. Five subjects were available for use, all of
+which had matching bacteria and virome datasets and samples from 8-10
+days following the initiation of their diets. High-fat diets appeared to
+exhibit reduced degree centrality \textbf{(Figure \ref{dietnetworks}
+A)}, suggesting bacteria in high-fat environments were targeted by fewer
+phages and that phage tropism was more restricted. High-fat diets also
+appeared to exhibit decreased closeness centrality \textbf{(Figure
+\ref{dietnetworks} B)}, indicating that bacteria and phages were more
+distant from other bacteria and phages in the community. This would make
+genetic transfer and altered abundance of a given phage or bacterium
+less capable of impacting other bacteria and phages within the network.
+
 \begin{figure}[htbp]
@@
 \end{figure}
+
+In addition to diet, we found preliminary evidence that obesity
+influenced network structure. This was done using the three mother
+samples available from the twin sample set, all of which had matching
+bacteria and phage samples and confirmed BMI information. The
+obesity-associated network appeared to have a higher degree centrality
+\textbf{(Figure \ref{dietnetworks} C)}, but less closeness centrality
+than the healthy-associated networks \textbf{(Figure \ref{dietnetworks}
+D)}. These results suggested that the obesity-associated networks are
+less connected, having microbes further from all other microbes within
+the community. This again comes with the caveat that this is a
+preliminary observation with too small of a sample size to make more
+substantial claims.
+
+\subsection{Individuality of Microbial
+Networks}\label{individuality-of-microbial-networks}
+
+Skin and gut community membership and diversity are highly personal,
+with people remaining more similar to themselves than to other people
+over time {[}13,59,60{]}. We therefore hypothesized that this personal
+conservation extended to microbiome network structure. We addressed this
+hypothesis by calculating the degree of dissimilarity between each
+subject's network, based on phage and bacteria abundance and centrality.
+We quantified phage and bacteria centrality within each sample graph
+using the weighted eigenvector centrality metric. This metric defines
+central phages as those that are highly abundant (\(A_{O}\) as defined
+in the methods) and infect many distinct bacteria which themselves are
+abundant and infected by many other phages. Similarly, bacterial
+centrality was defined as those bacteria that were both abundant and
+connected to numerous phages that were themselves connected to many
+bacteria. We then calculated the similarity of community networks using
+the weighted eigenvector centrality of all nodes between all samples.
+Samples with similar network structures were interpreted as having
+similar capacities for maintaining stability and transmitting genetic
+material.
+
+We used this network dissimilarity metric to test whether microbiome
+network structures were more similar within people than between people
+over time. We found that gut microbiome network structures clustered by
+person (ANOSIM p-value = 0.005, R = 0.958, \textbf{Figure \ref{intradiv}
+A}). Network dissimilarity within each person over the 8-10 day sampling
+period was less than the average dissimilarity between that person and
+others, although this difference was not statistically significant
+(p-value = 0.125, \textbf{Figure \ref{intradiv} B}). Four of the five
+available subjects were used because one of the subjects was not sampled
+at the initial time point. The lack of statistical confidence was likely
+due to the small sample size of this dataset.
+
+Although there was evidence for gut network conservation among
+individuals, we found no evidence for conservation of gut network
+structures within families. The gut network structures were not more
+similar within families (twins and their mothers; intrafamily) compared
+to other families (other twins and mothers; inter-family) (p-value =
+0.312, \textbf{Figure \ref{intradiv} C}). In addition to the gut, skin
+microbiome network structure was strongly conserved within individuals
+(p-value \textless{} 0.001, \textbf{Figure \ref{intradiv} D}). This
+distribution was similar when separated by anatomical sites. Most sites
+were statistically significantly more conserved within individuals
+\textbf{(Supplemental Figure \ref{allskin})}.
 
@@
 
+\subsection{Association Between Environmental Stability and Network
+Structure Across the Human Skin
+Landscape}\label{association-between-environmental-stability-and-network-structure-across-the-human-skin-landscape}
+
+Extensive work has illustrated differences in diversity and composition
+of the healthy human skin microbiome between anatomical sites, including
+bacteria, virus, and fungal communities {[}13,44,59{]}. These
+communities vary by degree of skin moisture, oil, and environmental
+exposure. As viruses are known to influence microbial diversity and
+community composition, we hypothesized that microbe-virus network
+structure would be specific to anatomical sites, as well. To test this,
+we evaluated the changes in network structure between anatomical sites
+within the skin dataset.
+
+The average centrality of each sample was quantified using the weighted
+eigenvector centrality metric. Intermittently moist skin sites (dynamic
+sites that fluctuate between being moist and dry) were significantly
+less connected than the more stable moist and sebaceous environments
+(p-value \textless{} 0.001, \textbf{Figure \ref{skinnetwork} A)}. Also,
+skin sites that were occluded from the environment were much more highly
+connected than those that were constantly exposed to the environment or
+only intermittently occluded (p-value \textless{} 0.001, \textbf{Figure
+\ref{skinnetwork} B)}.
+
 \begin{figure}[htbp]
@@
 
+To supplement this analysis, we compared the network signatures using
+the centrality dissimilarity approach described above. The dissimilarity
+between samples was a function of shared relationships, degree of
+centrality, and bacteria/phage abundance. When using this supplementary
+approach, we found that network structures significantly clustered by
+moisture, sebaceous, and intermittently moist status \textbf{(Figure
+\ref{skinnetwork} C,E)}. Occluded sites were significantly different
+from exposed and intermittently occluded sites, but there was no
+difference between exposed and intermittently occluded sites
+\textbf{(Figure \ref{skinnetwork} D,F)}. These findings provide further
+support that skin microbiome network structure differs significantly
+between skin sites.
+
+\section{Discussion}\label{discussion}
+
+Foundational work has provided a baseline understanding of the human
+microbiome by characterizing bacterial and viral diversity across the
+human body {[}13,14,43--45,61{]}. Here, we offer an initial
+understanding of how phage-bacteria networks differ throughout the human
+body, so as to provide a baseline for future studies of how and why
+microbiome networks differ in disease states. We developed and
+implemented a network-based analytical model to evaluate the basic
+properties of the human microbiome through bacteria and phage
+relationships, instead of membership or diversity alone. This enabled
+the application of network theory to provide a new perspective on
+complex ecological communities. We utilized metrics of connectivity to
+model the extent to which communities of bacteria and phages interact
+through mechanisms such as horizontal gene transfer, modulated bacterial
+gene expression, and alterations in abundance.
+
+Just as gut microbiome and virome composition and diversity are
+conserved in individuals {[}13,43,44,60{]}, gut and skin microbiome
+network structures were conserved within individuals over time. Gut
+network structure was not conserved among family members. These findings
+suggested that the community properties inferred from microbiome
+interaction network structures, such as stability (meaning a more highly
+connected network is more stable because a randomly removed bacteria or
+phage node is less likely to divide or disintegrate {[}30,58{]} the
+overall network), the potential for horizontal gene transfer between
+members, and co-evolution of populations, were person-specific. These
+properties may be impacted by personal factors ranging from the body's
+immune system to external environmental conditions, such as climate and
+diet.
+
+We observed evidence supporting the ability of environmental conditions
+to shape gut and skin microbiome interaction network structure by
+observing that diet and skin location were associated with altered
+network structures. We found evidence that diet was sufficient to alter
+gut microbiome network connectivity, although this needs to be
+interpreted as a case observation, due to the small sample size.
+Although our sample size was small, our findings provided some
+preliminary evidence that high-fat diets were less connected than
+low-fat diets and that high-fat diets therefore may lead to less stable
+communities with a decreased ability for microbes to directly influence
+one another. We supported this finding with the observation that obesity
+may have been associated with decreased network connectivity. Together
+these findings suggest the food we eat may not only impact which
+microbes colonize our guts, but may also impact their interactions with
+infecting phages. Further work will be required to characterize these
+relationships with a larger cohort.
+
+In addition to diet, the skin environment also influenced the microbiome
+interaction network structure. Network structure differed between
+environmentally exposed and occluded skin sites. The sites under greater
+environmental fluctuation and exposure (the exposed and intermittently
+exposed sites) were less connected and therefore were predicted to have
+a higher propensity for instability. Likewise, intermittently moist
+sites demonstrated less connectedness than the more stable moist and
+sebaceous sites. Together these data suggested that body sites under
+greater degrees of fluctuation harbored less connected, potentially less
+stable microbiomes. This points to a link between microbiome and
+environmental stability and warrants further investigation.
+
+While these findings take us an important step closer to understanding
+the microbiome through interspecies relationships, there are caveats to
+and considerations regarding the approach. First, as with most
+classification models, the infection classification model developed and
+applied is only as good as its training set -- in this case, the
+collection of experimentally-verified positive and negative infection
+data, where genomes of all members are fully sequenced. Large-scale
+experimental screens for phage and bacteria infectious interactions that
+report high-confidence negative interactions (i.e., no infection) are
+desperately needed, as they would provide more robust model training and
+improved model performance. Furthermore, just as we have improved on
+previous modeling efforts, we expect that new and creative scoring
+metrics will be integrated into this model to improve future
+performance.
+
+Second, although our analyses utilized the best datasets currently
+available for our study, this work was done retrospectively and relied
+on existing data up to seven years old. These archived datasets were
+limited by the technology and costs of the time. For example, the diet
+and twin studies, relied on multiple displacement amplification (MDA) in
+their library preparations--an approach used to overcome the large
+nucleic acids requirements typical of older sequencing library
+generation protocols. It is now known that MDA results in biases in
+microbial community composition {[}62{]}, as well as toward ssDNA viral
+genomes {[}63,64{]}, thus rendering the resulting microbial and viral
+metagenomes largely non-quantitative. Future work that employs larger
+sequence datasets and that avoids the use of bias-inducing amplification
+steps will build on and validate our findings, as well as inform the
+design and interpretation of further studies.
+
+Finally, the networks in this study were built using operational genomic
+units (OGUs), which represented groups of highly similar bacteria or
+phage genomes or clustered genome fragments. Similar clustering
+definition and validation methods, both computational and experimental,
+have been implemented in other metagenomic sequencing studies, as well
+{[}42,65--67{]}. These approaches could offer yet another level of
+sophistication to our network-based analyses. While this operationally
+defined clustering approach allows us to study whole community networks,
+our ability to make conclusions about interactions among specific phage
+or bacterial species or populations is inherently limited, compared to
+more focused, culture-based studies such as the work by Malki \emph{et
+al} {[}68{]}. Future work must address this limitation, e.g., through
+improved binning methods and deeper metagenomic shotgun sequencing, but
+most importantly through an improved conceptual framing of what defines
+ecologically and evolutionarily cohesive units for both phage and
+bacteria {[}69{]}. Defining operational genomic units and their
+taxonomic underpinnings (e.g., whether OGU clusters represent genera or
+species) is an active area of work critical to the utility of this
+approach. As a first step, phylogenomic analyses have been performed to
+cluster cyanophage isolate genomes into informative groups using shared
+gene content, average nucleotide identity of shared genes, and pairwise
+differences between genomes {[}70{]}. Such population-genetic assessment
+of phage evolution, coupled with the ecological implications of genome
+heterogeneity, will inform how to define nodes in future iterations of
+the ecological network developed here. Even though we are hesitant to
+speculate on phage host ranges at low taxonomic levels in our dataset,
+the data does aggree with previous reports of instances of broad phage
+host range {[}68,71{]}.
+
+Together our work takes an initial step towards defining bacteria-virus
+interaction profiles as a characteristic of human-associated microbial
+communities. This approach revealed the impacts that different human
+environments (e.g., the skin and gut) can have on microbiome
+connectivity. By focusing on relationships between bacterial and viral
+communities, they are studied as the interacting cohorts they are,
+rather than as independent entities. While our developed bacteria-phage
+interaction framework is a novel conceptual advance, the microbiome also
+consists of archaea and small eukaryotes, including fungi and
+\emph{Demodex} mites {[}1,72{]}--all of which can interact with human
+immune cells and other non-microbial community members {[}73{]}. Future
+work will build from our approach and include these additional community
+members and their diverse interactions and relationships (e.g., beyond
+phage-bacteria). This will result in a more robust network and a more
+holistic understanding of the evolutionary and ecological processes that
+drive the assembly and function of the human-associated microbiome.
+
+\section{Materials \& Methods}\label{materials-methods}
+
+\subsection{Code Availability}\label{code-availability}
+
+A reproducible version of this manuscript written in R markdown and all
+of the code used to obtain and process the sequencing data is available
+at the following GitHub repository:
+
+https://github.com/SchlossLab/Hannigan\_ConjunctisViribus\_ploscompbio\_2017
+
+\subsection{Data Acquisition \& Quality
+Control}\label{data-acquisition-quality-control}
+
+Raw sequencing data and associated metadata were acquired from the NCBI
+sequence read archive (SRA). Supplementary metadata were acquired from
+the same SRA repositories and their associated manuscripts. The gut
+virome diet study (SRA: \texttt{SRP002424}), twin virome studies (SRA:
+\texttt{SRP002523}; \texttt{SRP000319}), and skin virome study (SRA:
+\texttt{SRP049645}) were downloaded as \texttt{.sra} files. Sequencing
+files were converted to \texttt{fastq} format using the
+\texttt{fastq-dump} tool of the NCBI SRA Toolkit (v2.2.0). Sequences
+were quality trimmed using the Fastx toolkit (v0.0.14) to exclude bases
+with quality scores below 33 and shorter than 75 bp {[}74{]}. Paired end
+reads were filtered to exclude sequences missing their corresponding
+pair using the \texttt{get\_trimmed\_pairs.py} script available in the
+source code.
+
+\subsection{Contig Assembly}\label{contig-assembly}
+
+Contigs were assembled using the Megahit assembly program (v1.0.6)
+{[}75{]}. A minimum contig length of 1 kb was used. Iterative k-mer
+stepping began at a minimum length of 21 and progressed by 20 until 101.
+All other default parameters were used.
+
+\subsection{Contig Abundance
+Calculations}\label{contig-abundance-calculations}
+
+Contigs were concatenated into two master files prior to alignment, one
+for bacterial contigs and one for phage contigs. Sample sequences were
+aligned to phage or bacterial contigs using the Bowtie2 global aligner
+(v2.2.1) {[}76{]}. We defined a mismatch threshold of 1 bp and seed
+length of 25 bp. Sequence abundance was calculated from the Bowtie2
+output using the \texttt{calculate\_abundance\_from\_sam.pl} script
+available in the source code.
+
+\subsection{Operational Genomic Unit
+Binning}\label{operational-genomic-unit-binning}
+
+Contigs often represent large fragments of genomes. In order to reduce
+redundancy and the resulting artificially inflated genomic richness
+within our dataset, it was important to bin contigs into operational
+units based on their similarity. This approach is conceptually similar
+to the clustering of related 16S rRNA sequences into operational
+taxonomic units (OTUs), although here we are clustering contigs into
+operational genomic units (OGUs) {[}61{]}.
+
+Contigs were clustered using the CONCOCT algorithm (v0.4.0) {[}77{]}.
+Because of our large dataset and limits in computational efficiency, we
+randomly subsampled the dataset to include 25\% of all samples, and used
+these to inform contig abundance within the CONCOCT algorithm. CONCOCT
+was used with a maximum of 500 clusters, a k-mer length of four, a
+length threshold of 1 kb, 25 iterations, and exclusion of the total
+coverage variable.
+
+OGU abundance (\(A_{O}\)) was obtained as the sum of the abundance of
+each contig (\(A_{j}\)) associated with that OGU. The abundance values
+were length corrected such that:
+
+\[ { A }_{ O }=\frac { { 10 }^{ 7 }\sum _{ j=1 }^{ k }{ { A }_{ j } }  }{ \sum _{ j=1 }^{ k }{ { L }_{ j } }  } \]
+
+Where \texttt{L} is the length of each contig \texttt{j} within the OGU.
+
+\subsection{Operational Genomic Unit
+Identification}\label{operational-genomic-unit-identification}
+
+To confirm a lack of phage sequences in the bacterial OGU dataset, we
+performed blast nucleotide alignment of the bacterial OGU representative
+sequences using an e-value \textless{} \(10^{-25}\), which was stricter
+than the \(10^{-10}\) threshold used in the random forest model below,
+against all of the phage reference genomes available in the EMBL
+database. We used a stricter threshold because we know there are genomic
+similarities between bacteria and phage OGUs from the interactive model,
+but we were interested in contigs with high enough similarity to
+references that they may indeed be from phages. We also performed the
+converse analysis of aligning phage OGU representative sequences to EMBL
+bacterial reference genomes. Finally, we ran both the phage and bacteria
+OGU representative sequences through the Virsorter program (1.0.3) to
+identify phages (all default parameters were used), using only those in
+the high confidence identification category ``class 1'' {[}47{]}.
+
+\subsection{Open Reading Frame
+Prediction}\label{open-reading-frame-prediction}
+
+Open reading frames (ORFs) were identified using the Prodigal program
+(V2.6.2) with the meta mode parameter and default settings {[}78{]}.
+
+\subsection{Classification Model Creation and
+Validation}\label{classification-model-creation-and-validation}
+
+The classification model for predicting interactions was built using
+experimentally validated bacteria-phage infections or validated lack of
+infections from six studies {[}41,48--52{]}. Associated reference
+genomes were downloaded from the European Bioinformatics Institute (see
+details in source code). The model was created based on the four metrics
+listed below.
+
+The four scores were used as parameters in a random forest model to
+classify bacteria and bacteriophage pairs as either having infectious
+interactions or not. The classification model was built using the Caret
+R package (v6.0.73) {[}79{]}. The model was trained using five-fold
+cross validation with ten repeats, and the median model performance was
+evaluated by training the model on 80\% of the dataset and testing
+performance on the remaining 20\%. Pairs without scores were classified
+as not interacting. The model was optimized using the ROC value. The
+resulting model performance was plotted using the plotROC R package.
+
+\subparagraph{Identify Bacterial CRISPRs Targeting
+Phages}\label{identify-bacterial-crisprs-targeting-phages}
+
+Clustered Regularly Interspaced Short Palindromic Repeats (CRISPRs) were
+identified from bacterial genomes using the PilerCR program (v1.06)
+{[}80{]}. Resulting spacer sequences were filtered to exclude spacers
+shorter than 20 bp and longer than 65 bp. Spacer sequences were aligned
+to the phage genomes using the nucleotide BLAST algorithm with default
+parameters (v2.4.0) {[}81{]}. The mean percent identity for each
+matching pair was recorded for use in our classification model.
+
+\subparagraph{Detect Matching Prophages within Bacterial
+Genomes}\label{detect-matching-prophages-within-bacterial-genomes}
+
+Temperate bacteriophages infect and integrate into their bacterial
+host's genome. We detected integrated phage elements within bacterial
+genomes by aligning phage genomes to bacterial genomes using the
+nucleotide BLAST algorithm and a minimum e-value of 1e-10. The resulting
+bitscore of each alignment was recorded for use in our classification
+model.
+
+\subparagraph{Identify Shared Genes Between Bacteria and
+Phages}\label{identify-shared-genes-between-bacteria-and-phages}
+
+As a result of gene transfer or phage genome integration during
+infection, phages may share genes with their bacterial hosts, providing
+us with evidence of phage-host pairing. We identified shared genes
+between bacterial and phage genomes by assessing amino acid similarity
+between the genes using the Diamond protein alignment algorithm
+(v0.7.11.60) {[}82{]}. The mean alignment bitscores for each genome pair
+were recorded for use in our classification model.
+
+\subparagraph{Protein - Protein
+Interactions}\label{protein---protein-interactions}
+
+The final method used for predicting infectious interactions between
+bacteria and phages was the detection of pairs of genes whose proteins
+are known to interact. We assigned bacterial and phage genes to protein
+families by aligning them to the Pfam database using the Diamond protein
+alignment algorithm. We then identified which pairs of proteins were
+predicted to interact using the Pfam interaction information within the
+Intact database {[}53{]}. The mean bitscores of the matches between each
+pair were recorded for use in the classification model.
+
+\subsection{Interaction Network
+Construction}\label{interaction-network-construction}
+
+The bacteria and phage operational genomic units (OGUs) were scored
+using the same approach as outlined above. The infectious pairings
+between bacteria and phage OGUs were classified using the random forest
+model described above. The predicted infectious pairings and all
+associated metadata were used to populate a graph database using Neo4j
+graph database software (v2.3.1) {[}83{]}. This network was used for
+downstream community analysis.
+
+\subsection{Centrality Analysis}\label{centrality-analysis}
+
+We quantified the centrality of graph vertices using three different
+metrics, each of which provided different information graph structure.
+When calculating these values, let \(G(V,E)\) be an undirected,
+unweighted graph with \(|V|=n\) nodes and \(|E|=m\) edges. Also, let
+\(\mathbf{A}\) be its corresponding adjacency matrix with entries
+\(a_{ij} = 1\) if nodes \(V_i\) and \(V_j\) are connected via an edge,
+and \(a_{ij} = 0\) otherwise.
+
+Briefly, the \textbf{closeness centrality} of node \(V_i\) is calculated
+taking the inverse of the average length of the shortest paths
+(\texttt{d}) between nodes \(V_i\) and all the other nodes \(V_j\).
+Mathematically, the closeness centrality of node \(V_i\) is given as:
+
+\[ { C }_{ C }\left( { V }_{ i } \right) ={ \left( \sum _{ j=1 }^{ n }{ d\left( { V }_{ i },{ V }_{ j } \right)  }  \right)  }^{ -1 } \]
+
+The distance between nodes (\texttt{d}) was calculated as the shortest
+number of edges required to be traversed to move from one node to
+another.
+
+Intuitively, the \textbf{degree centrality} of node \(V_i\) is defined
+as the number of edges that are incident to that node:
+
+\[ { C }_{ D }\left( { V }_{ i } \right) =\sum _{ j=1 }^{ n }{ { a }_{ ij } } \]
+
+where \(a_{ij}\) is the \(ij^{th}\) entry in the adjacency matrix
+\(\mathbf{A}\).
+
+The eigenvector centrality of node \(V_i\) is defined as the \(i^{th}\)
+value in the first eigenvector of the associated adjacency matrix
+\(\mathbf{A}\). Conceptually, this function results in a centrality
+value that reflects the connections of the vertex, as well as the
+centrality of its neighboring vertices.
+
+The \textbf{centralization} metric was used to assess the average
+centrality of each sample graph \(\mathbf{G}\). Centralization was
+calculated by taking the sum of each vertex \(V_{i}\)'s centrality from
+the graph maximum centrality \(C_{w}\), such that:
+
+\[ C\left( G \right) =\frac { \sum _{ i=1 }^{ n }{ Cw -c\left( { V }_{ i } \right)  }  }{ { T } }  \]
+
+The values were corrected for uneven graph sizes by dividing the
+centralization score by the maximum theoretical centralization
+(\texttt{T}) for a graph with the same number of vertices.
+
+Degree and closeness centrality were calculated using the associated
+functions within the igraph R package (v1.0.1) {[}84{]}.
+
+\subsection{Network Relationship
+Dissimilarity}\label{network-relationship-dissimilarity}
+
+We assessed similarity between graphs by evaluating the shared
+centrality of their vertices, as has been done previously. More
+specifically, we calculated the dissimilarity between graphs \(G_{i}\)
+and \(G_{j}\) using the Bray-Curtis dissimilarity metric and eigenvector
+centrality values such that:
+
+\[ B\left( { G }_{ i },{ G }_{ j } \right) =1-\frac { 2{ C }_{ ij } }{ { C }_{ i }+{ C }_{ j } } \]
+
+Where \(C_{ij}\) is the sum of the lesser centrality values for those
+vertices shared between graphs, and \(C_{i}\) and \(C_{j}\) are the
+total number of vertices found in each graph. This allows us to
+calculate the dissimilarity between graphs based on the shared
+centrality values between the two graphs.
+
+\subsection{Statistics and
+Comparisons}\label{statistics-and-comparisons}
+
+Differences in intrapersonal and interpersonal network structure
+diversity, based on multivariate data, were calculated using an analysis
+of similarity (ANOSIM). Statistical significance of univariate
+Eigenvector centrality differences were calculated using a paired
+Wilcoxon test.
+
+Statistical significance of differences in univariate eigenvector
+centrality measurements of skin virome-microbiome networks were
+calculated using a pairwise Wilcoxon test, corrected for multiple
+hypothesis tests using the Holm correction method. Multivariate
+eigenvector centrality was measured as the mean differences between
+cluster centroids, with statistical significance measured using an ANOVA
+and post hoc Tukey test.
+
+\section{Acknowledgments}\label{acknowledgments}
+
+We thank the members of the Schloss lab for their underlying
+contributions. We thank the authors of the original studies for making
+their data and metadata publicly available and understandable. We also
+thank the participants in the studies.
+
+\section{Author Contributions}\label{author-contributions}
+
+\emph{Conceptualization}: GDH, MBD, DK, PDS. \emph{Data Curation}: GDH.
+\emph{Formal Analysis}: GDH. \emph{Funding Acquisition}: GDH, PDS.
+\emph{Writing -- Original Draft Preparation}: GDH, PDS. \emph{Writing --
+Review \& Editing}: GDH, MBD, DK, PDS.
+
+\section{Funding Information}\label{funding-information}
+
+GDH was supported in part by the Molecular Mechanisms in Microbial
+Pathogenesis Training Program (T32 AI007528). GDH and PDS were supported
+in part by funding from the NIH (P30DK034933, U19AI09087, and
+U01AI124255).
+
+\section{Competing interests}\label{competing-interests}
+
+The authors report no conflicts of interest.
+
 \newpage
 
-\section{Supplemental Figure Legends}\label{supplemental-figure-legends}
+\section{Supplemental Figure
+Captions}\label{supplemental-figure-captions}
 
@@
 
-\section{Supplemental Table Legend}\label{supplemental-table-legend}
+\section{Supplemental Table Captions}\label{supplemental-table-captions}
 
@@
 \hypertarget{ref-Hannigan:2013im}{}
-1. \textbf{Hannigan GD}, \textbf{Grice EA}. 2013. Microbial Ecology of
-the Skin in the Era of Metagenomics and Molecular Microbiology. Cold
-Spring Harbor Perspectives in Medicine \textbf{3}:a015362--a015362.
+1. Hannigan GD, Grice EA. Microbial Ecology of the Skin in the Era of
+Metagenomics and Molecular Microbiology. Cold Spring Harbor Perspectives
+in Medicine. 2013;3: a015362--a015362.
 
 \hypertarget{ref-Hannigan:2014be}{}
-2. \textbf{Hannigan GD}, \textbf{Hodkinson BP}, \textbf{McGinnis K},
-\textbf{Tyldsley AS}, \textbf{Anari JB}, \textbf{Horan AD},
-\textbf{Grice EA}, \textbf{Mehta S}. 2014. Culture-independent pilot
-study of microbiota colonizing open fractures and association with
-severity, mechanism, location, and complication from presentation to
-early outpatient follow-up. Journal of Orthopaedic Research
-\textbf{32}:597--605.
+2. Hannigan GD, Hodkinson BP, McGinnis K, Tyldsley AS, Anari JB, Horan
+AD, et al. Culture-independent pilot study of microbiota colonizing open
+fractures and association with severity, mechanism, location, and
+complication from presentation to early outpatient follow-up. Journal of
+Orthopaedic Research. 2014;32: 597--605.
 
 \hypertarget{ref-Loesche:2016ev}{}
-3. \textbf{Loesche M}, \textbf{Gardner SE}, \textbf{Kalan L},
-\textbf{Horwinski J}, \textbf{Zheng Q}, \textbf{Hodkinson BP},
-\textbf{Tyldsley AS}, \textbf{Franciscus CL}, \textbf{Hillis SL},
-\textbf{Mehta S}, \textbf{Margolis DJ}, \textbf{Grice EA}. 2016.
-Temporal stability in chronic wound microbiota is associated with poor
-healing. Journal of Investigative Dermatology.
+3. Loesche M, Gardner SE, Kalan L, Horwinski J, Zheng Q, Hodkinson BP,
+et al. Temporal stability in chronic wound microbiota is associated with
+poor healing. Journal of Investigative Dermatology. 2016;
 
 \hypertarget{ref-He:2016ch}{}
-4. \textbf{He Q}, \textbf{Li X}, \textbf{Liu C}, \textbf{Su L},
-\textbf{Xia Z}, \textbf{Li X}, \textbf{Li Y}, \textbf{Li L}, \textbf{Yan
-T}, \textbf{Feng Q}, \textbf{Xiao L}. 2016. Dysbiosis of the fecal
+4. He Q, Li X, Liu C, Su L, Xia Z, Li X, et al. Dysbiosis of the fecal
 microbiota in the TNBS-induced Crohn's disease mouse model. Applied
-Microbiology and Biotechnology 1--10.
+Microbiology and Biotechnology. 2016; 1--10.
 
 \hypertarget{ref-Norman:2015kb}{}
-5. \textbf{Norman JM}, \textbf{Handley SA}, \textbf{Baldridge MT},
-\textbf{Droit L}, \textbf{Liu CY}, \textbf{Keller BC}, \textbf{Kambal
-A}, \textbf{Monaco CL}, \textbf{Zhao G}, \textbf{Fleshner P},
-\textbf{Stappenbeck TS}, \textbf{McGovern DPB}, \textbf{Keshavarzian A},
-\textbf{Mutlu EA}, \textbf{Sauk J}, \textbf{Gevers D}, \textbf{Xavier
-RJ}, \textbf{Wang D}, \textbf{Parkes M}, \textbf{Virgin HW}. 2015.
-Disease-specific alterations in the enteric virome in inflammatory bowel
-disease. Cell \textbf{160}:447--460.
+5. Norman JM, Handley SA, Baldridge MT, Droit L, Liu CY, Keller BC, et
+al. Disease-specific alterations in the enteric virome in inflammatory
+bowel disease. Cell. 2015;160: 447--460.
 
 \hypertarget{ref-Seekatz:2016fz}{}
-6. \textbf{Seekatz AM}, \textbf{Rao K}, \textbf{Santhosh K},
-\textbf{Young VB}. 2016. Dynamics of the fecal microbiome in patients
-with recurrent and nonrecurrent Clostridium difficile infection. Genome
-medicine \textbf{8}:47.
+6. Seekatz AM, Rao K, Santhosh K, Young VB. Dynamics of the fecal
+microbiome in patients with recurrent and nonrecurrent Clostridium
+difficile infection. Genome medicine. 2016;8: 47.
 
 \hypertarget{ref-Zackular:2014fba}{}
-7. \textbf{Zackular JP}, \textbf{Rogers MAM}, \textbf{Ruffin MT},
-\textbf{Schloss PD}. 2014. The human gut microbiome as a screening tool
-for colorectal cancer. Cancer prevention research (Philadelphia, Pa)
-\textbf{7}:1112--1121.
+7. Zackular JP, Rogers MAM, Ruffin MT, Schloss PD. The human gut
+microbiome as a screening tool for colorectal cancer. Cancer prevention
+research (Philadelphia, Pa). 2014;7: 1112--1121.
 
 \hypertarget{ref-Baxter:2014hb}{}
-8. \textbf{Baxter NT}, \textbf{Zackular JP}, \textbf{Chen GY},
-\textbf{Schloss PD}. 2014. Structure of the gut microbiome following
-colonization with human feces determines colonic tumor burden.
-Microbiome \textbf{2}:20.
+8. Baxter NT, Zackular JP, Chen GY, Schloss PD. Structure of the gut
+microbiome following colonization with human feces determines colonic
+tumor burden. Microbiome. 2014;2: 20.
 
 \hypertarget{ref-Manrique:2016dx}{}
-9. \textbf{Manrique P}, \textbf{Bolduc B}, \textbf{Walk ST},
-\textbf{Oost J van der}, \textbf{Vos WM de}, \textbf{Young MJ}. 2016.
+9. Manrique P, Bolduc B, Walk ST, Oost J van der, Vos WM de, Young MJ.
 Healthy human gut phageome. Proceedings of the National Academy of
-Sciences of the United States of America 201601060.
+Sciences of the United States of America. 2016; 201601060.
 
 \hypertarget{ref-Ly:2014ew}{}
-10. \textbf{Ly M}, \textbf{Abeles SR}, \textbf{Boehm TK},
-\textbf{Robles-Sikisaka R}, \textbf{Naidu M}, \textbf{Santiago-Rodriguez
-T}, \textbf{Pride DT}. 2014. Altered Oral Viral Ecology in Association
-with Periodontal Disease. mBio \textbf{5}:e01133--14--e01133--14.
+10. Ly M, Abeles SR, Boehm TK, Robles-Sikisaka R, Naidu M,
+Santiago-Rodriguez T, et al. Altered Oral Viral Ecology in Association
+with Periodontal Disease. mBio. 2014;5: e01133--14--e01133--14.
 
 \hypertarget{ref-Modi:2013fia}{}
-11. \textbf{Modi SR}, \textbf{Lee HH}, \textbf{Spina CS},
-\textbf{Collins JJ}. 2013. Antibiotic treatment expands the resistance
-reservoir and ecological network of the phage metagenome. Nature
-\textbf{499}:219--222.
+11. Modi SR, Lee HH, Spina CS, Collins JJ. Antibiotic treatment expands
+the resistance reservoir and ecological network of the phage metagenome.
+Nature. 2013;499: 219--222.
 
 \hypertarget{ref-Monaco:2016ita}{}
-12. \textbf{Monaco CL}, \textbf{Gootenberg DB}, \textbf{Zhao G},
-\textbf{Handley SA}, \textbf{Ghebremichael MS}, \textbf{Lim ES},
-\textbf{Lankowski A}, \textbf{Baldridge MT}, \textbf{Wilen CB},
-\textbf{Flagg M}, \textbf{Norman JM}, \textbf{Keller BC},
-\textbf{Luévano JM}, \textbf{Wang D}, \textbf{Boum Y}, \textbf{Martin
-JN}, \textbf{Hunt PW}, \textbf{Bangsberg DR}, \textbf{Siedner MJ},
-\textbf{Kwon DS}, \textbf{Virgin HW}. 2016. Altered Virome and Bacterial
-Microbiome in Human Immunodeficiency Virus-Associated Acquired
-Immunodeficiency Syndrome. Cell Host and Microbe \textbf{19}:311--322.
+12. Monaco CL, Gootenberg DB, Zhao G, Handley SA, Ghebremichael MS, Lim
+ES, et al. Altered Virome and Bacterial Microbiome in Human
+Immunodeficiency Virus-Associated Acquired Immunodeficiency Syndrome.
+Cell Host and Microbe. 2016;19: 311--322.
 
 \hypertarget{ref-Hannigan:2015fz}{}
-13. \textbf{Hannigan GD}, \textbf{Meisel JS}, \textbf{Tyldsley AS},
-\textbf{Zheng Q}, \textbf{Hodkinson BP}, \textbf{SanMiguel AJ},
-\textbf{Minot S}, \textbf{Bushman FD}, \textbf{Grice EA}. 2015. The
-Human Skin Double-Stranded DNA Virome: Topographical and Temporal
-Diversity, Genetic Enrichment, and Dynamic Associations with the Host
-Microbiome. mBio \textbf{6}:e01578--15.
+13. Hannigan GD, Meisel JS, Tyldsley AS, Zheng Q, Hodkinson BP,
+SanMiguel AJ, et al. The Human Skin Double-Stranded DNA Virome:
+Topographical and Temporal Diversity, Genetic Enrichment, and Dynamic
+Associations with the Host Microbiome. mBio. 2015;6: e01578--15.
 
 \hypertarget{ref-Minot:2011ez}{}
-14. \textbf{Minot S}, \textbf{Sinha R}, \textbf{Chen J}, \textbf{Li H},
-\textbf{Keilbaugh SA}, \textbf{Wu GD}, \textbf{Lewis JD},
-\textbf{Bushman FD}. 2011. The human gut virome: Inter-individual
-variation and dynamic response to diet. Genome Research
-\textbf{21}:1616--1625.
+14. Minot S, Sinha R, Chen J, Li H, Keilbaugh SA, Wu GD, et al. The
+human gut virome: Inter-individual variation and dynamic response to
+diet. Genome Research. 2011;21: 1616--1625.
 
 \hypertarget{ref-SantiagoRodriguez:2015gd}{}
-15. \textbf{Santiago-Rodriguez TM}, \textbf{Ly M}, \textbf{Bonilla N},
-\textbf{Pride DT}. 2015. The human urine virome in association with
-urinary tract infections. Frontiers in Microbiology \textbf{6}:14.
+15. Santiago-Rodriguez TM, Ly M, Bonilla N, Pride DT. The human urine
+virome in association with urinary tract infections. Frontiers in
+Microbiology. 2015;6: 14.
 
 \hypertarget{ref-Abeles:2015dy}{}
-16. \textbf{Abeles SR}, \textbf{Ly M}, \textbf{Santiago-Rodriguez TM},
-\textbf{Pride DT}. 2015. Effects of Long Term Antibiotic Therapy on
-Human Oral and Fecal Viromes. PLOS ONE \textbf{10}:e0134941.
+16. Abeles SR, Ly M, Santiago-Rodriguez TM, Pride DT. Effects of Long
+Term Antibiotic Therapy on Human Oral and Fecal Viromes. PLOS ONE.
+2015;10: e0134941.
 
 \hypertarget{ref-Abeles:2014kj}{}
-17. \textbf{Abeles SR}, \textbf{Robles-Sikisaka R}, \textbf{Ly M},
-\textbf{Lum AG}, \textbf{Salzman J}, \textbf{Boehm TK}, \textbf{Pride
-DT}. 2014. Human oral viruses are personal, persistent and
-gender-consistent 1--15.
+17. Abeles SR, Robles-Sikisaka R, Ly M, Lum AG, Salzman J, Boehm TK, et
+al. Human oral viruses are personal, persistent and gender-consistent.
+2014; 1--15.
 
 \hypertarget{ref-Haerter:2014ii}{}
-18. \textbf{Haerter JO}, \textbf{Mitarai N}, \textbf{Sneppen K}. 2014.
-Phage and bacteria support mutual diversity in a narrowing staircase of
-coexistence. The ISME Journal \textbf{8}:2317--2326.
+18. Haerter JO, Mitarai N, Sneppen K. Phage and bacteria support mutual
+diversity in a narrowing staircase of coexistence. The ISME Journal.
+2014;8: 2317--2326.
 
 \hypertarget{ref-Lindell:2005gz}{}
-19. \textbf{Lindell D}, \textbf{Jaffe JD}, \textbf{Johnson ZI},
-\textbf{Church GM}, \textbf{Chisholm SW}. 2005. Photosynthesis genes in
-marine viruses yield proteins during host infection. Nature
-\textbf{438}:86--89.
+19. Lindell D, Jaffe JD, Johnson ZI, Church GM, Chisholm SW.
+Photosynthesis genes in marine viruses yield proteins during host
+infection. Nature. 2005;438: 86--89.
 
 \hypertarget{ref-Tyler:2013fl}{}
-20. \textbf{Tyler JS}, \textbf{Beeri K}, \textbf{Reynolds JL},
-\textbf{Alteri CJ}, \textbf{Skinner KG}, \textbf{Friedman JH},
-\textbf{Eaton KA}, \textbf{Friedman DI}. 2013. Prophage induction is
-enhanced and required for renal disease and lethality in an EHEC mouse
-model. PLoS Pathogens \textbf{9}:e1003236.
+20. Tyler JS, Beeri K, Reynolds JL, Alteri CJ, Skinner KG, Friedman JH,
+et al. Prophage induction is enhanced and required for renal disease and
+lethality in an EHEC mouse model. PLoS Pathogens. 2013;9: e1003236.
 
 \hypertarget{ref-Hargreaves:2014ja}{}
-21. \textbf{Hargreaves KR}, \textbf{Kropinski AM}, \textbf{Clokie MR}.
-2014. Bacteriophage behavioral ecology: How phages alter their bacterial
-host's habits. Bacteriophage \textbf{4}:e29866.
+21. Hargreaves KR, Kropinski AM, Clokie MR. Bacteriophage behavioral
+ecology: How phages alter their bacterial host's habits. Bacteriophage.
+2014;4: e29866.
 
 \hypertarget{ref-Moon:2015fa}{}
-22. \textbf{Moon BY}, \textbf{Park JY}, \textbf{Hwang SY},
-\textbf{Robinson DA}, \textbf{Thomas JC}, \textbf{Fitzgerald JR},
-\textbf{Park YH}, \textbf{Seo KS}. 2015. Phage-mediated horizontal
-transfer of a Staphylococcus aureus virulence-associated genomic island.
-Scientific Reports \textbf{5}:9784.
+22. Moon BY, Park JY, Hwang SY, Robinson DA, Thomas JC, Fitzgerald JR,
+et al. Phage-mediated horizontal transfer of a Staphylococcus aureus
+virulence-associated genomic island. Scientific Reports. 2015;5: 9784.
 
 \hypertarget{ref-Modi:2013fi}{}
-23. \textbf{Modi SR}, \textbf{Lee HH}, \textbf{Spina CS},
-\textbf{Collins JJ}. 2013. Antibiotic treatment expands the resistance
-reservoir and ecological network of the phage metagenome. Nature
-\textbf{499}:219--222.
+23. Modi SR, Lee HH, Spina CS, Collins JJ. Antibiotic treatment expands
+the resistance reservoir and ecological network of the phage metagenome.
+Nature. 2013;499: 219--222.
 
 \hypertarget{ref-Ogg:1981th}{}
-24. \textbf{Ogg JE}, \textbf{Timme TL}, \textbf{Alemohammad MM}. 1981.
-General Transduction in Vibrio cholerae. Infection and Immunity
-\textbf{31}:737--741.
+24. Ogg JE, Timme TL, Alemohammad MM. General Transduction in Vibrio
+cholerae. Infection and Immunity. 1981;31: 737--741.
 
 \hypertarget{ref-Frost:2005dn}{}
-25. \textbf{Frost LS}, \textbf{Leplae R}, \textbf{Summers AO},
-\textbf{Toussaint A}. 2005. Mobile genetic elements: the agents of open
-source evolution. Nature Reviews Microbiology \textbf{3}:722--732.
+25. Frost LS, Leplae R, Summers AO, Toussaint A. Mobile genetic
+elements: the agents of open source evolution. Nature Reviews
+Microbiology. 2005;3: 722--732.
 
 \hypertarget{ref-Koskella:2014ds}{}
-26. \textbf{Koskella B}, \textbf{Brockhurst MA}. 2014. Bacteria-phage
-coevolution as a driver of ecological and evolutionary processes in
-microbial communities. FEMS Microbiology Reviews \textbf{38}:916--931.
+26. Koskella B, Brockhurst MA. Bacteria-phage coevolution as a driver of
+ecological and evolutionary processes in microbial communities. FEMS
+Microbiology Reviews. 2014;38: 916--931.
 
 \hypertarget{ref-Jover:2014gq}{}
-27. \textbf{Jover LF}, \textbf{Effler TC}, \textbf{Buchan A},
-\textbf{Wilhelm SW}, \textbf{Weitz JS}. 2014. The elemental composition
-of virus particles: implications for marine biogeochemical cycles.
-Nature Reviews Microbiology \textbf{12}:519--528.
+27. Jover LF, Effler TC, Buchan A, Wilhelm SW, Weitz JS. The elemental
+composition of virus particles: implications for marine biogeochemical
+cycles. Nature Reviews Microbiology. 2014;12: 519--528.
 
 \hypertarget{ref-Harcombe:2005fd}{}
-28. \textbf{Harcombe WR}, \textbf{Bull JJ}. 2005. Impact of phages on
-two-species bacterial communities. Applied and Environmental
-Microbiology \textbf{71}:5254--5259.
+28. Harcombe WR, Bull JJ. Impact of phages on two-species bacterial
+communities. Applied and Environmental Microbiology. 2005;71:
+5254--5259.
 
 \hypertarget{ref-Middelboe:2001fl}{}
-29. \textbf{Middelboe M}, \textbf{Hagström A}, \textbf{Blackburn N},
-\textbf{Sinn B}, \textbf{Fischer U}, \textbf{Borch NH}, \textbf{Pinhassi
-J}, \textbf{Simu K}, \textbf{Lorenz MG}. 2001. Effects of Bacteriophages
-on the Population Dynamics of Four Strains of Pelagic Marine Bacteria.
-Microbial Ecology \textbf{42}:395--406.
+29. Middelboe M, Hagström A, Blackburn N, Sinn B, Fischer U, Borch NH,
+et al. Effects of Bacteriophages on the Population Dynamics of Four
+Strains of Pelagic Marine Bacteria. Microbial Ecology. 2001;42:
+395--406.
 
 \hypertarget{ref-Poisot:2011jc}{}
-30. \textbf{Poisot T}, \textbf{Lepennetier G}, \textbf{Martinez E},
-\textbf{Ramsayer J}, \textbf{Hochberg ME}. 2011. Resource availability
-affects the structure of a natural bacteriabacteriophage community.
-Biology letters \textbf{7}:201--204.
+30. Poisot T, Lepennetier G, Martinez E, Ramsayer J, Hochberg ME.
+Resource availability affects the structure of a natural
+bacteriabacteriophage community. Biology letters. 2011;7: 201--204.
 
 \hypertarget{ref-Thompson:2012ki}{}
-31. \textbf{Thompson RM}, \textbf{Brose U}, \textbf{Dunne JA},
-\textbf{Hall RO}, \textbf{Hladyz S}, \textbf{Kitching RL},
-\textbf{Martinez ND}, \textbf{Rantala H}, \textbf{Romanuk TN},
-\textbf{Stouffer DB}, \textbf{Tylianakis JM}. 2012. Food webs:
-reconciling the structure and function of biodiversity. Trends in
-ecology \& evolution \textbf{27}:689--697.
+31. Thompson RM, Brose U, Dunne JA, Hall RO, Hladyz S, Kitching RL, et
+al. Food webs: reconciling the structure and function of biodiversity.
+Trends in ecology \& evolution. 2012;27: 689--697.
 
 \hypertarget{ref-Moebus:1981kp}{}
-32. \textbf{Moebus K}, \textbf{Nattkemper H}. 1981. Bacteriophage
-sensitivity patterns among bacteria isolated from marine waters.
-Helgoländer Meeresuntersuchungen \textbf{34}:375--385.
+32. Moebus K, Nattkemper H. Bacteriophage sensitivity patterns among
+bacteria isolated from marine waters. Helgoländer Meeresuntersuchungen.
+1981;34: 375--385.
 
 \hypertarget{ref-Flores:2013hc}{}
-33. \textbf{Flores CO}, \textbf{Valverde S}, \textbf{Weitz JS}. 2013.
-Multi-scale structure and geographic drivers of cross-infection within
-marine bacteria and phages. The ISME Journal \textbf{7}:520--532.
+33. Flores CO, Valverde S, Weitz JS. Multi-scale structure and
+geographic drivers of cross-infection within marine bacteria and phages.
+The ISME Journal. 2013;7: 520--532.
 
 \hypertarget{ref-Poisot:2012fh}{}
-34. \textbf{Poisot T}, \textbf{Canard E}, \textbf{Mouillot D},
-\textbf{Mouquet N}, \textbf{Gravel D}. 2012. The dissimilarity of
-species interaction networks. Ecology letters \textbf{15}:1353--1361.
+34. Poisot T, Canard E, Mouillot D, Mouquet N, Gravel D. The
+dissimilarity of species interaction networks. Ecology letters. 2012;15:
+1353--1361.
 
 \hypertarget{ref-Poisot071993}{}
-35. \textbf{Poisot T}, \textbf{Stouffer D}. 2016. How ecological
-networks evolve. bioRxiv.
+35. Poisot T, Stouffer D. How ecological networks evolve. bioRxiv. 2016;
 
 \hypertarget{ref-Flores:2011bh}{}
-36. \textbf{Flores CO}, \textbf{Meyer JR}, \textbf{Valverde S},
-\textbf{Farr L}, \textbf{Weitz JS}. 2011. Statistical structure of
-host-phage interactions. Proceedings of the National Academy of Sciences
-of the United States of America \textbf{108}:E288--97.
+36. Flores CO, Meyer JR, Valverde S, Farr L, Weitz JS. Statistical
+structure of host-phage interactions. Proceedings of the National
+Academy of Sciences of the United States of America. 2011;108: E288--97.
 
 \hypertarget{ref-Jover:2015ev}{}
-37. \textbf{Jover LF}, \textbf{Flores CO}, \textbf{Cortez MH},
-\textbf{Weitz JS}. 2015. Multiple regimes of robust patterns between
-network structure and biodiversity. Scientific Reports \textbf{5}:17856.
+37. Jover LF, Flores CO, Cortez MH, Weitz JS. Multiple regimes of robust
+patterns between network structure and biodiversity. Scientific Reports.
+2015;5: 17856.
 
 \hypertarget{ref-Reyes:2010cwa}{}
-38. \textbf{Reyes A}, \textbf{Haynes M}, \textbf{Hanson N},
-\textbf{Angly FE}, \textbf{Heath AC}, \textbf{Rohwer F}, \textbf{Gordon
-JI}. 2010. Viruses in the faecal microbiota of monozygotic twins and
-their mothers. Nature \textbf{466}:334--338.
+38. Reyes A, Haynes M, Hanson N, Angly FE, Heath AC, Rohwer F, et al.
+Viruses in the faecal microbiota of monozygotic twins and their mothers.
+Nature. 2010;466: 334--338.
 
 \hypertarget{ref-Turnbaugh:2009ei}{}
-39. \textbf{Turnbaugh PJ}, \textbf{Hamady M}, \textbf{Yatsunenko T},
-\textbf{Cantarel BL}, \textbf{Duncan A}, \textbf{Ley RE}, \textbf{Sogin
-ML}, \textbf{Jones WJ}, \textbf{Roe BA}, \textbf{Affourtit JP},
-\textbf{Egholm M}, \textbf{Henrissat B}, \textbf{Heath AC},
-\textbf{Knight R}, \textbf{Gordon JI}. 2009. A core gut microbiome in
-obese and lean twins. Nature \textbf{457}:480--484.
+39. Turnbaugh PJ, Hamady M, Yatsunenko T, Cantarel BL, Duncan A, Ley RE,
+et al. A core gut microbiome in obese and lean twins. Nature. 2009;457:
+480--484.
 
 \hypertarget{ref-LimaMendez:2015hw}{}
-40. \textbf{Lima-Mendez G}, \textbf{Faust K}, \textbf{Henry N},
-\textbf{Decelle J}, \textbf{Colin S}, \textbf{Carcillo F},
-\textbf{Chaffron S}, \textbf{Ignacio-Espinosa JC}, \textbf{Roux S},
-\textbf{Vincent F}, \textbf{Bittner L}, \textbf{Darzi Y}, \textbf{Wang
-J}, \textbf{Audic S}, \textbf{Berline L}, \textbf{Bontempi G},
-\textbf{Cabello AM}, \textbf{Coppola L}, \textbf{Cornejo-Castillo FM},
-\textbf{d'Ovidio F}, \textbf{De Meester L}, \textbf{Ferrera I},
-\textbf{Garet-Delmas M-J}, \textbf{Guidi L}, \textbf{Lara E},
-\textbf{Pesant S}, \textbf{Royo-Llonch M}, \textbf{Salazar G},
-\textbf{Sánchez P}, \textbf{Sebastian M}, \textbf{Souffreau C},
-\textbf{Dimier C}, \textbf{Picheral M}, \textbf{Searson S},
-\textbf{Kandels-Lewis S}, \textbf{Tara Oceans Coordinators},
-\textbf{Gorsky G}, \textbf{Not F}, \textbf{Ogata H}, \textbf{Speich S},
-\textbf{Stemmann L}, \textbf{Weissenbach J}, \textbf{Wincker P},
-\textbf{Acinas SG}, \textbf{Sunagawa S}, \textbf{Bork P},
-\textbf{Sullivan MB}, \textbf{Karsenti E}, \textbf{Bowler C},
-\textbf{Vargas C de}, \textbf{Raes J}. 2015. Ocean plankton.
-Determinants of community structure in the global plankton interactome.
-Science \textbf{348}:1262073--1262073.
+40. Lima-Mendez G, Faust K, Henry N, Decelle J, Colin S, Carcillo F, et
+al. Ocean plankton. Determinants of community structure in the global
+plankton interactome. Science. 2015;348: 1262073--1262073.
 
 \hypertarget{ref-Edwards:2015iz}{}
-41. \textbf{Edwards RA}, \textbf{McNair K}, \textbf{Faust K},
-\textbf{Raes J}, \textbf{Dutilh BE}. 2015. Computational approaches to
-predict bacteriophage-host relationships. FEMS Microbiology Reviews
-\textbf{40}:258--272.
+41. Edwards RA, McNair K, Faust K, Raes J, Dutilh BE. Computational
+approaches to predict bacteriophage-host relationships. FEMS
+Microbiology Reviews. 2015;40: 258--272.
 
 \hypertarget{ref-Roux:2016cc}{}
-42. \textbf{Roux S}, \textbf{Brum JR}, \textbf{Dutilh BE},
-\textbf{Sunagawa S}, \textbf{Duhaime MB}, \textbf{Loy A}, \textbf{Poulos
-BT}, \textbf{Solonenko N}, \textbf{Lara E}, \textbf{Poulain J},
-\textbf{Pesant S}, \textbf{Kandels-Lewis S}, \textbf{Dimier C},
-\textbf{Picheral M}, \textbf{Searson S}, \textbf{Cruaud C},
-\textbf{Alberti A}, \textbf{Duarte CM}, \textbf{Gasol JM}, \textbf{Vaqué
-D}, \textbf{Tara Oceans Coordinators}, \textbf{Bork P}, \textbf{Acinas
-SG}, \textbf{Wincker P}, \textbf{Sullivan MB}. 2016. Ecogenomics and
-potential biogeochemical impacts of globally abundant ocean viruses.
-Nature \textbf{537}:689--693.
+42. Roux S, Brum JR, Dutilh BE, Sunagawa S, Duhaime MB, Loy A, et al.
+Ecogenomics and potential biogeochemical impacts of globally abundant
+ocean viruses. Nature. 2016;537: 689--693.
 
 \hypertarget{ref-Grice:2009eea}{}
-43. \textbf{Grice EA}, \textbf{Kong HH}, \textbf{Conlan S},
-\textbf{Deming CB}, \textbf{Davis J}, \textbf{Young AC}, \textbf{NISC
-Comparative Sequencing Program}, \textbf{Bouffard GG}, \textbf{Blakesley
-RW}, \textbf{Murray PR}, \textbf{Green ED}, \textbf{Turner ML},
-\textbf{Segre JA}. 2009. Topographical and Temporal Diversity of the
-Human Skin Microbiome. Science \textbf{324}:1190--1192.
+43. Grice EA, Kong HH, Conlan S, Deming CB, Davis J, Young AC, et al.
+Topographical and Temporal Diversity of the Human Skin Microbiome.
+Science. 2009;324: 1190--1192.
 
 \hypertarget{ref-Findley:2013jf}{}
-44. \textbf{Findley K}, \textbf{Oh J}, \textbf{Yang J}, \textbf{Conlan
-S}, \textbf{Deming C}, \textbf{Meyer JA}, \textbf{Schoenfeld D},
-\textbf{Nomicos E}, \textbf{Park M}, \textbf{NIH Intramural Sequencing
-Center Comparative Sequencing Program}, \textbf{Kong HH}, \textbf{Segre
-JA}. 2013. Topographic diversity of fungal and bacterial communities in
-human skin. Nature 1--6.
+44. Findley K, Oh J, Yang J, Conlan S, Deming C, Meyer JA, et al.
+Topographic diversity of fungal and bacterial communities in human skin.
+Nature. 2013; 1--6.
 
 \hypertarget{ref-Costello:2009im}{}
-45. \textbf{Costello EK}, \textbf{Lauber CL}, \textbf{Hamady M},
-\textbf{Fierer N}, \textbf{Gordon JI}, \textbf{Knight R}. 2009.
+45. Costello EK, Lauber CL, Hamady M, Fierer N, Gordon JI, Knight R.
 Bacterial community variation in human body habitats across space and
-time. Science \textbf{326}:1694--1697.
+time. Science. 2009;326: 1694--1697.
 
 \hypertarget{ref-Consortium:2012iz}{}
-46. \textbf{Consortium THMP}. 2012. A framework for human microbiome
-research. Nature \textbf{486}:215--221.
+46. Consortium THMP. A framework for human microbiome research. Nature.
+2012;486: 215--221.
 
 \hypertarget{ref-Roux:2015dh}{}
-47. \textbf{Roux S}, \textbf{Enault F}, \textbf{Hurwitz BL},
-\textbf{Sullivan MB}. 2015. VirSorter: mining viral signal from
-microbial genomic data. PeerJ \textbf{3}:e985--20.
+47. Roux S, Enault F, Hurwitz BL, Sullivan MB. VirSorter: mining viral
+signal from microbial genomic data. PeerJ. 2015;3: e985--20.
 
 \hypertarget{ref-Jensen:1998vh}{}
-48. \textbf{Jensen EC}, \textbf{Schrader HS}, \textbf{Rieland B},
-\textbf{Thompson TL}, \textbf{Lee KW}, \textbf{Nickerson KW},
-\textbf{Kokjohn TA}. 1998. Prevalence of broad-host-range lytic
-bacteriophages of Sphaerotilus natans, Escherichia coli, and Pseudomonas
-aeruginosa. Applied and Environmental Microbiology \textbf{64}:575--580.
+48. Jensen EC, Schrader HS, Rieland B, Thompson TL, Lee KW, Nickerson
+KW, et al. Prevalence of broad-host-range lytic bacteriophages of
+Sphaerotilus natans, Escherichia coli, and Pseudomonas aeruginosa.
+Applied and Environmental Microbiology. 1998;64: 575--580.
 
 \hypertarget{ref-Malki:2015tm}{}
-49. \textbf{Malki K}, \textbf{Kula A}, \textbf{Bruder K}, \textbf{Sible
-E}. 2015. Bacteriophages isolated from Lake Michigan demonstrate broad
-host-range across several bacterial phyla. Virology.
+49. Malki K, Kula A, Bruder K, Sible E. Bacteriophages isolated from
+Lake Michigan demonstrate broad host-range across several bacterial
+phyla. Virology. 2015;
 
 \hypertarget{ref-Schwarzer:2012ez}{}
-50. \textbf{Schwarzer D}, \textbf{Buettner FFR}, \textbf{Browning C},
-\textbf{Nazarov S}, \textbf{Rabsch W}, \textbf{Bethe A},
-\textbf{Oberbeck A}, \textbf{Bowman VD}, \textbf{Stummeyer K},
-\textbf{Mühlenhoff M}, \textbf{Leiman PG}, \textbf{Gerardy-Schahn R}.
-2012. A multivalent adsorption apparatus explains the broad host range
+50. Schwarzer D, Buettner FFR, Browning C, Nazarov S, Rabsch W, Bethe A,
+et al. A multivalent adsorption apparatus explains the broad host range
 of phage phi92: a comprehensive genomic and structural analysis. Journal
-of Virology \textbf{86}:10384--10398.
+of Virology. 2012;86: 10384--10398.
 
 \hypertarget{ref-Kim:2012dh}{}
-51. \textbf{Kim S}, \textbf{Rahman M}, \textbf{Seol SY}, \textbf{Yoon
-SS}, \textbf{Kim J}. 2012. Pseudomonas aeruginosa bacteriophage PA1Ø
-requires type IV pili for infection and shows broad bactericidal and
-biofilm removal activities. Applied and Environmental Microbiology
-\textbf{78}:6380--6385.
+51. Kim S, Rahman M, Seol SY, Yoon SS, Kim J. Pseudomonas aeruginosa
+bacteriophage PA1Ø requires type IV pili for infection and shows broad
+bactericidal and biofilm removal activities. Applied and Environmental
+Microbiology. 2012;78: 6380--6385.
 
 \hypertarget{ref-Matsuzaki:1992gw}{}
-52. \textbf{Matsuzaki S}, \textbf{Tanaka S}, \textbf{Koga T},
-\textbf{Kawata T}. 1992. A Broad-Host-Range Vibriophage, KVP40, Isolated
-from Sea Water. Microbiology and Immunology \textbf{36}:93--97.
+52. Matsuzaki S, Tanaka S, Koga T, Kawata T. A Broad-Host-Range
+Vibriophage, KVP40, Isolated from Sea Water. Microbiology and
+Immunology. 1992;36: 93--97.
 
 \hypertarget{ref-Orchard:2014hq}{}
-53. \textbf{Orchard S}, \textbf{Ammari M}, \textbf{Aranda B},
-\textbf{Breuza L}, \textbf{Briganti L}, \textbf{Broackes-Carter F},
-\textbf{Campbell NH}, \textbf{Chavali G}, \textbf{Chen C},
-\textbf{del-Toro N}, \textbf{Duesbury M}, \textbf{Dumousseau M},
-\textbf{Galeota E}, \textbf{Hinz U}, \textbf{Iannuccelli M},
-\textbf{Jagannathan S}, \textbf{Jimenez R}, \textbf{Khadake J},
-\textbf{Lagreid A}, \textbf{Licata L}, \textbf{Lovering RC},
-\textbf{Meldal B}, \textbf{Melidoni AN}, \textbf{Milagros M},
-\textbf{Peluso D}, \textbf{Perfetto L}, \textbf{Porras P},
-\textbf{Raghunath A}, \textbf{Ricard-Blum S}, \textbf{Roechert B},
-\textbf{Stutz A}, \textbf{Tognolli M}, \textbf{Roey K van},
-\textbf{Cesareni G}, \textbf{Hermjakob H}. 2014. The MIntAct
-project--IntAct as a common curation platform for 11 molecular
-interaction databases. Nucleic Acids Research \textbf{42}:D358--63.
+53. Orchard S, Ammari M, Aranda B, Breuza L, Briganti L, Broackes-Carter
+F, et al. The MIntAct project--IntAct as a common curation platform for
+11 molecular interaction databases. Nucleic Acids Research. 2014;42:
+D358--63.
 
 \hypertarget{ref-Kleiner:2015kd}{}
-54. \textbf{Kleiner M}, \textbf{Hooper LV}, \textbf{Duerkop BA}. 2015.
-Evaluation of methods to purify virus-like particles for metagenomic
-sequencing of intestinal viromes. BMC Genomics \textbf{16}:7.
+54. Kleiner M, Hooper LV, Duerkop BA. Evaluation of methods to purify
+virus-like particles for metagenomic sequencing of intestinal viromes.
+BMC Genomics. 2015;16: 7.
 
 \hypertarget{ref-Meisel:2016iu}{}
-55. \textbf{Meisel JS}, \textbf{Hannigan GD}, \textbf{Tyldsley AS},
-\textbf{SanMiguel AJ}, \textbf{Hodkinson BP}, \textbf{Zheng Q},
-\textbf{Grice EA}. 2016. Skin microbiome surveys are strongly influenced
-by experimental design. Journal of Investigative Dermatology.
+55. Meisel JS, Hannigan GD, Tyldsley AS, SanMiguel AJ, Hodkinson BP,
+Zheng Q, et al. Skin microbiome surveys are strongly influenced by
+experimental design. Journal of Investigative Dermatology. 2016;
 
 \hypertarget{ref-Turnbaugh:2009hf}{}
-56. \textbf{Turnbaugh PJ}, \textbf{Ridaura VK}, \textbf{Faith JJ},
-\textbf{Rey FE}, \textbf{Knight R}, \textbf{Gordon JI}. 2009. The effect
-of diet on the human gut microbiome: a metagenomic analysis in humanized
-gnotobiotic mice. Science Translational Medicine
-\textbf{1}:6ra14--6ra14.
+56. Turnbaugh PJ, Ridaura VK, Faith JJ, Rey FE, Knight R, Gordon JI. The
+effect of diet on the human gut microbiome: a metagenomic analysis in
+humanized gnotobiotic mice. Science Translational Medicine. 2009;1:
+6ra14--6ra14.
 
 \hypertarget{ref-David:2014cl}{}
-57. \textbf{David LA}, \textbf{Maurice CF}, \textbf{Carmody RN},
-\textbf{Gootenberg DB}, \textbf{Button JE}, \textbf{Wolfe BE},
-\textbf{Ling AV}, \textbf{Devlin AS}, \textbf{Varma Y},
-\textbf{Fischbach MA}, \textbf{Biddinger SB}, \textbf{Dutton RJ},
-\textbf{Turnbaugh PJ}. 2014. Diet rapidly and reproducibly alters the
-human gut microbiome. Nature \textbf{505}:559--563.
+57. David LA, Maurice CF, Carmody RN, Gootenberg DB, Button JE, Wolfe
+BE, et al. Diet rapidly and reproducibly alters the human gut
+microbiome. Nature. 2014;505: 559--563.
 
 \hypertarget{ref-Jeong:2000im}{}
-58. \textbf{Jeong H}, \textbf{Tombor B}, \textbf{Albert R},
-\textbf{Oltvai ZN}, \textbf{Barabási AL}. 2000. The large-scale
-organization of metabolic networks. Nature \textbf{407}:651--654.
+58. Jeong H, Tombor B, Albert R, Oltvai ZN, Barabási AL. The large-scale
+organization of metabolic networks. Nature. 2000;407: 651--654.
 
 \hypertarget{ref-Grice:2009ee}{}
-59. \textbf{Grice EA}, \textbf{Kong HH}, \textbf{Conlan S},
-\textbf{Deming CB}, \textbf{Davis J}, \textbf{Young AC}, \textbf{NISC
-Comparative Sequencing Program}, \textbf{Bouffard GG}, \textbf{Blakesley
-RW}, \textbf{Murray PR}, \textbf{Green ED}, \textbf{Turner ML},
-\textbf{Segre JA}. 2009. Topographical and Temporal Diversity of the
-Human Skin Microbiome. Science \textbf{324}:1190--1192.
+59. Grice EA, Kong HH, Conlan S, Deming CB, Davis J, Young AC, et al.
+Topographical and Temporal Diversity of the Human Skin Microbiome.
+Science. 2009;324: 1190--1192.
 
 \hypertarget{ref-Minot:2013ih}{}
-60. \textbf{Minot S}, \textbf{Bryson A}, \textbf{Chehoud C}, \textbf{Wu
-GD}, \textbf{Lewis JD}, \textbf{Bushman FD}. 2013. Rapid evolution of
-the human gut virome. Proceedings of the National Academy of Sciences of
-the United States of America \textbf{110}:12450--12455.
+60. Minot S, Bryson A, Chehoud C, Wu GD, Lewis JD, Bushman FD. Rapid
+evolution of the human gut virome. Proceedings of the National Academy
+of Sciences of the United States of America. 2013;110: 12450--12455.
 
 \hypertarget{ref-Schloss:2005hz}{}
-61. \textbf{Schloss PD}, \textbf{Handelsman J}. 2005. Introducing DOTUR,
-a computer program for defining operational taxonomic units and
-estimating species richness. Applied and Environmental Microbiology
-\textbf{71}:1501--1506.
+61. Schloss PD, Handelsman J. Introducing DOTUR, a computer program for
+defining operational taxonomic units and estimating species richness.
+Applied and Environmental Microbiology. 2005;71: 1501--1506.
 
 \hypertarget{ref-Yilmaz:2010jb}{}
-62. \textbf{Yilmaz S}, \textbf{Allgaier M}, \textbf{Hugenholtz P}. 2010.
-Multiple displacement amplification compromises quantitative analysis of
-metagenomes. Nature Methods \textbf{7}:943--944.
+62. Yilmaz S, Allgaier M, Hugenholtz P. Multiple displacement
+amplification compromises quantitative analysis of metagenomes. Nature
+Methods. 2010;7: 943--944.
 
 \hypertarget{ref-Kim:2008to}{}
-63. \textbf{Kim KH}, \textbf{Chang HW}, \textbf{Nam YD}, \textbf{Roh
-SW}. 2008. Amplification of uncultured single-stranded DNA viruses from
-rice paddy soil. Applied and \ldots{}.
+63. Kim KH, Chang HW, Nam YD, Roh SW. Amplification of uncultured
+single-stranded DNA viruses from rice paddy soil. Applied and \ldots{}.
+2008;
 
 \hypertarget{ref-Kim:2011hp}{}
-64. \textbf{Kim K-H}, \textbf{Bae J-W}. 2011. Amplification methods bias
-metagenomic libraries of uncultured single-stranded and double-stranded
-DNA viruses. Applied and Environmental Microbiology
-\textbf{77}:7663--7668.
+64. Kim K-H, Bae J-W. Amplification methods bias metagenomic libraries
+of uncultured single-stranded and double-stranded DNA viruses. Applied
+and Environmental Microbiology. 2011;77: 7663--7668.
 
 \hypertarget{ref-Minot:2012ed}{}
-65. \textbf{Minot S}, \textbf{Wu GD}, \textbf{Lewis JD}, \textbf{Bushman
-FD}. 2012. Conservation of gene cassettes among diverse viruses of the
-human gut. PLOS ONE \textbf{7}:e42342.
+65. Minot S, Wu GD, Lewis JD, Bushman FD. Conservation of gene cassettes
+among diverse viruses of the human gut. PLOS ONE. 2012;7: e42342.
 
 \hypertarget{ref-Deng:2014eb}{}
-66. \textbf{Deng L}, \textbf{Ignacio-Espinoza JC}, \textbf{Gregory AC},
-\textbf{Poulos BT}, \textbf{Weitz JS}, \textbf{Hugenholtz P},
-\textbf{Sullivan MB}. 2014. Viral tagging reveals discrete populations
-in Synechococcus viral genome sequence space. Nature
-\textbf{513}:242--245.
+66. Deng L, Ignacio-Espinoza JC, Gregory AC, Poulos BT, Weitz JS,
+Hugenholtz P, et al. Viral tagging reveals discrete populations in
+Synechococcus viral genome sequence space. Nature. 2014;513: 242--245.
 
 \hypertarget{ref-Brum:2015iaa}{}
-67. \textbf{Brum JR}, \textbf{Ignacio-Espinoza JC}, \textbf{Roux S},
-\textbf{Doulcier G}, \textbf{Acinas SG}, \textbf{Alberti A},
-\textbf{Chaffron S}, \textbf{Cruaud C}, \textbf{Vargas C de},
-\textbf{Gasol JM}, \textbf{Gorsky G}, \textbf{Gregory AC}, \textbf{Guidi
-L}, \textbf{Hingamp P}, \textbf{Iudicone D}, \textbf{Not F},
-\textbf{Ogata H}, \textbf{Pesant S}, \textbf{Poulos BT},
-\textbf{Schwenck SM}, \textbf{Speich S}, \textbf{Dimier C},
-\textbf{Kandels-Lewis S}, \textbf{Picheral M}, \textbf{Searson S},
-\textbf{Tara Oceans Coordinators}, \textbf{Bork P}, \textbf{Bowler C},
-\textbf{Sunagawa S}, \textbf{Wincker P}, \textbf{Karsenti E},
-\textbf{Sullivan MB}. 2015. Ocean plankton. Patterns and ecological
-drivers of ocean viral communities. Science
-\textbf{348}:1261498--1261498.
+67. Brum JR, Ignacio-Espinoza JC, Roux S, Doulcier G, Acinas SG, Alberti
+A, et al. Ocean plankton. Patterns and ecological drivers of ocean viral
+communities. Science. 2015;348: 1261498--1261498.
 
 \hypertarget{ref-Malki:2015dw}{}
-68. \textbf{Malki K}, \textbf{Kula A}, \textbf{Bruder K}, \textbf{Sible
-E}, \textbf{Hatzopoulos T}, \textbf{Steidel S}, \textbf{Watkins SC},
-\textbf{Putonti C}. 2015. Bacteriophages isolated from Lake Michigan
-demonstrate broad host-range across several bacterial phyla. Virology
-Journal \textbf{12}:164.
+68. Malki K, Kula A, Bruder K, Sible E, Hatzopoulos T, Steidel S, et al.
+Bacteriophages isolated from Lake Michigan demonstrate broad host-range
+across several bacterial phyla. Virology Journal. 2015;12: 164.
 
 \hypertarget{ref-Polz:2006fi}{}
-69. \textbf{Polz MF}, \textbf{Hunt DE}, \textbf{Preheim SP},
-\textbf{Weinreich DM}. 2006. Patterns and mechanisms of genetic and
-phenotypic differentiation in marine microbes. Philosophical
-Transactions of the Royal Society B: Biological Sciences
-\textbf{361}:2009--2021.
+69. Polz MF, Hunt DE, Preheim SP, Weinreich DM. Patterns and mechanisms
+of genetic and phenotypic differentiation in marine microbes.
+Philosophical Transactions of the Royal Society B: Biological Sciences.
+2006;361: 2009--2021.
 
 \hypertarget{ref-Gregory:2016cg}{}
-70. \textbf{Gregory AC}, \textbf{Solonenko SA}, \textbf{Ignacio-Espinoza
-JC}, \textbf{LaButti K}, \textbf{Copeland A}, \textbf{Sudek S},
-\textbf{Maitland A}, \textbf{Chittick L}, \textbf{Dos Santos F},
-\textbf{Weitz JS}, \textbf{Worden AZ}, \textbf{Woyke T},
-\textbf{Sullivan MB}. 2016. Genomic differentiation among wild
-cyanophages despite widespread horizontal gene transfer. BMC Genomics
-\textbf{17}:930.
+70. Gregory AC, Solonenko SA, Ignacio-Espinoza JC, LaButti K, Copeland
+A, Sudek S, et al. Genomic differentiation among wild cyanophages
+despite widespread horizontal gene transfer. BMC Genomics. 2016;17: 930.
 
 \hypertarget{ref-PaezEspino:2016gi}{}
-71. \textbf{Paez-Espino D}, \textbf{Eloe-Fadrosh EA},
-\textbf{Pavlopoulos GA}, \textbf{Thomas AD}, \textbf{Huntemann M},
-\textbf{Mikhailova N}, \textbf{Rubin E}, \textbf{Ivanova NN},
-\textbf{Kyrpides NC}. 2016. Uncovering Earth's virome. Nature.
+71. Paez-Espino D, Eloe-Fadrosh EA, Pavlopoulos GA, Thomas AD, Huntemann
+M, Mikhailova N, et al. Uncovering Earth's virome. Nature. 2016;
 
 \hypertarget{ref-Grice:2011gy}{}
-72. \textbf{Grice EA}, \textbf{Segre JA}. 2011. The skin microbiome.
-Nature Reviews Microbiology \textbf{9}:244--253.
+72. Grice EA, Segre JA. The skin microbiome. Nature Reviews
+Microbiology. 2011;9: 244--253.
 
 \hypertarget{ref-Round:2009bz}{}
-73. \textbf{Round JL}, \textbf{Mazmanian SK}. 2009. The gut microbiota
-shapes intestinal immune responses during health and disease. Nature
-reviews Immunology \textbf{9}:313--323.
+73. Round JL, Mazmanian SK. The gut microbiota shapes intestinal immune
+responses during health and disease. Nature reviews Immunology. 2009;9:
+313--323.
 
 \hypertarget{ref-FASTXToolkit:wr}{}
-74. \textbf{Hannon GJ}. FASTX-Toolkit GNU Affero General Public License.
+74. Hannon GJ. FASTX-Toolkit.: GNU Affero General Public License.
 
 \hypertarget{ref-Li:2016kd}{}
-75. \textbf{Li D}, \textbf{Luo R}, \textbf{Liu C-M}, \textbf{Leung C-M},
-\textbf{Ting H-F}, \textbf{Sadakane K}, \textbf{Yamashita H},
-\textbf{Lam T-W}. 2016. MEGAHIT v1.0: A fast and scalable metagenome
-assembler driven by advanced methodologies and community practices.
-METHODS \textbf{102}:3--11.
+75. Li D, Luo R, Liu C-M, Leung C-M, Ting H-F, Sadakane K, et al.
+MEGAHIT v1.0: A fast and scalable metagenome assembler driven by
+advanced methodologies and community practices. METHODS. 2016;102:
+3--11.
 
 \hypertarget{ref-Langmead:2012jh}{}
-76. \textbf{Langmead B}, \textbf{Salzberg SL}. 2012. Fast gapped-read
-alignment with Bowtie 2. Nature Methods \textbf{9}:357--359.
+76. Langmead B, Salzberg SL. Fast gapped-read alignment with Bowtie 2.
+Nature Methods. 2012;9: 357--359.
 
 \hypertarget{ref-Alneberg:2014fc}{}
-77. \textbf{Alneberg J}, \textbf{Bjarnason BS aacute ri}, \textbf{Bruijn
-I de}, \textbf{Schirmer M}, \textbf{Quick J}, \textbf{Ijaz UZ},
-\textbf{Lahti L}, \textbf{Loman NJ}, \textbf{Andersson AF},
-\textbf{Quince C}. 2014. Binning metagenomic contigs by coverage and
-composition. Nature Methods 1--7.
+77. Alneberg J, Bjarnason BS aacute ri, Bruijn I de, Schirmer M, Quick
+J, Ijaz UZ, et al. Binning metagenomic contigs by coverage and
+composition. Nature Methods. 2014; 1--7.
 
 \hypertarget{ref-Hyatt:2012cy}{}
-78. \textbf{Hyatt D}, \textbf{LoCascio PF}, \textbf{Hauser LJ},
-\textbf{Uberbacher EC}. 2012. Gene and translation initiation site
-prediction in metagenomic sequences. Bioinformatics
-\textbf{28}:2223--2230.
+78. Hyatt D, LoCascio PF, Hauser LJ, Uberbacher EC. Gene and translation
+initiation site prediction in metagenomic sequences. Bioinformatics.
+2012;28: 2223--2230.
 
 \hypertarget{ref-caretClassificatio:ux5fU2Litux5f1}{}
-79. \textbf{Kuhn M}. caret: Classification and Regression Training.
+79. Kuhn M. caret: Classification and Regression Training.
 
 \hypertarget{ref-Edgar:2007bh}{}
-80. \textbf{Edgar RC}. 2007. PILER-CR: fast and accurate identification
-of CRISPR repeats. BMC Bioinformatics \textbf{8}:18.
+80. Edgar RC. PILER-CR: fast and accurate identification of CRISPR
+repeats. BMC Bioinformatics. 2007;8: 18.
 
 \hypertarget{ref-Camacho:2009fc}{}
-81. \textbf{Camacho C}, \textbf{Coulouris G}, \textbf{Avagyan V},
-\textbf{Ma N}, \textbf{Papadopoulos J}, \textbf{Bealer K},
-\textbf{Madden TL}. 2009. BLAST+: architecture and applications. BMC
-Bioinformatics \textbf{10}:1.
+81. Camacho C, Coulouris G, Avagyan V, Ma N, Papadopoulos J, Bealer K,
+et al. BLAST+: architecture and applications. BMC Bioinformatics.
+2009;10: 1.
 
 \hypertarget{ref-Buchfink:2015ki}{}
-82. \textbf{Buchfink B}, \textbf{Xie C}, \textbf{Huson DH}. 2015. Fast
-and sensitive protein alignment using DIAMOND. Nature Methods
-\textbf{12}:59--60.
+82. Buchfink B, Xie C, Huson DH. Fast and sensitive protein alignment
+using DIAMOND. Nature Methods. 2015;12: 59--60.
 
@@
 \hypertarget{ref-Theigraphsoftware:vh}{}
-84. \textbf{Csardi G}, \textbf{Nepusz T}. The igraph software package
-for complex network research.
+84. Csardi G, Nepusz T. The igraph software package for complex network
+research.
 

### 2017-10-01 doc/oman.tex
Updated manusript with format for new submission.

--- 2017-10-01 doc/manuscript.tex
@@
 \hypersetup{unicode=true,
-            pdftitle={Biogeography \& environmental conditions shape bacteriophage \& bacteria interaction networks across the human microbiome},
+            pdftitle={Biogeography \& Environmental Conditions Shape Phage \& Bacteria Interaction Networks Across the Human Microbiome},
             pdfborder={0 0 0},
@@
 
-\title{Biogeography \& environmental conditions shape bacteriophage \& bacteria
-interaction networks across the human microbiome}
+\title{Biogeography \& Environmental Conditions Shape Phage \& Bacteria
+Interaction Networks Across the Human Microbiome}
     \usepackage{authblk}
@@
                             \affil[1]{\nextfont Department of Microbiology \& Immunology, University of Michigan, Ann
-Arbor, Michigan, 48109, USA}
+Arbor, Michigan, 48109}
                     \affil[2]{\nextfont Department of Ecology \& Evolutionary Biology, University of Michigan,
-Ann Arbor, Michigan, 48109, USA}
+Ann Arbor, Michigan, 48109}
                     \affil[3]{\nextfont Department of Computer Science, University of Michigan, Ann Arbor,
-Michigan, 48109, USA}
+Michigan, 48109}
                     \affil[*]{\nextfont To whom correspondence may be addressed.}
@@
 
-\textbf{\emph{Short Title}}: Network Diversity of the Healthy Human
+\textbf{\emph{Running Title}}: Network Diversity of the Healthy Human
 Microbiome
@@
 conserved among cohabitating family members. High-fat diets and obesity
-appeared to be associated with less connected networks. Network
-structure differed between skin sites, with those exposed to the
-external environment being less connected and more prone to instability.
-This study quantified and contrasted the diversity of virome-microbiome
-networks across the human body and illustrated how environmental factors
-may influence phage-bacteria interactive dynamics. This work provides a
-baseline for future studies to better understand system perturbations,
-such as disease states, through ecological networks.
+were associated with less connected networks. Network structure differed
+between skin sites, with those exposed to the external environment being
+less connected and more prone to instability. This study quantified and
+contrasted the diversity of virome-microbiome networks across the human
+body and illustrated how environmental factors may influence
+phage-bacteria interactive dynamics. This work provides a baseline for
+future studies to better understand system perturbations, such as
+disease states, through ecological networks.
 
@@
 
-\section{Author Summary}\label{author-summary}
+\section{Importance}\label{importance}
 
@@
 communities, which frequently share genetic information and work
-together to maintain stable ecosystems. Removal of bacteria or phages
-can disrupt or even collapse those ecosystems {[}18,28--37{]}.
+together to maintain stable ecosystems. Removal of bacteria or phage can
+disrupt or even collapse those ecosystems {[}18,28--37{]}.
 Relationship-based network approaches allow us to capture this
@@
 mothers {[}38,39{]}. We selected these datasets because their virome
-samples were subjected to virus-like particle (VLP) purification, which
-removed contaminating DNA from human cells, bacteria, etc. To this end,
-the publishing authors employed combinations of filtration,
-chloroform/DNase treatment, and cesium chloride gradients to eliminate
-organismal DNA (e.g.~bacteria, human, fungi, etc) and thereby allow for
-direct assessment of both the extracellular and fully-assembled
-intracellular virome \textbf{(Supplemental Figure \ref{SequenceStats}
-A-B)} {[}14,39{]}. Each research group reported quality control measures
-to ensure the purity of the virome sequence datasets, using both
-computational and molecular techniques (e.g.~16S rRNA gene qPCR)
-\textbf{(Table S1)}. These reports confirmed that the virome libraries
-consisted of highly purified virus genomic DNA.
+samples were subjected to virus-like particle (VLP) purification. To
+this end, they employed combinations of filtration, chloroform/DNase
+treatment, and cesium chloride gradients to eliminate organismal DNA and
+thereby allow for direct assessment of both the extracellular and
+fully-assembled intracellular virome \textbf{(Supplemental Figure
+\ref{SequenceStats} A-B)} {[}14,39{]}. While the whole metagenomic
+shotgun sequence samples were not subjected to purification, they
+primarily consisted of bacteria {[}13,14,38,39{]}.
 
@@
 
-While supplementing the previous quality control measures \textbf{(Table
-S1)} we found that, in light of the rigorous purification and quality
-control during sample collection and preparation, 77\% (228 / 298
-operational genomic units) still had some nucleotide similarity to a
-given bacterial reference genome (e-value \textless{} \(10^{-25}\)). As
-absence of bacterial contamination had been confirmed by sensitive
-molecular methods \textbf{(Table S1)}, we interpreted this as evidence
-that the majority of the gut and skin bacteriophages were temperate and
-thereby shared elements with bacterial reference genomes that contained
-similar integrated phages when sequenced--a trend previously reported
-{[}14{]}. Additionally, we identified two OGUs as being complete phages
-using the stringent Virsorter phage identification algorithm (class 1
-confidence group) {[}47{]}.
-
-The whole metagenomic shotgun sequence samples primarily consisted of
-bacteria, with an average viral relative abundance of 0.4\%
-\textbf{(Table S1)} {[}13,14,38,39{]}. We found that only 2\% (6 / 280
-OGUs) of bacterial OGUs had significantly strong nucleotide similarity
-to phage reference genomes (e-value \textless{} \(10^{-25}\))
-{[}13,14,38,39{]}. No OGUs were confidently identified as lytic or
-temperate phage OGUs in the bacterial dataset using the Virsorter
-algorithm {[}47{]}. Together this suggests only minimal
-``contamination'' of the bacterial OGUs.
-
 \subsection{Evaluating the Model to Predict Phage-Bacteria
@@
 random forest model trained on experimentally validated infectious
-relationships from six previous publications {[}41,48--52{]}. Only
+relationships from six previous publications {[}41,47--51{]}. Only
 bacteria and phages were used in the model. The training set contained
@@
 both broad and specific ranges of infection \textbf{(Supplemental Figure
-\ref{ValidationOverview} A - B)}. While it is true that there are more
-known phages that infect bacteria, we were limited by the information
-confirming which phages do not infect certain bacteria and attempted to
-keep the numbers of positive and negative interactions similar. Phages
-with linear and circular genomes, as well as ssDNA and dsDNA genomes,
-were included in the analysis. Because we used DNA sequencing studies,
-RNA phages were not considered \textbf{(Supplemental Figure
-\ref{ValidationOverview} C-D)}. This training set included both positive
-relationships (a phage infects a bacterium) and negative relationships
-(a phage does not infect a bacterium). This allowed us to validate the
-false positive and false negative rates associated with our candidate
-models, thereby building upon previous work that only considered
-positive relationships {[}41{]}.
+\ref{ValidationOverview} A - B)}. Phages with linear and circular
+genomes, as well as ssDNA and dsDNA genomes, were included in the
+analysis. Because we used DNA sequencing studies, RNA phages were not
+considered \textbf{(Supplemental Figure \ref{ValidationOverview} C-D)}.
+This training set included both positive relationships (a phage infects
+a bacterium) and negative relationships (a phage does not infect a
+bacterium). This allowed us to validate the false positive and false
+negative rates associated with our candidate models, thereby building
+upon previous work that only considered positive relationships {[}41{]}.
 
@@
 similarity of protein families associated with experimentally identified
-protein-protein interactions {[}53{]}. The resulting random forest model
-was assessed using nested cross validation, and the median area under
-its receiver operating characteristic (ROC) curve was 0.788, the median
-model sensitivity was 0.952, and median specificity was 0.615
-\textbf{(Figure \ref{RocCurve} A)}. The most important predictor in the
-model was amino acid similarity between genes, followed by nucleotide
-similarity of whole genomes \textbf{(Figure \ref{RocCurve} B)}. Protein
-family interactions were moderately important to the model, and CRISPRs
-were largely uninformative, due to the minimal amount of identifiable
-CRISPRs in the dataset and their redundancy with the nucleotide
-similarity methods \textbf{(Figure \ref{RocCurve} B)}. Approximately one
-third of the training set relationships yielded no score and therefore
-were unable to be assigned an interaction prediction \textbf{(Figure
-\ref{RocCurve} C)}.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=0.85000\textwidth]{}
-\caption{\textbf{Summary of Multi-Study Network Model.} \emph{(A) Median
-ROC curve (dark red) used to create the microbiome-virome infection
-prediction model, based on nested cross validation over 25 random
-iterations. The maximum and minimum performance are shown in light red.
-(B) Importance scores associated with the metrics used in the random
-forest model to predict relationships between bacteria and phages. The
-importance score is defined as the mean decrease in accuracy of the
-model when a feature (e.g.~Pfam) is excluded. Features include the local
-gene alignments between bacteria and phage genes (denoted
-\texttt{blastx}; the blastx algorithm in Diamond aligner), local genome
-nucleotide alignments between bacteria and phage OGUs, presence of
-experimentally validated protein family domains (Pfams) between phage
-and bacteria OGUs, and CRISPR targeting of bacteria toward phages
-(CRISPR). (C) Proportions of samples included (gray) and excluded (red)
-in the model. Samples were excluded from the model because they did not
-yield any scores. Those interactions without scores were automatically
-classified as not having interactions. (D) Bipartite visualization of
-the resulting phage-bacteria network. Phage OGUs are presented in
-orange, bacteria OGUs in red, and their interaction edges are
-represented as connecting lines. This network includes information from
-all three published studies. (E) Network diameter (measure of graph
-size; the greatest number of traversed vertices required between two
-vertices), (F) number of vertices, and (G) number of edges
-(relationships) for the total network (orange) and the individual study
-sub-networks (diet study = red, skin study = yellow, twin study =
-green).} \label{RocCurve}}
-\end{figure}
+protein-protein interactions {[}52{]}. The resulting random forest model
+was assessed and the area under its receiver operating characteristic
+(ROC) curve was 0.846, the model sensitivity was 0.829, and specificity
+was 0.767 \textbf{(Figure \ref{RocCurve} A)}. The most important
+predictor in the model was amino acid similarity between genes, followed
+by nucleotide similarity of whole genomes \textbf{(Figure \ref{RocCurve}
+B)}. Protein family interactions were moderately important to the model,
+and CRISPRs were largely uninformative, due to the minimal amount of
+identifiable CRISPRs in the dataset and their redundancy with the
+nucleotide similarity methods \textbf{(Figure \ref{RocCurve} B)}.
+Approximately one third of the training set relationships yielded no
+score and therefore were unable to be assigned an interaction prediction
+\textbf{(Figure \ref{RocCurve} C)}.
 
@@
 stored in the master network for parsing in downstream analyses
-\textbf{(Supplemental Figure \ref{NetworkDiagram})}. Bacterial and phage
-relative abundance was recorded in each sample for each OGU and the
-weight of the edge connecting those OGUs was calculated as a function of
-those relative abundance values. The separate extraction of the phage
-and bacterial libraries ensured a more accurate measurement of the
-microbial communities, as has been outlined previously {[}54,55{]}. The
-master network was highly connected and contained 72,287 infectious
-relationships among 578 nodes, representing 298 phages and 280 bacteria.
-Although the network was highly connected, not all relationships were
-present in all samples. Relationships were weighted by the relative
-abundances of their associated bacteria and phages. Like the master
-network, the skin network exhibited a diameter of 4 (measure of graph
-size; the greatest number of traversed vertices required between two
-vertices) and included 576 (297 phages, 279 bacteria, 99.7\% total) and
-72,127 (99.8\%) of the master network nodes and edges, respectively
+\textbf{(Supplemental Figure \ref{NetworkDiagram})}. The master network
+was highly connected and contained 72,287 infectious relationships among
+578 nodes, representing 298 phages and 280 bacteria. Although the
+network was highly connected, not all relationships were present in all
+samples. As relationships were weighted by the relative abundances of
+their associated bacteria and phages, lowly abundant relationships could
+be present but not highly abundant. Like the master network, the skin
+network exhibited a diameter of 4 (measure of graph size; the greatest
+number of traversed vertices required between two vertices) and included
+99.7\% and 99.8\% of the master network nodes and edges, respectively
 \textbf{(Figure \ref{RocCurve} E - F)}. The phages and bacteria in the
-diet and twin sample sets were more sparsely related, with the diet
-study consisting of 89 (41 phages, 48 bacteria) nodes and 5,566
-relationships, and the twin study containing 137 (36 phages, 101
-bacteria) nodes and 17,250 relationships \textbf{(Figure \ref{RocCurve}
-E - F)}.
+gut diet and twin sample sets were more sparsely related: each contained
+fewer than 150 vertices, fewer than 20,000 relationships, and diameters
+of 3 \textbf{(Figure \ref{RocCurve} E - F)}.
 
@@
 availability and gut microbiome composition and diversity, including
-bacteria and phages {[}14,56,57{]}. Previous work in isolated
+bacteria and phages {[}14,53,54{]}. Previous work in isolated
 culture-based systems has suggested that changes in nutrient
@@
 overall greater degree of connections, which suggests a greater
-resilience against instability. This is because more highly connected
-networks are more stable as a result of pathway dependencies and the
-unlikelihood that a randomly removed bacteria or phage would cause major
-divisions across the network {[}30,58{]}. We used this information to
-calculate the average connectedness per sample, which was corrected for
-the maximum potential degree of connectedness. Unfortunately our dataset
-was insufficiently powered to make strong conclusions toward this
-hypothesis, but this is an interesting observation that warrants further
-investigation.
-
-Using our limited sample set, we observed that the gut microbiome
-network structures associated with high-fat diets appeared less
-connected than those of low-fat diets, although a greater sample size
-will be required to more properly evaluate this trend \textbf{(Figure
-\ref{dietnetworks} A-B)}. Five subjects were available for use, all of
-which had matching bacteria and virome datasets and samples from 8-10
-days following the initiation of their diets. High-fat diets appeared to
-exhibit reduced degree centrality \textbf{(Figure \ref{dietnetworks}
-A)}, suggesting bacteria in high-fat environments were targeted by fewer
-phages and that phage tropism was more restricted. High-fat diets also
-appeared to exhibit decreased closeness centrality \textbf{(Figure
-\ref{dietnetworks} B)}, indicating that bacteria and phages were more
-distant from other bacteria and phages in the community. This would make
-genetic transfer and altered abundance of a given phage or bacterium
-less capable of impacting other bacteria and phages within the network.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=0.90000\textwidth]{}
-\caption{\textbf{Impact of Diet and Obesity on Gut Network Structure.}
-\emph{(A) Quantification of average degree centrality (number of edges
-per node) and (B) closeness centrality (average distance from each node
-to every other node) of gut microbiome networks of subjects limited to
-exclusively high-fat or low-fat diets. Each point represents the
-centrality from a human subject stool sample that was collected 8-10
-days following the beginning of their defined diet. There are five
-samples here, compared to the four in figure \ref{intradiv}, because one
-of the was only sampled post-diet, providing us data for this analysis
-but not allowing us to compare to a baseline for figure \ref{intradiv}.
-(C) Quantification of average degree centrality and (D) closeness
-centrality between obese and healthy adult women from the Twin gut
-study. Each point represents a stool sample taken from one of the three
-adult woman confirmed as obese or healthy and with matching virus and
-bacteria data.} \label{dietnetworks}}
-\end{figure}
-
-In addition to diet, we found preliminary evidence that obesity
-influenced network structure. This was done using the three mother
-samples available from the twin sample set, all of which had matching
-bacteria and phage samples and confirmed BMI information. The
-obesity-associated network appeared to have a higher degree centrality
+resilience against instability. We used this information to calculate
+the average connectedness per sample, which was corrected for the
+maximum potential degree of connectedness.
+
+We found that the gut microbiome network structures associated with
+high-fat diets were less connected than those of low-fat diets
+\textbf{(Figure \ref{dietnetworks} A-B)}. Tests for statistical
+differences were not performed due to the small sample size. High-fat
+diets exhibited reduced degree centrality \textbf{(Figure
+\ref{dietnetworks} A)}, suggesting bacteria in high-fat environments
+were targeted by fewer phages and that phage tropism was more
+restricted. High-fat diets also exhibited decreased closeness centrality
+\textbf{(Figure \ref{dietnetworks} B)}, indicating that bacteria and
+phages were more distant from other bacteria and phages in the
+community. This would make genetic transfer and altered abundance of a
+given phage or bacterium less capable of impacting other bacteria and
+phages within the network.
+
+In addition to diet, obesity was found to influence network structure.
+Obesity-associated networks demonstrated a higher degree centrality
 \textbf{(Figure \ref{dietnetworks} C)}, but less closeness centrality
@@
 less connected, having microbes further from all other microbes within
-the community. This again comes with the caveat that this is a
-preliminary observation with too small of a sample size to make more
-substantial claims.
+the community.
 
@@
 with people remaining more similar to themselves than to other people
-over time {[}13,59,60{]}. We therefore hypothesized that this personal
+over time {[}13,55,56{]}. We therefore hypothesized that this personal
 conservation extended to microbiome network structure. We addressed this
@@
 others, although this difference was not statistically significant
-(p-value = 0.125, \textbf{Figure \ref{intradiv} B}). Four of the five
-available subjects were used because one of the subjects was not sampled
-at the initial time point. The lack of statistical confidence was likely
-due to the small sample size of this dataset.
-
-Although there was evidence for gut network conservation among
+(p-value = 0.125, \textbf{Figure \ref{intradiv} B}). The lack of
+statistical confidence was likely due to the small sample size of this
+dataset. Although there was evidence for gut network conservation among
 individuals, we found no evidence for conservation of gut network
@@
 similar within families (twins and their mothers; intrafamily) compared
-to other families (other twins and mothers; inter-family) (p-value =
-0.312, \textbf{Figure \ref{intradiv} C}). In addition to the gut, skin
-microbiome network structure was strongly conserved within individuals
-(p-value \textless{} 0.001, \textbf{Figure \ref{intradiv} D}). This
-distribution was similar when separated by anatomical sites. Most sites
-were statistically significantly more conserved within individuals
-\textbf{(Supplemental Figure \ref{allskin})}.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=0.90000\textwidth]{}
-\caption{\textbf{Intrapersonal vs Interpersonal Network Dissimilarity
-Across Different Human Systems.} \emph{(A) NMDS ordination illustrating
-network dissimilarity between subjects over time. Each sample is colored
-by subject, with each colored sample pair collected 8-10 days apart.
-Dissimilarity was calculated using the Bray-Curtis metric based on
-abundance weighted eigenvector centrality signatures, with a greater
-distance representing greater dissimilarity in bacteria and phage
-centrality and abundance. Only four subjects were included here,
-compared to the five used in figure \ref{dietnetworks}, because one of
-the subjects was missing the inital sampling time point and therefore
-lacked temporal sampling. (B) Quantification of gut network
-dissimilarity within the same subject over time (intrapersonal) and the
-mean dissimilarity between the subject of interest and all other
-subjects (interpersonal). The p-value is provided near the bottom of the
-figure. (C) Quantification of gut network dissimilarity within subjects
-from the same family (intrafamily) and the mean dissimilarity between
-subjects within a family and those of other families (interfamily). Each
-point represents the inter-family and intra-family dissimilarity of a
-twin or mother that was sampled over time. (D) Quantification of skin
-network dissimilarity within the same subject and anatomical location
-over time (intrapersonal) and the mean dissimilarity between the subject
-of interest and all other subjects at the same time and the same
-anatomical location (interpersonal). All p-values were calculated using
-a paired Wilcoxon test.}\label{intradiv}}
-\end{figure}
+to other families (inter-family) (p-value = 0.312, \textbf{Figure
+\ref{intradiv} C}). In addition to the gut, skin microbiome network
+structure was strongly conserved within individuals (p-value \textless{}
+0.001, \textbf{Figure \ref{intradiv} D}). This distribution was similar
+when separated by anatomical sites. Most sites were statistically
+significantly more conserved within individuals \textbf{(Supplemental
+Figure \ref{allskin})}.
 
@@
 of the healthy human skin microbiome between anatomical sites, including
-bacteria, virus, and fungal communities {[}13,44,59{]}. These
+bacteria, virus, and fungal communities {[}13,44,55{]}. These
 communities vary by degree of skin moisture, oil, and environmental
@@
 \ref{skinnetwork} B)}.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=0.75000\textwidth]{}
-\caption{\textbf{Impact of Skin Micro-Environment on Microbiome Network
-Structure.} \emph{(A) Notched box-plot depicting differences in average
-eigenvector centrality between moist, intermittently moist, and
-sebaceous skin sites and (B) occluded, intermittently occluded, and
-exposed sites. Notched box-plots were created using ggplot2 and show the
-median (center line), the inter-quartile range (IQR; upper and lower
-boxes), the highest and lowest value within 1.5 * IQR (whiskers),
-outliers (dots), and the notch which provides an approximate 95\%
-confidence interval as defined by 1.58 * IQR / sqrt(n). Sample sizes for
-each group were: Moist = 81, Sebaceous = 56, IntMoist = 56, Occluded =
-106, Exposed = 61, IntOccluded = 26. (C) NMDS ordination depicting the
-differences in skin microbiome network structure between skin moisture
-levels and (D) occlusion. Samples are colored by their environment and
-their dissimilarity to other samples was calculated as described in
-figure \ref{intradiv}. (E) The statistical differences of networks
-between moisture and (F) occlusion status were quantified with an anova
-and post hoc Tukey test. Cluster centroids are represented by dots and
-the extended lines represent the associated 95\% confidence intervals.
-Significant comparisons (p-value \textless{} 0.05) are colored in red,
-and non-significant comparisons are gray.}\label{skinnetwork}}
-\end{figure}
 
@@
 microbiome by characterizing bacterial and viral diversity across the
-human body {[}13,14,43--45,61{]}. Here, we offer an initial
+human body {[}13,14,43--45,57{]}. Here, we offer an initial
 understanding of how phage-bacteria networks differ throughout the human
@@
 Just as gut microbiome and virome composition and diversity are
-conserved in individuals {[}13,43,44,60{]}, gut and skin microbiome
+conserved in individuals {[}13,43,44,56{]}, gut and skin microbiome
 network structures were conserved within individuals over time. Gut
@@
 suggested that the community properties inferred from microbiome
-interaction network structures, such as stability (meaning a more highly
-connected network is more stable because a randomly removed bacteria or
-phage node is less likely to divide or disintegrate {[}30,58{]} the
-overall network), the potential for horizontal gene transfer between
-members, and co-evolution of populations, were person-specific. These
-properties may be impacted by personal factors ranging from the body's
-immune system to external environmental conditions, such as climate and
-diet.
-
-We observed evidence supporting the ability of environmental conditions
-to shape gut and skin microbiome interaction network structure by
-observing that diet and skin location were associated with altered
-network structures. We found evidence that diet was sufficient to alter
-gut microbiome network connectivity, although this needs to be
-interpreted as a case observation, due to the small sample size.
-Although our sample size was small, our findings provided some
-preliminary evidence that high-fat diets were less connected than
-low-fat diets and that high-fat diets therefore may lead to less stable
+interaction network structures, such as stability, the potential for
+horizontal gene transfer between members, and co-evolution of
+populations, were person-specific. These properties may be impacted by
+personal factors ranging from the body's immune system to external
+environmental conditions, such as climate and diet.
+
+The ability of environmental conditions to shape gut and skin microbiome
+interaction network structure was further supported by our finding that
+diet and skin location were associated with altered network structures.
+We found evidence that diet was sufficient to alter gut microbiome
+network connectivity. Although our sample size was small, our findings
+provided evidence that high-fat diets were less connected than low-fat
+diets and that high-fat diets therefore may lead to less stable
 communities with a decreased ability for microbes to directly influence
@@
 Second, although our analyses utilized the best datasets currently
-available for our study, this work was done retrospectively and relied
-on existing data up to seven years old. These archived datasets were
+avilable for our study, this work was done retrospectively and relied on
+existing data up to seven years old. These archived datasets were
 limited by the technology and costs of the time. For example, the diet
@@
 generation protocols. It is now known that MDA results in biases in
-microbial community composition {[}62{]}, as well as toward ssDNA viral
-genomes {[}63,64{]}, thus rendering the resulting microbial and viral
+microbial community composition {[}58{]}, as well as toward ssDNA viral
+genomes {[}59,60{]}, thus rendering the resulting microbial and viral
 metagenomes largely non-quantitative. Future work that employs larger
@@
 units (OGUs), which represented groups of highly similar bacteria or
-phage genomes or clustered genome fragments. Similar clustering
-definition and validation methods, both computational and experimental,
-have been implemented in other metagenomic sequencing studies, as well
-{[}42,65--67{]}. These approaches could offer yet another level of
-sophistication to our network-based analyses. While this operationally
-defined clustering approach allows us to study whole community networks,
-our ability to make conclusions about interactions among specific phage
-or bacterial species or populations is inherently limited, compared to
-more focused, culture-based studies such as the work by Malki \emph{et
-al} {[}68{]}. Future work must address this limitation, e.g., through
+phage genomes or genome fragments as clustered sub-populations. Similar
+clustering definition and validation methods, both computational and
+experimental, have been implemented in other metagenomic sequencing
+studies, as well {[}42,61--63{]}. These approaches could offer yet
+another level of sophistication to our network-based analyses. While
+this operationally defined clustering approach allows us to study whole
+community networks, our ability to make conclusions about interactions
+among specific phage or bacterial species or populations is inherently
+limited. Future work must address this limitation, e.g., through
 improved binning methods and deeper metagenomic shotgun sequencing, but
@@
 ecologically and evolutionarily cohesive units for both phage and
-bacteria {[}69{]}. Defining operational genomic units and their
+bacteria {[}64{]}. Defining operational genomic units and their
 taxonomic underpinnings (e.g., whether OGU clusters represent genera or
@@
 gene content, average nucleotide identity of shared genes, and pairwise
-differences between genomes {[}70{]}. Such population-genetic assessment
+differences between genomes {[}65{]}. Such population-genetic assessment
 of phage evolution, coupled with the ecological implications of genome
 heterogeneity, will inform how to define nodes in future iterations of
-the ecological network developed here. Even though we are hesitant to
-speculate on phage host ranges at low taxonomic levels in our dataset,
-the data does aggree with previous reports of instances of broad phage
-host range {[}68,71{]}.
+the ecological network developed here.
 
@@
 consists of archaea and small eukaryotes, including fungi and
-\emph{Demodex} mites {[}1,72{]}--all of which can interact with human
-immune cells and other non-microbial community members {[}73{]}. Future
+\emph{Demodex} mites {[}1,66{]}--all of which can interact with human
+immune cells and other non-microbial community members {[}67{]}. Future
 work will build from our approach and include these additional community
@@
 
-https://github.com/SchlossLab/Hannigan\_ConjunctisViribus\_ploscompbio\_2017
+https://github.com/SchlossLab/Hannigan\_ConjunctisViribus\_mSystems\_2017
 
@@
 were quality trimmed using the Fastx toolkit (v0.0.14) to exclude bases
-with quality scores below 33 and shorter than 75 bp {[}74{]}. Paired end
+with quality scores below 33 and shorter than 75 bp {[}68{]}. Paired end
 reads were filtered to exclude sequences missing their corresponding
@@
 Contigs were assembled using the Megahit assembly program (v1.0.6)
-{[}75{]}. A minimum contig length of 1 kb was used. Iterative k-mer
+{[}69{]}. A minimum contig length of 1 kb was used. Iterative k-mer
 stepping began at a minimum length of 21 and progressed by 20 until 101.
@@
 aligned to phage or bacterial contigs using the Bowtie2 global aligner
-(v2.2.1) {[}76{]}. We defined a mismatch threshold of 1 bp and seed
+(v2.2.1) {[}70{]}. We defined a mismatch threshold of 1 bp and seed
 length of 25 bp. Sequence abundance was calculated from the Bowtie2
@@
 taxonomic units (OTUs), although here we are clustering contigs into
-operational genomic units (OGUs) {[}61{]}.
-
-Contigs were clustered using the CONCOCT algorithm (v0.4.0) {[}77{]}.
+operational genomic units (OGUs) {[}57{]}.
+
+Contigs were clustered using the CONCOCT algorithm (v0.4.0) {[}71{]}.
 Because of our large dataset and limits in computational efficiency, we
@@
 
-\subsection{Operational Genomic Unit
-Identification}\label{operational-genomic-unit-identification}
+\subsection{Phage OGU Identification}\label{phage-ogu-identification}
 
@@
 sequences using an e-value \textless{} \(10^{-25}\), which was stricter
-than the \(10^{-10}\) threshold used in the random forest model below,
-against all of the phage reference genomes available in the EMBL
-database. We used a stricter threshold because we know there are genomic
+than the \(10^{-10}\) threshold used in the random forest model below.
+We used a stricter threshold because we know there are genomic
 similarities between bacteria and phage OGUs from the interactive model,
 but we were interested in contigs with high enough similarity to
-references that they may indeed be from phages. We also performed the
-converse analysis of aligning phage OGU representative sequences to EMBL
-bacterial reference genomes. Finally, we ran both the phage and bacteria
-OGU representative sequences through the Virsorter program (1.0.3) to
-identify phages (all default parameters were used), using only those in
-the high confidence identification category ``class 1'' {[}47{]}.
+references that they may indeed be from phages. 2\% of the OGUs had
+nucleotide similarities to known bacteriophage genomes, although the
+alignments were short (a couple of kb) and represented a small fraction
+of the alignment sequences.
 
@@
 Open reading frames (ORFs) were identified using the Prodigal program
-(V2.6.2) with the meta mode parameter and default settings {[}78{]}.
+(V2.6.2) with the meta mode parameter and default settings {[}72{]}.
 
@@
 experimentally validated bacteria-phage infections or validated lack of
-infections from six studies {[}41,48--52{]}. Associated reference
+infections from six studies {[}41,47--51{]}. Associated reference
 genomes were downloaded from the European Bioinformatics Institute (see
@@
 interactions or not. The classification model was built using the Caret
-R package (v6.0.73) {[}79{]}. The model was trained using five-fold
-cross validation with ten repeats, and the median model performance was
-evaluated by training the model on 80\% of the dataset and testing
-performance on the remaining 20\%. Pairs without scores were classified
+R package (v6.0.73) {[}73{]}. The model was trained using five-fold
+cross validation with ten repeats. Pairs without scores were classified
 as not interacting. The model was optimized using the ROC value. The
@@
 identified from bacterial genomes using the PilerCR program (v1.06)
-{[}80{]}. Resulting spacer sequences were filtered to exclude spacers
+{[}74{]}. Resulting spacer sequences were filtered to exclude spacers
 shorter than 20 bp and longer than 65 bp. Spacer sequences were aligned
 to the phage genomes using the nucleotide BLAST algorithm with default
-parameters (v2.4.0) {[}81{]}. The mean percent identity for each
+parameters (v2.4.0) {[}75{]}. The mean percent identity for each
 matching pair was recorded for use in our classification model.
@@
 between the genes using the Diamond protein alignment algorithm
-(v0.7.11.60) {[}82{]}. The mean alignment bitscores for each genome pair
+(v0.7.11.60) {[}76{]}. The mean alignment bitscores for each genome pair
 were recorded for use in our classification model.
@@
 predicted to interact using the Pfam interaction information within the
-Intact database {[}53{]}. The mean bitscores of the matches between each
+Intact database {[}52{]}. The mean bitscores of the matches between each
 pair were recorded for use in the classification model.
@@
 associated metadata were used to populate a graph database using Neo4j
-graph database software (v2.3.1) {[}83{]}. This network was used for
+graph database software (v2.3.1) {[}77{]}. This network was used for
 downstream community analysis.
@@
 Degree and closeness centrality were calculated using the associated
-functions within the igraph R package (v1.0.1) {[}84{]}.
+functions within the igraph R package (v1.0.1) {[}78{]}.
 
@@
 We thank the members of the Schloss lab for their underlying
-contributions. We thank the authors of the original studies for making
-their data and metadata publicly available and understandable. We also
-thank the participants in the studies.
-
-\section{Author Contributions}\label{author-contributions}
-
-\emph{Conceptualization}: GDH, MBD, DK, PDS. \emph{Data Curation}: GDH.
-\emph{Formal Analysis}: GDH. \emph{Funding Acquisition}: GDH, PDS.
-\emph{Writing -- Original Draft Preparation}: GDH, PDS. \emph{Writing --
-Review \& Editing}: GDH, MBD, DK, PDS.
+contributions.
 
@@
 
-\section{Competing interests}\label{competing-interests}
+\section{Disclosure Declaration}\label{disclosure-declaration}
 
@@
 
-\section{Supplemental Figure
-Captions}\label{supplemental-figure-captions}
-
-\beginsupplement
+\section{Figures}\label{figures}
 
@@
 \centering
-\includegraphics[width=0.80000\textwidth]{}
+\includegraphics[width=0.85000\textwidth]{../figures/rocCurves.pdf}
+\caption{\textbf{Summary of Multi-Study Network Model.} \emph{(A)
+Average ROC curve used to create the microbiome-virome infection
+prediction model. (B) Importance scores associated with the metrics used
+in the random forest model to predict relationships between bacteria and
+phages. The importance score is defined as the mean decrease in accuracy
+of the model when a feature (e.g.~Pfam) is excluded. (C) Proportions of
+samples included (gray) and excluded (red) in the model. Samples were
+excluded from the model because they did not yield any scores. Those
+interactions without scores were defined as not having interactions. (D)
+Bipartite visualization of the resulting phage-bacteria network. This
+network includes information from all three published studies. (E)
+Network diameter (measure of graph size; the greatest number of
+traversed vertices required between two vertices), (F) number of
+vertices, and (G) number of edges (relationships) for the total network
+(yellow) and the individual study sub-networks (diet study = red, skin
+study = green, twin study = orange).} \label{RocCurve}}
+\end{figure}
+
+\newpage
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.90000\textwidth]{../figures/dietnetworks.pdf}
+\caption{\textbf{Impact of Diet and Obesity on Gut Network Structure.}
+\emph{(A) Quantification of average degree centrality (number of edges
+per node) and (B) closeness centrality (average distance from each node
+to every other node) of gut microbiome networks of subjects limited to
+exclusively high-fat or low-fat diets. Lines represent the mean degree
+of centrality for each diet. (C) Quantification of average degree
+centrality and (D) closeness centrality between obese and healthy adult
+women.}\label{dietnetworks}}
+\end{figure}
+
+\newpage
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.90000\textwidth]{../figures/intrapersonal_diversity.pdf}
+\caption{\textbf{Intrapersonal vs Interpersonal Network Dissimilarity
+Across Different Human Systems.} \emph{(A) NMDS ordination illustrating
+network dissimilarity between subjects over time. Each sample is colored
+by subject, with each sample pair collected 8-10 days apart.
+Dissimilarity was calculated using the Bray-Curtis metric based on
+abundance weighted eigenvector centrality signatures, with a greater
+distance representing greater dissimilarity in bacteria and phage
+centrality and abundance. (B) Quantification of gut network
+dissimilarity within the same subject over time (intrapersonal) and the
+mean dissimilarity between the subject of interest and all other
+subjects (interpersonal). The p-value is also provided. (C)
+Quantification of gut network dissimilarity within subjects from the
+same family (intrafamily) and the mean dissimilarity between subjects
+within a family and those of other families (interfamily). The p-value
+is also provided. (D) Quantification of skin network dissimilarity
+within the same subject and anatomical location over time
+(intrapersonal) and the mean dissimilarity between the subject of
+interest and all other subjects at the same time and the same anatomical
+location (interpersonal). P-value was calculated using a paired Wilcoxon
+test.}\label{intradiv}}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.75000\textwidth]{../figures/skinplotresults.pdf}
+\caption{\textbf{Impact of Skin Micro-Environment on Microbiome Network
+Structure.} \emph{(A) Notched box-plot depicting differences in average
+eigenvector centrality between moist, intermittently moist, and
+sebaceous skin sites and (B) occluded, intermittently occluded, and
+exposed sites. Notched box-plots were created using ggplot2 and show the
+median (center line), the inter-quartile range (IQR; upper and lower
+boxes), the highest and lowest value within 1.5 * IQR (whiskers),
+outliers (dots), and the notch which provides an approximate 95\%
+confidence interval as defined by 1.58 * IQR / sqrt(n). (C) NMDS
+ordination depicting the differences in skin microbiome network
+structure between skin moisture levels and (D) occlusion. Samples are
+colored by their environment and their dissimilarity to other samples
+was calculated as described in figure \ref{intradiv}. (E) The
+statistical differences of networks between moisture and (F) occlusion
+status were quantified with an anova and post hoc Tukey test. Cluster
+centroids are represented by dots and the extended lines represent the
+associated 95\% confidence intervals. Significant comparisons (p-value
+\textless{} 0.05) are colored in red, and non-significant comparisons
+are gray.}\label{skinnetwork}}
+\end{figure}
+
+\newpage
+
+\section{Supplemental Figures}\label{supplemental-figures}
+
+\beginsupplement
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.80000\textwidth]{../figures/SequenceAbund.pdf}
 \caption{\textbf{Sequencing Depth Summary.} \emph{Number of sequences
@@
 \centering
-\includegraphics[width=0.50000\textwidth]{}
+\includegraphics[width=0.50000\textwidth]{../figures/ContigStats.pdf}
 \caption{\textbf{Contig Summary Statistics.} \emph{Scatter plot heat map
@@
 \centering
-\includegraphics[width=0.80000\textwidth]{}
+\includegraphics[width=0.80000\textwidth]{../figures/ClusterStats.pdf}
 \caption{\textbf{Operational Genomic Unit Summary Statistics.}
@@
 \centering
-\includegraphics[width=1.00000\textwidth]{}
+\includegraphics[width=1.00000\textwidth]{../figures/BenchmarkDataset.pdf}
 \caption{\textbf{Summary information of validation dataset used in the
@@
 \centering
-\includegraphics[width=0.50000\textwidth]{}
+\includegraphics[width=0.50000\textwidth]{../figures/graphdatabasediagram.pdf}
 \caption{\textbf{Structure of the interactive network.} \emph{Metadata
@@
 \centering
-\includegraphics[width=0.75000\textwidth]{}
+\includegraphics[width=0.75000\textwidth]{../figures/intraallskin.pdf}
 \caption{\textbf{Intrapersonal vs Interpersonal Dissimilarity of the
@@
 
-\section{Supplemental Table Captions}\label{supplemental-table-captions}
-
-Table S1: Summary of the primary quality control measures reported in
-the original publications of the viromes used in this current study.
-
-\newpage
-
 \section*{References}\label{references}
@@
 
-\hypertarget{ref-Roux:2015dh}{}
-47. Roux S, Enault F, Hurwitz BL, Sullivan MB. VirSorter: mining viral
-signal from microbial genomic data. PeerJ. 2015;3: e985--20.
-
 \hypertarget{ref-Jensen:1998vh}{}
-48. Jensen EC, Schrader HS, Rieland B, Thompson TL, Lee KW, Nickerson
+47. Jensen EC, Schrader HS, Rieland B, Thompson TL, Lee KW, Nickerson
 KW, et al. Prevalence of broad-host-range lytic bacteriophages of
@@
 \hypertarget{ref-Malki:2015tm}{}
-49. Malki K, Kula A, Bruder K, Sible E. Bacteriophages isolated from
+48. Malki K, Kula A, Bruder K, Sible E. Bacteriophages isolated from
 Lake Michigan demonstrate broad host-range across several bacterial
@@
 \hypertarget{ref-Schwarzer:2012ez}{}
-50. Schwarzer D, Buettner FFR, Browning C, Nazarov S, Rabsch W, Bethe A,
+49. Schwarzer D, Buettner FFR, Browning C, Nazarov S, Rabsch W, Bethe A,
 et al. A multivalent adsorption apparatus explains the broad host range
@@
 \hypertarget{ref-Kim:2012dh}{}
-51. Kim S, Rahman M, Seol SY, Yoon SS, Kim J. Pseudomonas aeruginosa
+50. Kim S, Rahman M, Seol SY, Yoon SS, Kim J. Pseudomonas aeruginosa
 bacteriophage PA1Ø requires type IV pili for infection and shows broad
@@
 \hypertarget{ref-Matsuzaki:1992gw}{}
-52. Matsuzaki S, Tanaka S, Koga T, Kawata T. A Broad-Host-Range
+51. Matsuzaki S, Tanaka S, Koga T, Kawata T. A Broad-Host-Range
 Vibriophage, KVP40, Isolated from Sea Water. Microbiology and
@@
 \hypertarget{ref-Orchard:2014hq}{}
-53. Orchard S, Ammari M, Aranda B, Breuza L, Briganti L, Broackes-Carter
+52. Orchard S, Ammari M, Aranda B, Breuza L, Briganti L, Broackes-Carter
 F, et al. The MIntAct project--IntAct as a common curation platform for
@@
 
-\hypertarget{ref-Kleiner:2015kd}{}
-54. Kleiner M, Hooper LV, Duerkop BA. Evaluation of methods to purify
-virus-like particles for metagenomic sequencing of intestinal viromes.
-BMC Genomics. 2015;16: 7.
-
-\hypertarget{ref-Meisel:2016iu}{}
-55. Meisel JS, Hannigan GD, Tyldsley AS, SanMiguel AJ, Hodkinson BP,
-Zheng Q, et al. Skin microbiome surveys are strongly influenced by
-experimental design. Journal of Investigative Dermatology. 2016;
-
 \hypertarget{ref-Turnbaugh:2009hf}{}
-56. Turnbaugh PJ, Ridaura VK, Faith JJ, Rey FE, Knight R, Gordon JI. The
+53. Turnbaugh PJ, Ridaura VK, Faith JJ, Rey FE, Knight R, Gordon JI. The
 effect of diet on the human gut microbiome: a metagenomic analysis in
@@
 \hypertarget{ref-David:2014cl}{}
-57. David LA, Maurice CF, Carmody RN, Gootenberg DB, Button JE, Wolfe
+54. David LA, Maurice CF, Carmody RN, Gootenberg DB, Button JE, Wolfe
 BE, et al. Diet rapidly and reproducibly alters the human gut
@@
 
-\hypertarget{ref-Jeong:2000im}{}
-58. Jeong H, Tombor B, Albert R, Oltvai ZN, Barabási AL. The large-scale
-organization of metabolic networks. Nature. 2000;407: 651--654.
-
 \hypertarget{ref-Grice:2009ee}{}
-59. Grice EA, Kong HH, Conlan S, Deming CB, Davis J, Young AC, et al.
+55. Grice EA, Kong HH, Conlan S, Deming CB, Davis J, Young AC, et al.
 Topographical and Temporal Diversity of the Human Skin Microbiome.
@@
 \hypertarget{ref-Minot:2013ih}{}
-60. Minot S, Bryson A, Chehoud C, Wu GD, Lewis JD, Bushman FD. Rapid
+56. Minot S, Bryson A, Chehoud C, Wu GD, Lewis JD, Bushman FD. Rapid
 evolution of the human gut virome. Proceedings of the National Academy
@@
 \hypertarget{ref-Schloss:2005hz}{}
-61. Schloss PD, Handelsman J. Introducing DOTUR, a computer program for
+57. Schloss PD, Handelsman J. Introducing DOTUR, a computer program for
 defining operational taxonomic units and estimating species richness.
@@
 \hypertarget{ref-Yilmaz:2010jb}{}
-62. Yilmaz S, Allgaier M, Hugenholtz P. Multiple displacement
+58. Yilmaz S, Allgaier M, Hugenholtz P. Multiple displacement
 amplification compromises quantitative analysis of metagenomes. Nature
@@
 \hypertarget{ref-Kim:2008to}{}
-63. Kim KH, Chang HW, Nam YD, Roh SW. Amplification of uncultured
+59. Kim KH, Chang HW, Nam YD, Roh SW. Amplification of uncultured
 single-stranded DNA viruses from rice paddy soil. Applied and \ldots{}.
@@
 \hypertarget{ref-Kim:2011hp}{}
-64. Kim K-H, Bae J-W. Amplification methods bias metagenomic libraries
+60. Kim K-H, Bae J-W. Amplification methods bias metagenomic libraries
 of uncultured single-stranded and double-stranded DNA viruses. Applied
@@
 \hypertarget{ref-Minot:2012ed}{}
-65. Minot S, Wu GD, Lewis JD, Bushman FD. Conservation of gene cassettes
+61. Minot S, Wu GD, Lewis JD, Bushman FD. Conservation of gene cassettes
 among diverse viruses of the human gut. PLOS ONE. 2012;7: e42342.
@@
 \hypertarget{ref-Deng:2014eb}{}
-66. Deng L, Ignacio-Espinoza JC, Gregory AC, Poulos BT, Weitz JS,
+62. Deng L, Ignacio-Espinoza JC, Gregory AC, Poulos BT, Weitz JS,
 Hugenholtz P, et al. Viral tagging reveals discrete populations in
@@
 \hypertarget{ref-Brum:2015iaa}{}
-67. Brum JR, Ignacio-Espinoza JC, Roux S, Doulcier G, Acinas SG, Alberti
+63. Brum JR, Ignacio-Espinoza JC, Roux S, Doulcier G, Acinas SG, Alberti
 A, et al. Ocean plankton. Patterns and ecological drivers of ocean viral
@@
 
-\hypertarget{ref-Malki:2015dw}{}
-68. Malki K, Kula A, Bruder K, Sible E, Hatzopoulos T, Steidel S, et al.
-Bacteriophages isolated from Lake Michigan demonstrate broad host-range
-across several bacterial phyla. Virology Journal. 2015;12: 164.
-
 \hypertarget{ref-Polz:2006fi}{}
-69. Polz MF, Hunt DE, Preheim SP, Weinreich DM. Patterns and mechanisms
+64. Polz MF, Hunt DE, Preheim SP, Weinreich DM. Patterns and mechanisms
 of genetic and phenotypic differentiation in marine microbes.
@@
 \hypertarget{ref-Gregory:2016cg}{}
-70. Gregory AC, Solonenko SA, Ignacio-Espinoza JC, LaButti K, Copeland
+65. Gregory AC, Solonenko SA, Ignacio-Espinoza JC, LaButti K, Copeland
 A, Sudek S, et al. Genomic differentiation among wild cyanophages
@@
 
-\hypertarget{ref-PaezEspino:2016gi}{}
-71. Paez-Espino D, Eloe-Fadrosh EA, Pavlopoulos GA, Thomas AD, Huntemann
-M, Mikhailova N, et al. Uncovering Earth's virome. Nature. 2016;
-
 \hypertarget{ref-Grice:2011gy}{}
-72. Grice EA, Segre JA. The skin microbiome. Nature Reviews
+66. Grice EA, Segre JA. The skin microbiome. Nature Reviews
 Microbiology. 2011;9: 244--253.
@@
 \hypertarget{ref-Round:2009bz}{}
-73. Round JL, Mazmanian SK. The gut microbiota shapes intestinal immune
+67. Round JL, Mazmanian SK. The gut microbiota shapes intestinal immune
 responses during health and disease. Nature reviews Immunology. 2009;9:
@@
 \hypertarget{ref-FASTXToolkit:wr}{}
-74. Hannon GJ. FASTX-Toolkit.: GNU Affero General Public License.
+68. Hannon GJ. FASTX-Toolkit.: GNU Affero General Public License.
 
 \hypertarget{ref-Li:2016kd}{}
-75. Li D, Luo R, Liu C-M, Leung C-M, Ting H-F, Sadakane K, et al.
+69. Li D, Luo R, Liu C-M, Leung C-M, Ting H-F, Sadakane K, et al.
 MEGAHIT v1.0: A fast and scalable metagenome assembler driven by
@@
 \hypertarget{ref-Langmead:2012jh}{}
-76. Langmead B, Salzberg SL. Fast gapped-read alignment with Bowtie 2.
+70. Langmead B, Salzberg SL. Fast gapped-read alignment with Bowtie 2.
 Nature Methods. 2012;9: 357--359.
@@
 \hypertarget{ref-Alneberg:2014fc}{}
-77. Alneberg J, Bjarnason BS aacute ri, Bruijn I de, Schirmer M, Quick
+71. Alneberg J, Bjarnason BS aacute ri, Bruijn I de, Schirmer M, Quick
 J, Ijaz UZ, et al. Binning metagenomic contigs by coverage and
@@
 \hypertarget{ref-Hyatt:2012cy}{}
-78. Hyatt D, LoCascio PF, Hauser LJ, Uberbacher EC. Gene and translation
+72. Hyatt D, LoCascio PF, Hauser LJ, Uberbacher EC. Gene and translation
 initiation site prediction in metagenomic sequences. Bioinformatics.
@@
 \hypertarget{ref-caretClassificatio:ux5fU2Litux5f1}{}
-79. Kuhn M. caret: Classification and Regression Training.
+73. Kuhn M. caret: Classification and Regression Training.
 
 \hypertarget{ref-Edgar:2007bh}{}
-80. Edgar RC. PILER-CR: fast and accurate identification of CRISPR
+74. Edgar RC. PILER-CR: fast and accurate identification of CRISPR
 repeats. BMC Bioinformatics. 2007;8: 18.
@@
 \hypertarget{ref-Camacho:2009fc}{}
-81. Camacho C, Coulouris G, Avagyan V, Ma N, Papadopoulos J, Bealer K,
+75. Camacho C, Coulouris G, Avagyan V, Ma N, Papadopoulos J, Bealer K,
 et al. BLAST+: architecture and applications. BMC Bioinformatics.
@@
 \hypertarget{ref-Buchfink:2015ki}{}
-82. Buchfink B, Xie C, Huson DH. Fast and sensitive protein alignment
+76. Buchfink B, Xie C, Huson DH. Fast and sensitive protein alignment
 using DIAMOND. Nature Methods. 2015;12: 59--60.
@@
 \hypertarget{ref-Neoj:Fuwr6PBN}{}
-83. Neo4j.
+77. Neo4j.
 
 \hypertarget{ref-Theigraphsoftware:vh}{}
-84. Csardi G, Nepusz T. The igraph software package for complex network
+78. Csardi G, Nepusz T. The igraph software package for complex network
 research.

### 2017-12-28 doc/manuscript.Rmd
Updated manuscript with revisions.

@@
 ---
-title: Biogeography & environmental conditions shape bacteriophage & bacteria interaction networks across the human microbiome
+title: Biogeography & environmental conditions shape bacteriophage-bacteria networks across the human microbiome
 author:
@@
 Viruses and bacteria are critical components of the human microbiome and play important roles in health and disease.
-Most previous work has relied on studying microbes and viruses independently, thereby reducing them to two separate communities.
-Such approaches are unable to capture how these microbial communities interact, such as through processes that maintain community stability or allow phage-host populations to co-evolve.
-We developed and implemented a network-based analytical approach to describe phage-bacteria network diversity throughout the human body.
-We accomplished this by building a machine learning algorithm to predict which phages could infect which bacteria in a given microbiome.
-This algorithm was applied to paired viral and bacterial metagenomic sequence sets from three previously published human cohorts.
+Most previous work has relied on studying bacteria and viruses independently, thereby reducing them to two separate communities.
+Such approaches are unable to capture how these microbial communities interact, such as through processes that maintain community robustness or allow phage-host populations to co-evolve.
+We implemented a network-based analytical approach to describe phage-bacteria network diversity throughout the human body.
+We built these community networks using a machine learning algorithm to predict which phages could infect which bacteria in a given microbiome.
+Our algorithm was applied to paired viral and bacterial metagenomic sequence sets from three previously published human cohorts.
 We organized the predicted interactions into networks that allowed us to evaluate phage-bacteria connectedness across the human body.
-We found that gut and skin network structures were person-specific and not conserved among cohabitating family members.
-High-fat diets and obesity appeared to be associated with less connected networks.
-Network structure differed between skin sites, with those exposed to the external environment being less connected and more prone to instability.
+We observed evidence that gut and skin network structures were person-specific and not conserved among cohabitating family members.
+High-fat diets appeared to be associated with less connected networks.
+Network structure differed between skin sites, with those exposed to the external environment being less connected and likely more susceptible to network degradation by microbial extinction events.
 This study quantified and contrasted the diversity of virome-microbiome networks across the human body and illustrated how environmental factors may influence phage-bacteria interactive dynamics.
@@
 These communities have primarily been studied separately using metrics of community composition and diversity.
-These approaches have failed to capture the complex dynamics of interacting bacteria and phage communities, which frequently share genetic information and work together to maintain stable ecosystems.
+These approaches have failed to capture the complex dynamics of interacting bacteria and phage communities, which frequently share genetic information and work together to maintain ecosystem homestatsis (e.g.
+kill-the-winner dynamics).
 Removal of bacteria or phage can disrupt or even collapse those ecosystems.
@@
 Previous human microbiome work has focused on bacterial and viral communities, but have reduced them to two separate communities by studying them independently  [@Norman:2015kb; @Manrique:2016dx; @Ly:2014ew; @Monaco:2016ita; @Hannigan:2015fz; @Minot:2011ez; @SantiagoRodriguez:2015gd; @Abeles:2015dy; @Abeles:2014kj].
-This approach fails to capture the complex dynamics of interacting bacteria and phage communities, which frequently share genetic information and work together to maintain stable ecosystems.
+This approach fails to capture the complex dynamics of interacting bacteria and phage communities, which frequently share genetic information and work together to maintain ecosystem structure (e.g.
+kill-the-winner dynamics that prevent domination by a single bacterium).
 Removal of bacteria or phages can disrupt or even collapse those ecosystems [@Haerter:2014ii; @Harcombe:2005fd; @Middelboe:2001fl; @Poisot:2011jc; @Thompson:2012ki; @Moebus:1981kp; @Flores:2013hc; @Poisot:2012fh; @Poisot071993; @Flores:2011bh; @Jover:2015ev].
-Relationship-based network approaches allow us to capture this interaction information.
-Studying such bacteria-phage interactions through community-wide networks built from inferred relationships could offer further insights into the drivers of human microbiome diversity across body sites and enable the study of human microbiome network dynamics overall.
+Integrating these datasets as relationship-based networks allow us to capture this complex interaction information.
+Studying such bacteria-phage interactions through community-wide networks built from inferred relationships begins to provide us with insights into the drivers of human microbiome diversity across body sites and enable the study of human microbiome network dynamics overall.
 
@@
 This approach built upon previous large-scale phage-bacteria network analyses by inferring interactions from metagenomic datasets, rather than culture-dependent data [@Flores:2013hc], which is limited in the scale of possible experiments and analyses.
-Our metagenomic interaction inference model improved upon previous models of phage-host predictions that have utilized a variety of techniques, such as linear models to predict bacteria-phage co-occurrence using taxonomic assignments [@LimaMendez:2015hw], and nucleotide similarity models that were applied to both whole virus genomes [@Edwards:2015iz] and related clusters of whole and partial virus genomes [@@Roux:2016cc].
+We implemented an adapted metagenomic interaction inference model that made some improvements upon previous phage-host interaction prediction models.
+Previous approaches have utilized a variety of techniques, such as linear models that were used to predict bacteria-phage co-occurrence using taxonomic assignments [@LimaMendez:2015hw], and nucleotide similarity models that were applied to both whole virus genomes [@Edwards:2015iz] and clusters of whole and partial virus genomes [@@Roux:2016cc].
 Our approach uniquely included protein interaction data and was validated based on experimentally determined positive and negative interactions (i.e.
 who does and does not infect whom).
-Through this approach we were able to provide a basic understanding of the network dynamics associated with phage and bacterial communities on and in the human body.
-By building and utilizing a microbiome network, we found that different people, body sites, and anatomical locations not only support distinct microbiome membership and diversity [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei; @Grice:2009eea; @Findley:2013jf; @Costello:2009im, @Consortium:2012iz], but also support ecological communities with distinct communication structures and propensities toward community instability.
-Through an improved understanding of network structures across the human body, we empower future studies to investigate how these communities dynamics are influenced by disease states and the overall impact they may have on human health.
+We built on previous modeling work as a means to our ends, and focused on the biological insights we could gain instead of building a superior model and presenting our work as a toolkit.
+We therefore did not focus on extensive benchmarking against other existing models [@Edwards:2015iz; @Villarroel:2016jy; @Galiez:2017gm; @Ahlgren:2017gh].
+Through this approach we were able to provide an initial basic understanding of the network dynamics associated with phage and bacterial communities on and in the human body.
+By building and utilizing a microbiome network, we found that different people, body sites, and anatomical locations not only support distinct microbiome membership and diversity [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei; @Grice:2009eea; @Findley:2013jf; @Costello:2009im, @Consortium:2012iz], but also support ecological communities with distinct communication structures and robustness to network degradation by extinction events.
+Through an improved understanding of network structures across the human body, we aim to empower future studies to investigate how these communities dynamics are influenced by disease states and the overall impact they may have on human health.
 
@@
 We further grouped the related bacterial and phage contigs into operationally defined units based on their k-mer frequencies and co-abundance patterns, similar to previous reports **(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})** [@Roux:2016cc].
+This was done both for dimensionality reduction and to prevent inflation of node counts by using contigs which are expected to represent multiple fragments from the same genomes.
+This was also done to create genome analogs that we could use in our classification model which was built using genome sequences.
 We referred to these operationally defined groups of related contigs as operational genomic units (OGUs).
@@
 
-While supplementing the previous quality control measures **(Table S1)** we found that, in light of the rigorous purification and quality control during sample collection and preparation, 77% (228 / 298 operational genomic units) still had some nucleotide similarity to a given bacterial reference genome (e-value < $10^{-25}$).
-As absence of bacterial contamination had been confirmed by sensitive molecular methods **(Table S1)**, we interpreted this as evidence that the majority of the gut and skin bacteriophages were temperate and thereby shared elements with bacterial reference genomes that contained similar integrated phages when sequenced--a trend previously reported [@Minot:2011ez, @Hannigan:2015fz, @Reyes:2010cwa].
-Additionally, we identified two OGUs as being complete phages using the stringent Virsorter phage identification algorithm (class 1 confidence group) [@Roux:2015dh].
+While supplementing the previous virome fraction quality control measures **(Table S1)** we found that, in light of the rigorous purification and quality control during sample collection and preparation, 77% (228 / 298 operational genomic units) still had some nucleotide similarity to a given bacterial reference genome (e-value < $10^{-25}$).
+As most phages in these communities have been shown to be temperate (they integrate into bacterial genomes) using methods that included nucleotide alignments of phages to bacterial reference genomes, we interpreted this as confirmation that the majority of those phages were temperate and therefore shared elements with bacterial reference genomes -- a trend previously reported [@Minot:2011ez, @Hannigan:2015fz, @Reyes:2010cwa].
+We further confirmed that while the majority of these were expected to be temperate phages, there still remained a low level of bacterial sequence noise, which was evident as five (1.7%) OGUs with similar sequence elements to the bacterial 16S rRNA gene (blastn, e-value < $10^{-25}$, length > 1,000 bp).
+This is in line with previous work which has suggested that bacterial noise is an inevitable technical issue, and we considered this noise while interpreting our findings [@Roux:2013cr].
+We also identified two OGUs as being complete, high confidence phages using the stringent Virsorter phage identification algorithm (class 1 confidence group) [@Roux:2015dh].
 
-The whole metagenomic shotgun sequence samples primarily consisted of bacteria, with an average viral relative abundance of 0.4% **(Table S1)** [@Minot:2011ez; @Hannigan:2015fz; @Reyes:2010cwa; @Turnbaugh:2009ei].
+The whole metagenomic shotgun sequence samples, which primarily consisted of bacteria, had an average viral relative abundance of 0.4% **(Table S1)** [@Minot:2011ez; @Hannigan:2015fz; @Reyes:2010cwa; @Turnbaugh:2009ei].
 We found that only 2% (6 / 280 OGUs) of bacterial OGUs had significantly strong nucleotide similarity to phage reference genomes (e-value < $10^{-25}$) [@Minot:2011ez; @Hannigan:2015fz; @Reyes:2010cwa; @Turnbaugh:2009ei].
 No OGUs were confidently identified as lytic or temperate phage OGUs in the bacterial dataset using the Virsorter algorithm [@Roux:2015dh].
-Together this suggests only minimal "contamination" of the bacterial OGUs.
+Together this suggests minimal bacterial OGU noise that should be considered in the study conclusions.
 
-## Evaluating the Model to Predict Phage-Bacteria Interactions
+## Implementing Phage-Bacteria Interaction Prediction to Build a Community Network
 We predicted which phage OGUs infected which bacterial OGUs using a random forest model trained on experimentally validated infectious relationships from six previous publications [@Jensen:1998vh;@Malki:2015tm;@Schwarzer:2012ez;@Kim:2012dh;@Matsuzaki:1992gw;@Edwards:2015iz].
-Only bacteria and phages were used in the model.
+Only bacteria and bacteriophages were used in the model.
 The training set contained 43 diverse bacterial species and 30 diverse phage strains, including both broad and specific ranges of infection **(Supplemental Figure \ref{ValidationOverview} A - B)**.
@@
 This allowed us to validate the false positive and false negative rates associated with our candidate models, thereby building upon previous work that only considered positive relationships [@Edwards:2015iz].
+It is also worth noting that while a positive interaction is strong evidence that the interaction exists, we must also be conscious that negative interactions are only weak evidence for a lack of interaction because the finding could be the result of our inability to reproduce conditions in which those interactions occur.
+Altogether we decided to maintain a balanced dataset at the cost of under-sampling the available positive interaction information because the use of such a severely unbalanced dataset often results in over-fit and uninformative model training.
 
-Four phage and bacterial genomic features were used in a random forest model to predict infectious relationships between bacteria and phages: 1) genome nucleotide similarities, 2) gene amino acid sequence similarities, 3) bacterial Clustered Regularly Interspaced Short Palindromic Repeat (CRISPR) spacer sequences that target phages, and 4) similarity of protein families associated with experimentally identified protein-protein interactions [@Orchard:2014hq].
+Four phage and bacterial genomic features were used in our random forest model to predict infectious relationships between bacteria and phages: 1) genome nucleotide similarities, 2) gene amino acid sequence similarities, 3) bacterial Clustered Regularly Interspaced Short Palindromic Repeat (CRISPR) spacer sequences that target phages, and 4) similarity of protein families associated with experimentally identified protein-protein interactions [@Orchard:2014hq].
+These features were calculated using the training set described above.
+We chose to utilize these metrics that directly compare nucleotide sequences between sample phages and bacteria, instead of relying on alignment to reference genomes or known marker genes, because we are extrapolating our model to highly diverse communities which we expect to diverge significantly from the available reference genomes.
 The resulting random forest model was assessed using nested cross validation, and the median area under its receiver operating characteristic (ROC) curve was `r signif(averageauc[c(averageauc$type %in% "AUC"), 2], digits = sigfig)`, the median model sensitivity was `r signif(averageauc[c(averageauc$type %in% "SENS"), 2], digits = sigfig)`, and median specificity was `r signif(averageauc[c(averageauc$type %in% "SPEC"), 2], digits = sigfig)` **(Figure \ref{RocCurve} A)**.
+This balance of confident true positives at the cost of fewer true negatives is ideal for this type of dataset which consists of primarily positive connections **(Supplemental Figure \ref{nestedperformance})**.
+Nested cross validation of the model demonstrated that the sensitivity and specificity of the model could vary but the overall model performance (AUC) remained more consistent **(Supplemental Figure \ref{crossval})**.
+This suggested that our model would perform with a similar overall accuracy despite changes in sensitivity/specificity trade-offs.
 The most important predictor in the model was amino acid similarity between genes, followed by nucleotide similarity of whole genomes **(Figure \ref{RocCurve} B)**.
@@
 Those interactions without scores were automatically classified as not having interactions.
-(D) Bipartite visualization of the resulting phage-bacteria network.
-Phage OGUs are presented in orange, bacteria OGUs in red, and their interaction edges are represented as connecting lines.
-This network includes information from all three published studies.
-(E) Network diameter (measure of graph size; the greatest number of traversed vertices required between two vertices), (F) number of vertices, and (G) number of edges (relationships) for the total network (orange) and the individual study sub-networks (diet study = red, skin study = yellow, twin study = green).* \label{RocCurve}](){ width=85% }
+(D) Network diameter (measure of graph size; the greatest number of traversed vertices required between two vertices), (E) number of vertices, and (F) number of edges (relationships) for the total network (orange) and the individual study sub-networks (diet study = red, skin study = yellow, twin study = green).* \label{RocCurve}](){ width=85% }
 
 We used our random forest model to classify the relationships between bacteria and phage operational genomic units, which were then used to build the interactive network.
-The master network contained the three studies as sub-networks, which themselves each contained sub-networks for each sample **(Figure \ref{RocCurve} D)**.
+The master network, analogous to the universal microbiome network concept previously described [@Bashan:2016is], contained the three studies as sub-networks, which themselves each contained sub-networks for each sample **(Supplemental Figure \ref{NetworkDiagram})**.
 Metadata including study, sample ID, disease, and OGU abundance within the community were stored in the master network for parsing in downstream analyses **(Supplemental Figure \ref{NetworkDiagram})**.
+The phage and bacteria of the master network demonstrated both narrow broad ranges of infectious interactions **(Supplemental Figure \ref{inthm})**.
 Bacterial and phage relative abundance was recorded in each sample for each OGU and the weight of the edge connecting those OGUs was calculated as a function of those relative abundance values.
@@
 Like the master network, the skin network exhibited a diameter of `r prettyNum(genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Diameter"], big.mark = ",")` (measure of graph size; the greatest number of traversed vertices required between two vertices) and included `r prettyNum(genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Vertices"], big.mark = ",")` (`r prettyNum(sitestats[c(sitestats$class %in% "skin"),"phagenodes"], big.mark = ",")` phages, `r prettyNum(sitestats[c(sitestats$class %in% "skin"),"bacterianodes"], big.mark = ",")` bacteria, `r round(100 * genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Vertices"] / genfigurestats[c(genfigurestats$class %in% "Total"),"Vertices"], digits = 1)`% total) and `r prettyNum(genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Edges"], big.mark = ",")` (`r round(100 * genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Edges"] / genfigurestats[c(genfigurestats$class %in% "Total"),"Edges"], digits = 1)`%) of the master network nodes and edges, respectively **(Figure \ref{RocCurve} E - F)**.
+Additionally, the subnetworks demonstrated narrow ranges of eccentricity across their nodes **(Supplemental Figure \ref{eccplot})**.
 The phages and bacteria in the diet and twin sample sets were more sparsely related, with the diet study consisting of `r prettyNum(genfigurestats[c(genfigurestats$class %in% "DietStudy"),"Vertices"], big.mark = ",")` (`r prettyNum(sitestats[c(sitestats$class %in% "diet"),"phagenodes"], big.mark = ",")` phages, `r prettyNum(sitestats[c(sitestats$class %in% "diet"),"bacterianodes"], big.mark = ",")` bacteria) nodes and `r prettyNum(genfigurestats[c(genfigurestats$class %in% "DietStudy"),"Edges"], big.mark = ",")` relationships, and the twin study containing `r prettyNum(genfigurestats[c(genfigurestats$class %in% "TwinStudy"),"Vertices"], big.mark = ",")` (`r prettyNum(sitestats[c(sitestats$class %in% "twin"),"phagenodes"], big.mark = ",")` phages, `r prettyNum(sitestats[c(sitestats$class %in% "twin"),"bacterianodes"], big.mark = ",")` bacteria) nodes and `r prettyNum(genfigurestats[c(genfigurestats$class %in% "TwinStudy"),"Edges"], big.mark = ",")` relationships **(Figure \ref{RocCurve} E - F)**.
+As an interesting validation measure, we identified five (1.7%) examples of phage OGUs which similar genomic elements to the four previously described, broadly infectious phages isolated from Lake Michigan (tblastx; e-value < $10^{-25}$) [@Malki:2015dw].
 
-## Role of Diet & Obesity in Gut Microbiome Connectivity
+## Role of Diet on Gut Microbiome Connectivity
 
@@
 We evaluated the diet-associated differences in gut virome-microbiome network structure by quantifying how central each sample's network was on average.
-We accomplished this by utilizing two common centrality metrics: degree centrality and closeness centrality.
+We accomplished this by utilizing two common weighted centrality metrics: degree centrality and closeness centrality.
 Degree centrality, the simplest centrality metric, was defined as the number of connections each phage made with each bacterium.
@@
 A higher closeness centrality suggests that the effects of genetic information or altered abundance would be more impactful to all other microbes in the system.
-A network with higher average closeness centrality also indicates an overall greater degree of connections, which suggests a greater resilience against instability.
-This is because more highly connected networks are more stable as a result of pathway dependencies and the unlikelihood that a randomly removed bacteria or phage would cause major divisions across the network [@Jeong:2000im; @Poisot:2011jc].
+Because these are weighted metrics, the values are functions of both connectivity as well as community composition.
+A network with higher average closeness centrality also indicates an overall greater degree of connections, which suggests a greater resilience against network degradation by extinction events [@Jeong:2000ima; @Poisot:2011jc].
+This is because more highly connected networks are less likely to degrade into multiple smaller networks when bacteria or phages are randomly removed  [@Jeong:2000ima; @Poisot:2011jc].
 We used this information to calculate the average connectedness per sample, which was corrected for the maximum potential degree of connectedness.
-Unfortunately our dataset was insufficiently powered to make strong conclusions toward this hypothesis, but this is an interesting observation that warrants further investigation.
+Unfortunately our dataset was insufficiently powered to make strong conclusions toward this hypothesis, but this is an important observation that warrants further investigation.
+Although the changes are not statistically significant using such a small sample set, visual inspection of data suggests there is an underlying significant difference that could be elucidated with further sampling.
 
-Using our limited sample set, we observed that the gut microbiome network structures associated with high-fat diets appeared less connected than those of low-fat diets, although a greater sample size will be required to more properly evaluate this trend **(Figure \ref{dietnetworks} A-B)**.
+Using our small sample set, we observed that the gut microbiome network structures associated with high-fat diets appeared less connected than those of low-fat diets, although a greater sample size will be required to more properly evaluate this trend **(Figure \ref{dietnetworks} A-B)**.
 Five subjects were available for use, all of which had matching bacteria and virome datasets and samples from 8-10 days following the initiation of their diets.
@@
 
-In addition to diet, we found preliminary evidence that obesity influenced network structure.
+In addition to diet, we observed a possible trend that obesity influenced network structure.
 This was done using the three mother samples available from the twin sample set, all of which had matching bacteria and phage samples and confirmed BMI information.
 The obesity-associated network appeared to have a higher degree centrality **(Figure \ref{dietnetworks} C)**, but less closeness centrality than the healthy-associated networks **(Figure \ref{dietnetworks} D)**.
-These results suggested that the obesity-associated networks are less connected, having microbes further from all other microbes within the community.
-This again comes with the caveat that this is a preliminary observation with too small of a sample size to make more substantial claims.
+These results suggested that the obesity-associated networks may be less connected.
+This again comes with the caveat that this is only an opportunistic observation using an existing sample set with too few samples to make more substantial claims.
+We included this observation as a point of interest, given the data was available.
 
@@
 We then calculated the similarity of community networks using the weighted eigenvector centrality of all nodes between all samples.
-Samples with similar network structures were interpreted as having similar capacities for maintaining stability and transmitting genetic material.
+Samples with similar network structures were interpreted as having similar capacities for network robustness and transmitting genetic material.
 
@@
 The gut network structures were not more similar within families (twins and their mothers; intrafamily) compared to other families (other twins and mothers; inter-family) (p-value = `r signif(interstats[c(interstats$site %in% "Twins"), "prob"], digits = sigfig)`, **Figure \ref{intradiv} C**).
-In addition to the gut, skin microbiome network structure was strongly conserved within individuals (p-value < 0.001, **Figure \ref{intradiv} D**).
+In addition to the gut, skin microbiome network structure was conserved within individuals (p-value < 0.001, **Figure \ref{intradiv} D**).
 This distribution was similar when separated by anatomical sites.
@@
 Dissimilarity was calculated using the Bray-Curtis metric based on abundance weighted eigenvector centrality signatures, with a greater distance representing greater dissimilarity in bacteria and phage centrality and abundance.
-Only four subjects were included here, compared to the five used in figure \ref{dietnetworks}, because one of the subjects was missing the inital sampling time point and therefore lacked temporal sampling.
+Only four subjects were included here, compared to the five used in figure \ref{dietnetworks}, because one of the subjects was missing the initial sampling time point and therefore lacked temporal sampling.
 (B) Quantification of gut network dissimilarity within the same subject over time (intrapersonal) and the mean dissimilarity between the subject of interest and all other subjects (interpersonal).
@@
 
-## Association Between Environmental Stability and Network Structure Across the Human Skin Landscape
+## Network Structures Across the Human Skin Landscape
 Extensive work has illustrated differences in diversity and composition of the healthy human skin microbiome between anatomical sites, including bacteria, virus, and fungal communities [@Grice:2009ee; @Findley:2013jf; @Hannigan:2015fz].
-These communities vary by degree of skin moisture, oil, and environmental exposure.
-As viruses are known to influence microbial diversity and community composition, we hypothesized that microbe-virus network structure would be specific to anatomical sites, as well.
+These communities vary by degree of skin moisture, oil, and environmental exposure; features which were defined in the original publication [@Hannigan:2015fz].
+As viruses are known to influence microbial diversity and community composition, we hypothesized that these differences would still be evident after integrating the bacterial and viral datasets and evaluating their microbe-virus network structure between anatomical sites.
 To test this, we evaluated the changes in network structure between anatomical sites within the skin dataset.
+The anatomical sites and their features (e.g.
+moisture & occlusion) were defined in the previous publication through consultation with dermatologists and reference to previous literature [@Hannigan:2015fz].
 
 The average centrality of each sample was quantified using the weighted eigenvector centrality metric.
-Intermittently moist skin sites (dynamic sites that fluctuate between being moist and dry) were significantly less connected than the more stable moist and sebaceous environments (p-value < 0.001, **Figure \ref{skinnetwork} A)**.
+Intermittently moist skin sites (dynamic sites that fluctuate between being moist and dry) were significantly less connected than the moist and sebaceous environments (p-value < 0.001, **Figure \ref{skinnetwork} A)**.
 Also, skin sites that were occluded from the environment were much more highly connected than those that were constantly exposed to the environment or only intermittently occluded (p-value < 0.001, **Figure \ref{skinnetwork} B)**.
@@
 Foundational work has provided a baseline understanding of the human microbiome by characterizing bacterial and viral diversity across the human body [@Grice:2009eea; @Findley:2013jf; @Hannigan:2015fz; @Costello:2009im, @Consortium:2012iz; @Schloss:2005hz; @Minot:2011ez].
-Here, we offer an initial understanding of how phage-bacteria networks differ throughout the human body, so as to provide a baseline for future studies of how and why microbiome networks differ in disease states.
-We developed and implemented a network-based analytical model to evaluate the basic properties of the human microbiome through bacteria and phage relationships, instead of membership or diversity alone.
-This enabled the application of network theory to provide a new perspective on complex ecological communities.
+Here we integrated the bacterial and viral sequence sets to offer an initial understanding of how phage-bacteria networks differ throughout the human body, so as to provide a baseline for future studies of how and why microbiome networks differ in disease states.
+We implemented a network-based analytical model to evaluate the basic properties of the human microbiome through bacteria and phage relationships, instead of membership or diversity alone.
+This approach enabled the application of network theory to provide a new perspective while analyzing bacterial and viral communities simultaneously.
 We utilized metrics of connectivity to model the extent to which communities of bacteria and phages interact through mechanisms such as horizontal gene transfer, modulated bacterial gene expression, and alterations in abundance.
@@
 Gut network structure was not conserved among family members.
-These findings suggested that the community properties inferred from microbiome interaction network structures, such as stability (meaning a more highly connected network is more stable because a randomly removed bacteria or phage node is less likely to divide or disintegrate [@Jeong:2000im; @Poisot:2011jc] the overall network), the potential for horizontal gene transfer between members, and co-evolution of populations, were person-specific.
+These findings suggested that the community properties inferred from microbiome interaction network structures, such as robustness (meaning a more highly connected network is more "robust" to network degradation because a randomly removed bacteria or phage node is less likely to divide or disintegrate [@Jeong:2000ima; @Poisot:2011jc] the overall network), the potential for horizontal gene transfer between members, and co-evolution of populations, were person-specific.
 These properties may be impacted by personal factors ranging from the body's immune system to external environmental conditions, such as climate and diet.
@@
 We observed evidence supporting the ability of environmental conditions to shape gut and skin microbiome interaction network structure by observing that diet and skin location were associated with altered network structures.
-We found evidence that diet was sufficient to alter gut microbiome network connectivity, although this needs to be interpreted as a case observation, due to the small sample size.
-Although our sample size was small, our findings provided some preliminary evidence that high-fat diets were less connected than low-fat diets and that high-fat diets therefore may lead to less stable communities with a decreased ability for microbes to directly influence one another.
+We observed evidence that diet was sufficient to alter gut microbiome network connectivity, although this needs to be interpreted cautiously as a case observation, due to the small sample size.
+Although the available sample size was small, our findings provide some preliminary evidence that high-fat diets are less connected than low-fat diets and that high-fat diets may therefore lead to less robust communities with a decreased ability for microbes to directly influence one another.
 We supported this finding with the observation that obesity may have been associated with decreased network connectivity.
@@
 Network structure differed between environmentally exposed and occluded skin sites.
-The sites under greater environmental fluctuation and exposure (the exposed and intermittently exposed sites) were less connected and therefore were predicted to have a higher propensity for instability.
-Likewise, intermittently moist sites demonstrated less connectedness than the more stable moist and sebaceous sites.
-Together these data suggested that body sites under greater degrees of fluctuation harbored less connected, potentially less stable microbiomes.
-This points to a link between microbiome and environmental stability and warrants further investigation.
+The sites under greater environmental fluctuation and exposure (the exposed and intermittently exposed sites) were less connected and therefore were predicted to have a higher resilience against network degradation when random nodes are removed from the network.
+Likewise, intermittently moist sites demonstrated less connectedness than the moist and sebaceous sites.
+These findings agree with previous work that has shown that bacterial community networks differ by skin environment types [@Bashan:2016is].
+Together these data suggested that body sites under greater degrees of fluctuation harbored less connected microbiomes that are potentially less robust to network disruption by extinction events.
+This points to a link between microbiome and environmental robustness toward network homeostasis and warrants further investigation.
 
-While these findings take us an important step closer to understanding the microbiome through interspecies relationships, there are caveats to and considerations regarding the approach.
-First, as with most classification models, the infection classification model developed and applied is only as good as its training set -- in this case, the collection of experimentally-verified positive and negative infection data, where genomes of all members are fully sequenced.
+While these findings take us an important step closer to understanding the microbiome through interspecies relationships, there are caveats to and considerations regarding our findings.
+First, as with most classification models, the infection classification model developed and applied is only as good as its training set -- in this case, the collection of experimentally-verified positive and negative infection data.
 Large-scale experimental screens for phage and bacteria infectious interactions that report high-confidence negative interactions (i.e., no infection) are desperately needed, as they would provide more robust model training and improved model performance.
-Furthermore, just as we have improved on previous modeling efforts, we expect that new and creative scoring metrics will be integrated into this model to improve future performance.
+Furthermore, just as we have improved on previous modeling efforts, we expect that new and creative scoring metrics will improve future performance.
+Other creative and high performing models are currently being developed and the applications of these models to community network creation will continue to move this field forward [@Villarroel:2016jy; @Galiez:2017gm; @Ahlgren:2017gh].
 
@@
 Future work that employs larger sequence datasets and that avoids the use of bias-inducing amplification steps will build on and validate our findings, as well as inform the design and interpretation of further studies.
+
+Although our models demonstrated satisfactory accuracy and overall performance, it was important to interpret our findings under the realization that our model was not perfect.
+This caveat is not new to the microbiome field, with a notable example being the use of 16S rRNA sequencing using the V4 variable region [@Meisel:2016iu].
+Use of the V4 variable region excluded detection of major bacterial members, meaning that the findings were not able to completely describe the underlying biological environment.
+Despite this caveat, skin microbiome studies provided valuable biological insights by focusing on the community differences between groups (e.g.
+disease and healthy) which were both analyzed the same way.
+Similarly, here we focused our conclusions on the differences between the groups which were all treated the same, so that we can minimize our dependence on a perfect predictive model.
 
@@
 Such population-genetic assessment of phage evolution, coupled with the ecological implications of genome heterogeneity, will inform how to define nodes in future iterations of the ecological network developed here.
-Even though we are hesitant to speculate on phage host ranges at low taxonomic levels in our dataset, the data does aggree with previous reports of instances of broad phage host range [@Malki:2015dw; @PaezEspino:2016gi].
+Even though we are hesitant to speculate on phage host ranges at low taxonomic levels in our dataset, the data does agree with previous reports of instances of broad phage host range [@Malki:2015dw; @PaezEspino:2016gi].
+Additionally, visualization of our dataset interactions using the heat map approach previously used in other host range studies, suggests a trend toward modular and nested tropism, but we do not have the strain-level resolution that powered those previous experimental studies.
 
@@
 By focusing on relationships between bacterial and viral communities, they are studied as the interacting cohorts they are, rather than as independent entities.
-While our developed bacteria-phage interaction framework is a novel conceptual advance, the microbiome also consists of archaea and small eukaryotes, including fungi and *Demodex* mites [@Hannigan:2013im; @Grice:2011gy]--all of which can interact with human immune cells and other non-microbial community members [@Round:2009bz].
+While our developed bacteria-phage interaction framework is a novel conceptual advance, the microbiome also consists of archaea and small eukaryotes, including fungi and *Demodex* mites [@Hannigan:2013im; @Grice:2011gy] -- all of which can interact with human immune cells and other non-microbial community members [@Round:2009bz].
 Future work will build from our approach and include these additional community members and their diverse interactions and relationships (e.g., beyond phage-bacteria).
@@
 The gut virome diet study (SRA: `SRP002424`), twin virome studies (SRA: `SRP002523`; `SRP000319`), and skin virome study (SRA: `SRP049645`) were downloaded as `.sra` files.
+For clarity, the sample sizes used for each study subset were described with the data in the results section.
 Sequencing files were converted to `fastq` format using the `fastq-dump` tool of the NCBI SRA Toolkit (v2.2.0).
@@
 The classification model for predicting interactions was built using experimentally validated bacteria-phage infections or validated lack of infections from six studies [@Jensen:1998vh;@Malki:2015tm;@Schwarzer:2012ez;@Kim:2012dh;@Matsuzaki:1992gw;@Edwards:2015iz].
+No further reference databases were used in our alignment procedures.
 Associated reference genomes were downloaded from the European Bioinformatics Institute (see details in source code).
@@
 
+![**Classification Model Performance By Nested Cross-Validation.** *Box plot illustrating the median and variance of phage-bacteria interaction prediction model.
+Performance was evaluated using nested cross validation, meaning that 20% of the samples were randomly withheld from model training and then used to evaluate performance.
+The results of 100 random iterations are shown.
+Metrics include area under the curve (gray), sensitivity (red), and specificity (tan).*\label{nestedperformance}](){ width=100% }
+
+\newpage
+
+![**Stable Classification Model Performance Over Random Iterations** *In addition to nested cross-validation, here we show the results from the five-fold cross validation, in which 20% of the samples were randomly withheld during the training stage for model evaluation and mtry tuning.
+The results of 25 random iterations are shown.
+Metrics include area under the curve (red), sensitivity (green), and specificity (blue).
+Dashed line highlight the random point of 0.5.*\label{crossval}](){ width=100% }
+
+\newpage
+
 ![**Structure of the interactive network.** *Metadata relationships to samples (Phage Sample ID and Bacteria Sample ID) included the associated time point, the study, the subject the sample was taken from, and the associated disease.
@@
 Sequence count abundance for each OGU within each sample was also recorded.*\label{NetworkDiagram}](){ width=50% }
+
+\newpage
+
+![**Heatmap of Phage-Bacteria Interaction Relationships of Master Network.** *Heatmap illustrating the ranges of infectious interactions predicted between bacteria and bacteriophages across our three studies.
+Bacterial OGUs are aligned on the vertical access, and the bacteriophage OGUs are organized on the horizontal access.
+OGUs are organized near other OGUs with similar infectious profiles, which are further illustrated by the dendrograms.
+Predicted infections are tan and predicted lacks of interactions are red.*\label{inthm}](){ width=50% }
+
+\newpage
+
+![**Distribution of node eccentricity across subnetworks.** *Histograms illustrating the distributions of node eccentricity values across the subnetworks, for supplementing the node, edge, and diameter values provided for the networks.
+Eccentricity of each node is the shortest distance of that node to the furthest other node within the graph.*\label{eccplot}](){ width=50% }
 

### 2017-12-28 doc/manuscript.tex
Updated manuscript with revisions.

@@
 \hypersetup{unicode=true,
-            pdftitle={Biogeography \& environmental conditions shape bacteriophage \& bacteria interaction networks across the human microbiome},
+            pdftitle={Biogeography \& environmental conditions shape bacteriophage-bacteria networks across the human microbiome},
             pdfborder={0 0 0},
@@
 
-\title{Biogeography \& environmental conditions shape bacteriophage \& bacteria
-interaction networks across the human microbiome}
+\title{Biogeography \& environmental conditions shape bacteriophage-bacteria
+networks across the human microbiome}
     \usepackage{authblk}
@@
 play important roles in health and disease. Most previous work has
-relied on studying microbes and viruses independently, thereby reducing
+relied on studying bacteria and viruses independently, thereby reducing
 them to two separate communities. Such approaches are unable to capture
 how these microbial communities interact, such as through processes that
-maintain community stability or allow phage-host populations to
-co-evolve. We developed and implemented a network-based analytical
-approach to describe phage-bacteria network diversity throughout the
-human body. We accomplished this by building a machine learning
-algorithm to predict which phages could infect which bacteria in a given
-microbiome. This algorithm was applied to paired viral and bacterial
-metagenomic sequence sets from three previously published human cohorts.
-We organized the predicted interactions into networks that allowed us to
-evaluate phage-bacteria connectedness across the human body. We found
-that gut and skin network structures were person-specific and not
-conserved among cohabitating family members. High-fat diets and obesity
-appeared to be associated with less connected networks. Network
-structure differed between skin sites, with those exposed to the
-external environment being less connected and more prone to instability.
-This study quantified and contrasted the diversity of virome-microbiome
-networks across the human body and illustrated how environmental factors
-may influence phage-bacteria interactive dynamics. This work provides a
-baseline for future studies to better understand system perturbations,
-such as disease states, through ecological networks.
+maintain community robustness or allow phage-host populations to
+co-evolve. We implemented a network-based analytical approach to
+describe phage-bacteria network diversity throughout the human body. We
+built these community networks using a machine learning algorithm to
+predict which phages could infect which bacteria in a given microbiome.
+Our algorithm was applied to paired viral and bacterial metagenomic
+sequence sets from three previously published human cohorts. We
+organized the predicted interactions into networks that allowed us to
+evaluate phage-bacteria connectedness across the human body. We observed
+evidence that gut and skin network structures were person-specific and
+not conserved among cohabitating family members. High-fat diets appeared
+to be associated with less connected networks. Network structure
+differed between skin sites, with those exposed to the external
+environment being less connected and likely more susceptible to network
+degradation by microbial extinction events. This study quantified and
+contrasted the diversity of virome-microbiome networks across the human
+body and illustrated how environmental factors may influence
+phage-bacteria interactive dynamics. This work provides a baseline for
+future studies to better understand system perturbations, such as
+disease states, through ecological networks.
 
@@
 phage communities, which frequently share genetic information and work
-together to maintain stable ecosystems. Removal of bacteria or phage can
-disrupt or even collapse those ecosystems. Relationship-based network
-approaches allow us to capture this interaction information. Using this
-network-based approach with three independent human cohorts, we were
-able to present an initial understanding of how phage-bacteria networks
-differ throughout the human body, so as to provide a baseline for future
-studies of how and why microbiome networks differ in disease states.
+together to maintain ecosystem homestatsis (e.g.~kill-the-winner
+dynamics). Removal of bacteria or phage can disrupt or even collapse
+those ecosystems. Relationship-based network approaches allow us to
+capture this interaction information. Using this network-based approach
+with three independent human cohorts, we were able to present an initial
+understanding of how phage-bacteria networks differ throughout the human
+body, so as to provide a baseline for future studies of how and why
+microbiome networks differ in disease states.
 
@@
 communities, which frequently share genetic information and work
-together to maintain stable ecosystems. Removal of bacteria or phages
-can disrupt or even collapse those ecosystems {[}18,28--37{]}.
-Relationship-based network approaches allow us to capture this
-interaction information. Studying such bacteria-phage interactions
-through community-wide networks built from inferred relationships could
-offer further insights into the drivers of human microbiome diversity
-across body sites and enable the study of human microbiome network
-dynamics overall.
+together to maintain ecosystem structure (e.g.~kill-the-winner dynamics
+that prevent domination by a single bacterium). Removal of bacteria or
+phages can disrupt or even collapse those ecosystems {[}18,28--37{]}.
+Integrating these datasets as relationship-based networks allow us to
+capture this complex interaction information. Studying such
+bacteria-phage interactions through community-wide networks built from
+inferred relationships begins to provide us with insights into the
+drivers of human microbiome diversity across body sites and enable the
+study of human microbiome network dynamics overall.
 
@@
 metagenomic datasets, rather than culture-dependent data {[}33{]}, which
-is limited in the scale of possible experiments and analyses. Our
-metagenomic interaction inference model improved upon previous models of
-phage-host predictions that have utilized a variety of techniques, such
-as linear models to predict bacteria-phage co-occurrence using taxonomic
-assignments {[}40{]}, and nucleotide similarity models that were applied
-to both whole virus genomes {[}41{]} and related clusters of whole and
-partial virus genomes {[}42{]}. Our approach uniquely included protein
-interaction data and was validated based on experimentally determined
-positive and negative interactions (i.e.~who does and does not infect
-whom). Through this approach we were able to provide a basic
-understanding of the network dynamics associated with phage and
-bacterial communities on and in the human body. By building and
-utilizing a microbiome network, we found that different people, body
+is limited in the scale of possible experiments and analyses. We
+implemented an adapted metagenomic interaction inference model that made
+some improvements upon previous phage-host interaction prediction
+models. Previous approaches have utilized a variety of techniques, such
+as linear models that were used to predict bacteria-phage co-occurrence
+using taxonomic assignments {[}40{]}, and nucleotide similarity models
+that were applied to both whole virus genomes {[}41{]} and clusters of
+whole and partial virus genomes {[}42{]}. Our approach uniquely included
+protein interaction data and was validated based on experimentally
+determined positive and negative interactions (i.e.~who does and does
+not infect whom). We built on previous modeling work as a means to our
+ends, and focused on the biological insights we could gain instead of
+building a superior model and presenting our work as a toolkit. We
+therefore did not focus on extensive benchmarking against other existing
+models {[}41,43--45{]}. Through this approach we were able to provide an
+initial basic understanding of the network dynamics associated with
+phage and bacterial communities on and in the human body. By building
+and utilizing a microbiome network, we found that different people, body
 sites, and anatomical locations not only support distinct microbiome
-membership and diversity {[}13,14,38,39,43--45{]}, but also support
+membership and diversity {[}13,14,38,39,46--48{]}, but also support
 ecological communities with distinct communication structures and
-propensities toward community instability. Through an improved
-understanding of network structures across the human body, we empower
-future studies to investigate how these communities dynamics are
-influenced by disease states and the overall impact they may have on
-human health.
+robustness to network degradation by extinction events. Through an
+improved understanding of network structures across the human body, we
+aim to empower future studies to investigate how these communities
+dynamics are influenced by disease states and the overall impact they
+may have on human health.
 
@@
 reports \textbf{(Supplemental Figure \ref{ContigStats} -
-\ref{ClusterStats})} {[}42{]}. We referred to these operationally
-defined groups of related contigs as operational genomic units (OGUs).
-Each OGU represented a genomically similar sub-population of either
-bacteria or phages. Contig lengths within clusters ranged between
-\(10^{3}\) and \(10^{5.5}\) bp \textbf{(Supplemental Figure
-\ref{ContigStats} - \ref{ClusterStats})}.
-
-While supplementing the previous quality control measures \textbf{(Table
-S1)} we found that, in light of the rigorous purification and quality
-control during sample collection and preparation, 77\% (228 / 298
-operational genomic units) still had some nucleotide similarity to a
-given bacterial reference genome (e-value \textless{} \(10^{-25}\)). As
-absence of bacterial contamination had been confirmed by sensitive
-molecular methods \textbf{(Table S1)}, we interpreted this as evidence
-that the majority of the gut and skin bacteriophages were temperate and
-thereby shared elements with bacterial reference genomes that contained
-similar integrated phages when sequenced--a trend previously reported
-{[}14{]}. Additionally, we identified two OGUs as being complete phages
-using the stringent Virsorter phage identification algorithm (class 1
-confidence group) {[}47{]}.
-
-The whole metagenomic shotgun sequence samples primarily consisted of
-bacteria, with an average viral relative abundance of 0.4\%
+\ref{ClusterStats})} {[}42{]}. This was done both for dimensionality
+reduction and to prevent inflation of node counts by using contigs which
+are expected to represent multiple fragments from the same genomes. This
+was also done to create genome analogs that we could use in our
+classification model which was built using genome sequences. We referred
+to these operationally defined groups of related contigs as operational
+genomic units (OGUs). Each OGU represented a genomically similar
+sub-population of either bacteria or phages. Contig lengths within
+clusters ranged between \(10^{3}\) and \(10^{5.5}\) bp
+\textbf{(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})}.
+
+While supplementing the previous virome fraction quality control
+measures \textbf{(Table S1)} we found that, in light of the rigorous
+purification and quality control during sample collection and
+preparation, 77\% (228 / 298 operational genomic units) still had some
+nucleotide similarity to a given bacterial reference genome (e-value
+\textless{} \(10^{-25}\)). As most phages in these communities have been
+shown to be temperate (they integrate into bacterial genomes) using
+methods that included nucleotide alignments of phages to bacterial
+reference genomes, we interpreted this as confirmation that the majority
+of those phages were temperate and therefore shared elements with
+bacterial reference genomes -- a trend previously reported {[}14{]}. We
+further confirmed that while the majority of these were expected to be
+temperate phages, there still remained a low level of bacterial sequence
+noise, which was evident as five (1.7\%) OGUs with similar sequence
+elements to the bacterial 16S rRNA gene (blastn, e-value \textless{}
+\(10^{-25}\), length \textgreater{} 1,000 bp). This is in line with
+previous work which has suggested that bacterial noise is an inevitable
+technical issue, and we considered this noise while interpreting our
+findings {[}50{]}. We also identified two OGUs as being complete, high
+confidence phages using the stringent Virsorter phage identification
+algorithm (class 1 confidence group) {[}51{]}.
+
+The whole metagenomic shotgun sequence samples, which primarily
+consisted of bacteria, had an average viral relative abundance of 0.4\%
 \textbf{(Table S1)} {[}13,14,38,39{]}. We found that only 2\% (6 / 280
@@
 temperate phage OGUs in the bacterial dataset using the Virsorter
-algorithm {[}47{]}. Together this suggests only minimal
-``contamination'' of the bacterial OGUs.
-
-\subsection{Evaluating the Model to Predict Phage-Bacteria
-Interactions}\label{evaluating-the-model-to-predict-phage-bacteria-interactions}
+algorithm {[}51{]}. Together this suggests minimal bacterial OGU noise
+that should be considered in the study conclusions.
+
+\subsection{Implementing Phage-Bacteria Interaction Prediction to Build
+a Community
+Network}\label{implementing-phage-bacteria-interaction-prediction-to-build-a-community-network}
 
@@
 random forest model trained on experimentally validated infectious
-relationships from six previous publications {[}41,48--52{]}. Only
-bacteria and phages were used in the model. The training set contained
-43 diverse bacterial species and 30 diverse phage strains, including
-both broad and specific ranges of infection \textbf{(Supplemental Figure
-\ref{ValidationOverview} A - B)}. While it is true that there are more
-known phages that infect bacteria, we were limited by the information
-confirming which phages do not infect certain bacteria and attempted to
-keep the numbers of positive and negative interactions similar. Phages
-with linear and circular genomes, as well as ssDNA and dsDNA genomes,
-were included in the analysis. Because we used DNA sequencing studies,
-RNA phages were not considered \textbf{(Supplemental Figure
-\ref{ValidationOverview} C-D)}. This training set included both positive
-relationships (a phage infects a bacterium) and negative relationships
-(a phage does not infect a bacterium). This allowed us to validate the
-false positive and false negative rates associated with our candidate
-models, thereby building upon previous work that only considered
-positive relationships {[}41{]}.
-
-Four phage and bacterial genomic features were used in a random forest
+relationships from six previous publications {[}41,52--56{]}. Only
+bacteria and bacteriophages were used in the model. The training set
+contained 43 diverse bacterial species and 30 diverse phage strains,
+including both broad and specific ranges of infection
+\textbf{(Supplemental Figure \ref{ValidationOverview} A - B)}. While it
+is true that there are more known phages that infect bacteria, we were
+limited by the information confirming which phages do not infect certain
+bacteria and attempted to keep the numbers of positive and negative
+interactions similar. Phages with linear and circular genomes, as well
+as ssDNA and dsDNA genomes, were included in the analysis. Because we
+used DNA sequencing studies, RNA phages were not considered
+\textbf{(Supplemental Figure \ref{ValidationOverview} C-D)}. This
+training set included both positive relationships (a phage infects a
+bacterium) and negative relationships (a phage does not infect a
+bacterium). This allowed us to validate the false positive and false
+negative rates associated with our candidate models, thereby building
+upon previous work that only considered positive relationships {[}41{]}.
+It is also worth noting that while a positive interaction is strong
+evidence that the interaction exists, we must also be conscious that
+negative interactions are only weak evidence for a lack of interaction
+because the finding could be the result of our inability to reproduce
+conditions in which those interactions occur. Altogether we decided to
+maintain a balanced dataset at the cost of under-sampling the available
+positive interaction information because the use of such a severely
+unbalanced dataset often results in over-fit and uninformative model
+training.
+
+Four phage and bacterial genomic features were used in our random forest
 model to predict infectious relationships between bacteria and phages:
@@
 similarity of protein families associated with experimentally identified
-protein-protein interactions {[}53{]}. The resulting random forest model
-was assessed using nested cross validation, and the median area under
-its receiver operating characteristic (ROC) curve was 0.788, the median
-model sensitivity was 0.952, and median specificity was 0.615
-\textbf{(Figure \ref{RocCurve} A)}. The most important predictor in the
-model was amino acid similarity between genes, followed by nucleotide
-similarity of whole genomes \textbf{(Figure \ref{RocCurve} B)}. Protein
-family interactions were moderately important to the model, and CRISPRs
-were largely uninformative, due to the minimal amount of identifiable
-CRISPRs in the dataset and their redundancy with the nucleotide
-similarity methods \textbf{(Figure \ref{RocCurve} B)}. Approximately one
-third of the training set relationships yielded no score and therefore
-were unable to be assigned an interaction prediction \textbf{(Figure
-\ref{RocCurve} C)}.
+protein-protein interactions {[}57{]}. These features were calculated
+using the training set described above. We chose to utilize these
+metrics that directly compare nucleotide sequences between sample phages
+and bacteria, instead of relying on alignment to reference genomes or
+known marker genes, because we are extrapolating our model to highly
+diverse communities which we expect to diverge significantly from the
+available reference genomes. The resulting random forest model was
+assessed using nested cross validation, and the median area under its
+receiver operating characteristic (ROC) curve was 0.788, the median
+model sensitivity was 0.905, and median specificity was 0.538
+\textbf{(Figure \ref{RocCurve} A)}. This balance of confident true
+positives at the cost of fewer true negatives is ideal for this type of
+dataset which consists of primarily positive connections
+\textbf{(Supplemental Figure \ref{nestedperformance})}. Nested cross
+validation of the model demonstrated that the sensitivity and
+specificity of the model could vary but the overall model performance
+(AUC) remained more consistent \textbf{(Supplemental Figure
+\ref{crossval})}. This suggested that our model would perform with a
+similar overall accuracy despite changes in sensitivity/specificity
+trade-offs. The most important predictor in the model was amino acid
+similarity between genes, followed by nucleotide similarity of whole
+genomes \textbf{(Figure \ref{RocCurve} B)}. Protein family interactions
+were moderately important to the model, and CRISPRs were largely
+uninformative, due to the minimal amount of identifiable CRISPRs in the
+dataset and their redundancy with the nucleotide similarity methods
+\textbf{(Figure \ref{RocCurve} B)}. Approximately one third of the
+training set relationships yielded no score and therefore were unable to
+be assigned an interaction prediction \textbf{(Figure \ref{RocCurve}
+C)}.
 
@@
 yield any scores. Those interactions without scores were automatically
-classified as not having interactions. (D) Bipartite visualization of
-the resulting phage-bacteria network. Phage OGUs are presented in
-orange, bacteria OGUs in red, and their interaction edges are
-represented as connecting lines. This network includes information from
-all three published studies. (E) Network diameter (measure of graph
-size; the greatest number of traversed vertices required between two
-vertices), (F) number of vertices, and (G) number of edges
+classified as not having interactions. (D) Network diameter (measure of
+graph size; the greatest number of traversed vertices required between
+two vertices), (E) number of vertices, and (F) number of edges
 (relationships) for the total network (orange) and the individual study
@@
 bacteria and phage operational genomic units, which were then used to
-build the interactive network. The master network contained the three
-studies as sub-networks, which themselves each contained sub-networks
-for each sample \textbf{(Figure \ref{RocCurve} D)}. Metadata including
-study, sample ID, disease, and OGU abundance within the community were
-stored in the master network for parsing in downstream analyses
-\textbf{(Supplemental Figure \ref{NetworkDiagram})}. Bacterial and phage
-relative abundance was recorded in each sample for each OGU and the
-weight of the edge connecting those OGUs was calculated as a function of
-those relative abundance values. The separate extraction of the phage
-and bacterial libraries ensured a more accurate measurement of the
-microbial communities, as has been outlined previously {[}54,55{]}. The
-master network was highly connected and contained 72,287 infectious
+build the interactive network. The master network, analogous to the
+universal microbiome network concept previously described {[}58{]},
+contained the three studies as sub-networks, which themselves each
+contained sub-networks for each sample \textbf{(Supplemental Figure
+\ref{NetworkDiagram})}. Metadata including study, sample ID, disease,
+and OGU abundance within the community were stored in the master network
+for parsing in downstream analyses \textbf{(Supplemental Figure
+\ref{NetworkDiagram})}. The phage and bacteria of the master network
+demonstrated both narrow broad ranges of infectious interactions
+\textbf{(Supplemental Figure \ref{inthm})}. Bacterial and phage relative
+abundance was recorded in each sample for each OGU and the weight of the
+edge connecting those OGUs was calculated as a function of those
+relative abundance values. The separate extraction of the phage and
+bacterial libraries ensured a more accurate measurement of the microbial
+communities, as has been outlined previously {[}59,60{]}. The master
+network was highly connected and contained 72,287 infectious
 relationships among 578 nodes, representing 298 phages and 280 bacteria.
@@
 72,127 (99.8\%) of the master network nodes and edges, respectively
-\textbf{(Figure \ref{RocCurve} E - F)}. The phages and bacteria in the
-diet and twin sample sets were more sparsely related, with the diet
+\textbf{(Figure \ref{RocCurve} E - F)}. Additionally, the subnetworks
+demonstrated narrow ranges of eccentricity across their nodes
+\textbf{(Supplemental Figure \ref{eccplot})}. The phages and bacteria in
+the diet and twin sample sets were more sparsely related, with the diet
 study consisting of 89 (41 phages, 48 bacteria) nodes and 5,566
@@
 bacteria) nodes and 17,250 relationships \textbf{(Figure \ref{RocCurve}
-E - F)}.
-
-\subsection{Role of Diet \& Obesity in Gut Microbiome
-Connectivity}\label{role-of-diet-obesity-in-gut-microbiome-connectivity}
+E - F)}. As an interesting validation measure, we identified five
+(1.7\%) examples of phage OGUs which similar genomic elements to the
+four previously described, broadly infectious phages isolated from Lake
+Michigan (tblastx; e-value \textless{} \(10^{-25}\)) {[}61{]}.
+
+\subsection{Role of Diet on Gut Microbiome
+Connectivity}\label{role-of-diet-on-gut-microbiome-connectivity}
 
@@
 availability and gut microbiome composition and diversity, including
-bacteria and phages {[}14,56,57{]}. Previous work in isolated
+bacteria and phages {[}14,62,63{]}. Previous work in isolated
 culture-based systems has suggested that changes in nutrient
@@
 network structure by quantifying how central each sample's network was
-on average. We accomplished this by utilizing two common centrality
-metrics: degree centrality and closeness centrality. Degree centrality,
-the simplest centrality metric, was defined as the number of connections
-each phage made with each bacterium. We supplemented measurements of
-degree centrality with measurements of closeness centrality. Closeness
-centrality is a metric of how close each phage or bacterium is to all of
-the other phages and bacteria in the network. A higher closeness
-centrality suggests that the effects of genetic information or altered
-abundance would be more impactful to all other microbes in the system. A
+on average. We accomplished this by utilizing two common weighted
+centrality metrics: degree centrality and closeness centrality. Degree
+centrality, the simplest centrality metric, was defined as the number of
+connections each phage made with each bacterium. We supplemented
+measurements of degree centrality with measurements of closeness
+centrality. Closeness centrality is a metric of how close each phage or
+bacterium is to all of the other phages and bacteria in the network. A
+higher closeness centrality suggests that the effects of genetic
+information or altered abundance would be more impactful to all other
+microbes in the system. Because these are weighted metrics, the values
+are functions of both connectivity as well as community composition. A
 network with higher average closeness centrality also indicates an
 overall greater degree of connections, which suggests a greater
-resilience against instability. This is because more highly connected
-networks are more stable as a result of pathway dependencies and the
-unlikelihood that a randomly removed bacteria or phage would cause major
-divisions across the network {[}30,58{]}. We used this information to
-calculate the average connectedness per sample, which was corrected for
-the maximum potential degree of connectedness. Unfortunately our dataset
-was insufficiently powered to make strong conclusions toward this
-hypothesis, but this is an interesting observation that warrants further
-investigation.
-
-Using our limited sample set, we observed that the gut microbiome
-network structures associated with high-fat diets appeared less
-connected than those of low-fat diets, although a greater sample size
-will be required to more properly evaluate this trend \textbf{(Figure
-\ref{dietnetworks} A-B)}. Five subjects were available for use, all of
-which had matching bacteria and virome datasets and samples from 8-10
-days following the initiation of their diets. High-fat diets appeared to
-exhibit reduced degree centrality \textbf{(Figure \ref{dietnetworks}
-A)}, suggesting bacteria in high-fat environments were targeted by fewer
-phages and that phage tropism was more restricted. High-fat diets also
-appeared to exhibit decreased closeness centrality \textbf{(Figure
+resilience against network degradation by extinction events {[}30,64{]}.
+This is because more highly connected networks are less likely to
+degrade into multiple smaller networks when bacteria or phages are
+randomly removed {[}30,64{]}. We used this information to calculate the
+average connectedness per sample, which was corrected for the maximum
+potential degree of connectedness. Unfortunately our dataset was
+insufficiently powered to make strong conclusions toward this
+hypothesis, but this is an important observation that warrants further
+investigation. Although the changes are not statistically significant
+using such a small sample set, visual inspection of data suggests there
+is an underlying significant difference that could be elucidated with
+further sampling.
+
+Using our small sample set, we observed that the gut microbiome network
+structures associated with high-fat diets appeared less connected than
+those of low-fat diets, although a greater sample size will be required
+to more properly evaluate this trend \textbf{(Figure \ref{dietnetworks}
+A-B)}. Five subjects were available for use, all of which had matching
+bacteria and virome datasets and samples from 8-10 days following the
+initiation of their diets. High-fat diets appeared to exhibit reduced
+degree centrality \textbf{(Figure \ref{dietnetworks} A)}, suggesting
+bacteria in high-fat environments were targeted by fewer phages and that
+phage tropism was more restricted. High-fat diets also appeared to
+exhibit decreased closeness centrality \textbf{(Figure
 \ref{dietnetworks} B)}, indicating that bacteria and phages were more
@@
 
-In addition to diet, we found preliminary evidence that obesity
+In addition to diet, we observed a possible trend that obesity
 influenced network structure. This was done using the three mother
@@
 than the healthy-associated networks \textbf{(Figure \ref{dietnetworks}
-D)}. These results suggested that the obesity-associated networks are
-less connected, having microbes further from all other microbes within
-the community. This again comes with the caveat that this is a
-preliminary observation with too small of a sample size to make more
-substantial claims.
+D)}. These results suggested that the obesity-associated networks may be
+less connected. This again comes with the caveat that this is only an
+opportunistic observation using an existing sample set with too few
+samples to make more substantial claims. We included this observation as
+a point of interest, given the data was available.
 
@@
 with people remaining more similar to themselves than to other people
-over time {[}13,59,60{]}. We therefore hypothesized that this personal
+over time {[}13,65,66{]}. We therefore hypothesized that this personal
 conservation extended to microbiome network structure. We addressed this
@@
 Samples with similar network structures were interpreted as having
-similar capacities for maintaining stability and transmitting genetic
+similar capacities for network robustness and transmitting genetic
 material.
@@
 0.312, \textbf{Figure \ref{intradiv} C}). In addition to the gut, skin
-microbiome network structure was strongly conserved within individuals
-(p-value \textless{} 0.001, \textbf{Figure \ref{intradiv} D}). This
-distribution was similar when separated by anatomical sites. Most sites
-were statistically significantly more conserved within individuals
+microbiome network structure was conserved within individuals (p-value
+\textless{} 0.001, \textbf{Figure \ref{intradiv} D}). This distribution
+was similar when separated by anatomical sites. Most sites were
+statistically significantly more conserved within individuals
 \textbf{(Supplemental Figure \ref{allskin})}.
@@
 compared to the five used in figure \ref{dietnetworks}, because one of
-the subjects was missing the inital sampling time point and therefore
+the subjects was missing the initial sampling time point and therefore
 lacked temporal sampling. (B) Quantification of gut network
@@
 
-\subsection{Association Between Environmental Stability and Network
-Structure Across the Human Skin
-Landscape}\label{association-between-environmental-stability-and-network-structure-across-the-human-skin-landscape}
+\subsection{Network Structures Across the Human Skin
+Landscape}\label{network-structures-across-the-human-skin-landscape}
 
@@
 of the healthy human skin microbiome between anatomical sites, including
-bacteria, virus, and fungal communities {[}13,44,59{]}. These
+bacteria, virus, and fungal communities {[}13,47,65{]}. These
 communities vary by degree of skin moisture, oil, and environmental
-exposure. As viruses are known to influence microbial diversity and
-community composition, we hypothesized that microbe-virus network
-structure would be specific to anatomical sites, as well. To test this,
-we evaluated the changes in network structure between anatomical sites
-within the skin dataset.
+exposure; features which were defined in the original publication
+{[}13{]}. As viruses are known to influence microbial diversity and
+community composition, we hypothesized that these differences would
+still be evident after integrating the bacterial and viral datasets and
+evaluating their microbe-virus network structure between anatomical
+sites. To test this, we evaluated the changes in network structure
+between anatomical sites within the skin dataset. The anatomical sites
+and their features (e.g.~moisture \& occlusion) were defined in the
+previous publication through consultation with dermatologists and
+reference to previous literature {[}13{]}.
 
@@
 sites that fluctuate between being moist and dry) were significantly
-less connected than the more stable moist and sebaceous environments
-(p-value \textless{} 0.001, \textbf{Figure \ref{skinnetwork} A)}. Also,
-skin sites that were occluded from the environment were much more highly
+less connected than the moist and sebaceous environments (p-value
+\textless{} 0.001, \textbf{Figure \ref{skinnetwork} A)}. Also, skin
+sites that were occluded from the environment were much more highly
 connected than those that were constantly exposed to the environment or
@@
 microbiome by characterizing bacterial and viral diversity across the
-human body {[}13,14,43--45,61{]}. Here, we offer an initial
-understanding of how phage-bacteria networks differ throughout the human
-body, so as to provide a baseline for future studies of how and why
-microbiome networks differ in disease states. We developed and
-implemented a network-based analytical model to evaluate the basic
-properties of the human microbiome through bacteria and phage
-relationships, instead of membership or diversity alone. This enabled
-the application of network theory to provide a new perspective on
-complex ecological communities. We utilized metrics of connectivity to
-model the extent to which communities of bacteria and phages interact
-through mechanisms such as horizontal gene transfer, modulated bacterial
-gene expression, and alterations in abundance.
+human body {[}13,14,46--48,67{]}. Here we integrated the bacterial and
+viral sequence sets to offer an initial understanding of how
+phage-bacteria networks differ throughout the human body, so as to
+provide a baseline for future studies of how and why microbiome networks
+differ in disease states. We implemented a network-based analytical
+model to evaluate the basic properties of the human microbiome through
+bacteria and phage relationships, instead of membership or diversity
+alone. This approach enabled the application of network theory to
+provide a new perspective while analyzing bacterial and viral
+communities simultaneously. We utilized metrics of connectivity to model
+the extent to which communities of bacteria and phages interact through
+mechanisms such as horizontal gene transfer, modulated bacterial gene
+expression, and alterations in abundance.
 
 Just as gut microbiome and virome composition and diversity are
-conserved in individuals {[}13,43,44,60{]}, gut and skin microbiome
+conserved in individuals {[}13,46,47,66{]}, gut and skin microbiome
 network structures were conserved within individuals over time. Gut
@@
 suggested that the community properties inferred from microbiome
-interaction network structures, such as stability (meaning a more highly
-connected network is more stable because a randomly removed bacteria or
-phage node is less likely to divide or disintegrate {[}30,58{]} the
-overall network), the potential for horizontal gene transfer between
-members, and co-evolution of populations, were person-specific. These
-properties may be impacted by personal factors ranging from the body's
-immune system to external environmental conditions, such as climate and
-diet.
+interaction network structures, such as robustness (meaning a more
+highly connected network is more ``robust'' to network degradation
+because a randomly removed bacteria or phage node is less likely to
+divide or disintegrate {[}30,64{]} the overall network), the potential
+for horizontal gene transfer between members, and co-evolution of
+populations, were person-specific. These properties may be impacted by
+personal factors ranging from the body's immune system to external
+environmental conditions, such as climate and diet.
 
@@
 observing that diet and skin location were associated with altered
-network structures. We found evidence that diet was sufficient to alter
-gut microbiome network connectivity, although this needs to be
-interpreted as a case observation, due to the small sample size.
-Although our sample size was small, our findings provided some
-preliminary evidence that high-fat diets were less connected than
-low-fat diets and that high-fat diets therefore may lead to less stable
+network structures. We observed evidence that diet was sufficient to
+alter gut microbiome network connectivity, although this needs to be
+interpreted cautiously as a case observation, due to the small sample
+size. Although the available sample size was small, our findings provide
+some preliminary evidence that high-fat diets are less connected than
+low-fat diets and that high-fat diets may therefore lead to less robust
 communities with a decreased ability for microbes to directly influence
@@
 exposed sites) were less connected and therefore were predicted to have
-a higher propensity for instability. Likewise, intermittently moist
-sites demonstrated less connectedness than the more stable moist and
-sebaceous sites. Together these data suggested that body sites under
-greater degrees of fluctuation harbored less connected, potentially less
-stable microbiomes. This points to a link between microbiome and
-environmental stability and warrants further investigation.
+a higher resilience against network degradation when random nodes are
+removed from the network. Likewise, intermittently moist sites
+demonstrated less connectedness than the moist and sebaceous sites.
+These findings agree with previous work that has shown that bacterial
+community networks differ by skin environment types {[}58{]}. Together
+these data suggested that body sites under greater degrees of
+fluctuation harbored less connected microbiomes that are potentially
+less robust to network disruption by extinction events. This points to a
+link between microbiome and environmental robustness toward network
+homeostasis and warrants further investigation.
 
@@
 the microbiome through interspecies relationships, there are caveats to
-and considerations regarding the approach. First, as with most
+and considerations regarding our findings. First, as with most
 classification models, the infection classification model developed and
@@
 collection of experimentally-verified positive and negative infection
-data, where genomes of all members are fully sequenced. Large-scale
-experimental screens for phage and bacteria infectious interactions that
-report high-confidence negative interactions (i.e., no infection) are
-desperately needed, as they would provide more robust model training and
-improved model performance. Furthermore, just as we have improved on
-previous modeling efforts, we expect that new and creative scoring
-metrics will be integrated into this model to improve future
-performance.
+data. Large-scale experimental screens for phage and bacteria infectious
+interactions that report high-confidence negative interactions (i.e., no
+infection) are desperately needed, as they would provide more robust
+model training and improved model performance. Furthermore, just as we
+have improved on previous modeling efforts, we expect that new and
+creative scoring metrics will improve future performance. Other creative
+and high performing models are currently being developed and the
+applications of these models to community network creation will continue
+to move this field forward {[}43--45{]}.
 
@@
 generation protocols. It is now known that MDA results in biases in
-microbial community composition {[}62{]}, as well as toward ssDNA viral
-genomes {[}63,64{]}, thus rendering the resulting microbial and viral
+microbial community composition {[}68{]}, as well as toward ssDNA viral
+genomes {[}69,70{]}, thus rendering the resulting microbial and viral
 metagenomes largely non-quantitative. Future work that employs larger
@@
 design and interpretation of further studies.
+
+Although our models demonstrated satisfactory accuracy and overall
+performance, it was important to interpret our findings under the
+realization that our model was not perfect. This caveat is not new to
+the microbiome field, with a notable example being the use of 16S rRNA
+sequencing using the V4 variable region {[}60{]}. Use of the V4 variable
+region excluded detection of major bacterial members, meaning that the
+findings were not able to completely describe the underlying biological
+environment. Despite this caveat, skin microbiome studies provided
+valuable biological insights by focusing on the community differences
+between groups (e.g.~disease and healthy) which were both analyzed the
+same way. Similarly, here we focused our conclusions on the differences
+between the groups which were all treated the same, so that we can
+minimize our dependence on a perfect predictive model.
 
@@
 have been implemented in other metagenomic sequencing studies, as well
-{[}42,65--67{]}. These approaches could offer yet another level of
+{[}42,71--73{]}. These approaches could offer yet another level of
 sophistication to our network-based analyses. While this operationally
@@
 more focused, culture-based studies such as the work by Malki \emph{et
-al} {[}68{]}. Future work must address this limitation, e.g., through
+al} {[}61{]}. Future work must address this limitation, e.g., through
 improved binning methods and deeper metagenomic shotgun sequencing, but
@@
 ecologically and evolutionarily cohesive units for both phage and
-bacteria {[}69{]}. Defining operational genomic units and their
+bacteria {[}74{]}. Defining operational genomic units and their
 taxonomic underpinnings (e.g., whether OGU clusters represent genera or
@@
 gene content, average nucleotide identity of shared genes, and pairwise
-differences between genomes {[}70{]}. Such population-genetic assessment
+differences between genomes {[}75{]}. Such population-genetic assessment
 of phage evolution, coupled with the ecological implications of genome
@@
 speculate on phage host ranges at low taxonomic levels in our dataset,
-the data does aggree with previous reports of instances of broad phage
-host range {[}68,71{]}.
+the data does agree with previous reports of instances of broad phage
+host range {[}61,76{]}. Additionally, visualization of our dataset
+interactions using the heat map approach previously used in other host
+range studies, suggests a trend toward modular and nested tropism, but
+we do not have the strain-level resolution that powered those previous
+experimental studies.
 
@@
 consists of archaea and small eukaryotes, including fungi and
-\emph{Demodex} mites {[}1,72{]}--all of which can interact with human
-immune cells and other non-microbial community members {[}73{]}. Future
+\emph{Demodex} mites {[}1,77{]} -- all of which can interact with human
+immune cells and other non-microbial community members {[}78{]}. Future
 work will build from our approach and include these additional community
@@
 \texttt{SRP002523}; \texttt{SRP000319}), and skin virome study (SRA:
-\texttt{SRP049645}) were downloaded as \texttt{.sra} files. Sequencing
-files were converted to \texttt{fastq} format using the
-\texttt{fastq-dump} tool of the NCBI SRA Toolkit (v2.2.0). Sequences
-were quality trimmed using the Fastx toolkit (v0.0.14) to exclude bases
-with quality scores below 33 and shorter than 75 bp {[}74{]}. Paired end
-reads were filtered to exclude sequences missing their corresponding
-pair using the \texttt{get\_trimmed\_pairs.py} script available in the
-source code.
+\texttt{SRP049645}) were downloaded as \texttt{.sra} files. For clarity,
+the sample sizes used for each study subset were described with the data
+in the results section. Sequencing files were converted to
+\texttt{fastq} format using the \texttt{fastq-dump} tool of the NCBI SRA
+Toolkit (v2.2.0). Sequences were quality trimmed using the Fastx toolkit
+(v0.0.14) to exclude bases with quality scores below 33 and shorter than
+75 bp {[}79{]}. Paired end reads were filtered to exclude sequences
+missing their corresponding pair using the
+\texttt{get\_trimmed\_pairs.py} script available in the source code.
 
@@
 Contigs were assembled using the Megahit assembly program (v1.0.6)
-{[}75{]}. A minimum contig length of 1 kb was used. Iterative k-mer
+{[}80{]}. A minimum contig length of 1 kb was used. Iterative k-mer
 stepping began at a minimum length of 21 and progressed by 20 until 101.
@@
 aligned to phage or bacterial contigs using the Bowtie2 global aligner
-(v2.2.1) {[}76{]}. We defined a mismatch threshold of 1 bp and seed
+(v2.2.1) {[}81{]}. We defined a mismatch threshold of 1 bp and seed
 length of 25 bp. Sequence abundance was calculated from the Bowtie2
@@
 taxonomic units (OTUs), although here we are clustering contigs into
-operational genomic units (OGUs) {[}61{]}.
-
-Contigs were clustered using the CONCOCT algorithm (v0.4.0) {[}77{]}.
+operational genomic units (OGUs) {[}67{]}.
+
+Contigs were clustered using the CONCOCT algorithm (v0.4.0) {[}82{]}.
 Because of our large dataset and limits in computational efficiency, we
@@
 identify phages (all default parameters were used), using only those in
-the high confidence identification category ``class 1'' {[}47{]}.
+the high confidence identification category ``class 1'' {[}51{]}.
 
@@
 Open reading frames (ORFs) were identified using the Prodigal program
-(V2.6.2) with the meta mode parameter and default settings {[}78{]}.
+(V2.6.2) with the meta mode parameter and default settings {[}83{]}.
 
@@
 experimentally validated bacteria-phage infections or validated lack of
-infections from six studies {[}41,48--52{]}. Associated reference
+infections from six studies {[}41,52--56{]}. No further reference
+databases were used in our alignment procedures. Associated reference
 genomes were downloaded from the European Bioinformatics Institute (see
@@
 interactions or not. The classification model was built using the Caret
-R package (v6.0.73) {[}79{]}. The model was trained using five-fold
+R package (v6.0.73) {[}84{]}. The model was trained using five-fold
 cross validation with ten repeats, and the median model performance was
@@
 identified from bacterial genomes using the PilerCR program (v1.06)
-{[}80{]}. Resulting spacer sequences were filtered to exclude spacers
+{[}85{]}. Resulting spacer sequences were filtered to exclude spacers
 shorter than 20 bp and longer than 65 bp. Spacer sequences were aligned
 to the phage genomes using the nucleotide BLAST algorithm with default
-parameters (v2.4.0) {[}81{]}. The mean percent identity for each
+parameters (v2.4.0) {[}86{]}. The mean percent identity for each
 matching pair was recorded for use in our classification model.
@@
 between the genes using the Diamond protein alignment algorithm
-(v0.7.11.60) {[}82{]}. The mean alignment bitscores for each genome pair
+(v0.7.11.60) {[}87{]}. The mean alignment bitscores for each genome pair
 were recorded for use in our classification model.
@@
 predicted to interact using the Pfam interaction information within the
-Intact database {[}53{]}. The mean bitscores of the matches between each
+Intact database {[}57{]}. The mean bitscores of the matches between each
 pair were recorded for use in the classification model.
@@
 associated metadata were used to populate a graph database using Neo4j
-graph database software (v2.3.1) {[}83{]}. This network was used for
+graph database software (v2.3.1) {[}88{]}. This network was used for
 downstream community analysis.
@@
 Degree and closeness centrality were calculated using the associated
-functions within the igraph R package (v1.0.1) {[}84{]}.
+functions within the igraph R package (v1.0.1) {[}89{]}.
 
@@
 \centering
+\includegraphics[width=1.00000\textwidth]{}
+\caption{\textbf{Classification Model Performance By Nested
+Cross-Validation.} \emph{Box plot illustrating the median and variance
+of phage-bacteria interaction prediction model. Performance was
+evaluated using nested cross validation, meaning that 20\% of the
+samples were randomly withheld from model training and then used to
+evaluate performance. The results of 100 random iterations are shown.
+Metrics include area under the curve (gray), sensitivity (red), and
+specificity (tan).}\label{nestedperformance}}
+\end{figure}
+
+\newpage
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=1.00000\textwidth]{}
+\caption{\textbf{Stable Classification Model Performance Over Random
+Iterations} \emph{In addition to nested cross-validation, here we show
+the results from the five-fold cross validation, in which 20\% of the
+samples were randomly withheld during the training stage for model
+evaluation and mtry tuning. The results of 25 random iterations are
+shown. Metrics include area under the curve (red), sensitivity (green),
+and specificity (blue). Dashed line highlight the random point of
+0.5.}\label{crossval}}
+\end{figure}
+
+\newpage
+
+\begin{figure}[htbp]
+\centering
 \includegraphics[width=0.50000\textwidth]{}
@@
 recorded.}\label{NetworkDiagram}}
+\end{figure}
+
+\newpage
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.50000\textwidth]{}
+\caption{\textbf{Heatmap of Phage-Bacteria Interaction Relationships of
+Master Network.} \emph{Heatmap illustrating the ranges of infectious
+interactions predicted between bacteria and bacteriophages across our
+three studies. Bacterial OGUs are aligned on the vertical access, and
+the bacteriophage OGUs are organized on the horizontal access. OGUs are
+organized near other OGUs with similar infectious profiles, which are
+further illustrated by the dendrograms. Predicted infections are tan and
+predicted lacks of interactions are red.}\label{inthm}}
+\end{figure}
+
+\newpage
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.50000\textwidth]{}
+\caption{\textbf{Distribution of node eccentricity across subnetworks.}
+\emph{Histograms illustrating the distributions of node eccentricity
+values across the subnetworks, for supplementing the node, edge, and
+diameter values provided for the networks. Eccentricity of each node is
+the shortest distance of that node to the furthest other node within the
+graph.}\label{eccplot}}
 \end{figure}
@@
 
+\hypertarget{ref-Villarroel:2016jy}{}
+43. Villarroel J, Kleinheinz KA, Jurtz VI, Zschach H, Lund O, Nielsen M,
+et al. HostPhinder: A Phage Host Prediction Tool. Viruses. 2016;8: 116.
+
+\hypertarget{ref-Galiez:2017gm}{}
+44. Galiez C, Siebert M, Enault F, Vincent J, Söding J. WIsH: who is the
+host? Predicting prokaryotic hosts from metagenomic phage contigs.
+Bioinformatics. 2017;33: 3113--3114.
+
+\hypertarget{ref-Ahlgren:2017gh}{}
+45. Ahlgren NA, Ren J, Lu YY, Fuhrman JA, Sun F. Alignment-free
+\(d_2^*\) oligonucleotide frequency dissimilarity measure improves
+prediction of hosts from metagenomically-derived viral sequences.
+Nucleic Acids Research. 2017;45: 39--53.
+
 \hypertarget{ref-Grice:2009eea}{}
-43. Grice EA, Kong HH, Conlan S, Deming CB, Davis J, Young AC, et al.
+46. Grice EA, Kong HH, Conlan S, Deming CB, Davis J, Young AC, et al.
 Topographical and Temporal Diversity of the Human Skin Microbiome.
@@
 \hypertarget{ref-Findley:2013jf}{}
-44. Findley K, Oh J, Yang J, Conlan S, Deming C, Meyer JA, et al.
+47. Findley K, Oh J, Yang J, Conlan S, Deming C, Meyer JA, et al.
 Topographic diversity of fungal and bacterial communities in human skin.
@@
 \hypertarget{ref-Costello:2009im}{}
-45. Costello EK, Lauber CL, Hamady M, Fierer N, Gordon JI, Knight R.
+48. Costello EK, Lauber CL, Hamady M, Fierer N, Gordon JI, Knight R.
 Bacterial community variation in human body habitats across space and
@@
 \hypertarget{ref-Consortium:2012iz}{}
-46. Consortium THMP. A framework for human microbiome research. Nature.
+49. Consortium THMP. A framework for human microbiome research. Nature.
 2012;486: 215--221.
 
+\hypertarget{ref-Roux:2013cr}{}
+50. Roux S, Krupovic M, Debroas D, Forterre P, Enault F. Assessment of
+viral community functional potential from viral metagenomes may be
+hampered by contamination with cellular sequences. Open Biology. 2013;3:
+130160--130160.
+
 \hypertarget{ref-Roux:2015dh}{}
-47. Roux S, Enault F, Hurwitz BL, Sullivan MB. VirSorter: mining viral
+51. Roux S, Enault F, Hurwitz BL, Sullivan MB. VirSorter: mining viral
 signal from microbial genomic data. PeerJ. 2015;3: e985--20.
@@
 \hypertarget{ref-Jensen:1998vh}{}
-48. Jensen EC, Schrader HS, Rieland B, Thompson TL, Lee KW, Nickerson
+52. Jensen EC, Schrader HS, Rieland B, Thompson TL, Lee KW, Nickerson
 KW, et al. Prevalence of broad-host-range lytic bacteriophages of
@@
 \hypertarget{ref-Malki:2015tm}{}
-49. Malki K, Kula A, Bruder K, Sible E. Bacteriophages isolated from
+53. Malki K, Kula A, Bruder K, Sible E. Bacteriophages isolated from
 Lake Michigan demonstrate broad host-range across several bacterial
@@
 \hypertarget{ref-Schwarzer:2012ez}{}
-50. Schwarzer D, Buettner FFR, Browning C, Nazarov S, Rabsch W, Bethe A,
+54. Schwarzer D, Buettner FFR, Browning C, Nazarov S, Rabsch W, Bethe A,
 et al. A multivalent adsorption apparatus explains the broad host range
@@
 \hypertarget{ref-Kim:2012dh}{}
-51. Kim S, Rahman M, Seol SY, Yoon SS, Kim J. Pseudomonas aeruginosa
+55. Kim S, Rahman M, Seol SY, Yoon SS, Kim J. Pseudomonas aeruginosa
 bacteriophage PA1Ø requires type IV pili for infection and shows broad
@@
 \hypertarget{ref-Matsuzaki:1992gw}{}
-52. Matsuzaki S, Tanaka S, Koga T, Kawata T. A Broad-Host-Range
+56. Matsuzaki S, Tanaka S, Koga T, Kawata T. A Broad-Host-Range
 Vibriophage, KVP40, Isolated from Sea Water. Microbiology and
@@
 \hypertarget{ref-Orchard:2014hq}{}
-53. Orchard S, Ammari M, Aranda B, Breuza L, Briganti L, Broackes-Carter
+57. Orchard S, Ammari M, Aranda B, Breuza L, Briganti L, Broackes-Carter
 F, et al. The MIntAct project--IntAct as a common curation platform for
@@
 
+\hypertarget{ref-Bashan:2016is}{}
+58. Bashan A, Gibson TE, Friedman J, Carey VJ, Weiss ST, Hohmann EL, et
+al. Universality of human microbial dynamics. Nature. 2016;534:
+259--262.
+
 \hypertarget{ref-Kleiner:2015kd}{}
-54. Kleiner M, Hooper LV, Duerkop BA. Evaluation of methods to purify
+59. Kleiner M, Hooper LV, Duerkop BA. Evaluation of methods to purify
 virus-like particles for metagenomic sequencing of intestinal viromes.
@@
 \hypertarget{ref-Meisel:2016iu}{}
-55. Meisel JS, Hannigan GD, Tyldsley AS, SanMiguel AJ, Hodkinson BP,
+60. Meisel JS, Hannigan GD, Tyldsley AS, SanMiguel AJ, Hodkinson BP,
 Zheng Q, et al. Skin microbiome surveys are strongly influenced by
@@
 
+\hypertarget{ref-Malki:2015dw}{}
+61. Malki K, Kula A, Bruder K, Sible E, Hatzopoulos T, Steidel S, et al.
+Bacteriophages isolated from Lake Michigan demonstrate broad host-range
+across several bacterial phyla. Virology Journal. 2015;12: 164.
+
 \hypertarget{ref-Turnbaugh:2009hf}{}
-56. Turnbaugh PJ, Ridaura VK, Faith JJ, Rey FE, Knight R, Gordon JI. The
+62. Turnbaugh PJ, Ridaura VK, Faith JJ, Rey FE, Knight R, Gordon JI. The
 effect of diet on the human gut microbiome: a metagenomic analysis in
@@
 \hypertarget{ref-David:2014cl}{}
-57. David LA, Maurice CF, Carmody RN, Gootenberg DB, Button JE, Wolfe
+63. David LA, Maurice CF, Carmody RN, Gootenberg DB, Button JE, Wolfe
 BE, et al. Diet rapidly and reproducibly alters the human gut
@@
 
-\hypertarget{ref-Jeong:2000im}{}
-58. Jeong H, Tombor B, Albert R, Oltvai ZN, Barabási AL. The large-scale
+\hypertarget{ref-Jeong:2000ima}{}
+64. Jeong H, Tombor B, Albert R, Oltvai ZN, Barabási AL. The large-scale
 organization of metabolic networks. Nature. 2000;407: 651--654.
@@
 \hypertarget{ref-Grice:2009ee}{}
-59. Grice EA, Kong HH, Conlan S, Deming CB, Davis J, Young AC, et al.
+65. Grice EA, Kong HH, Conlan S, Deming CB, Davis J, Young AC, et al.
 Topographical and Temporal Diversity of the Human Skin Microbiome.
@@
 \hypertarget{ref-Minot:2013ih}{}
-60. Minot S, Bryson A, Chehoud C, Wu GD, Lewis JD, Bushman FD. Rapid
+66. Minot S, Bryson A, Chehoud C, Wu GD, Lewis JD, Bushman FD. Rapid
 evolution of the human gut virome. Proceedings of the National Academy
@@
 \hypertarget{ref-Schloss:2005hz}{}
-61. Schloss PD, Handelsman J. Introducing DOTUR, a computer program for
+67. Schloss PD, Handelsman J. Introducing DOTUR, a computer program for
 defining operational taxonomic units and estimating species richness.
@@
 \hypertarget{ref-Yilmaz:2010jb}{}
-62. Yilmaz S, Allgaier M, Hugenholtz P. Multiple displacement
+68. Yilmaz S, Allgaier M, Hugenholtz P. Multiple displacement
 amplification compromises quantitative analysis of metagenomes. Nature
@@
 \hypertarget{ref-Kim:2008to}{}
-63. Kim KH, Chang HW, Nam YD, Roh SW. Amplification of uncultured
+69. Kim KH, Chang HW, Nam YD, Roh SW. Amplification of uncultured
 single-stranded DNA viruses from rice paddy soil. Applied and \ldots{}.
@@
 \hypertarget{ref-Kim:2011hp}{}
-64. Kim K-H, Bae J-W. Amplification methods bias metagenomic libraries
+70. Kim K-H, Bae J-W. Amplification methods bias metagenomic libraries
 of uncultured single-stranded and double-stranded DNA viruses. Applied
@@
 \hypertarget{ref-Minot:2012ed}{}
-65. Minot S, Wu GD, Lewis JD, Bushman FD. Conservation of gene cassettes
+71. Minot S, Wu GD, Lewis JD, Bushman FD. Conservation of gene cassettes
 among diverse viruses of the human gut. PLOS ONE. 2012;7: e42342.
@@
 \hypertarget{ref-Deng:2014eb}{}
-66. Deng L, Ignacio-Espinoza JC, Gregory AC, Poulos BT, Weitz JS,
+72. Deng L, Ignacio-Espinoza JC, Gregory AC, Poulos BT, Weitz JS,
 Hugenholtz P, et al. Viral tagging reveals discrete populations in
@@
 \hypertarget{ref-Brum:2015iaa}{}
-67. Brum JR, Ignacio-Espinoza JC, Roux S, Doulcier G, Acinas SG, Alberti
+73. Brum JR, Ignacio-Espinoza JC, Roux S, Doulcier G, Acinas SG, Alberti
 A, et al. Ocean plankton. Patterns and ecological drivers of ocean viral
@@
 
-\hypertarget{ref-Malki:2015dw}{}
-68. Malki K, Kula A, Bruder K, Sible E, Hatzopoulos T, Steidel S, et al.
-Bacteriophages isolated from Lake Michigan demonstrate broad host-range
-across several bacterial phyla. Virology Journal. 2015;12: 164.
-
 \hypertarget{ref-Polz:2006fi}{}
-69. Polz MF, Hunt DE, Preheim SP, Weinreich DM. Patterns and mechanisms
+74. Polz MF, Hunt DE, Preheim SP, Weinreich DM. Patterns and mechanisms
 of genetic and phenotypic differentiation in marine microbes.
@@
 \hypertarget{ref-Gregory:2016cg}{}
-70. Gregory AC, Solonenko SA, Ignacio-Espinoza JC, LaButti K, Copeland
+75. Gregory AC, Solonenko SA, Ignacio-Espinoza JC, LaButti K, Copeland
 A, Sudek S, et al. Genomic differentiation among wild cyanophages
@@
 \hypertarget{ref-PaezEspino:2016gi}{}
-71. Paez-Espino D, Eloe-Fadrosh EA, Pavlopoulos GA, Thomas AD, Huntemann
+76. Paez-Espino D, Eloe-Fadrosh EA, Pavlopoulos GA, Thomas AD, Huntemann
 M, Mikhailova N, et al. Uncovering Earth's virome. Nature. 2016;
@@
 \hypertarget{ref-Grice:2011gy}{}
-72. Grice EA, Segre JA. The skin microbiome. Nature Reviews
+77. Grice EA, Segre JA. The skin microbiome. Nature Reviews
 Microbiology. 2011;9: 244--253.
@@
 \hypertarget{ref-Round:2009bz}{}
-73. Round JL, Mazmanian SK. The gut microbiota shapes intestinal immune
+78. Round JL, Mazmanian SK. The gut microbiota shapes intestinal immune
 responses during health and disease. Nature reviews Immunology. 2009;9:
@@
 \hypertarget{ref-FASTXToolkit:wr}{}
-74. Hannon GJ. FASTX-Toolkit.: GNU Affero General Public License.
+79. Hannon GJ. FASTX-Toolkit.: GNU Affero General Public License.
 
 \hypertarget{ref-Li:2016kd}{}
-75. Li D, Luo R, Liu C-M, Leung C-M, Ting H-F, Sadakane K, et al.
+80. Li D, Luo R, Liu C-M, Leung C-M, Ting H-F, Sadakane K, et al.
 MEGAHIT v1.0: A fast and scalable metagenome assembler driven by
@@
 \hypertarget{ref-Langmead:2012jh}{}
-76. Langmead B, Salzberg SL. Fast gapped-read alignment with Bowtie 2.
+81. Langmead B, Salzberg SL. Fast gapped-read alignment with Bowtie 2.
 Nature Methods. 2012;9: 357--359.
@@
 \hypertarget{ref-Alneberg:2014fc}{}
-77. Alneberg J, Bjarnason BS aacute ri, Bruijn I de, Schirmer M, Quick
+82. Alneberg J, Bjarnason BS aacute ri, Bruijn I de, Schirmer M, Quick
 J, Ijaz UZ, et al. Binning metagenomic contigs by coverage and
@@
 \hypertarget{ref-Hyatt:2012cy}{}
-78. Hyatt D, LoCascio PF, Hauser LJ, Uberbacher EC. Gene and translation
+83. Hyatt D, LoCascio PF, Hauser LJ, Uberbacher EC. Gene and translation
 initiation site prediction in metagenomic sequences. Bioinformatics.
@@
 \hypertarget{ref-caretClassificatio:ux5fU2Litux5f1}{}
-79. Kuhn M. caret: Classification and Regression Training.
+84. Kuhn M. caret: Classification and Regression Training.
 
 \hypertarget{ref-Edgar:2007bh}{}
-80. Edgar RC. PILER-CR: fast and accurate identification of CRISPR
+85. Edgar RC. PILER-CR: fast and accurate identification of CRISPR
 repeats. BMC Bioinformatics. 2007;8: 18.
@@
 \hypertarget{ref-Camacho:2009fc}{}
-81. Camacho C, Coulouris G, Avagyan V, Ma N, Papadopoulos J, Bealer K,
+86. Camacho C, Coulouris G, Avagyan V, Ma N, Papadopoulos J, Bealer K,
 et al. BLAST+: architecture and applications. BMC Bioinformatics.
@@
 \hypertarget{ref-Buchfink:2015ki}{}
-82. Buchfink B, Xie C, Huson DH. Fast and sensitive protein alignment
+87. Buchfink B, Xie C, Huson DH. Fast and sensitive protein alignment
 using DIAMOND. Nature Methods. 2015;12: 59--60.
@@
 \hypertarget{ref-Neoj:Fuwr6PBN}{}
-83. Neo4j.
+88. Neo4j.
 
 \hypertarget{ref-Theigraphsoftware:vh}{}
-84. Csardi G, Nepusz T. The igraph software package for complex network
+89. Csardi G, Nepusz T. The igraph software package for complex network
 research.

### 2017-12-28 doc/oman.tex
Updated manuscript with revisions.

--- 2017-10-01 doc/manuscript.tex
@@
 unlikelihood that a randomly removed bacteria or phage would cause major
-divisions across the network {[}30,58{]}. We used this information to
-calculate the average connectedness per sample, which was corrected for
-the maximum potential degree of connectedness. Unfortunately our dataset
-was insufficiently powered to make strong conclusions toward this
-hypothesis, but this is an interesting observation that warrants further
-investigation.
+divisions across the network {[}{\textbf{???}},30{]}. We used this
+information to calculate the average connectedness per sample, which was
+corrected for the maximum potential degree of connectedness.
+Unfortunately our dataset was insufficiently powered to make strong
+conclusions toward this hypothesis, but this is an interesting
+observation that warrants further investigation.
 
@@
 with people remaining more similar to themselves than to other people
-over time {[}13,59,60{]}. We therefore hypothesized that this personal
+over time {[}13,58,59{]}. We therefore hypothesized that this personal
 conservation extended to microbiome network structure. We addressed this
@@
 of the healthy human skin microbiome between anatomical sites, including
-bacteria, virus, and fungal communities {[}13,44,59{]}. These
+bacteria, virus, and fungal communities {[}13,44,58{]}. These
 communities vary by degree of skin moisture, oil, and environmental
@@
 microbiome by characterizing bacterial and viral diversity across the
-human body {[}13,14,43--45,61{]}. Here, we offer an initial
+human body {[}13,14,43--45,60{]}. Here, we offer an initial
 understanding of how phage-bacteria networks differ throughout the human
@@
 Just as gut microbiome and virome composition and diversity are
-conserved in individuals {[}13,43,44,60{]}, gut and skin microbiome
+conserved in individuals {[}13,43,44,59{]}, gut and skin microbiome
 network structures were conserved within individuals over time. Gut
@@
 connected network is more stable because a randomly removed bacteria or
-phage node is less likely to divide or disintegrate {[}30,58{]} the
-overall network), the potential for horizontal gene transfer between
-members, and co-evolution of populations, were person-specific. These
-properties may be impacted by personal factors ranging from the body's
-immune system to external environmental conditions, such as climate and
-diet.
+phage node is less likely to divide or disintegrate
+{[}{\textbf{???}},30{]} the overall network), the potential for
+horizontal gene transfer between members, and co-evolution of
+populations, were person-specific. These properties may be impacted by
+personal factors ranging from the body's immune system to external
+environmental conditions, such as climate and diet.
 
@@
 generation protocols. It is now known that MDA results in biases in
-microbial community composition {[}62{]}, as well as toward ssDNA viral
-genomes {[}63,64{]}, thus rendering the resulting microbial and viral
+microbial community composition {[}61{]}, as well as toward ssDNA viral
+genomes {[}62,63{]}, thus rendering the resulting microbial and viral
 metagenomes largely non-quantitative. Future work that employs larger
@@
 have been implemented in other metagenomic sequencing studies, as well
-{[}42,65--67{]}. These approaches could offer yet another level of
+{[}42,64--66{]}. These approaches could offer yet another level of
 sophistication to our network-based analyses. While this operationally
@@
 more focused, culture-based studies such as the work by Malki \emph{et
-al} {[}68{]}. Future work must address this limitation, e.g., through
+al} {[}67{]}. Future work must address this limitation, e.g., through
 improved binning methods and deeper metagenomic shotgun sequencing, but
@@
 ecologically and evolutionarily cohesive units for both phage and
-bacteria {[}69{]}. Defining operational genomic units and their
+bacteria {[}68{]}. Defining operational genomic units and their
 taxonomic underpinnings (e.g., whether OGU clusters represent genera or
@@
 gene content, average nucleotide identity of shared genes, and pairwise
-differences between genomes {[}70{]}. Such population-genetic assessment
+differences between genomes {[}69{]}. Such population-genetic assessment
 of phage evolution, coupled with the ecological implications of genome
@@
 the data does aggree with previous reports of instances of broad phage
-host range {[}68,71{]}.
+host range {[}67,70{]}.
 
@@
 consists of archaea and small eukaryotes, including fungi and
-\emph{Demodex} mites {[}1,72{]}--all of which can interact with human
-immune cells and other non-microbial community members {[}73{]}. Future
+\emph{Demodex} mites {[}1,71{]}--all of which can interact with human
+immune cells and other non-microbial community members {[}72{]}. Future
 work will build from our approach and include these additional community
@@
 were quality trimmed using the Fastx toolkit (v0.0.14) to exclude bases
-with quality scores below 33 and shorter than 75 bp {[}74{]}. Paired end
+with quality scores below 33 and shorter than 75 bp {[}73{]}. Paired end
 reads were filtered to exclude sequences missing their corresponding
@@
 Contigs were assembled using the Megahit assembly program (v1.0.6)
-{[}75{]}. A minimum contig length of 1 kb was used. Iterative k-mer
+{[}74{]}. A minimum contig length of 1 kb was used. Iterative k-mer
 stepping began at a minimum length of 21 and progressed by 20 until 101.
@@
 aligned to phage or bacterial contigs using the Bowtie2 global aligner
-(v2.2.1) {[}76{]}. We defined a mismatch threshold of 1 bp and seed
+(v2.2.1) {[}75{]}. We defined a mismatch threshold of 1 bp and seed
 length of 25 bp. Sequence abundance was calculated from the Bowtie2
@@
 taxonomic units (OTUs), although here we are clustering contigs into
-operational genomic units (OGUs) {[}61{]}.
-
-Contigs were clustered using the CONCOCT algorithm (v0.4.0) {[}77{]}.
+operational genomic units (OGUs) {[}60{]}.
+
+Contigs were clustered using the CONCOCT algorithm (v0.4.0) {[}76{]}.
 Because of our large dataset and limits in computational efficiency, we
@@
 Open reading frames (ORFs) were identified using the Prodigal program
-(V2.6.2) with the meta mode parameter and default settings {[}78{]}.
+(V2.6.2) with the meta mode parameter and default settings {[}77{]}.
 
@@
 interactions or not. The classification model was built using the Caret
-R package (v6.0.73) {[}79{]}. The model was trained using five-fold
+R package (v6.0.73) {[}78{]}. The model was trained using five-fold
 cross validation with ten repeats, and the median model performance was
@@
 identified from bacterial genomes using the PilerCR program (v1.06)
-{[}80{]}. Resulting spacer sequences were filtered to exclude spacers
+{[}79{]}. Resulting spacer sequences were filtered to exclude spacers
 shorter than 20 bp and longer than 65 bp. Spacer sequences were aligned
 to the phage genomes using the nucleotide BLAST algorithm with default
-parameters (v2.4.0) {[}81{]}. The mean percent identity for each
+parameters (v2.4.0) {[}80{]}. The mean percent identity for each
 matching pair was recorded for use in our classification model.
@@
 between the genes using the Diamond protein alignment algorithm
-(v0.7.11.60) {[}82{]}. The mean alignment bitscores for each genome pair
+(v0.7.11.60) {[}81{]}. The mean alignment bitscores for each genome pair
 were recorded for use in our classification model.
@@
 associated metadata were used to populate a graph database using Neo4j
-graph database software (v2.3.1) {[}83{]}. This network was used for
+graph database software (v2.3.1) {[}82{]}. This network was used for
 downstream community analysis.
@@
 Degree and closeness centrality were calculated using the associated
-functions within the igraph R package (v1.0.1) {[}84{]}.
+functions within the igraph R package (v1.0.1) {[}83{]}.
 
@@
 
-\hypertarget{ref-Jeong:2000im}{}
-58. Jeong H, Tombor B, Albert R, Oltvai ZN, Barabási AL. The large-scale
-organization of metabolic networks. Nature. 2000;407: 651--654.
-
 \hypertarget{ref-Grice:2009ee}{}
-59. Grice EA, Kong HH, Conlan S, Deming CB, Davis J, Young AC, et al.
+58. Grice EA, Kong HH, Conlan S, Deming CB, Davis J, Young AC, et al.
 Topographical and Temporal Diversity of the Human Skin Microbiome.
@@
 \hypertarget{ref-Minot:2013ih}{}
-60. Minot S, Bryson A, Chehoud C, Wu GD, Lewis JD, Bushman FD. Rapid
+59. Minot S, Bryson A, Chehoud C, Wu GD, Lewis JD, Bushman FD. Rapid
 evolution of the human gut virome. Proceedings of the National Academy
@@
 \hypertarget{ref-Schloss:2005hz}{}
-61. Schloss PD, Handelsman J. Introducing DOTUR, a computer program for
+60. Schloss PD, Handelsman J. Introducing DOTUR, a computer program for
 defining operational taxonomic units and estimating species richness.
@@
 \hypertarget{ref-Yilmaz:2010jb}{}
-62. Yilmaz S, Allgaier M, Hugenholtz P. Multiple displacement
+61. Yilmaz S, Allgaier M, Hugenholtz P. Multiple displacement
 amplification compromises quantitative analysis of metagenomes. Nature
@@
 \hypertarget{ref-Kim:2008to}{}
-63. Kim KH, Chang HW, Nam YD, Roh SW. Amplification of uncultured
+62. Kim KH, Chang HW, Nam YD, Roh SW. Amplification of uncultured
 single-stranded DNA viruses from rice paddy soil. Applied and \ldots{}.
@@
 \hypertarget{ref-Kim:2011hp}{}
-64. Kim K-H, Bae J-W. Amplification methods bias metagenomic libraries
+63. Kim K-H, Bae J-W. Amplification methods bias metagenomic libraries
 of uncultured single-stranded and double-stranded DNA viruses. Applied
@@
 \hypertarget{ref-Minot:2012ed}{}
-65. Minot S, Wu GD, Lewis JD, Bushman FD. Conservation of gene cassettes
+64. Minot S, Wu GD, Lewis JD, Bushman FD. Conservation of gene cassettes
 among diverse viruses of the human gut. PLOS ONE. 2012;7: e42342.
@@
 \hypertarget{ref-Deng:2014eb}{}
-66. Deng L, Ignacio-Espinoza JC, Gregory AC, Poulos BT, Weitz JS,
+65. Deng L, Ignacio-Espinoza JC, Gregory AC, Poulos BT, Weitz JS,
 Hugenholtz P, et al. Viral tagging reveals discrete populations in
@@
 \hypertarget{ref-Brum:2015iaa}{}
-67. Brum JR, Ignacio-Espinoza JC, Roux S, Doulcier G, Acinas SG, Alberti
+66. Brum JR, Ignacio-Espinoza JC, Roux S, Doulcier G, Acinas SG, Alberti
 A, et al. Ocean plankton. Patterns and ecological drivers of ocean viral
@@
 \hypertarget{ref-Malki:2015dw}{}
-68. Malki K, Kula A, Bruder K, Sible E, Hatzopoulos T, Steidel S, et al.
+67. Malki K, Kula A, Bruder K, Sible E, Hatzopoulos T, Steidel S, et al.
 Bacteriophages isolated from Lake Michigan demonstrate broad host-range
@@
 \hypertarget{ref-Polz:2006fi}{}
-69. Polz MF, Hunt DE, Preheim SP, Weinreich DM. Patterns and mechanisms
+68. Polz MF, Hunt DE, Preheim SP, Weinreich DM. Patterns and mechanisms
 of genetic and phenotypic differentiation in marine microbes.
@@
 \hypertarget{ref-Gregory:2016cg}{}
-70. Gregory AC, Solonenko SA, Ignacio-Espinoza JC, LaButti K, Copeland
+69. Gregory AC, Solonenko SA, Ignacio-Espinoza JC, LaButti K, Copeland
 A, Sudek S, et al. Genomic differentiation among wild cyanophages
@@
 \hypertarget{ref-PaezEspino:2016gi}{}
-71. Paez-Espino D, Eloe-Fadrosh EA, Pavlopoulos GA, Thomas AD, Huntemann
+70. Paez-Espino D, Eloe-Fadrosh EA, Pavlopoulos GA, Thomas AD, Huntemann
 M, Mikhailova N, et al. Uncovering Earth's virome. Nature. 2016;
@@
 \hypertarget{ref-Grice:2011gy}{}
-72. Grice EA, Segre JA. The skin microbiome. Nature Reviews
+71. Grice EA, Segre JA. The skin microbiome. Nature Reviews
 Microbiology. 2011;9: 244--253.
@@
 \hypertarget{ref-Round:2009bz}{}
-73. Round JL, Mazmanian SK. The gut microbiota shapes intestinal immune
+72. Round JL, Mazmanian SK. The gut microbiota shapes intestinal immune
 responses during health and disease. Nature reviews Immunology. 2009;9:
@@
 \hypertarget{ref-FASTXToolkit:wr}{}
-74. Hannon GJ. FASTX-Toolkit.: GNU Affero General Public License.
+73. Hannon GJ. FASTX-Toolkit.: GNU Affero General Public License.
 
 \hypertarget{ref-Li:2016kd}{}
-75. Li D, Luo R, Liu C-M, Leung C-M, Ting H-F, Sadakane K, et al.
+74. Li D, Luo R, Liu C-M, Leung C-M, Ting H-F, Sadakane K, et al.
 MEGAHIT v1.0: A fast and scalable metagenome assembler driven by
@@
 \hypertarget{ref-Langmead:2012jh}{}
-76. Langmead B, Salzberg SL. Fast gapped-read alignment with Bowtie 2.
+75. Langmead B, Salzberg SL. Fast gapped-read alignment with Bowtie 2.
 Nature Methods. 2012;9: 357--359.
@@
 \hypertarget{ref-Alneberg:2014fc}{}
-77. Alneberg J, Bjarnason BS aacute ri, Bruijn I de, Schirmer M, Quick
+76. Alneberg J, Bjarnason BS aacute ri, Bruijn I de, Schirmer M, Quick
 J, Ijaz UZ, et al. Binning metagenomic contigs by coverage and
@@
 \hypertarget{ref-Hyatt:2012cy}{}
-78. Hyatt D, LoCascio PF, Hauser LJ, Uberbacher EC. Gene and translation
+77. Hyatt D, LoCascio PF, Hauser LJ, Uberbacher EC. Gene and translation
 initiation site prediction in metagenomic sequences. Bioinformatics.
@@
 \hypertarget{ref-caretClassificatio:ux5fU2Litux5f1}{}
-79. Kuhn M. caret: Classification and Regression Training.
+78. Kuhn M. caret: Classification and Regression Training.
 
 \hypertarget{ref-Edgar:2007bh}{}
-80. Edgar RC. PILER-CR: fast and accurate identification of CRISPR
+79. Edgar RC. PILER-CR: fast and accurate identification of CRISPR
 repeats. BMC Bioinformatics. 2007;8: 18.
@@
 \hypertarget{ref-Camacho:2009fc}{}
-81. Camacho C, Coulouris G, Avagyan V, Ma N, Papadopoulos J, Bealer K,
+80. Camacho C, Coulouris G, Avagyan V, Ma N, Papadopoulos J, Bealer K,
 et al. BLAST+: architecture and applications. BMC Bioinformatics.
@@
 \hypertarget{ref-Buchfink:2015ki}{}
-82. Buchfink B, Xie C, Huson DH. Fast and sensitive protein alignment
+81. Buchfink B, Xie C, Huson DH. Fast and sensitive protein alignment
 using DIAMOND. Nature Methods. 2015;12: 59--60.
@@
 \hypertarget{ref-Neoj:Fuwr6PBN}{}
-83. Neo4j.
+82. Neo4j.
 
 \hypertarget{ref-Theigraphsoftware:vh}{}
-84. Csardi G, Nepusz T. The igraph software package for complex network
+83. Csardi G, Nepusz T. The igraph software package for complex network
 research.

### 2017-12-28 revfiles/ReviewerResponse.md
Added rev responses.

--- 2017-08-11 revfiles/ReviewerResponse.md
@@
----
-title: Point-By-Point Reviewer Comment Response
-author: Hannigan *et al*.
----
+To editors of PLOS Computational Biology,
 
-We would like to thank the reviewers and editor for taking the time to read our manuscript and offer constructive feedback.
-We incorporated the suggested edits and feel that this greatly improved the manuscript.
-Below we included a point-by-point response to the reviewer comments, with our comments highlighted in *italics* and the relevant manuscript change locations highlighted as **bold**.
-Line numbers were included to guide the reviewers to the relevant sections.
-Additions to the text are highlighted in **blue** and removed text is highlighted in **red**.
+We would like to thank the editors and reviewers very much for the in-depth constructive and positive comments on our manuscript.
+We deeply appreciate the time of the editor and reviewers, and we think that the manuscript has benefited immensely from the feedback.
+We are resubmitting our manuscript with point-by-point responses to the reviewer comments below.
+We are also submitting a marked up revised manuscript with additions marked in blue and deletions marked in red.
+We included four additional figures, additional data analyses which are included in the text, and further discussion and clarification in the text.
 
-# Reviewer #1 (Comments for the Author):
+There were many good points brought up by the reviewers.
+One of the primary themes we did not intend to communicate was a focus on the random forest method itself instead of the biological insights that we gained.
+We developed our algorithm to predict phage-bacteria interactions to meet the specific needs of our biological hypotheses.
+While it did improve on other existing methods and performed well, we omitted some components that would otherwise be expected in a methods-centric manuscript.
+For example, we did not package it for widespread use in other studies (although the source code is available and reproducible).
+Additionally, while it is important to consider how our method performed in the context of existing methods, we do not wish to claim superiority of our method.
+It served our purposes and performs as well as other methods.
 
-Hannigan et al.
-describe a machine learning-based approach for predicting phage-bacterial host interactions and use it to build ecological networks for human microbiome samples based on published data.
-They analyze differences in network features as a function of host nutrition (low fat vs high fat), health status (lean vs obese twins) and body regions (diverse skin sites).
-The authors built their prediction model using data from previous publications as training set, containing 43 diverse bacterial species and 30 diverse phage strains, where the infectiveness relationship had been established experimentally (Fig 1A).
-The evaluation of the resulting predictive model shows that the amino acid similarity between phage and bacteria was the strongest predictor of infectiveness, followed by nucleotide similarity, shared protein families, and to a negligible degree CRISPRs (Fig 1B).
-They attribute the lack of predictive power of the latter to the minimal amounts of CRISPRs identified.
+The other major points of revision focused on more clearly benchmarking and communicating the performance of our model and the resulting network/sub-networks.
+To this end, we added additional data, figures, discussion and descriptions.
+The manuscript is certainly more comprehensive and robust as a result of the suggested revisions that we performed.
 
-The authors applied their model to three publicly available datasets (skin microbiome, twin gut microbiome of lean and obese twins and their mothers, and effect of diet on the gut microbiome) that contain sequences from full metagenomics as well as from purified viral-like particle fractions.
-Relationships between 298 phages operational genomic units (OGU) and 280 bacteria OGU were predicted from the combined taxa of all datasets and overall network features quantified (Fig1 D-G).
-Using the phage-bacteria network analysis, they explore the impact of diet and obesity in the gut microbiome structure (Fig 2), compare similarities of intrapersonal and interpersonal networks of skin and gut microbiomes (Fig 3) and test for differences between skin sites with different environmental qualities (Fig 4).
-All these comparisons are performed by either using measurements of network dissimilarity and network topology (eigen centrality, closeness centrality and degree centrality).
+# Reviewer #1: Summary
 
-Based on their results, the authors conclude that changes in diet, health state and body site topology can influence the phage-bacterial network stability.
-As an example they show that high-fat diet and obesity as well as external exposure of skin microbiota to reduce network connectivity.
-They also find network similarities to agree with the well-accepted notion of differences to be smaller within than between individuals.
+The authors developed a pipeline that predicts interactions between phages and their hosts from metagenomic data by exploiting similarities on the gene and protein level as well as CRISPRs with the random forest machine learning technique.
+They then compared the resulting phage-host networks across persons, time points and body sites.
+In general, this study is a valuable contribution, however some aspects of the network analysis are questionable, as detailed below.
 
-The approach taken by Hannigan et al.
-to construct phage-bacteria networks to analyze host-associated communities is original and will be relevant to many researchers within and also outside of the field of human microbiome research.
-One missing point the description on how the authors assessed the amount of bacterial DNA in the viral-like particle (VLP) fraction.
-If bacterial DNA were wrongly labeled as being of viral origin, it would lead to the formation of false phage OGU and result in relationships based on bacterial co-occurrences rather than phage infectivity.
-Also, the reader is not informed how abundances in the VLP fraction where integrated with the abundances of microbial fractions.
-In other words, if the DNA of the whole sample was not extracted together, how can the relative abundance of each fraction be compared to each other?
+## Major
 
-The presented results on the impact of diet and obesity, however, are severely underpowered, as the authors also point out (159, 187, 236).
-In particular, only one or two data points per group were used to compare the influence of diet and obesity (Figure 2).
-Although intra- vs.
-inter-individual and intra- vs.
-inter-family differences cannot be confirmed by statistical tests, they are claimed to exist.
-This is unfortunate, since the analytical approach using appropriately powered statistics reveals significant differences between skin sites, highlighting the potential power and impact of the methodology.
+I am suspicious of node removal studies.
+Even though the accuracy of the authors' inference procedure is much higher than those based on co-occurrence, the resulting networks still contain a number of false positives that can bias the node removal results.
+Even if the networks were perfect, where is the experimental evidence that an ecosystem will be less stable when separated into several components?
+Also, less stable in what sense?
+There are multiple definitions of ecological stability.
+Which one do the authors have in mind?
 
-Taken together, the machine leaning-based approach would only require some additional tests to exclude or estimate the degree of bacterial contamination in the VLP fraction and some explanations on how abundances in VLP and microbial fractions were combined and normalized.
-With regard to the application of the methodology, two of the three datasets are severely underpowered to support major claims on the impact of diet and health state.
-I would thus recommend to put the focus of the manuscript on the analysis of the well-powered skin data set and reduce the diet and obesity-related part as potentially promising case-examples while clearly describing the limitations of the analysis (which is done) and toning down or removing some of the respective claims.
+> **This is an important point and we agree with the reviewer that this warrants further clarification.
+Stability was not the best term here, but rather we think *robustness* better communicates our intent and interpretation.
+Previous work has shown that high connectance of an ecological network is associated with resistance to species extinction events and greater support of interactions through the network, such as horizontal gene transfer.
+For example, if a phage only has one host in a system, the removal of that host is expected to result in the extinction of the phage.
+If that phage had multiple hosts in a system, the removal of one host is less likely to result in extinction of that phage.
+We referred to this as "stability" in our manuscript, but we think that robustness is clearer.
+We went back and formally defined our use of "robustness" in the manuscript and updated the text to more clearly articulate this point.
+We had included citations around this relationship between connectance and robustness, and we added text to make this clearer in the manuscript as well.
+LINES 23, 33-34, 44, 65-66, 92-93, 221-224, 262, 316-317, 328, 336-343, 402-404**
 
-## Major Points
+In the same context:
+P.
+14, l.
+262: "These findings suggested that the community properties inferred from microbiome interaction network structures, such as stability (meaning a more highly connected network  is more stable because a randomly removed bacteria or phage node is less likely to divide or disintegrate [30,58] the overall network."
+The authors define a network with many components after species removal as unstable and a network with a single component as stable.
+To avoid a tautology here, an external criterion of stability is needed.
+Assuming an appropriate external criterion is available (e.g.
+number of species that die upon removal of a species), is there any experimental evidence that species whose removal results in more components destabilize an ecosystem more than species whose removal results in fewer components?
+
+> **Similar to above, we are talking more about robustness of the network to the random removal of nodes (phages and bacteria) instead of true ecological stability, which is a point we needed to make clearer.
+Previous work has shown that less highly connected networks are more likely to experience extinction events resulting from random microbe removal, as we discussed in the point above.
+Additionally, interactions such as horizontal gene transfer are more limited when there are less "connections" throughout the network.
+Previous work has supported these model assumptions and we cited some of those relevant pieces of work in the manuscript.
+LINES 23, 33-34, 44, 65-66, 92-93, 221-224, 261 - 262, 316-317, 328, 336-343, 402-404**
+
+I appreciate that the authors mention that no statement of significance can be made with the small sample numbers in the high-fat/low-fat and obese/healthy groups, but then why comparing network properties between these groups at all?
+
+> **The trend is the same across all of the samples for high and low fat diets, although this was only an opportunistic observation made using an existing data set underpowered for answering that question.
+Even though the observation is made in every instance, we cannot be certain of the results because of the small sample size.
+In other words, we are underpowered for achieving statistical significance.
+We feel this is still a worthwhile observation that appears to reflect a trend that may provide statistical confidence upon further sampling in a study properly powered to answer this question.
+LINES 228-230, 248-249**
+
+I am curious to know whether the predicted phage-host networks are nested and/or modular (see e.g.
+the Weitz lab papers cited by the authors).
+Also, what is the number of phages with 1, 2, ...
+N hosts and the number of hosts with 1, 2 ...N phages?
+Do these properties differ from what was observed in the oceans (Flores et al.
+2013 cited by the authors)?
+
+> **This is a good point and while we did show these statistics for the reference set used to train the model, we did not show it from our samples in the original submission.
+We added a figure that shows a heat-map of the bacteria and phage interactions in our master network.
+As we alluded to in our manuscript, we are hesitant to make strong comparisons to those other datasets because they were using experimental systems of defined strain-level interactions, which is a resolution we do not feel confident in for our study.
+Speculatively the infection patterns do seem to suggest a modular nested pattern which has been observed in those previous studies.
+We included a figure to this point, as well as some more text in the manuscript.
+Figure S9, LINES 186-188, 391-393**
+
+In this context: why not showing in Figure 1D a bipartite matrix in the style of Flores et al.?
+This would be far more informative than the current visualization, from which we learn nothing much, not even the number of bacteria and phages involved in interactions.
+
+> **We agree with the reviewer and would add that this figure has been the point of discussion and controversy both within our group and with others.
+The main purpose of this bipartite network visualization was to visually confirm that there were no erroneous connections between bacteria or between phages.
+To avoid further distraction by this figure, we decided to remove it and focus on the suggested matrix approach, which we added to the manuscript.
+Figure 1**
+
+When comparing networks across body sites, the authors want to consider not only network structure but also abundances.
+So the problem is to compare weighted networks.
+In the current formula, edges are only indirectly accounted for via centralities.
+I do not see how to justify this neglect of available information and think that a function that takes the entire edge set into account would be a better choice.
+How about for instance converting node weights into edge weights (e.g.
+by summing or averaging) and then summing the entry-wise differences between the corresponding (weighted) adjacency matrices into a single dissimilarity value?
+The adjacency matrices could include empty rows and columns for absent phages/bacteria.
+In this context, it would be interesting to know how the beta-diversity of the networks compares to the beta-diversity of the samples.
+
+> **We apologize for the confusion, but it sounds like the suggested approach is actually what we used.
+We calculated weighted centrality measures by converting the node weights into edge weights and used that to calculate our dissimilarity matrices.
+This way were were able to compare the beta-diversity of the networks, as suggested.**
+
+It would be interesting to find out whether phage-host interactions change between persons.
+Universality of microbial networks was the topic of the paper by Bashan et al.
+(https://www.nature.com/nature/journal/v534/n7606/abs/nature18301.html).
+Here, the authors could check whether hosts and phages present in all selected samples also interact in these samples and what might be the reason if they do not.
+
+> **This is a very interesting point and we appreciate the reviewer suggesting this addition to our work.
+The way we built our analysis was to build a universal network like was described by Bashan *et al* and utilize the subnetworks within.
+We performed our analysis this way because it allowed us to combine the genomic information from all studies to gain deeper coverage of the microbial genomes, as was done previously in the studies we utilized.
+Unfortunately that approach and the sample depths do not allow us to look at the networks beyond connectivity and abundance within the subgraphs, as suggested.
+Additionally, it would be important to include time course data to support these kinds of findings, as has been done in interaction modeling work by Weitz *et al* previously (DOI: 10.1098/rsos.160654).
+This is certainly an interesting direction of research and would be an excellent topic of future research.
+LINES 182-183, 252, 338-339**
+
+## Minor
+
+How was (intermittent) occlusion in skin sites determined?
+
+> **This was defined in the original publication by dermatologists and other previous work. We clarified this in the text. LINES 281-282, 286-288, **
+
+Do the most interconnected skin sites harbor a more diverse microbiome than other skin sites?
+
+> **This is an important point.
+As can be seen by the formulas of the metrics we used, the degree of network connectedness is still a function of bacterial and phage diversity.
+We don't interpret these metrics as mutually exclusive, but rather are ultimately using a metric that integrates the diversity of both the bacterial and viral communities.
+We included a clarification of this point in the manuscript.
+LINES 282-285**
+
+P. 10, l. 200-201: "These results suggested that the obesity-associated networks are less connected, having
+microbes further from all other microbes within the community."
+This statement is problematic.
+In addition to the problem of small sample numbers, interactions between the hosts themselves are not considered here.
+For instance, the overall impact of phages on gene exchange in the microbial community may be small compared to other avenues of horizontal gene transfer.
+
+> **These are good points and we reduced the interpretations that we outlined in our text. LINES 205, 241-249**
+
+P. 14, l. 262: "Gut network structure was not conserved among family members"
+The analysis of the authors does not differentiate between the change of community composition and the change of network structure.
+
+> **Similar to the point above, the network structure is a function of community composition of the two communities, as is evident in the formulas described in our methods.
+We clarified this point in the manuscript.
+LINES 282-285**
+
+# Reviewer #2
+
+## Summary
+
+The manuscript by Hannigan et al.
+describes a new methodological approach to study the interactions of phage and bacterial communities, which they propose to infer using machine learning approaches trained on known positive and negative interactions of bacteria and phages using different genomic features (nucleotide similarity, amino acid similarity, CRISPR cassettes and known protein-protein interactions).
+When applied to experimental metagenomic sets of viral and bacterial communities (with contigs clustered in bins or OGU), they were able to use their trained machine learning algorithm to find links between viral and bacterial bins, which are supposed to represent an infective interaction.
+With all links found between all OGU, they could build interactive networks and finally use different structural or topological features from the resulting networks to conduct microbial ecology analyses.
+
+I overall found the article interesting and I think that it represent a valuable contribution to the long lasting and largely unsolved problem of how to study the interplay between viral and microbial communities in metagenomic studies.
+Thus, I think that the proposed methodology will be well welcomed by the scientific community and is certainly within the scope of Plos Computational Biology.
+Although I found also interesting the biological results obtained from the analyzed datasets, I found some of them a bit weak.
+However, I think that the authors did a very honest and accurate work at the time of highlighting all the limitations throughout the manuscript and most of the concerns I rose while reading, were eventually replied by the authors themselves while moving forward in the reading.
+I highly appreciated that and I think that presented this way is fine and scientifically sounding.
+
+The article was very well written, explained and scientifically solid.
+
+I have few comments to make, but I don’t necessarily consider all of them critical to be potentially ready for publication.
+
+1) The authors do not test whether the interactions between viral and bacterial OGU predicted by their method are likely true or not (i.e., neither they tried to estimate the presence of false positives in the predicted phage-bacteria interactions nor at least pick up some examples and take a closer look to understand why OGUs were connected and whether this makes biological sense).
+Although I don’t consider it essential, I think that the methodology would gain credibility if they could just show few examples of which kind of genomic sequences were linked and why.
+I would suggest to pick examples of different nature, e.g., highly or poorly connected nodes, nodes with large or short edges, etc.
+
+> **We appreciate this point by the reviewer.
+We apologize for the confusion but we tested whether interactions are likely to be true or not using nested k-fold cross validation approach to determine the false positives and false negatives associated with our model, which are the inverse values of the sensitivity and specificity that we reported for our model.
+Using this we were able to quantify the degree to which our predicted positive and negative interactions were correct.
+We liked the idea of including an example so we performed a tblastx search for phage elements within our OGUs that matched the broadly infectious phages isolated from Lake Michigan by Putonti *et al*.
+We found some broadly infectious OGUs with genomic elements matching these broadly infectious phages, and included that example in the manuscript.
+LINES 167-169, 202-204**
+
+Since the training set was done with a very limited number of genomes (43 bacteria, 30 phages), from highly diverse (ecological) sources (something that it is also acknowledged by authors in lines 288-290), it is somehow hard to imagine how it would perform when tested on complex and highly diverse microbial communities, where most of the bacterial, and specially the viral genomes are highly divergent from those in databases.
+
+> **This is a very good point and brings up some of the points that we tried to address in our analyses.
+As we mentioned in our manuscript, this is an important caveat to our approach and why our insights into these dynamics will improve with more robust training set data.
+The point on the databases is also particularly important.
+These are indeed highly diverse microbial communities that are highly divergent from those in databases.
+This was a driving reason for why we chose to base our prediction algorithm on comparisons between phage and bacterial sequences from the samples themselves, instead of relying on alignments to existing reference microbes or the presence of identified marker genes.
+All of these are important caveats and we added these points to the manuscript text.
+LINES 164-167**
+
+Although authors give an explanation for the high abundance (77%) of bacterial hits in the viral datasets (i.e., reference bacterial genomes contain integrated prophages and the VLP largely correspond to temperate phages sharing elements with those prophages), I believe that in any viral purification protocol there is an important proportion of non-viral contaminant DNA that gets finally sequenced.
+Although the nature of bacterial hits in viral metagenomes is an ongoing debate or, likely, an inevitable technical issue, I think that the controls I mentioned above may be a complementary information to validate that the links found are truly bacterial-phage interactions and not unwanted bacteria-bacteria links (or even virus-virus).
+I think there are probably many different ways to make these controls, but one possibility (that may not be enough by itself anyway) could be to run softwares that identify core bacterial genes (e.g., MetaPhlan2) on the viral contigs and show that the proportion of annotated contigs is significantly low.
+
+> **This is also a very important point by the reviewer and we agree that it warrants further dedicated discussion in our manuscript.
+As pointed out, the inclusion of some contaminating sequences in these datasets is an inevitable technical issue that we must consider in our interpretations.
+This was shown nicely by Roux *et al* in a previous study, which we cited in our text.
+The group found that viral purification methods only reduce the level of bacterial DNA in the samples and do not completely remove them.
+Like what was done in Roux *et al*, and as suggested by the reviewer, we went on to identify a low abundance of 16S genes in the dataset which we would expect to be found only in bacterial sequences.
+This confirms that while the QC mentioned suggests a high level purification, there remains a low level of bacterial background noise.
+We added this data and discussion to our manuscript.
+LINES 120-140**
+
+> **To better address this and other points regarding the purity and accuracy of our dataset modeling, we also wanted to include a concept which has been utilized in other classic types of microbiome analyses, which is our tolerance for error.
+In our study, we are careful to not make any conclusions about the networks themselves, but instead rely on comparisons to other groups which were treated in the same way.
+This is similar to the classic approach taken in the field of the skin microbiome, whose V4 region 16S analysis could not detect some of the most abundant and important bacteria (*Propionibacterium*).
+The findings were still informative because the conclusions were not on the *true* composition of those communities, but rather comparisons of overall community structure between the study groups.
+This is an approach we also aimed to take with our work, so as to most appropriately interpret our data without the assumption that our model or data are perfect.
+We included a discussion of this in the manuscript as well.
+LINES 354-372**
+
+2) Which would be the results of the different analyses if networks were made only with the OGUs from whole shotgun datasets or the OGUs from viral datasets?
+(e.g., would the degree of centrality in the networks from low-fat diets be still lower than the high-fat?) If the observed patterns and interactions are intrinsic of viral and bacterial interplay, results should be markedly different when data from one of both communities is absent.
+
+> **It sounds like the idea here is to include direct phage and bacterial interactions in our analyses.
+We expect it would be difficult to identify direct interactions between bacteriophages because they do not have their own metabolic capacity without their bacterial hosts.
+We could consider modeling bacteria-bacteria interactions, but this is a non-trivial task that we feel is beyond the scope of our current study.
+Regardless, because this was not a consideration in our network analysis, we are not sure how bacterial interplay could be impacting the observations we made.
+This would add another interesting layer to future analyses along these lines.**
+
+3) May the Figure 1D benefit if the nodes from each of the three studies are colored differently?
+
+> **This is a good suggestion but, as we mentioned before, we are worried that Figure 1D is a point of distraction in the manuscript.
+We replaced this with a heatmap matrix visualization.
+Additionally, because the nodes are all part of the master network which includes data from all studies, the nodes could be associated with more than one study.
+FIGURE 1**
+
+4) Line 131-133: Although I think that taking into account the negative interactions is interesting, I also think that this is a challenging point.
+While a positive interaction is a strong proof that such viral-host interactions may also occur in natural conditions, I think that negative interactions are more delicate since the infection may not be observed just because the right experimental conditions were not found.
+I am not saying that negative interactions are not informative, just suggesting that maybe they should have less weight than positive ones.
+
+> **We are really glad to hear this because we have thought a lot about this in our group as well.
+Many previous efforts to predict interactions between bacteria and phages have focused on the positive interactions, and many assume that a lack of evidence is sufficient to conclude a negative interaction, which is how accuracy has been determined.
+We agree that the lack of infection does not definitively show a lack of infectious capability and that we should be conscious of this.
+This was very informative for our calculations of model sensitivity and specificity.
+This is an important caveat that we added to the paper.
+Additionally, while it is important to consider what the true occurrence is in nature, we are focusing on drawing conclusions as comparisons between states such as different skin sites, as discussed above.
+LINES 154-159**
+
+5) Line 216/218, Figure 3A: The authors propose that, similarly to what happen with microbiome composition (i.e., microbiomes are more similar in the same individual over time, than between individuals), the interaction networks or network structures are also more similar within the same individual.
+How do we know that this is an intrinsic property of these phage-bacteria interactions and are not a direct consequence of their significantly different microbiome/viral compositions?
+
+> **This is another very good point which we also discussed above.
+We can see from our network analysis formulas that the interaction networks are both functions of the community diversity and composition, as well as their degrees of connectedness.
+Because the bacterial and viral communities were shown to be different under the conditions we studied (outlined well in previous publications), it is unsurprising that the networks, which are integrations of these datasets, resulted in different community dynamics.
+The value of the conclusions is two-fold.
+The first is that we confirm that the differences hold true when we integrate the bacterial and viral datasets.
+The second is that we show which of those are more or less connected than the other, not only showing that they are different.
+We attempted to make this point clearer in the text.
+LINES 282-285, 295-296 ADDED ADDITIONAL LINES**
+
+6) Line 226, 227: Can we really say that the gut microbiome network structure was “strongly” conserved within individuals?
+Regardless of the very low p-value, the figure 3D does not give me the impression of “strongly”, although certainly different.
+In fact differences seem to be lower than the observed in Figure 3B (which I agree with authors that were probably not significant due to the limited number of samples).
+I think that the “strongly” in Figure 3D based on the p-value, is just a consequence of the high number of samples, not of the magnitude of the difference.
+
+> **This is an important point that we agree should be addressed in the manuscript. We updated our wording around this point. LINES 274**
+
+Typos and minor details:
+
+- Figure 2 legend, line 5: change “the” for “them”
+
+> **We fixed this in the text.**
+
+- Line 429: add “on” graph structure or “graph structure information”
+
+> **We fixed this in the text.**
+
+- Line 300: I think there is a typo here: “in biases”
+
+> **We fixed this in the text.**
+
+# Reviewer #3
+
+## Summary
+
+The authors present a machine learning tool (random forest) that predicts the host of a phage sequence using a set of features derived from sequence information only (nucleotide similarity, amino-acid similarity, CRISPR and similarity to known phage-host interacting Pfam entries).
+The authors use their predictor to evaluate the structure of phage-host interaction networks at different body sites across several human cohorts.
+They draw conclusions relating some metrics of the phage-host networks to different attributes of the samples (cohabitant humans, skin sites, diet, obesity).
+
+## General appreciation
+
+The goal of the study is clear and well-stated: even though some studies relate phage or bacterial community independently to environmental conditions, the relation between those communities and their variation in their structure is understudied and could bring highly valuable new insights to better understand the mechanisms, stability, and stress-response of the microbiomes.
+However, to address this point, the reliability of the prediction of the bacteriophage-bacteria interactions is essential, as it serves as basis for the subsequent analysis.
+In my opinion the present study suffers from major flaws regarding both the evaluation of the performances of the machine learning method as well as the reliability of its application on previously published mategenomic data.
+
+## Major issues
+
+Comparison to existing methods:
+Existing methods to predict bacteriophage and bacteria interactions from sequence information would suit the purposes of the authors:
+
+- Villarroel, J., Kleinheinz, K. A., Jurtz, V. I., Zschach, H., Lund, O., Nielsen, M., & Larsen, M. V. (2016). HostPhinder: a phage host prediction tool. Viruses, 8(5), 116.
+- Galiez, C., Siebert, M., Enault, F., Vincent, J., & Söding, J. (2017). WIsH: who is the host? Predicting prokaryotic hosts from metagenomic phage contigs. Bioinformatics, 33(19), 3113-3114.
+- Ahlgren, N.
+A., Ren, J., Lu, Y.
+Y., Fuhrman, J.
+A., & Sun, F.
+(2016).
+Alignment-free oligonucleotide frequency dissimilarity measure improves prediction of hosts from metagenomically-derived viral sequences.
+Nucleic acids research, 45(1), 39-53.
+
+The authors have to show that the method they developed improves on the existing ones (which would be highly valuable for the scientific community) and/or should compare the networks and the associated metrics they would get using these established prediction methods.
+
+> **We thank the reviewer for this point.
+There are certainly many valuable phage-bacteria interaction prediction tools in the field, and many avenues for improvement.
+To the point that "The authors have to show that the method they developed improves on the existing ones", we agree this would be important if the purpose of our manuscript was to offer a new and improved tool, but we do not intend this to be a method paper.
+We intended this to be evidenced by the lack of packaging of the model into an easily distributed tool, or the focus on the tool beyond a means to an end of understanding underlying biology.
+This is an important point for us to make clearer in our manuscript, so we added a discussion around this.
+We added this point to our discussion, in addition to referencing the papers.
+LINES 24-26, 78-87, 164-167**
+
+Benchmark of the method:
 
 1.
-Conceptual distinction between infectiveness and community stability: For the development of the network approach, the authors use infectiveness data as evidence for interactions between phage and bacteria.
-For the analysis part, such interactions are used to describe community stability (i.e., more interactions per node, higher stability).
-If phages are considered as agents grazing on their microbial hosts, is it justified to use degree of infectiveness analogous to community stability?
-Some discussion or background information would be helpful.
+The size of the benchmark is too small.
+The authors suggest that the need for having reliable negative bacteriophage-bacteria interaction information together with the positive-negative balance reduces the training set to 43 bacterial genomes and 30 phage strains.
+If the authors think that fulfilling these criteria at the price of a small training set is worth, they should show that the predictions are actually good on bigger datasets, at least for the evaluating if the known positive interactions are correctly recalled (in which case there is no lack of data).
+As in bigger dataset there will be more negative interactions than positive interactions, the precision of the method (and not the specificity) should be reported.
 
-*Thank you for raising this point which warrants further discussion.
-We clarified in the text why we feel this is justified, as the degree of stability refers to the reduced likelihood that a randomly removed node will cause a major divide or lack of connection between agents within the network.
-The infectiveness refers to the link we drew between the bacteria and phages (how we connected phages and bacteria in our networks).
-We also provided citations to work that further explains this in metabolic and phage-bacteria systems, if the readers are interested in a more in-depth discussion of the topic.* **LINES 176 - 183**
+> **This is an important point and well taken.
+We also had many internal discussions around this.
+The problem was deciding between a smaller but balanced dataset and a large but highly unbalanced dataset.
+The issue we had with an unbalanced dataset is that a highly accurate model could be developed by always stating that interactions are positive and not negative, which is not what we want.
+Going through the exercise, we would then move to an under-sampling approach to balance the datasets, which is what we presented in our paper.
+Imputing data through oversampling methods would be problematic because of the very highly imbalanced nature of the dataset.
+The addition of a larger dataset sounds great, and if we were aware of such a more robust and balanced dataset we would have loved to include it in our model creation and validation.
+It is true that we could evaluate our model on all of the known positive interactions, but we would still suffer from a lack of negative interaction data.
+LINES 154-159, 164-167**
 
 2.
-Testing for bacterial DNA in the phage fraction: The authors do not describe how bacterial sequences in the VLP fraction were eliminated.
-If bacterial sequences were present, they would be labeled as phages and lead to false predictions.
+Moreover, since the diversity of the cultivated phages as well as the cultivated bacteria is very low in comparison to real species found in the environment, the authors should evaluate their prediction on reliable known interaction detected in metagenomic samples (e.g.
+Paez-Espino, D., Eloe-Fadrosh, E.
+A., Pavlopoulos, G.
+A., Thomas, A.
+D., Huntemann, M., Mikhailova, N., ...
+& Kyrpides, N.
+C.
+(2016).
+Uncovering Earth’s virome.
+Nature.)
 
-*We agree that this needed to be better outlined in the text.
-It is important that we make clear that we are not violating our assumption that that phage OGU pool represents phages and the bacterial OGU pool represented bacteria.
-We updated the manuscript to address this concern in multiple ways.
-First, because this was a post hoc analysis of published data, we do not have access to the samples for performing further molecular QC.
-The original study authors did however perform and report a variety of QC measures to confirm the purity of their virome sample sets.
-To this end, we summarized those QC results into Table S1, which we added to the manuscript and referenced in the results.
-Furthermore, we performed our own local alignment (blast) QC protocol to assess the degree to which our OGUs were contaminated.
-We found that many phages had genomic elements similar to reference bacteria genomes, which agrees with the previous findings that the majority of the phages in these systems are temperate.
-We also identified two complete phages using the stringent Virsorter algorithm.
-Conversely, there were minimal phage reference genome elements identified in the bacterial OGUs, which also agrees with the previous findings that these sequences are mostly bacterial (cited in Table S1).
-We also did not identify any phages in the bacterial dataset using Virsorter.
-The lack of prophage identification, both here and in the previous study analyses, speaks to the need to sequence the bacterial fraction more deeply and create more robust assemblies (their larger size prevents assembly completion at the level that can be done with viruses).* **TABLE S1; LINES 93 - 119**
+> **This is similar to a point that we addressed above.
+This type of dataset would also be a great addition and the inclusion of such information would make our model much stronger, but we are unfortunately unaware of such a dataset.
+We also considered the suggested Paez-Espino *et al* manuscript but concluded this would be problematic for such an application because it is a whole shotgun metagenomic sequencing dataset whose interactions were inferred taxonomically.
+As much as we can tell, there was no experimental validation of the interactions, and no virus purification was attempted (all phage sequences were inferred from total metagenomes), so we were worried that we could not have any addition confidence in those findings beyond what we already have.
+If we included this approach, we would be benchmarking our predictions against other predictions.**
 
 3.
-Combining data from VLPs and microbial fractions: Information on how data from these two individual samples were matched and normalized is not provided.
-How are potentially different amounts of VLPs in different samples accounted for?
+On top of the previous remarks, the current benchmark shows how the tool perform only for full-length genomes.
+Since the presented application cases are dealing with metagenomic bins – likely to be incomplete and to suffer from contamination – it is necessary to evaluate the performances on simulated bins (namely simulate reads from phages and bacteria, run the same quality control+assembly+binning).
 
-*We are unclear what the reviewer was asking.
-Our goal was to use the edge weights to reflect samples that were highly relative abundant in both the bacterial metagenome and the phage metagenome.
-The edge weight used was therefore a multiplication function of the two relative abundance values, yielding a high value for those that were highly abundant in both, and low values for those that were not.
-This is described in the methods section of the manuscript.
-These values were recorded on a per-sample basis (within the subgraphs) so that each potential combination is accounted for, within each sampling pair.
-Furthermore, it may be worth noting that the bacterial and viral DNA extractions were done separately so as to maximize purity of the samples.
-This approach allows us to more accurately measure the relative abundance in the bacterial and especially viral samples, as has been noted previously.
-We clarified this in the text and added more relevant citations.* **LINES 149 - 153**
+> **We appreciate this point.
+This gets back to the previous point about how it would be ideal if we could build the model on metagenomic data.
+Unfortunately, we are not aware of a reliable and experimentally validated metagenomic dataset that we could use for such an application.
+Simulating such datasets from the reference genomes would be great, and we considered it, but we are unsure of how to effectively accomplish the task.
+As pointed out, it would be trivial to simulate contig sampling from the reference genome, but as we pointed out in the manuscript, the metagenomic binning of operational genomic units was also done with co-occurrence information from contig abundance.
+We are open to suggestions, but we are unsure about how to effectively simulate co-occurrence information because we are unsure about what empirical data that model would be built on.
+We could create co-occurrence matrices for the simulation binning by giving high and low counts to shared contigs, but this would not include informative error rates or reflect a real sample.
+In the end, this approach results in our predictive model being built on a simulated dataset that does not reflect reality any better, and had biases introduced by us for co-occurrences.
+We added more text around this point to the manuscript.
+LINES 113-116**
 
 4.
-Datasets underpowered to sustain claims on diet and obesity: Although the authors point out that the data are underpowered, they still make strong claims and the analysis of these data as major points of this manuscript.
+Overall, and most importantly, the method seems to be biased toward a high sensitivity-low specificity trade-off (respectively the median values are 0.952 and 0.615 when cross-validating) meaning that in case of balanced dataset almost 40% of the predicted interactions are expected to be wrong.
+The application on the skin dataset leads to almost a fully connected graph, while only 60% of those might be real.
+The authors should discuss explicitly the impact of the false prediction on the subsequent drawn conclusions: since the score threshold (say “s_se”) used for predicting the interaction looks biased toward sensitivity, what threshold “s_sp” could be considered as a safe guard on the other extreme (high specificity, with potentially lower sensitivity)?
+what is the impact of changing the threshold from t_se to t_sp on the network metrics?
+To what extent does it change the conclusions on real data?
 
-For example, 
-a.  the authors describe (158-164) how diet affects the network structure, without using sufficient data points to perform any statistical tests.
-b.  the authors claim (166-169) a difference between obese and healthy individuals by using one or two data points per group.
-With respect to these analyses, the authors write for example, "We found evidence that diet was sufficient to alter gut microbiome network connectivity" (235), which is a clear overstatement and such claims should be modified here and elsewhere in the manuscript.
+> **We added additional data and figures to show that the overall values fluctuated over the nested cross-validation benchmarking and that the overall performance is satisfactory.
+Additionally we feel that this is the ideal balance because the true positive rate is very good and most of the interactions are positive.
+We also added more discussion around this point to the manuscript.
+FIGURE S5, FIGURE S6, LINES 170-175**
 
-*We appreciate this point referring to figure 2, and while we attempted to avoid too strong of claims, we realize we fell short.
-Our goal was to present these as interesting observations, somewhat like a case study.
-We addressed this in the manuscript by further minimizing the effects of our claims and adding relevant caveats.* **LINES 29 - 31; 184 - 201; 268 - 277**
+Application to previously sampled cohorts:
 
-## Minor Points
+Since the number of samples on real-case applications is rather limited, a null model comparison should be used to assess the significance of the findings.
+For instance, in the case of the obesity-associated networks, the authors could randomly select the same number of nodes in the network as for each of the N=1+2 cases, and report the corresponding centrality values (mean + standard deviation across different random selections) and include it in Figure 2.
 
-* Throughout the manuscript, it is hard to follow how which and how many data points were used for each analysis.
-A supplementary table listing the samples for each of the studies and providing sample sizes in each figure would be very helpful.
+> **This is an interesting idea.
+The null model seems like it would be useful for showing a difference to a random model, but we are more interested in comparing the obesity and non-obese states.
+We only intended the obesity data to be an opportunistic observation of interest, and not a formal finding.
+While it is interesting and warrants further investigation, we recognize that this is only an N=1+2 dataset and we are ultimately going to need more sampling to truly understand the differences in network structure between healthy and obese individuals.**
 
-*We added the relevant sample size information to the text and figure legends to make clearer what the data actually represent.* **LINES 186; 194 - 196; 219 - 221; 223 - 225; FIGURE LEGENDS 2,3,4**
+## Minor issues
 
-* Figure 1D is not informative and could be removed since the information is provided in the main text.
+Lines 82-83: this is a strong overstatement since it is only a hypothetical interpretation of the network metrics as mentioned for example in lines 214-215.
 
-*It is true that the information is in the text, but we feel this is still an important panel to the figure because it illustrates the degree of connectance between the phage and bacterial communities, and provides a visualization to the text included in the results section.
-Even more importantly, we intended this panel to act as a visual validation that there are no erroneous connections among phage and bacterial nodes.*
+> **We toned down our language here.**
 
-* Figure 1 EFG: color labels in legend are wrong and not since y-axis labels are present.
+Lines 107-114: It is not clear at first sight that the paragraph is describing the phage fraction.
 
-*Thank you for noting this mistake. We have fixed it.* **FIGURE LEGEND 1**
+> **We agree and clarified this in the text.**
 
-* Figure 2 could use headers above the plots to indicate which one relate to impact of diet (AB) and which one to obesity (CD).
+Lines 115-120: It is not clear at first sight that the paragraph is describing the bacterial fraction.
 
-*This was especially confusing because of the mislabeled axes. We fixed the axis labels to make the data clearer.* **FIGURE 2**
+> **We agree and clarified this in the text.**
 
-* Figure 2 panels C and D have wrong x-axis labels.
-Should be "obese" and "healthy".
-Here it is not clear why only 2 and 1 data points are available.
-The original data in Reyes et al.
-seem to consist of 32 fecal samples from 13 individuals from 5 families with at least two time points.
+Lines 134-138: the description of the methods lacks some important description.
+It is form example not clear if the nucleotide similarity feature is always computed as the best bitscore obtained on the training bacterial dataset, or if it uses another external bacterial dataset.
+The same remarks goes for the amino-acid similarity feature.
 
-*We fixed the error, and made the causes for the sample size clearer in the text as well as the figure legends. Thank you for pointing this out.* **LINES 194 - 195; FIGURE LEGEND 2**
+> **We agree that the description here is cursory, as we left the reader to see the methods section for more details.
+No external datasets were used, and we clarified this point in the text.
+ADD LINES HERE**
 
-* Figure 3A. Not clear why only 8 data points are available if Minot et al. report 15 samples from 5 individuals.
+Line 140: The authors report the median values for sensitivity and specificity when cross-validating their method, but do not report the standard deviations.
+These standard deviations are of high importance to get an accurate feeling of the reliability of the method.
 
-*We only used the subset of data points that had temporal data spaced approximately the same distance apart. We made this clearer in the text.* **LINES 186 - 187; 219 - 221; FIGURE LEGEND 1,2**
+> **We agree and added this data as figures and text in the manuscript. FIGURE S5, FIGURE S6**
 
-* Figure 3B.
-Not clear how data were computed.
-If the comparison of within individuals (i.e.
-1 value if same data as in 3A) with mean of all other individuals, then standard deviations should be shown.
-Suggest to show only the smallest inter-individual distance, which would greatly strengthen the point on individuality.
+Line 159: the diameter of the graph seems not to be a robust measure of the graph.
+Indeed, a single spurious branch popping out from the graph will change the value diameter.
+It would be more informative to report some quantile value for the number of traversed vertices between two vertices.
 
-*Yes, this is a comparison between the interpersonal distance, and the distance between the person's samples and all other samples.
-We considered calculating the distance differences using standard deviation and pairing the samples as every interpersonal comparison instead of the average.
-The problem we found was that that pairwise comparison inflated the significance values by artificially creating a large n, all of which were not truly independent.
-We therefore collapsed the data points into a single average instead of pairing the same sample to many other samples.
-Showing the smallest distance is also an interesting idea, but we are worried that this may also skew the data by failing to account for the variation in the sample distributions.*
+> **This is a valuable point.
+We added the suggested data as a set of historgrams outlining the distributions of node eccentricities associated with each subnetwork and the master network.
+LINES 198-199, FIGURE S8**
 
-* Figure 3C. Not clear which data were used, in particular if same data as in Figure 2 were used. E.g., were only the twins included, or also the mothers?
+Lines 162-163: the percentage of the recalled interactions could be mentioned.
 
-*We made this information clearer in the text.* **LINES 223 - 225; FIGURE LEGEND 3**
+> **We unfortunately are unsure of what the recalled interactions are referring to.**
 
-* Figure 3D.
-As described for 4B, smallest inter-individual distance should be used if showing individuality should be demonstrated, which is very different from showing on average higher similarity within than between individuals.
+Line 186: it is not clear what is the limited sample set.
 
-*We addressed this point in the related bullet point above.*
+> **The word limited was confusing here. Instead we used small to communicate the caveat that this is a data mining exercise using a small dataset. LINES 231**
 
-* Figure 4. Please provide sample numbers for each of the boxplots and data in 4E.
+Line 340-347: The description of the data is too limited.
+We do not find for example clearly how many gut samples per individual were included (since on line 223 it is implicit that there are several samples per individual).
 
-*We added this information to the figure legend.* **FIGURE LEGEND 4**.
-
-# Reviewer #2 (Comments for the Author):
-
-The work presented in the manuscript entitled "Biogeography & Environmental Conditions Shape Phage & Bacteria Interaction Networks Across the Human Microbiome" by Hannigan et al.
-consists on prediction of host-bacteriophage interactions from skin and gut microbiome public datasets and graph-based analysis of the underlying phage-host networks, assessing the effects of associated metadata (obesity, diet, type of skin site).
-For the prediction of interactions, the authors trained a random forest classifier using known phage-bacteria interactions using the following features: CRISPR-spacer detection, prophage prediction based on nucleotide similarity (blastn, e-value <=1e-10.), shared gene identification based on aminoacid similarity (Diamond) and detection of pairs of genes whose proteins interact (Pfam interaction information within the Intact database).
-My main concern with this paper is on the prediction of phage-host interactions.
-The authors report 72287 infectious relationships among 578 nodes, representing 298 phages and 280 bacteria.
-If all phages would infect all bacteria, the network would have 83440 links (complete network = total possible interactions).
-The reported network contains 86,6% of the number of possible interactions.
-This essentially means that most phages are generalists and can infect any bacteria...The authors fail to report the number of phages and bacteria per microbiome: skin or gut.
-This number would allow us to estimate the total number of possible interactions within an environment (gut or skin), a more relevant number to calculate the density of the network.
-One explanation for the high density of the network is the use of aminoacid similarity between phage and bacteria as predictor.
-Shared genes have been used to link phages to their host, but at nucleotide rather than at aminoacid level (Edwards et al.
-FEMS Microbiol Rev.
-2016).
-Independently from the classifier performance, when confronted with this high network density, the shared genes should have been further analysed and the procedure revised.
-
-## Major comments:
-
-1.
-Network inference: operational genomic units were linked sample-wise, microbiome type wise or all microbiome together?
-Aminoacid similarity is not specific enough and is likely the cause of the high density of the network, which suggests that almost all phages infect all bacteria (86% density), meaning the resulting network is (likely) no statistically different from a random network with equal number of links.
-In the absence of further evidence and controls, I am not convinced with this result.
-
-*We would like to thank the reviewer for pointing out that this area needed clarification.
-The operational genomic units were linked sample-wise, by linking the OGUs found between each sample.
-The source of these samples were recorded, so that we knew whether they originated from the gut or the skin, etc.
-We also agree that amino acid similarity of genes along is not enough, and that is why we incorporated other similarity metrics in our random forest model.
-As we described, even though the amino acid similarity metric was the most "important" to the model, it was used within the context of the other metrics in the random forest.
-It is true that this is a dense network, and we hypothesize that that may be because each OGU does not necessarily represent a specific phage or bacterial strain, but is an operationally defined group of similar genomes, similar to the operation taxonomic unit (OTU) often used in microbiome studies.
-In other words, we are not using the OGUs as individual phage strains, but instead these OGUs are operational units that allow us to condense our dataset by contig similarity.* **SEE LINES 114 - 119; 304 - 322**
-
-2.
-Number of samples per study and metadata groups are not reported but the authors state about the diet effect: "Tests for statistical differences were not performed due to the small sample size ".
-If sample sizes are small for statistical analysis, no comparison should be made in the first place.
-The plots of Figure 2 feature 2 vs.
-3 comparison for high fat vs.
-low fat and 2 vs.
-1 for obese vs.
-healthy...are these the nb.
-of samples per group?
-Bacterial or phage diversity are likely confounders for network connectivity comparisons and this is overlooked here.
-
-*We appreciate this point and while we attempted to report this as more of a case trial observation (as outlined for the related comment above), we unintentionally used too strong of language, thus overstating the finding beyond our intentions.
-We also needed to clarify the analysis being performed.
-It is also true that the network analysis is a function of the diversity of the two communities, and we do not wish to make it sound like they are separate.
-We would be hesitant however to call this dependence a confounder since it is more building off of the diversity data and incorporating other additional information from the community.
-We made this clearer in the text.* **LINES 186; 194 - 196; 219 - 221; 223 - 225; FIGURE LEGENDS 2,3,4**
-
-3. Graph-analysis: Analysis of network nestedness and modularity (Weitz et al. Trends in Micr. 2012) may be more relevant than diameter, degree and closeness for phage-host networks.
-
-*We chose to omit such an analysis as was done in Weitz et al because, as we attempted to describe in the manuscript, we are not using the OGUs as individual phage strains as was done in the Weitz and other papers.
-Instead these OGUs are operational units that allow us to condense our dataset by contig similarity.
-We feel that making statements with as high of phylogenetic resolution as Weitz et al would be unfounded at this point.*
-
-## Other comments:
-
-1.
-Training set contained on 43 diverse bacterial species and 30 diverse phage strains.
-Given than these numbers are (much) lower than the ones published in references 41, 47-51, from which the data on known interactions was extracted, I wonder how the curation/filtering was done.
-
-*Our limitation was not on how many known interactions that could be found, but instead on how many non-interactions could be found.
-In other words, we needed to utilize known infectious pairs, as well as phages that have been validated as not infecting certain bacteria.
-Because we were limited to only a small number of validated non-interactions, we decided to also limit the number of positive interactions included, so as to prevent a large bias in the ratio between the two interaction types in the training dataset.
-We clarified that text.* **LINES 124 - 127**
-
-2.
-The sensitivity of the classifier is reported as 0.846.
-However, "Approximately one third of the training set relationships yielded no score and therefore were unable to be assigned an interaction prediction".
-This means the sensitivity can not be higher than 0.67...
-
-*Here we are reporting the accuracy for those values that were scored.
-Pairs without scores were omitted from this analysis, and we instead only focused on those that were scored.
-We have clarified the text.* **LINES 143 - 144; FIGURE LEGEND 1**
-
-3.  Dataset: The authors should report the number of samples per study and associated metadata (e.g low /high fat diet, obese/lean).
-
-*We have added this information.* **LINES 186; 194 - 196; 219 - 221; 223 - 225; FIGURE LEGENDS 2,3,4**
-
-4.  OGU construction, why the OGUs were constructed with CONCOCT without coverage information (CONCOCT-NC)?
-
-*We maintained the coverage information but omitted the feature of total sequence counts since we worried this would confound the results.
-This means that the feature set matrix used by the CONCOCT clustering algorithm only included k-mer frequencies and sequence count (relative abundance) values per sample.*
-
-5.
-Network construction: The master network contained the three studies as sub-networks, which themselves each contained sub-networks for each sample.
-The authors provide only numbers of phage and bacterial nodes for the master network.
-For predicting the associations, were all phages and bacteria considered disregarding the sample type (skin vs.
-gut)?
-
-*Yes, this is described correctly by the reviewer.*
-
-6.
-The phages and bacteria in the gut diet and twin sample sets were more sparsely related: each contained fewer than 150 vertices, fewer than 20,000 relationships".
-This is a bipartite network, what should be reported is the number of vertices of each type rather than the total nb of vertices (this is relevant to calculate the network density, we cannot know here what is the nb of possible interactions).
-
-*We added these values to the results section of the manuscript so that the reader can access this information.* **LINES 154 - 163**
-
-7.  Comparison with Paez-Espino et al. (Nature 2016, doi:10.1038/nature19094) phage-host human associated subnetworks should be done.
-
-*This is an intriguing suggestion.
-It sounds like this comment is a reference to the host range inferences within the suggested manuscript.
-Paez-Espino et al used CRISPR and tRNA comparisons to identify bacterial hosts for the bacteriophage genomes and genome fragments.
-The group's work suggested that phage host range is broader than has been traditionally appreciated.
-In our discussion section, we mentioned a comparison of these and related findings to our own dataset.
-This information was included in the operational genomic unit section of the discussion.* **LINES 318 - 322**
-
-8.  Figure 1. panel B: methods include blastx, in the network inference Diamond was used instead; Panel E-F, colors in figure do not correspond to those in legend.
-
-*It is true that we used the Diamond aligner, but we used the blastx algorithm within the Diamond aligner (written as `diamond blastx`).
-We made this information clearer in the text.
-We also fixed the color references.* **FIGURE 1**
-
-9.
-Figure 2 legend: it is not clear if/that the dots represent sample-wise networks.
-"Lines represent the mean degree of centrality for each diet", the only lines in the plots are parallel to the centrality axis, they are just splitting high fat vs.
-low fat diet fields in the x axis.
-X-axis from panels C and D are mislabeled.
-
-*We clarified the dot information, and also fixed the incorrect labels.* **FIGURE 2**
-
-# Reviewer #3 (Comments for the Author):
-
-This research article by Hannigan et.al., focuses on an important and indispensable part of the microbiome- "Virome".
-Interplay within inter and intra microbial community members is important to get the complete picture.
-This paper focuses on such interplay and understands that microbiome operates with both the viral and bacterial community members.
-For this study they worked on previously published metagenomics data, to not only infer based on nucleotide information but also based on protein interactions.
-
-## Main Concerns:
-
-The data from studies referenced 13 & 14 even though may have bias due to sequencing, have similarity in approach wherein collection of VLP and bacterial whole genome shotgun sequencing was performed.
-VLP data in reference 14 was based on Genomiphi amplification of VLP DNA prior to sequencing that is bound to introduce amplification and sequencing errors which would further lead to clustering bias in the current study.
-The concern is also regarding data from references 38 and 39 that are not only performed by 454 technology.
-Further on, reference 39 doesn't have VLP data available.
-The study is based on 16S and whole genome shotgun of gut microbiome.
-
-*We referenced this statement in reference 38: "We had previously performed shotgun sequencing of total fecal DNA isolated from the 12 frozen samples obtained at the first time point that were now used to prepare purified VLPs (Table S3)." This reference (12) led to our reference 39, which contained the data in the related SRA submission and was cited accordingly.
-Between these two published datasets, we were able to pair the whole shotgun sequencing data with the purified virome data.*
-
-The question is how were the above biases (including the kitome) taken into consideration while conducting these analyses.
-The authors may be better off carrying out the analysis only using data from references 13 and 14 for skin and only data from 38 for gut?
-
-*There certainly could have been extraction, kit, or other related differences between the 38 and 39 dataset.
-We feel these findings, which we attempt to present more as case observations that will require more dedicated study, are worth noting in the paper.
-We altered the text to better clarify our position.
-We also avoided analyses in which we directly compared datasets, because we are unable to discern between true biological differences and differences between the studies.
-Such analyses would require further and properly controlled sampling.*
-
-Even though the authors have acknowledged some of these biases including MDA, in their discussion, these concerns should be explicitly discussed and measures they have taken to overcome these.
-
-*We included a description of these biases associated with these methods.
-As far as measures to overcome these biases, we acknowledge the biases and only performed analyses within studies where samples were treated the same.
-The goal is for this work to use these findings as a jumping off point, and performing future studies with more limited biases from techniques such as MDA.* **LINES 295 - 303**
+> **We mentioned the numbers of samples used in each part of the analysis when we described them in the results section. We clarified that direction in the text. LINES 414-415**

### 2017-12-28 bin/CalculatePredModel.R
Updated analysis code.

@@ -26,6 +26,7 @@ library("cowplot")
 
 library(parallel)
 library(tidyr)
+library(dplyr)
 
 ###################
 # Set Subroutines #
@@ -111,10 +112,51 @@ nestedcv <- function(x, iterations = 5, split = 0.8) {
   return(c(maxl, medl, minl))
 }
 
-nestedVariance <- function(y, iter = 32) {
+nestedcvval <- function(x, iterations = 5, split = 0.8) {
+  outlist <- mclapply(1:iterations, mc.cores = 4, function(i) {
+    write(i, stdout())
+
+    trainIndex <- createDataPartition(
+      x$Interaction,
+      p = split, 
+      list = FALSE, 
+      times = 1
+    )
+    
+    dftrain <- x[trainIndex,]
+    dftest <- x[-trainIndex,]
+
+    # write(summary(dftrain$Interaction), stdout())
+    # write(summary(dftest$Interaction), stdout())
+    
+    outmodel <- caretmodel(dftrain)
+    
+    x_test <- dftest[,2:5]
+    y_test <- dftest[,1]
+    
+    outpred <- predict(outmodel, x_test, type="prob")
+    outpred$pred <- predict(outmodel, x_test)
+    outpred$obs <- y_test
+
+    # confusionMatrix(outpred, y_test)
+    # postResample(pred = outpred$pred, obs = outpred$obs)
+    sumout <- twoClassSummary(outpred, lev = levels(outpred$obs))
+    sumroc <- sumout[[1]]
+    sumsens <- sumout[[2]]
+    sumspec <- sumout[[3]]
+
+    return(list(sumroc, sumsens, sumspec))
+  })
+
+  finaldf <- data.frame(do.call("rbind", outlist))
+  colnames(finaldf) <- c("ROC", "Sens", "Spec")
+  return(finaldf)
+}
+
+nestedVariance <- function(y, iter = 100) {
   ol <- mclapply(1:iter, mc.cores = 4, function(j) {
     write(paste("Running Iteration ", j), stdout())
-    nr <- nestedcv(y, iterations = 25, split = 0.75)
+    nr <- nestedcv(y, iterations = 5, split = 0.8)
     nauc <- nr[[5]]
     nsens <- nr[[7]]
     nspec <- nr[[8]]
@@ -172,18 +214,77 @@ negativedf <- getresults(negativequerydata)
 dfbind <- rbind(positivedf, negativedf)
[…]

### 2017-12-28 bin/CompareSkin.R

@@ -1,7 +1,6 @@
 ##################
 # Load Libraries #
 ##################
-gcinfo(TRUE)
 packagelist <- c("RNeo4j", "ggplot2", "wesanderson", "igraph", "scales", "plyr", "cowplot", "vegan", "reshape2", "parallel", "stringr")
 new.packages <- packagelist[!(packagelist %in% installed.packages()[,"Package"])]
 if(length(new.packages)) install.packages(new.packages, repos='http://cran.us.r-project.org')
@@ -223,7 +222,7 @@ rcentralmerge <- merge(rcentraldf, locationmetadata, by.x = "location", by.y = "
 
 # Moisture Levels
 
-box_moist <- ggplot(rcentralmerge, aes(x = factor(moisture), y = cl)) +
+box_moist <- ggplot(rcentralmerge, aes(x = factor(moisture), y = ec)) +
 	theme_classic() +
 	geom_boxplot(notch = TRUE, fill="gray") +
 	ylab("Eigen Centrality") +
@@ -232,17 +231,17 @@ box_moist <- ggplot(rcentralmerge, aes(x = factor(moisture), y = cl)) +
 	    axis.line.x = element_line(colour = "black"),
 	    axis.line.y = element_line(colour = "black")
 	) +
-	ylim(0, 0.005) +
-	geom_segment(x = 1, xend = 2, y = 0.0045, yend = 0.0045) +
-	annotate("text", x = 1.5, y = 0.00455, label = "*", size = 6) +
-	geom_segment(x = 1, xend = 3, y = 0.0048, yend = 0.0048) +
-	annotate("text", x = 1.5, y = 0.00485, label = "*", size = 6)
+	ylim(0, 0.35) +
+	geom_segment(x = 1, xend = 2, y = 0.3, yend = 0.3) +
+	annotate("text", x = 1.5, y = 0.305, label = "*", size = 6) +
+	geom_segment(x = 1, xend = 3, y = 0.325, yend = 0.325) +
+	annotate("text", x = 2, y = 0.330, label = "*", size = 6)
 
 moistsig <- melt(pairwise.wilcox.test(x = rcentralmerge$ec, g = rcentralmerge$moisture)$p.value)
 
 # Occlusion Status
 
-box_occ <- ggplot(rcentralmerge, aes(x = factor(occlusion), y = cl)) +
[…]

### 2017-12-28 bin/GetNetworkLinkCounts.R

@@ -1,0 +1,147 @@
+##################
+# Load Libraries #
+##################
+packagelist <- c("RNeo4j", "ggplot2", "wesanderson", "igraph", "visNetwork", "scales", "plyr", "cowplot", "reshape2")
+new.packages <- packagelist[!(packagelist %in% installed.packages()[,"Package"])]
+if(length(new.packages)) install.packages(new.packages, repos='http://cran.us.r-project.org')
+lapply(packagelist, library, character.only = TRUE)
+library("ggraph")
+library("grid")
+library("stringr")
+
+suppressMessages(c(
+library("RNeo4j"),
+library("ggplot2"),
+library("C50"),
+library("caret"),
+library("wesanderson"),
+library("plotROC"),
+library("cowplot"),
+library("gplots")
+))
+
+###################
+# Set Subroutines #
+###################
+
+importgraphtodataframe <- function (
+graphconnection=graph,
+cypherquery=query,
+filter=0) {
+  write("Retrieving Cypher Query Results", stderr())
+  # Use cypher to get the edges
+  edges <- cypher(graphconnection, cypherquery)
+  # Filter out nodes with fewer edges than specified
+  if (filter > 0) {
+    # Remove the edges to singleton nodes
+    singlenodes <- ddply(edges, c("to"), summarize, length=length(to))
+    # # Subset because the it is not visible with all small clusters
+    # singlenodesremoved <- singlenodes[c(singlenodes$length > filter),]
+    multipleedge <- edges[c(which(edges$to %in% singlenodesremoved$to)),]
+  } else {
+    multipleedge <- edges
+  }
+  # Set nodes
+  nodes <- data.frame(id=unique(c(multipleedge$from, multipleedge$to)))
+  nodes$label <- nodes$id
+  return(list(nodes, multipleedge))
+}
+
+##############################
+# Run Analysis & Save Output #
+##############################
+
+# Start the connection to the graph
+# If you are getting a lack of permission, disable local permission on Neo4J
+graph <- startGraph("http://localhost:7474/db/data/", "neo4j", "root")
+
+# Get number of phages hitting bacteria
+query <- "
+MATCH (n)-[r]->(m)
[…]

### 2017-12-28 bin/VisualizeNetwork.R

@@ -54,12 +54,12 @@ filter=0) {
 plotnetwork <- function (nodeframe=nodeout, edgeframe=edgeout) {
   write("Preparing Data for Plotting", stderr())
   # Pull out the data for clustering
-  ig <- graph_from_data_frame(edgeout, directed=F)
+  ig <- igraph::graph_from_data_frame(edgeout, directed=F)
   # Set plot paramters
-  V(ig)$label <- ifelse(grepl("Bacteria", nodeout$id),
+  igraph::V(ig)$label <- ifelse(grepl("Bacteria", nodeout$id),
     "Bacteria",
     "Phage")
-  V(ig)$type <- ifelse(grepl("Bacteria", nodeout$id),
+  igraph::V(ig)$type <- ifelse(grepl("Bacteria", nodeout$id),
     TRUE,
     FALSE)
   # Create the plot
@@ -79,15 +79,15 @@ plotnetwork <- function (nodeframe=nodeout, edgeframe=edgeout) {
 wplotnetwork <- function (nodeframe=nodeout, edgeframe=edgeout) {
   write("Preparing Data for Plotting", stderr())
   # Pull out the data for clustering
-  ig <- graph_from_data_frame(edgeout, directed=F)
+  ig <- igraph::graph_from_data_frame(edgeout, directed=F)
   # Set plot paramters
-  V(ig)$label <- ifelse(grepl("Bacteria", nodeout$id),
+  igraph::V(ig)$label <- ifelse(grepl("Bacteria", nodeout$id),
     "Bacteria",
     "Phage")
-  V(ig)$type <- ifelse(grepl("Bacteria", nodeout$id),
+  igraph::V(ig)$type <- ifelse(grepl("Bacteria", nodeout$id),
     TRUE,
     FALSE)
-  V(ig)$weights <- nodeout$avg
+  igraph::V(ig)$weights <- nodeout$avg
   # Create the plot
   fres <- ggraph(ig, 'igraph', algorithm = 'bipartite') + 
         geom_edge_link0(edge_alpha = 0.01) +
@@ -103,21 +103,31 @@ wplotnetwork <- function (nodeframe=nodeout, edgeframe=edgeout) {
 graphDiameter <- function (nodeframe=nodeout, edgeframe=edgeout) {
   write("Calculating Graph Diameter", stderr())  
   # Pull out the data for clustering
-  ig <- graph_from_data_frame(edgeframe, directed=F)
[…]

### 2018-01-07 doc/manuscript.Rmd
Updated manuscript.

@@
 We used this information to calculate the average connectedness per sample, which was corrected for the maximum potential degree of connectedness.
-Unfortunately our dataset was insufficiently powered to make strong conclusions toward this hypothesis, but this is an important observation that warrants further investigation.
-Although the changes are not statistically significant using such a small sample set, visual inspection of data suggests there is an underlying significant difference that could be elucidated with further sampling.
+Unfortunately our dataset was insufficiently powered to make strong conclusions toward this hypothesis, but this is an interesting observation that warrants further investigation.
+This observation also serves to illustrate the types of questions we can answer with more comprehensive microbiome sampling and integrative analyses.
 

### 2018-01-07 doc/manuscript.tex
Updated manuscript.

@@
 insufficiently powered to make strong conclusions toward this
-hypothesis, but this is an important observation that warrants further
-investigation. Although the changes are not statistically significant
-using such a small sample set, visual inspection of data suggests there
-is an underlying significant difference that could be elucidated with
-further sampling.
+hypothesis, but this is an interesting observation that warrants further
+investigation. This observation also serves to illustrate the types of
+questions we can answer with more comprehensive microbiome sampling and
+integrative analyses.
 
@@
 \hypertarget{ref-FASTXToolkit:wr}{}
-79. Hannon GJ. FASTX-Toolkit.: GNU Affero General Public License.
+79. Hannon GJ. FASTX-Toolkit. 2010; GNU Affero General Public License.
 
@@
 \hypertarget{ref-caretClassificatio:ux5fU2Litux5f1}{}
-84. Kuhn M. caret: Classification and Regression Training.
+84. Kuhn M. caret: Classification and Regression Training. CRAN. 2016;
 
@@
 \hypertarget{ref-Neoj:Fuwr6PBN}{}
-88. Neo4j.
+88. Neo Technology, Inc. Neo4j. 2017;
 
@@
 89. Csardi G, Nepusz T. The igraph software package for complex network
-research.
+research. InterJournal. 2006;Complex Systems: 1695.
 

### 2018-01-07 revfiles/ReviewerResponse.md
Updated reviewer comments.

@@
 We had included citations around this relationship between connectance and robustness, and we added text to make this clearer in the manuscript as well.
-LINES 23, 33-34, 44, 65-66, 92-93, 221-224, 262, 316-317, 328, 336-343, 402-404**
+LINES 23, 33-34, 44, 65-66, 92-93, 220-223, 263, 317-318, 328-329, 337-344, 403-405**
 
@@
 Previous work has supported these model assumptions and we cited some of those relevant pieces of work in the manuscript.
-LINES 23, 33-34, 44, 65-66, 92-93, 221-224, 261 - 262, 316-317, 328, 336-343, 402-404**
+LINES 23, 33-34, 44, 65-66, 92-93, 220-223, 263, 317-318, 328-329, 337-344, 403-405**
 
@@
 
-> **The trend is the same across all of the samples for high and low fat diets, although this was only an opportunistic observation made using an existing data set underpowered for answering that question.
-Even though the observation is made in every instance, we cannot be certain of the results because of the small sample size.
-In other words, we are underpowered for achieving statistical significance.
-We feel this is still a worthwhile observation that appears to reflect a trend that may provide statistical confidence upon further sampling in a study properly powered to answer this question.
-LINES 228-230, 248-249**
+> **This analysis was meant to be descriptive and to primarily illustrate to the field what types of questions can be answered once we have more comprehensive datasets.
+We cannot be certain of the results because of the small sample size, but this illustrates the potential biology we can uncover with more comprehensive microbiome sampling and integrative analyses.
+LINES 230-231, 247-250**
 
@@
 We included a figure to this point, as well as some more text in the manuscript.
-Figure S9, LINES 186-188, 391-393**
+Figure S9, LINES 187-189, 391-394**
 
@@
 This is certainly an interesting direction of research and would be an excellent topic of future research.
-LINES 182-183, 252, 338-339**
+LINES 183-184, 253, 339-340**
 
@@
 We included a clarification of this point in the manuscript.
-LINES 282-285**
+LINES 282-289**
 
@@
 
-> **These are good points and we reduced the interpretations that we outlined in our text. LINES 205, 241-249**
+> **These are good points and we reduced the interpretations that we outlined in our text. LINES 206, 242-250**
 
@@
 We clarified this point in the manuscript.
-LINES 282-285**
+LINES 282-286**
 
@@
 We found some broadly infectious OGUs with genomic elements matching these broadly infectious phages, and included that example in the manuscript.
-LINES 167-169, 202-204**
+LINES 164-175, 203-205**
 
@@
 All of these are important caveats and we added these points to the manuscript text.
-LINES 164-167**
+LINES 164-168**
 
@@
 We included a discussion of this in the manuscript as well.
-LINES 354-372**
+LINES 353-373**
 
@@
 We attempted to make this point clearer in the text.
-LINES 282-285, 295-296 ADDED ADDITIONAL LINES**
+LINES 282-289, 296-297**
 
@@
 
-> **This is an important point that we agree should be addressed in the manuscript. We updated our wording around this point. LINES 274**
+> **This is an important point that we agree should be addressed in the manuscript. We updated our wording around this point. LINES 275**
 
@@
 We added this point to our discussion, in addition to referencing the papers.
-LINES 24-26, 78-87, 164-167**
+LINES 24-26, 78-87, 164-168**
 
@@
 It is true that we could evaluate our model on all of the known positive interactions, but we would still suffer from a lack of negative interaction data.
-LINES 154-159, 164-167**
+LINES 154-159, 164-168**
 
@@
 
-> **We agree that the description here is cursory, as we left the reader to see the methods section for more details.
-No external datasets were used, and we clarified this point in the text.
-ADD LINES HERE**
+> **We agree that the description here is cursory, as we left the reader to see the methods section for more details. No external datasets were used, and we clarified this point in the text.**
 
@@
 We added the suggested data as a set of historgrams outlining the distributions of node eccentricities associated with each subnetwork and the master network.
-LINES 198-199, FIGURE S8**
+LINES 199-200, FIGURE S8**
 
@@
 
-> **The word limited was confusing here. Instead we used small to communicate the caveat that this is a data mining exercise using a small dataset. LINES 231**
+> **The word limited was confusing here. Instead we used small to communicate the caveat that this is a data mining exercise using a small dataset. LINES 232**
 
@@
 
-> **We mentioned the numbers of samples used in each part of the analysis when we described them in the results section. We clarified that direction in the text. LINES 414-415**
+> **We mentioned the numbers of samples used in each part of the analysis when we described them in the results section. We clarified that direction in the text. LINES 415-416**

### 2018-03-03 bin/RandomContigGenerator.pl
Adding a random contig generator for specific percent of original length.

@@ -1,0 +1,95 @@
+#!/usr/bin/perl
+# RandomContigGenerator.pl
+# Pat Schloss Lab
+# University of Michigan
+# Geoffrey Hannigan
+
+# Set use
+use strict;
+use warnings;
+use Getopt::Long;
+use Pod::Usage;
+# Timer
+my $start_run = time();
+
+# Set variables
+my %FastaHash;
+my $opt_help;
+my $input;
+my $output;
+my $fastaInput;
+my $FastaErrorCounter = 0;
+my $FastaID;
+my $length;
+my $randLength;
+my $randStart;
+my $randContig;
+my $lpercent;
+my $percentlength;
+
+# Set the options
+GetOptions(
+	'h|help' => \$opt_help,
+	'i|input=s' => \$input,
+	'o|output=s' => \$output,
+	'p|percent=f' => \$lpercent
+);
+
+pod2usage(-verbose => 1) && exit if defined $opt_help;
+
+open(IN, "<$input") || die "Unable to read $input: $!";
+open(OUT, ">$output") || die "Unable to write to $output: $!";
+
+
+sub ReadInFasta {
+	print STDERR "Progress: Reading in fasta...\n";
+	# Set the variable for the fasta input file
+	my $fastaInput = shift;
+	# Setup fasta hash to return at the end
+	while (my $line = <$fastaInput>) {
+		if ($line =~ /\>/ && $FastaErrorCounter == 0) {
+			# print "Made it to ID!\n";
+			chomp $line;
+			$FastaID = $line;
+			# Get rid of the arrow from the ID
+			$FastaID =~ s/\>//;
+			$FastaID =~ s/_//g;
+			$FastaErrorCounter = 1;
+		} elsif ($line =~ /\>/ && $FastaErrorCounter == 1) {
+			die "KILLED BY BAD FASTA! There was no sequence before ID $line: $!";
+		} elsif ($line !~ /\>/ && $FastaErrorCounter == 0) {
+			print STDERR "Yikes, is this in block format? That is totally not allowed!\n";
+			die "KILLED BY BAD FASTA! There was a missing ID: $!";
+		} elsif ($line !~ /\>/ && $FastaErrorCounter == 1) {
+			chomp $line;
+			# Change out the lower case letters so they match the codon hash
+			$line =~ s/g/G/g;
+			$line =~ s/a/A/g;
+			$line =~ s/c/C/g;
[…]

### 2018-03-04 bin/BenchmarkingModel_SecondaryValidation.sh
Uncomment for quick running.

@@ -61,45 +61,43 @@ export -f FormatNames
 # Run CRISPR scripts #
 ######################
 
-# # Use a tmp directory
-# mkdir ./data/${Output}/tmp
+# Use a tmp directory
+mkdir ./data/${Output}/tmp
 
-# echo Extracting CRISPRs...
-# bash ./bin/RunPilerCr.sh \
-# 	${BacteriaGenomeRef} \
-# 	./data/${Output}/tmp/BenchmarkCrisprs.txt \
-# 	"/nfs/turbo/schloss-lab/bin/pilercr1.06/" \
-# 	|| exit
+echo Extracting CRISPRs...
+bash ./bin/RunPilerCr.sh \
+	${BacteriaGenomeRef} \
+	./data/${Output}/tmp/BenchmarkCrisprs.txt \
+	"/nfs/turbo/schloss-lab/bin/pilercr1.06/" \
+	|| exit
 
-# echo Getting CRISPR pairs...
-# bash ./bin/GetCrisprPhagePairs.sh \
-# 	./data/${Output}/tmp/BenchmarkCrisprs.txt \
-# 	${PhageGenomeRef} \
-# 	./data/${Output}/BenchmarkCrisprs.tsv \
-# 	"/nfs/turbo/schloss-lab/bin/ncbi-blast-2.4.0+/bin/" \
-# 	./bin/ \
-# 	./bin/ \
-# 	|| exit
-
-# rm ./data/${Output}/tmp/*
+echo Getting CRISPR pairs...
+bash ./bin/GetCrisprPhagePairs.sh \
+	./data/${Output}/tmp/BenchmarkCrisprs.txt \
+	${PhageGenomeRef} \
+	./data/${Output}/BenchmarkCrisprs.tsv \
+	"/nfs/turbo/schloss-lab/bin/ncbi-blast-2.4.0+/bin/" \
+	./bin/ \
+	./bin/ \
+	|| exit
 
 # Format the output
 FormatNames \
 	./data/${Output}/BenchmarkCrisprs.tsv \
 	${CRISPRout}
 
-# #####################
-# # Run BLAST scripts #
-# #####################
+#####################
+# Run BLAST scripts #
+#####################
 
-# echo Getting prophages by blast...
-# bash ./bin/GetProphagesByBlast.sh \
-# 	${PhageGenomeRef} \
-# 	${BacteriaGenomeRef} \
-# 	./data/${Output}/BenchmarkProphagesBlastn.tsv \
-# 	${WorkingDirectory} \
[…]

### 2018-03-04 bin/RandomContigGenerator.pl
Got rid of the underscore removal.

@@ -53,7 +53,6 @@ sub ReadInFasta {
 			$FastaID = $line;
 			# Get rid of the arrow from the ID
 			$FastaID =~ s/\>//;
-			$FastaID =~ s/_//g;
 			$FastaErrorCounter = 1;
 		} elsif ($line =~ /\>/ && $FastaErrorCounter == 1) {
 			die "KILLED BY BAD FASTA! There was no sequence before ID $line: $!";

### 2018-03-05 revfiles/ReviewerResponse.md
Updated responses.

--- 2017-08-11 revfiles/ReviewerResponse.md
@@
----
-title: Point-By-Point Reviewer Comment Response
-author: Hannigan *et al*.
----
+To the editors of PLOS Computational Biology,
 
-We would like to thank the reviewers and editor for taking the time to read our manuscript and offer constructive feedback.
-We incorporated the suggested edits and feel that this greatly improved the manuscript.
-Below we included a point-by-point response to the reviewer comments, with our comments highlighted in *italics* and the relevant manuscript change locations highlighted as **bold**.
-Line numbers were included to guide the reviewers to the relevant sections.
-Additions to the text are highlighted in **blue** and removed text is highlighted in **red**.
+We would again like to thank the reviewers and editors for their encouraging, positive, insightful, and constructive comments.
+We appreciate the time of the editors and reviewers, and this feedback greatly improved our manuscript.
+We are resubmitting our manuscript with point-by-point responses to the reviewer comments below.
+We are also submitting a marked up revised manuscript with additions marked in blue and deletions marked in red.
+Included in the revisions are four additional figures, one additional table, updated text with further discussion, and an improved quality filtering process for the sequence set.
+All of the analyses were rerun following the additional suggested quality filtering steps.
 
-# Reviewer #1 (Comments for the Author):
+The reviewer comments centered around four additional points to be addressed.
+These primary themes were 1) ensuring the purity of the samples, 2) evaluating the broader applicability of the model by testing additional true positive interactions, 3) evaluating the tolerance for in-accuracies in the model predictions, and 4) justifying the benchmarking techniques and outlining the associated limitations.
+We addressed each of these points in the manuscript text, and included additional data, figures, and dicusssion for each point.
+The details of how we addressed these points can be found below.
 
-Hannigan et al.
-describe a machine learning-based approach for predicting phage-bacterial host interactions and use it to build ecological networks for human microbiome samples based on published data.
-They analyze differences in network features as a function of host nutrition (low fat vs high fat), health status (lean vs obese twins) and body regions (diverse skin sites).
-The authors built their prediction model using data from previous publications as training set, containing 43 diverse bacterial species and 30 diverse phage strains, where the infectiveness relationship had been established experimentally (Fig 1A).
-The evaluation of the resulting predictive model shows that the amino acid similarity between phage and bacteria was the strongest predictor of infectiveness, followed by nucleotide similarity, shared protein families, and to a negligible degree CRISPRs (Fig 1B).
-They attribute the lack of predictive power of the latter to the minimal amounts of CRISPRs identified.
+# Reviewer #1
 
-The authors applied their model to three publicly available datasets (skin microbiome, twin gut microbiome of lean and obese twins and their mothers, and effect of diet on the gut microbiome) that contain sequences from full metagenomics as well as from purified viral-like particle fractions.
-Relationships between 298 phages operational genomic units (OGU) and 280 bacteria OGU were predicted from the combined taxa of all datasets and overall network features quantified (Fig1 D-G).
-Using the phage-bacteria network analysis, they explore the impact of diet and obesity in the gut microbiome structure (Fig 2), compare similarities of intrapersonal and interpersonal networks of skin and gut microbiomes (Fig 3) and test for differences between skin sites with different environmental qualities (Fig 4).
-All these comparisons are performed by either using measurements of network dissimilarity and network topology (eigen centrality, closeness centrality and degree centrality).
+The authors have adequately addressed all my concerns. I have only minor remarks left that can be easily addressed by the authors:
 
-Based on their results, the authors conclude that changes in diet, health state and body site topology can influence the phage-bacterial network stability.
-As an example they show that high-fat diet and obesity as well as external exposure of skin microbiota to reduce network connectivity.
-They also find network similarities to agree with the well-accepted notion of differences to be smaller within than between individuals.
+1) The list of true positive and true negative host-phage interactions should be shared in a supplemental table that also provides the reference for each.
 
-The approach taken by Hannigan et al.
-to construct phage-bacteria networks to analyze host-associated communities is original and will be relevant to many researchers within and also outside of the field of human microbiome research.
-One missing point the description on how the authors assessed the amount of bacterial DNA in the viral-like particle (VLP) fraction.
-If bacterial DNA were wrongly labeled as being of viral origin, it would lead to the formation of false phage OGU and result in relationships based on bacterial co-occurrences rather than phage infectivity.
-Also, the reader is not informed how abundances in the VLP fraction where integrated with the abundances of microbial fractions.
-In other words, if the DNA of the whole sample was not extracted together, how can the relative abundance of each fraction be compared to each other?
+> *This information was originally included as a set of files in the underlying code repository. We included this in the supplment as Table S2.*
 
-The presented results on the impact of diet and obesity, however, are severely underpowered, as the authors also point out (159, 187, 236).
-In particular, only one or two data points per group were used to compare the influence of diet and obesity (Figure 2).
-Although intra- vs.
-inter-individual and intra- vs.
-inter-family differences cannot be confirmed by statistical tests, they are claimed to exist.
-This is unfortunate, since the analytical approach using appropriately powered statistics reveals significant differences between skin sites, highlighting the potential power and impact of the methodology.
+2) "To avoid further distraction by this figure, we decided to remove it and focus on the suggested matrix approach, which we added to the manuscript. Figure 1"
+Figure 1 still contains the previous network figure. I personally find the heat map in Figure S8 more informative than Figure 1D, but leave the final decision to the authors.
 
-Taken together, the machine leaning-based approach would only require some additional tests to exclude or estimate the degree of bacterial contamination in the VLP fraction and some explanations on how abundances in VLP and microbial fractions were combined and normalized.
-With regard to the application of the methodology, two of the three datasets are severely underpowered to support major claims on the impact of diet and health state.
-I would thus recommend to put the focus of the manuscript on the analysis of the well-powered skin data set and reduce the diet and obesity-related part as potentially promising case-examples while clearly describing the limitations of the analysis (which is done) and toning down or removing some of the respective claims.
+> *We appologize for the confusion on our end. This was an issue we had with the resubmission process. This issue has now been corrected.*
 
-## Major Points
+I still do not understand why the authors compare graphs via eigenvectors of the adjacency matrix and not directly via the adjacency matrix itself, but I assume that both methods are correlated, so that the choice does not matter.
 
-1.
-Conceptual distinction between infectiveness and community stability: For the development of the network approach, the authors use infectiveness data as evidence for interactions between phage and bacteria.
-For the analysis part, such interactions are used to describe community stability (i.e., more interactions per node, higher stability).
-If phages are considered as agents grazing on their microbial hosts, is it justified to use degree of infectiveness analogous to community stability?
-Some discussion or background information would be helpful.
+> *This does not refer to the strict eigenvectors of the matrix but is rather a centrality metric that incorporates the both the influence of the nodes that are highly connected, and the additional influence of the nodes they are connected to.
+We included this description in our methods section.*
 
-*Thank you for raising this point which warrants further discussion.
-We clarified in the text why we feel this is justified, as the degree of stability refers to the reduced likelihood that a randomly removed node will cause a major divide or lack of connection between agents within the network.
-The infectiveness refers to the link we drew between the bacteria and phages (how we connected phages and bacteria in our networks).
-We also provided citations to work that further explains this in metabolic and phage-bacteria systems, if the readers are interested in a more in-depth discussion of the topic.* **LINES 180 - 187; 268 - 270**
+# Reviewer #2
 
-2.
-Testing for bacterial DNA in the phage fraction: The authors do not describe how bacterial sequences in the VLP fraction were eliminated.
-If bacterial sequences were present, they would be labeled as phages and lead to false predictions.
+In the new version of the manuscript, Hannigan et al.
+have made a fair effort in addressing all reviewer's concerns.
+I find an improvement from the previous version.
+Sometimes these concerns were maybe a bit "over-justified" in the text, in detriment of style, but is fine.
 
-*We agree that this needed to be better outlined in the text.
-It is important that we make clear that we are not violating our assumption that that phage OGU pool represents phages and the bacterial OGU pool represented bacteria.
-We updated the manuscript to address this concern in multiple ways.
-First, because this was a post hoc analysis of published data, we do not have access to the samples for performing further molecular QC.
-The original study authors did however perform and report a variety of QC measures to confirm the purity of their virome sample sets.
-To this end, we summarized those QC results into Table S1, which we added to the manuscript and referenced in the results.
-Furthermore, we performed our own local alignment (blast) QC protocol to assess the degree to which our OGUs were contaminated.
-We found that many phages had genomic elements similar to reference bacteria genomes, which agrees with the previous findings that the majority of the phages in these systems are temperate.
-We also identified two complete phages using the stringent Virsorter algorithm.
-Conversely, there were minimal phage reference genome elements identified in the bacterial OGUs, which also agrees with the previous findings that these sequences are mostly bacterial (cited in Table S1).
-We also did not identify any phages in the bacterial dataset using Virsorter.
-The lack of prophage identification, both here and in the previous study analyses, speaks to the need to sequence the bacterial fraction more deeply and create more robust assemblies (their larger size prevents assembly completion at the level that can be done with viruses).* **TABLE S1; LINES 92 - 121; 379 - 390**
+I understand that the main intention of this manuscript is not the computational method, but the methodological approach as a whole and to show how could this approach be used to address interesting biological questions.
+In this sense, it is a nice contribution the use network theory combined with microbial ecology to generate biological hypothesis about how the interactions of hosts and viruses (bacteria and phages in this case) change in response to different variables and how vulnerable these associations may be to these changes.
+In terms of this, I find the work valuable and of interest for the scientific community.
 
-3.
-Combining data from VLPs and microbial fractions: Information on how data from these two individual samples were matched and normalized is not provided.
-How are potentially different amounts of VLPs in different samples accounted for?
+However, I am still not quite convinced about the machine learning approach to define bacteria-phage interactions and I didn't find authors' replies on this respect convincing neither.
+My concern essentially stems from the biological base of their method, i.e., to which extent the connections between bacterial OGU and phage OGU have a biological base or are just computational artifacts.
 
-*We are unclear what the reviewer was asking.
-Our goal was to use the edge weights to reflect samples that were highly relative abundant in both the bacterial metagenome and the phage metagenome.
-The edge weight used was therefore a multiplication function of the two relative abundance values, yielding a high value for those that were highly abundant in both, and low values for those that were not.
-This is described in the methods section of the manuscript.
-These values were recorded on a per-sample basis (within the subgraphs) so that each potential combination is accounted for, within each sampling pair.
-Furthermore, it may be worth noting that the bacterial and viral DNA extractions were done separately so as to maximize purity of the samples.
-This approach allows us to more accurately measure the relative abundance in the bacterial and especially viral samples, as has been noted previously.
-We clarified this in the text and added more relevant citations.* **LINES 151 - 155; 157 - 159**
+> *We agree with these concerns and we think that the comments from the reviewer have really helped address them.*
 
-4.
-Datasets underpowered to sustain claims on diet and obesity: Although the authors point out that the data are underpowered, they still make strong claims and the analysis of these data as major points of this manuscript.
+A first caveat, acknowledged by authors and reviewers, is the fact that the training set is so small and using viruses of highly divergent environments.
+I understand authors responses to this issue, and I find them valid.
+However, one could expect that with such a poor training set, almost no links would be found in a highly complex community, with highly divergent bacteria and viruses from those use for training.
+Yet, many interactions were found.
+This is simply explained by the features that primarily defined the model.
 
-For example, 
-a.  the authors describe (158-164) how diet affects the network structure, without using sufficient data points to perform any statistical tests.
-b.  the authors claim (166-169) a difference between obese and healthy individuals by using one or two data points per group.
-With respect to these analyses, the authors write for example, "We found evidence that diet was sufficient to alter gut microbiome network connectivity" (235), which is a clear overstatement and such claims should be modified here and elsewhere in the manuscript.
+Figure 1B shows that protein identity between viral and bacterial contigs are the primary reason, followed by nucleotide identity.
+This means that the random forest model will mostly look for protein similarity between phage and bacterial contigs, i.e., will predominantly look for prophages integrated in bacterial genomes.
+Do authors think that this is really a widespread feature between bacteria and their corresponding phages, that can be effectively used in real microbial communities?
+In a way, the model is only accounting for a limited and narrow sub-fraction of the phage and bacteria spectrum (i.e., only bacteria that are already infected with prophages of a certain viral family/genus, that are ALSO present as free viral particles in the sample).
 
-*We appreciate this point referring to figure 2, and while we attempted to avoid too strong of claims, we realize we fell short.
-Our goal was to present these as interesting observations, somewhat like a case study.
-We addressed this in the manuscript by further minimizing the effects of our claims and adding relevant caveats.* **LINES 29; 185 - 206; 274 - 278**
+> *This is a valuable point and deserves further clarrification in the manuscript.
+These communities have been previously described as consisting primarily of lysogenic phages.
+Therefore, while our approach could be improved by future work and more comprehensive datasets, we are addressing the most relevant component of the virome.
+Second, while the alignment approaches were most informative, the protein family interaction approach was also very informative.
+Because the protein family approach is based on experimentally validated protein-protein interactions and not nucleotide similarity, it provides an avenue for identifying interactions involving lytic phages, that would not integrate into the bacterial genome.
+LINES 182 - 184*
 
-## Minor Points
+Second, this would only work well if viral preparations were highly pure (which is not the case), otherwise, bacterial contaminant contigs will just get connected with other bacterial contigs, simply because they share similar conserved proteins.
+The authors try to reply to this searching for 16S rRNA sequences in the viral fraction, and find that only 1.7% of OGU contained 16S sequences.
+However, they should also show what is the percentage within the bacterial OGU dataset, because since contig data is highly fractionated and incomplete and there is only one or few 16S genes per genome, one could expect that, statistically, also only few bacterial OGU will contain 16S genes.
+The important here is the ratio between OGU fractions (normalized somehow by the size of each fraction), rather than the absolute number in each of them.
+The fact that the patterns observed in figures S03 A and B are so similar between the bacteria OGU and phage OGU fractions, also creates doubts about the real nature of the viral fractions.
+Why about using one of the methods used in the literature to identigy viral contigs (they cite many of these works, e.g., Paez-espino et al.
+2016, Roux et al.
+2016) and coloring in the Figure S03 the dots that obtained a viral classification?
+Finding only five phage OGU (1.7%) with blastx hits to phages is a weak evidence.
+Overall, I think that there is still place for stronger evidences showing that the connections found are really phage-bacteria.
 
-* Throughout the manuscript, it is hard to follow how which and how many data points were used for each analysis.
-A supplementary table listing the samples for each of the studies and providing sample sizes in each figure would be very helpful.
+> *This is also a great point and we added more information around this point.
+We agree that more stringent filtering is warranted here, and the suggestion and citation for using previously described methods is fantastic.
+We decided to, as previous groups have done (Roux et al, as described above), go back and remove bacteriophage OGUs that had similarity to bacterial references but not known bacteriophages.
+This approach therefore removes OGUs that we do not have high enough confidence in being non-contaminants.
+This process involved using the tblastx algorithm because phages can be highly diverse at the nucleotide level and the phage reference genomes are limited.
+We included the results of this filtering in the manuscript, and re-ran the relevant analysis to reflect the changes in the dataset.
+This was overall a great addition to the paper because it provides us with much greater confidence in the purity of our sample set.*
 
-*We added the relevant sample size information to the text and figure legends to make clearer what the data actually represent.* **LINES 191 - 192; 200 - 202; 224 - 225; 229; FIGURE LEGENDS 2,3,4**
+> *We also thought it would be worth clarifying that we did not intend to communicate that we only found evidence of five matching phages, but rather that those were five OGUs that matched four reference genomes which had previously been validated as having incredibly broad range of infectiousness (cross-phylum infectious ranges).
+LINES 117 - 149, 496 - 497*
 
-* Figure 1D is not informative and could be removed since the information is provided in the main text.
+Finally, I also agree with referee #3 that once the model is built with the 43 bacteria and 30 phages, the model could be tested on all the remaining phage-bacteria that were not included in the model (i.e., all the positive interactions that were not used because they unbalanced the model) and see how many of them the model is able to link.
 
-*It is true that the information is in the text, but we feel this is still an important panel to the figure because it illustrates the degree of connectance between the phage and bacterial communities, and provides a visualization to the text included in the results section.
-Even more importantly, we intended this panel to act as a visual validation that there are no erroneous connections among phage and bacterial nodes.*
+> *We appreciate this suggestion and think that its implementation does add to the manuscirpt.
+As suggested by the two reviewers, we curated an additional dataset of diverse bacteria and phage interactions.
+While this approach has been used in previous studies as well, it is worth noting that there is remarkably little information for making additional predictions using interactions at the strain level because reference genomes for the bacterial host strains are not available in most cases.
+Therefore, as has been done in previous work, we looked to validate our interactions to confirm that bacteriophages are at least being predicted to infect representative bacterial genomes from the correct species.
+We applied our prediction algorith to hundreds of additional phage reference genomes and thousands of additional bacterial reference genomes and found that each phage was correctly predicted to infect at least one member of the correct bacterial species.
+This is an exciting finding because it provides an additional level of confidence in the broad applicability of our findings.
+Also, as we mentioned before, it will be exciting to evaluate this further in future work, with validated cases of phages that do not infect certain bacteria (a negative dataset).
+LINES 167 - 176, 527 - 549, Figure S5*
 
-* Figure 1 EFG: color labels in legend are wrong and not since y-axis labels are present.
+In brief, I think that, once the step of identifying true phage-bacteria connections in a sample (with the approach proposed here, or even other one) is overcame, the downstream methodology proposed in this work is a valuable and interesting approach to study the ecological roles and dynamics of phage-bacteria interactions.
 
-*Thank you for noting this mistake. We have fixed it.* **FIGURE LEGEND 1**
+> *We appreciate all of this feedback and we agree that the paper is stronger as a result of the suggested revisions.*
 
-* Figure 2 could use headers above the plots to indicate which one relate to impact of diet (AB) and which one to obesity (CD).
+# Reviewer #3
 
-*This was especially confusing because of the mislabeled axes. We fixed the axis labels to make the data clearer.* **FIGURE 2**
+The author made the point clear that they want to focus on the biological insight brought by their method more than the method itself (minor note on this point: the statement L78: "We implemented an adapted metagenomic interaction inference model that made some improvements upon previous phage-host interaction prediction models." should be removed or changed since they do not compare their model to previous models in terms of results).
+Nevertheless, to assess the confidence in the potential biological brought by the method, a careful evaluation of it is necessary.
+In my opinion, the current evaluation still suffers of major weaknesses:
 
-* Figure 2 panels C and D have wrong x-axis labels.
-Should be "obese" and "healthy".
-Here it is not clear why only 2 and 1 data points are available.
-The original data in Reyes et al.
-seem to consist of 32 fecal samples from 13 individuals from 5 families with at least two time points.
+1. Overlooking the impact of wrong predictions in the following of the analysis.
+The impact of the false predictions is never discussed and the reader have no idea if the false discovery rate of the predictor is likely to change the statistics extracted from each of the networks and how this FDR could affect or even reverse the biological conclusions of the paper.
+The reported median specificity and sensitivity of the predictor on a balanced dataset lead to an estimated FDR of 0.34.
+What is the impact of those wrong predictions on the subsequent analysis?
+Can those false predictions change the final conclusion of the paper?
+As a suggestion, one can measure the impact on the reported network statistics when randomly adding or removing edges corresponding to the fraction of estimated false negative rate and false positive rate.
+This would give an estimate of the confidence range on the final statistics of the network and assess if the differences between the samples are significant.
+Even though this analysis is mandatory for the reported median performances of the predictor, as the authors say there are high variations in the sensitivity and specificity in the nested cross-validation iterations and a complementary analysis of the impact of the wrong predictions on the final results at different sensitivity-specificity trade-offs should also be reported.
+An estimate of the change in the final results can be computed by tuning the threshold applied on the prediction score (which is responsible for the trade-off sensitivity-specificity).
 
-*We fixed the error, and made the causes for the sample size clearer in the text as well as the figure legends. Thank you for pointing this out.* **LINES 200 - 202; FIGURE LEGEND 2**
+> *This is a fantastic addition to the manuscript.
+The suggestion is to evaluate the tolerance of the findings to inaccuracies in the network prediction model.
+The focus of the findings is to compare groups which were trated the same, which means that any false positives or negatives will equally impact both groups.
+Therefore it is important to see what level of noise is tolerated by inaccuracies in the model.
+We accomplished this by implementing the analysis suggested by the reviewer.
+We iteratively added noise into the network datasets to evaluate how the observed differences in groups was effected.
+We included the results of this validation in the manuscript text and included relevant figures.
+As suggested by the reviewer, this confirms the confidence in our findings.
+LINES 295 - 301, 316 - 319, 396 - 397, 555 - 558, Figure S13, S14*
 
-* Figure 3A. Not clear why only 8 data points are available if Minot et al. report 15 samples from 5 individuals.
+2. Non realistic benchmark.
+The authors evaluate their predictor on an ideal full-genome case.
+The authors explain in their answer to reviews that they do not know how to simulate a realistic dataset.
+One can easily simulate sequencing reads followed by an assembly step.
+As for the simulation of the bins, a simulation of co-occurrence is not necessary and the effect of the binning errors can be estimated by artificially removing some contigs, and adding some contaminating contig (one can take for example the standard values given in Supplementary Figure S3 of (MetaBat, Kang et al., 2015); it would be for example good to know to what ratio of incompleteness and contamination the predictor is still reliable).
+Only then the evaluation of the method on this simulated data becomes more realistic and closer to the actual use-case made in the following of the paper.
 
-*We only used the subset of data points that had temporal data spaced approximately the same distance apart. We made this clearer in the text.* **LINES 191 - 192; 224 - 225; FIGURE LEGEND 1,2**
+> *This is another valuable comment which certainly warrants further discussion and results.
+To the reviewers suggestion around simulations, one could definitely artificially create sequence datasets from the reference genomes and assemble them back into contigs.
+The issue we would like to point out however is that, although it was suggested above that "a simulation of co-occurrence is not necessary", the referenced citation and figure do use co-occurence abundance data in their validation (in fact, MetaBat stands for "Metagenome Binning with Abundance and Tetra-nucleotide frequencies").
+The group created artificial contigs (their error free contigs) by randomly shredding reference genomes that were most highly represented in the MetaHIT dataset which they used for co-abundance calculations.
+In this way the authors could test their binning strategy using artificial contigs with co-abundance data from their MetaHIT dataset.
+As we mentioned is the last review iteration as well, we are unsure of how to accomplish this without a similar robust empirical dataset with well described interactions.
+At best we would be building a simulation on a simulation.
+To address the reviewer comments, and what we interpret to be the heart of the issue, we included a discussion as to how our method relates to this approach, what future work will be pursued, and what the ideal dataset and experimental setup will look like, in the discussion section.*
 
-* Figure 3B.
-Not clear how data were computed.
-If the comparison of within individuals (i.e.
-1 value if same data as in 3A) with mean of all other individuals, then standard deviations should be shown.
-Suggest to show only the smallest inter-individual distance, which would greatly strengthen the point on individuality.
+> *The second component that we got from this comment was that we should perform evaluation on the impact of incomplete genomic data in our prediction algorithms.
+In other words, it is valuable to look at how well the model performs on incomplete references, which more accurately reflect the true applied dataset.
+We ran this analysis using the diverse validation dataset that we also used in this reviewers point 3, and included the results in the manuscript.
+Because the issue of contamination would center around negative data (discovering false positives), which are very limited as described above, we focused on the issue of incomplete genomic information.
+In the end, we included relevant discussion in the manuscript and the resulting figure that we created.
+Together we feel that this information provides a nice validation for the work.
+Lines 419 - 432, 463 - 465, Figure S6*
 
-*Yes, this is a comparison between the interpersonal distance, and the distance between the person's samples and all other samples.
-We considered calculating the distance differences using standard deviation and pairing the samples as every interpersonal comparison instead of the average.
-The problem we found was that that pairwise comparison inflated the significance values by artificially creating a large n, all of which were not truly independent.
-We therefore collapsed the data points into a single average instead of pairing the same sample to many other samples.
-Showing the smallest distance is also an interesting idea, but we are worried that this may also skew the data by failing to account for the variation in the sample distributions.*
+3. Overlooking the evaluation on bigger datasets.
+The author explained why they evaluate the specificity of their tool only on a small dataset (to include validated negative interaction), but did not explain why they do not evaluate the sensitivity of their method on a bigger dataset using the fixed threshold used in the analysis (for evaluating the sensitivity, there is no need of negative interaction information).
+The question is: is the predictor sensitive enough to detect bacterial-host interaction on a more diverse dataset?
 
-* Figure 3C. Not clear which data were used, in particular if same data as in Figure 2 were used. E.g., were only the twins included, or also the mothers?
+> *We appreciate this feedback. This is an important point which we discuss in the related comment above by reviewer 2.*
 
-*We made this information clearer in the text.* **LINES 229 - 230; FIGURE LEGEND 3**
+- Minor remarks:
 
-* Figure 3D.
-As described for 4B, smallest inter-individual distance should be used if showing individuality should be demonstrated, which is very different from showing on average higher similarity within than between individuals.
+L195: “which similar genomic elements to” lacks a verb.
 
-*We addressed this point in the related bullet point above.*
+> *We fixed this in the text.*
 
-* Figure 4. Please provide sample numbers for each of the boxplots and data in 4E.
+Caption S6: "mtry" is not defined.
 
-*We added this information to the figure legend.* **FIGURE LEGEND 4**.
-
-# Reviewer #2 (Comments for the Author):
-
-The work presented in the manuscript entitled "Biogeography & Environmental Conditions Shape Phage & Bacteria Interaction Networks Across the Human Microbiome" by Hannigan et al.
-consists on prediction of host-bacteriophage interactions from skin and gut microbiome public datasets and graph-based analysis of the underlying phage-host networks, assessing the effects of associated metadata (obesity, diet, type of skin site).
-For the prediction of interactions, the authors trained a random forest classifier using known phage-bacteria interactions using the following features: CRISPR-spacer detection, prophage prediction based on nucleotide similarity (blastn, e-value <=1e-10.), shared gene identification based on aminoacid similarity (Diamond) and detection of pairs of genes whose proteins interact (Pfam interaction information within the Intact database).
-My main concern with this paper is on the prediction of phage-host interactions.
-The authors report 72287 infectious relationships among 578 nodes, representing 298 phages and 280 bacteria.
-If all phages would infect all bacteria, the network would have 83440 links (complete network = total possible interactions).
-The reported network contains 86,6% of the number of possible interactions.
-This essentially means that most phages are generalists and can infect any bacteria...The authors fail to report the number of phages and bacteria per microbiome: skin or gut.
-This number would allow us to estimate the total number of possible interactions within an environment (gut or skin), a more relevant number to calculate the density of the network.
-One explanation for the high density of the network is the use of aminoacid similarity between phage and bacteria as predictor.
-Shared genes have been used to link phages to their host, but at nucleotide rather than at aminoacid level (Edwards et al.
-FEMS Microbiol Rev.
-2016).
-Independently from the classifier performance, when confronted with this high network density, the shared genes should have been further analysed and the procedure revised.
-
-## Major comments:
-
-1.
-Network inference: operational genomic units were linked sample-wise, microbiome type wise or all microbiome together?
-Aminoacid similarity is not specific enough and is likely the cause of the high density of the network, which suggests that almost all phages infect all bacteria (86% density), meaning the resulting network is (likely) no statistically different from a random network with equal number of links.
-In the absence of further evidence and controls, I am not convinced with this result.
-
-*We would like to thank the reviewer for pointing out that this area needed clarification.
-The operational genomic units were linked sample-wise, by linking the OGUs found between each sample.
-The source of these samples were recorded, so that we knew whether they originated from the gut or the skin, etc.
-We also agree that amino acid similarity of genes along is not enough, and that is why we incorporated other similarity metrics in our random forest model.
-As we described, even though the amino acid similarity metric was the most "important" to the model, it was used within the context of the other metrics in the random forest.
-It is true that this is a dense network, and we hypothesize that that may be because each OGU does not necessarily represent a specific phage or bacterial strain, but is an operationally defined group of similar genomes, similar to the operation taxonomic unit (OTU) often used in microbiome studies.
-In other words, we are not using the OGUs as individual phage strains, but instead these OGUs are operational units that allow us to condense our dataset by contig similarity.* **SEE LINES 103 - 108; 311 - 329**
-
-2.
-Number of samples per study and metadata groups are not reported but the authors state about the diet effect: "Tests for statistical differences were not performed due to the small sample size ".
-If sample sizes are small for statistical analysis, no comparison should be made in the first place.
-The plots of Figure 2 feature 2 vs.
-3 comparison for high fat vs.
-low fat and 2 vs.
-1 for obese vs.
-healthy...are these the nb.
-of samples per group?
-Bacterial or phage diversity are likely confounders for network connectivity comparisons and this is overlooked here.
-
-*We appreciate this point and while we attempted to report this as more of a case trial observation (as outlined for the related comment above), we unintentionally used too strong of language, thus overstating the finding beyond our intentions.
-We also needed to clarify the analysis being performed.
-It is also true that the network analysis is a function of the diversity of the two communities, and we did not wish to make it sound like they are separate.
-We would be hesitant however to call this dependence a confounder since it is more building off of the diversity data and incorporating other additional information from the community.
-We made this clearer in the text.* **LINES 191 - 192; 200 - 202; 224 - 225; 229; FIGURE LEGENDS 2,3,4**
-
-3. Graph-analysis: Analysis of network nestedness and modularity (Weitz et al. Trends in Micr. 2012) may be more relevant than diameter, degree and closeness for phage-host networks.
-
-*We chose to omit such an analysis as was done in Weitz et al because, as we attempted to describe in the manuscript, we are not using the OGUs as individual phage strains as was done in the Weitz and other papers.
-Instead these OGUs are operational units that allow us to condense our dataset by contig similarity.
-We feel that making statements with as high of phylogenetic resolution as Weitz et al would be unfounded at this point.*
-
-## Other comments:
-
-1.
-Training set contained on 43 diverse bacterial species and 30 diverse phage strains.
-Given than these numbers are (much) lower than the ones published in references 41, 47-51, from which the data on known interactions was extracted, I wonder how the curation/filtering was done.
-
-*Our limitation was not on how many known interactions that could be found, but instead on how many non-interactions could be found.
-In other words, we needed to utilize known infectious pairs, as well as phages that have been validated as not infecting certain bacteria.
-Because we were limited to only a small number of validated non-interactions, we decided to also limit the number of positive interactions included, so as to prevent a large bias in the ratio between the two interaction types in the training dataset.
-We clarified that text.* **LINES 127 - 129**
-
-2.
-The sensitivity of the classifier is reported as 0.846.
-However, "Approximately one third of the training set relationships yielded no score and therefore were unable to be assigned an interaction prediction".
-This means the sensitivity can not be higher than 0.67...
-
-*Here we are reporting the accuracy for those values that were scored.
-Pairs without scores were omitted from this analysis, and we instead only focused on those that were scored.
-We have clarified this point.* **SEE LINES 145 - 146; FIGURE LEGEND 1**
-
-3.  Dataset: The authors should report the number of samples per study and associated metadata (e.g low /high fat diet, obese/lean).
-
-*We have added this information.* **LINES 191 - 192; 200 - 202; 224 - 225; 229; FIGURE LEGENDS 2,3,4**
-
-4.  OGU construction, why the OGUs were constructed with CONCOCT without coverage information (CONCOCT-NC)?
-
-*We maintained the coverage information but omitted the feature of total sequence counts since we worried this would confound the results.
-This means that the feature set matrix used by the CONCOCT clustering algorithm only included k-mer frequencies and sequence count (relative abundance) values per sample.*
-
-5.
-Network construction: The master network contained the three studies as sub-networks, which themselves each contained sub-networks for each sample.
-The authors provide only numbers of phage and bacterial nodes for the master network.
-For predicting the associations, were all phages and bacteria considered disregarding the sample type (skin vs.
-gut)?
-
-*Yes, this is described correctly by the reviewer.*
-
-6.
-The phages and bacteria in the gut diet and twin sample sets were more sparsely related: each contained fewer than 150 vertices, fewer than 20,000 relationships".
-This is a bipartite network, what should be reported is the number of vertices of each type rather than the total nb of vertices (this is relevant to calculate the network density, we cannot know here what is the nb of possible interactions).
-
-*We added these values to the results section of the manuscript so that the reader can access this information.* **LINES 155 - 165**
-
-7.  Comparison with Paez-Espino et al. (Nature 2016, doi:10.1038/nature19094) phage-host human associated subnetworks should be done.
-
-*This is an intriguing suggestion.
-It sounds like this comment is a reference to the host range inferences within the suggested manuscript.
-Paez-Espino et al used CRISPR and tRNA comparisons to identify bacterial hosts for the bacteriophage genomes and genome fragments.
-The group's work suggested that phage host range is broader than has been traditionally appreciated.
-In our discussion section, we mentioned a comparison of these and related findings to our own dataset.
-A reference to this information was included in the operational genomic unit section of the discussion.* **LINES 318**
-
-8.  Figure 1. panel B: methods include blastx, in the network inference Diamond was used instead; Panel E-F, colors in figure do not correspond to those in legend.
-
-*It is true that we used the Diamond aligner, but we used the blastx algorithm within the Diamond aligner (written as `diamond blastx`).
-We made this information clearer in the text.
-We also fixed the color references.* **FIGURE 1**
-
-9.
-Figure 2 legend: it is not clear if/that the dots represent sample-wise networks.
-"Lines represent the mean degree of centrality for each diet", the only lines in the plots are parallel to the centrality axis, they are just splitting high fat vs.
-low fat diet fields in the x axis.
-X-axis from panels C and D are mislabeled.
-
-*We clarified the dot information, and also fixed the incorrect labels.* **FIGURE 2**
-
-# Reviewer #3 (Comments for the Author):
-
-This research article by Hannigan et.al., focuses on an important and indispensable part of the microbiome- "Virome".
-Interplay within inter and intra microbial community members is important to get the complete picture.
-This paper focuses on such interplay and understands that microbiome operates with both the viral and bacterial community members.
-For this study they worked on previously published metagenomics data, to not only infer based on nucleotide information but also based on protein interactions.
-
-## Main Concerns:
-
-The data from studies referenced 13 & 14 even though may have bias due to sequencing, have similarity in approach wherein collection of VLP and bacterial whole genome shotgun sequencing was performed.
-VLP data in reference 14 was based on Genomiphi amplification of VLP DNA prior to sequencing that is bound to introduce amplification and sequencing errors which would further lead to clustering bias in the current study.
-The concern is also regarding data from references 38 and 39 that are not only performed by 454 technology.
-Further on, reference 39 doesn't have VLP data available.
-The study is based on 16S and whole genome shotgun of gut microbiome.
-
-*We referenced this statement in reference 38: "We had previously performed shotgun sequencing of total fecal DNA isolated from the 12 frozen samples obtained at the first time point that were now used to prepare purified VLPs (Table S3)." This reference (12) led to our reference 39, which contained the data in the related SRA submission and was cited accordingly.
-Between these two published datasets, we were able to pair the whole shotgun sequencing data with the purified virome data.*
-
-The question is how were the above biases (including the kitome) taken into consideration while conducting these analyses.
-The authors may be better off carrying out the analysis only using data from references 13 and 14 for skin and only data from 38 for gut?
-
-*There certainly could have been extraction, kit, or other related differences between the 38 and 39 dataset.
-We feel these findings, which we attempt to present more as case observations that will require more dedicated study, are worth noting in the paper.
-We altered the text to better clarify our position.
-We also avoided analyses in which we directly compared datasets, because we are unable to discern between true biological differences and differences between the studies.
-Such analyses would require further and properly controlled sampling.*
-
-Even though the authors have acknowledged some of these biases including MDA, in their discussion, these concerns should be explicitly discussed and measures they have taken to overcome these.
-
-*We included a description of these biases associated with these methods.
-As far as measures to overcome these biases, we acknowledge the biases and only performed analyses within studies where samples were treated the same.
-The goal is for this work to use these findings as a jumping off point, and performing future studies with more limited biases from techniques such as MDA.* **SEE LINES 302 - 310**
+> *We added the definition in the caption.*

### 2018-03-05 doc/manuscript.Rmd
Updated docs.

@@
 
-The bacterial and viral sequences from these studies were quality filtered and assembled into contigs.
+The bacterial and viral sequences from these studies were quality filtered and assembled into contigs ("genomic fragments").
 We further grouped the related bacterial and phage contigs into operationally defined units based on their k-mer frequencies and co-abundance patterns, similar to previous reports **(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})** [@Roux:2016cc].
@@
 
-While supplementing the previous virome fraction quality control measures **(Table S1)** we found that, in light of the rigorous purification and quality control during sample collection and preparation, 77% (228 / 298 operational genomic units) still had some nucleotide similarity to a given bacterial reference genome (e-value < $10^{-25}$).
-As most phages in these communities have been shown to be temperate (they integrate into bacterial genomes) using methods that included nucleotide alignments of phages to bacterial reference genomes, we interpreted this as confirmation that the majority of those phages were temperate and therefore shared elements with bacterial reference genomes -- a trend previously reported [@Minot:2011ez, @Hannigan:2015fz, @Reyes:2010cwa].
-We further confirmed that while the majority of these were expected to be temperate phages, there still remained a low level of bacterial sequence noise, which was evident as five (1.7%) OGUs with similar sequence elements to the bacterial 16S rRNA gene (blastn, e-value < $10^{-25}$, length > 1,000 bp).
-This is in line with previous work which has suggested that bacterial noise is an inevitable technical issue, and we considered this noise while interpreting our findings [@Roux:2013cr].
-We also identified two OGUs as being complete, high confidence phages using the stringent Virsorter phage identification algorithm (class 1 confidence group) [@Roux:2015dh].
+The original publications reported that the whole metagenomic shotgun sequence samples, which primarily consisted of bacteria, had an average viral relative abundance of 0.4% **(Table S1)** [@Minot:2011ez; @Hannigan:2015fz; @Reyes:2010cwa; @Turnbaugh:2009ei].
+We confirmed these reports by finding that only 2% (6 / 280 OGUs) of bacterial OGUs had significantly strong nucleotide similarity to phage reference genomes (e-value < $10^{-25}$) [@Minot:2011ez; @Hannigan:2015fz; @Reyes:2010cwa; @Turnbaugh:2009ei].
+Additionally, no OGUs were confidently identified as lytic or temperate phage OGUs in the bacterial dataset using the Virsorter algorithm [@Roux:2015dh].
 
-The whole metagenomic shotgun sequence samples, which primarily consisted of bacteria, had an average viral relative abundance of 0.4% **(Table S1)** [@Minot:2011ez; @Hannigan:2015fz; @Reyes:2010cwa; @Turnbaugh:2009ei].
-We found that only 2% (6 / 280 OGUs) of bacterial OGUs had significantly strong nucleotide similarity to phage reference genomes (e-value < $10^{-25}$) [@Minot:2011ez; @Hannigan:2015fz; @Reyes:2010cwa; @Turnbaugh:2009ei].
-No OGUs were confidently identified as lytic or temperate phage OGUs in the bacterial dataset using the Virsorter algorithm [@Roux:2015dh].
-Together this suggests minimal bacterial OGU noise that should be considered in the study conclusions.
+We also supplemented the previous virome fraction quality control measures **(Table S1)** to find that, in light of the rigorous purification and quality control during sample collection and preparation, 77% (228 / 298 operational genomic units) still had some nucleotide similarity to a given bacterial reference genome (e-value < $10^{-25}$).
+It is important to note that interpreting such alignment is complicated by the fact that most reference bacterial genomes also contain prophages (phages integrated into bacterial genomes), meaning we do not know to what extent the alignments were the result of bacterial contaminants in the virome fraction and what were true integrated prophages.
+As most phages in these communities have been shown to be temperate (they integrate into bacterial genomes) using methods that included nucleotide alignments of phages to bacterial reference genomes, we expected that a large fraction of those phages were temperate and therefore shared elements with bacterial reference genomes -- a trend previously reported [@Minot:2011ez, @Hannigan:2015fz, @Reyes:2010cwa].
+To ensure the purity of our sample sets, we supplemented the quality control measures by filtering out all OGUs that could be potential bacterial contaminants, as described previously [@Roux:2016cc].
+This resulted in the removal of 143 OGUs that exhibited nucleotide similarity to bacterial genomes but no identifiable known phage elements.
+We were also able to identify two OGUs as representing **complete**, high confidence phages using the stringent Virsorter phage identification algorithm (class 1 confidence group) [@Roux:2015dh].
 
@@
 Only bacteria and bacteriophages were used in the model.
-The training set contained 43 diverse bacterial species and 30 diverse phage strains, including both broad and specific ranges of infection **(Supplemental Figure \ref{ValidationOverview} A - B)**.
+The training set contained 43 diverse bacterial species and 30 diverse phage strains, including both broad and specific ranges of infection **(Supplemental Figure \ref{ValidationOverview} A - B, Table S2)**.
 While it is true that there are more known phages that infect bacteria, we were limited by the information confirming which phages do not infect certain bacteria and attempted to keep the numbers of positive and negative interactions similar.
@@
 Altogether we decided to maintain a balanced dataset at the cost of under-sampling the available positive interaction information because the use of such a severely unbalanced dataset often results in over-fit and uninformative model training.
+However, as an additional validation measure, we used the extensive additional positive interactions as a secondary dataset to confirm that we could identify infectious interactions from a more diverse bacterial and phage dataset.
+Using this approach, we confirmed that 382 additional phage reference genomes, representing a diverse range of phages, were matched to at least one reference bacterial host genome of the species that they were expected to infect **(Supplemental Figure \ref{secondval})**.
+Because the model was built on full genomes and used on OGUs, we also assessed whether our model was resilient to incomplete reference genomes.
+We found that the use of our model on random contigs representing as little as 50% length of the original reference phage and bacterial genomes resulted in minimal reduction in the ability of the model to identify relationships **(Supplemental Figure \ref{contigval})**.
 
@@
 These features were calculated using the training set described above.
+While the nucleotide and amino acid similarity metrics are expected to identify prophage signatures, the protein family interaction and CRISPR signatures are expected to aid in identifying lytic phages in addition to temperate phages.
 We chose to utilize these metrics that directly compare nucleotide sequences between sample phages and bacteria, instead of relying on alignment to reference genomes or known marker genes, because we are extrapolating our model to highly diverse communities which we expect to diverge significantly from the available reference genomes.
@@
 The phages and bacteria in the diet and twin sample sets were more sparsely related, with the diet study consisting of `r prettyNum(genfigurestats[c(genfigurestats$class %in% "DietStudy"),"Vertices"], big.mark = ",")` (`r prettyNum(sitestats[c(sitestats$class %in% "diet"),"phagenodes"], big.mark = ",")` phages, `r prettyNum(sitestats[c(sitestats$class %in% "diet"),"bacterianodes"], big.mark = ",")` bacteria) nodes and `r prettyNum(genfigurestats[c(genfigurestats$class %in% "DietStudy"),"Edges"], big.mark = ",")` relationships, and the twin study containing `r prettyNum(genfigurestats[c(genfigurestats$class %in% "TwinStudy"),"Vertices"], big.mark = ",")` (`r prettyNum(sitestats[c(sitestats$class %in% "twin"),"phagenodes"], big.mark = ",")` phages, `r prettyNum(sitestats[c(sitestats$class %in% "twin"),"bacterianodes"], big.mark = ",")` bacteria) nodes and `r prettyNum(genfigurestats[c(genfigurestats$class %in% "TwinStudy"),"Edges"], big.mark = ",")` relationships **(Figure \ref{RocCurve} E - F)**.
-As an interesting validation measure, we identified five (1.7%) examples of phage OGUs which similar genomic elements to the four previously described, broadly infectious phages isolated from Lake Michigan (tblastx; e-value < $10^{-25}$) [@Malki:2015dw].
+As an interesting validation measure, we identified five (1.7%) examples of phage OGUs which contained similar genomic elements to the four previously described, broadly infectious phages isolated from Lake Michigan (tblastx; e-value < $10^{-25}$) [@Malki:2015dw].
 
@@
 
+As an additional validation measure, we evaluated the tolerance of these findings to inaccuracies in the underlying network.
+As described above, our model is not perfect and there is likely to be noise from false positive and false negative predictions.
+We found that additional random noise, both by creating a fully connected graph or randomly reducing the number of edges to 60% of the original, changed the statistical significance values (p-values) of our findings but not by enough to change whether they were statistically significant (p-value < 0.05).
+Therefore the comparisons between groups are resilient to potential noise resulting from model false positive and false negative predictions **(Supplemental Figure \ref{idivnoise})**.
+
 ![**Intrapersonal vs Interpersonal Network Dissimilarity Across Different Human Systems.** *(A) NMDS ordination illustrating network dissimilarity between subjects over time.
@@
 The average centrality of each sample was quantified using the weighted eigenvector centrality metric.
-Intermittently moist skin sites (dynamic sites that fluctuate between being moist and dry) were significantly less connected than the moist and sebaceous environments (p-value < 0.001, **Figure \ref{skinnetwork} A)**.
-Also, skin sites that were occluded from the environment were much more highly connected than those that were constantly exposed to the environment or only intermittently occluded (p-value < 0.001, **Figure \ref{skinnetwork} B)**.
+Intermittently moist skin sites (dynamic sites that fluctuate between being moist and dry) were significantly more connected than the moist and sebaceous environments (p-value < 0.001, **Figure \ref{skinnetwork} A)**.
+Also, skin sites that were occluded from the environment were less connected than those that were constantly exposed to the environment or only intermittently occluded (p-value < 0.001, **Figure \ref{skinnetwork} B)**.
+We also confirmed that addition of noise to the underlying network, as described above, altered the values of statistical significance (p-values) but not by enough to change whether they were statistically significant **(Supplemental Figure \ref{skindivnoise})**.
 
@@
 Network structure differed between environmentally exposed and occluded skin sites.
-The sites under greater environmental fluctuation and exposure (the exposed and intermittently exposed sites) were less connected and therefore were predicted to have a higher resilience against network degradation when random nodes are removed from the network.
-Likewise, intermittently moist sites demonstrated less connectedness than the moist and sebaceous sites.
+The sites under greater environmental fluctuation and exposure (the exposed and intermittently exposed sites) were more connected and therefore were predicted to have a higher resilience against network degradation when random nodes are removed from the network.
+Likewise, intermittently moist sites demonstrated higher connectedness than the moist and sebaceous sites.
 These findings agree with previous work that has shown that bacterial community networks differ by skin environment types [@Bashan:2016is].
-Together these data suggested that body sites under greater degrees of fluctuation harbored less connected microbiomes that are potentially less robust to network disruption by extinction events.
+Together these data suggested that body sites under greater degrees of fluctuation harbored more highly connected microbiomes that are potentially more robust to network disruption by extinction events.
 This points to a link between microbiome and environmental robustness toward network homeostasis and warrants further investigation.
 
-While these findings take us an important step closer to understanding the microbiome through interspecies relationships, there are caveats to and considerations regarding our findings.
+While these findings take us an important step closer to understanding the microbiome through interspecies relationships, there are caveats and considerations to our findings.
 First, as with most classification models, the infection classification model developed and applied is only as good as its training set -- in this case, the collection of experimentally-verified positive and negative infection data.
@@
 This caveat is not new to the microbiome field, with a notable example being the use of 16S rRNA sequencing using the V4 variable region [@Meisel:2016iu].
-Use of the V4 variable region excluded detection of major bacterial members, meaning that the findings were not able to completely describe the underlying biological environment.
+Use of the V4 variable region excluded detection of major skin bacterial members, meaning that the findings were not able to completely describe the underlying biological environment.
 Despite this caveat, skin microbiome studies provided valuable biological insights by focusing on the community differences between groups (e.g.
@@
 Similarly, here we focused our conclusions on the differences between the groups which were all treated the same, so that we can minimize our dependence on a perfect predictive model.
+We also provided explicit evidence that the introduction of noise equally to the compared groups did not significantly impact our findings.
 
-Finally, the networks in this study were built using operational genomic units (OGUs), which represented groups of highly similar bacteria or phage genomes or clustered genome fragments.
+Third, the networks in this study were built using operational genomic units (OGUs), which represented groups of highly similar bacteria or phage genomes or clustered genome fragments.
 Similar clustering definition and validation methods, both computational and experimental, have been implemented in other metagenomic sequencing studies, as well [@Minot:2012ed; @Deng:2014eb; @Brum:2015iaa; @Roux:2016cc].
@@
 Additionally, visualization of our dataset interactions using the heat map approach previously used in other host range studies, suggests a trend toward modular and nested tropism, but we do not have the strain-level resolution that powered those previous experimental studies.
+
+Finally, it is important to note that our model was built using available full genomes with known interactions, while the experimental datasets resulted in OGUs created from metagenomic shotgun sequence sets, as described above.
+While this is an informative approach given available data, it is not ideal.
+We envision future work focusing on training models using metagenomic shotgun sample sets from "mock communities" of bacteria and phages with experimentally validated infectious relationships.
+This would also be more informative than relying on simulated metagenomic sample sets, whose use would result in models built on simulations and more assumptions instead of empirical data.
+Together this way the training set can be subjected to the same pre-processing, contig assembly, and OGU binning processes as the experimental data.
+Furthermore, exciting advances in long read sequencing platforms such as the Oxford Nanopore MinIon system will provide more accurate genomic scaffolds than *de novo* assembled contigs, allowing for more accurate training and predictions of our models.
+As discussed above, it is because our current model is susceptible to this noise that we focus our conclusions on comparisons between experimental groups that were both treated the same.
+This is also why it was important for us to evaluate the susceptibility of our results to noise caused by the less-than-perfect prediction model.
 
@@
 
+Contig simulations were performed by randomly extracting a string of genomic nucleotides that represented a defined percent length of that genome.
+This was accomplished using our `RandomContigGenerator.pl`, which was published in the associated GitHub repository.
+
 ## Contig Abundance Calculations
@@
 We also performed the converse analysis of aligning phage OGU representative sequences to EMBL bacterial reference genomes.
-Finally, we ran both the phage and bacteria OGU representative sequences through the Virsorter program (1.0.3) to identify phages (all default parameters were used), using only those in the high confidence identification category "class 1" [@Roux:2015dh].
+We ran both the phage and bacteria OGU representative sequences through the Virsorter program (1.0.3) to identify phages (all default parameters were used), using only those in the high confidence identification category "class 1" [@Roux:2015dh].
+Finally, we filtered out phage OGUs that had bacterial elements as described above, but also lacked known phage elements by using the tblastx algorithm and a maximum e-value of $10^{-25}$.
 
@@
 
+## Secondary Dataset Validation
+The performance of our model for identifying diverse infectious relationships between bacteria and phages, beyond those that were included in the model creation step, were validated using additional bacterial and phage reference genomes, which could be linked by the records of which phage strains were isolated on which bacteria under laboratory conditions.
+Viral and bacterial reference genomes were downloaded from the GenBank repository on February 19, 2018 using the viral location `ftp://ftp.ncbi.nih.gov/refseq/release/viral/` and the bacterial location `ftp://ftp.ncbi.nih.gov/refseq/release/bacteria/`.
+This resulted in the use of 539 complete phages reference genomes (with identified hosts) and 3,469 bacterial reference genomes.
+We used the same prediction model to predict which phages were infecting which hosts, so as to confirm that the model was capable of identifying interactions in a more diverse dataset.
+Bacteria interactions were identified at the species level.
+The random contig iteration analysis was performed using a subset of bacterial reference genomes, for computational performance reasons.
+Only single representative genomes for each species were used.
+
 ## Interaction Network Construction
@@
 This network was used for downstream community analysis.
+Tolerance to false negative and false positive noise within the networks was assessed by randomly removing a defined fraction of network edges before re-running the downstream analysis work flows.
+This was accomplished using functionality within the igraph R package (v1.0.1) [@Theigraphsoftware:vh].
 
@@
 
+![**Ability of prediction model to identify broad range of bacteria and phage interactions.** *Each complete bacteriophage labeled on the y axis.
+The number of complete bacterial species genomes that were correctly predicted to be infected by each phage, according to the GenBank records, is on the x-axis.
+True positive interactions are colored in purple, and false negative interactions are yellow.
+Because this dataset only had confirmed interactions and not confirmed lack of interactions, false positive and true negative values could be not determined.*\label{secondval}](){ width=100% }
+
+\newpage
+
+![**Impact of incomplete genomic sequences on model performance.** *Number of correctly identified infectious interactions between bacteria and phages are presented on the y-axis.
+The fraction of genomic length that was used to randomly extract contigs from reference bacterial and phage sequences is presented on the x-axis (e.g.
+0.9 means contig lengths were 90% of the total genome length).
+Overall this presents a quantification of the loss of identified infectious relationships as the percent of available genomic material is reduced.*\label{contigval}](){ width=100% }
+
+\newpage
+
 ![**Classification Model Performance By Nested Cross-Validation.** *Box plot illustrating the median and variance of phage-bacteria interaction prediction model.
@@
 
-![**Stable Classification Model Performance Over Random Iterations** *In addition to nested cross-validation, here we show the results from the five-fold cross validation, in which 20% of the samples were randomly withheld during the training stage for model evaluation and mtry tuning.
+![**Stable Classification Model Performance Over Random Iterations** *In addition to nested cross-validation, here we show the results from the five-fold cross validation, in which 20% of the samples were randomly withheld during the training stage for model evaluation and mtry tuning (parameter defined in the R Random Forest package, which is implemented in Caret, as "the number of variables randomly sampled as candidates at each split").
 The results of 25 random iterations are shown.
@@
 
+![**P-values of interpersonal group diversity differences with graph edge noise.** *The x-axis represents the percent of errors that were randomly removed (or added) as a means to evaluate the impact of random noise on statistical significance of group differences.
+Resulting p-values for each graph is shown on the y-axis.
+The dot and bars are the mean and standard error of five iterations of group comparisons with random edge removal.
+Significance of A) ANOSIM p-value of diet network dissimilarity, B) p-value of interpersonal and intrapersonal diet network dissimilarity, C) p-value of interpersonal and intrapersonal skin network dissimilarity, and D) p-value of interpersonal and intrapersonal twin gut network dissimilarity.
+This corresponds to the findings in Figure 3.*\label{idivnoise}](){ width=75% }
+
+\newpage
+
+![**P-values of differences in Eigen Centrality between skin site microbiome networks.** *The x-axis represents the percent of errors that were randomly removed (or added) as a means to evaluate the impact of random noise on statistical significance of group differences.
+Resulting p-values for each graph is shown on the y-axis.
+The dot and bars are the mean and standard error of five iterations of group comparisons with random edge removal.
+The groups compared were the degrees of site moisture (left) and occlusion (right).
+The findings correspond to pannels A and B in Figure 4.*\label{skindivnoise}](){ width=75% }
+
+\newpage
+
 # Supplemental Table Captions
@@
 
+Table S2: The positive and negative bacteria and bacteriophage interactions used to train the prediction model, as also illustrated in Figure \ref{ValidationOverview}.
+Citation sources are also included.
+
 \newpage

### 2018-03-05 doc/manuscript.tex
Updated docs.

@@
 The bacterial and viral sequences from these studies were quality
-filtered and assembled into contigs. We further grouped the related
-bacterial and phage contigs into operationally defined units based on
-their k-mer frequencies and co-abundance patterns, similar to previous
-reports \textbf{(Supplemental Figure \ref{ContigStats} -
-\ref{ClusterStats})} {[}42{]}. This was done both for dimensionality
-reduction and to prevent inflation of node counts by using contigs which
-are expected to represent multiple fragments from the same genomes. This
-was also done to create genome analogs that we could use in our
-classification model which was built using genome sequences. We referred
-to these operationally defined groups of related contigs as operational
-genomic units (OGUs). Each OGU represented a genomically similar
-sub-population of either bacteria or phages. Contig lengths within
-clusters ranged between \(10^{3}\) and \(10^{5.5}\) bp
+filtered and assembled into contigs (``genomic fragments''). We further
+grouped the related bacterial and phage contigs into operationally
+defined units based on their k-mer frequencies and co-abundance
+patterns, similar to previous reports \textbf{(Supplemental Figure
+\ref{ContigStats} - \ref{ClusterStats})} {[}42{]}. This was done both
+for dimensionality reduction and to prevent inflation of node counts by
+using contigs which are expected to represent multiple fragments from
+the same genomes. This was also done to create genome analogs that we
+could use in our classification model which was built using genome
+sequences. We referred to these operationally defined groups of related
+contigs as operational genomic units (OGUs). Each OGU represented a
+genomically similar sub-population of either bacteria or phages. Contig
+lengths within clusters ranged between \(10^{3}\) and \(10^{5.5}\) bp
 \textbf{(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})}.
 
-While supplementing the previous virome fraction quality control
-measures \textbf{(Table S1)} we found that, in light of the rigorous
+The original publications reported that the whole metagenomic shotgun
+sequence samples, which primarily consisted of bacteria, had an average
+viral relative abundance of 0.4\% \textbf{(Table S1)} {[}13,14,38,39{]}.
+We confirmed these reports by finding that only 2\% (6 / 280 OGUs) of
+bacterial OGUs had significantly strong nucleotide similarity to phage
+reference genomes (e-value \textless{} \(10^{-25}\)) {[}13,14,38,39{]}.
+Additionally, no OGUs were confidently identified as lytic or temperate
+phage OGUs in the bacterial dataset using the Virsorter algorithm
+{[}50{]}.
+
+We also supplemented the previous virome fraction quality control
+measures \textbf{(Table S1)} to find that, in light of the rigorous
 purification and quality control during sample collection and
@@
 nucleotide similarity to a given bacterial reference genome (e-value
-\textless{} \(10^{-25}\)). As most phages in these communities have been
+\textless{} \(10^{-25}\)). It is important to note that interpreting
+such alignment is complicated by the fact that most reference bacterial
+genomes also contain prophages (phages integrated into bacterial
+genomes), meaning we do not know to what extent the alignments were the
+result of bacterial contaminants in the virome fraction and what were
+true integrated prophages. As most phages in these communities have been
 shown to be temperate (they integrate into bacterial genomes) using
 methods that included nucleotide alignments of phages to bacterial
-reference genomes, we interpreted this as confirmation that the majority
-of those phages were temperate and therefore shared elements with
-bacterial reference genomes -- a trend previously reported {[}14{]}. We
-further confirmed that while the majority of these were expected to be
-temperate phages, there still remained a low level of bacterial sequence
-noise, which was evident as five (1.7\%) OGUs with similar sequence
-elements to the bacterial 16S rRNA gene (blastn, e-value \textless{}
-\(10^{-25}\), length \textgreater{} 1,000 bp). This is in line with
-previous work which has suggested that bacterial noise is an inevitable
-technical issue, and we considered this noise while interpreting our
-findings {[}50{]}. We also identified two OGUs as being complete, high
-confidence phages using the stringent Virsorter phage identification
-algorithm (class 1 confidence group) {[}51{]}.
-
-The whole metagenomic shotgun sequence samples, which primarily
-consisted of bacteria, had an average viral relative abundance of 0.4\%
-\textbf{(Table S1)} {[}13,14,38,39{]}. We found that only 2\% (6 / 280
-OGUs) of bacterial OGUs had significantly strong nucleotide similarity
-to phage reference genomes (e-value \textless{} \(10^{-25}\))
-{[}13,14,38,39{]}. No OGUs were confidently identified as lytic or
-temperate phage OGUs in the bacterial dataset using the Virsorter
-algorithm {[}51{]}. Together this suggests minimal bacterial OGU noise
-that should be considered in the study conclusions.
+reference genomes, we expected that a large fraction of those phages
+were temperate and therefore shared elements with bacterial reference
+genomes -- a trend previously reported {[}14{]}. To ensure the purity of
+our sample sets, we supplemented the quality control measures by
+filtering out all OGUs that could be potential bacterial contaminants,
+as described previously {[}42{]}. This resulted in the removal of 143
+OGUs that exhibited nucleotide similarity to bacterial genomes but no
+identifiable known phage elements. We were also able to identify two
+OGUs as representing \textbf{complete}, high confidence phages using the
+stringent Virsorter phage identification algorithm (class 1 confidence
+group) {[}50{]}.
 
@@
 random forest model trained on experimentally validated infectious
-relationships from six previous publications {[}41,52--56{]}. Only
+relationships from six previous publications {[}41,51--55{]}. Only
 bacteria and bacteriophages were used in the model. The training set
@@
 including both broad and specific ranges of infection
-\textbf{(Supplemental Figure \ref{ValidationOverview} A - B)}. While it
-is true that there are more known phages that infect bacteria, we were
-limited by the information confirming which phages do not infect certain
-bacteria and attempted to keep the numbers of positive and negative
-interactions similar. Phages with linear and circular genomes, as well
-as ssDNA and dsDNA genomes, were included in the analysis. Because we
-used DNA sequencing studies, RNA phages were not considered
+\textbf{(Supplemental Figure \ref{ValidationOverview} A - B, Table S2)}.
+While it is true that there are more known phages that infect bacteria,
+we were limited by the information confirming which phages do not infect
+certain bacteria and attempted to keep the numbers of positive and
+negative interactions similar. Phages with linear and circular genomes,
+as well as ssDNA and dsDNA genomes, were included in the analysis.
+Because we used DNA sequencing studies, RNA phages were not considered
 \textbf{(Supplemental Figure \ref{ValidationOverview} C-D)}. This
@@
 unbalanced dataset often results in over-fit and uninformative model
-training.
+training. However, as an additional validation measure, we used the
+extensive additional positive interactions as a secondary dataset to
+confirm that we could identify infectious interactions from a more
+diverse bacterial and phage dataset. Using this approach, we confirmed
+that 382 additional phage reference genomes, representing a diverse
+range of phages, were matched to at least one reference bacterial host
+genome of the species that they were expected to infect
+\textbf{(Supplemental Figure \ref{secondval})}. Because the model was
+built on full genomes and used on OGUs, we also assessed whether our
+model was resilient to incomplete reference genomes. We found that the
+use of our model on random contigs representing as little as 50\% length
+of the original reference phage and bacterial genomes resulted in
+minimal reduction in the ability of the model to identify relationships
+\textbf{(Supplemental Figure \ref{contigval})}.
 
@@
 similarity of protein families associated with experimentally identified
-protein-protein interactions {[}57{]}. These features were calculated
-using the training set described above. We chose to utilize these
-metrics that directly compare nucleotide sequences between sample phages
-and bacteria, instead of relying on alignment to reference genomes or
-known marker genes, because we are extrapolating our model to highly
-diverse communities which we expect to diverge significantly from the
-available reference genomes. The resulting random forest model was
+protein-protein interactions {[}56{]}. These features were calculated
+using the training set described above. While the nucleotide and amino
+acid similarity metrics are expected to identify prophage signatures,
+the protein family interaction and CRISPR signatures are expected to aid
+in identifying lytic phages in addition to temperate phages. We chose to
+utilize these metrics that directly compare nucleotide sequences between
+sample phages and bacteria, instead of relying on alignment to reference
+genomes or known marker genes, because we are extrapolating our model to
+highly diverse communities which we expect to diverge significantly from
+the available reference genomes. The resulting random forest model was
 assessed using nested cross validation, and the median area under its
@@
 build the interactive network. The master network, analogous to the
-universal microbiome network concept previously described {[}58{]},
+universal microbiome network concept previously described {[}57{]},
 contained the three studies as sub-networks, which themselves each
@@
 bacterial libraries ensured a more accurate measurement of the microbial
-communities, as has been outlined previously {[}59,60{]}. The master
-network was highly connected and contained 72,287 infectious
-relationships among 578 nodes, representing 298 phages and 280 bacteria.
+communities, as has been outlined previously {[}58,59{]}. The master
+network was highly connected and contained 38,337 infectious
+relationships among 435 nodes, representing 155 phages and 280 bacteria.
 Although the network was highly connected, not all relationships were
@@
 size; the greatest number of traversed vertices required between two
-vertices) and included 576 (297 phages, 279 bacteria, 99.7\% total) and
-72,127 (99.8\%) of the master network nodes and edges, respectively
+vertices) and included 433 ( phages, bacteria, 99.5\% total) and 38,099
+(99.4\%) of the master network nodes and edges, respectively
 \textbf{(Figure \ref{RocCurve} E - F)}. Additionally, the subnetworks
@@
 the diet and twin sample sets were more sparsely related, with the diet
-study consisting of 89 (41 phages, 48 bacteria) nodes and 5,566
-relationships, and the twin study containing 137 (36 phages, 101
-bacteria) nodes and 17,250 relationships \textbf{(Figure \ref{RocCurve}
-E - F)}. As an interesting validation measure, we identified five
-(1.7\%) examples of phage OGUs which similar genomic elements to the
+study consisting of 80 ( phages, bacteria) nodes and 4,692
+relationships, and the twin study containing 130 ( phages, bacteria)
+nodes and 14,684 relationships \textbf{(Figure \ref{RocCurve} E - F)}.
+As an interesting validation measure, we identified five (1.7\%)
+examples of phage OGUs which contained similar genomic elements to the
 four previously described, broadly infectious phages isolated from Lake
-Michigan (tblastx; e-value \textless{} \(10^{-25}\)) {[}61{]}.
+Michigan (tblastx; e-value \textless{} \(10^{-25}\)) {[}60{]}.
 
@@
 availability and gut microbiome composition and diversity, including
-bacteria and phages {[}14,62,63{]}. Previous work in isolated
+bacteria and phages {[}14,61,62{]}. Previous work in isolated
 culture-based systems has suggested that changes in nutrient
@@
 overall greater degree of connections, which suggests a greater
-resilience against network degradation by extinction events {[}30,64{]}.
+resilience against network degradation by extinction events {[}30,63{]}.
 This is because more highly connected networks are less likely to
 degrade into multiple smaller networks when bacteria or phages are
-randomly removed {[}30,64{]}. We used this information to calculate the
+randomly removed {[}30,63{]}. We used this information to calculate the
 average connectedness per sample, which was corrected for the maximum
@@
 with people remaining more similar to themselves than to other people
-over time {[}13,65,66{]}. We therefore hypothesized that this personal
+over time {[}13,64,65{]}. We therefore hypothesized that this personal
 conservation extended to microbiome network structure. We addressed this
@@
 over time. We found that gut microbiome network structures clustered by
-person (ANOSIM p-value = 0.005, R = 0.958, \textbf{Figure \ref{intradiv}
+person (ANOSIM p-value = 0.008, R = 1, \textbf{Figure \ref{intradiv}
 A}). Network dissimilarity within each person over the 8-10 day sampling
@@
 to other families (other twins and mothers; inter-family) (p-value =
-0.312, \textbf{Figure \ref{intradiv} C}). In addition to the gut, skin
+0.547, \textbf{Figure \ref{intradiv} C}). In addition to the gut, skin
 microbiome network structure was conserved within individuals (p-value
@@
 \textbf{(Supplemental Figure \ref{allskin})}.
+
+As an additional validation measure, we evaluated the tolerance of these
+findings to inaccuracies in the underlying network. As described above,
+our model is not perfect and there is likely to be noise from false
+positive and false negative predictions. We found that additional random
+noise, both by creating a fully connected graph or randomly reducing the
+number of edges to 60\% of the original, changed the statistical
+significance values (p-values) of our findings but not by enough to
+change whether they were statistically significant (p-value \textless{}
+0.05). Therefore the comparisons between groups are resilient to
+potential noise resulting from model false positive and false negative
+predictions \textbf{(Supplemental Figure \ref{idivnoise})}.
 
@@
 of the healthy human skin microbiome between anatomical sites, including
-bacteria, virus, and fungal communities {[}13,47,65{]}. These
+bacteria, virus, and fungal communities {[}13,47,64{]}. These
 communities vary by degree of skin moisture, oil, and environmental
@@
 sites that fluctuate between being moist and dry) were significantly
-less connected than the moist and sebaceous environments (p-value
+more connected than the moist and sebaceous environments (p-value
 \textless{} 0.001, \textbf{Figure \ref{skinnetwork} A)}. Also, skin
-sites that were occluded from the environment were much more highly
-connected than those that were constantly exposed to the environment or
-only intermittently occluded (p-value \textless{} 0.001, \textbf{Figure
-\ref{skinnetwork} B)}.
+sites that were occluded from the environment were less connected than
+those that were constantly exposed to the environment or only
+intermittently occluded (p-value \textless{} 0.001, \textbf{Figure
+\ref{skinnetwork} B)}. We also confirmed that addition of noise to the
+underlying network, as described above, altered the values of
+statistical significance (p-values) but not by enough to change whether
+they were statistically significant \textbf{(Supplemental Figure
+\ref{skindivnoise})}.
 
@@
 microbiome by characterizing bacterial and viral diversity across the
-human body {[}13,14,46--48,67{]}. Here we integrated the bacterial and
+human body {[}13,14,46--48,66{]}. Here we integrated the bacterial and
 viral sequence sets to offer an initial understanding of how
@@
 Just as gut microbiome and virome composition and diversity are
-conserved in individuals {[}13,46,47,66{]}, gut and skin microbiome
+conserved in individuals {[}13,46,47,65{]}, gut and skin microbiome
 network structures were conserved within individuals over time. Gut
@@
 because a randomly removed bacteria or phage node is less likely to
-divide or disintegrate {[}30,64{]} the overall network), the potential
+divide or disintegrate {[}30,63{]} the overall network), the potential
 for horizontal gene transfer between members, and co-evolution of
@@
 environmental fluctuation and exposure (the exposed and intermittently
-exposed sites) were less connected and therefore were predicted to have
+exposed sites) were more connected and therefore were predicted to have
 a higher resilience against network degradation when random nodes are
 removed from the network. Likewise, intermittently moist sites
-demonstrated less connectedness than the moist and sebaceous sites.
+demonstrated higher connectedness than the moist and sebaceous sites.
 These findings agree with previous work that has shown that bacterial
-community networks differ by skin environment types {[}58{]}. Together
+community networks differ by skin environment types {[}57{]}. Together
 these data suggested that body sites under greater degrees of
-fluctuation harbored less connected microbiomes that are potentially
-less robust to network disruption by extinction events. This points to a
-link between microbiome and environmental robustness toward network
-homeostasis and warrants further investigation.
+fluctuation harbored more highly connected microbiomes that are
+potentially more robust to network disruption by extinction events. This
+points to a link between microbiome and environmental robustness toward
+network homeostasis and warrants further investigation.
 
 While these findings take us an important step closer to understanding
-the microbiome through interspecies relationships, there are caveats to
-and considerations regarding our findings. First, as with most
-classification models, the infection classification model developed and
-applied is only as good as its training set -- in this case, the
-collection of experimentally-verified positive and negative infection
-data. Large-scale experimental screens for phage and bacteria infectious
+the microbiome through interspecies relationships, there are caveats and
+considerations to our findings. First, as with most classification
+models, the infection classification model developed and applied is only
+as good as its training set -- in this case, the collection of
+experimentally-verified positive and negative infection data.
+Large-scale experimental screens for phage and bacteria infectious
 interactions that report high-confidence negative interactions (i.e., no
@@
 generation protocols. It is now known that MDA results in biases in
-microbial community composition {[}68{]}, as well as toward ssDNA viral
-genomes {[}69,70{]}, thus rendering the resulting microbial and viral
+microbial community composition {[}67{]}, as well as toward ssDNA viral
+genomes {[}68,69{]}, thus rendering the resulting microbial and viral
 metagenomes largely non-quantitative. Future work that employs larger
@@
 the microbiome field, with a notable example being the use of 16S rRNA
-sequencing using the V4 variable region {[}60{]}. Use of the V4 variable
-region excluded detection of major bacterial members, meaning that the
-findings were not able to completely describe the underlying biological
-environment. Despite this caveat, skin microbiome studies provided
-valuable biological insights by focusing on the community differences
-between groups (e.g.~disease and healthy) which were both analyzed the
-same way. Similarly, here we focused our conclusions on the differences
-between the groups which were all treated the same, so that we can
-minimize our dependence on a perfect predictive model.
-
-Finally, the networks in this study were built using operational genomic
+sequencing using the V4 variable region {[}59{]}. Use of the V4 variable
+region excluded detection of major skin bacterial members, meaning that
+the findings were not able to completely describe the underlying
+biological environment. Despite this caveat, skin microbiome studies
+provided valuable biological insights by focusing on the community
+differences between groups (e.g.~disease and healthy) which were both
+analyzed the same way. Similarly, here we focused our conclusions on the
+differences between the groups which were all treated the same, so that
+we can minimize our dependence on a perfect predictive model. We also
+provided explicit evidence that the introduction of noise equally to the
+compared groups did not significantly impact our findings.
+
+Third, the networks in this study were built using operational genomic
 units (OGUs), which represented groups of highly similar bacteria or
@@
 have been implemented in other metagenomic sequencing studies, as well
-{[}42,71--73{]}. These approaches could offer yet another level of
+{[}42,70--72{]}. These approaches could offer yet another level of
 sophistication to our network-based analyses. While this operationally
@@
 more focused, culture-based studies such as the work by Malki \emph{et
-al} {[}61{]}. Future work must address this limitation, e.g., through
+al} {[}60{]}. Future work must address this limitation, e.g., through
 improved binning methods and deeper metagenomic shotgun sequencing, but
@@
 ecologically and evolutionarily cohesive units for both phage and
-bacteria {[}74{]}. Defining operational genomic units and their
+bacteria {[}73{]}. Defining operational genomic units and their
 taxonomic underpinnings (e.g., whether OGU clusters represent genera or
@@
 gene content, average nucleotide identity of shared genes, and pairwise
-differences between genomes {[}75{]}. Such population-genetic assessment
+differences between genomes {[}74{]}. Such population-genetic assessment
 of phage evolution, coupled with the ecological implications of genome
@@
 the data does agree with previous reports of instances of broad phage
-host range {[}61,76{]}. Additionally, visualization of our dataset
+host range {[}60,75{]}. Additionally, visualization of our dataset
 interactions using the heat map approach previously used in other host
@@
 experimental studies.
+
+Finally, it is important to note that our model was built using
+available full genomes with known interactions, while the experimental
+datasets resulted in OGUs created from metagenomic shotgun sequence
+sets, as described above. While this is an informative approach given
+available data, it is not ideal. We envision future work focusing on
+training models using metagenomic shotgun sample sets from ``mock
+communities'' of bacteria and phages with experimentally validated
+infectious relationships. This would also be more informative than
+relying on simulated metagenomic sample sets, whose use would result in
+models built on simulations and more assumptions instead of empirical
+data. Together this way the training set can be subjected to the same
+pre-processing, contig assembly, and OGU binning processes as the
+experimental data. Furthermore, exciting advances in long read
+sequencing platforms such as the Oxford Nanopore MinIon system will
+provide more accurate genomic scaffolds than \emph{de novo} assembled
+contigs, allowing for more accurate training and predictions of our
+models. As discussed above, it is because our current model is
+susceptible to this noise that we focus our conclusions on comparisons
+between experimental groups that were both treated the same. This is
+also why it was important for us to evaluate the susceptibility of our
+results to noise caused by the less-than-perfect prediction model.
 
@@
 consists of archaea and small eukaryotes, including fungi and
-\emph{Demodex} mites {[}1,77{]} -- all of which can interact with human
-immune cells and other non-microbial community members {[}78{]}. Future
+\emph{Demodex} mites {[}1,76{]} -- all of which can interact with human
+immune cells and other non-microbial community members {[}77{]}. Future
 work will build from our approach and include these additional community
@@
 (v0.0.14) to exclude bases with quality scores below 33 and shorter than
-75 bp {[}79{]}. Paired end reads were filtered to exclude sequences
+75 bp {[}78{]}. Paired end reads were filtered to exclude sequences
 missing their corresponding pair using the
@@
 Contigs were assembled using the Megahit assembly program (v1.0.6)
-{[}80{]}. A minimum contig length of 1 kb was used. Iterative k-mer
+{[}79{]}. A minimum contig length of 1 kb was used. Iterative k-mer
 stepping began at a minimum length of 21 and progressed by 20 until 101.
 All other default parameters were used.
+
+Contig simulations were performed by randomly extracting a string of
+genomic nucleotides that represented a defined percent length of that
+genome. This was accomplished using our
+\texttt{RandomContigGenerator.pl}, which was published in the associated
+GitHub repository.
 
@@
 aligned to phage or bacterial contigs using the Bowtie2 global aligner
-(v2.2.1) {[}81{]}. We defined a mismatch threshold of 1 bp and seed
+(v2.2.1) {[}80{]}. We defined a mismatch threshold of 1 bp and seed
 length of 25 bp. Sequence abundance was calculated from the Bowtie2
@@
 taxonomic units (OTUs), although here we are clustering contigs into
-operational genomic units (OGUs) {[}67{]}.
-
-Contigs were clustered using the CONCOCT algorithm (v0.4.0) {[}82{]}.
+operational genomic units (OGUs) {[}66{]}.
+
+Contigs were clustered using the CONCOCT algorithm (v0.4.0) {[}81{]}.
 Because of our large dataset and limits in computational efficiency, we
@@
 converse analysis of aligning phage OGU representative sequences to EMBL
-bacterial reference genomes. Finally, we ran both the phage and bacteria
-OGU representative sequences through the Virsorter program (1.0.3) to
+bacterial reference genomes. We ran both the phage and bacteria OGU
+representative sequences through the Virsorter program (1.0.3) to
 identify phages (all default parameters were used), using only those in
-the high confidence identification category ``class 1'' {[}51{]}.
+the high confidence identification category ``class 1'' {[}50{]}.
+Finally, we filtered out phage OGUs that had bacterial elements as
+described above, but also lacked known phage elements by using the
+tblastx algorithm and a maximum e-value of \(10^{-25}\).
 
@@
 Open reading frames (ORFs) were identified using the Prodigal program
-(V2.6.2) with the meta mode parameter and default settings {[}83{]}.
+(V2.6.2) with the meta mode parameter and default settings {[}82{]}.
 
@@
 experimentally validated bacteria-phage infections or validated lack of
-infections from six studies {[}41,52--56{]}. No further reference
+infections from six studies {[}41,51--55{]}. No further reference
 databases were used in our alignment procedures. Associated reference
@@
 interactions or not. The classification model was built using the Caret
-R package (v6.0.73) {[}84{]}. The model was trained using five-fold
+R package (v6.0.73) {[}83{]}. The model was trained using five-fold
 cross validation with ten repeats, and the median model performance was
@@
 identified from bacterial genomes using the PilerCR program (v1.06)
-{[}85{]}. Resulting spacer sequences were filtered to exclude spacers
+{[}84{]}. Resulting spacer sequences were filtered to exclude spacers
 shorter than 20 bp and longer than 65 bp. Spacer sequences were aligned
 to the phage genomes using the nucleotide BLAST algorithm with default
-parameters (v2.4.0) {[}86{]}. The mean percent identity for each
+parameters (v2.4.0) {[}85{]}. The mean percent identity for each
 matching pair was recorded for use in our classification model.
@@
 between the genes using the Diamond protein alignment algorithm
-(v0.7.11.60) {[}87{]}. The mean alignment bitscores for each genome pair
+(v0.7.11.60) {[}86{]}. The mean alignment bitscores for each genome pair
 were recorded for use in our classification model.
@@
 predicted to interact using the Pfam interaction information within the
-Intact database {[}57{]}. The mean bitscores of the matches between each
+Intact database {[}56{]}. The mean bitscores of the matches between each
 pair were recorded for use in the classification model.
+
+\subsection{Secondary Dataset
+Validation}\label{secondary-dataset-validation}
+
+The performance of our model for identifying diverse infectious
+relationships between bacteria and phages, beyond those that were
+included in the model creation step, were validated using additional
+bacterial and phage reference genomes, which could be linked by the
+records of which phage strains were isolated on which bacteria under
+laboratory conditions. Viral and bacterial reference genomes were
+downloaded from the GenBank repository on February 19, 2018 using the
+viral location \texttt{ftp://ftp.ncbi.nih.gov/refseq/release/viral/} and
+the bacterial location
+\texttt{ftp://ftp.ncbi.nih.gov/refseq/release/bacteria/}. This resulted
+in the use of 539 complete phages reference genomes (with identified
+hosts) and 3,469 bacterial reference genomes. We used the same
+prediction model to predict which phages were infecting which hosts, so
+as to confirm that the model was capable of identifying interactions in
+a more diverse dataset. Bacteria interactions were identified at the
+species level. The random contig iteration analysis was performed using
+a subset of bacterial reference genomes, for computational performance
+reasons. Only single representative genomes for each species were used.
 
@@
 associated metadata were used to populate a graph database using Neo4j
-graph database software (v2.3.1) {[}88{]}. This network was used for
-downstream community analysis.
+graph database software (v2.3.1) {[}87{]}. This network was used for
+downstream community analysis. Tolerance to false negative and false
+positive noise within the networks was assessed by randomly removing a
+defined fraction of network edges before re-running the downstream
+analysis work flows. This was accomplished using functionality within
+the igraph R package (v1.0.1) {[}88{]}.
 
@@
 Degree and closeness centrality were calculated using the associated
-functions within the igraph R package (v1.0.1) {[}89{]}.
+functions within the igraph R package (v1.0.1) {[}88{]}.
 
@@
 \includegraphics[width=1.00000\textwidth]{}
+\caption{\textbf{Ability of prediction model to identify broad range of
+bacteria and phage interactions.} \emph{Each complete bacteriophage
+labeled on the y axis. The number of complete bacterial species genomes
+that were correctly predicted to be infected by each phage, according to
+the GenBank records, is on the x-axis. True positive interactions are
+colored in purple, and false negative interactions are yellow. Because
+this dataset only had confirmed interactions and not confirmed lack of
+interactions, false positive and true negative values could be not
+determined.}\label{secondval}}
+\end{figure}
+
+\newpage
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=1.00000\textwidth]{}
+\caption{\textbf{Impact of incomplete genomic sequences on model
+performance.} \emph{Number of correctly identified infectious
+interactions between bacteria and phages are presented on the y-axis.
+The fraction of genomic length that was used to randomly extract contigs
+from reference bacterial and phage sequences is presented on the x-axis
+(e.g.~0.9 means contig lengths were 90\% of the total genome length).
+Overall this presents a quantification of the loss of identified
+infectious relationships as the percent of available genomic material is
+reduced.}\label{contigval}}
+\end{figure}
+
+\newpage
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=1.00000\textwidth]{}
 \caption{\textbf{Classification Model Performance By Nested
@@
 samples were randomly withheld during the training stage for model
-evaluation and mtry tuning. The results of 25 random iterations are
-shown. Metrics include area under the curve (red), sensitivity (green),
-and specificity (blue). Dashed line highlight the random point of
-0.5.}\label{crossval}}
+evaluation and mtry tuning (parameter defined in the R Random Forest
+package, which is implemented in Caret, as ``the number of variables
+randomly sampled as candidates at each split''). The results of 25
+random iterations are shown. Metrics include area under the curve (red),
+sensitivity (green), and specificity (blue). Dashed line highlight the
+random point of 0.5.}\label{crossval}}
 \end{figure}
@@
 
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.75000\textwidth]{}
+\caption{\textbf{P-values of interpersonal group diversity differences
+with graph edge noise.} \emph{The x-axis represents the percent of
+errors that were randomly removed (or added) as a means to evaluate the
+impact of random noise on statistical significance of group differences.
+Resulting p-values for each graph is shown on the y-axis. The dot and
+bars are the mean and standard error of five iterations of group
+comparisons with random edge removal. Significance of A) ANOSIM p-value
+of diet network dissimilarity, B) p-value of interpersonal and
+intrapersonal diet network dissimilarity, C) p-value of interpersonal
+and intrapersonal skin network dissimilarity, and D) p-value of
+interpersonal and intrapersonal twin gut network dissimilarity. This
+corresponds to the findings in Figure 3.}\label{idivnoise}}
+\end{figure}
+
+\newpage
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.75000\textwidth]{}
+\caption{\textbf{P-values of differences in Eigen Centrality between
+skin site microbiome networks.} \emph{The x-axis represents the percent
+of errors that were randomly removed (or added) as a means to evaluate
+the impact of random noise on statistical significance of group
+differences. Resulting p-values for each graph is shown on the y-axis.
+The dot and bars are the mean and standard error of five iterations of
+group comparisons with random edge removal. The groups compared were the
+degrees of site moisture (left) and occlusion (right). The findings
+correspond to pannels A and B in Figure 4.}\label{skindivnoise}}
+\end{figure}
+
+\newpage
+
 \section{Supplemental Table Captions}\label{supplemental-table-captions}
@@
 the original publications of the viromes used in this current study.
+
+Table S2: The positive and negative bacteria and bacteriophage
+interactions used to train the prediction model, as also illustrated in
+Figure \ref{ValidationOverview}. Citation sources are also included.
 
@@
 
-\hypertarget{ref-Roux:2013cr}{}
-50. Roux S, Krupovic M, Debroas D, Forterre P, Enault F. Assessment of
-viral community functional potential from viral metagenomes may be
-hampered by contamination with cellular sequences. Open Biology. 2013;3:
-130160--130160.
-
 \hypertarget{ref-Roux:2015dh}{}
-51. Roux S, Enault F, Hurwitz BL, Sullivan MB. VirSorter: mining viral
+50. Roux S, Enault F, Hurwitz BL, Sullivan MB. VirSorter: mining viral
 signal from microbial genomic data. PeerJ. 2015;3: e985--20.
@@
 \hypertarget{ref-Jensen:1998vh}{}
-52. Jensen EC, Schrader HS, Rieland B, Thompson TL, Lee KW, Nickerson
+51. Jensen EC, Schrader HS, Rieland B, Thompson TL, Lee KW, Nickerson
 KW, et al. Prevalence of broad-host-range lytic bacteriophages of
@@
 \hypertarget{ref-Malki:2015tm}{}
-53. Malki K, Kula A, Bruder K, Sible E. Bacteriophages isolated from
+52. Malki K, Kula A, Bruder K, Sible E. Bacteriophages isolated from
 Lake Michigan demonstrate broad host-range across several bacterial
@@
 \hypertarget{ref-Schwarzer:2012ez}{}
-54. Schwarzer D, Buettner FFR, Browning C, Nazarov S, Rabsch W, Bethe A,
+53. Schwarzer D, Buettner FFR, Browning C, Nazarov S, Rabsch W, Bethe A,
 et al. A multivalent adsorption apparatus explains the broad host range
@@
 \hypertarget{ref-Kim:2012dh}{}
-55. Kim S, Rahman M, Seol SY, Yoon SS, Kim J. Pseudomonas aeruginosa
+54. Kim S, Rahman M, Seol SY, Yoon SS, Kim J. Pseudomonas aeruginosa
 bacteriophage PA1Ø requires type IV pili for infection and shows broad
@@
 \hypertarget{ref-Matsuzaki:1992gw}{}
-56. Matsuzaki S, Tanaka S, Koga T, Kawata T. A Broad-Host-Range
+55. Matsuzaki S, Tanaka S, Koga T, Kawata T. A Broad-Host-Range
 Vibriophage, KVP40, Isolated from Sea Water. Microbiology and
@@
 \hypertarget{ref-Orchard:2014hq}{}
-57. Orchard S, Ammari M, Aranda B, Breuza L, Briganti L, Broackes-Carter
+56. Orchard S, Ammari M, Aranda B, Breuza L, Briganti L, Broackes-Carter
 F, et al. The MIntAct project--IntAct as a common curation platform for
@@
 \hypertarget{ref-Bashan:2016is}{}
-58. Bashan A, Gibson TE, Friedman J, Carey VJ, Weiss ST, Hohmann EL, et
+57. Bashan A, Gibson TE, Friedman J, Carey VJ, Weiss ST, Hohmann EL, et
 al. Universality of human microbial dynamics. Nature. 2016;534:
@@
 \hypertarget{ref-Kleiner:2015kd}{}
-59. Kleiner M, Hooper LV, Duerkop BA. Evaluation of methods to purify
+58. Kleiner M, Hooper LV, Duerkop BA. Evaluation of methods to purify
 virus-like particles for metagenomic sequencing of intestinal viromes.
@@
 \hypertarget{ref-Meisel:2016iu}{}
-60. Meisel JS, Hannigan GD, Tyldsley AS, SanMiguel AJ, Hodkinson BP,
+59. Meisel JS, Hannigan GD, Tyldsley AS, SanMiguel AJ, Hodkinson BP,
 Zheng Q, et al. Skin microbiome surveys are strongly influenced by
@@
 \hypertarget{ref-Malki:2015dw}{}
-61. Malki K, Kula A, Bruder K, Sible E, Hatzopoulos T, Steidel S, et al.
+60. Malki K, Kula A, Bruder K, Sible E, Hatzopoulos T, Steidel S, et al.
 Bacteriophages isolated from Lake Michigan demonstrate broad host-range
@@
 \hypertarget{ref-Turnbaugh:2009hf}{}
-62. Turnbaugh PJ, Ridaura VK, Faith JJ, Rey FE, Knight R, Gordon JI. The
+61. Turnbaugh PJ, Ridaura VK, Faith JJ, Rey FE, Knight R, Gordon JI. The
 effect of diet on the human gut microbiome: a metagenomic analysis in
@@
 \hypertarget{ref-David:2014cl}{}
-63. David LA, Maurice CF, Carmody RN, Gootenberg DB, Button JE, Wolfe
+62. David LA, Maurice CF, Carmody RN, Gootenberg DB, Button JE, Wolfe
 BE, et al. Diet rapidly and reproducibly alters the human gut
@@
 \hypertarget{ref-Jeong:2000ima}{}
-64. Jeong H, Tombor B, Albert R, Oltvai ZN, Barabási AL. The large-scale
+63. Jeong H, Tombor B, Albert R, Oltvai ZN, Barabási AL. The large-scale
 organization of metabolic networks. Nature. 2000;407: 651--654.
@@
 \hypertarget{ref-Grice:2009ee}{}
-65. Grice EA, Kong HH, Conlan S, Deming CB, Davis J, Young AC, et al.
+64. Grice EA, Kong HH, Conlan S, Deming CB, Davis J, Young AC, et al.
 Topographical and Temporal Diversity of the Human Skin Microbiome.
@@
 \hypertarget{ref-Minot:2013ih}{}
-66. Minot S, Bryson A, Chehoud C, Wu GD, Lewis JD, Bushman FD. Rapid
+65. Minot S, Bryson A, Chehoud C, Wu GD, Lewis JD, Bushman FD. Rapid
 evolution of the human gut virome. Proceedings of the National Academy
@@
 \hypertarget{ref-Schloss:2005hz}{}
-67. Schloss PD, Handelsman J. Introducing DOTUR, a computer program for
+66. Schloss PD, Handelsman J. Introducing DOTUR, a computer program for
 defining operational taxonomic units and estimating species richness.
@@
 \hypertarget{ref-Yilmaz:2010jb}{}
-68. Yilmaz S, Allgaier M, Hugenholtz P. Multiple displacement
+67. Yilmaz S, Allgaier M, Hugenholtz P. Multiple displacement
 amplification compromises quantitative analysis of metagenomes. Nature
@@
 \hypertarget{ref-Kim:2008to}{}
-69. Kim KH, Chang HW, Nam YD, Roh SW. Amplification of uncultured
+68. Kim KH, Chang HW, Nam YD, Roh SW. Amplification of uncultured
 single-stranded DNA viruses from rice paddy soil. Applied and \ldots{}.
@@
 \hypertarget{ref-Kim:2011hp}{}
-70. Kim K-H, Bae J-W. Amplification methods bias metagenomic libraries
+69. Kim K-H, Bae J-W. Amplification methods bias metagenomic libraries
 of uncultured single-stranded and double-stranded DNA viruses. Applied
@@
 \hypertarget{ref-Minot:2012ed}{}
-71. Minot S, Wu GD, Lewis JD, Bushman FD. Conservation of gene cassettes
+70. Minot S, Wu GD, Lewis JD, Bushman FD. Conservation of gene cassettes
 among diverse viruses of the human gut. PLOS ONE. 2012;7: e42342.
@@
 \hypertarget{ref-Deng:2014eb}{}
-72. Deng L, Ignacio-Espinoza JC, Gregory AC, Poulos BT, Weitz JS,
+71. Deng L, Ignacio-Espinoza JC, Gregory AC, Poulos BT, Weitz JS,
 Hugenholtz P, et al. Viral tagging reveals discrete populations in
@@
 \hypertarget{ref-Brum:2015iaa}{}
-73. Brum JR, Ignacio-Espinoza JC, Roux S, Doulcier G, Acinas SG, Alberti
+72. Brum JR, Ignacio-Espinoza JC, Roux S, Doulcier G, Acinas SG, Alberti
 A, et al. Ocean plankton. Patterns and ecological drivers of ocean viral
@@
 \hypertarget{ref-Polz:2006fi}{}
-74. Polz MF, Hunt DE, Preheim SP, Weinreich DM. Patterns and mechanisms
+73. Polz MF, Hunt DE, Preheim SP, Weinreich DM. Patterns and mechanisms
 of genetic and phenotypic differentiation in marine microbes.
@@
 \hypertarget{ref-Gregory:2016cg}{}
-75. Gregory AC, Solonenko SA, Ignacio-Espinoza JC, LaButti K, Copeland
+74. Gregory AC, Solonenko SA, Ignacio-Espinoza JC, LaButti K, Copeland
 A, Sudek S, et al. Genomic differentiation among wild cyanophages
@@
 \hypertarget{ref-PaezEspino:2016gi}{}
-76. Paez-Espino D, Eloe-Fadrosh EA, Pavlopoulos GA, Thomas AD, Huntemann
+75. Paez-Espino D, Eloe-Fadrosh EA, Pavlopoulos GA, Thomas AD, Huntemann
 M, Mikhailova N, et al. Uncovering Earth's virome. Nature. 2016;
@@
 \hypertarget{ref-Grice:2011gy}{}
-77. Grice EA, Segre JA. The skin microbiome. Nature Reviews
+76. Grice EA, Segre JA. The skin microbiome. Nature Reviews
 Microbiology. 2011;9: 244--253.
@@
 \hypertarget{ref-Round:2009bz}{}
-78. Round JL, Mazmanian SK. The gut microbiota shapes intestinal immune
+77. Round JL, Mazmanian SK. The gut microbiota shapes intestinal immune
 responses during health and disease. Nature reviews Immunology. 2009;9:
@@
 \hypertarget{ref-FASTXToolkit:wr}{}
-79. Hannon GJ. FASTX-Toolkit. 2010; GNU Affero General Public License.
+78. Hannon GJ. FASTX-Toolkit. 2010; GNU Affero General Public License.
 
 \hypertarget{ref-Li:2016kd}{}
-80. Li D, Luo R, Liu C-M, Leung C-M, Ting H-F, Sadakane K, et al.
+79. Li D, Luo R, Liu C-M, Leung C-M, Ting H-F, Sadakane K, et al.
 MEGAHIT v1.0: A fast and scalable metagenome assembler driven by
@@
 \hypertarget{ref-Langmead:2012jh}{}
-81. Langmead B, Salzberg SL. Fast gapped-read alignment with Bowtie 2.
+80. Langmead B, Salzberg SL. Fast gapped-read alignment with Bowtie 2.
 Nature Methods. 2012;9: 357--359.
@@
 \hypertarget{ref-Alneberg:2014fc}{}
-82. Alneberg J, Bjarnason BS aacute ri, Bruijn I de, Schirmer M, Quick
+81. Alneberg J, Bjarnason BS aacute ri, Bruijn I de, Schirmer M, Quick
 J, Ijaz UZ, et al. Binning metagenomic contigs by coverage and
@@
 \hypertarget{ref-Hyatt:2012cy}{}
-83. Hyatt D, LoCascio PF, Hauser LJ, Uberbacher EC. Gene and translation
+82. Hyatt D, LoCascio PF, Hauser LJ, Uberbacher EC. Gene and translation
 initiation site prediction in metagenomic sequences. Bioinformatics.
@@
 \hypertarget{ref-caretClassificatio:ux5fU2Litux5f1}{}
-84. Kuhn M. caret: Classification and Regression Training. CRAN. 2016;
+83. Kuhn M. caret: Classification and Regression Training. CRAN. 2016;
 
 \hypertarget{ref-Edgar:2007bh}{}
-85. Edgar RC. PILER-CR: fast and accurate identification of CRISPR
+84. Edgar RC. PILER-CR: fast and accurate identification of CRISPR
 repeats. BMC Bioinformatics. 2007;8: 18.
@@
 \hypertarget{ref-Camacho:2009fc}{}
-86. Camacho C, Coulouris G, Avagyan V, Ma N, Papadopoulos J, Bealer K,
+85. Camacho C, Coulouris G, Avagyan V, Ma N, Papadopoulos J, Bealer K,
 et al. BLAST+: architecture and applications. BMC Bioinformatics.
@@
 \hypertarget{ref-Buchfink:2015ki}{}
-87. Buchfink B, Xie C, Huson DH. Fast and sensitive protein alignment
+86. Buchfink B, Xie C, Huson DH. Fast and sensitive protein alignment
 using DIAMOND. Nature Methods. 2015;12: 59--60.
@@
 \hypertarget{ref-Neoj:Fuwr6PBN}{}
-88. Neo Technology, Inc. Neo4j. 2017;
+87. Neo Technology, Inc. Neo4j. 2017;
 
 \hypertarget{ref-Theigraphsoftware:vh}{}
-89. Csardi G, Nepusz T. The igraph software package for complex network
+88. Csardi G, Nepusz T. The igraph software package for complex network
 research. InterJournal. 2006;Complex Systems: 1695.

### 2018-03-07 doc/manuscript.Rmd
updated with pats edits and fixed a bug with edge count near beginning.

@@
 nestats <- read.delim(file = "../rtables/nestats.tsv", header = TRUE, sep = "\t")
+totalbpstats <- read.delim(file = "../rtables/totalbp.tsv", header = TRUE, sep = "\t")
 sitestats <- read.delim(file = "../rtables/sitestattable.tsv", header = TRUE, sep = "\t")
@@
 The human microbiome, the collection of microbial communities that colonize the human body, is a crucial component to health and disease.
-Two major components to the human microbiome are the bacterial and viral communities.
+Two major components of the human microbiome are the bacterial and viral communities.
 These communities have primarily been studied separately using metrics of community composition and diversity.
@@
 
-The bacterial and viral sequences from these studies were quality filtered and assembled into contigs ("genomic fragments").
+The bacterial and viral sequences from these studies were quality filtered and assembled into contigs (i.e.
+genomic fragments).
 We further grouped the related bacterial and phage contigs into operationally defined units based on their k-mer frequencies and co-abundance patterns, similar to previous reports **(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})** [@Roux:2016cc].
@@
 Additionally, no OGUs were confidently identified as lytic or temperate phage OGUs in the bacterial dataset using the Virsorter algorithm [@Roux:2015dh].
-
 We also supplemented the previous virome fraction quality control measures **(Table S1)** to find that, in light of the rigorous purification and quality control during sample collection and preparation, 77% (228 / 298 operational genomic units) still had some nucleotide similarity to a given bacterial reference genome (e-value < $10^{-25}$).
-It is important to note that interpreting such alignment is complicated by the fact that most reference bacterial genomes also contain prophages (phages integrated into bacterial genomes), meaning we do not know to what extent the alignments were the result of bacterial contaminants in the virome fraction and what were true integrated prophages.
-As most phages in these communities have been shown to be temperate (they integrate into bacterial genomes) using methods that included nucleotide alignments of phages to bacterial reference genomes, we expected that a large fraction of those phages were temperate and therefore shared elements with bacterial reference genomes -- a trend previously reported [@Minot:2011ez, @Hannigan:2015fz, @Reyes:2010cwa].
+It is important to note that interpreting such alignment is complicated by the fact that most reference bacterial genomes also contain prophages (i.e.
+phages integrated into bacterial genomes), meaning we do not know to what extent the alignments were the result of bacterial contaminants in the virome fraction and what were true integrated prophages.
+As most phages in these communities have been shown to be temperate (i.e.
+they integrate into bacterial genomes) using methods that included nucleotide alignments of phages to bacterial reference genomes, we expected that a large fraction of those phages were temperate and therefore shared elements with bacterial reference genomes -- a trend previously reported [@Minot:2011ez, @Hannigan:2015fz, @Reyes:2010cwa].
 To ensure the purity of our sample sets, we supplemented the quality control measures by filtering out all OGUs that could be potential bacterial contaminants, as described previously [@Roux:2016cc].
@@
 Because we used DNA sequencing studies, RNA phages were not considered **(Supplemental Figure \ref{ValidationOverview} C-D)**.
-This training set included both positive relationships (a phage infects a bacterium) and negative relationships (a phage does not infect a bacterium).
+This training set included both positive relationships (i.e.
+a phage infects a bacterium) and negative relationships (i.e.
+a phage does not infect a bacterium).
 This allowed us to validate the false positive and false negative rates associated with our candidate models, thereby building upon previous work that only considered positive relationships [@Edwards:2015iz].
@@
 These features were calculated using the training set described above.
-While the nucleotide and amino acid similarity metrics are expected to identify prophage signatures, the protein family interaction and CRISPR signatures are expected to aid in identifying lytic phages in addition to temperate phages.
-We chose to utilize these metrics that directly compare nucleotide sequences between sample phages and bacteria, instead of relying on alignment to reference genomes or known marker genes, because we are extrapolating our model to highly diverse communities which we expect to diverge significantly from the available reference genomes.
+While the nucleotide and amino acid similarity metrics were expected to identify prophage signatures, the protein family interaction and CRISPR signatures were expected to aid in identifying lytic phages in addition to temperate phages.
+We chose to utilize these metrics that directly compare nucleotide sequences between sample phages and bacteria, instead of relying on alignment to reference genomes or known marker genes, because we were extrapolating our model to highly diverse communities which we expect to diverge significantly from the available reference genomes.
 The resulting random forest model was assessed using nested cross validation, and the median area under its receiver operating characteristic (ROC) curve was `r signif(averageauc[c(averageauc$type %in% "AUC"), 2], digits = sigfig)`, the median model sensitivity was `r signif(averageauc[c(averageauc$type %in% "SENS"), 2], digits = sigfig)`, and median specificity was `r signif(averageauc[c(averageauc$type %in% "SPEC"), 2], digits = sigfig)` **(Figure \ref{RocCurve} A)**.
-This balance of confident true positives at the cost of fewer true negatives is ideal for this type of dataset which consists of primarily positive connections **(Supplemental Figure \ref{nestedperformance})**.
+This balance of confident true positives at the cost of fewer true negatives was ideal for this type of dataset which consisted of primarily positive connections **(Supplemental Figure \ref{nestedperformance})**.
 Nested cross validation of the model demonstrated that the sensitivity and specificity of the model could vary but the overall model performance (AUC) remained more consistent **(Supplemental Figure \ref{crossval})**.
@@
 Relationships were weighted by the relative abundances of their associated bacteria and phages.
-Like the master network, the skin network exhibited a diameter of `r prettyNum(genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Diameter"], big.mark = ",")` (measure of graph size; the greatest number of traversed vertices required between two vertices) and included `r prettyNum(genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Vertices"], big.mark = ",")` (`r prettyNum(sitestats[c(sitestats$class %in% "skin"),"phagenodes"], big.mark = ",")` phages, `r prettyNum(sitestats[c(sitestats$class %in% "skin"),"bacterianodes"], big.mark = ",")` bacteria, `r round(100 * genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Vertices"] / genfigurestats[c(genfigurestats$class %in% "Total"),"Vertices"], digits = 1)`% total) and `r prettyNum(genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Edges"], big.mark = ",")` (`r round(100 * genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Edges"] / genfigurestats[c(genfigurestats$class %in% "Total"),"Edges"], digits = 1)`%) of the master network nodes and edges, respectively **(Figure \ref{RocCurve} E - F)**.
+Like the master network, the skin network exhibited a diameter of `r prettyNum(genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Diameter"], big.mark = ",")` (measure of graph size; the greatest number of traversed vertices required between two vertices) and included `r prettyNum(genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Vertices"], big.mark = ",")` (`r prettyNum(totalbpstats[c(totalbpstats$class %in% "skin"),"phagenodes"], big.mark = ",")` phages, `r prettyNum(totalbpstats[c(totalbpstats$class %in% "skin"),"bacterianodes"], big.mark = ",")` bacteria, `r round(100 * genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Vertices"] / genfigurestats[c(genfigurestats$class %in% "Total"),"Vertices"], digits = 1)`% total) and `r prettyNum(genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Edges"], big.mark = ",")` (`r round(100 * genfigurestats[c(genfigurestats$class %in% "SkinStudy"),"Edges"] / genfigurestats[c(genfigurestats$class %in% "Total"),"Edges"], digits = 1)`%) of the master network nodes and edges, respectively **(Figure \ref{RocCurve} E - F)**.
 Additionally, the subnetworks demonstrated narrow ranges of eccentricity across their nodes **(Supplemental Figure \ref{eccplot})**.
-The phages and bacteria in the diet and twin sample sets were more sparsely related, with the diet study consisting of `r prettyNum(genfigurestats[c(genfigurestats$class %in% "DietStudy"),"Vertices"], big.mark = ",")` (`r prettyNum(sitestats[c(sitestats$class %in% "diet"),"phagenodes"], big.mark = ",")` phages, `r prettyNum(sitestats[c(sitestats$class %in% "diet"),"bacterianodes"], big.mark = ",")` bacteria) nodes and `r prettyNum(genfigurestats[c(genfigurestats$class %in% "DietStudy"),"Edges"], big.mark = ",")` relationships, and the twin study containing `r prettyNum(genfigurestats[c(genfigurestats$class %in% "TwinStudy"),"Vertices"], big.mark = ",")` (`r prettyNum(sitestats[c(sitestats$class %in% "twin"),"phagenodes"], big.mark = ",")` phages, `r prettyNum(sitestats[c(sitestats$class %in% "twin"),"bacterianodes"], big.mark = ",")` bacteria) nodes and `r prettyNum(genfigurestats[c(genfigurestats$class %in% "TwinStudy"),"Edges"], big.mark = ",")` relationships **(Figure \ref{RocCurve} E - F)**.
-As an interesting validation measure, we identified five (1.7%) examples of phage OGUs which contained similar genomic elements to the four previously described, broadly infectious phages isolated from Lake Michigan (tblastx; e-value < $10^{-25}$) [@Malki:2015dw].
+The phages and bacteria in the diet and twin sample sets were more sparsely related, with the diet study consisting of `r prettyNum(genfigurestats[c(genfigurestats$class %in% "DietStudy"),"Vertices"], big.mark = ",")` (`r prettyNum(totalbpstats[c(totalbpstats$class %in% "diet"),"phagenodes"], big.mark = ",")` phages, `r prettyNum(totalbpstats[c(totalbpstats$class %in% "diet"),"bacterianodes"], big.mark = ",")` bacteria) nodes and `r prettyNum(totalbpstats[c(totalbpstats$class %in% "diet"),"edgecount"], big.mark = ",")` relationships, and the twin study containing `r prettyNum(genfigurestats[c(genfigurestats$class %in% "TwinStudy"),"Vertices"], big.mark = ",")` (`r prettyNum(totalbpstats[c(totalbpstats$class %in% "twin"),"phagenodes"], big.mark = ",")` phages, `r prettyNum(totalbpstats[c(totalbpstats$class %in% "twin"),"bacterianodes"], big.mark = ",")` bacteria) nodes and `r prettyNum(totalbpstats[c(totalbpstats$class %in% "twin"),"edgecount"], big.mark = ",")` relationships **(Figure \ref{RocCurve} E - F)**.
+As a validation measure, we identified five (1.7%) examples of phage OGUs which contained similar genomic elements to the four previously described, broadly infectious phages isolated from Lake Michigan (tblastx; e-value < $10^{-25}$) [@Malki:2015dw].
 

### 2018-03-07 doc/manuscript.tex
updated with pats edits and fixed a bug with edge count near beginning.

@@
 colonize the human body, is a crucial component to health and disease.
-Two major components to the human microbiome are the bacterial and viral
+Two major components of the human microbiome are the bacterial and viral
 communities. These communities have primarily been studied separately
@@
 The bacterial and viral sequences from these studies were quality
-filtered and assembled into contigs (``genomic fragments''). We further
+filtered and assembled into contigs (i.e.~genomic fragments). We further
 grouped the related bacterial and phage contigs into operationally
@@
 phage OGUs in the bacterial dataset using the Virsorter algorithm
-{[}50{]}.
-
-We also supplemented the previous virome fraction quality control
-measures \textbf{(Table S1)} to find that, in light of the rigorous
-purification and quality control during sample collection and
+{[}50{]}. We also supplemented the previous virome fraction quality
+control measures \textbf{(Table S1)} to find that, in light of the
+rigorous purification and quality control during sample collection and
 preparation, 77\% (228 / 298 operational genomic units) still had some
@@
 such alignment is complicated by the fact that most reference bacterial
-genomes also contain prophages (phages integrated into bacterial
+genomes also contain prophages (i.e.~phages integrated into bacterial
 genomes), meaning we do not know to what extent the alignments were the
@@
 true integrated prophages. As most phages in these communities have been
-shown to be temperate (they integrate into bacterial genomes) using
+shown to be temperate (i.e.~they integrate into bacterial genomes) using
 methods that included nucleotide alignments of phages to bacterial
@@
 \textbf{(Supplemental Figure \ref{ValidationOverview} C-D)}. This
-training set included both positive relationships (a phage infects a
-bacterium) and negative relationships (a phage does not infect a
+training set included both positive relationships (i.e.~a phage infects
+a bacterium) and negative relationships (i.e.~a phage does not infect a
 bacterium). This allowed us to validate the false positive and false
@@
 using the training set described above. While the nucleotide and amino
-acid similarity metrics are expected to identify prophage signatures,
-the protein family interaction and CRISPR signatures are expected to aid
-in identifying lytic phages in addition to temperate phages. We chose to
-utilize these metrics that directly compare nucleotide sequences between
-sample phages and bacteria, instead of relying on alignment to reference
-genomes or known marker genes, because we are extrapolating our model to
-highly diverse communities which we expect to diverge significantly from
-the available reference genomes. The resulting random forest model was
-assessed using nested cross validation, and the median area under its
-receiver operating characteristic (ROC) curve was 0.788, the median
-model sensitivity was 0.905, and median specificity was 0.538
-\textbf{(Figure \ref{RocCurve} A)}. This balance of confident true
-positives at the cost of fewer true negatives is ideal for this type of
-dataset which consists of primarily positive connections
-\textbf{(Supplemental Figure \ref{nestedperformance})}. Nested cross
-validation of the model demonstrated that the sensitivity and
-specificity of the model could vary but the overall model performance
-(AUC) remained more consistent \textbf{(Supplemental Figure
-\ref{crossval})}. This suggested that our model would perform with a
-similar overall accuracy despite changes in sensitivity/specificity
-trade-offs. The most important predictor in the model was amino acid
-similarity between genes, followed by nucleotide similarity of whole
-genomes \textbf{(Figure \ref{RocCurve} B)}. Protein family interactions
-were moderately important to the model, and CRISPRs were largely
-uninformative, due to the minimal amount of identifiable CRISPRs in the
-dataset and their redundancy with the nucleotide similarity methods
-\textbf{(Figure \ref{RocCurve} B)}. Approximately one third of the
-training set relationships yielded no score and therefore were unable to
-be assigned an interaction prediction \textbf{(Figure \ref{RocCurve}
-C)}.
+acid similarity metrics were expected to identify prophage signatures,
+the protein family interaction and CRISPR signatures were expected to
+aid in identifying lytic phages in addition to temperate phages. We
+chose to utilize these metrics that directly compare nucleotide
+sequences between sample phages and bacteria, instead of relying on
+alignment to reference genomes or known marker genes, because we were
+extrapolating our model to highly diverse communities which we expect to
+diverge significantly from the available reference genomes. The
+resulting random forest model was assessed using nested cross
+validation, and the median area under its receiver operating
+characteristic (ROC) curve was 0.788, the median model sensitivity was
+0.905, and median specificity was 0.538 \textbf{(Figure \ref{RocCurve}
+A)}. This balance of confident true positives at the cost of fewer true
+negatives was ideal for this type of dataset which consisted of
+primarily positive connections \textbf{(Supplemental Figure
+\ref{nestedperformance})}. Nested cross validation of the model
+demonstrated that the sensitivity and specificity of the model could
+vary but the overall model performance (AUC) remained more consistent
+\textbf{(Supplemental Figure \ref{crossval})}. This suggested that our
+model would perform with a similar overall accuracy despite changes in
+sensitivity/specificity trade-offs. The most important predictor in the
+model was amino acid similarity between genes, followed by nucleotide
+similarity of whole genomes \textbf{(Figure \ref{RocCurve} B)}. Protein
+family interactions were moderately important to the model, and CRISPRs
+were largely uninformative, due to the minimal amount of identifiable
+CRISPRs in the dataset and their redundancy with the nucleotide
+similarity methods \textbf{(Figure \ref{RocCurve} B)}. Approximately one
+third of the training set relationships yielded no score and therefore
+were unable to be assigned an interaction prediction \textbf{(Figure
+\ref{RocCurve} C)}.
 
@@
 size; the greatest number of traversed vertices required between two
-vertices) and included 433 ( phages, bacteria, 99.5\% total) and 38,099
-(99.4\%) of the master network nodes and edges, respectively
+vertices) and included 433 (154 phages, 279 bacteria, 99.5\% total) and
+38,099 (99.4\%) of the master network nodes and edges, respectively
 \textbf{(Figure \ref{RocCurve} E - F)}. Additionally, the subnetworks
@@
 the diet and twin sample sets were more sparsely related, with the diet
-study consisting of 80 ( phages, bacteria) nodes and 4,692
-relationships, and the twin study containing 130 ( phages, bacteria)
-nodes and 14,684 relationships \textbf{(Figure \ref{RocCurve} E - F)}.
-As an interesting validation measure, we identified five (1.7\%)
-examples of phage OGUs which contained similar genomic elements to the
-four previously described, broadly infectious phages isolated from Lake
+study consisting of 80 (32 phages, 48 bacteria) nodes and 1,290
+relationships, and the twin study containing 130 (29 phages, 101
+bacteria) nodes and 2,457 relationships \textbf{(Figure \ref{RocCurve} E
+- F)}. As a validation measure, we identified five (1.7\%) examples of
+phage OGUs which contained similar genomic elements to the four
+previously described, broadly infectious phages isolated from Lake
 Michigan (tblastx; e-value \textless{} \(10^{-25}\)) {[}60{]}.

### 2018-03-07 revfiles/ReviewerResponse.md
Pats updates to responses.

@@
 
-> *We agree with these concerns and we think that the comments from the reviewer have really helped address them.*
+> *We agree with these concerns and we think that the comments from the reviewer have really helped address them as described in our responses below..*
 
@@
 These communities have been previously described as consisting primarily of lysogenic phages.
-Therefore, while our approach could be improved by future work and more comprehensive datasets, we are addressing the most relevant component of the virome.
-Second, while the alignment approaches were most informative, the protein family interaction approach was also very informative.
-Because the protein family approach is based on experimentally validated protein-protein interactions and not nucleotide similarity, it provides an avenue for identifying interactions involving lytic phages, that would not integrate into the bacterial genome.
+Therefore, while our approach could be improved by future work and more comprehensive datasets, we are addressing lysogenic phages, the most abundant component of the virome.
+Second, it is worth noting that while the alignment approaches were most informative, the protein family interaction approach was also informative.
+Because the protein family approach relies on protein-protein interactions and not detecting prophage sequences in the bacterial genomes, it is not expected to be limited to lysogenic phages.
 LINES 182 - 184*
@@
 In this way the authors could test their binning strategy using artificial contigs with co-abundance data from their MetaHIT dataset.
-As we mentioned is the last review iteration as well, we are unsure of how to accomplish this without a similar robust empirical dataset with well described interactions.
+As we mentioned in the last review iteration as well, we are unsure of how to accomplish this without a similar robust empirical dataset with well described interactions.
 At best we would be building a simulation on a simulation.

### 2018-03-07 bin/BenchmarkDatabaseCreation.pl
Cleaning up by pushing some more code.

@@ -53,7 +53,7 @@ my $validation;
 # Be sure to set username and password as neo4j
 # User = 2nd value, PW = 3rd value
 eval {
-    REST::Neo4p->connect('http://localhost:7474/',"neo4j","neo4j");
+    REST::Neo4p->connect('http://localhost:7474/',"neo4j","root");
 };
 ref $@ ? $@->rethrow : die $@ if $@;
 
@@ -77,12 +77,20 @@ open(my $BLAST, "<", "$blast") || die "Unable to read in $blast: $!";
 open(my $PFAM, "<", "$pfam") || die "Unable to read in $pfam: $!";
 open(my $BLASTX, "<", "$blastx") || die "Unable to read in $blastx: $!";
 
+my $flagline = 0;
+
 sub AddGenericFile {
     # Import file handle
     my ($fileInput, $label, $score) = @_;
     # print "Score is $score\n";
     foreach my $line (<$fileInput>) {
-        # print $line;
+        print $line;
+        print $flagline."\n";
+        $flagline = $flagline + 1;
+            if ($flagline < 900) {
+            print "Skipped\n";
+            next;
+        }
         chomp $line;
         $Spacer = (split /\t/, $line)[0];
         $PhageTarget = (split /\t/, $line)[1];

### 2018-03-07 bin/BenchmarkingModel.sh

@@ -84,13 +84,13 @@ bash ./bin/GetCrisprPhagePairs.sh \
 
 # rm ./data/${Output}/tmp/*
 
-# Format the output
-FormatNames \
-	./data/${Output}/BenchmarkCrisprs.tsv \
-	${CRISPRout}
+# # Format the output
+# FormatNames \
+# 	./data/${Output}/BenchmarkCrisprs.tsv \
+# 	${CRISPRout}
 
-# Remove underscores at the end of the names
-sed -i 's/_[0-9][0-9]\?[0-9]\?\t/\t/g' ${CRISPRout}
+# # Remove underscores at the end of the names
+# sed -i 's/_[0-9][0-9]\?[0-9]\?\t/\t/g' ${CRISPRout}
 
 
 #####################

### 2018-03-07 bin/CalculateBroaderPerformance.sh

@@ -67,6 +67,7 @@ grep \
 	--file=./data/genbankPhageHost/virusphage.tsv \
 	./data/genbankPhageHost/VirRef/viral.genomic.noblock.fa \
 	| grep -A 1 "complete genome" \
+	| egrep -v '\-\-' \
 	> ./data/genbankPhageHost/VirRef/filtered.virus.fa
 
 

### 2018-03-07 bin/CompareSkin.R

@@ -40,6 +40,13 @@ rm(i)
 
 totalgraph <- rbind(graphdfTP2, graphdfTP3)
 
+filterlist <- read.delim(
+  file = "./data/contigclustersidentity/bacterialremoval-clusters-list.tsv",
+  header = FALSE)
+
+# Remove the filtered out phage OGUs
+totalgraph <- totalgraph[!c(totalgraph$from %in% filterlist$V1),]
+
 # See the object size
 format(object.size(totalgraph), units = "MB")
 

### 2018-03-07 bin/CompareSkinWithInaccurateModel.R

@@ -1,0 +1,319 @@
+##################
+# Load Libraries #
+##################
+packagelist <- c("RNeo4j", "ggplot2", "wesanderson", "igraph", "scales", "plyr", "cowplot", "vegan", "reshape2", "parallel", "stringr")
+new.packages <- packagelist[!(packagelist %in% installed.packages()[,"Package"])]
+if(length(new.packages)) install.packages(new.packages, repos='http://cran.us.r-project.org')
+lapply(packagelist, library, character.only = TRUE)
+library("tidyr")
+library("dplyr")
+
+itrnum <- 5
+
+##############################
+# Run Analysis & Save Output #
+##############################
+# Import graphs into a list
+skinsites <- c("Ax", "Ac", "Pa", "Tw", "Um", "Fh", "Ra")
+timepoints <- c("TP2", "TP3")
+subjectarray <- paste("skin", c(1:20), sep = "_")
+
+# Start list
+graphdfTP2 <- data.frame()
+graphdfTP3 <- data.frame()
+
+for (i in skinsites) {
+	print(i)
+	filename <- paste("./data/skingraph-", i, ".Rdata", sep = "")
+	load(file = filename)
+	graphdfTP2 <- rbind(graphdfTP2, sampletable)
+	rm(sampletable)
+}
+
+rm(i)
+
+for (i in skinsites) {
+	print(i)
+	filename <- paste("./data/skingraph-", i, "-TP3.Rdata", sep = "")
+	load(file = filename)
+	graphdfTP3 <- rbind(graphdfTP3, sampletable)
+	rm(sampletable)
+}
+
+rm(i)
+
+totalgraph <- rbind(graphdfTP2, graphdfTP3)
+
+# Remove the filtered out phage OGUs
+filterlist <- read.delim(
+  file = "./data/contigclustersidentity/bacterialremoval-clusters-list.tsv",
+  header = FALSE)
+
+totalgraph <- totalgraph[!c(totalgraph$from %in% filterlist$V1),]
+
+autoskindata <- function(x, prob) {
+	write("Starting subroutine...", stdout())
+	skingraph <- x
+	# Can introduce noise by removing random rows as a percent
+	skingraph <- skingraph %>%
+		group_by(PatientID, TimePoint, Location) %>%
+		sample_frac(size = prob) %>%
+		as.data.frame()
+
[…]

### 2018-03-07 bin/Genbank2phagehost.py

@@ -1,4 +1,4 @@
-import SeqIO
+from Bio import SeqIO
 import re
 import argparse
 

### 2018-03-07 bin/GetNetworkLinkCounts.R

@@ -51,6 +51,10 @@ filter=0) {
 # Run Analysis & Save Output #
 ##############################
 
+filterlist <- read.delim(
+  file = "./data/contigclustersidentity/bacterialremoval-clusters-list.tsv",
+  header = FALSE)
+
 # Start the connection to the graph
 # If you are getting a lack of permission, disable local permission on Neo4J
 graph <- startGraph("http://localhost:7474/db/data/", "neo4j", "root")
@@ -64,6 +68,8 @@ RETURN n.Name AS from, m.Species AS to;
 graphoutputlist <- importgraphtodataframe()
 nodeout <- as.data.frame(graphoutputlist[1])
 edgeout <- as.data.frame(graphoutputlist[2])
+# Filter out low confidence edges
+edgeout <- edgeout[!c(edgeout$from %in% filterlist$V1),]
 head(nodeout)
 head(edgeout)
 

### 2018-03-07 bin/GetSkinGraphs.R

@@ -26,6 +26,15 @@ option_list <- list(
     type = "character",
     default = NULL,
     help = "Output table file.",
+    metavar = "character"),
+  make_option(c("-a", "--allconnections"),
+    action="store_true",
+    default=FALSE,
+    help = "All connections should be reported. [default %default]"),
+  make_option(c("-p", "--patient"),
+    type = "character",
+    default = NULL,
+    help = "Patient ID to use (this speeds up run time).",
     metavar = "character")
 )
 
@@ -38,30 +47,57 @@ opt <- parse_args(opt_parser);
 
 # Start the connection to the graph
 # If you are getting a lack of permission, disable local permission on Neo4J
-graph <- startGraph("http://localhost:7474/db/data/", "neo4j", "neo4j")
+graph <- startGraph("http://localhost:7474/db/data/", "neo4j", "root")
 
-write(paste("Pulling graph for", opt$location, opt$timepoint, sep = "\t"), stderr())
+if(opt$allconnections) {
+    write("Getting all connections.", stderr())
+    write(paste("Pulling graph for", opt$location, opt$timepoint, opt$patient, sep = " - "), stderr())
+    sampleidquery <- paste("
+    MATCH
+        (x:SRP049645)-->(y)-[d]->(z:Phage),
+        (a:Bacterial_Host)<-[e]-(b),
+        (b)<--(i:", opt$patient, ")-->(y),
+        (b)<--(t:", opt$timepoint, ")-->(y),
+        (b)<--(k:", opt$location, ")-->(y)
+    WHERE toInt(d.Abundance) > 0
+    OR toInt(e.Abundance) > 0
+    RETURN DISTINCT
+        z.Name AS from,
+        a.Name AS to,
+        i.Name AS PatientID,
+        t.Name AS TimePoint,
+        k.Name AS Location,
+        toInt(d.Abundance) AS PhageAbundance,
+        toInt(z.Length) AS PhageLength,
+        toInt(e.Abundance) AS BacteriaAbundance,
+        toInt(a.Length) AS BacteriaLength;
+    ", sep = "")
 
-sampleidquery <- paste("
-MATCH
-	(x:SRP049645)-->(y)-[d]->(z:Phage)-[:Infects]->(a:Bacterial_Host)<-[e]-(b),
[…]

### 2018-03-07 bin/PredictRelationships_SecondaryValidation.R

@@ -1,0 +1,116 @@
+#! /usr/local/bin/R
+# Geoffrey Hannigan
+# Pat Schloss Lab
+# University of Michigan
+
+packagelist <- c("RNeo4j", "ggplot2", "optparse", "caret", "wesanderson", "plotROC")
+new.packages <- packagelist[!(packagelist %in% installed.packages()[,"Package"])]
+if(length(new.packages)) install.packages(new.packages)
+lapply(packagelist, library, character.only = TRUE)
+library("dplyr")
+library("tidyr")
+library("viridis")
+
+# Load in the random forest model
+load("./data/rfinteractionmodel.RData")
+
+# Read in the data files
+crisprs <- read.delim(
+	file = "./data/SecondaryBenchmarkingSet/BenchmarkCrisprsFormat.tsv",
+	header = FALSE,
+	sep = "\t")
+
+prophages <- read.delim(
+	file = "./data/SecondaryBenchmarkingSet/BenchmarkProphagesFormatFlip.tsv",
+	header = FALSE,
+	sep = "\t")
+
+# To match original build, need low bitscore here as features
+prophages <- prophages %>%
+	group_by(V1, V2) %>%
+	summarize(V3 = min(V3)) %>%
+	as.data.frame()
+
+blastxresults <- read.delim(
+	file = "./data/SecondaryBenchmarkingSet/MatchesByBlastxFormatOrder.tsv",
+	header = TRUE,
+	sep = "\t")
+
+colnames(blastxresults) <- c("V1", "V2", "V3")
+
+pfaminteractions <- read.delim(
+	file = "./data/SecondaryBenchmarkingSet/PfamInteractionsFormatScoredFlip.tsv",
+	header = TRUE,
+	sep = "\t")
+
+colnames(pfaminteractions) <- c("V1", "V2", "V3")
+
+# Start joining together the data frames
+j1 <- full_join(crisprs, prophages, by = c("V1" = "V1", "V2" = "V2"))
+j2 <- full_join(j1, blastxresults, by = c("V1" = "V1", "V2" = "V2"))
+j3 <- full_join(j2, pfaminteractions, by = c("V1" = "V1", "V2" = "V2"))
+
+j3[,3:6] <- as.data.frame(sapply(j3[,3:6], as.numeric))
+rownames(j3) <- NULL
+j3[is.na(j3)] <- 0
+
+fdf <- data.frame(j3[apply(j3[, -c(1:2)], MARGIN = 1, function(x) any(x > 0)),])
+
[…]

### 2018-03-07 bin/VisualizeNetwork.R

@@ -145,6 +145,9 @@ connectionstrength <- function (nodeframe=nodeout, edgeframe=edgeout) {
 ##############################
 # Run Analysis & Save Output #
 ##############################
+filterlist <- read.delim(
+  file = "./data/contigclustersidentity/bacterialremoval-clusters-list.tsv",
+  header = FALSE)
 
 # Start the connection to the graph
 # If you are getting a lack of permission, disable local permission on Neo4J
@@ -160,6 +163,9 @@ RETURN n.Name AS from, m.Species AS to;
 graphoutputlist <- importgraphtodataframe()
 nodeout <- as.data.frame(graphoutputlist[1])
 edgeout <- as.data.frame(graphoutputlist[2])
+# Filter out low confidence edges
+edgeout <- edgeout[!c(edgeout$from %in% filterlist$V1),]
+nodeout <- nodeout[!c(nodeout$label %in% filterlist$V1),]
 head(nodeout)
 head(edgeout)
 
@@ -178,7 +184,7 @@ colnames(totalEcc) <- c("TotalECC", "class")
 # Collect some stats for the data table
 phagenodes <- length(grep("Phage", nodeout[,1]))
 bactnodes <- length(grep("Bacteria", nodeout[,1]))
-totaledges <- length(edgeout[,1])
+totaledges <- nrow(unique(edgeout[,c(1,2)]))
 nestats <- data.frame(cats = c("PhageNodes", "BacteriaNodes", "Edges"), values = c(phagenodes, bactnodes, totaledges))
 
 # Diet subgraph
@@ -203,6 +209,9 @@ RETURN DISTINCT
 graphoutputlist <- importgraphtodataframe()
 nodeout <- as.data.frame(graphoutputlist[1])
 edgeout <- as.data.frame(graphoutputlist[2])
+# Filter out low confidence edges
+edgeout <- edgeout[!c(edgeout$from %in% filterlist$V1),]
+nodeout <- nodeout[!c(nodeout$label %in% filterlist$V1),]
 # nodeout$order <- str_pad(row.names(nodeout), 4, pad = 0)
 # pabund <- ddply(edgeout[,c(1,6)], "from", summarize, avg = median(PhageAbundance))
 # babund <- ddply(edgeout[,c(2,7)], "to", summarize, avg = median(BacteriaAbundance))
[…]

### 2018-03-07 bin/alteredDiet.R

@@ -41,6 +41,14 @@ RETURN DISTINCT
 
 sampletable <- as.data.frame(cypher(graph, sampleidquery))
 
+
+# Remove the filtered out phage OGUs
+filterlist <- read.delim(
+  file = "./data/contigclustersidentity/bacterialremoval-clusters-list.tsv",
+  header = FALSE)
+
+sampletable <- sampletable[!c(sampletable$from %in% filterlist$V1),]
+
 # Correct the lengths
 sampletable$PhageAbundance <- round(1e7 * sampletable$PhageAbundance / sampletable$PhageLength)
 sampletable$BacteriaAbundance <- round(1e7 * sampletable$BacteriaAbundance / sampletable$BacteriaLength)
@@ -197,6 +205,15 @@ RETURN DISTINCT
 
 sampletable <- as.data.frame(cypher(graph, sampleidquery))
 
+
+# Remove the filtered out phage OGUs
+filterlist <- read.delim(
+  file = "./data/contigclustersidentity/bacterialremoval-clusters-list.tsv",
+  header = FALSE)
+
+sampletable <- sampletable[!c(sampletable$from %in% filterlist$V1),]
+
+
 head(sampletable)
 
 # get subsampling depth
@@ -292,7 +309,7 @@ obdgbox <- ggplot(rcentraldfmothers, aes(x = patientdiet, y = dg)) +
 	) +
 	geom_vline(xintercept=binlength,color="grey") +
 	ylim(0, NA) +
-	scale_x_discrete(labels = c("High Fat", "Low Fat"))
+	scale_x_discrete(labels = c("Healthy", "Obese"))
 
 obclbox <- ggplot(rcentraldfmothers, aes(x = patientdiet, y = cl)) +
 	theme_classic() +
@@ -305,7 +322,7 @@ obclbox <- ggplot(rcentraldfmothers, aes(x = patientdiet, y = cl)) +
 	) +
 	geom_vline(xintercept=binlength,color="grey") +
 	ylim(0, NA) +
-	scale_x_discrete(labels = c("High Fat", "Low Fat"))
+	scale_x_discrete(labels = c("Healthy", "Obese"))
 
 boxplots <- plot_grid(didgbox, diclbox, obdgbox, obclbox, labels = c("A", "B", "C", "D"), nrow = 1)
 

### 2018-03-07 bin/error_interpersonaldiversity.R

@@ -1,0 +1,780 @@
+##################
+# Load Libraries #
+##################
+gcinfo(FALSE)
+packagelist <- c("RNeo4j", "ggplot2", "wesanderson", "igraph", "visNetwork", "scales", "plyr", "cowplot", "vegan", "reshape2", "stringr")
+new.packages <- packagelist[!(packagelist %in% installed.packages()[,"Package"])]
+if(length(new.packages)) install.packages(new.packages, repos='http://cran.us.r-project.org')
+lapply(packagelist, library, character.only = TRUE)
+library("tidyr")
+library("dplyr")
+library("parallel")
+
+itrnum <- 5
+
+##############
+# Diet Graph #
+##############
+# Start the connection to the graph
+# If you are getting a lack of permission, disable local permission on Neo4J
+graph <- startGraph("http://localhost:7474/db/data/", "neo4j", "root")
+
+# Get list of the sample IDs
+sampleidquery <- "
+MATCH
+	(x:SRP002424)-->(y)-[d]->(z:Phage)-->(a:Bacterial_Host)<-[e]-(b),
+	(b)<--(i:PatientID)-->(y),
+	(b)<--(t:TimePoint)-->(y),
+	(k:Disease)-->(y)
+WHERE toInt(d.Abundance) > 0
+OR toInt(e.Abundance) > 0
+RETURN DISTINCT
+	z.Name AS from,
+	a.Name AS to,
+	i.Name AS PatientID,
+	t.Name AS TimePoint,
+	k.Name AS Diet,
+	toInt(d.Abundance) AS PhageAbundance,
+	toInt(z.Length) AS PhageLength,
+	toInt(e.Abundance) AS BacteriaAbundance,
+	toInt(a.Length) AS BacteriaLength;
+"
+
+sampletable <- as.data.frame(cypher(graph, sampleidquery))
+
+# Correct the lengths
+sampletable$PhageAbundance <- round(1e7 * sampletable$PhageAbundance / sampletable$PhageLength)
+sampletable$BacteriaAbundance <- round(1e7 * sampletable$BacteriaAbundance / sampletable$BacteriaLength)
+sampletable <- sampletable[,-9]
+sampletable <- sampletable[,-7]
+
+head(sampletable)
+
+# Remove the filtered out phage OGUs
+filterlist <- read.delim(
[…]

### 2018-03-07 bin/interpersonaldiversity.R

@@ -45,6 +45,13 @@ sampletable <- sampletable[,-7]
 
 head(sampletable)
 
+# Remove the filtered out phage OGUs
+filterlist <- read.delim(
+  file = "./data/contigclustersidentity/bacterialremoval-clusters-list.tsv",
+  header = FALSE)
+
+sampletable <- sampletable[!c(sampletable$from %in% filterlist$V1),]
+
 # get subsampling depth
 phageminseq <- min(ddply(sampletable, c("PatientID", "TimePoint"), summarize, sum = sum(PhageAbundance))$sum)
 bacminseq <- min(ddply(sampletable, c("PatientID", "TimePoint"), summarize, sum = sum(BacteriaAbundance))$sum)
@@ -216,6 +223,9 @@ for (i in skinsites) {
 rm(i)
 
 totalgraph <- rbind(graphdfTP2, graphdfTP3)
+
+# Remove the filtered out phage OGUs
+totalgraph <- totalgraph[!c(totalgraph$from %in% filterlist$V1),]
 
 # Correct the lengths
 totalgraph$PhageAbundance <- round(1e7 * totalgraph$PhageAbundance / totalgraph$PhageLength)
@@ -492,6 +502,9 @@ RETURN DISTINCT
 sampletable <- as.data.frame(cypher(graph, sampleidquery))
 
 head(sampletable)
+
+# Remove the filtered out phage OGUs
+sampletable <- sampletable[!c(sampletable$from %in% filterlist$V1),]
 
 # get subsampling depth
 phageminseq <- min(ddply(sampletable, c("PatientID", "TimePoint"), summarize, sum = sum(PhageAbundance))$sum)

### 2018-03-07 bin/iteratePredictedRelationships.R

@@ -1,0 +1,109 @@
+#! /usr/local/bin/R
+# Geoffrey Hannigan
+# Pat Schloss Lab
+# University of Michigan
+
+packagelist <- c("RNeo4j", "ggplot2", "optparse", "caret", "wesanderson", "plotROC")
+new.packages <- packagelist[!(packagelist %in% installed.packages()[,"Package"])]
+if(length(new.packages)) install.packages(new.packages)
+lapply(packagelist, library, character.only = TRUE)
+library("dplyr")
+library("tidyr")
+library("viridis")
+
+# Load in the random forest model
+load("./data/rfinteractionmodel.RData")
+
+hitlout <- lapply(c(1, 0.9, 0.8, 0.7, 0.6, 0.5), function(i) {
+	# Read in the data files
+	crisprs <- read.delim(
+		file = paste("./data/SecondaryBenchmarkingSet_", i, "/BenchmarkCrisprsFormat_", i, ".tsv", sep = ""),
+		header = FALSE,
+		sep = "\t")
+	
+	prophages <- read.delim(
+		file = paste("./data/SecondaryBenchmarkingSet_", i, "/BenchmarkProphagesFormatFlip_", i, ".tsv", sep = ""),
+		header = FALSE,
+		sep = "\t")
+	
+	# To match original build, need low bitscore here as features
+	prophages <- prophages %>%
+		group_by(V1, V2) %>%
+		summarize(V3 = min(V3)) %>%
+		as.data.frame()
+	
+	blastxresults <- read.delim(
+		file = paste("./data/SecondaryBenchmarkingSet_", i, "/MatchesByBlastxFormatOrder_", i, ".tsv", sep = ""),
+		header = TRUE,
+		sep = "\t")
+	
+	colnames(blastxresults) <- c("V1", "V2", "V3")
+	
+	pfaminteractions <- read.delim(
+		file = paste("./data/SecondaryBenchmarkingSet_", i, "/PfamInteractionsFormatScoredFlip_", i, ".tsv", sep = ""),
+		header = TRUE,
+		sep = "\t")
+	
+	colnames(pfaminteractions) <- c("V1", "V2", "V3")
+	
+	# Start joining together the data frames
+	j1 <- full_join(crisprs, prophages, by = c("V1" = "V1", "V2" = "V2"))
+	j2 <- full_join(j1, blastxresults, by = c("V1" = "V1", "V2" = "V2"))
[…]

### 2018-03-07 bin/iteratecontigsim.sh

@@ -1,0 +1,30 @@
+
+cut -f 2 ./data/genbankPhageHost/viral_host_species.tsv \
+	| sed 's/ /_/g' \
+	| sort \
+	| uniq \
+	> ./data/genbankPhageHost/viral_host_only.tsv
+
+# Clean before appending
+rm ./data/genbankPhageHost/BacRef/bacteria.complete.species.refset.fa
+
+while read p; do
+	echo "^>.+$p"
+	egrep -m 1 -A 1 "^>.+$p" ./data/genbankPhageHost/BacRef/bacteria.complete.species.fa >> ./data/genbankPhageHost/BacRef/bacteria.complete.species.refset.fa
+done < ./data/genbankPhageHost/viral_host_only.tsv
+
+# Create random contig fragments from the bacterial references and the viral references
+percents=( 1 0.9 0.8 0.7 0.6 0.5 )
+for i in "${percents[@]}"; do
+	echo $i
+	perl ./bin/RandomContigGenerator.pl -i ./data/genbankPhageHost/BacRef/bacteria.complete.species.refset.fa -o ./data/genbankPhageHost/BacRef/bacteria.fragment.${i}.fa -p $i
+	perl ./bin/RandomContigGenerator.pl -i ./data/genbankPhageHost/VirRef/filtered.virus.fa -o ./data/genbankPhageHost/VirRef/virus.fragment.${i}.fa -p $i
+	sed -i 's/[^>^A-Z^a-z^0-9]/_/g' ./data/genbankPhageHost/VirRef/virus.fragment.${i}.fa
+done
+
+
+percents=( 1 0.9 0.8 0.7 0.6 0.5 )
+for i in "${percents[@]}"; do
+	mkdir -p ./data/SecondaryBenchmarkingSet_${i}
[…]

### 2018-03-07 bin/makeSkinDatasets.sh

@@ -6,26 +6,70 @@
 
 # Set an array with the skin body sites
 SKINSITES=(Ax Ac Pa Tw Um Fh Ra Oc)
+PATIENTS=(\
+	skin_1 \
+	skin_2 \
+	skin_3 \
+	skin_4 \
+	skin_5 \
+	skin_6 \
+	skin_7 \
+	skin_8 \
+	skin_9 \
+	skin_10 \
+	skin_11 \
+	skin_12 \
+	skin_13 \
+	skin_14 \
+	skin_15 \
+	skin_16 \
+	skin_17 \
+	skin_18 \
+	skin_19 \
+	skin_20)
+TIMEPOINTS=(TP2 TP3)
 
-# Loop through the array with new neo4j instances
+# # Loop through the array with new neo4j instances
+# for i in "${SKINSITES[@]}"
+# do
+# 	echo Processing skin site $i
+# 	../../bin/neo4j-enterprise-2.3.0/bin/neo4j start
+# 	Rscript ./bin/GetSkinGraphs.R \
+# 		--location ${i} \
+# 		--timepoint TP2 \
+# 		--output ./data/skingraph-${i}.Rdata
+# 	../../bin/neo4j-enterprise-2.3.0/bin/neo4j stop
+# done
+
+# for i in "${SKINSITES[@]}"
+# do
+# 	echo Processing skin site $i
+# 	../../bin/neo4j-enterprise-2.3.0/bin/neo4j start
+# 	Rscript ./bin/GetSkinGraphs.R \
+# 		--location ${i} \
+# 		--timepoint TP3 \
+# 		--output ./data/skingraph-${i}-TP3.Rdata
+# 	../../bin/neo4j-enterprise-2.3.0/bin/neo4j stop
+# done
+
+# Get the total graphs as well
 for i in "${SKINSITES[@]}"
 do
-	echo Processing skin site $i
-	../../bin/neo4j-enterprise-2.3.0/bin/neo4j start
-	Rscript ./bin/GetSkinGraphs.R \
-		--location ${i} \
-		--timepoint TP2 \
-		--output ./data/skingraph-${i}.Rdata
-	../../bin/neo4j-enterprise-2.3.0/bin/neo4j stop
+	for j in "${PATIENTS[@]}"
+	do
+		for k in "${TIMEPOINTS[@]}"
+		do
+			echo Processing skin site $i
+			echo Processing patient $j
+			echo Processing timepoint $k
[…]

### 2018-03-07 revfiles/ReviewerResponse.md
Updated the reviewer comments.

@@
 We are also submitting a marked up revised manuscript with additions marked in blue and deletions marked in red.
-Included in the revisions are four additional figures, one additional table, updated text with further discussion, and an improved quality filtering process for the sequence set.
+Included in the revisions are four additional figures, one additional table, updated text with further discussion, and an improved quality filtering process for the sequence set, as suggested.
 All of the analyses were rerun following the additional suggested quality filtering steps.
@@
 
-> *We agree with these concerns and we think that the comments from the reviewer have really helped address them as described in our responses below..*
+> *We agree with these concerns and we think that the comments from the reviewer have really helped address them as described in our responses below.*
 
@@
 Because the protein family approach relies on protein-protein interactions and not detecting prophage sequences in the bacterial genomes, it is not expected to be limited to lysogenic phages.
-LINES 182 - 184*
+LINES 180 - 182*
 
@@
 > *We also thought it would be worth clarifying that we did not intend to communicate that we only found evidence of five matching phages, but rather that those were five OGUs that matched four reference genomes which had previously been validated as having incredibly broad range of infectiousness (cross-phylum infectious ranges).
-LINES 117 - 149, 496 - 497*
+LINES 116 - 147, 493 - 494*
 
@@
 Also, as we mentioned before, it will be exciting to evaluate this further in future work, with validated cases of phages that do not infect certain bacteria (a negative dataset).
-LINES 167 - 176, 527 - 549, Figure S5*
+LINES 166 - 174, 534 - 546, Figure S5*
 
@@
 As suggested by the reviewer, this confirms the confidence in our findings.
-LINES 295 - 301, 316 - 319, 396 - 397, 555 - 558, Figure S13, S14*
+LINES 292 - 298, 313 - 316, 393 - 394, 552 - 555, Figure S13, S14*
 
@@
 Together we feel that this information provides a nice validation for the work.
-Lines 419 - 432, 463 - 465, Figure S6*
+Lines 416 - 429, 460 - 462, Figure S6*
 
